### 2018-05-10 dynamicrangepaper/elsarticle-template/extendingdynamic_range.tex
Start of thesis template. And edits for dynamic range paper

%\documentclass[3p,twocolumn]{elsarticle}
\documentclass[review]{elsarticle}
\usepackage{graphicx}
\usepackage[export]{adjustbox}%for left right graphic adjust
\usepackage{amsmath}

\graphicspath{ {images/} }

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

%\title{Extending Dynamic Range of Generic Electronics for Time Projection Chambers (GET)}
\title{Extending the Dynamic Range of Electronics in a Time Projection Chamber}
%\title{Extending Dynamic Range of GET Electronics}

%% Group authors per affiliation:
%\author{Elsevier\fnref{myfootnote}}
%\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[msu,nscl]{J. Estee}
\author[msu,nscl]{W.G. Lynch}

\address[msu]{Michigan State University, Dept. Physics and Astronomy }
\address[nscl]{National Superconducting Cyclotron Laboratory}

\begin{abstract}
When applying Time Projection Chambers  (TPCs) to low and intermediate nuclear collisions one can face challenges in addressing large dynamic ranges in the energy losses of the detected particles. For recent experiments with the SAMURAI Pion-Reconstruction and Ion-Tracker (S$\pi$RIT) TPC, it was important to detect and analyze the small energy losses of relativistic pions as well as the large energy losses of, slow moving, heavy ions.  We illustrate how such large dynamic ranges can be handled in TPC with a wire plane readout  by effective use and knowledge the tails of the pad response function. By employing this technique, we are able to increase the signal to noise ratio by a factor of 2x.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction} 
The challenges of measuring large dynamic ranges in the energy losses of detected particles is widespread and a several techniques have been employed to address this issue in TPCs. At the time of publication it is common for TPC electronics to have at best a signal to noise ratio of 1000:1. If the signal to noise ratio of at least 20:1 is required to measure minimum ionizing particles, (m.i.p.), around $\beta\gamma=0.4$ this means the maximum range we could expect would be 50x greater than m.i.p.  In heavy ion collisions (HIC) of 300 MeVA fixed target beam energies the range of particle velocities ranges from m.i.p. to ~0.1$\beta\gamma$. These velocities cover about 40x the dE/dx of m.i.p. Also the energy loss scales like $z^2$ for larger charged particles. Also, if the track enters the TPC at an obtuse angle, the deposited charge on one pad could be higher by a factor of about 4x.  It becomes clear the large dynamic range that must be covered in such intermediate HIC experiments. 
A couple of TPC's such as the EOS TPC or the AT-TPC were able to circumvent the measuring of large energy deposits by lowering the gain in certain regions either by lowering the voltage on the anode wires or electronics gain settings. Of course the dE/dx resolution and accuracy of the momentum determination goes down for light particles as only a subset of the whole TPC can be used. Also there may be no dE/dx measurement for particles that avoid these regions. Here we illustrate how to expand the dynamic range without these drawbacks within the context of a standard multi-wire TPC.

\begin{figure}[ht]
\label{fig:padwire}
\includegraphics[width=\linewidth]{padwire}
\caption{Cartoon graphic showing the 3 wire planes and a section of the pad plane. h3 =14 mm, h2 = 8 mm, h1 = 4 mm, d2 = 1 mm, d1 = 4 mm. The actual orientation is inverted from this picture but for ease of displaying this orientation fits best.}
\end{figure}

\paragraph{1.1 TPC Overview}
\paragraph{Wire planes}
As seen in figure \ref{fig:padwire} the S$\pi$RIT TPC consists of three wire grids above the two dimensional array of charge sensitive readout pads, the pad plane. The first two wire grids operate as a gate and a shielding, or ground grid, with 1 mm spacing and they are not important for the discussion of this paper. The wire grid closest to the pad plane is the high voltage anode wire grid consisting of 20 $\mu$m wires spaced at 4 mm apart and set at a height of 4 mm from the pad plane. In the near vicinity of these wires the avalanche of the preliminary electrons occurs. The electrons deposited from tracks in the detector gas are multiplied on the order of 2000  times and the slow moving ions moving away induce a signal on the read out pads below. The resulting distribution on the pad plane is fixed by the geometry of the anode wire grid and its distance from the pad plane. The anode wire planes were sectioned off into individual distribution boards power 14 independent sections. 12 sections of the anode wire plane was held at 1460 V. This setting was optimized to ensure the small signal of pions would at least have a signal to noise ratio of 20:1. Two of the sections were held to 1214 V, this was due to some high currents which at the time of experiment were not fully understood. The reduction in voltage resulted in a reduction of about 10x gain compared to the higher anode wire sections. We will return to these two sections of lower gain later to validate the method that will be described. 
\paragraph{Pad plane} 
The S$\pi$RIT TPC pad plane consists of a 2-dimensional plane of charge sensitive pads. Each pad being rectangular in shape with a dimension of .8 cm x 1.2 cm. It is laid out on a grid measuring 112 by 108 pads with a total area of 134 cm x 86 cm. For convenience we have chosen the +x axis to point to the right of beam, the -y axis as the direction pointing down into the drift volume, and the +z axis along the incoming beam. The avalanche wires run perpendicular to the beam axis along the x axis as seen in the figure \ref{fig:padwire}. 
\paragraph{Generic Electronics for TPCs}
Signals from the pads in the S$\pi$RIT TPC are amplified and digitized by the newly developed Generic Electronics for TPCs (GET) \cite{get}.  Short cables transmit the signals from the pads to the inputs of the AGET chips. Each AGET chip can service 64 pads and contains a Preamp (PA), and a Switched Capacitor Array (SCA) with a maximum of 512 time buckets sampling at 1 to 100 MHz. Four AGET chips are mounted on one AsAd (Asic and Adc) motherboard. The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range. Also the peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns. The 12 bit ADC's on the AsAd boards deliver and effective 10.5 bit resolution. For the first series of experiments the gain was set to the highest setting, 0.12 pC, and the peaking time was set to 117 ns. At such a high gain, the pion signal was able to fully be measured. To put this gain setting into perspective, the minimum velocity that would cause saturated pads was $\beta\gamma\approx$ 0.2 for z=1 particles, $\beta\gamma\approx$ 0.6 for z=2 particles, and z$\geq$3 particles were always saturating.

\section{Pad Response Function}
\paragraph{Experimental PRF}
The fractional charge seen by each pad is referred to as the Pad Response Function (PRF). Some simple wire plane geometries have analytical expressions for the PRF which are well studied and may be looked up using a Gatti distribution \cite{blumrol}. Though theoretical PRF’s may be available, Blum and Rolandi still suggest an effective PRF may be required \cite{blumrol}. When analytical PRFs do not exist, an effective PRF may be calculated from experimental data. We postulate that the PRF is only a function of the total charge deposited on the wire Q and the displacement $\lambda$.

\begin{figure}[h]
\label{fig:av}
\includegraphics[width=\linewidth]{defofavalanche}
\caption{Cartoon graphic of avalanche event on an anode wire over one layer of pads. The estimate of the position of the avalanche is given by $\bar{x}$  the weighted mean. The position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
\end{figure}

\begin{equation}\label{eq:1}
\begin{split}
PRF(\lambda_i) = \frac{q_i(\lambda_i)}{Q}\\
where \ Q=\sum_i q_i\\
and \   \lambda_i=x_i-\bar{x}\\
\end{split}
\end{equation}

When calculating the effective PRF from experiment, we selected data that is not saturated, as this will only distort the shape of the PRF. Since the beam comes in along the z direction, the x direction gives the best momentum resolution and was the natural choice for clustering and calculating the PRF.

The PRF is given in equation \ref{eq:1} where i is the index over the pads and Q is the total charge within the layer. In figure \ref{fig:av}, the estimate for the avalanche position along the wire is given by the weighted mean position $\bar{x}$. Also seen is $\lambda_i$, defined as the difference, in position, of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x}$.

\begin{figure}[ht]
\label{fig:expprf}
\includegraphics[width=\linewidth]{expprf}
\caption{Experimental pad response function. Constructed from total number of pads $>$=3. }
\end{figure}

Calculating the PRF in the way described above, the resulting experimental PRF for the S$\pi$RIT TPC is seen in figure \ref{fig:exprf}. The PRF obtained from a of the experimental data gives rise to a well behaved function.  It is this function we will be fitting the data to extend the saturated pads.

\paragraph{Method of Desaturaiton}

Figure \ref{fig:satpad} shows a typical situation of saturated signals. This will be used as an example to explain the method of extrapolating the saturated pads, which I refer to in this paper as desaturation. When an avalanche causes an induced signal so large, the pads directly underneath collect the largest charge and typically would be saturated. These are represented as $q_{1'}$ and $q_{2'}$ in the figure \ref{fig:satpad}. The pads further away would experience smaller, non-saturated signals.

Though we don't know the charge of the saturated pads, we know that all the pads charges must satisfy the PRF distribution described in figure \ref{fig:expprf}. Thus, using these small non-saturated tails, we perform a $\chi^2$ fit to find the unknown charge on the saturated pads.

\begin{figure}[ht]
\label{fig:satpad}
\includegraphics[width=\linewidth]{saturated_pads}
\end{figure}

\section{Experimental data}

Two sets of data will be discussed in this paper. The first set was a tuned cocktail beam consisting of (p,d,t,${}^3He$,${}^4He$,${}^6Li$,${}^7Li$) light charged particles which was injected into the TPC for calibration purposes. The cocktail beam was tuned to two different $\beta\rho$ settings and the momentum resolution was approximately 1\% as determined by the slits of the BigRIPS fragment separator of the Radioactive Isotope Beam Factory (RIBF) facility in RIKEN. A thick 21mm thick aluminum target was inserted for part of the lower $\beta\rho$ setting, further reducing the energy of the beam for a third calibration point.

In a typical cocktail event one particle enters the TPC volume at a time and mostly parallel to the pad plane. Thus, the cocktail beam data represents an ideal case for the momentum and dE/dx determination. Also the cocktail data does not suffer from inefficiencies related to high multiplicity events we see in the collision data.

\begin{figure}[ht]
\label{fig:data}
\includegraphics[width=\linewidth]{data}
\caption{Pad plane projection for a collision event in the TPC. Highlighted by red arrows are two regions of anode wires which had a reduced voltage of 1214 V. The voltage of the rest of the TPC anode wires are 1460 V. The reduction in voltage reduces the gain by a factor of $\approx$10. }
\end{figure}

Shown in figure \ref{fig:data} is the typical pad plane response for a central nuclear collision of the system ${}^{}$Sn + ${}^{}$Sn. During the experiment two anode section voltages were turned down from 1460 V to 1214 V. The relative drop in gain is about 10. The anode sections which are lowered are denoted by the red arrows in figure \ref{fig:data}.

\section{Results}
\paragraph{Low gain vs corrected high gain}

As mentioned above, two anode sections, covering 12 layers of pads in total, had their gain lowered by approximately a factor of 10. The signals in this lowered gain region have effectively 10x the dynamic range as compared with the high gain regions. That is to say when a track would saturate pads in the high gain region, the signal in this low gain region is still preserved and can be measured. By comparing the $\frac{dE}{dx}$ values of the high gain sections with the low gain section we can determine whether the desaturation correction described above is successful.

\begin{figure}[ht]
\label{fig:lowvshigh_raw}
\includegraphics[width=\linewidth]{lowvshigh_raw_cuts}
\caption{The uncorrected high gain dE/dx vs low gain dE/dx for ~??? events of collision data.  }
\end{figure}

It is clear from figure \ref{fig:lowvshigh_raw} the effect of saturation has on the high gain channels. The data starts to saturate and deviates from the y=x line around about 400 ADC/mm in the high gain channels and eventually plateaus.  After applying the desaturation correction the correlation between the high gain and low gain sections is significantly restored as seen in figure \ref{fig:lowvshigh_desat}. Judging by the corrected correlation plot we can believe the correction to at least about 900 ADC/mm. It seems the 1:1 correlation is mostly restored and the increase in dynamic range in the high gain pads is about a factor of 2x.

\begin{figure}[ht]
\label{fig:lowvshigh_desat}
\includegraphics[width=\linewidth]{lowvshigh_desat}
\caption{The corrected high gain dE/dx vs low gain dE/dx for ~??? events of collision data.  }
\end{figure}

\paragraph{Particle Identification (PID)}

Comparing the low to high gain sections provides a direct comparison for determining the success of the extrapolation. The true goal of any correction would be to improve the PID. 
\begin{figure}[h]
\label{fig:cocktail_raw}
\includegraphics[width=\linewidth]{cocktail_raw}
\caption{Uncorrected cocktail data.}
\end{figure}

\begin{figure}[ht]
\label{fig:data_desat}
\includegraphics[width=\linewidth]{cocktail_desat}
\caption{Corrected (desaturated) cocktail data.}
\end{figure}

Looking firstly at the ideal case of the cocktail beam in figures \ref{fig:cocktail_raw} and \ref{fig:cocktail_desat} we note the very clean pronounced PID lines of several particle species. Once can clearly see the three  energy settings set by the magnetic rigidity of the fragment separator leading to three "blobs" around 1700 and two near 900 [MeV/c/z]. The remainder portion of the PID lines result form the particle losing its initial energy by passing through the walls and other materials outside the main detector volume. Looking at the uncorrected data in figure \ref{fig:cocktail_raw} we can see the effects of saturation turn the PID lines from the expected values from theory, given by red lines, and all seem to tend towards some kind of plateau. When looking at the uncorrected collision data in figure \ref{fig:data_raw}, one can see the lines more complete an the issue of saturation. One can see the stark difference after applying the desaturation method in figures \ref{fig:cocktail_desat} and \ref{fig:data_desat}. Most notable is the difference between He and Li particle species which suffer the most from saturation. Perhaps the more subtle improvement of the lighter particles, (p,d,t), can also be seen in the PID values at lower momentum.

\begin{figure}[ht]
\label{fig:data_raw}
\includegraphics[width=\linewidth]{data_raw}
\caption{Uncorrected collision data.}
\end{figure}

\begin{figure}[ht]
\label{fig:data_desat}
\includegraphics[width=\linewidth]{data_desat}
\caption{Corrected (desaturated) collision data.}
\end{figure}

\section{Conclusion}
We have shown that one can easily construct the Pad Response Function from unsaturated experimental data. Using this function and applying a simple $\chi^2$ fit to the unsaturated tails of data, one can recover the saturated pad charge in the middle of the distribution. Simply because the true PRF is fixed by the anode wire geometry. We also demonstrated the success of this method by a direct comparison of dE/dx of the high gain sections, which we are extending, to the low gain sections which give the true unsaturated dE/dx value. In the end this simple method gives significant improvement to the PID of particles. The effect improves the dynamic range by a factor of at least 2x. 
\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-05-14 dynamicrangepaper/elsarticle-template/extendingdynamic_range.tex
Incorperated betty's edits. need to review

@@
 \begin{abstract}
-When applying Time Projection Chambers  (TPCs) to low and intermediate nuclear collisions one can face challenges in addressing large dynamic ranges in the energy losses of the detected particles.
-For recent experiments with the SAMURAI Pion-Reconstruction and Ion-Tracker (S$\pi$RIT) TPC, it was important to detect and analyze the small energy losses of relativistic pions as well as the large energy losses of, slow moving, heavy ions.
-We illustrate how such large dynamic ranges can be handled in TPC with a wire plane readout  by effective use and knowledge the tails of the pad response function.
-By employing this technique, we are able to increase the signal to noise ratio by a factor of 2x.
+As Time Projection Chambers (TPCs) become widely utilized in low and intermediate nuclear physics experiments,  it becomes important to extend the dynamic range to cover  the large range in energy losses.
+In a recent set of experiments using the SAMURAI Pion-Reconstruction and Ion-Tracker (S$\pi$RIT) TPC,it was important to simultaneously measure relativistic pions and heavy ions from the same collisions.
+A track which saturates the TPC electronics only does so for several pads near to the track while pads further away are not saturated.
+By performing a $\chi^2$ fit using the pad response function on the unsaturated pads, we can recover the saturated pads charge.
+This resulted in an increase of the dynamic range by a factor of 2.
 \end{abstract}
@@
 \section{Introduction} 
-The challenges of measuring large dynamic ranges in the energy losses of detected particles is widespread and a several techniques have been employed to address this issue in TPCs.
+ At low to intermediate heavy ion collision (HIC) energies, it may be necessary to measure particles that have a large variation in energy losses.
+Several techniques have been employed to address this issue in time projection chambers (TPC).
 At the time of publication it is common for TPC electronics to have at best a signal to noise ratio of 1000:1.
 If the signal to noise ratio of at least 20:1 is required to measure minimum ionizing particles, (m.i.p.), around $\beta\gamma=0.4$ this means the maximum range we could expect would be 50x greater than m.i.p.
-In heavy ion collisions (HIC) of 300 MeVA fixed target beam energies the range of particle velocities ranges from m.i.p.
+In heavy ion collisions (HIC) of 300 AMeV beam energies the range of particle velocities ranges from m.i.p.
 to ~0.1$\beta\gamma$.
 These velocities cover about 40x the dE/dx of m.i.p.
-Also the energy loss scales like $z^2$ for larger charged particles.
+The energy loss scales like $z^2$ for larger charged particles.
 Also, if the track enters the TPC at an obtuse angle, the deposited charge on one pad could be higher by a factor of about 4x.
@@
 \begin{figure}[ht]
-\label{fig:padwire}
 \includegraphics[width=\linewidth]{padwire}
@@
 The actual orientation is inverted from this picture but for ease of displaying this orientation fits best.}
+\label{fig:padwire}
 \end{figure}
@@
 In the near vicinity of these wires the avalanche of the preliminary electrons occurs.
-The electrons deposited from tracks in the detector gas are multiplied on the order of 2000  times and the slow moving ions moving away induce a signal on the read out pads below.
+The electrons deposited from tracks in the detector gas are multiplied on the order of 2000  times and the slow moving ions induce a signal on the read out pads below.
 The resulting distribution on the pad plane is fixed by the geometry of the anode wire grid and its distance from the pad plane.
-The anode wire planes were sectioned off into individual distribution boards power 14 independent sections.
+The anode wire planes were sectioned off into 14 independent sections.
 12 sections of the anode wire plane was held at 1460 V.
 This setting was optimized to ensure the small signal of pions would at least have a signal to noise ratio of 20:1.
-Two of the sections were held to 1214 V, this was due to some high currents which at the time of experiment were not fully understood.
+Two of the sections were held to 1214 V.
 The reduction in voltage resulted in a reduction of about 10x gain compared to the higher anode wire sections.
-We will return to these two sections of lower gain later to validate the method that will be described.
+These two sections of lower gain allow validation of the method that will be described.
 \paragraph{Pad plane} 
 The S$\pi$RIT TPC pad plane consists of a 2-dimensional plane of charge sensitive pads.
-Each pad being rectangular in shape with a dimension of .8 cm x 1.2 cm.
+Each pad is rectangular in shape with a dimension of .8 cm x 1.2 cm.
 It is laid out on a grid measuring 112 by 108 pads with a total area of 134 cm x 86 cm.
@@
 Some simple wire plane geometries have analytical expressions for the PRF which are well studied and may be looked up using a Gatti distribution \cite{blumrol}.
-Though theoretical PRF’s may be available, Blum and Rolandi still suggest an effective PRF may be required \cite{blumrol}.
 When analytical PRFs do not exist, an effective PRF may be calculated from experimental data.
@@
 \begin{figure}[h]
-\label{fig:av}
 \includegraphics[width=\linewidth]{defofavalanche}
@@
 The position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
+\label{fig:av}
 \end{figure}
@@
 \begin{figure}[ht]
-\label{fig:expprf}
 \includegraphics[width=\linewidth]{expprf}
 \caption{Experimental pad response function. Constructed from total number of pads $>$=3. }
+\label{fig:expprf}
 \end{figure}
@@
 Calculating the PRF in the way described above, the resulting experimental PRF for the S$\pi$RIT TPC is seen in figure \ref{fig:exprf}.
-The PRF obtained from a of the experimental data gives rise to a well behaved function.
-It is this function we will be fitting the data to extend the saturated pads.
+The PRF obtained from averaging many events gives rise to a well behaved function.
+It is this function we will be using to fit the data to extend the saturated pads.
 
@@
 Figure \ref{fig:satpad} shows a typical situation of saturated signals.
-This will be used as an example to explain the method of extrapolating the saturated pads, which I refer to in this paper as desaturation.
-When an avalanche causes an induced signal so large, the pads directly underneath collect the largest charge and typically would be saturated.
+This will be used as an example to explain the method of extrapolating the saturated pads, referred to in this paper as desaturation.
+When an avalanche causes large induced signal, the pads directly underneath collect the largest charge and typically would be saturated.
 These are represented as $q_{1'}$ and $q_{2'}$ in the figure \ref{fig:satpad}.
@@
 
-Though we don't know the charge of the saturated pads, we know that all the pads charges must satisfy the PRF distribution described in figure \ref{fig:expprf}.
-Thus, using these small non-saturated tails, we perform a $\chi^2$ fit to find the unknown charge on the saturated pads.
+Since all the pad's charges must satisfy the PRF distribution described in figure \ref{fig:expprf}, then using these small non-saturated tails we perform a $\chi^2$ fit to find the unknown charge on the saturated pads.
 
 \begin{figure}[ht]
+\includegraphics[width=\linewidth]{saturated_pads}
+\caption{•}
 \label{fig:satpad}
-\includegraphics[width=\linewidth]{saturated_pads}
 \end{figure}
@@
 In a typical cocktail event one particle enters the TPC volume at a time and mostly parallel to the pad plane.
-Thus, the cocktail beam data represents an ideal case for the momentum and dE/dx determination.
-Also the cocktail data does not suffer from inefficiencies related to high multiplicity events we see in the collision data.
+Thus, the cocktail beam data represents an ideal case for the momentum and dE/dx determination as it does not suffer from inefficiencies related to high multiplicity events we see in the collision experimental data.
 
 \begin{figure}[ht]
-\label{fig:data}
 \includegraphics[width=\linewidth]{data}
@@
 }
+\label{fig:data}
 \end{figure}
 
-Shown in figure \ref{fig:data} is the typical pad plane response for a central nuclear collision of the system ${}^{}$Sn + ${}^{}$Sn.
-During the experiment two anode section voltages were turned down from 1460 V to 1214 V.
-The relative drop in gain is about 10.
-The anode sections which are lowered are denoted by the red arrows in figure \ref{fig:data}.
+Shown in figure \ref{fig:data} is the typical pad plane response for a central nuclear collision of the system ${132}^{}$Sn + ${124}^{}$Sn.
+During the experiment the voltages of two anode sections (as indicated by red arrows in figure \ref{fig:data}) were turned down from 1460 V to 1214 V.
+Since the anode voltages are dropped, the gain of these sections is also reduced by a factor of about 10.
 
@@
 \begin{figure}[ht]
+\includegraphics[width=\linewidth]{lowvshigh_raw_cuts}
+\caption{The uncorrected high gain dE/dx vs low gain dE/dx collision data.  }
 \label{fig:lowvshigh_raw}
-\includegraphics[width=\linewidth]{lowvshigh_raw_cuts}
-\caption{The uncorrected high gain dE/dx vs low gain dE/dx for ~??? events of collision data.  }
 \end{figure}
@@
 After applying the desaturation correction the correlation between the high gain and low gain sections is significantly restored as seen in figure \ref{fig:lowvshigh_desat}.
-Judging by the corrected correlation plot we can believe the correction to at least about 900 ADC/mm.
+Judging by the corrected correlati on plot we can believe the correction to at least about 900 ADC/mm.
 It seems the 1:1 correlation is mostly restored and the increase in dynamic range in the high gain pads is about a factor of 2x.
@@
 \begin{figure}[ht]
-\label{fig:lowvshigh_desat}
 \includegraphics[width=\linewidth]{lowvshigh_desat}
 \caption{The corrected high gain dE/dx vs low gain dE/dx for ~??? events of collision data.  }
+\label{fig:lowvshigh_desat}
 \end{figure}
@@
 \begin{figure}[h]
-\label{fig:cocktail_raw}
 \includegraphics[width=\linewidth]{cocktail_raw}
 \caption{Uncorrected cocktail data.}
+\label{fig:cocktail_raw}
 \end{figure}
@@
 \begin{figure}[ht]
-\label{fig:data_desat}
 \includegraphics[width=\linewidth]{cocktail_desat}
 \caption{Corrected (desaturated) cocktail data.}
+\label{fig:data_desat}
 \end{figure}
@@
 Looking firstly at the ideal case of the cocktail beam in figures \ref{fig:cocktail_raw} and \ref{fig:cocktail_desat} we note the very clean pronounced PID lines of several particle species.
-Once can clearly see the three  energy settings set by the magnetic rigidity of the fragment separator leading to three "blobs" around 1700 and two near 900 [MeV/c/z].
-The remainder portion of the PID lines result form the particle losing its initial energy by passing through the walls and other materials outside the main detector volume.
-Looking at the uncorrected data in figure \ref{fig:cocktail_raw} we can see the effects of saturation turn the PID lines from the expected values from theory, given by red lines, and all seem to tend towards some kind of plateau.
-When looking at the uncorrected collision data in figure \ref{fig:data_raw}, one can see the lines more complete an the issue of saturation.
-One can see the stark difference after applying the desaturation method in figures \ref{fig:cocktail_desat} and \ref{fig:data_desat}.
+Once can clearly see the three  B$\rho$ settings of the fragment separator leading to three ovals around 1700 and two near 900 [MeV/c/Z].
+The tails of the PID lines leading away from these three ovals are resulting from the particle losing its initial energy by passing through the walls and other materials outside the main detector volume.
+Looking at the uncorrected data in figure \ref{fig:cocktail_raw} we can see the effects of saturation.The experimental PID deviates from the theoretical values given by the red curves.
+It seems the dE/dx values plateau starting around 400 ADC/mm as we also saw in figure \ref{fig:lowvshigh_raw}.
+After applying the desaturation method we see a stark difference in figures \ref{fig:cocktail_desat} and \ref{fig:data_desat}.
 Most notable is the difference between He and Li particle species which suffer the most from saturation.
@@
 \begin{figure}[ht]
-\label{fig:data_raw}
 \includegraphics[width=\linewidth]{data_raw}
 \caption{Uncorrected collision data.}
+\label{fig:data_raw}
 \end{figure}
@@
 \begin{figure}[ht]
-\label{fig:data_desat}
 \includegraphics[width=\linewidth]{data_desat}
 \caption{Corrected (desaturated) collision data.}
+\label{fig:data_desat}
 \end{figure}
@@
 We have shown that one can easily construct the Pad Response Function from unsaturated experimental data.
-Using this function and applying a simple $\chi^2$ fit to the unsaturated tails of data, one can recover the saturated pad charge in the middle of the distribution.
-Simply because the true PRF is fixed by the anode wire geometry.
-We also demonstrated the success of this method by a direct comparison of dE/dx of the high gain sections, which we are extending, to the low gain sections which give the true unsaturated dE/dx value.
-In the end this simple method gives significant improvement to the PID of particles.
-The effect improves the dynamic range by a factor of at least 2x.
+Using this function and applying a simple $\chi^2$ fit to the unsaturated tails of data, one can recover the saturated pad charge in the middle of the distribution because the true PRF is fixed by the anode wire geometry.
+We also demonstrated the success of this method by a direct comparison of dE/dx obtained in the high gain sections,to the true unsaturated dE/dx value obtained in the low gain sections which give the true unsaturated dE/dx value.
+This simple method gives significant improvement to the PID of particle and improves the dynamic range by a factor of at least 2.
 \section*{References}

### 2018-10-05 dynamicrangepaper/elsarticle-template/extendingdynamic_range.tex
Before bills edits and bettys merge. New vector graphics

@@
 \usepackage{amsmath}
-
+\usepackage{float}
 \graphicspath{ {images/} }
@@
 \author[msu,nscl]{W.G. Lynch}
+\author[msu,nscl]{J. Barney}
+\author[msu,nscl]{G. Cerizza}
+\author[kor]{B. Hong}
+\author[riken]{T. Isobe}
+\author[nscl]{G. Jhang}
+\author[kyoto]{M. Kaneko}
+\author[riken]{M. Kurata-Nishimura}
+\author[krakow]{P. Lasko}
+\author[kor]{J. W. Lee}
+\author[krakow]{J. Lukasik}
+\author[a&m]{A.B. McIntosh}
+\author[kyoto]{T. Murakami}
+\author[krakow]{P. Pawlowski}
+\author[poland]{K. Pelczar}
+\author[nscl]{C. Santamaria}
+\author[riken]{D. Suzuki}
+\author[nscl]{M. B. Tsang}
+\author[a&m]{S.J. Yennello}
+\author[tsing]{Y. Zhang}
+\author[]{and the S$\pi$RIT collaboration}
 
@@
 \address[nscl]{National Superconducting Cyclotron Laboratory}
+\address[kor]{Department of Physics, Korea University}
+\address[riken]{RIKEN Nishina Center}
+\address[kyoto]{Department of Physics, Kyoto University}
+\address[krakow]{IFJ PAN, Krak\'{o}w}
+\address[a&m]{Dept. of Physics and Astronomy, Texas A$\&$M University}
+\address[tsing]{Department of Physics, Tsinghua University}
+\address[poland]{Faculty of Physics, Astronomy and Applied Computer Science, Jagiellonian University}
 
 \begin{abstract}
-As Time Projection Chambers (TPCs) become widely utilized in low and intermediate nuclear physics experiments,  it becomes important to extend the dynamic range to cover  the large range in energy losses.
-In a recent set of experiments using the SAMURAI Pion-Reconstruction and Ion-Tracker (S$\pi$RIT) TPC,it was important to simultaneously measure relativistic pions and heavy ions from the same collisions.
+As Time Projection Chambers (TPCs) become widely used in low to intermediate nuclear physics experiments,  it becomes important to extend the dynamic range to cover  the large range in energy losses.
+In a recent set of experiments using the SAMURAI Pion-Reconstruction and Ion-Tracker (S$\pi$RIT) TPC, it was important to simultaneously measure relativistic pions and heavy ions from the same collisions.
 A track which saturates the TPC electronics only does so for several pads near to the track while pads further away are not saturated.
-By performing a $\chi^2$ fit using the pad response function on the unsaturated pads, we can recover the saturated pads charge.
+By performing a $\chi^2$ fit using the pad response function on the unsaturated pads, we can recover the saturated pad's charges.
 This resulted in an increase of the dynamic range by a factor of 2.
@@
  At low to intermediate heavy ion collision (HIC) energies, it may be necessary to measure particles that have a large variation in energy losses.
-Several techniques have been employed to address this issue in time projection chambers (TPC).
-At the time of publication it is common for TPC electronics to have at best a signal to noise ratio of 1000:1.
-If the signal to noise ratio of at least 20:1 is required to measure minimum ionizing particles, (m.i.p.), around $\beta\gamma=0.4$ this means the maximum range we could expect would be 50x greater than m.i.p.
+At the time of publication it is common for Time Projection Chamber (TPC) electronics to have a dynamic range, the maximum signal to noise ratio, of around 1000:1.
+If a signal to noise ratio of at least 20:1 is required to measure minimum ionizing particles (m.i.p.), around $\beta\gamma=0.4$, this means the maximum range we could expect would be 50x greater than m.i.p.
 In heavy ion collisions (HIC) of 300 AMeV beam energies the range of particle velocities ranges from m.i.p.
 to ~0.1$\beta\gamma$.
-These velocities cover about 40x the dE/dx of m.i.p.
-The energy loss scales like $z^2$ for larger charged particles.
-Also, if the track enters the TPC at an obtuse angle, the deposited charge on one pad could be higher by a factor of about 4x.
+These velocities alone cover about 40x the dE/dx of m.i.p.
+Since the energy loss scales like $z^2$ for larger charged particles we could expect a factor of 9x more for z=3 charged particles.
+Also, if the track enters the TPC at an obtuse angle, the collected charge on one pad could be higher by a factor of about 4x.
 It becomes clear the large dynamic range that must be covered in such intermediate HIC experiments.
-A couple of TPC's such as the EOS TPC or the AT-TPC were able to circumvent the measuring of large energy deposits by lowering the gain in certain regions either by lowering the voltage on the anode wires or electronics gain settings.
-Of course the dE/dx resolution and accuracy of the momentum determination goes down for light particles as only a subset of the whole TPC can be used.
-Also there may be no dE/dx measurement for particles that avoid these regions.
+
+ \begin{figure}[H]
+\includegraphics[width=\linewidth]{intrographic}
+\caption{}
+\label{fig:intro}
+\end{figure}
+
+Several techniques have been employed to address this issue in TPCs.
+A couple of TPCs such as the EOS TPC \cite{eos} or the AT-TPC were able to circumvent the measuring of large energy deposits by lowering the gain in certain regions either by lowering the voltage on the anode wires or electronics gain settings.
+Lowering the gain in large regions of the TPC worsens the dE/dx resolution and momentum determination for m.i.p.
+as only a subset of the whole TPC can be used.
+Also there may be no dE/dx measurement for tracks that avoid these regions.
 Here we illustrate how to expand the dynamic range without these drawbacks within the context of a standard multi-wire TPC.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{padwire}
@@
 h3 =14 mm, h2 = 8 mm, h1 = 4 mm, d2 = 1 mm, d1 = 4 mm.
-The actual orientation is inverted from this picture but for ease of displaying this orientation fits best.}
+This graphic is inverted from the actual wire planes and pad plane to display the perspective easier.}
 \label{fig:padwire}
@@
 \paragraph{Wire planes}
-As seen in figure \ref{fig:padwire} the S$\pi$RIT TPC consists of three wire grids above the two dimensional array of charge sensitive readout pads, the pad plane.
+As seen in figure \ref{fig:padwire} the S$\pi$RIT TPC consists of three wire grids below the two dimensional array of charge sensitive readout pads, the pad plane.
 The first two wire grids operate as a gate and a shielding, or ground grid, with 1 mm spacing and they are not important for the discussion of this paper.
@@
 The anode wire planes were sectioned off into 14 independent sections.
-12 sections of the anode wire plane was held at 1460 V.
-This setting was optimized to ensure the small signal of pions would at least have a signal to noise ratio of 20:1.
-Two of the sections were held to 1214 V.
-The reduction in voltage resulted in a reduction of about 10x gain compared to the higher anode wire sections.
-These two sections of lower gain allow validation of the method that will be described.
+12 sections of the anode wire planes were held at 1460 V.
+This setting was optimized to ensure the small signal of pions would at have a least a signal to noise ratio of 20:1.
+Two of the sections were held to 1214 V due to a concern of the current each section was drawing.
+The reduction in voltage resulted in a reduction in gain of about 10x compared to the higher anode wire sections.
+These two sections of lower gain allowed for a direct validation of the method that will be described.
+
 \paragraph{Pad plane} 
-The S$\pi$RIT TPC pad plane consists of a 2-dimensional plane of charge sensitive pads.
-Each pad is rectangular in shape with a dimension of .8 cm x 1.2 cm.
+The S$\pi$RIT TPC pad plane consists of a 2-dimensional plane of charge sensitive pads which are rectangular in shape with a dimension of 0.8 cm x 1.2 cm.
 It is laid out on a grid measuring 112 by 108 pads with a total area of 134 cm x 86 cm.
-For convenience we have chosen the +x axis to point to the right of beam, the -y axis as the direction pointing down into the drift volume, and the +z axis along the incoming beam.
+For convenience we have chosen the +x axis to point to the left of beam, the -y axis as the direction pointing down into the drift volume, and the +z axis along the incoming beam.
 The avalanche wires run perpendicular to the beam axis along the x axis as seen in the figure \ref{fig:padwire}.
@@
 Also the peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns.
-The 12 bit ADC's on the AsAd boards deliver and effective 10.5 bit resolution.
-For the first series of experiments the gain was set to the highest setting, 0.12 pC, and the peaking time was set to 117 ns.
+The ADCs on the AsAd boards deliver a 12 bit resolution.
+For the first series of experiments the gain was set to the highest setting, 0.12 pC, and the peaking time was set to 117 ns, and the sampling time set to 25 MHz.
 At such a high gain, the pion signal was able to fully be measured.
@@
 \paragraph{Experimental PRF}
+When electrons terminate on the anode wires they induce a 2-dimensional charge distribution on the pad plane.
 The fractional charge seen by each pad is referred to as the Pad Response Function (PRF).
 Some simple wire plane geometries have analytical expressions for the PRF which are well studied and may be looked up using a Gatti distribution \cite{blumrol}.
-When analytical PRFs do not exist, an effective PRF may be calculated from experimental data.
-We postulate that the PRF is only a function of the total charge deposited on the wire Q and the displacement $\lambda$.
+When analytic PRFs do not exist, an effective PRF may be calculated from experimental data.
+We postulate that the PRF is only a function of the total charge deposited on the wire Q and the displacement, $\lambda$, from the mean avalanche position $\bar{x}$.
 
-\begin{figure}[h]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{defofavalanche}
@@
 
-When calculating the effective PRF from experiment, we selected data that is not saturated, as this will only distort the shape of the PRF.
-Since the beam comes in along the z direction, the x direction gives the best momentum resolution and was the natural choice for clustering and calculating the PRF.
+For the purpose of calculating the effective PRF from experimental data, we selected only the pads which were not saturated.
+Since the beam comes in along the z direction, the x direction gives the best momentum resolution and was the natural choice for combining hits into clusters and calculating the PRF.
 
-The PRF is given in equation \ref{eq:1} where i is the index over the pads and Q is the total charge within the layer.
-In figure \ref{fig:av}, the estimate for the avalanche position along the wire is given by the weighted mean position $\bar{x}$.
-Also seen is $\lambda_i$, defined as the difference, in position, of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x}$.
+The way we calculate the PRF is given by equation \ref{eq:1} where i is the index over the pads and Q is the total charge within the layer.
+Figure \ref{fig:av}  illustrates the estimate for the avalanche position along the wire given by, the weighted mean position $\bar{x}$.
+Also seen is $\lambda_i$, defined as the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x}$.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{expprf}
@@
 
-Calculating the PRF in the way described above, the resulting experimental PRF for the S$\pi$RIT TPC is seen in figure \ref{fig:exprf}.
-The PRF obtained from averaging many events gives rise to a well behaved function.
+The resulting experimental PRF for the S$\pi$RIT TPC is shown in figure \ref{fig:expprf} and is obtained from averaging many events and describes a well behaved function.
 It is this function we will be using to fit the data to extend the saturated pads.
@@
 
+For convenience we will use the term desaturation for when we estimate and recover the charge values of the saturated pads.
 Figure \ref{fig:satpad} shows a typical situation of saturated signals.
-This will be used as an example to explain the method of extrapolating the saturated pads, referred to in this paper as desaturation.
-When an avalanche causes large induced signal, the pads directly underneath collect the largest charge and typically would be saturated.
+When an avalanche causes a large induced signal, the pads directly underneath collect the largest charge and typically would be saturated.
 These are represented as $q_{1'}$ and $q_{2'}$ in the figure \ref{fig:satpad}.
-The pads further away would experience smaller, non-saturated signals.
+The pads further away would experience smaller, non-saturated signals shown as $q_{1}$ and $q_{4}$.
 
-Since all the pad's charges must satisfy the PRF distribution described in figure \ref{fig:expprf}, then using these small non-saturated tails we perform a $\chi^2$ fit to find the unknown charge on the saturated pads.
+Since the charge deposited on each  pad must satisfy the PRF distribution described in figure \ref{fig:expprf}, then using these small non-saturated tails we perform a $\chi^2$ fit to find the unknown charge on the saturated pads.
+The data points of the $\chi^2$ fit are the unsaturated pads,  $q_{1}$ and $q_{4}$, and the unknown parameters of the fit are $q_{2'}$ and $q_{3'}$ with the expected values coming from the PRF described above.
+The values of $q_{2'}$ and $q_{3'}$ at the minimum of the $\chi^2$ would be the best estimate for the saturated pads.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{saturated_pads}
@@
 
-Two sets of data will be discussed in this paper.
+Two sets of data were used for testing and validation of this method.
 The first set was a tuned cocktail beam consisting of (p,d,t,${}^3He$,${}^4He$,${}^6Li$,${}^7Li$) light charged particles which was injected into the TPC for calibration purposes.
@@
 In a typical cocktail event one particle enters the TPC volume at a time and mostly parallel to the pad plane.
-Thus, the cocktail beam data represents an ideal case for the momentum and dE/dx determination as it does not suffer from inefficiencies related to high multiplicity events we see in the collision experimental data.
+Therefore the cocktail beam data represents an ideal case for the momentum and dE/dx determination as it does not suffer from inefficiencies related to high multiplicity events we see in the collision experimental data.
 
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{data}
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{data.pdf}
 \caption{Pad plane projection for a collision event in the TPC.
@@
 The voltage of the rest of the TPC anode wires are 1460 V.
-The reduction in voltage reduces the gain by a factor of $\approx$10.
+The reduction in voltage reduces the gain by a factor of about 10x.
 }
@@
 
-Shown in figure \ref{fig:data} is the typical pad plane response for a central nuclear collision of the system ${132}^{}$Sn + ${124}^{}$Sn.
-During the experiment the voltages of two anode sections (as indicated by red arrows in figure \ref{fig:data}) were turned down from 1460 V to 1214 V.
-Since the anode voltages are dropped, the gain of these sections is also reduced by a factor of about 10.
+The other type of data was where we collided a ${}^{132}$Sn beam onto a ${}^{124}$Sn target in which we triggered on central nuclear collisions.
+Shown in figure \ref{fig:data} is the typical pad plane response for a central nuclear collision.
+During the experiment the voltages of two anode sections (as indicated by red arrows in figure \ref{fig:data} were turned down from 1460 V to 1214 V.
+Since the anode voltages were dropped, the gain of these sections were also reduced by a factor of about 10x.
 
@@
 
-As mentioned above, two anode sections, covering 12 layers of pads in total, had their gain lowered by approximately a factor of 10.
-The signals in this lowered gain region have effectively 10x the dynamic range as compared with the high gain regions.
-That is to say when a track would saturate pads in the high gain region, the signal in this low gain region is still preserved and can be measured.
-By comparing the $\frac{dE}{dx}$ values of the high gain sections with the low gain section we can determine whether the desaturation correction described above is successful.
+As mentioned above two anode sections, covering 12 layers of pads in total, had their gain lowered by approximately a factor of 10x.
+Thus, the signals in this lowered gain region have effectively 10x the dynamic range as compared with the high gain regions.
+That is to say when a track would saturate pads in the high gain region, the signal in this low gain region was still preserved and could be measured.
+By comparing the dE/dx values of the high gain sections with the low gain section we can determine whether the desaturation correction described above is successful.
 
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{lowvshigh_raw_cuts}
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{dedxcompare_nodesat}
 \caption{The uncorrected high gain dE/dx vs low gain dE/dx collision data.  }
@@
 
-It is clear from figure \ref{fig:lowvshigh_raw} the effect of saturation has on the high gain channels.
-The data starts to saturate and deviates from the y=x line around about 400 ADC/mm in the high gain channels and eventually plateaus.
-After applying the desaturation correction the correlation between the high gain and low gain sections is significantly restored as seen in figure \ref{fig:lowvshigh_desat}.
-Judging by the corrected correlati on plot we can believe the correction to at least about 900 ADC/mm.
-It seems the 1:1 correlation is mostly restored and the increase in dynamic range in the high gain pads is about a factor of 2x.
+When plotting the raw data in figure \ref{fig:lowvshigh_raw}, the effect of saturation can be seen on the high gain channels.
+For signals in size below 400 ADC/mm the electronics are not saturated and therefore the high and low gain sections agree.
+The data starts to saturate and deviates above 400 ADC/mm in the high gain channels and eventually plateaus.
+Though these events saturate the high gain sections the low gain sections still have not saturated and provide the true dE/dx values.
+After applying the desaturation method, the correction the correlation between the high gain and low gain sections is significantly restored as seen in figure \ref{fig:lowvshigh_desat}.
+Judging by the data in the corrected correlation plot we believe the correction to at least about 900 ADC/mm.
+It seems the 1:1 correlation which was a plateau before is mostly restored and the increase in dynamic range in the high gain pads by at least a factor of 2x.
 
-\begin{figure}[ht]
-\includegraphics[width=\linewidth]{lowvshigh_desat}
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{dedxcompare_new}
 \caption{The corrected high gain dE/dx vs low gain dE/dx for ~??? events of collision data.  }
@@
 
-Comparing the low to high gain sections provides a direct comparison for determining the success of the extrapolation. The true goal of any correction would be to improve the PID. 
-\begin{figure}[h]
+Comparing the low to high gain sections provides a direct comparison for determining the success of the extrapolation. But, the true goal of any correction would be to improve the PID.
+
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{cocktail_raw}
@@
 
-\begin{figure}[ht]
+Looking firstly at the ideal case of the cocktail beam in figures \ref{fig:cocktail_raw} and \ref{fig:cocktail_desat} we note the very clean pronounced PID lines of several particle species.
+Once can clearly see the three  B$\rho$ settings of the fragment separator leading to three ovals around 1700 and two near 900 [MeV/c/Z].
+The tails of the PID lines leading away from these three ovals are resulting from the particle losing its initial energy by passing through the walls and other materials outside the main detector volume, therefore lowering their initial momentum.
+The red lines are the theoretical PID lines representing the most probable energy loss as given by Bichsel straggling functions after calibration to the experimental data.
+One can also get similar curves from Geant4.
+Looking at the uncorrected data in figure \ref{fig:cocktail_raw} we can see the effects of saturation.
+It seems the dE/dx values plateau and the PID lines deviate from their theoretical expectations starting around 400 ADC/mm as we also saw in figure \ref{fig:lowvshigh_raw}.
+After applying the desaturation method we see a stark difference in figures \ref{fig:cocktail_raw} and \ref{fig:cocktail_desat}.
+Most notable is the difference between He and Li particle species which suffer the most from saturation.
+Also a more subtle improvement of the lighter particles, (p,d,t), can also be seen in the PID values at lower momentum.
+
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{cocktail_desat}
 \caption{Corrected (desaturated) cocktail data.}
-\label{fig:data_desat}
+\label{fig:cocktail_desat}
 \end{figure}
 
-Looking firstly at the ideal case of the cocktail beam in figures \ref{fig:cocktail_raw} and \ref{fig:cocktail_desat} we note the very clean pronounced PID lines of several particle species.
-Once can clearly see the three  B$\rho$ settings of the fragment separator leading to three ovals around 1700 and two near 900 [MeV/c/Z].
-The tails of the PID lines leading away from these three ovals are resulting from the particle losing its initial energy by passing through the walls and other materials outside the main detector volume.
-Looking at the uncorrected data in figure \ref{fig:cocktail_raw} we can see the effects of saturation.The experimental PID deviates from the theoretical values given by the red curves.
-It seems the dE/dx values plateau starting around 400 ADC/mm as we also saw in figure \ref{fig:lowvshigh_raw}.
-After applying the desaturation method we see a stark difference in figures \ref{fig:cocktail_desat} and \ref{fig:data_desat}.
-Most notable is the difference between He and Li particle species which suffer the most from saturation.
-Perhaps the more subtle improvement of the lighter particles, (p,d,t), can also be seen in the PID values at lower momentum.
+Looking to the collision data in figures $\ref{fig:data_raw}$ and $\ref{fig:data_desat}$ we also see a similar result.
+Of course the collision data PID is much less clean than the ideal case of the cocktail beam, nevertheless we can see a similar improvement in the PID lines of the real data when comparing the raw to after the desaturation has been applied.
+Most notably in the separation of particle species at lower momenta.
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{data_raw}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
 \includegraphics[width=\linewidth]{data_desat}
@@
 \section{Conclusion}
-We have shown that one can easily construct the Pad Response Function from unsaturated experimental data.
-Using this function and applying a simple $\chi^2$ fit to the unsaturated tails of data, one can recover the saturated pad charge in the middle of the distribution because the true PRF is fixed by the anode wire geometry.
-We also demonstrated the success of this method by a direct comparison of dE/dx obtained in the high gain sections,to the true unsaturated dE/dx value obtained in the low gain sections which give the true unsaturated dE/dx value.
-This simple method gives significant improvement to the PID of particle and improves the dynamic range by a factor of at least 2.
+The Pad Response Function (PRF) is fixed by the anode wire geometry of the TPC and an experimental PRF can be calculated from unsaturated experimental data.
+It is also true that for all pad charges follow this PRF independent of the initial charge of deposited on the wire.
+By applying a simple $\chi^2$ fit to the unsaturated tails of pad distribution, one can recover the saturated pad's charges in the middle of this distribution.
+We also demonstrated the success of this method by a direct comparison of dE/dx obtained in the high gain sections, to the dE/dx value obtained in the low gain sections.
+Also significant improvement of the PID of both cocktail calibration and nuclear collision data is seen.
+Using this simple method we were able to extend the dynamic range by at least a factor of 2x.
 \section*{References}

### 2018-10-05 thesis/chapters/ch_tpc.tex
Before bills edits and bettys merge. New vector graphics

\chapter{Time Projection Chamber}
\section{Overview}

\subsection{Field cage}

\subsection{Wire planes}

\section{Gas Properties of P10 Gas}

\subsection{Drift velocity}

\subsection{Avalanche process}

Near the vicinity of the anode wires primary electrons under go acceleration and pick up kinetic energy. This kinetic energy can liberate secondary electrons in the gas which can then go on to also accelerate and liberate more secondary electrons resulting in an avalanche of electrons. Each electron in the avalanche process undergoes a random stochastic process where eventually they will terminate on the anode wire. Since the secondary electrons have no memory of the initial collision of the primary electrons, we can think of each secondary electrons as independent in the avalanche process. The total number of secondary electrons will give rise to the amplification factor, or gain, of the anode wires. 
Using Garfield++ a simple 2D model of the wire plane structure was created. In this simplified model the wire planes lie within two parallel plates which model the pad plane and cathode \ref{table:gar_model}.
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|}
\hline
 Element & Voltage (V) & y distance (cm) \\
\hline
Pad plane    & 0 & 0 \\
\hline
Anode plane  & 1470 & -.4 \\
\hline
Ground plane & 0 & -.8 \\
\hline
Gating grid  & -171. & -1.4 \\
\hline
Cathode  & -6036.7 & -51.02 \\
\hline
\end{tabular}
\caption{Geometry and voltages used in Garfiled++ model}
\label{table:gar_model}
\end{table}

### 2018-10-15 MSU-thesis-estee.tex
Split papers from thesis. now just thesis

% This is a template for use with the MSU Thesis class
% Vesion 2.8 2017/12/13
%
% Class options: 
%[PhD]	Doctor of Philosophy (default) 
%[DEd]	Doctor of Education
%[DMA]	Doctor of Musical Arts
%[MA]	Master of Arts
%[MS]	Master of Science
%[MAT]	Master of Arts for Teachers
%[MBA]	Master of Business Administration
%[MFA]	Master of Fine Arts
%[MIPS]	Master of International Planning Studies
%[MHRL]	Master of Human Resources and Labor Relations 
%[MMus]	Master of Music
%[MPH]	Master of Public Health
%[MPP]	Master of Public Policy
%[MSW]	Master of Social Work
%[MURP]	Master in Urban and Regional Planning 
%%
% Default is PhD
%
%
% This template has everything in the right order.
% Just add real content and you're done!
%
\documentclass[]{msu-thesis}
%
% for a prettier, but possibly non-compliant table of contents use the [mixedtoc] option
% for a plain table of contents use the [plaintoc] option
% for a horrendous looking, but possibly required table of contents, use the [boldtoc] option
%
% If you have large tables/figures that need to be in landscape mode, add the [lscape] option

% This is standard fontenc/inputenc for pdflatex
% If you use LuaLaTeX or XeLaTeX you should replace this with the fontspec package
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%
% If the thesis office requires Times, we'll give them Times
% You can experiment with other font packages here if you like.
% If you are using XeLaTeX or LuaLaTeX load the Times or Times New Roman font with \setmainfont
\usepackage{newtxtext,newtxmath} 
%
% Load any extra packages here
%
% You must specify the title of your thesis, your name, the field of study (not department), and the year
\title{The syntax and semantics of phonology}
\author{Joe Linguist}
\fieldofstudy{Linguistics} % This should be in sentence case
\date{2017}

% If you want a dedication page, specify the text of the dedication here and uncomment the next command.
%
%\dedication{This thesis is dedicated to someone.}
%
\begin{document}

% All the stuff before your actual chapters is called the front matter
\frontmatter
% First make the title page
\maketitlepage
% Next make the abstract
\begin{abstract}
% Your abstract goes here.  Master's 1 page max. PhD 2 page max.
\end{abstract}

% Force a newpage
\clearpage
% Make the copyright page. The Graduate School ridiculously prohibits you
% from having a copyright page unless you pay ProQuest to register the copyright.
% This should be illegal, but I didn't make up the rule.

\makecopyrightpage

% If you have a dedication page, uncomment the next command to print the dedication page
%
%\makededicationpage
%
\clearpage
% Your Acknowledgements are formatted like a chapter, but with no number
\chapter*{Acknowledgements}
\DoubleSpacing % Acknowledgements should be double spaced
Your acknowledgements here.
%
\clearpage
% We need to turn single spacing back on for the contents/figures/tables lists
\SingleSpacing
\tableofcontents* % table of contents will not be listed in the TOC
\clearpage
\listoftables % comment this out if you have no tables
\clearpage
\listoffigures % comment this out if you have no figures
%
% If you have a list of abbreviations/symbols it would go here preceded by a \clearpage
% See the class documentation and the Memoir manual for how to create other lists 
%
\mainmatter
%
% The next line removes the dots in chapter headings in the TOC
% May violate thesis office rules
%\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}

\include{chapters/ch_nucleareos}
\include{chapters/ch_piontheory}
\include{chapters/ch_embed}
\include{chapters/ch_theorymodel}
\include{chapters/ch_tpc}
\include{chapters/ch_software}
\include{chapters/ch_experiment}
\include{chapters/ch_analysis}
\include{chapters/ch_results}

%
% If you have pages that must appear in landscape mode, use the [lscape] documentclass option
% and enclose the pages in a {landscape} environment.
%\clearpage\pagestyle{lscape} % first clear the page and change the pagestyle
%\begin{landscape}
%
% your landscape table(s) or figure(s) here
%
%\end{landscape}
%\pagestyle{plain} % remember to change the pagestyle back to plain
%
%
% If you have appendices, they would go here.  
% Comment these lines out if you don't
% If you have more than one appendix, you need to use 
%   \begin{appendices}
%   \chapter{First appendix}
%   \chapter{Second appendix}
%   \end{appendices}
%ch_nucleareos.txt~
\appendix
\chapter{Your appendix}
%
\backmatter
% The next lines add the dots back into the References/Bibliography heading
% of the TOC.  Only uncomment this if you need to put the dots back in having removed them for Chapter headings.
%
%\addtocontents{toc}{%
%   \protect\renewcommand{\protect\cftchapterdotsep} {\cftdotsep}}
%
\makebibliographypage % make the bibliography cover page
% Bibliography can be single spaced
%
\SingleSpacing
%
% Your bibliography command here (e.g. \bibliography{your-bib-file}) if using natbib
%
% Remember that although the bibliography is single spaced, there needs to
% be a blank line between entries. This is set by your bibliography package
% If you are using natbib it is \bibsep; if using biblatex it's \bibitemsep
\end{document}

### 2019-06-09 Documentation/PionMultiplicityAnalysis.tex
Added Pion multiplicity document

\documentclass[12pt, letterpaper, twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{amsmath} 
\bibliographystyle{elsarticle-num}

\title{Pion Multiplicity Analysis}
\author{Justin Estee}
\date{June 2019}
\graphicspath{ {Images/} }

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

 \section{Introduction}
 The purpose of this document is to outline (in detail) the method used to get the total pion multiplicity and the multiplicity spectra.

 In our data we have several corrections we must account for to get the multiplicity spectra and the total pion multiplicity per event. This includes getting the total number of events measured, correcting for the efficiency, and correcting for the total solid angle not covered.

 \section{Center of Mass Transformation}
 After rotating the beam event by the beam angles provided by Jon, our system is now rotated event-by-event to align with the TPC z-axis. Now using the K.E. of the beam calculated from Jon (BigRIPS analysis), we can transform to the COM system by performing a Lorentz transformation to the COM by applying $-\beta$ to the Lorentz transform matrices in the ROOT class TLorentzVector. The calculation for $\beta$ is given in Eq. \ref{eq:beta} where $E_{Beam}=~K.E._{Beam} + M_{Beam}$ and $p_{Beam}$ is the momentum of the beam in the Lab frame.

 \begin{equation}
 \beta = \frac{\vert p_{Beam} \vert}{E_{Beam} + M_{Target}}
 \label{eq:beta}
 \end{equation}

 \section{Cuts on the Data}
The data has been cut in the good region of efficiency in the lab frame ($\theta_{Lab}$ $<$ 70 \&\& ($\phi$ $<$ 30 \&\& $\phi$ $>$ 330) $||$ 150 $<$ $\phi$ $<$ 210) and then transformed to the Center of Mass (COM) system as shown in Fig. \ref{fig:data}.

Several other quality cuts were made:
\begin{itemize}
\item ${}^{132}$Sn beam PID cut
\item On target vertex cut
\item Distance-to-vertex $<$ 20mm
\item Number-of-clusters $>$ 15
\item Multiplicity $>$ 50
\item No Gating Grid fast close events
\end{itemize}

These are the typical data cuts we take to get clean PID. All of these cuts are reasonable quality cuts and do not place too much of a bias on the data.

 \begin{figure*}
\begin{adjustbox}{max width=1.2\linewidth,center}
\centering
\subfigure[Lab frame]{\label{fig:a}\includegraphics[width=0.7\textwidth]{dataLab.png}}%
\subfigure[Transformed to Center of Mass]{\label{fig:b}\includegraphics[width=0.7\textwidth]{dataCOM.png}}%
\end{adjustbox}
\label{fig:data}
 \caption{$\pi^+$ data within cut, $\theta_{Lab}$ $<$ 70 \&\& ($\phi$ $<$ 30 \&\& $\phi$ $>$ 330) $||$ 150 $<$ $\phi$ $<$ 210 }
 \label{fig:data}
\end{figure*}

 \section{Efficiency Correction}
 \subsection{Embedding MC Tracks}
 We are embedding simulated MC tracks (which take into account most of the TPC detector physics) into real data events. This way we can incorporate effects such as saturation, track multiplicity, detector bias, detector acceptance, software inefficiency, and other effects that are otherwise too hard to directly simulate. A measured known MC track is digitized and embedded and the whole event is reconstructed. The embedded hits are tagged and tracked through the full software reconstruction. We embedded into real data events that are in the vertex cut and the beam PID cut condition (i.e. ${}^{132}$Sn beam PID cut from Jon's BigRIPS analysis).

  After the event is reconstructed there are three possible outcomes, there are no embedded tracks reconstructed, there is one embedded track found, or there are several embedded tracks found. The first two cases are very straightforward (either the embedded track is found or not), in the latter case is the case of track splitting. Several effects in the TPC (both physical and in software) which can cause a track to split. All reconstructed embedded tracks are saved in an array which can be studied in detail.

 For the purpose of this paper we are not interested in track splitting and we need to make a choice on which track measured corresponds to input MC track, without biasing the results. It would be easiest to just match the known MC input momentum with the final track momentum, but this would bias the results to be more like the MC input momentum. Since we are interested in studying what happens to the momentum the track with the smallest distance-to-vertex is taken as the track which is most likely resembles the input MC track. This makes logical sense since all the MC track must originate from the TPC vertex.

 Once we have the reconstructed track, combined with the information for the MC input track, we can approach the situation of calculating efficiency.

 \subsection{Method to Calculate Efficiency}
 The efficiency calculation is done bin-by-bin. A uniform distribution of pions in momentum, $p$, and angles $\theta$, and $\phi$ is input into the embedding analysis. A database of the input and output tracks from the embedding is made. The efficiency is calculated simply by subdividing the phase space ($p$,$\theta$,$\phi$) into several bins. 
 Then calculating the efficiency is simply looping over the database from the embedding results and calculating the ratio of tracks detected divided by tracks input for each bin, as in Eq. \ref{eq:eff_cal}. Where $N(p,\theta,\phi)_{detected}$ is the total number of tracks detected and $N(p,\theta,\phi)_{total}$ represents the total number of tracks input.

 A track which is ``$detected$" is classified as a track which was reconstructed in the embedding process and survives the cuts placed on the data set. This could mean any cut, $dE/dx$, distance to vertex, multiplicity, or cluster number. The total number of tracks is the total number of tracks within the multiplicity cut and the phase space region of interest.

 \begin{equation} 
 \label{eq:eff_cal}
 \epsilon_{(p,\theta,\phi)} = \frac{N(p,\theta,\phi)_{detected}}{N(p,\theta,\phi)_{total}}
 \end{equation}

\subsection{Efficiency Corrected Spectra}

 \section{Correcting for the Integrated Solid Angle}
 Since the TPC is a fixed target, "box`` shaped detector, the angular coverage does not cover the full solid angle of $4\pi$. Because of the quality angle cuts placed in Fig~\ref{fig:data}, only around 20\% of $4\pi$ (in the COM) is covered. Furthermore, since the angular coverage also depends slightly on the momentum of the pion, we must also correct for the solid angle covered by the TPC for a given momentum.

 Using a uniform distribution in $\theta$,$\phi$, and momentum $p$, as an input to the MC embedding procedure, we can determine the angular coverage of the TPC by putting the same cuts in the Lab frame (as Fig.~\ref{fig:data}) then transforming to the COM system to see the angular coverage of the TPC.

 Since the MC embedding has no information of the beam, remember we are just evaluating efficiency in different lab angles, we assume a value of $\beta = -.364$ for transforming the MC embedded tracks to the COM frame. The momentum acceptance calculated by Jon shows that the $\beta$ distribution is sharply peaked, with little variation, around this value.

For this analysis I have not put in the angle rotation for the MC track coverage calculation. Seen in Fig.~\ref{fig:cutcov} is the coverage in the COM system for the given Lab frame cuts in Fig.~\ref{fig:data}. The coverage depends on the momentum of the system but for these cuts the $\pi{-}$ and $\pi^+$ coverages are almost the same for a given momentum.

\begin{figure*}
\begin{adjustbox}{max width=1.2\linewidth,center}
\centering
\subfigure[70 MeV/c $\pi^{+}$]{\label{fig:a}\includegraphics[width=0.35\textwidth]{70Mev}}%
\subfigure[310 MeV/c $\pi^{+}$]{\label{fig:b}\includegraphics[width=0.35\textwidth]{310MeV}}%
\subfigure[410 MeV/c $\pi^{+}$]{\label{fig:c}\includegraphics[width=0.35\textwidth]{510MeV}}%
\end{adjustbox}
\label{fig:cutcov}

\medskip

\begin{adjustbox}{max width=1.2\linewidth,center}
\centering
\subfigure[70 MeV/c $\pi^{-}$]{\label{fig:d}\includegraphics[width=0.35\textwidth]{pim70MeV}}%
\subfigure[310 MeV/c $\pi^{-}$]{\label{fig:e}\includegraphics[width=0.35\textwidth]{pim310MeV}}%
\subfigure[410 MeV/c $\pi^{-}$]{\label{fig:f}\includegraphics[width=0.35\textwidth]{pim510MeV}}%
\end{adjustbox}
\caption{$\theta_{Lab}$ $<$ 70 \&\& ($\phi$ $<$ 30 \&\& $\phi$ $>$ 330) $||$ 150 $<$ $\phi$ $<$ 210 cut coverage in COM coordinates. From MC embedding of uniform distribution in $\phi$, $\theta$, and $p$}
\end{figure*}

 We now can calculate the solid angle covered in the COM system by summing the total solid angle covered by each bin as given in Eq.~\ref{eq:solid}, where $\theta_H$ and $\theta_L$ are the high and low limits of the bin in $\theta$, and $\phi_H$ and $\phi_L$ are the high and low limits in $\phi$ respectively.

 \begin{equation}
 \Omega = \sum (\cos(\theta_L) - \cos(\theta_H))(\phi_H - \phi_L)
 \label{eq:solid}
 \end{equation}

 This is done for each momentum bin, which is assigned a total angular coverage of $\Omega_p$, for each momentum value, $p$. It is a good assumption that in the COM system the $\pi$ distribution is nearly isotropic. Correcting for the areas of the solid angle coverage we correct the counts in each momentum bin, $N_p$, by ${4\pi/\Omega_p}$, since this is the inverse of the fraction of space we measure in the TPC.

 \subsection{Solid Angle and Efficiency Corrected Spectra}

 \section{Calculation of Impact Parameter}
 I refer to Jon for details of how he gets the impact parameter from cross section measurements. What he and Genie has provided us is with a tool to calculate the impact parameter estimate from the multiplicity distribution. For a multiplicity cut of $>$ 50, this corresponds to an impact parameter range of 0 $<$ b $<$ 3.5 [fm] with an average of 2.4 [fm]. This is important for later where we will calculate the total number of participants for these numbers. The range in $b$ will help us to estimate our error bar on the total number of participants.

 \section{Calculation of Number of Participants $A_{part}$}
 The total number of particles in the participant region of the collision is calculated using the Glauber model where each nuclei is given a random isotropic distribution in $\phi$ and $\theta$, where a random number between 0 and 1 is chosen, $rand[0,1]$, and input into Eq.~\ref{eq:randphi} and Eq.~\ref{eq:randtheta}, to generate random isotropic points in spherical angles. The radius is randomly sampled from $4\pi r^{2}\rho(r)$, where the absolute normalization is not important. $\rho(r)$ is given in Eq~\ref{eq:radial} where $R=5.45$ and $a=.4835$ for ${}^{132}$Sn and $R=5.38$ and $a=.4835$ for the ${}^{124}$Sn, as plotted in Fig.~\ref{fig:radial}.

 \begin{equation}
 \phi = rand[0,1] * 2\pi
 \label{eq:randphi}
 \end{equation}

 \begin{equation}
  \theta = (2*rand[0,1] - 1)
  \label{eq:randtheta}
 \end{equation}

 \begin{equation}
 \rho(r) = \rho_{o} \frac{1}{1+ exp(\frac{r-R}{a})}
 \label{eq:radial}
 \end{equation}

  \begin{figure*}
\includegraphics[width=\textwidth]{density.png}
 \caption{Radial nucleon density.}
 \label{fig:radial}
\end{figure*}

 The nuclei are initialized to be separated by a distance of $b$, center to center, as given by the impact parameter you choose. The nuclei are then collided by small time steps and at each step the distance to each pair of nucleons, $d$, is calculated. If ,$d < \sqrt{ \frac{ \sigma_{nn} }{\pi}}$, then the nucleons are said to have collided (and are therefore are "participants``), where $\sigma_{nn}$ is the nucleon-nucleon free cross section, which is a constant 33$mb$ (or 3.3 $fm^2$), for a range of incident energies we cover in our experiments. Thus the distance in which nucleons are said to collide is about $d < 1fm$. I then measure the total number of particles which collided to get the number of participants, $A_{part}$.

 Since this is a MC method it is averaged over several MC trials to arrive at the average number for $A_{part}$. Table~\ref{tb:apart} shows a list of the $A_{part}$ for various impact parameters. The total number of MC Glauber model collisions was 1000 to reduce the statistical error bar in the calculation.

\begin{table}[!htb]
\begin{center}
 \begin{tabular}{||c c||}
 \hline
 ${}^{132}$Sn + ${}^{124}$Sn \\ [0.5ex] 
 \hline\hline
 b [fm] & $A_{part}$ \\ [0.5ex]
 \hline\hline
 0 &  255.6 \\
 \hline
 1 & 254.0 \\ 
 \hline
 2.4 & 244.2\\
 \hline
 3.45 & 232.0\\
 \hline
 4 & 220.6 \\
 \hline
 5 & 197.4 \\ 
 \hline
 6 & 158.0 \\ [1ex] 
 \hline
\end{tabular}
\end{center}
\caption{$A_{part}$ for a given impact parameter b in the Glauber model}
\label{tb:apart}
\end{table}

Shown in Fig.~\ref{fig:coll} is the Glauber collision (projected onto the x-z plane) for the initial, middle, and final stages of the collision.

\begin{figure}[!hbt]

\begin{adjustbox}{max width=\linewidth,center}
\centering
\subfigure[Initial stage]{\includegraphics[width=0.35\textwidth]{step_0.png}}%
\subfigure[Middle stage]{\includegraphics[width=0.35\textwidth]{step_50.png}}%
\subfigure[Final stage]{\includegraphics[width=0.35\textwidth]{step_99.png}}%
\end{adjustbox}
\caption{Glauber model in initial, middle, and final stage of the collision. Light colors denote spectators and vivid colors participants. Blue denotes target and red denotes projectile.}
\label{fig:coll}

\end{figure}

\clearpage

\section{Pion Multiplicity}
In \cite{fopi}, the total pion multiplicity is defined as $M(\pi)=1.5(M(\pi^-)+M(\pi^+))$. Integrating the pion efficiency (and solid angle corrected) pion spectra in FIGURE, we can get the total pion multiplicities for ${}^{132}$Sn.

\begin{table}[!htb]
\begin{center}
 \begin{tabular}{||c c c c c||}
 \hline
  & ${}^{132}$Sn + ${}^{124}$Sn & ${}^{108}$Sn + ${}^{112}$Sn & ${}^{124}$Sn + ${}^{112}$Sn & ${}^{112}$Sn + ${}^{124}$Sn\\ [0.5ex] 
 \hline\hline
  $M(\pi^-)$ & 2  & 2 & 2 & \\ [0.5ex]
 \hline\hline
 $M(\pi^+)$ &  255.6  &  2 & 2 & \\
 \hline
\end{tabular}
\end{center}
\caption{Pion multiplicities.}
\label{tb:pionmult}
\end{table}

\section{Comparing With FOPI Data}

\subsection{Correction for $A_{part}$ Effect on $M(\pi)$}

\begin{figure}[!hbt]

\begin{adjustbox}{max width=\linewidth,center}
\centering
\subfigure[With $A_{part}$ correction]{\includegraphics[width=0.5\textwidth]{multiplicitypion.png}}%
\subfigure[No correction]{\includegraphics[width=0.5\textwidth]{multiplicitypion_scaled.png}}%
\end{adjustbox}
\caption{Pion multiplicity before and after correction, compared with FOPI data.}
\label{fig:mult}

\end{figure}

\subsection{N/Z Effect for $\pi^-/\pi^+$}
It is well known that the $\pi^-/\pi^+ \propto (n/z)^2$, at low energies. Because of this dependence we need to correct the ${}^{197}$Au + ${}^{197}$Au FOPI data (which has $N/Z= 1.49$), and extrapolate the  $\pi^-/\pi^+$ to $N/Z=1.56$, for the ${}^{132}$Sn~+~${}^{124}$Sn system. Plotted in Fig.~\ref{fig:nzdepend} is the N/Z dependence published in the FOPI data set \cite{fopi}. The measured data points are given in the blue circles. The paper cites a linear fit was applied to the data which I have plotted at the black dotted line. The paper also says a linear extrapolation was performed to extrapolate to $N/Z=1.56$. One can see the blue "X`` are the points which they calculate, which matches the linear extrapolation. For .8 and 1 A GeV the linear extrapolation seems reasonable but for .4 A GeV we know the dependence should be $(n/z)^2$, as is clear from the trend of the data. I extrapolated the $N/Z=1.56$ for .4 AGeV which is shown as the green point in Fig.~\ref{fig:nzdepend}, and also in .8 and 1 AGeV.

\begin{figure}[!hbt]
\centering
\includegraphics[width=\textwidth]{NZDependence.png}
\caption{$\pi^-/\pi^+$ N/Z dependence }
\label{fig:nzdepend}
\end{figure}

 In Fig.~\ref{fig:pionratio}, the total pion ratio extended to $N/Z=1.56$ is plotted in the blue and red points, the blue points representing the pion ratio from the FOPI experiment (with their fit), and the red points representing the error in extrapolating .4 AGeV which I described above. The ${}^{132}$Sn+${}^{124}$Sn is plotted in green. We can see a good agreement (considering the red points are not data but extrapolated data), also considering the correct extrapolation at .4 AGeV is made. Error bars to come.

\begin{figure}[!hbt]
\centering
\includegraphics[width=\textwidth]{pionratio.png}
\caption{$\pi^-/\pi^+$ for $N/Z=1.56$}
\label{fig:pionratio}
\end{figure}

\newpage
\bibliography{mybibfile}

\end{document}

### 2019-06-09 chapters/ch_data_analysis.tex
Added Pion multiplicity document

\chapter{Data Analysis I: Calibration and Corrections}
\section{Software}

The S$\pi$RITROOT software is modular tasked based code based on the FAIRROOT package written in C++ \cite{fairroot}. The main tasks in the S$\pi$RITROOT software reconstruction are:
\begin{itemize}
  \item Decoder task
  \item Pulse Shape Algorithm (PSA Task)
  \item Helix Track Finding Algorithm
  \item Clustering Algorithm
  \item Track Fitting (GENFIT package)
  \item Vertex Fitting (RAVE package)
\end{itemize}

The decoder task converts the binary data file into a container class which maps the electronics channels into the corresponding pads and (x,z) coordinates.

There may be several pulses in a pad coming from two tracks passing under the same pad separated  by arrival time. Using an expected pulse shape the PSA task fits the signal pulses within a pad, giving the arrival time of the drifted electrons from each particular track. The height of the fitted pulse is proportional to the total charge of that event, Q and the y-coordinate is calculated as $y = v\cdot t_0$ where $v$ is the drift velocity and $t_0$ the arrival time. Combining the information from these first two tasks, (x,y,z,Q), we construct what is called a "hit".

 The Helix Track Finding Algorithm finds the collection of hits belonging to one track out of all the hits in an event. The hits within a track are then reduced into clusters. A cluster's position is the average position of the hits within a cluster, with the total charge of the cluster being the sum of the hits charges.

 A tracks average position is estimated by the cluster's average position. The clusters are then fitted in the GENFIT track fitting package \cite{genfit}, giving the final momentum of the track. A final vertex of the event is fitted from all tracks using the package RAVE \cite{rave}.

\paragraph{Definition of clustering}

A brief description of the method of clustering is illustrated in Figure \ref{fig:topview}. It is impractical to cluster in both the x and z-axis and we only cluster the hits along one axis. The three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis and the upper three are along the z-axis, as shown by the bolded pads for one of the clusters in each direction.

 The clustering direction depends on the angle  of the track with respects to the x-axis, defined as $\theta$. For example, a track going along the z-axis the crossing angle is defined as $90^{\circ}$, and a track going along the x-axis defined as $0^{\circ}$. In the case that the crossing angle is $45^{\circ} < \theta \leq 90^{\circ} $ the clustering direction is along the x-axis. For $0^{\circ} < \theta \leq 45^{\circ}$ it is along the z-axis.

 The position along the clustering direction is calculated by weighting the individual hit's positions by their charges $q_i$ and getting the mean value. The other direction is set to the center of the pad. For example if we are clustering along the x-axis for a cluster, the z-position is set to the center of the pad in the z-direction and vice versa.

Clustering in this way gives us better position resolution for calculating the position of each cluster. You could imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined. By clustering in the direction most perpendicular to the track, we get a better position resolution.

\begin{figure}[H]
\includegraphics[scale=.5]{top_view_helix_ext.pdf}
\caption{Cartoon graphic of a top down view of a fit to a track passing through several pads. The bolded pads and the charges $q_i$ represent the hits belonging to that pad and the clusters of the track representing the average position of the track. The three clusters at the bottom are clustered in the x-direction and for the upper three clustered in the z-direction. The estimate of the position of the avalanche is given by the track fit and the position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
\label{fig:topview}
\end{figure}

\begin{figure}[H]
\includegraphics[width=\linewidth]{padsat_Large}
\caption{A cartoon illustration of the charge distribution resulting from an electron avalanche on one wire and the projections of the distribution onto the two axis $\rho(x)$ onto the x-axis and $\rho(z)$ onto the z-axis. The orientation of the wire planes is flipped upside down to display the perspective better.}
\label{fig:prf}
\end{figure}

\section{Calibrations}

\paragraph{Electronics calibration}

\paragraph{Anode gain calibration}

\section{Corrections}
\paragraph{Extending the dynamic range of Electronics}

\section{Monte Carlo Simulation}
We use Geant4 as and event generator for performing Monte Carlo Simulations in the TPC. A scale model of field cage, front window, front window frame, pad plane, and aluminum top plate are modeled. The correct materials are used as well as the field cage is modeled with P-10 gas at a density of ????. By using Geant4 we can also input the the magnetic field map of the SAMURAI dipole magnet (as calculated by the SAMURAI group via a TOSCA simulation).  In this way any particle type may be studied and the full interactions (scattering, decay, energy loss, path taken, etc.), are accounted for. The output of this simulation is a series of energy loss points which contain the amount of energy lost in $keV/cm$ and the position in $(x,y,z)$.

Separate software tasks model the converting energy loss into electrons, the drifting of electrons, the avalanche process, and the electronics response.

\begin{figure}
\includegraphics[scale=.01]{place}
\caption{A summary of all the effects modeled in the TPC MC simulation.}
\label{fig:place}
\end{figure}

\section{Drift Task}
The drift task takes the energy loss points calculated from Geant4 and converts them into electrons.  and then modeling their drift behavior in the field cage. As discussed previously the various processes[CITE To formula about drifting electronc]

 A full microscopic treatment of the stochastic nature of each electron would be too cumbersome; most of the properties of the electrons are described by macroscopic quantities can be described by 
The charge of each MC point is converted to the total number of electrons liberated in the P-10 gas. This is a well understood property of proportional counters and is stable over a wide range of velocities and particle types [CITE BOOK]. The conversion factor of P-10 can be calculated by considering the partial volumes of each component of the gas.  \ref{eq:kev2el}.

\begin{equation}
Number of e^{-} = 28 keV/cm
\label{eq:kev2el}
\end{equation}

\section{Pad Response Task}
After the avalanche process of the previous task, the total charge of the event is split over several pads defined by the Pad Response Function (as discussed in section ????). As shown in Fig.~\ref{fig:onepad}, there are three wires that lie directly above a given pad. The $z$ coordinate of the avalanche can only be one of these three wires, where as the $x$ coordinate can be any value along the wire. The functional form of the software PRF (given in Eq.~\ref{eq:softPRF}), was tuned to match the experimental data. Shown in Fig.~\ref{fig:mcdataPRF}, we see the tuned software PRF can match the experimental PRF from data over several crossing angles (as mentioned in Chapter ???? governs the shape of the PRF).

\begin{equation}
PRF(x,z) = \frac{1}{2\pi\sigma_z\sigma_x}\exp^{\frac{-{(x-x_o)}^2}{2\sigma_x^2}}\exp^{\frac{-{(z-z_o)}^2}{2\sigma_z^2}}
\label{eq:softPRF}
\end{equation}

\begin{figure}
\includegraphics[scale=.005]{place}
\caption{A cartoon of the wires over one pad. }
\label{fig:onepad}
\end{figure}

\begin{figure}
\includegraphics[scale=.005]{place}
\caption{Comparison of MC and data PRF}
\label{fig:mcdataPRF}
\end{figure}

\section{Electronics Task}
The electronics task takes the total charge on each pad and simulates the electronics response, converting electronics into ADC channels. Accounting for the pedestal, the measured output in ADC channels (for a given gain setting) is given in Eq.~\ref{eq:etoADC}.

\begin{equation}
\mathrm{1 ADC }= \frac{ADC_{\mathrm{Max}} - ADC_{\mathrm{Pedestal}}}{G*f_c}
\label{eq:etoADC}
\end{equation}

\section{Track Embedding}
Track embedding is the process of taking a simulated MC track from Geant4 and embedding its response into a real data event. After reconstructing this new embedded event we match the input MC track embedded to its corresponding final reconstructed track.  By doing so we can evaluate the response of the entire TPC system to any given input value. The TPC system is composed of three major components (each which can introduce errors and or biases) the software, the detector, and the experimental setup.

As discussed in [SOFTWARE CHAPTER] the software is composed of several task, each which introduce some error. Table {REF TABLE] listing some of the errors each task may introduce illustrating how difficult propagating the errors would be through each system.

The detector system itself introduces errors related to the physical processes of the measurement itself. To address this we model the TPC (and its materials) in a Geant4 simulation which provides an accurate description of various interactions of a particle traversing the materials (including the gas volume) of the TPC.

 software is the most straight forward; let the software routine process an input and measure the result. Understanding the measurement requires modeling the physics involved in the theory and operations of TPC's and the  electronics. The experimental setup itself is quite large and complex, several ancillary detectors such as the Kyoto multiplicity array, Krakow veto array, Active veto array, beam identification detectors, etc. Even if a full accurate model could be constructed the complex trigger logic of the DAQ system would be impossible to model. If we notice that the biases and errors of the entire experimental setup is contained in the measured experimental data. Therefore, by inputting the MC data into a real experimental event (and measuring the output of that MC track) we can estimate the errors of the experimental setup.

The software analysis routine and the bias introduced by the trigger settings of the experiment introduce systematic errors in the reconstruction of tracks

\section{Efficiency Corrections}
Since the \spirit TPC is a fixed target experiment it's angular coverage is certainly not 4$\pi$. Because the target is several cm away from the widow of the field cage the geometric acceptance is not even 2$\pi$. The rectangular design complicates the calculation of the geometric acceptance, or the efficiency.

### 2019-06-09 chapters/ch_data_analysis_2.tex
Added Pion multiplicity document

\chapter{Data Analysis II: Extracting Physics}
\section{Beam Particle Identification}
\section{Event Mixing}
\section{Reaction Plane Determination}

### 2019-06-09 chapters/ch_experiment.tex
Added Pion multiplicity document

\chapter{Experiment}

\section{S$\pi$RIT TPC Overview}
The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (S$\pi$RI TPC) is a Multi-Wire Proportional Counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiments.  The TPC is enclosed in a thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach the arrays of scintillating bars detectors on the sides and downstream of the TPC. The S$\pi$RI TPC was developed to fit inside the Samurai dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC. More detail and specifications of the Samurai dipole magnet are given in \cite{samurai}.

A target ladder allowed for up to 5 fixed targets to be mounted. A ACME worm gear allowed for the x-axis motion for changing the targets. The motion of this worm gear was translated through the target motion feed-through by several brass gears and non-magnetic gear boxes. The motion of the target ladder could be controlled by hand or by operation of the drill. The targets were mounted on stand-offs on the target ladder and also had z-axis motion. This allowed for the targets to be positioned as close as possible to the thin window of the field cage, maximizing the geometric acceptance.

The electronics were mounted to the aluminum top plate. Several aluminum ribs were mounted to the top of the plate to bring structural rigidity. Because of this the top plate flatness was within 150$\mu$m across the whole top plate as measured by a laser position system. The charge sensitive pads of the pad plane were etched into several circuit boards were recessed and glued to the bottom portion of the top plate. Vias through the pad plane circuit boards brought the signal traces from the pads to surface mount pads on the other side of the boards. Several holes cut through the top plate allowed for the interface cables of the electronics to be connected to these surface pads.

Just below the pad plane were a set of three wire planes; the gating grid, ground, and anode wire planes. The detailed function of these wires will be explained later but they served to separate the boundary between the drift volume and the avalanche volume.

The front and sides of the field cage were assembled from 8 independent rigid circuit boards. The downstream window served as the downstream wall of the field cage all though it was was a large polycarbonate frame in which a removable kapton window could be installed. This thin window minimized the scattering of exiting neutrons and charged particles for downstream detectors.

The voltage step down takes the high voltage of the field cage cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.

\begin{table}
 \begin{tabular}{||c c c c||} 
 \hline
 \multicolumn{4}{|c|}{S$\pi$RIT TPC Overview} \\
 \hline
 Pad Plane Area & 1.3 m x 0.9 m & Gas Gain & 1000\\
 \hline
 Pad Size & 1.2 cm x 0.8 cm & Drift Velocity & 5.5 cm/$\mu$s \\
 \hline
 Number of pads & 12096 (112x108) & E-field & 135 V/cm  \\
 \hline
  Gas composition & 90\% Ar + 10\% CH${}_4$ & Multiplicity limit & 200  \\
 \hline
 Drift length & 50 cm & dE/dx range & Z=1-8, $\pi$, p,d,t,He,Li-O \\ [1ex] 
 \hline
\end{tabular}
\caption{An overview of the properties of the S$\pi$RIT TPC.}
\label{tb:spiritoverview}
\end{table}

\section{Considerations when constructing a TPC}
Several considerations went into the construction of the S$\pi$RI TPC which I wish to summarize and document here. All materials and glue of the TPC were selected as low out-gassing materials. Several materials that are common place in nuclear labs, from vacuum grease to o-rings, out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time. The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments. If a material we wished to used was not on these lists we placed the material in a clean chamber with the counter gas and flowed this counter gas through a small proportional counter making sure the gain did not drop at high collection rates when exposed to a high rate alpha Americium source.

Sparking
Two volumes of gas.

\section{Field Cage}

The field cage was designed to hang from the top plate and therefore needed to be of a lightweight construction. Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors. Therefore instead of a downstream wall, a large thin exit window was constructed. The cathode was constructed of an aluminum honeycomb laminate. Two sheets of ??? aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.

\begin{figure}[H]
\includegraphics[width=\linewidth]{fc_explode}
\caption{Exploded view of the S$\pi$RIT TPC}
\label{fig:fc_explode}
\end{figure}

The field cage was constructed from several panels of printed circuit boards (PCBs). The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires [CITE]. The halogen free material chosen was Rodgers ????. We build the TPC with the option to run explosive gases such  hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure. While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.

The front of the field cage was made of two PCBs and each side was constructed of three PCBs. The exit window was a 10$\mu$m thick Kapton window with evaporated aluminum strips on the inside and outside; mounted on a polycarbonate frame.

\section{Voltage Step Down}

To lower the risk of a high voltage spark around the cathode region a voltage step down was developed. A series of 7 concentric copper rings with 20 MOhm resistors between them steps the cathode voltage down to ground. This minimizes sharp transitions of the electric field and therefore sparking.

\section{Wire Planes}

\section{Pad Plane}

\section{Electronics}

Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.  Short cables transmit the signals from the pads to the inputs of the AGET chips. Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz. Four AGET chips are mounted on one AsAd (ASIC and ADC) motherboard. The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range, and the ADCs on each AsAd board provides 12 bit resolution. The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns. In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets). The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used. The variations in the electronics were calibrated by measuring the response of each channel to a injected reference pulse, covering the full dynamic range of each channel.

\begin{table}
 \begin{tabular}{||c c c c||} 
 \hline
 \multicolumn{4}{|c|}{GET Electronics Settings} \\
 \hline
 GET settings & All options & ${}^{132}$Sn + ${}^{124}$Sn & ${}^{124}$Sn + ${}^{112}$Sn \\
 \hline
 ADC bit range      & 14 bits &  &  \\
 \hline
 Sampling frequency & 1 - 100 MHz &   &    \\
 \hline
 Dynamic range      & .12,.24,1.0,10 pC & & \\
 \hline
 Peaking time       &  69,117,232,501,720, 1014 ns &  &   \\
 \hline
 Time bucket        & 512 & 270 & 270  \\
 \hline
\end{tabular}
\caption{Summary of the optional settings of the GET electronics and those used during the 20?????? experimental runs}
\label{tb:spiritoverview}
\end{table}

\section{S$\pi$RIT at RIBF}

\section{Kyoto Multiplicity Trigger}
\section{Krakow ?????? (KATANA)}
\section{Ancillary Detectors }
\section{Radio Isotope Beam Factory (RIBF) Facility }

SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
\begin{figure}[H]
\includegraphics[width=\linewidth]{SAMURAI-beamline.png}
\caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
\label{fig:sambl}
\end{figure}

\section{Experimental setup}
\section{Trigger Condition}

### 2019-06-09 chapters/ch_intro.tex
Added Pion multiplicity document

\chapter{Introduction}
\section{From Nuclear Forces to the Equation of State}
\section{Phases of Nuclear Matter}
\section{The Nuclear Equation of State}
\section{Previous Constraints}
\section{Motivation for building the S$\pi$RIT TPC}

### 2019-06-09 chapters/ch_results.tex
Added Pion multiplicity document

\chapter{Results}
\section{Comparison to Theory}
\section{$\pi$ Yields}
\section{$\pi^-$/$\pi^+$ Ratio}
\section{Constraint on the Symmetry Energy}

### 2019-08-07 Documentation/PionMultiplicityAnalysis.tex
Adding better tables and gas section. Macros folder added

@@
 
-\subsection{Efficiency Corrected Spectra}
+  \subsection{Efficiency Correction}
+The efficiency is corrected only for bins of $\theta$ and not bins in $\phi$ to improve statistics.
+The efficiency function is calculated only for the angular data cut region shown later.
+The zero count region is not corrected for and will be corrected by doing a solid angle correction later.
+Figure~\ref{fig:eff} shows the "effective`` efficiency for the pion vs momentum.
+Effective means we just take the efficiency corrected spectra and divide by the non-efficiency corrected spectra.
+The efficiency itself depends on  three parameters ($p$,$\theta$,$\phi$) and we are reducing it down to one parameter plot vs $p$.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{effEff.png}
+ \caption{Effective efficiency for pions}
+ \label{fig:eff}
+\end{figure}
 
@@
 Correcting for the areas of the solid angle coverage we correct the counts in each momentum bin, $N_p$, by ${4\pi/\Omega_p}$, since this is the inverse of the fraction of space we measure in the TPC.
+Figure~\ref{fig:solid} shows this $d\Omega$ made for each momentum bin, as a fraction of $4\pi$.
 
- \subsection{Solid Angle and Efficiency Corrected Spectra}
+\begin{figure}[!hbt]
+\includegraphics[width=\textwidth]{omega.png}
+ \caption{Solid angle correction for each momentum bin.}
+ \label{fig:solid}
+\end{figure}
+
+ \subsection{Efficiency Corrected and Solid Angle Corrected Spectra}
+ Figure~\ref{fig:spectra} shows the efficiency corrected, and solid angle corrected, spectra.
+The efficiency corrected pion multiplicities are $\pi^- = .63$ and $\pi^+ = .12$ and the uncorrected multiplicities are $\pi^- = .63$ and $\pi^+ = .12$
+
+\begin{figure}[!hbt]
+\includegraphics[width=\textwidth]{spectra.png}
+\caption{Efficiency and solid angle corrected spectra.}
+\label{fig:spectra}
+\end{figure}
+
+ \clearpage
 
@@
 \subsection{Correction for $A_{part}$ Effect on $M(\pi)$}
+In \cite{fopi}, there is a small correction for the dependence of the multiplicity on the number of participants.
+I will leave it to the reader to read the paper.
+The function used to correct is given in Eq.~\ref{eq:corr}, where $\tau = .86$, the correction factor being, $f=1.05$.
+We must scale our answer by this estimate to get compare with the  ${}^{197}$Au + ${}^{197}$Au FOPI data.
+The only issue here is this scaling law is not for the beam energy we are at.
+This is for 1 AGeV but roughly speaking it should not depend so much on beam energy.
+
+\begin{equation}
+f = \frac{256}{394 *.9}^{\tau -1}
+\label{eq:corr}
+\end{equation}
+
+Figure~\ref{fig:mult} shows the corrected (and uncorrected) pion multiplicity, as calculated from Table~\ref{tb:apart} and Table~\ref{tb:pionmult}, plotted against the FOPI data.
+The blue dotted curve is the fit from the published data, and the red line fit is with our new point added in.
+The trend is in good agreement with our measured data.
+The error bar on the scaled multiplicity is coming from the error bar in $A_{part}$ from the uncertainty in the range in impact parameters probed.
 
 \begin{figure}[!hbt]
-
 \begin{adjustbox}{max width=\linewidth,center}
@@
 \end{figure}
+
+\clearpage
 

### 2019-08-07 chapters/ch_data_analysis.tex
Adding better tables and gas section. Macros folder added

@@
 
-\section{Calibrations}
+\section{Calibrations and Corrections}
+Cocktail Calibration
 
-\paragraph{Electronics calibration}
+\subsection{Electronics calibration}
+Pulse the ground grid. 
+Picture of calibration of channels before and after.
 
-\paragraph{Anode gain calibration}
+\subsection{Anode gain calibration}
+Picture of low vs high gain channels and fit
 
-\section{Corrections}
-\paragraph{Extending the dynamic range of Electronics}
+\subsection{Extending the dynamic range of the Electronics}
+Add paper here.
+
+\subsection{Space Charge Corrections}
+Overview
+Discuss the relevant time scales, drift lengths, magnitudes, and locations of space charge
+Figure of cartoon sheet charge
+Figure of beam path locations for different beam types
+Table of energy loss tables for energies and beam types
+Reference appendix for poisson solver 
+Tables for ion drift velocity in P-10 Gas reference Sauli
+Figure of sDAC or POCA 
+Figure of cartoon of what is happening to tracks
+Figure of correction map in TPC and MC map 
+Figure of before and after correction BDC vs reco momentum
+Figure of track residuals before and after?
+
+In theory, a TPC functions with the electric and magnetic fields parallel to each other.
+In this way the electrons move opposite to the electric field winding tightly around the magnetic field lines, reducing transverse diffusion in the process.
+In practice, due to the finite size of the dipole magnet and field cage, there are traverse components to both the electric and magnetic fields.
+These transverse components introduce drift velocities in the transverse directions, causing a shift in the measured cluster positions of the track.
+Thereby introducing systematic shifts in the calculation of the momentum and the vertex calculation.
+
+Most of the time the beam does not undergo a nuclear collision with the target and passes through the TPC drift volume.
+The KATANA array threshold was set to veto such events, ensuring the gating grid remain closed to prevent the large amount of charge deposited by the beam into the avalanche region.
+While the electron drift velocity is fast enough for the electrons produced by the unreacted beam to terminate on the closed gating grid, the drift velocities of the positive ions produced are of order $10^{-4}$ times slower.
+At higher beam rates the positive charge is allowed to pile up producing a region of positive space charge, introducing perturbations to the nominal electric field.
+
+Since the beam comes in along the z-axis, and the drift direction of the ions is in the -y direction, we can estimate the charge density as an uniform sheet charge.
+The surface charge density is related to the beam rate, ion drift length, ion drift velocity, and the energy loss of the beam.
+Though the incoming beam comes randomly, the slow drift velocity combined with the high beam rate makes the uniform approximation valid.
+Tracking or estimating the beam rate as a function of time with in a given run would provide a better estimate of the space charge.
+Or experimentally a laser system could be pulsed after each event (throughout the drift volume), giving the experimental correction map for the drift.
+While the potential for a laser system was implemented in the field cage design, a laser system ultimately was not developed for the \spirit tpc.
+
+The beam rate within a run is roughly constant, therefore we can estimate the space charge and provide a first order correction for the space charge effect.
+
+As mentioned in CHAPTER ???, the cocktail beam momentum was well known to within 1\% as set by the BIGRips spectrometer.
+Also time of flight analysis of upstream and downstream scintilators also independently confirmed the beam rigidity setting.
+The expected momentum is given in TABLE ???
+as calculated by from the beam rigidity setting of the magnet (WHICH MAGNET).
+The measured momenta as determined by the TPC software (given in FIG TABLE ), shows a disagreement on the level of 5\% too high.
+We noticed that if one only uses the first 90 layers (out of 112) of the TPC, the momenta is lower; one should expect the momentum to go higher as the track length is shorter (short tracks effectively are straight lines).
+In the cocktail beam there is no un-reacted beam causing any significant space charge.
+The magnetic field map of the SAMURAI magnet has been calculated by TOSCA simulation CITE ???, and several points have been verified experimentally with a hall probe to be within XXX \%.
+We assume that the electric field (to first order) is uniform in the y-direction.
+Using Garfield++ CITE ???, we can model the transverse drift of the electrons in the presence of such fields.
+
+A grid of electrons uniformly distributed in the TPC model space were drifted to the gating grid position. The final shifts in the x and z positions was measured.
 
@@
 
-\section{Drift Task}
+\subsection{Drift Task}
 The drift task takes the energy loss points calculated from Geant4 and converts them into electrons.
@@
 
-\section{Pad Response Task}
+\subsection{Pad Response Task}
 After the avalanche process of the previous task, the total charge of the event is split over several pads defined by the Pad Response Function (as discussed in section ????).
@@
 
-\section{Electronics Task}
+\subsection{Electronics Task}
 The electronics task takes the total charge on each pad and simulates the electronics response, converting electronics into ADC channels.
@@
 
-\section{Track Embedding}
+\subsection{Simulating Saturation}
+Add figure showing saturated time bucket spectra with location of saturation identified and with pulse shape from embedding I would like to add and how it blocks it.
+Add Figure with 2D pad plane response with and without saturation flag
+
+\section{Monte Carlo Track Embedding}
+Add Figure of MC track embedding
+
 Track embedding is the process of taking a simulated MC track from Geant4 and embedding its response into a real data event.
@@
 
+\subsection{MC and Data Comparison}
+Add Figure of Pad response function for pion,proton.... for MC vs Data vs angles...
+Add Figure of Number of clusters of MC vs Data
+Add Figure of dEdx MC vs Data
+Add Figure of Momentum resolution MC vs Data
+Add Figure of track residuals? MC vs data?
+
 \section{Efficiency Corrections}
+Add Figures of efficiency vs angles in TPC polar angle plot for pions
+
 Since the \spirit TPC is a fixed target experiment it's angular coverage is certainly not 4$\pi$.
@@
 The rectangular design complicates the calculation of the geometric acceptance, or the efficiency.
+
+\section{Beam Particle Identification}
+Figures of beam contaminants in our beam PID line. 
+Table of beam purity reference to Jon's thesis paper
+
+\section{Solid angle coverage}

### 2019-08-07 chapters/ch_experiment.tex
Adding better tables and gas section. Macros folder added

@@
 \section{S$\pi$RIT TPC Overview}
+Add Overview image with labels
+
 The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (S$\pi$RI TPC) is a Multi-Wire Proportional Counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiments.
@@
 A target ladder allowed for up to 5 fixed targets to be mounted.
-A ACME worm gear allowed for the x-axis motion for changing the targets.
+A ACME worm gear allowed for the x-axis motion for changing the targets during the experiment, without needing to open or move the TPC.
 The motion of this worm gear was translated through the target motion feed-through by several brass gears and non-magnetic gear boxes.
@@
 Several aluminum ribs were mounted to the top of the plate to bring structural rigidity.
-Because of this the top plate flatness was within 150$\mu$m across the whole top plate as measured by a laser position system.
+A flatness within 150$\mu$m was achieved across the whole top plate as measured by a laser position system.
 The charge sensitive pads of the pad plane were etched into several circuit boards were recessed and glued to the bottom portion of the top plate.
@@
 The voltage step down takes the high voltage of the field cage cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{\spirit TPC Overview} \\
+ \midrule
+Pad plane area & 1.3 m x .9 m\\
+Pad size       & 1.2 cm x .8 cm \\
+Number of pads & 12096 (112 x 108) \\
+Gas composition& 90\% Ar + 10\% CH${}_4$  \\
+Multiplicity limit & 200  \\
+dE/dx range        & Z=1-8, $\pi$, p,d,t,He,Li-O \\
+Drift length       & 50 cm \\
+\bottomrule
+\end{tabular}
+\caption{Caption}
+\end{table*}
 
@@
 
-\section{Considerations when constructing a TPC}
-Several considerations went into the construction of the S$\pi$RI TPC which I wish to summarize and document here.
-All materials and glue of the TPC were selected as low out-gassing materials.
-Several materials that are common place in nuclear labs, from vacuum grease to o-rings, out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time.
-The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments.
-If a material we wished to used was not on these lists we placed the material in a clean chamber with the counter gas and flowed this counter gas through a small proportional counter making sure the gain did not drop at high collection rates when exposed to a high rate alpha Americium source.
+\subsection{Enclosure}
 
-Sparking
-Two volumes of gas.
+\subsection{Voltage Step Down}
 
-\section{Field Cage}
+\subsection{Field Cage}
 
@@
 The halogen free material chosen was Rodgers ????.
-We build the TPC with the option to run explosive gases such  hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
+We built the TPC with the option to run explosive gases such  hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
 While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.
@@
 
-\section{Voltage Step Down}
+\subsection{Voltage Step Down}
 
@@
 
-\section{Wire Planes}
+\subsection{Wire Planes}
+Add figure of gating grid transparency closed and open configuration
 
-\section{Pad Plane}
+\subsection{Pad Plane}
+Add figure of close up of pad plane
 
-\section{Electronics}
+\subsection{Electronics}
 
@@
 \caption{Summary of the optional settings of the GET electronics and those used during the 20?????? experimental runs}
-\label{tb:spiritoverview}
+\label{tb:getoverview}
 \end{table}
 
+\subsection{Considerations when constructing a TPC}
+Several considerations went into the construction of the S$\pi$RI TPC which I wish to summarize and document here.
+All materials and glues of the TPC were selected as low out-gassing materials.
+Several materials (that are common place in nuclear labs), such as vacuum grease, viton o-rings, all out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time.
+The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments.
+If a material we wished to used was not on these lists we placed the material in a clean chamber with the counter gas and flowed this counter gas through a small proportional counter making sure the gain did not drop at high collection rates when exposed to a high rate alpha Americium source.
+
+Sparking
+Two volumes of gas.
+
+\subsection{Gas Properties}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{driftvel.png}
+\caption{Drift velocity of electrons in P10 gas.}
+\label{fig:driftvel}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{gain.png}
+\caption{Number of electrons produced in a single avalanche on an anode wire.
+Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
+The expected Polya distribution fit is also given in yellow.}
+\label{fig:anodegain}
+\end{figure}
+
+\begin{table}[!htp] % not just 'h!'
+\centering % not a center environment
+\begin{tabular}{
+  @{}
+  l
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=5.2]
+  S[table-format=5.2]
+  @{}
+}
+\toprule
+Gas properties &
+ {$\sigma_{t}$} &
+ {$\sigma_{l}$} &
+ {$v_{d}$} &
+ {$G_{h}$} &
+ {$G_{l}$} \\
+&
+  {($\si{\centi\meter}^{-1/2}$)} &
+  {($\si{\centi\meter}^{-1/2}$)} &
+  {(\si{\centi\meter\per\micro\second})} \\
+\midrule
+\phantom{abc}   &.024   &.034  &5.43   &  903   &150     \\
+\bottomrule
+\end{tabular}
+
+\caption{}
+\label{tb:gasprop}
+\end{table}
+
+Add table for gas diffusion
+
+\section{Ancillary Detectors }
+\subsection{Kyoto Multiplicity Trigger}
+\subsection{Krakow ?????? (KATANA)}
+
+\section{Radio Isotope Beam Factory (RIBF) Facility }
+Cyclotron facility overview.
+Samurai line overview.
+Beam line element overview.
+Big rips beam PID. reference
+
 \section{S$\pi$RIT at RIBF}
-
-\section{Kyoto Multiplicity Trigger}
-\section{Krakow ?????? (KATANA)}
-\section{Ancillary Detectors }
-\section{Radio Isotope Beam Factory (RIBF) Facility }
+Picture of setup.
 
@@
 
-\section{Experimental setup}
+\section{Experimental Setup}
+
 \section{Trigger Condition}
+How it was made using kyoto krakow
+
+\section{Collision Data Taken}

### 2019-08-07 chapters/ch_intro.tex
Adding better tables and gas section. Macros folder added

@@
 \chapter{Introduction}
+
 \section{From Nuclear Forces to the Equation of State}
+Figure showing the nuclear forces for the pp, pn, nn 
+Figure showing the binding energy vs p-n asym
+
 \section{Phases of Nuclear Matter}
+
 \section{The Nuclear Equation of State}
+
+\section{Boltzmann Ulong Uhlenbeck (BUU) Transport Code}
+Figure of transport simulation code showing collision progression
+
+\section{Observables}
+Figure showing p-n observable lessens at high energy and from all densities
+
+\section{The $\pi$ observable}
+Figure showing delta resonance 
+Figure showing pion's produced at 2po 
+Figure showing fermi motion effect
+
 \section{Previous Constraints}
+Figure showing GW constraint and other previous constraints 
+Can you group all the constraints to one nice plot???
+
 \section{Motivation for building the S$\pi$RIT TPC}

### 2019-08-07 chapters/ch_results.tex
Adding better tables and gas section. Macros folder added

@@
 \section{Comparison to Theory}
+Add figure of Pion spectra for different theory 
+Reference paper for multiple theories
+
 \section{$\pi$ Yields}
+Add figure on Glauber model 
+Add figure on comparison to FOPI data
+Add figure of angular coverage of TPC spectra
+
 \section{$\pi^-$/$\pi^+$ Ratio}
+Add figure of Theory for pion ratios
+
+\section{Pion Inferometry}
+
 \section{Constraint on the Symmetry Energy}
+maybe not...

### 2019-08-07 macrosForThesis/driftvelocity/getDriftVel.C

@@ -1,0 +1,141 @@
+#include <iostream>
+#include <fstream>
+
+#include "TCanvas.h"
+#include "TROOT.h"
+#include "TMath.h"
+#include "TFile.h"
+#include "TH3D.h"
+#include "DriftLineRKF.hh"
+#include "ComponentUser.hh"
+
+
+#include "ViewField.hh"
+#include "ComponentAnalyticField.hh"
+#include "MediumMagboltz.hh"
+#include "SolidTube.hh"
+#include "SolidBox.hh"
+#include "GeometrySimple.hh"
+#include "Sensor.hh"
+#include "AvalancheMC.hh"
+#include "AvalancheMicroscopic.hh"
+#include "TrackHeed.hh"
+#include "Random.hh"
+#include "Plotting.hh"
+#include <sys/stat.h>
+#include <sys/types.h>
+
+using namespace std;
+using namespace Garfield;
+
+
+TGraph *make(TGraph *h)
+{
+  int fICvs=0;
+  int fWCvs=6000;
+  int fHCvs=4000;
+
+  double fWUnit=0.006;
+  double fHUnit=0.035;
+
+  double fWStat=0.25;
+  double fHStat=0.18;
+  double fXStat=0.95;
+  double fYStat=0.88;
+
+  double fTMargin=0.01;
+  double fBMargin=0.305;
+  double fLMargin=0.12;
+  double fRMargin=0.12;
+  double fRMarginH1=0.03;
+  double fRMarginH2=0.155;
+
+  double fMainTitleSize=0.08;
+  double fAxisLabelSize=0.05;
+  double fAxisTitleSize=0.05;
+  double fXTitleOffset=.9;
+  double fYTitleOffset=.8;
+  h->SetTitle("");
+  h->GetXaxis()->CenterTitle();
+  h->GetXaxis()->SetTitleOffset(fXTitleOffset);
+  h->GetXaxis()->SetTitleSize(fAxisTitleSize);
+  h->GetXaxis()->SetLabelSize(fAxisLabelSize);
+  h->GetYaxis()->CenterTitle();
+  h->GetYaxis()->SetTitleOffset(fYTitleOffset);
+  h->GetYaxis()->SetTitleSize(fAxisTitleSize);
+  h->GetYaxis()->SetLabelSize(fAxisLabelSize);
+
+  return h;
+}
+
+int main(int argc, char *argv[])
+{
[…]

### 2019-08-07 macrosForThesis/driftvelocity/publishgasplots.C

@@ -1,0 +1,12 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishgasplots()
+{
+
+  TFile *inf = new TFile("gasplots.root","UPDATE");
+  TGraph *e_drift = (TGraph *)inf->Get("e_driftvel");
+  style::save(e_drift,"test.png");
+
+}

### 2019-08-07 thesis_style.h

@@ -1,0 +1,173 @@
+namespace style
+{
+  TCanvas *stdcvs    ();
+
+      TH1 *make (TH1     *h);
+      TH2 *make (TH2     *h);
+      TF1 *make (TF1     *f);
+  TLegend *make (TLegend *legend);
+  TCanvas *make (TCanvas *cvs);
+  
+
+  void make (TGraph  *graph);
+
+  void save  (TH1 *h,TString filename, bool logz);
+  void save2  (TH2 *h,TString filename, bool logz);
+  void save  (TGraph *h,TString filename, bool square, bool logy);
+  /********************************************************/
+
+  void init();
+
+     int fICvs=0;
+     int fWCvs=1500;
+     int fHCvs=1000;
+
+  double fWUnit=0.006;
+  double fHUnit=0.035;
+
+  double fWStat=0.25;
+  double fHStat=0.18;
+  double fXStat=0.95;
+  double fYStat=0.88;
+
+  double fTMargin=0.06;
+  double fBMargin=0.14;
+  double fLMargin=0.12;
+  double fRMargin=0.06;
+  double fRMarginH1=0.03;
+  double fRMarginH2=0.155;
+
+  double fMainTitleSize=0.08;
+  double fAxisLabelSize=0.06;
+  double fAxisTitleSize=0.06;
+  double fXTitleOffset=1.1;
+  double fYTitleOffset=.9;
+
+};
+
+void style::init() {
+  fXStat=1.-fRMargin;
+  fYStat=1.-fTMargin;
+  gStyle->SetStatStyle(0);
+  gStyle->SetStatW(fWStat);
+  gStyle->SetStatH(fHStat);
+  gStyle->SetStatX(fXStat);
+  gStyle->SetStatY(fYStat);
+  gStyle->SetTitleFontSize(fMainTitleSize);
+  gStyle->SetLineWidth(10);
+
+}
+
+TCanvas *style::stdcvs()
+{
+  init();
+  auto cvs=new TCanvas("cvs","cvs",fWCvs,fHCvs);
+  make(cvs);
+  return cvs;
+}
+
+TCanvas *style::make(TCanvas *cvs) {
+  cvs->SetMargin(fLMargin,fRMargin,fBMargin,fTMargin);
+  cvs->SetFrameLineWidth(10);
+  
+  return cvs;
[…]

### 2019-08-14 MSU-thesis-estee.tex
Added Pion multiplicity document
Added more figures

@@
 
+%IS THIS THE RIGHT BIB STYLE TO USE????!?!?!??!?!?!??!
+%!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+\bibliographystyle{elsarticle-num}
+
 % This is standard fontenc/inputenc for pdflatex
@@
 \usepackage[T1]{fontenc}
+\usepackage{float}
+\usepackage{graphicx}
+\graphicspath{ {images/} }
+
 %
@@
 % If you are using XeLaTeX or LuaLaTeX load the Times or Times New Roman font with \setmainfont
-\usepackage{newtxtext,newtxmath} 
+\usepackage{newtxtext,newtxmath,booktabs,siunitx} 
 %
@@
 % You must specify the title of your thesis, your name, the field of study (not department), and the year
-\title{The syntax and semantics of phonology}
-\author{Joe Linguist}
-\fieldofstudy{Linguistics} % This should be in sentence case
-\date{2017}
+\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
+\author{Justin Estee}
+\fieldofstudy{Physics} % This should be in sentence case
+\date{2019}
 
@@
 %
-%\dedication{This thesis is dedicated to someone.}
+\dedication{This thesis is dedicated to someone.}
 %
+
+%User defined commands
+\newcommand{\spirit}{S$\pi$RIT }
+\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+\includeonly{
+%chapters/ch_intro,
+%chapters/ch_theory,
+chapters/ch_experiment,
+chapters/ch_data_analysis,
+%chapters/ch_data_analysis_2,
+%chapters/ch_results
+}
+
 \begin{document}
+%\sisetup{quotient-mode=fraction} % Output a/b as \frac{a}{b}
 
@@
 
-\include{chapters/ch_nucleareos}
-\include{chapters/ch_piontheory}
-\include{chapters/ch_embed}
-\include{chapters/ch_theorymodel}
-\include{chapters/ch_tpc}
-\include{chapters/ch_software}
+\include{chapters/ch_intro}
+\include{chapters/ch_theory}
 \include{chapters/ch_experiment}
-\include{chapters/ch_analysis}
+\include{chapters/ch_data_analysis}
+\include{chapters/ch_data_analysis_2}
 \include{chapters/ch_results}
@@
 % If you are using natbib it is \bibsep; if using biblatex it's \bibitemsep
+\bibliography{mybibfile}
+
 \end{document}

### 2019-08-14 chapters/ch_data_analysis.tex
Added more figures

@@
 \chapter{Data Analysis I: Calibration and Corrections}
+
+Need to explain pedestal subtraction 
+GG noise subtraction
+
 \section{Software}
@@
 \section{Calibrations and Corrections}
+
+\subsection{Cocktail calibration}
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrcrrrcrrr@{}}\toprule
+& \multicolumn{3}{c}{$100 MeV$} & \phantom{abc}& \multicolumn{3}{c}{$100 MeV$} &
+\phantom{abc} & \multicolumn{3}{c}{$300 MeV$}\\
+\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
+& &\multicolumn{2}{c}{Measured}\\
+Particle &\phantom{abc} & Raw & E$\times$B\\
+\midrule
+p   & 882.8 & d & d & 903.5 & e & e &\phantom{abcdef} & f & f \\
+d   & 817.1 & d & d & 898.5 & e & e & 1621.1 & f & f\\
+t   & 589.5 & d & d & 887 & e & e & 1612.4 & f & f  \\
+$^{3}$He  & 1617.3  & d & d & 1795.2 & e & e & 3236.4 & f & f\\
+$^{4}$He  & 1405.6  & d & d & 1782.9 & e & e & 3226.4 & f & f \\
+\bottomrule
+\end{tabular}
+\caption{Summary of expected cocktail. }
+\label{tb:cocktailsummary}
+\end{table*}
+
 Cocktail Calibration 
+Picture of cocktail before and after ExB effect
+Table of LISE++ expected cocktail energies ridigity setting of dipole magnets (reference big rips line)
 
@@
 Pulse the ground grid. 
-Picture of calibration of channels before and after.
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{gaincalib.png}
+\caption{Calibration of electronics}
+\label{fig:gaincalib}
+\end{figure}
 
 \subsection{Anode gain calibration}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{highlowcal.png}
+\caption{Calibration of low high.}
+\label{fig:highlowcal}
+\end{figure}
+
 Picture of low vs high gain channels and fit
@@
 Discuss the relevant time scales, drift lengths, magnitudes, and locations of space charge
-Figure of cartoon sheet charge
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{spacechg_cartoon.png}
+\caption{Location of space charge in 132 Sn}
+\label{fig:spacechg_cartoon}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{beampath.png}
+\caption{Beam path of the experiments}
+\label{fig:beampaths}
+\end{figure}
+
 Figure of beam path locations for different beam types

### 2019-08-14 chapters/ch_experiment.tex
Added more figures

@@
 \end{tabular}
-\caption{Caption}
+\caption{An overview of the properties of the \spirit TPC}
+\label{tb:spiritoverview}
 \end{table*}
-
-\begin{table}
- \begin{tabular}{||c c c c||} 
- \hline
- \multicolumn{4}{|c|}{S$\pi$RIT TPC Overview} \\
- \hline
- Pad Plane Area & 1.3 m x 0.9 m & Gas Gain & 1000\\
- \hline
- Pad Size & 1.2 cm x 0.8 cm & Drift Velocity & 5.5 cm/$\mu$s \\
- \hline
- Number of pads & 12096 (112x108) & E-field & 135 V/cm  \\
- \hline
-  Gas composition & 90\% Ar + 10\% CH${}_4$ & Multiplicity limit & 200  \\
- \hline
- Drift length & 50 cm & dE/dx range & Z=1-8, $\pi$, p,d,t,He,Li-O \\ [1ex] 
- \hline
-\end{tabular}
-\caption{An overview of the properties of the S$\pi$RIT TPC.}
-\label{tb:spiritoverview}
-\end{table}
 
@@
 
-\begin{table}
- \begin{tabular}{||c c c c||} 
- \hline
- \multicolumn{4}{|c|}{GET Electronics Settings} \\
- \hline
- GET settings & All options & ${}^{132}$Sn + ${}^{124}$Sn & ${}^{124}$Sn + ${}^{112}$Sn \\
- \hline
- ADC bit range      & 14 bits &  &  \\
- \hline
- Sampling frequency & 1 - 100 MHz &   &    \\
- \hline
- Dynamic range      & .12,.24,1.0,10 pC & & \\
- \hline
- Peaking time       &  69,117,232,501,720, 1014 ns &  &   \\
- \hline
- Time bucket        & 512 & 270 & 270  \\
- \hline
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{GET electronics settings}\\
+\midrule
+ADC bit range       & 14 bits \\
+Sampling frequency  & 1-100 MHz \\
+Dynamic range       & .12, .24, 1.0, 10pC \\
+Peaking time        & 69,117,232,501,720,1014 ns \\
+Time bucket range   & 512\\
+\bottomrule
 \end{tabular}
-\caption{Summary of the optional settings of the GET electronics and those used during the 20?????? experimental runs}
+\caption{Summary of range of GET electronics settings. }
 \label{tb:getoverview}
-\end{table}
+\end{table*}
 

### 2019-08-14 macrosForThesis/GatingGrid/gatinggrid.py

@@ -1,0 +1,163 @@
+"""
+Created on Fri Mar  6 10:33:15 2015
+
+Gating grid simulation
+
+@author: tangwanc
+"""
+##################
+from numpy import matrix
+from numpy import linalg
+from numpy import log
+from numpy import arange
+from numpy import arccos
+from numpy import sin
+from matplotlib import pyplot as plt
+from math import pi
+import numpy as np
+
+# Electric field in [V/cm]
+Ef = 133.3
+# Electric field in [V/mm]
+Efmm = Ef/10.0
+
+# Electric Permativity [Farad/mm]
+eps0 = 8.854187817e-15
+
+# Heights of wire are referrenced from the pad plane (z = 0)
+
+# Anode distance
+z1 = 3.96
+
+# Ground distance
+z2 = 8.011
+
+# Gating grid distance
+z3 = 14.061
+
+# Cathode distance
+zp = 510.24
+
+###### Pitch of wires ######
+
+# Anode pitch
+s1 = 4.0
+
+# Ground pitch
+s2 = 1.0
+
+# Gating grid pitch
+s3 = 1.0
+
+###### Radii of wires ######
+
+# Anode radius
+rs = 0.01
+
+# Ground radius
+rz = 0.0381
+
+# Gating grid radius
+rg = 0.0381
+
+###### Voltages [V]######
+
+# Anode voltage
+vs = 680
+
+# Ground voltage
+vz = 0
+
+# Gating grid voltage
+vg = -114.8
+
+# Cathode voltage
+vp = -6632.50
+
+###### Matrices #######
+A = (1/eps0)*matrix([[z1-(s1/(2*pi))*log(2*pi*rs/s1),z1,z1,z1],
+                    [z1, z2-(s2/(2*pi))*log(2*pi*rz/s2), z2, z2],
+                    [z1, z2, z3-(s3/(2*pi))*log(2*pi*rg/s3), z3],
+                    [z1, z2, z3, zp]])
+
+# Inverse of A
+invA = linalg.inv(A)
+
+# Voltage matrix
+Vol = matrix([[vs], [vz], [vg], [vp]])
+
+####### Solving for surface charge densities ######
+sigma = invA*Vol
+
[…]

### 2019-08-14 macrosForThesis/SpaceCharge/PlotBeamPaths/getBeamPath.C

@@ -1,0 +1,84 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+void openBeamFile(string filename, vector<double> &xx, vector<double> &yy, vector<double> &zz)
+{
+ string line;
+  ifstream file(filename);
+  while(getline(file,line))
+    {
+      stringstream sstring(line);
+      double x,y,z;
+      sstring >> x >> y >> z;
+      xx.push_back(x/10.);
+      yy.push_back(y/10.);
+      zz.push_back(z/10.); //convert to cm
+    }
+}
+
+void getBeamPath()
+{
+  vector<double> x_132, y_132, z_132;
+  vector<double> x_108, y_108, z_108;
+  vector<double> x_124, y_124, z_124;
+  vector<double> x_112, y_112, z_112;
+      
+  openBeamFile("_132Sn_BeamTrack.data",x_132,y_132,z_132);
+  TGraph *beam_132 = new TGraph(x_132.size(), z_132.data(),x_132.data());
+  beam_132 -> GetYaxis()->SetRangeUser(-43.2,43.2);
+  beam_132 -> GetXaxis()->SetRangeUser(0,132);
+
+  openBeamFile("_108Sn_BeamTrack.data",x_108,y_108,z_108);
+  TGraph *beam_108 = new TGraph(x_108.size(), z_108.data(),x_108.data());
+  beam_108 -> GetYaxis()->SetRangeUser(-43.2,43.2);
+  beam_108 -> GetXaxis()->SetRangeUser(0,108);
+
+  openBeamFile("_124Sn_BeamTrack.data",x_124,y_124,z_124);
+  TGraph *beam_124 = new TGraph(x_124.size(), z_124.data(),x_124.data());
+  beam_124 -> GetYaxis()->SetRangeUser(-43.2,43.2);
+  beam_124 -> GetXaxis()->SetRangeUser(0,124);
+
+  openBeamFile("_112Sn_BeamTrack.data",x_112,y_112,z_112);
+  TGraph *beam_112 = new TGraph(x_112.size(), z_112.data(),x_112.data());
+  beam_132 -> GetYaxis()->SetRangeUser(-14,1);
+  beam_132 -> GetXaxis()->SetRangeUser(0,112);
+  beam_132 -> GetYaxis()->SetTitle("x (cm)");
+  beam_132 -> GetXaxis()->SetTitle("z (cm)");
+
[…]

### 2019-08-14 macrosForThesis/SpaceCharge/PlotBeamPaths/getBeamPath_cir.C

@@ -1,0 +1,70 @@
+
+
+void getBeamPath()
+{
+  string line;
+  ifstream file("_132Sn_BeamTrack.data");
+  vector<double>xx,yy,zz;
+  while(getline(file,line))
+    {
+      stringstream sstring(line);
+      double x,y,z;
+      sstring >> x >> y >> z;
+      xx.push_back(x/10.);
+      yy.push_back(y/10.);
+      zz.push_back(z/10.); //convert to cm
+      //      cout<<x<<" "<<y<<" "<<z<<endl;
+    }
+
+  TGraph *beam = new TGraph(xx.size(), zz.data(),xx.data());
+  beam->GetYaxis()->SetRangeUser(-43.2,43.2);
+  beam->Draw("ALO");
+
+  TF1 *beampath = new TF1("beampath","[0] + [1]*x + [2]*pow(x,2)",0,130); 
+  //  TF1 *beampath = new TF1("beampath","[0] + [1]*x + [2]*pow(x,2)",0,130); 
+  beampath->SetParameters(-0.0685578,-0.046186, -0.000395632);
+  //  beampath->SetParameters(-0.685578,-0.046186,-3.95632e-05);
+  //  beam->Fit(beampath);
+  beampath->Draw("same");
+
+   auto chi2Function = [&](const Double_t *par) {
+      //minimisation function computing the sum of squares of residuals
+      // looping at the beampathaph points
+      Int_t np = beam->GetN();
+      Double_t f = 0;
+      Double_t *x = beam->GetX();
+      Double_t *y = beam->GetY();
+      for (Int_t i=0;i<np;i++) {
+         Double_t u = x[i] - par[0];
+         Double_t v = y[i] - par[1];
+         Double_t dr = par[2] - std::sqrt(u*u+v*v);
+         f += dr*dr;
+      }
+      return f;
+   };
+   // wrap chi2 funciton in a function object for the fit
+   // 3 is the number of fit parameters (size of array par)
+   ROOT::Math::Functor fcn(chi2Function,3);
+   ROOT::Fit::Fitter  fitter;
+   double pStart[3] = {0,0,1};
+   fitter.SetFCN(fcn, pStart);
+   fitter.Config().ParSettings(0).SetName("x0");
+   fitter.Config().ParSettings(1).SetName("y0");
+   fitter.Config().ParSettings(2).SetName("R");
+   // do the fit 
+   bool ok = fitter.FitFCN();
+   if (!ok) {
[…]

### 2019-08-20 chapters/ch_data_analysis.tex
Added some preliminary delta res

@@
 \begin{tabular}{@{}rrrrcrrrcrrr@{}}\toprule
-& \multicolumn{3}{c}{$100 MeV$} & \phantom{abc}& \multicolumn{3}{c}{$100 MeV$} &
-\phantom{abc} & \multicolumn{3}{c}{$300 MeV$}\\
+& \multicolumn{3}{c}{$100 MeV$} & \multicolumn{3}{c}{$100 MeV$} & \multicolumn{3}{c}{$300 MeV$}\\
 \cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
-& &\multicolumn{2}{c}{Measured}\\
+& &\multicolumn{2}{c}{Measured} & & \multicolumn{2}{c}{Measured} & & \multicolumn{2}{c}{Measured}\\
+\cmidrule{3-4} \cmidrule{7-8} \cmidrule{11-12}
 Particle &\phantom{abc} & Raw & E$\times$B\\
@@
 
-Figure of beam path locations for different beam types
-Table of energy loss tables for energies and beam types
+\begin{table}[!htp] % not just 'h!'
+\centering % not a center environment
+\begin{tabular}{
+  @{}
+  l
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=5.2]
+  S[table-format=5.2]
+  @{}
+}
+\toprule
+Beam Energy Loss  &
+ {${}^{132}$Sn} &
+ {${}^{124}$Sn} &
+ {${}^{112}$Sn} &
+ {${}^{108}$Sn} &
+  {Avg.}\\
+
+\midrule
+$\si{\kilo\eV\per\centi\meter}$ & 11.2   &.034  &5.43   &  903   &150     \\
+\bottomrule
+\end{tabular}
+
+\caption{Average energy loss of each beam.}
+\label{tb:beameloss}
+\end{table}
+
 Reference appendix for poisson solver 

### 2019-10-15 macrosForThesis/ElectronBackground/electron_bkg_article.tex
Electron Background and some MC matching

\documentclass[12pt, letterpaper, twoside]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{/home/justin/mythesis/Graphics/}}

\title{Estimating the Electron Background}
\author{Justin Estee}
\date{August 2019}

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

 \newcommand\elec{\mathrm{e^-}}
\newcommand\postr{\mathrm{e^+}}
\newcommand\piz{\mathrm{\pi_0}}

\section{Electron background}
The $\piz$ decays through two main branches, 
\begin{equation}
\piz \rightarrow \gamma\gamma
\label{eq:branch1}
\end{equation}

 \begin{equation}
\piz \rightarrow \elec\postr\gamma
\label{eq:branch2}
\end{equation}.

Where the branching ratios are 98.8\% for Eq.~\ref{eq:branch1} and 1.2\% for Eq.~\ref{eq:branch2}, with all other branching ratios being several order of magnitude less. 
The main source of the $\elec \postr$ background in our PID spectra are coming from the so called Dalitz decay, Eq.~\ref{eq:branch2}.

\section{Dalitz Decay Plots}

\begin{figure}[!htb]
\centering
\includegraphics[]{pizerodecay.pdf}
\label{fig:pizdecay}
\caption{Figure for convention of 3 Body decay.}
\end{figure}

The physics of the 3-body decay can be completely described by the 4-vector relation,

\begin{equation}
p^\mu = p^\mu_1 + p^\mu_2 + p^\mu_3,
\label{eq:dalitz_fourvec}
\end{equation}
where in our decay,

\begin{equation}
p^\mu = \{M_\Delta,\vec{0}\}
\end{equation}

\begin{equation}
p^\mu_1 = \{E_\gamma,\vec{p}_\gamma\}
\end{equation}

\begin{equation}
p^\mu_2 = \{E_{e^-},\vec{p}_{e^-}\}
\end{equation}

\begin{equation}
p^\mu_3 = \{E_{e^+},\vec{p}_{e^-}\},
\end{equation}
which are given in the rest frame of the decay particle.

2-body decays are completely determined by energy and momentum conservation whereas 3-body decays have additional degrees of freedom. The phase space of a 3-body decay are best described by what is referred to as a Dalitz plot. There are 12 d.o.f. from the momentum vectors of 3 particles. Since the 3-body decay must occur within a plane we can remove 3 d.o.f. by conveniently selecting $p_{i,z} = 0$, 3 d.o.f. from energy momentum relations, 3 d.o.f. from momentum conservation, and 1 d.o.f. by noticing rotations in the x-y plane do not change the result. There are only two degrees of freedom that completely determine the system. In a Dalitz plot these variables are the invariant masses of two particles,

\begin{equation}
m^2_{12} = (p^\mu_1 + p^\mu_2)^2 = (p^\mu - p^\mu_3)^2 = M^2 + m_3^2 - 2ME_3
\label{eq:dalitz_m12}
\end{equation}

\begin{equation}
m^2_{23} = (p^\mu_2 + p^\mu_3)^2 = (p^\mu - p^\mu_1)^2 = M^2 + m_1^2 - 2ME_1
\label{eq:dalitz_m23}
\end{equation}

\begin{equation}
m^2_{13} = (p^\mu_1 + p^\mu_3)^2 = (p^\mu - p^\mu_2)^2 = M^2 + m_2^2 - 2ME_2.
\label{eq:dalitz_m13}
\end{equation}

Only two of these invariant masses are required to fully simulate the entire 3-body decay, where as the remaining invariant mass is determined by energy conservation,

\begin{equation}
m^2_{12} + m^2_{23} + m^2_{13} = M^2 + \sum_i m_i^2.
\label{eq:dalitz_conserve}
\end{equation}

In our simulation the indicies are defined to each particle as such: 1$\leftrightarrow~e_-$, 2$ \leftrightarrow e^+$, and 3$ \leftrightarrow \gamma$. The invariant masses $m_{13}$ and $m_{23}$ are randomly generated and $m_{12}$ is given by Eq.~\ref{eq:dalitz_conserve}. The space in which $m_{13}$ and $m_{23}$ covers is generated uniformly. If $m_{12} \leq 0$ then the event is skipped as it did not meet the conservation criteria. The energies of each particle, $E_i$, can be solved for in Eq's~\ref{eq:dalitz_m12} \ref{eq:dalitz_m13} \ref{eq:dalitz_m23}. Also, the magnitude of the momentum $p_i$ is given by,

\begin{equation}
p_i^2 = E_i^2 - m_i^2.
\end{equation}

The angle of emission in the plane of the decay is fixed by these quantities. Modifying Eq.~\ref{eq:dalitz_fourvec}, we can solve for the angles between two particles $\gamma_{13}$ and $\gamma_{23}$, as defined in in Fig.~\ref{fig:pizdecay};

\begin{equation}
\cos(\gamma_{23}) =  \frac{2E_\gamma E_{e^+} + 2M_{\pi^0}E_{e^-} -M_{\piz}^2}{2p_{\gamma}p_{e^+}}
\end{equation}

\begin{equation}
\cos(\gamma_{13}) =  \frac{2E_\gamma E_{e^-} + 2M_{\piz}E_{e^+} -M_{\piz}^2}{2p_{\gamma}p_{e^-}}.
\end{equation}

The entire 4-vector of each particle, in the rest frame of the $\piz$, can be described in the plane of the decay. The axis along the decaying $\gamma$ must align with the momentum vector of the $\piz$ in the center of mass frame, by conservation of momentum. The direction of decay of the $\gamma$ has no preference and may decay forwards or backwards with respects to the incoming $\piz$ momentum, as illustrated in Fig.~\ref{fig:pizdecay}. Due to the symmetry along this same axis, the plane angle may be randomly oriented around the $\piz$ axis. In the simulation the angle of the plane is randomly rotated and half of the time the $\gamma$ decays forward and vice versa.

The rest frame of the $\piz$ is defined by its momentum in the COM frame of the heavy ion collision where the Lorentz factor $\beta_{\piz}$ is given by the momentum and total energy of the pion,
\begin{equation}
\beta_{\piz} = \frac{ p_{\piz} }{E_{\piz}}.
\end{equation}

We assume the pion emission in the COM system is isotropic. Two Lorentz boosts are relevant in this simulation. First, the Lorentz boost from the rest frame of the pion to the COM of the h.i.c. Secondly from the COM system to the Lab frame which is defined by the Lorentz boost factor $\beta_{\mathrm{COM}}$,

The resolution and acceptance of the TPC may smear or shift the $\elec$ and $\postr$ momentum distribution. We use the theoretically calculated momentum distribution in the lab frame as an input into the MC simulation of the TPC. The simulated spectra from the TPC is given in Fig.~ . The peak position has shifted significantly due systematic shifts in the momentum resolution.

\end{document}

### 2019-10-15 macrosForThesis/ElectronBackground/JonsBkg/distribution.C

@@ -1,0 +1,52 @@
+void distribution(){
+    TH1D *histDist=new TH1D("histDist","",1000,0,1);
+    Double_t me=0.511;
+    Double_t mpi0=135.0;
+    Double_t scale=38.3;
+    Double_t widthintegral=0;
+    ofstream csvout;
+    csvout.open("eedist.txt");
+    Double_t temp=0;
+    Int_t counter=0;
+    for(int i=0;i<1400;i++){
+        Double_t mee=double(i)/10.;
+        Double_t width=1./mee*TMath::Sqrt(1.-4.*me*me/mee/mee)*(1+2*me*me/mee/mee)*TMath::Power((1-mee*mee/mpi0/mpi0),3)*TMath::Power(TMath::Abs(1+0.0305*mee*mee/mpi0/mpi0),2)/scale;
+	if(width>0) widthintegral=widthintegral+width;
+	if(widthintegral>0 && widthintegral > temp+0.0001 ){
+	  csvout << widthintegral << "\t" << mee << endl;
+	  counter++;
+	}
+    if(widthintegral > temp+0.0001)	temp=widthintegral;
+    }
+    cout << "fill done" << endl;
+    cout << "lines: " << counter << endl;
+    cout << widthintegral << endl;
+    //histDist->Draw();
+}
+/*
+void intPion(){
+  TFile *f=new TFile("histP5.root");
+  TH1D * histP = new TH1D("histP","",1000,0,500);
+  histP=(TH1D*)f->Get("histP5");
+  Double_t scale = 1/((double)histP->Integral());
+  histP->Scale(scale);
+  ofstream csvout;
+  csvout.open("pionInt.txt");
+  int counter=0;
+  Double_t mom, n,n1;
+  for(int i=0;i<1000;i++){
+    mom=((double)i)*0.5;
+
+    n+=histP->GetBinContent(i);
+    if(n>1E-100 && n>n1){
+      csvout << n << "\t" << mom << endl;
+      counter++;
+    }
+    n1=n;
+  }
+  //csvout.Close();
+  TCanvas *c1=new TCanvas("c1","",800,800);
+  histP->Draw();
+  cout << counter << endl;
+}
+*/

### 2019-10-15 macrosForThesis/ElectronBackground/gausSumData.C

@@ -1,0 +1,57 @@
+TH1D *hist = new TH1D("hist","hist",60,-3,3);
+TF1 *g = new TF1("g","[0]*TMath::Gaus(x,[1],[2])",-3,3);
+
+double gausSummer(double norm, double mean, double sigma)
+{
+  g->SetParameters(norm,mean,sigma);
+
+  double total_g = 0;
+
+  for(int i =1 ; i <= hist->GetNbinsX(); i++)
+    {
+      double x = hist->GetXaxis()->GetBinCenter(i);
+      total_g += g->Eval(x);
+    }
+
+  return total_g;
+}
+
+void gausSumData()
+{
+
+
+  /*
+//pi- data
+  vector<double> sigma = {0.123871,0.153382,0.127992,0.0974235,0.119985,0.117022,0.120162,0.121615,0.117738,0.113552,0.1114,0.116035,0.11313,0.116668,0.115118,0.114094,0.113988,0.11522,0.115849,0.115047,0.114424,0.113906,0.113974,0.106661,0.108766,0.109621,0.11205};
+
+  vector<double> mean = {-0.0181331,0.00874019,-0.00365319,0.0024157,0.0038113,0.00645765,0.00175903,0.0153028,0.00866258,0.00305236,0.000726118,0.0019443,0.00284441,-0.00103479,-0.00248062,-0.00346796,-0.00421854,-0.00476472,-0.00526699,-0.00519617,-0.00567455,-0.00554713,-0.0054933,0.00444418,0.00174947,0.000538114,-0.0023042};
+
+  vector<double> norm = {11.5918,16.3924,29.1153,48.7323,59.0943,87.3287,122.671,206.021,294.693,467.275,685.861,945.807,1423.65,1984.24,2703.72,3752.03,4857.35,6117.76,7481.96,8988.27,10226.2,11623.8,12156.1,12954.1,3018.45,2870.49,2621.79};
+  */
+  
+  /*
+  //e+ Genie fit
[…]

### 2019-10-15 macrosForThesis/ElectronBackground/gausSummer.C

@@ -1,0 +1,20 @@
+
+void gausSummer(double norm, double mean, double sigma)
+{
+  TH1D *hist = new TH1D("hist","hist",60,-3,3);
+  TF1 *g = new TF1("g","[0]*TMath::Gaus(x,[1],[2])",-3,3);
+  g->SetParameters(norm,mean,sigma);
+  
+  double total_g = 0;
+
+  for(int i =1 ; i <= hist->GetNbinsX(); i++)
+    {
+      double x = hist->GetXaxis()->GetBinCenter(i);
+      total_g += g->Eval(x);
+    }
+
+  cout<<"Total is "<<total_g<<endl;
+
+}
+
+

### 2019-10-15 macrosForThesis/ElectronBackground/plotpipBkg.C

@@ -1,0 +1,58 @@
+
+void plotpipBkg()
+{
+  //  TFile *f = TFile::Open("data.root");
+  //  TH1D *data = (TH1D *)f->Get("data");
+  
+  TFile *m = TFile::Open("MChistPID_ep_newBin.root");
+  TH2D *pid = (TH2D *)m->Get("histPIDLog");
+  auto mc = pid->ProjectionX();
+
+  vector<double> bins = {40,60,62,64,66,68,70,72,74,76,78,80,82,84,86,88,90,92,94,96,98,100,102,104,106,108,110,112,116,120,124,128};
+  
+  vector<double> counts = {57.2325,56.671,50.1775,52.5155,50.471,56.7945,51.942,48.4285,56.2485,59.8715,39.3344,45.64655,53.3265,54.042,54.35,40.4222,45.58955,44.79085,44.9633,48.83245,48.9515,50.7965,44.70755,43.54125,42.15, 45.21,31.98,42.72,35.59,32.48,34.74,22};
+
+  //  vector<double> counts = {57.2325,56.671,50.1775,52.5155,50.471,56.7945,51.942,48.4285,56.2485,59.8715,39.3344,45.64655,53.3265,54.042,54.35,40.4222,45.58955,44.79085,44.9633,48.83245,48.9515,50.7965,44.70755,43.54125,49.47665,60.0325,69.695,50.62225,36.4035,39.39875,27.35875,40.11925};
+
+  //  vector<double> counts = {165.508,201.849,212.804,269.342,213.14,248.459,213.508,210.374,220.89,216.526,202.761,205.096,221.826,177.13,180.191,198.889,190.988,185.62};
+
+  //  vector<double> bins = {40,60,62,64,66,68,70,72,74,76,78,80,82,84,86,88,90,92,94,96,98,100,102,104,106,108,110,112,116,120,124,128};
+
+  //  vector<double> counts =  {57.2325,56.671,50.1775,52.5155,50.471,56.7945,51.942,48.4285,56.2485,59.8715,39.3344,45.64655,53.3265,54.042,54.35,40.4222,45.58955,44.79085,44.9633,48.83245,48.9515,50.7965,44.70755,43.54125,49.47665,60.0325,69.695,50.62225,36.4035,39.39875,27.35875,40.11925};
+
[…]

### 2019-10-15 macrosForThesis/MatchMCtoData/MatchCluster/publishClusterNum.C

@@ -1,0 +1,54 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishClusterNum()
+{
+  TFile *f = TFile::Open("mcpim_DBhist.root");
+  TH1D *mc = (TH1D *)f->Get("ndf_dist");
+  TH1D *data = (TH1D *)f->Get("ndf_data");
+  mc->GetYaxis()->SetTitle("Probability");
+  mc->GetXaxis()->SetTitle("# of Clusters");
+  data->GetYaxis()->SetTitle("Probability");
+  data->GetXaxis()->SetTitle("# of Clusters");
+
+  data->GetYaxis()->SetMaxDigits(2);
+  mc->GetYaxis()->SetMaxDigits(2);
+  
+  TCanvas *cvs = style::stdcvs();
+  make(data);
+  make(mc);
+  
+  data->SetLineColor(2);
+  mc->SetLineColor(4);
+  //  data = (TH1D *)data->Rebin(20,"data",xbins_h);
+  //  mc = (TH1D *)mc->Rebin(20,"mc",xbins_l);
+  //  data->Scale(1./data->Integral(),"width");
+  //  mc->Scale(1./mc->Integral(),"width");
+
+  //  data->Fit(polya_h);
+  //mc->Fit(polya_l);
+
+  data->GetXaxis()->SetRangeUser(0,130);
+  //  data->GetYaxis()->SetRangeUser(.01,6e-3);
+  data->Draw("hist");
+  mc->Draw("same hist");
+
+  data->SetMarkerStyle(20);
+  mc->SetMarkerStyle(20);
+  data->SetMarkerSize(5);
+  mc->SetMarkerSize(5);
+  data->SetMarkerColor(2);
+  mc->SetMarkerColor(4);
+  
+  auto leg = new TLegend(.3,.2,.7,.5);
+  leg->AddEntry(data,"Data","le");
+  leg->AddEntry(mc,"MC","le");
+  leg->SetBorderSize(0);
+  leg->Draw("same");
+  cvs->SetLogy();
+  cvs->SaveAs("numcluster.png");
+  cvs->SaveAs("numcluster.jpg");
+
+}
+

### 2019-10-15 macrosForThesis/MatchMCtoData/MatchdEdx/publishdEdx.C

@@ -1,0 +1,85 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishdEdx()
+{
+  vector<double> xvec_d = {450,430,410,390,370,350,330,310,290,270,250,230,210,190,170,150,130,110,90,70,50,30};
+  vector<double> mean_d_vec = {21.0945,21.1656,21.0155,21.084,21.2802,21.5517,21.6691,21.7089,21.9353,22.5101,23.0205,23.7143,24.6391,25.9423,27.7671,30.4992,34.6803,41.7807,54.6574,77.6,123,220};
+  vector<double> sigma_d_vec = {2.57073,2.5559,2.43828,2.14023,2.45448,2.63387,2.57544,2.49798,2.45187,2.77636,2.88779,2.99449,3.19529,3.41103,3.73242,4.22408,5.10251,6.73288,8.94627,15.6,18.4,14.99};
+
+  vector<double> xvec_mc = {390,370,350,330,310,290,270,250,230,210,190,170,150,130,110,90,70,50,30};
+    
+  vector<double> mean_mc_vec = {20.8646,20.8,20.8584,20.9096,21.0242,21.1889,21.4688,21.8518,22.4023,23.2309,24.4311,26.1693,28.771,32.8933,39.7323,51.9551,76.7896,124.426,262};
+    
+  vector<double> sigma_mc_vec = {2.56569,2.49788,2.54408,2.54976,2.53057,2.54595,2.5621,2.57777,2.5949,2.62692,2.74507,2.88023,3.16469,3.66627,4.69026,7.0187,15.69,18.764,27.7};
+  
+  TGraph *sigma_d = new TGraph(xvec_d.size(),xvec_d.data(),sigma_d_vec.data());
+  TGraph *mean_d = new TGraph(xvec_d.size(),xvec_d.data(),mean_d_vec.data());
+ 
+  TGraph *sigma_mc = new TGraph(xvec_mc.size(),xvec_mc.data(),sigma_mc_vec.data());
+  TGraph *mean_mc = new TGraph(xvec_mc.size(),xvec_mc.data(),mean_mc_vec.data());
+
+  sigma_d->GetYaxis()->SetTitle("dE/dx arb. units");
+  sigma_d->GetXaxis()->SetTitle("Momentum MeV/c");  
[…]

### 2019-10-15 macrosForThesis/MatchMCtoData/MatchdEdx/publishdEdxMean.C

@@ -1,0 +1,53 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishdEdx()
+{
+  TFile *f = TFile::Open("mc_dedx.root");
+  TGraph *mean_mc = (TGraph *)f->Get("mean_g");
+  TGraph *sigma_mc = (TGraph *)f->Get("sigma_g");  
+
+  TFile *g = TFile::Open("data_dedx.root");
+  TGraph *mean_d = (TGraph *)g->Get("mean_g");
+  TGraph *sigma_d = (TGraph *)g->Get("sigma_g");  
+
+  mean_d->GetYaxis()->SetTitle("dE/dx arb. units");
+  mean_d->GetXaxis()->SetTitle("Momentum MeV/c");  
+  
+  TCanvas *cvs = style::stdcvs();
+  make(mean_d);
+  make(mean_mc);
+  make(sigma_d);
+  make(sigma_mc);
+  
+  mean_d->SetLineColor(2);
+  mean_mc->SetLineColor(4);
+  sigma_d->SetLineColor(2);
+  sigma_mc->SetLineColor(4);
+
+  mean_d->GetXaxis()->SetRangeUser(40,400);
+  mean_d->GetYaxis()->SetRangeUser(10,300);
+
+  mean_d->SetMarkerStyle(20);
+  mean_mc->SetMarkerStyle(20);
+  mean_d->SetMarkerSize(5);
+  mean_mc->SetMarkerSize(5);
+  mean_d->SetMarkerColor(2);
+  mean_mc->SetMarkerColor(4);
+  
+  mean_d->Draw("APO");
+  mean_mc->Draw("same PO");
+
+  auto leg = new TLegend(.5,.6,.7,.5);
+  leg->AddEntry(mean_d,"Data","le");
+  leg->AddEntry(mean_mc,"MC","le");
+  leg->SetBorderSize(0);
+  leg->Draw("same");
+  cvs->SetLogx();
+  cvs->SetLogy();
+  cvs->SaveAs("dedxcompare.png");
+  cvs->SaveAs("dedxcompare.jpg");
+
+}
+

### 2019-10-15 macrosForThesis/MatchMCtoData/isotropic.C

@@ -1,0 +1,32 @@
+
+
+void isotropic()
+{
+  TRandom3 *ran = new TRandom3(12345);
+  TH1D *angles = new TH1D("angles","angles",90,0,90);
+
+  int nMC = 1e6;
+
+  TF1 *theta = new TF1("theta","[0] + [1]*x + [2]*pow(x,2) + [3]*pow(x,3) + [4]*pow(x,4)",0,90);
+  theta->SetParameters(-29.4261,427.701,-1.66007,-0.0734422,0.000540305);
+  		       
+  TF1 *mom = new TF1("mom","[0] + [1]*x + [2]*pow(x,2) + [3]*pow(x,3) + [4]*pow(x,4) + [5]*pow(x,5) + [6]*pow(x,6) + [7]*pow(x,7) + [8]*pow(x,8) + [9]*pow(x,9)",34,550);
+  mom->SetParameters(1325.43,-106.233,2.85887,-0.0313853,0.000185054,-6.54095e-07,1.43438e-09,-1.91671e-12,1.43189e-15,-4.58979e-19);
+
+  for(int i = 0; i < nMC; i++)
+    {
+      double phi = ran->Uniform(0,2*TMath::Pi());
+      double theta = acos(2*ran->Uniform() -1);
+      double p = mom->GetRandom();
+      cout<<"input "<<theta*TMath::RadToDeg()<<endl;
+      TLorentzVector vec;
+      vec.SetXYZM(p*cos(phi)*sin(theta),p*sin(phi)*sin(theta),p*cos(theta),140.);
+      vec.Boost(0,0,.36);
+      angles->Fill(vec.Theta()*TMath::RadToDeg());
+      cout<<vec.Theta()*TMath::RadToDeg()<<endl;
+    }
+
+  angles->Draw();
+  
+
+}

### 2019-10-15 macrosForThesis/ThermalSpectralPionRatio/plotThermalRatio_paper.C

@@ -1,0 +1,91 @@
+
+TSpline3 *bp_spl;
+TSpline3 *kp_spl;
+TSpline3 *bm_spl;
+TSpline3 *km_spl;
+
+double ratio_fcn(double *x, double *par)
+{
+  double ke = x[0];
+  double a = 138;
+  double z = 57;
+  double r_o = 1.25; //fm
+  double T = 29; //temperature 
+  
+  double sigma_pip = kp_spl->Eval(ke)*pow(2*a,bp_spl->Eval(ke));
+  double sigma_pim = km_spl->Eval(ke)*pow(2*a,bm_spl->Eval(ke));
+
+  if(ke < 37 || ke > 280)
+    sigma_pim = 0;
+
+  double up_un = 1.44*z * (pow(2,2/3.) -1 )/r_o/pow(a,1/3.);
+
+
+  //  return TMath::Exp(-2*up_un/T);
+  return (sigma_pim/sigma_pip) * TMath::Exp(-2*up_un/T);
+}
+
+
+
+
+void plotThermalRatio()
+{
+
+  vector<double> b_plus = {.799  , .697 , .621  , .629  , .646  , .671  , .674  , .695 , .692   , .694};
+  vector<double> k_plus = {12.11 , 24.5 , 36.35 , 37.76 , 35.06 , 32.26 , 32.09 , 28.1 , 24.24 , 22.4};
+  vector<double> t_ke   = {37,52,68,83,100,125,151,187,233,280};
+
+  vector<double> b_minus = {1, .833, .805, .78, .751, .731, .759, .793, .735, .765};
+  vector<double> k_minus = {10.8, 22.12, 23.9, 25.64, 28.52, 29.43, 24.2, 17.46, 21.56, 16.12};
+
+
+  cout<<b_plus.size()<<" "<<k_plus.size()<<" "<<t_ke.size()<<endl;
+  cout<<b_minus.size()<<" "<<k_minus.size()<<endl;
+
+  bp_spl = new TSpline3("bp_spl",t_ke.data(),b_plus.data(),t_ke.size());
+  TGraph *bp_g = new TGraph(t_ke.size(),t_ke.data(),b_plus.data());
+
+  kp_spl = new TSpline3("kp_spl",t_ke.data(),k_plus.data(),t_ke.size());
+  TGraph *kp_g = new TGraph(t_ke.size(),t_ke.data(),k_plus.data());
+
+  bm_spl = new TSpline3("bm_spl",t_ke.data(),b_minus.data(),t_ke.size());
[…]

### 2020-02-08 chapters/ch_results.tex
Pion spectra added

@@
 
-\section{$\pi$ Yields}
+\section{Pion Spectra and Yields}
 Add figure on Glauber model 
@@
 
+\section{Pion Double Ratio}
+Add figure of Theory for pion ratios
+
 \section{Pion Inferometry}

### 2020-02-08 macrosForThesis/DigiTasks/particleCalib.C

@@ -1,3 +1,7 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
 auto me = 0.5109989461;                                                                                
 auto mp = 938.2720813;                                                                                 
 auto mn = 939.565346;                                                                                  
@@ -16,17 +20,32 @@ void particleCalib()
   TF1 *p_lin = new TF1("p_lin","TMath::Exp([0] + [1]*x)",0,max);
   p_lin->SetParameters(-.05823,-1.78728e-4);
 
+  TF1 *p_lin_new = new TF1("p_lin_new","[0] + [1]*x",0,max);
+  p_lin_new->SetParameters(0.861124,-9.88284e-05);
+
   TF1 *d_lin = new TF1("d_lin","TMath::Exp([0] + [1]*x)",0,max);
   d_lin->SetParameters(-.0487855,-1.68159e-4);
+
+  TF1 *d_lin_2 = new TF1("d_lin_2","[0] + [1]*x",0,max);
+  d_lin_2->SetParameters(0.920522,-0.000133518);
 
   TF1 *t_lin = new TF1("t_lin","TMath::Exp([0] + [1]*x)",0,max);
   t_lin->SetParameters(-.0715918,-1.20368e-4);
 
+  TF1 *t_lin_new = new TF1("t_lin_new","[0] + [1]*x",0,max);
+  t_lin_new->SetParameters(0.911872,-0.000109902);
+
   TF1 *he3_lin = new TF1("he3_lin","TMath::Exp([0] + [1]*x)",0,max);
   he3_lin->SetParameters(-.0486119,-1.137e-4);
 
+  TF1 *he3_lin_new = new TF1("he3_lin_new","[0] + [1]*x",0,max);
+  he3_lin_new->SetParameters(0.964982,-0.000137747);
+
   TF1 *he4_lin = new TF1("he4_lin","TMath::Exp([0] + [1]*x)",0,max);
   he4_lin->SetParameters(-.0628344,-1.05124e-4);
+
+  TF1 *he4_lin_new = new TF1("he4_lin_new","[0] + [1]*x",0,max);
+  he4_lin_new->SetParameters(0.964982,-0.000137747);
 
   p_lin->SetLineColor(kRed);
[…]

### 2020-02-08 macrosForThesis/MatchMCtoData/MatchPOCA/publishPOCA.C

@@ -2,30 +2,28 @@
 
 using namespace style;
 
-void publishClusterNum()
+void publishPOCA()
 {
   gStyle->SetErrorX(0);
 
   TFile *f = TFile::Open("../angles_poca.root");
-  TH1D *mc_p = (TH1D *)f->Get("ndf_mc_p");
-  TH1D *data_p = (TH1D *)f->Get("ndf_p");
+  TH1D *mc_p = (TH1D *)f->Get("poca_mc_p");
+  TH1D *data_p = (TH1D *)f->Get("poca_p");
 
-  TH1D *mc_d = (TH1D *)f->Get("ndf_mc_d");
-  TH1D *data_d = (TH1D *)f->Get("ndf_d");
+  TH1D *mc_d = (TH1D *)f->Get("poca_mc_d");
+  TH1D *data_d = (TH1D *)f->Get("poca_d");
 
-  TH1D *mc_t = (TH1D *)f->Get("ndf_mc_t");
-  TH1D *data_t = (TH1D *)f->Get("ndf_t");
+  TH1D *mc_t = (TH1D *)f->Get("poca_mc_t");
+  TH1D *data_t = (TH1D *)f->Get("poca_t");
 
-  TH1D *mc_he3 = (TH1D *)f->Get("ndf_mc_he3");
-  TH1D *data_he3 = (TH1D *)f->Get("ndf_he3");
+  TH1D *mc_he3 = (TH1D *)f->Get("poca_mc_he3");
+  TH1D *data_he3 = (TH1D *)f->Get("poca_he3");
 
-  TH1D *mc_he4 = (TH1D *)f->Get("ndf_mc_he4");
-  TH1D *data_he4 = (TH1D *)f->Get("ndf_he4");
+  TH1D *mc_he4 = (TH1D *)f->Get("poca_mc_he4");
+  TH1D *data_he4 = (TH1D *)f->Get("poca_he4");
 
-  mc_p->GetYaxis()->SetTitle("Probability");
-  mc_p->GetXaxis()->SetTitle("# of Clusters");
   data_p->GetYaxis()->SetTitle("Probability");
-  data_p->GetXaxis()->SetTitle("# of Clusters");
+  data_p->GetXaxis()->SetTitle("Dist. to Vertex [mm]");
 
   data_p->GetYaxis()->SetMaxDigits(2);
   mc_p->GetYaxis()->SetMaxDigits(2);
[…]

### 2020-02-08 macrosForThesis/MatchMCtoData/MatchPOCA/publishPOCA_divide.C

@@ -1,0 +1,277 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishPOCA_divide()
+{
+  gStyle->SetErrorX(0);
+
+  TFile *f = TFile::Open("../angles_poca.root");
+  TH1D *mc_p = (TH1D *)f->Get("poca_mc_p");
+  TH1D *data_p = (TH1D *)f->Get("poca_p");
+
+  TH1D *mc_d = (TH1D *)f->Get("poca_mc_d");
+  TH1D *data_d = (TH1D *)f->Get("poca_d");
+
+  TH1D *mc_t = (TH1D *)f->Get("poca_mc_t");
+  TH1D *data_t = (TH1D *)f->Get("poca_t");
+
+  TH1D *mc_he3 = (TH1D *)f->Get("poca_mc_he3");
+  TH1D *data_he3 = (TH1D *)f->Get("poca_he3");
+
+  TH1D *mc_he4 = (TH1D *)f->Get("poca_mc_he4");
+  TH1D *data_he4 = (TH1D *)f->Get("poca_he4");
+
+  data_p->GetYaxis()->SetTitle("Probability");
+  data_p->GetXaxis()->SetTitle("Dist. to Vertex [mm]");
+
+  //  data_p->GetYaxis()->SetMaxDigits(2);
+  //  mc_p->GetYaxis()->SetMaxDigits(2);
+  
+  TCanvas *cvs = style::stdcvs();
+  make(data_p);
+  make(mc_p);
+  make(data_d);
+  make(mc_d);
+  make(data_t);
+  make(mc_t);
+  make(data_he3);
+  make(mc_he3);
+  make(data_he4);
+  make(mc_he4);
+
+  data_p->SetLineColor(kBlue);
+  mc_p->SetLineColor(kRed);
+
+  data_d->SetLineColor(kBlue - 4);
+  mc_d->SetLineColor(kRed - 4);
+
+  data_t->SetLineColor(kBlue - 7);
+  mc_t->SetLineColor(kRed - 7);
+
+  data_he3->SetLineColor(kBlue - 9);
+  mc_he3->SetLineColor(kRed - 9);
+
+  data_he4->SetLineColor(kBlue - 10);
+  mc_he4->SetLineColor(kRed - 10);
+
+
+  double scale_p = 6.25;
+  double scale_d = 6.25;
+  double scale_t = 6.25;
[…]

### 2020-02-08 macrosForThesis/PionSpectra/publishPionRatios.C

@@ -1,0 +1,101 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishPionRatios()
+{
+
+  TFile *f = TFile::Open("publishPions.root");
+  TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
+  TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
+
+  TH1D *doubleRatio = (TH1D *)f->Get("doubleRatio");
+
+  make(singleRatio_sn108);
+  make(singleRatio_sn132);
+  make(doubleRatio);
+
+  gStyle->SetEndErrorSize(0);
+
+  //style
+  int lineS132 = 1; 
+  int lineS108 = 7;
+
+  int markS132 = 21;
+  int markS108 = 25;  
+
+  //color
+  int lineC132 = kRed ; 
+  int lineC108 = kRed - 7;
+
+  int markC132 = kRed ;
+  int markC108 = kRed - 7;  
+
+  //size
+  int markSz132 = 2; //marker size
+  int markSz108 = 2; //marker size 
+
+  //  TLegend *leg = new TLegend(.2,.2,.55,.5);
+  TLegend *leg = new TLegend(.3,.5,.8,.8);
+  leg->AddEntry(singleRatio_sn132,"{}^{132}Sn +{}^{124}Sn","lpe");
+  leg->AddEntry(singleRatio_sn108,"{}^{108}Sn +{}^{112}Sn","lpe");
+
+  /*  TLegend *leg_2 = new TLegend(.75,.75,.95,.9);
+  leg_2->SetHeader("{}^{108}Sn +{}^{112}Sn");
+  leg_2->AddEntry(pim_sn108,"#pi^{-}","lpe");
+  leg_2->AddEntry(pip_sn108,"#pi^{+} (x10^{-1} )","lpe");
+  */
+
+  TCanvas *cvs = style::stdcvs();
+  //  cvs->SetLogy();
+  //  singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
+  
+  singleRatio_sn132->GetYaxis()->SetRangeUser(0,15);
+  singleRatio_sn132->GetYaxis()->SetTitle("#pi^{-}/#pi^{+}");
+  singleRatio_sn132->GetYaxis()->CenterTitle();
+  singleRatio_sn132->GetXaxis()->SetTitle("T_{#scale[.6]{COM}} (MeV)");
[…]

### 2020-03-05 chapters/ch_data_analysis.tex
Electron Background and some MC matching
Added space charge sections to thesis. match clusters data retrieved

@@
 \midrule
-p   & 882.8 & d & d & 903.5 & e & e &\phantom{abcdef} & f & f \\
-d   & 817.1 & d & d & 898.5 & e & e & 1621.1 & f & f\\
+p   & 882.8 & 929.5 & 877.3 & 903.5 & 929.5 & 889 &\phantom{abcdef} & f & f \\
+d   & 817.1 & d & d & 898.5 & e & e & 1621.1 & 1704 & 1612\\
 t   & 589.5 & d & d & 887 & e & e & 1612.4 & f & f  \\
@@
 \subsection{Space Charge Corrections}
-Overview
-Discuss the relevant time scales, drift lengths, magnitudes, and locations of space charge
 
@@
 \label{fig:beampaths}
+\end{figure}
+
+As the beam passes through the field cage it ionizes the gas along the way creating electron-ion pairs.
+The drift velocities of the ions are \num{1e4} times slower than the electron drift velocities.
+Because ions move very slow, the potential build up of positive ions may create a space charge which distorts the drifting electrons of the tracks, impacting the measurement of their momentum.
+There are several regions of the TPC in which electron-ion pairs are created.
+The largest source of positive ions are created in the avalanche process near the anode wires.
+These ions slowly drift toward the cathode and are captured by the gating grid.
+The other source of ions come from the primary ionization produced by the beam and reaction products in the detector gas.
+The energy loss <dE/dx>$\propto Z^2$, where Z is the charge of the particle type.
+Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2.5e3} times that of the light charged particles which mostly are of charge Z~1.
+ The beam is positioned about 25\si{\centi\metre} below the anode plane in the TPC.
+It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane.
+It takes the ions \num{5e4}\si{\micro\sec} to travel.
+The beam rate in the experiment varied around a value of about 10\si{\kilo\hertz}, which has an average occurrence of 1 beam every 100\si{\micro\sec}, which is much shorter than the time it takes for the ions created by each beam to terminate on the cathode plane.
+This results in a build up of positive ions in the shape of a sheet charge, carved out by the beam path as shown in Fig.~\ref{fig:spacechg_cartoon}.
+The average distance between sequential ion paths created by each beam is a spacing of about 50\si{\micro\metre} apart, and the average number of beam paths that compose the sheet charge is around 500 tracks.
+Though the arrival time of each beam track is random, the large number of tracks, and small inter beam spacing, allows us to approximate the sheet charge as an uniform sheet charge.
+
+The electric field in the presence of the sheet charge can be calculated by solving Poisson's equation,
+
+\begin{equation}
+\nabla^2 \phi = \rho,
+\end{equation}
+
+ where $\phi$ is the electric potential and $\rho$ is the free space charge; given the Dirichlet boundary conditions of the field cage.
+The electric field is given as the gradient of the potential,  $\vec{E}= -\nabla \phi$.
+
+ To reduce computation time, we notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a reference free charge $\rho_o$ and scale the solution for any other free charge.
+The full magnetic field map is provided by the SAMURAI collaboration \cite{magnet}.
+The velocity field map is given by Eq.~\ref{eq:elecdrift} and the electron drift through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration.
+The correction map is calculated by starting from the anode y-position and the measured (x,z) on the pad-plane and stepping backward in time in the Runge-Kutta integration through the velocity field map until the electron reaches the measured y-position.
+
+ It has been shown before that the amount of space charge present in the chamber is related to observables such as the distance of closest approach of each track to the vertex point \cite{starSC}.
+This is easily understood, as the space charge distorts the electrons as they drift through the chamber eventually distorting the overall shape of the measured track.
+In the presence of no space charge, you would correctly expect the distance of closest approach of each track to the vertex point would be a distribution centered around zero.
+In the presence of space charge the tracks are distorted and the distribution has some mean bias.
+
+Shown in Fig.~\ref{fig:sc_shift} is an example of the distortion map in the TPC for left-going tracks in blue, and right-going tracks in green, with the vectors of distortion magnified.
+The space charge affects left and right going tracks differently, with the right-going tracks going to higher momentum values and the left-going tracks going to lower momentum values.
+The inset figure of Fig.~\ref{fig:sc_shift} shows the shift in the x-position distance to vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
+We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR}$.
+
+Overview
+Discuss the relevant time scales, drift lengths, magnitudes, and locations of space charge
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{DVTP_raw.pdf}
+\caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks in the TPC for the }
+\label{fig:spacechg_cartoon}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{PeakLocation_vs_BR.pdf}
+\caption{$\Delta\mathrm{V}_\mathrm{LR}$ as a function of the beam rate. }
+\label{fig:spacechg_br}
+\end{figure}
+
+The average beam rate was recorded in each experimental run and slightly varied from run to run, due to beam production variations.
+The amount of space charge present in the field cage is directly proportional to the beam rate; therefore $\Delta\mathrm{V}_\mathrm{LR}$ is also proportional to the beam rate as shown in Fig.~\ref{fig:spacechg_br}.
+The only parameter in the space charge correction algorithm is the surface charge density $\sigma_{\mathrm{SC}}$.
+By varying $\sigma_{\mathrm{SC}}$ for a wide range of values the $\Delta\mathrm{V}_\mathrm{LR}$ observable is measured and plotted in the left panel of Fig.~\ref{fig:spacechg_relation}.
+The surface charge density which gives $\Delta\mathrm{V}_\mathrm{LR} = 0$ is taken to be the estimate for the average amount of space charge present.
+This is done for several runs which vary in beam intensity.
+Since the relation is linear between the surface charge density and the beam rate, a linear fit gives good agreement for interpolating the surface charge values as a function of beam rate, as seen in the right figure of Fig.~\ref{fig:spacechg_relation}.
+
+This is done for all the systems. I NEED TO PUT THE ASSUMPTIONS OF 132 108 HERE!!!
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{SC_Relation.pdf}
+\caption{Space charge relation}
+\label{fig:spacechg_relation}
+\end{figure}
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{Effect_SC.pdf}
+\caption{Shift of tracks}
+\label{fig:sc_shift}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_small_angle.pdf} 
+        \caption{Generic} \label{fig:mom_S_before}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_large_angle.pdf} 
+        \caption{Competitors} \label{fig:mom_L_before}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_aftercor_small_angle.pdf} 
+        \caption{Generic} \label{fig:mom_S_after}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_aftercor_large_angle.pdf} 
+        \caption{Competitors} \label{fig:mom_L_after}
+    \end{subfigure}
+\label{fig:mom_sc}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_frac_small_angle.pdf} 
+        \caption{Generic} \label{fig:mom_S_1D}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{BDC_P_frac_large_angle.pdf} 
+        \caption{Competitors} \label{fig:mom_L_1D}
+    \end{subfigure}
+
+\label{fig:mom_1D}
 \end{figure}
@@
 
+\begin{figure}[!hbt]
+\includegraphics[width=\linewidth]{pulseshape}
+\caption{Pulse shapes for normal and saturated signals. Both are normalized to a max height of 1.}
+\label{fig:pulseshape}
+\end{figure}
+
 \begin{equation}

### 2020-03-05 macrosForThesis/MatchMCtoData/MatchCluster/publishClusterNum.C

@@ -94,20 +94,20 @@ void publishClusterNum()
 
   data_p->GetXaxis()->SetRangeUser(18,110);
   data_p->GetYaxis()->SetRangeUser(.001,1);
-  //  data_p->Draw("");
-  //  mc_p->Draw("same");
+    data_p->Draw("");
+    mc_p->Draw("same");
 
-  //  data_d->Draw("same");
-  //  mc_d->Draw("same");
+    data_d->Draw("same");
+    mc_d->Draw("same");
 
-  //  data_t->Draw("same");
-  //  mc_t->Draw("same");
+  data_t->Draw("same");
+  mc_t->Draw("same");
 
   data_he3->Draw("same");
   mc_he3->Draw("same");
 
-  //  data_he4->Draw("same");
-  //  mc_he4->Draw("same");
+    data_he4->Draw("same");
+    mc_he4->Draw("same");
 
   data_p->SetMarkerStyle(20);
   mc_p->SetMarkerStyle(20);

### 2020-03-05 macrosForThesis/PionSpectra/CoulombPotential/fitCoulombPotential.C

@@ -1,0 +1,94 @@
+
+
+void fitCoulombPotential()
+{
+ TFile *f = TFile::Open("/home/justin/mythesis/macrosForThesis/PionSpectra/rootfiles/publishPions_sm.root");
+  TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
+  TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
+
+ TFile *g = TFile::Open("/home/justin/mythesis/macrosForThesis/PionSpectra/rootfiles/publishPions.root");
+  TH1D *pip_132 = (TH1D *)g->Get("pip_sn132");
+  TH1D *pim_132 = (TH1D *)g->Get("pim_sn132");
+  TH1D *pip_108 = (TH1D *)g->Get("pip_sn108");
+  TH1D *pim_108 = (TH1D *)g->Get("pim_sn108");
+
+  double s132_n = 132 + 124 - 100;
+  double s132_p = 100;
+
+  double s108_n = 108 + 112 - 100;
+  double s108_p = 100;
+
+  /*
+  double r132 = (5*pow(s132_n,2) + s132_n*s132_p)/(5*pow(s132_p,2) + s132_n*s132_p);
+  double r108 = (5*pow(s108_n,2) + s108_n*s108_p)/(5*pow(s108_p,2) + s108_n*s108_p);
+  */
+
+  double r132 = 2.4;
+  double r108 = 1.5;
+
+  cout<<"Sn132 ratio "<<r132<<endl;
+  cout<<"Sn108 ratio "<<r108<<endl;
+
+  double pi_mass = 139.57;
+  double sn132_temp = 0;
+  double sn108_temp = 0;
+  
+  /*
+  TF1 *blz = new TF1("blz","[0]*(x + [3] - [1])*TMath::Exp(-(x + [3])/[2])",0,300);
+  blz->SetParameters(100,10,30);
+  blz->FixParameter(3,pi_mass);
+  */
+  TF1 *blz = new TF1("blz","[0]*TMath::Exp(-(x + [1])/[2])",0,300);
+  blz->SetParameter(0,1000);
+  blz->SetParameter(2,20);
+  blz->FixParameter(1,pi_mass);
+
+  pip_132->Fit(blz,"","",100,300);
[…]

### 2020-03-05 macrosForThesis/PionSpectra/CutVariation/publishCutVariations_DR.C

@@ -92,11 +92,15 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 void publishCutVariations_DR( int this_var)
 {
+  //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
+
+  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
+  vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
 
   int bins = 6;
   int var  = 4;
   int default_p = 3;
-  double frac = 1.2;
+  double frac = 2;
 
   int lineSt = 6;
   int markCl = 1;
@@ -105,12 +109,12 @@ void publishCutVariations_DR( int this_var)
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("cutVariaiton_DR.root");
+  TFile *f = TFile::Open("cutVariaiton_DR_extend.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.25,.04,.3,.04,.02,"<|");
-  TArrow *arrowR = new TArrow(.75,.04,.8,.04,.02,"|>");
+  TArrow *arrowL = new TArrow(.22,.04,.28,.04,.02,"<|");
+  TArrow *arrowR = new TArrow(.78,.04,.84,.04,.02,"|>");
 
   gStyle->SetOptStat(0);
 
@@ -179,6 +183,16 @@ void publishCutVariations_DR( int this_var)
 	      // Size factors
 	      Float_t xFactor = pad[0][0]->GetAbsWNDC()/pad[i][iBin]->GetAbsWNDC();
 	      Float_t yFactor = pad[0][0]->GetAbsHNDC()/pad[i][iBin]->GetAbsHNDC();
+
+	      singleRatio[Ny - iBin][iVar]->GetXaxis()->ChangeLabel(1,-1,-1,-1,-1,-1,Form("%d",labels.at(iVar).at(0)));
[…]

### 2020-03-05 macrosForThesis/PionSpectra/CutVariation/publishCutVariations_DRall.C

@@ -17,7 +17,7 @@ void publishCutVariations_DRall()
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("cutVariaiton_DR.root");
+  TFile *f = TFile::Open("cutVariaiton_DR_extend.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];

### 2020-03-05 macrosForThesis/PionSpectra/CutVariation/publishCutVariations_sn108.C

@@ -90,13 +90,17 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn108(int this_var = 0)
+void publishCutVariations_sn108(int this_var = 0,TString filename = "blank")
 {
+  // vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
+
+  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
+  vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
 
   int bins = 6;
   int var  = 4;
   int default_p = 3;
-  double frac = 1.2;
+  double frac = 10;
 
   int lineSt = 6;
   int markCl = 1;
@@ -105,12 +109,12 @@ void publishCutVariations_sn108(int this_var = 0)
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("cutVariaiton_singleRatio.root");
+  TFile *f = TFile::Open("cutVariaiton_singleRatio_extendMult.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.25,.04,.3,.04,.02,"<|");
-  TArrow *arrowR = new TArrow(.75,.04,.8,.04,.02,"|>");
+  TArrow *arrowL = new TArrow(.22,.04,.28,.04,.02,"<|");
+  TArrow *arrowR = new TArrow(.78,.04,.84,.04,.02,"|>");
 
   gStyle->SetOptStat(0);
 
@@ -174,10 +178,20 @@ void publishCutVariations_sn108(int this_var = 0)
 	      pad[i][iBin]->SetFillStyle(4000);
 	      pad[i][iBin]->SetFrameFillStyle(4000);
 	      pad[i][iBin]->cd();
-
+	      // pad[i][iBin]->SetLogy();
 	      // Size factors
[…]

### 2020-03-05 macrosForThesis/PionSpectra/CutVariation/publishCutVariations_sn132.C

@@ -90,13 +90,17 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132(int this_var = 0)
+void publishCutVariations_sn132(int this_var = 1,TString filename = "blank")
 {
+  //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
+  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
+  vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
+    
   int bins = 6;
   int var  = 4;
   int default_p = 3;
-  double frac = 1.2;
+  double frac = 4.;
 
   int lineSt = 6;
   int markCl = 1;
@@ -105,12 +109,12 @@ void publishCutVariations_sn132(int this_var = 0)
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("cutVariaiton_singleRatio.root");
+  TFile *f = TFile::Open("cutVariaiton_singleRatio_extendMult.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.25,.04,.3,.04,.02,"<|");
-  TArrow *arrowR = new TArrow(.75,.04,.8,.04,.02,"|>");
+  TArrow *arrowL = new TArrow(.22,.04,.28,.04,.02,"<|");
+  TArrow *arrowR = new TArrow(.78,.04,.84,.04,.02,"|>");
 
   gStyle->SetOptStat(0);
 
@@ -147,7 +151,6 @@ void publishCutVariations_sn132(int this_var = 0)
   gStyle->SetFrameLineWidth(linew);
   gStyle->SetEndErrorSize(5);
 
-
   for(int iBin = 1; iBin <= 6; iBin++)
     {
       for(int iVar = 0; iVar < var; iVar++)
[…]

### 2020-03-05 macrosForThesis/PionSpectra/PlotDataTheory/publishDoubleRatio.C

@@ -259,12 +259,12 @@ TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K265EM070/input
   p_stiff->SetFillStyle(0);
   p_stiff->SetBorderSize(0);
 
-  TLegend *leg_soft = new TLegend(.15,.55,.4,.85);
-  leg_soft->AddEntry(pBUU_soft_no_DR,"pBUU","lpe");
+  TLegend *leg_soft = new TLegend(.15,.7,.4,.85);
+  //  leg_soft->AddEntry(pBUU_soft_no_DR,"pBUU","lpe");
   leg_soft->AddEntry(xbuu_DR,"#chiBUU","lpe");
   leg_soft->AddEntry(tuQMD_05_DR,"TuQMD","lpe");
-  leg_soft->AddEntry(amdJAM55_DR,"AMD+JAM","lpe");
-  leg_soft->AddEntry(urQMD_46_DR,"urQMD","lpe");
+  //  leg_soft->AddEntry(amdJAM55_DR,"AMD+JAM","lpe");
+  //  leg_soft->AddEntry(urQMD_46_DR,"urQMD","lpe");
 
   leg_soft->SetBorderSize(0);
   leg_soft->SetFillStyle(0);
@@ -272,38 +272,40 @@ TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K265EM070/input
   p_soft->Draw();
   p_stiff->Draw();
   
-  TLegend *leg_stiff = new TLegend(.15,.55,.4,.85);
-  leg_stiff->AddEntry(pBUU_stiff_no_DR,"pBUU","lpe");
+  TLegend *leg_stiff = new TLegend(.35,.7,.6,.85);
+  //  leg_stiff->AddEntry(pBUU_stiff_no_DR,"pBUU","lpe");
   leg_stiff->AddEntry(xbuu120_DR,"#chiBUU","lpe");
   leg_stiff->AddEntry(tuQMD_20_DR,"TuQMD","lpe");
-  leg_stiff->AddEntry(amdJAM152_DR,"AMD+JAM","lpe");
-  leg_stiff->AddEntry(urQMD_104_DR,"urQMD","lpe");
+  //  leg_stiff->AddEntry(amdJAM152_DR,"AMD+JAM","lpe");
+  //  leg_stiff->AddEntry(urQMD_104_DR,"urQMD","lpe");
 
   leg_stiff->SetFillStyle(0);
[…]

### 2020-03-05 macrosForThesis/PionSpectra/PlotDataTheory/publishSingleSumAll.C

@@ -1,0 +1,384 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+void publishSingleSumAll()
+{
+  TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
+  TGraphErrors *pBUU_soft_no_132 = (TGraphErrors *) pBUU_soft_no_f->Get("piRatio132");
+  TGraphErrors *pBUU_soft_no_108 = (TGraphErrors *) pBUU_soft_no_f->Get("piRatio108");
+  
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff_r-1.root");
+  TGraphErrors *pBUU_stiff_no_132 = (TGraphErrors *) pBUU_stiff_no_f->Get("piRatio132");
+  TGraphErrors *pBUU_stiff_no_108 = (TGraphErrors *) pBUU_stiff_no_f->Get("piRatio108");
+  
+  TFile *amdJAM55_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML55new_r-1.root");
[…]

### 2020-03-05 macrosForThesis/PionSpectra/PlotDataTheory/publishSymEn.C

@@ -1,0 +1,217 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishSymEn()
+{
+  TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
+  TF1 *pBUU_soft_no_DR = (TF1 *) pBUU_soft_no_f->Get("Srho");
+
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff_r-1.root");
+  TF1 *pBUU_stiff_no_DR = (TF1 *) pBUU_stiff_no_f->Get("Srho");
+
+  TFile *amdJAM55_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML55new_r-1.root");
+  TF1 *amdJAM55_DR = (TF1 *) amdJAM55_f->Get("Srho");
+
+  TFile *amdJAM152_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML152new_r-1.root");
+  TF1 *amdJAM152_DR = (TF1 *) amdJAM152_f->Get("Srho");
+
+
+  TFile *urQMD_46_f = TFile::Open("/home/justin/Homework/UrQMD/input_for_esym.info/UrQMDL46_r-1.root");
+  TF1 *urQMD_46_DR = (TF1 *) urQMD_46_f->Get("Srho");
+
+  TFile *urQMD_104_f = TFile::Open("/home/justin/Homework/UrQMD/input_for_esym.info/UrQMDL104_r-1.root");
+  TF1 *urQMD_104_DR = (TF1 *) urQMD_104_f->Get("Srho");
+
+
+  TFile *imQMD_05_f = TFile::Open("/home/justin/Homework/IQMD-BNU/input_for_esym.info/IQMD-BNUg05_r-1.root");
+  TF1 *imQMD_05_DR = (TF1 *) imQMD_05_f->Get("Srho");
+
[…]

### 2020-03-05 macrosForThesis/PionSpectra/PlotDataTheory/test.C

@@ -1,0 +1,595 @@
+void test()
+{
+//=========Macro generated from canvas: cvs/cvs
+//=========  (Mon Feb 24 18:44:50 2020) by ROOT version 6.13/02
+
+   gStyle->SetCanvasPreferGL(kTRUE);
+
+   TCanvas *cvs = new TCanvas("cvs", "cvs",100,64,1500,1000);
+   gStyle->SetOptStat(0);
+   cvs->Range(-35.12195,-0.511751,257.561,0.993399);
+   cvs->SetFillColor(0);
+   cvs->SetBorderMode(0);
+   cvs->SetBorderSize(2);
+   cvs->SetLogy();
+   cvs->SetLeftMargin(0.12);
+   cvs->SetRightMargin(0.06);
+   cvs->SetTopMargin(0.06);
+   cvs->SetBottomMargin(0.14);
+   cvs->SetFrameLineWidth(10);
+   cvs->SetFrameBorderMode(0);
+   cvs->SetFrameLineWidth(10);
+   cvs->SetFrameBorderMode(0);
+   Double_t xAxis1[7] = {0, 30, 60, 90, 120, 170, 240}; 
+   
+   TH1D *singleRatio_sn108__1 = new TH1D("singleRatio_sn108__1","",6, xAxis1);
+   singleRatio_sn108__1->SetBinContent(1,7.120791);
+   singleRatio_sn108__1->SetBinContent(2,2.282283);
+   singleRatio_sn108__1->SetBinContent(3,1.311681);
+   singleRatio_sn108__1->SetBinContent(4,0.9909349);
+   singleRatio_sn108__1->SetBinContent(5,0.8208762);
+   singleRatio_sn108__1->SetBinContent(6,0.6800425);
+   singleRatio_sn108__1->SetBinContent(7,0.8924082);
+   singleRatio_sn108__1->SetBinError(1,0.2198919);
+   singleRatio_sn108__1->SetBinError(2,0.04890967);
+   singleRatio_sn108__1->SetBinError(3,0.03067687);
+   singleRatio_sn108__1->SetBinError(4,0.02925379);
+   singleRatio_sn108__1->SetBinError(5,0.02937922);
+   singleRatio_sn108__1->SetBinError(6,0.0422841);
+   singleRatio_sn108__1->SetBinError(7,0.1593297);
+   singleRatio_sn108__1->SetMinimum(0.5);
[…]

### 2020-03-05 macrosForThesis/PionSpectra/SpaceChargeVariation/plotLeftRight.C

@@ -1,0 +1,154 @@
+
+
+void plotLeftRight()
+{
+
+  TFile *sn108_sc = TFile::Open("analysis-Sn108-nominal_LR.root");
+  TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
+
+  TFile *sn108_nosc = TFile::Open("analysis-Sn108-noSC.root");
+  TFile *sn132_nosc = TFile::Open("analysis-Sn132-noSC.root");
+
+  TH1D *pim_108_L = (TH1D *)sn108_sc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R = (TH1D *)sn108_sc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L = (TH1D *)sn108_sc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R = (TH1D *)sn108_sc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L = (TH1D *)sn132_sc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R = (TH1D *)sn132_sc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L = (TH1D *)sn132_sc->Get("pipkecm_ec_L");
+  TH1D *pip_132_R = (TH1D *)sn132_sc->Get("pipkecm_ec_R");
+
+  TH1D *pim_108_L_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_L");
+  TH1D *pip_132_R_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_R");
+
+  pim_108_L->Scale(1./pim_108_L->Integral(),"width");
+  pim_108_R->Scale(1./pim_108_R->Integral(),"width");
[…]

### 2020-03-05 macrosForThesis/PionSpectra/TheoryBootstrap/theoryBootStrapBCA_108Single.C

@@ -1,0 +1,204 @@
+TRandom3 *ran = new TRandom3(12345);
+
+void loadVec(vector< vector<double>> &d, TGraphErrors *g)
+{
+  int num_bins = g->GetN();
+   for(int iBin = 0; iBin < num_bins; iBin++)
+    {
+      double x,y;
+      g->GetPoint(iBin,x,y);
+      d.at(iBin).push_back(y);
+    }
+}
+
+
+vector<double> SampleBoot(vector< vector<double>> data)
+{
+  vector<double> boot(data.size(),0);
+
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      int num_th = data.at(iBin).size(); //# of theories
+
+      for(int i = 0; i < num_th; i++)
+	boot.at(iBin) += data.at(iBin).at(ran->Integer(num_th));
+
+      boot.at(iBin) /= num_th;
+    }
+  
+  return boot;
+}
+
+
+vector<TH1D *> BootStrap(int n_throws, vector< vector<double>> data)
+{
+  vector<TH1D *> hist_vec;
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      TH1D *h;
+      if(iBin < 1)
+	h = new TH1D(Form("bin_%d",iBin),Form("bin_%d",iBin),10,0,40);
+      else
+	h = new TH1D(Form("bin_%d",iBin),Form("bin_%d",iBin),100,0,15);
+      hist_vec.push_back(h);
+    }
+
+  for(int iBoot = 0; iBoot < n_throws; iBoot++)
+    {
+      auto vec = SampleBoot(data);
+      for(int iBin = 0; iBin < vec.size(); iBin++)
+	hist_vec.at(iBin)->Fill(vec.at(iBin));
+    }
+
+  return hist_vec;
+}
+
+
+void theoryBootStrapBCA_108Single()
+{
+
+    TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
+  TGraphErrors *pBUU_soft_no_DR = (TGraphErrors *) pBUU_soft_no_f->Get("piRatio108");
+
[…]

### 2020-04-04 chapters/ch_data_analysis.tex
General Backup

@@
 
+MAYBE ADD A LITTLE NAPKIN CALCULATION OF SPACE CHARGE TO SHOW ORDER OF EFFECT.
+
+\begin{table}[!htp] % not just 'h!'
+\centering % not a center environment
+\begin{tabular}{
+  @{}
+  l
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=5.2]
+  S[table-format=5.2]
+  @{}
+}
+\toprule
+Beam Energy Loss  &
+ {${}^{132}$Sn} &
+ {${}^{124}$Sn} &
+ {${}^{112}$Sn} &
+ {${}^{108}$Sn} &
+  {Avg.}\\
+
+\midrule
+$\si{\kilo\eV\per\centi\meter}$ & 11.2   &.034  &5.43   &  903   &150     \\
+\bottomrule
+\end{tabular}
+
+\caption{Average energy loss of each beam.}
+\label{tb:beameloss}
+\end{table}
+
 \begin{figure}[H]
@@
 
-This is done for all the systems. I NEED TO PUT THE ASSUMPTIONS OF 132 108 HERE!!!
+This is done for all the systems. I NEED TO PUT THE ASSUMPTIONS of SPACE CHARGE SHEET DENSITY OF 132 108 HERE!!!
+
+From the BDC tracking, the precise interaction point on target is known.
+Using this point greatly improves the momentum resolution of the track fitting, but there is a systematic shift when comparing the momentum value without using the vertex BDC point in the fit.
+The space charge affects right-going and left-going tracks differently making right-going tracks more rigid (higher momentum) and left-going tracks less rigid (lower momentum).
+For polar angles of $\theta < 40 \deg$, the disagreement between momentum values fitted with the BDC, versus without, are much less because the projection of the track does not disagree as much with the BDC point as tracks with polar angles of $\theta > 40 \deg$.
+This is seen in
 
@@
 \end{figure}
-
-\begin{table}[!htp] % not just 'h!'
-\centering % not a center environment
-\begin{tabular}{
-  @{}
-  l
-  S[table-format=1.2]
-  S[table-format=1.2]
-  S[table-format=1.2]
-  S[table-format=5.2]
-  S[table-format=5.2]
-  @{}
-}
-\toprule
-Beam Energy Loss  &
- {${}^{132}$Sn} &
- {${}^{124}$Sn} &
- {${}^{112}$Sn} &
- {${}^{108}$Sn} &
-  {Avg.}\\
-
-\midrule
-$\si{\kilo\eV\per\centi\meter}$ & 11.2   &.034  &5.43   &  903   &150     \\
-\bottomrule
-\end{tabular}
-
-\caption{Average energy loss of each beam.}
-\label{tb:beameloss}
-\end{table}
 
@@
 \subsection{MC and Data Comparison}
+
+\begin{figure}[!hbt]
+\includegraphics[width=\linewidth]{numcluster.png}
+\caption{Comparison of MC to data.}
+\label{fig:clustcomp}
+\end{figure}
+
+\begin{figure}[!hbt]
+\includegraphics[width=\linewidth]{poca.png}
+\caption{Comparison of MC to data.}
+\label{fig:pocacomp}
+\end{figure}
+
 Add Figure of Pad response function for pion,proton.... for MC vs Data vs angles...
@@
 Add Figure of track residuals? MC vs data?
+
+\section{Aligning TPC}
+
+\section{CoBo timing correction}
 

### 2020-04-04 chapters/ch_data_analysis_2.tex
Adding better tables and gas section. Macros folder added
Pion spectra added
General Backup

@@
 \chapter{Data Analysis II: Extracting Physics}
-\section{Beam Particle Identification}
-\section{Event Mixing}
-\section{Reaction Plane Determination}
+
+\section{Edge Cuts}
+As seen in Fig.~\ref{fig:edge}, edges of the measurement volume of the TPC we can see several edge effects distortion the clusters positions.
+The last pads on the edges of the pad-plane naturally have no neighbors on the outsides, therefore their cluster position is biased towards the inside of the TPC.
+Also since we measure a finite window in time for the vertical direction, pulses near the edge of the time-bucket window are clipped and the reconstructed times are biased towards inside the TPC volume.
+
+The number of clusters affected is small but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
+Simple cuts around the left,right, top, and bottom of the TPC remove these clusters from the analysis,
+
+\begin{equation*}
+  |x|\geq420~\mathrm{mm},\quad y\leq-522+\mathrm{y_o}~\mathrm{mm},
+  \quad\mathrm{and}\quad y\geq-64+\mathrm{(Hit\ Shift)}~\mathrm{mm}.
+  \label{eq:hitshift}
+\end{equation*}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{clusterLR.png} 
+        \caption{Generic} \label{fig:mom_S_before}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{clusterTop.png} 
+        \caption{Competitors} \label{fig:mom_L_before}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{clusterBottom.png} 
+        \caption{Generic} \label{fig:mom_S_after}
+    \end{subfigure}
+
+\label{fig:edge}
+\end{figure}
+
+\section{High Density Cut}
+The density of tracks near the target region is very high and most of electronics information from this region is inaccurate in both timing and charge value.
+Hits lying within an semi-ellipsoidal cut region around the target are removed from the software and not included in the track and momentum reconstruction.
+
+Add picture of ellipsoid
+
+\section{Beam angle selection}
+The beam is bent by the magnetic field before hitting the target.
+The BDCs track the beam outside of the magnet.
+The beam is then propagated through the magnetic field until the target position.
+The beam angle on target is categorized by two angles $\theta_{a,proj}$ and $\theta_
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
+        \caption{Generic} \label{fig:mom_S_before}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
+        \caption{Competitors} \label{fig:mom_L_before}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
+        \caption{Generic} \label{fig:mom_S_after}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
+        \caption{Competitors} \label{fig:mom_L_after}
+    \end{subfigure}
+\label{fig:mom_sc}
+\end{figure}
+
+\section{Vertex Cut}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{vertex-Sn108.png} 
+        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{vertex-Sn112.png} 
+        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{vertex-Sn124.png} 
+        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{vertex-Sn132.png} 
+        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
+    \end{subfigure}
+\label{fig:vertexdist}
+\end{figure}
+
+\section{Impact Parameter Selection}
+
+\section{Track Quality Cuts}
+
+\section{Angular Quality Cuts}

### 2020-04-04 chapters/ch_intro.tex
General Backup

--- 2019-06-09 chapters/ch_intro.tex
@@
 \chapter{Introduction}
+
 \section{From Nuclear Forces to the Equation of State}
+Isospin as a good quantum number at low energies 
+Figure showing the nuclear forces for the pp, pn, nn 
+Pauli exclusion 
+inter particle distance , saturation density (maybe figure of all densities?)
+Building the infinite matter
+Statistical model 
+Forces manifest in Energy/particle
+
+\section{The Nuclear Equation of State}
+Figure showing the binding energy vs p-n asym
+Liquid drop model, mass equation, move to higher densities
+Symmetric EoS asymmetric EoS(symmetry energy)
+Density dependence of the symmetry energy
+
 \section{Phases of Nuclear Matter}
-\section{The Nuclear Equation of State}
+gas liquid phase, gas, where are we
+Progression through the heavy ion collision 
+liquid, liquid gas, gas
+
+\section{Studying EoS through Heavy Ion Collisions}
+going from infinite matter to finite matter 
+approximate nuclear matter with neutron rich radioactive ion beams
+probe different asymmetries 
+probe different densities with different beam energies 
+Symmetry energy goes down at higher beam energies
+
+\section{Boltzmann Ulong Uhlenbeck (BUU) Transport Code}
+Does not reach equilibrium for all observable (hadrons) what about pions?
+Build up a mean field picture (momentum dependent) unknown quantities here
+Non equilibrium can be solved with Boltzman equation with collision term and solved by MC
+Figure of transport simulation code showing collision progression 
+Clustering is an issue 
+Meson, resonance production
+
+\section{Observables of interest to the EoS}
+Particle yields of isospin opposites 
+Flow??
+Figure showing p-n observable lessens at high energy and from all densities
+
+\section{Pion Production }
+Figure showing delta resonance 
+Figure showing pion's produced at 2po 
+Figure showing fermi motion effect (not really sub threshold but ok)
+(chemical potential model, delta isobar) prediction for pion ratio
+pion mass is small momentum shifted by coulomb affecting shape of pion spectra
+
 \section{Previous Constraints}
+Figure showing GW constraint and other previous constraints 
+FOPI data at 400 A MeV not as sensitive to symmetry energy 
+Saturation density and lower constrained but issues at high density
+conflicting analysis on FOPI data which was not intended to be used for Symm Energy
+Can you group all the constraints to one nice plot???
+
 \section{Motivation for building the S$\pi$RIT TPC}
+Based on EOS TPC
+High efficiency for detecting pions and other light charged particles 
+Effort to reduce the experimental error bars on such spectra which theory could use

### 2020-04-04 macrosForThesis/PionSpectra/CoulombPotential/fitCoulombPotential.C

@@ -23,12 +23,17 @@ void fitCoulombPotential()
   double r108 = (5*pow(s108_n,2) + s108_n*s108_p)/(5*pow(s108_p,2) + s108_n*s108_p);
   */
 
-  double r132 = 2.4;
-  double r108 = 1.5;
+  //r132 from N/Z  2.09
+  //r108 from N/Z  1.36
+
+  //Data driven
+  double r132 = 4.83;
+  double r108 = 2.04;
 
   cout<<"Sn132 ratio "<<r132<<endl;
   cout<<"Sn108 ratio "<<r108<<endl;
 
+  double coul_factor = 6 * 100 * 1.44/5; //fm * MeV
   double pi_mass = 139.57;
   double sn132_temp = 0;
   double sn108_temp = 0;
@@ -79,16 +84,52 @@ void fitCoulombPotential()
   cout<<"Sn132 avg. "<<sn132_temp<<endl;
   cout<<"Sn108 avg. "<<sn108_temp<<endl;
 
+  //Total E
   //  TF1 *fcn = new TF1("fcn","[0]*TMath::Exp(-2*[1]/[2])*(x + [1])/(x - [1])*TMath::Sqrt( (pow(x + [1],2) - pow([3],2))/(pow(x - [1],2) - pow([3],2)) )",100,300);
-  TF1 *fcn = new TF1("fcn","[0]*TMath::Exp(-2*[1]/[2])*(x + [3] + [1])/(x + [3] - [1])*TMath::Sqrt( (pow(x + [3] + [1],2) - pow([3],2))/(pow(x + [3] - [1],2) - pow([3],2)) )",30,300);
 
-  //  TF1 *fcn = new TF1("fcn","[0]*(1 + 2*[1]*( [2]/(pow(x + [2],2) - pow([2],2)) - 1/[3]))",20,300);
+
+  //K.E.
+  //  TF1 *fcn = new TF1("fcn","[0]*TMath::Exp(-2*[1]/[2])*(x + [3] + [1])/(x + [3] - [1])*TMath::Sqrt( (pow(x + [3] + [1],2) - pow([3],2))/(pow(x + [3] - [1],2) - pow([3],2)) )",30,300);
+
+  //Solve root
+ 
[…]

### 2020-04-04 macrosForThesis/PionSpectra/PlotDataTheory/DanCozma/April03/publishPion_132Ratio.C

@@ -1,0 +1,199 @@
+
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishPion_132Ratio()
+{
+  TFile *f = TFile::Open("/home/justin/mythesis/macrosForThesis/PionSpectra/rootfiles/publishPions_sm.root");
+  TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
+  TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
+
+  TH1D *doubleRatio = (TH1D *)f->Get("doubleRatio");
+
+  make(singleRatio_sn108);
+  make(singleRatio_sn132);
+  make(doubleRatio);
+
+  gStyle->SetEndErrorSize(0);
+
+  //style
+  int lineS132 = 1; 
+  int lineS108 = 7;
+
+  int markS132 = 21;
+  int markS108 = 25;  
+
+  //color
+  int lineC132 = kRed ; 
+  int lineC108 = kRed - 7;
+
+  int markC132 = kRed ;
+  int markC108 = kRed - 7;  
+
+  //size
+  int markSz132 = 2; //marker size
+  int markSz108 = 2; //marker size 
+
+  //  TLegend *leg = new TLegend(.2,.2,.55,.5);
+  TLegend *leg = new TLegend(.3,.5,.8,.8);
+  leg->AddEntry(singleRatio_sn132,"{}^{132}Sn +{}^{124}Sn","lpe");
+  leg->AddEntry(singleRatio_sn108,"{}^{108}Sn +{}^{112}Sn","lpe");
+
+  /*  TLegend *leg_2 = new TLegend(.75,.75,.95,.9);
+  leg_2->SetHeader("{}^{108}Sn +{}^{112}Sn");
+  leg_2->AddEntry(pim_sn108,"#pi^{-}","lpe");
+  leg_2->AddEntry(pip_sn108,"#pi^{+} (x10^{-1} )","lpe");
+  */
+
+  TCanvas *cvs = style::stdcvs();
+  //  cvs->SetLogy();
+  //  singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
+  
+  singleRatio_sn132->GetYaxis()->SetRangeUser(0,15);
+  singleRatio_sn132->GetYaxis()->SetTitle("#pi^{-}/#pi^{+}");
+  singleRatio_sn132->GetYaxis()->CenterTitle();
[…]

### 2020-04-04 macrosForThesis/PionSpectra/PlotDataTheory/publishPion_132Ratio.C

@@ -415,16 +415,21 @@ TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K265EM070/input
   singleRatio_sn132->Draw("E1");
   boot_stiff->GetXaxis()->SetRangeUser(0,240);
   boot_stiff->GetYaxis()->SetRangeUser(.9,40);
-  boot_stiff->Draw("CO E3");
+  //  boot_stiff->Draw("CO E3");
   
   boot_soft->SetFillColorAlpha(kBlue,.2);
+  /*
   boot_soft->Draw("same CO E3");
   soft_upper->Draw("C");
   soft_lower->Draw("C");
   
   stiff_upper->Draw("C");
   stiff_lower->Draw("C");
+  */
 
+
+  singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
+  singleRatio_sn132->GetXaxis()->SetRangeUser(0,300);
   singleRatio_sn132->DrawCopy("same E1");
 
   cvs->SetLogy();

### 2020-04-04 macrosForThesis/PionSpectra/SpaceChargeVariation/plotLeftRight.C

@@ -2,10 +2,12 @@
 
 void plotLeftRight()
 {
-
+  gStyle->SetOptStat(0);
+  
   TFile *sn108_sc = TFile::Open("analysis-Sn108-nominal_LR.root");
-  TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
-
+  //TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
+  TFile *sn132_sc = TFile::Open("analysis-Sn132-sn132_1_2_LR.root");
+  
   TFile *sn108_nosc = TFile::Open("analysis-Sn108-noSC.root");
   TFile *sn132_nosc = TFile::Open("analysis-Sn132-noSC.root");
 
@@ -29,6 +31,7 @@ void plotLeftRight()
   TH1D *pip_132_L_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_L");
   TH1D *pip_132_R_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_R");
 
+
   pim_108_L->Scale(1./pim_108_L->Integral(),"width");
   pim_108_R->Scale(1./pim_108_R->Integral(),"width");
   pip_108_L->Scale(1./pip_108_L->Integral(),"width");
@@ -38,7 +41,7 @@ void plotLeftRight()
   pim_132_R->Scale(1./pim_132_R->Integral(),"width");
   pip_132_L->Scale(1./pip_132_L->Integral(),"width");
   pip_132_R->Scale(1./pip_132_R->Integral(),"width");
-  
+
   pim_108_L_no->Scale(1./pim_108_L_no->Integral(),"width");
   pim_108_R_no->Scale(1./pim_108_R_no->Integral(),"width");
   pip_108_L_no->Scale(1./pip_108_L_no->Integral(),"width");
@@ -48,6 +51,30 @@ void plotLeftRight()
   pim_132_R_no->Scale(1./pim_132_R_no->Integral(),"width");
   pip_132_L_no->Scale(1./pip_132_L_no->Integral(),"width");
   pip_132_R_no->Scale(1./pip_132_R_no->Integral(),"width");
+
+
+  /*
+  pim_108_L->Scale(1.,"width");
+  pim_108_R->Scale(1.,"width");
+  pip_108_L->Scale(1.,"width");
[…]

### 2020-04-04 macrosForThesis/PionSpectra/TheoryBootstrap/Weight/DeltPot/theoryBootStrapBCA_108Single.C

@@ -1,0 +1,254 @@
+TRandom3 *ran = new TRandom3(12345);
+
+void loadVec(vector< vector<double>> &d, TGraphErrors *g)
+{
+  int num_bins = g->GetN();
+   for(int iBin = 0; iBin < num_bins; iBin++)
+    {
+      double x,y;
+      g->GetPoint(iBin,x,y);
+      d.at(iBin).push_back(y);
+    }
+}
+
+
+vector<double> SampleBoot(vector< vector<double>> data, int idx_max)
+{
+  double weight = .5;
+  vector<double> boot(data.size(),0);
+
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      int num_th = data.at(iBin).size(); //# of theories
+
+      for(int i = 0; i < num_th; i++)
+	{
+	  int rand_idx = ran->Integer(num_th);//ranges from 0 to num_th - 1; vector index
+	  //data that is > idx_max is weighted lower
+	  //compare the vector indicies
+	  
+	  if(rand_idx > idx_max)
+	    {
+	      if(ran->Uniform() > weight)
+		rand_idx = ran->Integer(idx_max + 1); //samples from 0 to idx_max
+	    }
+
+	  boot.at(iBin) += data.at(iBin).at(rand_idx);
+	}
+
+      boot.at(iBin) /= num_th;
+    }
+  
+  return boot;
+}
+
+
+vector<TH1D *> BootStrap(int n_throws, vector< vector<double>> data,int idx_max = 1e9)
+{
+  vector<TH1D *> hist_vec;
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      TH1D *h;
+      if(iBin < 1)
+	h = new TH1D(Form("bin_%d",iBin),Form("bin_%d",iBin),10,0,40);
+      else
+	h = new TH1D(Form("bin_%d",iBin),Form("bin_%d",iBin),100,0,15);
+      hist_vec.push_back(h);
+    }
+
+  for(int iBoot = 0; iBoot < n_throws; iBoot++)
+    {
+      auto vec = SampleBoot(data,idx_max);
+      for(int iBin = 0; iBin < vec.size(); iBin++)
+	hist_vec.at(iBin)->Fill(vec.at(iBin));
+    }
+
[…]

### 2020-04-04 macrosForThesis/SpaceCharge/SpaceChargeDensity/plotChargeDensity.C

@@ -1,0 +1,80 @@
+#include "/home/justin/mythesis/thesis_style.h"
+using namespace style;
+
+void plotChargeDensity()
+{
+
+  vector<double> run = {2272,2273,2274,2275,2276,2283,2284,2285,2286,2288,2289,2291,2310,2311,2314,2315,2320,2322,2323,2324,2325,2331,2332,2333,2334,2335,2336,2337,2340,2341,2362,2363,2368,2369,2370,2371,2372,2373,2374,2375,2378,2379,2380,2381,2382,2383,2384,2385,2386,2387,2388,2389,2391,2392,2393,2394,2395,2396,2397,2398,2399,2400,2401,2402,2429,2432,2433,2434,2437,2438,2439,2440,2442,2453,2461,2462,2463,2501,2502,2503,2505,2506,2507,2508,2509,2542,2543,2544,2546,2547,2548,2552,2553,2554,2555,2556,2557,2558,2559,2560,2562,2563,2564,2565,2566,2567,2568,2569,2570,2571,2572,2573,2574,2575,2578,2579,2580,2581,2582,2583,2584,2585,2586,2587,2588,2589,2590,2591,2592,2593,2594,2595,2596,2597,2598,2599,2600,2601,2617,2618,2619,2620,2621,2622,2623,2841,2843,2844,2845,2846,2848,2849,2850,2851,2852,2855,2856,2857,2858,2859,2860,2861,2875,2877,2878,2879,2880,2881,2882,2883,2884,2887,2888,2889,2890,2891,2892,2893,2894,2896,2898,2899,2900,2901,2902,2903,2904,2905,2907,2914,2916,2917,2919,2920,2921,2922,2924,2925,2926,2927,2929,2930,2931,2932,2933,2934,2935,2936,2939,2940,2941,2942,2943,2944,2945,2946,2948,2955,2956,2958,2959,2960,2961,2962,2964,2965,2966,2968,2969,2970,2971,2972,2973,2975,2976,2977,2978,2979};
+
+
[…]

### 2020-04-09 MSU-thesis-estee.tex
Added some preliminary delta res
Added space charge sections to thesis. match clusters data retrieved
General Backup
Routine Backup

@@
 \usepackage{graphicx}
-\graphicspath{ {images/} }
+\usepackage{cleveref}
+\usepackage{overpic}
+\usepackage{contour}
+\usepackage{color}
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/}}
 
@@
 % If you are using XeLaTeX or LuaLaTeX load the Times or Times New Roman font with \setmainfont
-\usepackage{newtxtext,newtxmath,booktabs,siunitx} 
+\usepackage{newtxtext,newtxmath,booktabs,siunitx,caption,subcaption} 
 %
@@
 \newcommand{\spirit}{S$\pi$RIT }
+\newcommand{\tin}[2]{{}^{#1}\mathrm{Sn} + {}^{#2}\mathrm{Sn}}
 \newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
+
 \includeonly{
@@
 chapters/ch_data_analysis,
-%chapters/ch_data_analysis_2,
+chapters/ch_data_analysis_2,
 %chapters/ch_results
@@
 \include{chapters/ch_data_analysis}
-\include{chapters/ch_data_analysis_2}
+%\include{chapters/ch_data_analysis_2}
 \include{chapters/ch_results}
@@
   % \end{appendices}
-%ch_nucleareos.txt~
+
 \appendix
-\chapter{Your appendix}
+\input{chapters/ch_appendix}
+
 %

### 2020-04-09 chapters/ch_appendix.tex
Routine Backup

\chapter{Runs analyzed in this data}
\newcommand{\hsn}{$^{132}$Sn+$^{124}$Sn}
\newcommand{\mhsn}{$^{124}$Sn+$^{112}$Sn}
\newcommand{\mlsn}{$^{112}$Sn+$^{124}$Sn}
\newcommand{\lsn}{$^{108}$Sn+$^{112}$Sn}

\begin{table}[H]
  \begin{center}
    \begin{tabular}{ccl}
      \hline
      System & \#runs & Run numbers \\
      \hline\hline
      \hsn & 113 & 2841, 2843, 2844, 2845, 2846, 2848, 2849, 2850, 2851, 2852, 2855, 2856, \\
      & & 2857, 2858, 2859, 2860, 2861, 2875, 2877, 2878, 2879, 2880, 2881, 2882, \\
      & & 2883, 2884, 2887, 2888, 2889, 2890, 2891, 2892, 2893, 2894, 2896, 2898, \\
      & & 2899, 2900, 2901, 2902, 2903, 2904, 2905, 2907, 2914, 2916, 2917, 2919, \\
      & & 2920, 2921, 2922, 2924, 2925, 2926, 2927, 2929, 2930, 2931, 2932, 2933, \\
      & & 2934, 2935, 2936, 2939, 2940, 2941, 2942, 2943, 2944, 2945, 2946, 2948, \\
      & & 2955, 2956, 2958, 2959, 2960, 2961, 2962, 2964, 2965, 2966, 2968, 2969, \\
      & & 2970, 2971, 2972, 2973, 2975, 2976, 2977, 2978, 2979, 2980, 2981, 2982, \\
      & & 2983, 2984, 2985, 2986, 2988, 2989, 2990, 2991, 2992, 2993, 2997, 2999, \\
      & & 3000, 3002, 3003, 3007, 3039 \\
      \hline
      \mhsn & 60 & 2542, 2543, 2544, 2546, 2547, 2548, 2552, 2553, 2554, 2555, 2556, 2557, \\
      & & 2558, 2559, 2560, 2562, 2563, 2564, 2565, 2566, 2567, 2568, 2569, 2570, \\
      & & 2571, 2572, 2573, 2574, 2575, 2578, 2579, 2580, 2581, 2582, 2583, 2584, \\
      & & 2585, 2586, 2587, 2588, 2589, 2590, 2591, 2592, 2593, 2594, 2595, 2596, \\
      & & 2597, 2598, 2599, 2600, 2601, 2617, 2618, 2619, 2620, 2621, 2622, 2623  \\
      \hline
      \mlsn & 68 & 3059, 3061, 3062, 3065, 3066, 3068, 3069, 3071, 3074, 3075, 3076, 3077, \\
      & & 3078, 3080, 3081, 3082, 3083, 3084, 3085, 3087, 3088, 3089, 3090, 3091, \\
      & & 3092, 3093, 3094, 3095, 3097, 3098, 3102, 3103, 3138, 3139, 3140, 3141, \\
      & & 3142, 3143, 3144, 3145, 3146, 3148, 3149, 3150, 3151, 3152, 3153, 3154, \\
      & & 3155, 3156, 3157, 3158, 3159, 3165, 3166, 3167, 3168, 3169, 3170, 3171, \\
      & & 3172, 3177, 3179, 3180, 3181, 3182, 3183, 3184 \\
      \hline
      \lsn & 85 & 2272, 2273, 2274, 2275, 2276, 2283, 2284, 2285, 2286, 2288, 2289, 2291, \\
      & & 2310, 2311, 2314, 2315, 2320, 2322, 2323, 2324, 2325, 2331, 2332, 2333, \\
      & & 2334, 2335, 2336, 2337, 2340, 2341, 2362, 2363, 2368, 2369, 2370, 2371, \\
      & & 2372, 2373, 2374, 2375, 2378, 2379, 2380, 2381, 2382, 2383, 2384, 2385, \\
      & & 2386, 2387, 2388, 2389, 2391, 2392, 2393, 2394, 2395, 2396, 2397, 2398, \\
      & & 2399, 2400, 2401, 2402, 2429, 2432, 2433, 2434, 2437, 2438, 2439, 2440, \\
      & & 2442, 2453, 2461, 2462, 2463, 2501, 2502, 2503, 2505, 2506, 2507, 2508, \\
      & & 2509 \\
      \hline
    \end{tabular}
    \caption{List of runs for the analysis. \label{tb:runList}}
  \end{center}
\end{table}
\chapter{Second appendix}

### 2020-04-09 chapters/ch_data_analysis.tex
Routine Backup

@@
 \ra{1.3}
+\begin{tabular}{@{}rrrrrrr@{}}\toprule
+& \multicolumn{3}{c}{$100 MeV Target$}\\
+\cmidrule{2-4}
+Particle &\phantom{abc} & Measured & Corrected & \% Difference & \% Difference\\
+\midrule
+p   & 882.8 & 929.5 & 877.3   &  3.7  & -1.0  \\
+d   & 817.1 & 831.15 & 797.94 &  1.7  & -2.3\\
+\bottomrule
+\end{tabular}
+\caption{Summary of expected cocktail. }
+\label{tb:cocktail100tar}
+\end{table*}
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrrrr@{}}\toprule
+& \multicolumn{3}{c}{$100 MeV$}\\
+\cmidrule{2-4}
+Particle & Expected & Measured & Corrected & \% Difference & \% Difference\\
+\midrule
+p   & 882.8 & 903.5 & 889   &  2.0   & -1.6  \\
+d   & 817.1 & 898.5 & 874.5 &  2.1   & -2.7\\
+\bottomrule
+\end{tabular}
+\caption{Summary of expected cocktail. }
+\label{tb:cocktail100}
+\end{table*}
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrrrr@{}}\toprule
+& \multicolumn{3}{c}{$300 MeV$}\\
+\cmidrule{2-4}
+Particle & Expected & Measured & Corrected & \% Difference Raw & \% Difference Corrected\\
+\midrule
+d   & 1621 & 1704 & 1612   &  5.1 & -0.6  \\
+t   & 1612 & 1691 & 1596   &  4.9  & -1.0\\
+${}^{4}$He   & 1613 & 1698 &  5.3 & 1595  & -1.1\\
+
+\bottomrule
+\end{tabular}
+\caption{Summary of expected cocktail. }
+\label{tb:cocktail300}
+\end{table*}
+
+\begin{table*}\centering
+\ra{1.3}
 \begin{tabular}{@{}rrrrcrrrcrrr@{}}\toprule
@@
 \cmidrule{3-4} \cmidrule{7-8} \cmidrule{11-12}
-Particle &\phantom{abc} & Raw & E$\times$B\\
+Particle &\phantom{abc} & Measured & E$\times$B\\
 \midrule
 p   & 882.8 & 929.5 & 877.3 & 903.5 & 929.5 & 889 &\phantom{abcdef} & f & f \\
-d   & 817.1 & d & d & 898.5 & e & e & 1621.1 & 1704 & 1612\\
+d   & 817.1 & 831.15 & d & 898.5 & e & e & 1621.1 & 1704 & 1612\\
 t   & 589.5 & d & d & 887 & e & e & 1612.4 & f & f  \\
@@
 
-Cocktail Calibration 
+Light charged particles (p,d,t,${}^{3}$He,${}^{4}$He), beams were produced and measured in the TPC.
+The magnetic and slit settings of the dipoles in the BIGRIPS spectrometer was set so that the measured momentum resolution of the beam was $\frac{\delta p}{p}$ < 1\%.
+Two magnetic rigidity settings were studied, with an empty target.
+A thick Aluminum target was used to provide a slightly lower point for part of the lower rigidity setting, effectively creating three calibration points over several particle species.
+The production of certain particles (t,${}^{3}$He), produced too few counts to make a good measurement, and in the high momentum rigidity setting protons could not propagate down the line and there were no counts.
+
+Since the expected momentum resolution resulting from the spectrometer was less than 1\%, the observed momentum resolution measured by the TPC is a good measurement of the combined momentum resolution of the software and TPC (intrinsic detection) system.
+The momentum resolution of the TPC depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
+This calibration beam represents and ideal situation where the track was parallel to the pad plane and only one particle was measured at one time.
+The energy loss resolution can also be directly inferred from the measurement since each energy setting represents a monochromatic source of each particle species, which has a well defined energy loss distribution.
+An average momentum resolution of ~2\% and the energy loss resolution of ~ 5\% was measured for particles ranging from protons to ${}^{4}$He, over the range of momenta measured in the calibration beam as summarized in Table~\ref{tb:momresolution}.
+
+Since magnetic dipole setting of the BIGRIPS spectrometer define the energy of each particle type we can calculate the expected momenta of each particle species measured.
+Small corrections to the momenta were propagated using LISE++ software which can calculate the energy loss through several materials in the beam line.
+These corrections resulted in a small change in the momenta.
+The measured momenta of the calibration beam differed significantly from the expected values as seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
+This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocity in the direction of $\vec{E}\times\vec{B}$ direction.
+The $\vec{E}\times\vec{B}$ dift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
+The disagreement in measured and expected momenta is upwards of ~5\% difference in the higher momentum calibration settings.
+The details of the correction technique are discussed in the later Section~\ref{sec:spacecharge} are discussed in a more general correction which also includes correction for the space charge; the same correction technique was applied here in the special case of zero space charge which is the special case of only having $\vec{E}\times\vec{B}$ components .
+
+The values under the corrected column of seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data correcting for $\vec{E}\times\vec{B}$.
+A significant improvement is seen in the high momentum setting going from around ~5\% disagreement to within ~1\% agreement in the corrected data.
+For the lower momentum settings (Tables~\cref{tb:cocktail100tar,tb:cocktail100}), protons see a slight improvement of about ~1\% where as the deuterons are over corrected in both settings.
+The level of agreement of the all corrected values is still within the estimated momentum resolution of the TPC.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule
+Momentum Resolution \% & <dE/dx> Resolution \% \\
+\midrule
+1.6  & 4.6\\
+\bottomrule
+\end{tabular}
+\caption{Summary of expected cocktail. }
+\label{tb:momresolution}
+\end{table*}
+
 Picture of cocktail before and after ExB effect
@@
 \subsection{Electronics calibration}
-Pulse the ground grid. 
+The channel number of the electronics was calibrated by measuring the response of each channel to an input pulse supplied by a pulse generator.
+The pulse was injected into the ground plane of the TPC.
+This distributed the pulse evenly across the entire pad plane over a range of input voltages.
+The input voltage is plotted as a function of the measured ADC channel in Fig.~\ref{fig:gaincalib} for every channel.
+The small variation in each channel can be seen as the wide band around each measurement point.
+A linear fit is performed to get the best fit line which provides a reference line which each channel is calibrated to.
+The right panel shows the resulting distribution of channels after calibration.
+This is a relative calibration technique meant to calibrate the varying gains in each channel relative to one another.
+
 \begin{figure}[H]
@@
 
+As mentioned earlier, the anode wires were separated into 14 independently biased sections.
+The high voltage of sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires.
+The wire section voltages were lowered once and adjusted once again.
+Out of all the runs used in the analysis in this thesis \ref{tb:runList}, the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all the other runs.
+By lowering the voltage on these anode wires, the gas gain is lowered as compared with all the other anode wire plane sections which operate at \SI{1460}{\volt}.
+To account for the drop in gain, in the software we increase the gain of the pads which lie above these anode wires.
+To calibrate these sections we perform a relative calibration to the high gain anode sections by comparing the energy loss values of the high and low gain sections.
+
 \begin{figure}[H]
@@
 
-Picture of low vs high gain channels and fit
+\subsection{Extending the dynamic range of the Electronics}
+Using a TPC for measurements of HIC in nuclear physics presents a different set of challenges as opposed to higher energy experiments.
+Typically in higher energy experiments the charge of particles is Z$e$, where Z=1.
+Also the particles are traveling at higher energies in which the energy losses in a detector range from the minimum of the energy loss curve to the small logarithmic climb CITE FIGURE HERE!!!.
+The dynamic range of electronics in such experiments can cover a wide range of particle energies in the energy loss curve.
+In nuclear HICs, we are interested in cluster of nucleons typically from Z=1-3 resulting from the collision, and even higher in some applications.
+As seen in Eq.~\ref{eq:bb}, the energy loss is proportional to $Z^2$; the energy values of the resulting particles are at much lower energies which are in the $1/\beta^2$ region.
+The energy range and particle types covered by the electronics are significantly limited as they quickly reach the dynamic range limitations of the electronics as the charge of a particle increases and the velocity decreases.
 
-\subsection{Extending the dynamic range of the Electronics}
-Add paper here.
+Several TPCs have tried to address this issue by having regions of low and high gain, either in amplification gain or in electronics gain.
+Here we run into the issue that high velocity particles will have little to no signal in the low gain regions.
+At lower velocities, particles will deposit much charge over the low gain regions but will saturate the high gain regions, whose charge values are usually lost.
+Only within the dynamic range will tracks have the best momentum resolution, outside the dynamic range clusters will be missing in the sections which depend on the track velocity, lowering the momentum resolution of tracking.
+There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these persistent issues in TPC electronics CITE HERE, by being able to switch to a lower gain value when the maximum range is reached.
+Though, it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of external hardware, espeically in experiments which have already been performed with older electronics technologies.
+
+In TPCs, the effective dynamic range can be very different from the single pad dynamic range.
+Typically TPCs are operated inside of a magnetic field for reconstructing momentum of each particle, which requires sub-millimeter precision in the position determination of the track path.
+This is achieved by averaging the charge distribution over several pads which is discussed in Section CITE HERE.
+Therefore the effective dynamic range is related to the relative charge values of adjacent pads (between the center pad and outside pads).
+For example, to measure minimum ionizing particles the signal to noise ratio of the pad with the smallest charge in the distribution should be some reasonable value, say at least 6:1.
+From Section CITE HERE we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
+If we require the adjacent pads to have a signal-to-noise ratio of 6:1, then the central pad would have a signal to noise ratio of 50:1.
+Considering this is the signal-to-noise ratio for minimum ionizing particles, and the maximum signal-to-noise ratio is 800:1, this means the effective dynamic range in the TPC is roughly 16 times that of minimum ionizing particles.
+The dashed lines and vertical blue bar in Fig.~ \ref{fig:intro} are separated by a factor of 20, representing the typical effective dynamic range in a TPC.
+This dynamic range estimate should be regarded as approximate because the energy loss fluctuates significantly about the most probable energy loss, with a long ``Landau'' like tail, as described by Bichsel \cite{bichsel}.
+Nevertheless, the blue dashed lines and vertical blue bar illustrate that the range of energy losses sampled in a fixed gain readout system is limited.
+One can change the gain and shift the energy loss range that can be sampled, but the dynamic range itself cannot be increased.
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{intrographic}
+%\includegraphics[width=\linewidth]{intrographic_2}
+\caption{The expected $dE/dx$ lines of different particles are given in red as calculated by Geant4.
+The approximate dynamic range of the TPC is shown by the vertical bar for the gain setting used in the experiment.
+Anything outside of this region would be saturated to some degree.}
+\label{fig:intro}
+\end{figure}
+
+The rapid increase in the stopping powers at low momentum illustrates the degree to which the effective dynamic range can be exceeded and highlights the problems encountered in studies of intermediate HICs, in which light particles with low momenta are abundantly produced along with highly charged particles.
+Similar problems are encountered when TPCs are used as active targets in direct reaction studies with rare isotope beams \cite{pattpc}.
+
+Several techniques have been employed to increase the observable range of energy losses.
+This can be done by lowering the electronics gain of selected readout channels, or by changing the gas amplification at the readout plane in certain areas of the TPC.
+In the EOS TPC \cite{eos} this was done by decreasing the voltages in select anode wires in the multi-wire readout.
+With the prototype Active Target TPC lowering or increasing the gain was achieved by decreasing or increasing the electric fields on selected pads within a Micromegas \cite{pattpc}.
+The results of changing the gas-gain, or the electronics gain, are rather similar in that reducing the gain to sample a range of higher energy loss makes the TPC effectively blind to minimum ionizing particles in  the regions of lower gain.
+
+\begin{figure}[ht!]
+\includegraphics[scale=.5]{top_view_helix_ext.pdf}
+\caption{Cartoon of a top down view of a fit to a track passing through several pads.
+The bolded pads and the charges $q_i$ represent the hits belonging to that pad and the clusters of the track representing the average position of the track.
+The three clusters at the bottom are clustered in the $x$-direction for the upper three are clustered in the $z$-direction.
+The estimated position of the avalanche is given by the track fit, and the position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
+\label{fig:topview}
+\end{figure}
+
+We define the clustering direction depending on the angle of the track, at the point of each cluster, with respects to the $x$ axis, defined as $\theta$.
+For example, the crossing angle is defined as $90^{\circ}$ for a track going along the $z$ axis, and $0^{\circ}$ for a track going along the $x$ axis.
+In the case that the crossing angle is $45^{\circ} < \theta \leq 90^{\circ} $ the clustering direction is along the $x$ axis.
+For $0^{\circ} < \theta \leq 45^{\circ}$ it is along the $z$ axis.
+
+ The position along the clustering direction is calculated by weighting the individual hit's positions by their charges $q_i$ and getting the mean value.
+The other direction is set to the center of the pad.
+For example, if we are clustering along the $x$ axis for a cluster, the $z$-position is set to the center of the pad in the $z$-direction and vice versa.
+
+\subsection{Method of Desaturation}
+We will use the term ``desaturation'' for our process of correcting the charge values of the saturated pads.
+Figure \ref{fig:satpad} shows a typical situation of saturated signals.
+When an avalanche causes a large induced signal, the pads directly underneath collect the largest charge becoming saturated, denoted as $q_{2'}$ and $q_{3'}$.
+Pads further away experience smaller, non saturating charges, denoted as $q_{1}$ and $q_{4}$.
+Though we do not know the saturated charge values, the distribution of all charges must follow the PRF which we have experimentally measured.
+From the clusters crossing angle, we can get the corresponding parameters for the PRF as described above and in Fig.~\ref{fig:normsigma}.
+
+We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as given by the $PRF(\lambda_i)$ function.
+
+\begin{equation}\label{eq:chi}
+\chi^2 = \sum_i \frac{(q_i^{\mathrm{obs}} - q_i^{\mathrm{expect}})^2}{q_i^{\mathrm{expect}}}
+\end{equation}
+
+To determine the best estimate for the charge values of each saturated pad, a chi squared function is minimized, given in  Equation \ref{eq:chi}, where $q_i^{\mathrm{obs}}$ are the observed, non-saturated charges $q_{1}$ and $q_{4}$, and $q_i^{\mathrm{expect}}$ are the charges we expect to observe, calculated as $q_i^{\mathrm{expected}} = Q\cdot PRF(\lambda_i)$.
+The charges $q_{2'}$ and $q_{3'}$, make up the unknown variable and are allowed to vary in the $\chi^2$ minimization, where they are added to make up the total expected charge $Q$.
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{saturated_pads}
+\caption{A typical case of a saturating event.
+The red pulses represent the time bucket signal for each collected charge.
+The pads directly underneath the avalanche point, $q_{2'}$ and $q_{3'}$, are saturated while pads farther away, $q_1$ and $q_4$ are not saturated.}
+\label{fig:satpad}
+\end{figure}
+
+Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{cocktail.png}
+\caption{Pad plane projection for a cocktail event in the TPC.}
+\label{fig:cocktail}
+\end{figure}
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{data.png}
+\caption{Pad plane projection for a collision event in the TPC.
+Highlighted by red arrows are two regions of anode wires which had a reduced voltage of 1214 V.
+The voltage of the rest of the TPC anode wires are 1460 V.
+The reduction in voltage reduces the gain by a factor of about 10.
+}
+\label{fig:data}
+\end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\linewidth]{dedx_compare}
+\caption{The left panel shows the high gain stopping power vs low gain when the method of desaturation was not applied.
+In the right panel the desaturation technique was applied to the high gain region.
+The low gain does not suffer from saturation and represents the true $dE/dx$ value.}
+\label{fig:lowvshigh}
+\end{figure*}
+
+\begin{figure*}[t]
+\includegraphics[width=\linewidth]{cocktail_combine.png}
+\caption{Uncorrected (left panel) and desaturated (right panel) cocktail data.}
+\label{fig:cocktail_combine}
+\end{figure*}
+
+\begin{figure*}[t]
+\includegraphics[width=\linewidth]{data_combine}
+\caption{Uncorrected (left panel) and desaturated (right panel) collision data at polar angles of $\theta < 40^{\circ}$ and azimuthal angles between $-80^{\circ} < \phi < 80^{\circ}$}
+\label{fig:data_combine}
+\end{figure*}
+
+\subsection{Low gain vs corrected high gain}
+
+Tracks which saturate pads in the high gain region are not saturated in the low gain region.
+By comparing the dE/dx values of these two sections, we can directly measure the success of the desaturation in the high gain regions using the method described above.
+
+In Fig.~\ref{fig:lowvshigh}, the effect of saturation can be seen in the high gain region for the uncorrected data.
+For signals below 400 ADC/mm \footnote{Un-calibrated ADC channels in arbitrary units.} the electronics are not saturated, and therefore the high and low gain sections agree.
+The data starts to saturate above 400 ADC/mm in the high gain channels eventually reaching a plateau while the low gain sections are not saturated and provide the true $dE/dx$ values.
+ After applying the desaturation method, the correlation between the high and low gain sections is restored, as seen in Fig.~\ref{fig:lowvshigh}.
+From this comparison, we infer that the correction works up to signals of 2000 ADC/mm, increasing the dynamic range by a factor of at least 5.
+
+\subsection{Particle Identification (PID)}
+
+Comparing the low to high gain sections directly validates the desaturation technique, but the goal  of this exercise is to improve the particle identification (PID).
+In the following PID plots the red lines represent the most probable energy loss as given by Geant4 straggling functions.
+A linear calibration was performed to convert keV in Geant4 to ADC in the experiment given by $ADC/mm = 19\;keV/cm$.
+
+There are pronounced PID lines of several particle species in both the uncorrected and corrected cocktail beam PID shown in the subplots of Fig.~\ref{fig:cocktail_combine}.
+Three ovals around a momentum of 1700~MeV/$c$/q and two near 900~MeV/$c$/q correspond to the three $B\rho$ settings injected into the TPC.
+The tails of the PID lines are resulting from the particles passing through the walls and other materials outside the main detector volume, therefore lowering the initial momentum.
+
+The uncorrected data in Fig.~\ref{fig:cocktail_combine} shows the effects of saturation; the PID lines deviate from their theoretical expectations starting at around 400~ADC/mm eventually reaching a plateau.
+After applying the desaturation technique, we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation.
+A more subtle improvement of the lighter particles, (p, d, t), can also be seen in the PID lines at lower momenta.
+
+Looking at the collision data, shown in Fig.~\ref{fig:data_combine}, we also see a similar result.
+In the collision data, the PID suffers from more background and inefficiencies than the cocktail beam, nevertheless we can see a similar improvement in the PID lines when comparing before and after applying desaturation.
+Notably the largest improvement is the separation of particle species at lower momenta and the separation of the Li species into ${}^{6}$Li and ${}^{7}$Li.
+In these regions, there was little to no PID resolution before desaturation.
+
+The dynamic range was extended by at least a factor of 5, as demonstrated by the improved PID lines, and quantified by direct comparison to low gain sections of the TPC.
+This improved PID will allow for us to extend the momentum distributions of all species to lower momenta and to heavier ions than what was previously available.
 
 \subsection{Space Charge Corrections}
+\label{sec:spacecharge}
 

### 2020-04-09 chapters/ch_experiment.tex
Added some preliminary delta res
Added space charge sections to thesis. match clusters data retrieved
Routine Backup

@@
 \chapter{Experiment}
+
+\section{Energy loss in material}
+The average energy loss in a material can be described by the Bethe-Bloch equation,
+\begin{equation}\label{eq:bb}
+\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2).
+\end{equation}
+Where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle.
+There is a large variation in energy loss around this mean value, with a long high energy loss tail.
+A full description of the energy loss distribution can be described in CITE HERE.
+
+\section{Pad Response Function}
+The charge distribution on the pad plane resulting from an electron avalanche is two dimensional, as illustrated in Fig.~\ref{fig:prf}, with the projections onto the $x$ and $z$ axis of this distribution labeled as $\rho(x)$ and $\rho(z)$ respectively.
+The charge observed on each pad is the integral of this distribution over the pad's dimensions, which is called the Pad Response Function (PRF).
+
+In this example the clustering direction would be along the $x$ axis, as illustrated by the bolded pads.
+By choosing to cluster only in one direction, we have not included the charge in adjacent pads along the $z$ axis resulting from the tails of $\rho(z)$.
+Therefore the charge not included in the the bolded pads will be incorporated into adjacent clusters, introducing small correlations in charge between neighboring clusters.
+
+\subsection{Gatti Pad Response Function}
+Gatti \cite{gatti} derived a semi-empirical formula for the charge distribution in a simple multi-wire TPC.
+The function given in Eq.~\ref{eq:gatti}, depends only on the width of a pad ($w$), the distance of the anode plane to the pad plane ($h$), and the distance of the pad center to the avalanche point $\lambda$.
+It is a single parameter equation where the two parameters $K_1 = \frac{K_{2}\sqrt{K_3}}{4 \arctan(\sqrt{K_3})}$ and $K_2 = \frac{\pi}{2}\left(1-\frac{\sqrt{K_{3}}}{2}\right)$ depend on the parameter $K_3$, which is a function of the ratio of the anode wire diameter to the distance of the anode wires to the pad plane.
+$K_3$ can be looked up in a graph in \cite{blumrol} and \cite{gatti}.
+
+\begin{equation}\label{eq:gatti}
+\begin{split}
+PRF_{\mathrm{Gatti}}(\lambda)
+& = \frac{K_{1}}{K_{2}\sqrt{K_{3}}}\bigl[\arctan(\sqrt{K_{3}}\tanh\bigl[K_{2}\bigl(\frac{\lambda}{h}+\frac{w}{2h}\bigr)\bigr]) \\
+& - \arctan(\sqrt{K_{3}}\tanh\bigl[K_{2}\bigl(\frac{\lambda}{h}-\frac{w}{2h}\bigr)\bigr])\bigr] \\
+\end{split}
+\end{equation}
+
+\subsection{Experimental Pad Response Function}
+
+The correlations we introduced by only clustering along one direction do not play a significant role in the particle identification, but cause deviations from the expected Gatti distribution.
+Also, analytic PRFs only exist for classical multi-wire TPCs.
+For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
+
+As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as $\lambda_i = x_i-\bar{x}$.
+The PRF is simply defined as the charge fraction of each pad as a function of $\lambda$, as shown in Equation \ref{eq:prf}.
+
+\begin{equation}\label{eq:prf}
+PRF(\lambda_i) = \frac{q_i(\lambda_i)}{Q}
+\end{equation}
+
+Averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
+Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
+
+\begin{equation}\label{eq:gaus}
+PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
+\end{equation}
+
+\begin{figure}[ht!]
+\begin{overpic}[width=\linewidth]{expprf}
+\put(61,55){\contour{white}{ PRF${}_{\mathrm{Gaus}}(\lambda)$ eq. \ref{eq:gaus}  }}
+\put(61,49){\contour{white}{ PRF${}_{\mathrm{Gatti}}(\lambda)$ eq. \ref{eq:gatti} }}
+\end{overpic}
+\caption{Experimental pad response function of many events for a crossing angle of $85^{\circ} < \theta \leq 90^{\circ}$.  }
+\label{fig:expprf}
+\end{figure}
+
+\section{Pad Response Function vs. crossing angle}
+The shape of the PRF depends on the crossing angle of the track \cite{gatti}.
+Plotted in Fig.~\ref{fig:prfangle} is the PRF of $\pi^-$ tracks vs.
+the crossing angle $\theta$.
+The PRF gets wider starting from $90^{\circ}$  and going to $45^{\circ}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
+Since we switch the clustering direction from $x$ to the $z$ direction at $45^{\circ}$, the opposite trend is seen where the PRF becomes narrower as the position resolution gets better going from $45^{\circ}$ to $0^{\circ}$.
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{PRF_vsAngle}
+\caption{Experimental pad response function as a function of the crossing angle $\theta$ with the corresponding fits.}
+\label{fig:prfangle}
+\end{figure}
+
+\begin{figure}[ht!]
+\begin{overpic}[width=\linewidth]{normsigma}
+\end{overpic}
+\caption{Parameters $N_{0}$ and $\sigma$ as a function of the crossing angle $\theta$ with the $4^{th}$ order polynomial fits.}
+\label{fig:normsigma}
+\end{figure}
+
+\begin{comment}
+\begin{table}
+\centering
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Coefficient & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ [0.5ex] 
+ \hline\hline
+ $0 < \theta < 45$ & & & & &  \\ [.25ex]
+ \hline
+ $N_0$ & .897 & 5.766E-3 & -4.263E-4 & 7.444E-6 & 5.705E-8 \\ 
+ \hline
+ $\sigma$ & 5.496 & -3.920E-2 & 2.693E-3 & -5.208E-5 & 5.334E-7\\
+ \hline
+ $45 < \theta < 90$ & & & &  & \\ [.25ex]
+ \hline	
+ $N_0$ & 1.220 & -6.258E-2 & 1.608E-3 & -1.492E-5  & 4.654E-8 \\
+ \hline
+ $\sigma$ & 31.368 & -1.109 & 1.779E-2 & -1.336E-4 & 3.940E-7\\
+ \hline
+\end{tabular}
+\caption{Coefficients of the $4_th$ order polynomial fit to the Gaussian parameters $N_0$ and $\sigma$. The polynomial form is given as $c_0 + c_1 x + c_2 x^2 + c_3 x^3 + c_4 x^4$}
+\label{tb:coeff}
+\end{table}
+\end{comment}
+
+Fits were performed to the experimental data with  $5^{\circ}$ width bins from $0^{\circ} < \theta \leq 90^{\circ}$.
+The two parameters of the Gaussian fits are plotted versus $\theta$ in Fig.~\ref{fig:normsigma}; a $4^{th}$ order polynomial fit between these points allowed for interpolating between $\theta$.
 
@@
 
+\begin{equation}
+\frac{d\vec{x}}{dt} = \frac{\mu}{1+(\omega\tau)^2}\Big(\vec{E} + \omega\tau\frac{\vec{E}\times\vec{B}}{|\vec{B}|}+\omega^2\tau^2\frac{\vec{E}\cdot\vec{B}}{|\vec{B}|^2}\vec{B}\Big)
+\label{eq:elecdrift}
+\end{equation}
+
 \begin{figure}[H]
@@
   S[table-format=1.2]
+  S[table-format=1.2]
   S[table-format=5.2]
@@
  {$v_{d}$} &
+ {$v_{i}$}  &
  {$G_{h}$} &
@@
   {($\si{\centi\meter}^{-1/2}$)} &
+  {(\si{\centi\meter\per\micro\second})} &
  {(\si{\centi\meter\per\micro\second})} \\
+
 \midrule
-\phantom{abc}   &.024   &.034  &5.43   &  903   &150     \\
+\phantom{abc}   &.024   &.034  &5.43  &  \num{2.05e-4} &  903   &150     \\
 \bottomrule

### 2020-04-09 macrosForThesis/CocktailBeam/plotCocktailSummary.C

@@ -1,0 +1,19 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+void plotCocktailSummary()
+{
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+}

### 2020-04-09 macrosForThesis/PionSpectra/PlotDataTheory/DanCozma/April03/publishPion_132Ratio.C

@@ -168,24 +168,28 @@ void publishPion_132Ratio()
 
   //  TF1 *fit = new TF1("fit","[0] + [1]*x + [2]*x*x + [3]*pow(x,3) + [4]*pow(x,4) + [5]*pow(x,5) + [6]*pow(x,6)",0,300);
   //  fit->SetParameters(1,1,1,1,1,1,1);
-  TF1 *fit = new TF1("fit","[0]*pow(x,[1])",0,300);
-  fit->SetParameters(1,1);
+  TF1 *fit = new TF1("fit","[0]*TMath::Exp([1]*x + [2]*pow(x,2.5))",0,300);
+  fit->SetParameters(12,-1e-2,-1e-6);
 
   singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
   singleRatio_sn132->GetXaxis()->SetRangeUser(0,300);
   singleRatio_sn132->Draw();
+
+  sim4_g->Draw("same E3");
   /*
-  sim1_g->Draw("same E3");
-  sim2_g->Draw("same E3");
+  sim5_g->Fit(fit);
+  sim5_g->Draw("same E3");
+
   sim3_g->Draw("same E3");
   sim4_g->Draw("same E3");
   sim5_g->Draw("same E3");
   sim6_g->Draw("same E3");
 
   sim7_g->Draw("same E3");
-  sim8_g->Draw("same E3");*/
-  
+  */
+  sim8_g->Draw("same E3");
   sim9_g->Draw("same E3");
+
 
   singleRatio_sn132->DrawCopy("same E1");
 

### 2020-04-25 chapters/ch_data_analysis_2.tex
Routine Backup
Regular update

@@
-\chapter{Data Analysis II: Extracting Physics}
+\chapter{Data Analysis II}
+
+\section{Beam Particle Identification}
+Figures of beam contaminants in our beam PID line. 
+Table of beam purity reference to Jon's thesis paper
 
@@
 
-The number of clusters affected is small but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
+The number of affected clusters is small, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
 Simple cuts around the left,right, top, and bottom of the TPC remove these clusters from the analysis,
@@
 The beam is then propagated through the magnetic field until the target position.
-The beam angle on target is categorized by two angles $\theta_{a,proj}$ and $\theta_
+The beam angle on target is categorized by two angles $\theta_{a_{proj}}$ and $\theta$
 
@@
 \section{Vertex Cut}
+A vertex cut on verticies originating from the target region ensure we are selecting on nuclear collisions with the target.
+When looking at the z-component of the verticies of all events, we can see several features which correspond to different elements within the beam line and TPC.
+We also note that near the expected position of the target, WHAT POSITION??, we see a large enhancement of counts which corresponds to nuclear collisions on the target.
+The z-component of all runs in each beam type are plotted in the range of the target region in Fig.~\ref{fig:vertexdist}.
+Since the target thickness was less than 1 mm, the vertex resolution achieved can be experimentally measured by fitting the vertex distribution, assuming a Gaussian distribution.
+The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrr@{}}\toprule
+\multicolumn{3}{c}{Vertex Resolution}\\
+\cmidrule{1-3}
+System & Mean (cm) & Sigma (cm) \\
+\midrule
+$\tin{132}{124}$ & -14.79  & 1.2 \\
+$\tin{124}{112}$ & -14.71  & 1.1 \\
+$\tin{112}{124}$ & -14.78  & 1.2 \\
+$\tin{108}{112}$ & -14.75  & 1.3 \\ 
+\bottomrule
+\end{tabular}
+\caption{Summary of measured verticies and their resolution.}
+\label{tb:vertexresol}
+\end{table*}
 
@@
 \section{Impact Parameter Selection}
+The impact parameter which is defined as the distance (center-to-center), between two nuclei in a collision is a pure theoretical concept.
+Though experimentally there is no way in experiment to measure the impact parameter directly, several quantities have been indirectly related to the centrality of the collision CITE HERE.
+As discussed in Section CITE HERE, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events.
+This is under the reasonable assumption that the number of charged particles in a collision grows as the impact parameter decreases.
+This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons.
+In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters (peripheral collisions), less nucleons participate.
+
+Certainly the Kyoto array introduces a bias in our measurement of the multiplicity of all the events.
+At low track multiplicities, the orientation of the event starts to play a major role.
+Theoretically the orientation of the event (the reaction plane) is random, but due to the fact that we only measure the multiplicity on the sides of the TPC, we are preferentially triggering on events with reaction planes that emit particles in these directions.
+Therefore events with low track multiplicity (therefore more peripheral), that emit preferentially up and down in the TPC will not be triggered on.
+This becomes less of an issue as even in mid-peripheral collisions were the number of particles emitted becomes much greater than the Kyoto multiplicity selection.
+
+Assuming a geometric cross section $\sigma$ and the impact parameter of the collision $b$,
+
+\begin{equation}
+\sigma = \pi \cdot b^2,
+\label{eq:crossSect}
+\end{equation}
+
+we can estimate the maximum impact parameter $b_{\mathrm{max}}$, by measuring the total cross section of each system.
+A detailed analysis on determining the cross section is given in CITE HERE.
+I have summarized the finding of the cross section and $b_{\mathrm{max}}$ for all the systems.
+Assuming $b_{\mathrm{max}}$ represents the most peripheral collisions we can define an estimate for the impact parameter $\hat{b}$, which we can define from the cumulative integral of the measured multiplicity distribution,
+
+\begin{equation}
+\hat{b} = b_{\mathrm{max}} \frac{\int_{\infty}^{M_x} f(M) dM}{\int_{\infty}^{0} f(M) dM}.
+\end{equation}
 
 \section{Track Quality Cuts}
+%graphic showing distance to vertex, number of clusters
+%number of cluster cut
+%distance to vertex
+In this section we will discuss the quality of the reconstructed track and will simply refer to it as the ``track quality".
+Certainly tracks with more clusters are of higher quality than those with few clusters.
+Occasionally the software will associate clusters and construct a track whose vertex does not coincide with the event vertex.
+This can occur for a variety of reasons.
+For a low momentum pion the radius of curvature is such that several rotations can occur in the event which the software can interpret as several tracks; while only one of the reconstructed track appears to originate from the event vertex.
+It could also happen that false tracks are created by the software from clusters which originate from several different real tracks.
+These are fairly obvious as they usually traverse the detector horizontally and the number of clusters is very low.
+In this situation the false track does not originate from anywhere near the vertex.
+
+To reduce the number of fake tracks, or multiple counting of tracks, a simple cut on the measurement of a track's distance to vertex can be used.
+The vector representing the minimum distance to vertex for each track can be calculated and its corresponding magnitude.
+We define the distance to vertex cut as a small sphere centered on the vertex location with a radius R.
+If the distance to vertex of a particular track $d_i$ < R the track is assumed to have originated from the vertex; tracks that are outside are thrown out of the analysis.
+
+Assuming tracks are mostly continuous in clusters, i.e. only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track.
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{distancetovertex.pdf}
+\label{fig:poca}
+\caption{Distance to vertex for tracks.}
+\end{figure}
 

### 2020-04-28 chapters/ch_data_analysis.tex
Regular update
Update saturation section

@@
 \label{fig:topview}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{padsat_Large}
-\caption{A cartoon illustration of the charge distribution resulting from an electron avalanche on one wire and the projections of the distribution onto the two axis $\rho(x)$ onto the x-axis and $\rho(z)$ onto the z-axis.
-The orientation of the wire planes is flipped upside down to display the perspective better.}
-\label{fig:prf}
 \end{figure}
@@
 
-\begin{figure}[ht!]
-\includegraphics[scale=.5]{top_view_helix_ext.pdf}
-\caption{Cartoon of a top down view of a fit to a track passing through several pads.
+%DOUBLY DEFINED? IS THIS SUPPOSED TO BE HERE
+%\begin{figure}[ht!]
+%\includegraphics[scale=.5]{top_view_helix_ext.pdf}
+%\caption{Cartoon of a top down view of a fit to a track passing through several pads.
 The bolded pads and the charges $q_i$ represent the hits belonging to that pad and the clusters of the track representing the average position of the track.
@@
 The estimated position of the avalanche is given by the track fit, and the position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
-\label{fig:topview}
-\end{figure}
+%\label{fig:topview}
+%\end{figure}
 
@@
 
-\subsection{Low gain vs corrected high gain}
-
 Tracks which saturate pads in the high gain region are not saturated in the low gain region.
@@
 From this comparison, we infer that the correction works up to signals of 2000 ADC/mm, increasing the dynamic range by a factor of at least 5.
-
-\subsection{Particle Identification (PID)}
 
@@
 The inset figure of Fig.~\ref{fig:sc_shift} shows the shift in the x-position distance to vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
-We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR}$.
+We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR}$, see Fig.~\ref{fig:VLR}.mom
 
@@
 \caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks in the TPC for the }
-\label{fig:spacechg_cartoon}
+\label{fig:VLR}
 \end{figure}
@@
 \subsection{Pad Response Task}
-After the avalanche process of the previous task, the total charge of the event is split over several pads defined by the Pad Response Function (as discussed in section ????).
-As shown in Fig.~\ref{fig:onepad}, there are three wires that lie directly above a given pad.
-The $z$ coordinate of the avalanche can only be one of these three wires, where as the $x$ coordinate can be any value along the wire.
-The functional form of the software PRF (given in Eq.~\ref{eq:softPRF}), was tuned to match the experimental data.
-Shown in Fig.~\ref{fig:mcdataPRF}, we see the tuned software PRF can match the experimental PRF from data over several crossing angles (as mentioned in Chapter ????
-governs the shape of the PRF).
+The total charge of the event is split over several pads defined by thePRF (as discussed in section \ref{sec:prf}.
+As shown in Fig.~\ref{fig:2DPRF}, there are three wires that lie directly above a given pad.
+The $z$ coordinate of the avalanche is limited to one of these three wires, where as the $x$ coordinate can be any value along the wire.
+A 2-dimensional Gaussian was enough to describe the experimental PRF,
 
 \begin{equation}
-PRF(x,z) = \frac{1}{2\pi\sigma_z\sigma_x}\exp^{\frac{-{(x-x_o)}^2}{2\sigma_x^2}}\exp^{\frac{-{(z-z_o)}^2}{2\sigma_z^2}}
-\label{eq:softPRF}
+PRF(x,z) = \frac{1}{2\pi\sigma_z\sigma_x}\exp^{\frac{-{(x-x_o)}^2}{2\sigma_x^2}}\exp^{\frac{-{(z-z_o)}^2}{2\sigma_z^2}},
+\label{eq:2DPRF}
 \end{equation}
 
-\begin{figure}
-\includegraphics[scale=.005]{place}
-\caption{A cartoon of the wires over one pad. }
-\label{fig:onepad}
-\end{figure}
+where the location of the avalanche is ($x_o$,$z_o$).
+The $\sigma$ values were tuned to match the experimental data.
+Figure~\ref{mcdataPRF} shows the PRF response of $\pi^-$ tracks.
+The black line is the PRF fit to the experimental data.
+We can see a good agreement between the MC, Fig.~\ref{fig:prfpimMC}, and experimental data, Fig.~\ref{fig:prfpimData}.
 
-\begin{figure}
-\includegraphics[scale=.005]{place}
-\caption{Comparison of MC and data PRF}
-\label{fig:mcdataPRF}
+\begin{figure}[!htb]
+         \centering
+         \includegraphics[width=\textwidth]{PRFs_MC_wcut.png}
+         \caption{PRF response of Monte Carlo simulation of $\pi^-$}
+         \label{fig:prfpimMC}
 \end{figure}
@@
 \subsection{Electronics Task}
-The electronics task takes the total charge on each pad and simulates the electronics response, converting electronics into ADC channels.
-Accounting for the pedestal, the measured output in ADC channels (for a given gain setting) is given in Eq.~\ref{eq:etoADC}.
+The electronics task simulates the electronics response to induced charge on a pad, converting electronics into ADC channels.
+The ADC value of a channel, R, is modeled as,
+
+\begin{equation}
+\mathrm{R}= \frac{ADC_{\mathrm{Max}} - ADC_{\mathrm{Pedestal}}}{G*f_c},
+\label{eq:etoADC}
+\end{equation}
+
+where $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), $\mathrm{ADC_{Pedestal}}$ is the baseline or pedestal value used in the experiment (???), $G$ is the gas gain, and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
+Later a section is dedicated to small particle dependent corrections which account for small changes in the charge calibration which may be due to a whole host of detector effects that are not modeled.
+
+The induced current in the pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape read out.
+The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type; for a given shaping constant.
+This allows us to assume the pulse shape is constant with the only two variables being the height of the pulse, Q, and the starting time bucket of the pulse, $t_o$.
+The shaping constant was set to \SI{117}{\nano\second} for most of the data.
+Figure~\ref{fig:pulseshape} shows the pulse shape for a normal pulse that is within the dynamic range as the dashed blue curve, the pulse is normalized so the maximum height is 1.
 
@@
 
+\subsection{Simulating Saturation}
+There are several types of saturation that must be simulated or accounted for to correctly model the TPC response.
+They all are varying degrees of the same effect which manifest in different ways in the detector.
+In general all amplifiers have a finite range of output values set by a positive and negative rail (typically -12V and +12V).
+If the input charge into a charge sensitive pre-amplifier causes the output voltage to reach the max (or min) output voltage, the electronics is considered saturated and the response may be non-linear.
+The picture is complicated by the fact that there is usually an RC feedback loop which dissipates the input charge.
+The time in which a pre-amplifier returns to its typical linear behavior depends on the input charge and how quickly it can be dissipated \cite{akiGET}.
+The pad is otherwise considered dead and no further charge can be measured until the electronics recover.
+
+Due to the long high energy tail in energy loss distributions of a particle traversing matter, it is common to see pads along the track which have very high energy loss values and saturate the electronics of that pad, even for minimum ionizing tracks.
+Saturation in this case is infrequent and not an issue as many other clusters are not saturated.
+As the particle's charge gets higher, or the momentum gets lower, the mean energy loss value gets much higher and a significant fraction of pads in a track may be saturated.
+While earlier we outlined and algorithm to correct the saturation for the track measurement itself, saturation also affects surrounding tracks.
+
+As mentioned earlier the pad is dead for some time, which depends on the total input charge.
+For charges that are relatively at or above the threshold for saturation the pad is certainly dead for the remainder of time of the measurement; i.e.
+$T_{dead}$ > \SI{10}{\micro\second}.
+This will cause signals from other tracks passing underneath the offending track to not be measured; the saturated pad ``shadows" any future signal.
+In the case where two tracks are separated in time due to different polar angles, but have significant overlap in a pad-plane projection, a significant portion of the later track may be shadowed and the track itself will suffer in quality significantly.
+For events with high track multiplicity this causes a significant shadowing effects, in that the upward going tracks recorded at earlier times shadow the downward going tracks recorded at later times.
+This will change event by event and is very difficult to simulate except through a MC track embedding approach which will be discussed later.
+
+In the case where the input signal is very large, the electronics can be dead for up to \SI{35}{\milli\second} \cite{akiGET}.
+The beam rate in the experiment was about \SI{10}{\kilo\hertz}, which corresponds to an average of \SI{10}{\micro\second} between subsequent beams.
+Pads can be effectively dead for several events before recovering.
+Due to the large charge of the beam, a large amount of very high energy electrons are produced via scattering from the beam passing through the gas.
+In the presence of the magnetic field the radius of curvature of most of the electrons are within several \si{cm}.
+Figure~\ref{fig:deltaE} shows the horizontal extent of the electrons in a top down view of the TPC.
+While some of the electrons can stop in the gas, a significant fraction can travel to the top and bottom of the TPC.
+These electron can pass through the gating grid without being blocked and either terminate on the pad or possibly deposit their charge directly in the pad.
+The charge induced on the pad was large enough to kill the pad for a time long enough to last until at least the next event.
+These pads are dead for the whole event and are randomly distributed around the beam path projection on the pad plane.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{deltaE_topview} 
+        \caption{Top view} \label{fig:deltaE_topview}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{deltaE_sideview} 
+        \caption{Side view} \label{fig:deltaE_sideview}
+    \end{subfigure}
+    \caption{Geant4 simulation of ${}^{108}$Sn beam at 270 MeV/A in P10 gas. Notice the extent of the delta electrons in the vertical direction as compared to the horizontal extent. }
+\label{fig:deltaE}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{padplane_sat.png}
+\caption{Tagging the saturated pads.}
+\label{fig:satTag}
+\end{figure}
+
+Simulation of saturation is handled naturally by embedding MC tracks into real data, by correctly identifying when in the time bucket spectrum saturation has occurred in each pad.
+Once the time of saturation has been identified, no further signals are embedded as the pad is assumed to be dead.
+In the case a pad is dead for the whole event the time of saturation is set to the first time bucket and no signal is embedded.
+The characteristic signature of a saturated signal is the fast fall time of the falling edge of the pulse going quickly to zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
+The long exponential tail can be effectively removed by a simple software technique which is similar to the electronics concept of ``pole-zero compensation".
+If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,
+
 \begin{equation}
-\mathrm{1 ADC }= \frac{ADC_{\mathrm{Max}} - ADC_{\mathrm{Pedestal}}}{G*f_c}
-\label{eq:etoADC}
+f_i^{'} = \frac{-b_1\cdot f_{i-1}^{'} + a_o\cdot f_i + a_1 \cdot f_{i-1}}{b_o}
+\label{eq:satpolez}
 \end{equation}
 
-\subsection{Simulating Saturation}
-Add figure showing saturated time bucket spectra with location of saturation identified and with pulse shape from embedding I would like to add and how it blocks it.
-Add Figure with 2D pad plane response with and without saturation flag
+where $a_o$ = .9723, $a_1$=-.9453, $b_o$=.9545, and $b_1$=-.9203.
+These numbers are experimentally fit to give the fastest, non-zero, correction to the tail of a normal pulse.
+This correction produces a negative undershoot for the saturated pulse, since there was no long exponential tail to begin with.
+Figure~\ref{fig:pulseshape} shows before the correction technique as the red curves, and after the technique as the blue curves for both the normal and saturated pulse.
+As the program is looking for and identifying positive peaks in the spectrum, we also calculate the corrected ADC value of that particular time bucket according to Eq.~\ref{eq:satpolez}.
+A peak is identified as a saturated signal if it is less than -20$\cdot G$ for more than 8 Tbs, where $G$ is the gain calibration of the pad, and the max ADC value is > $G\cdot$500 to eliminate false negative pulses which come from the gating grid subtraction in dead pads.
+The saturated flag is set to true and the data time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket of the negative peak.
+The MC time bucket position is set to $t_{peak} - 30$ to ensure that the MC signals do not overlap the saturating data signal.
+
+Saturation can also occur if the sum of all the pulse heights are over the max ADC threshold of 3500 ADC.
+This is because the fall time of the pre-amplifier circuit is much longer than the time bucket window and pulses are allowed to pile up.
+While pads that saturate via this method still are identified through the same algorithm described above, nothing is done in the MC class to take this into account when embedding signals.
+We assume this method of saturation is a higher order correction and the current algorithm approximates most of the saturation effects.
+Figure~\ref{fig:satTag} shows the tagging algorithm for an event in the TPC.
+The max ADC values in a given pad are colored and the black dots are pads tagged as saturated.
+Some pads appear to be below 3500 ADC, but upon further inspection they satisfy the sum of all the signals causes saturation though the max ADC never goes over 3500.
+
+Dead pads are identified earlier in the software in the STCore.cc class.
+A dead pad is classified as having a r.m.s.
+value  < 50 ADC and a max ADC < 50.
+The saturated flag is set to true, and the time bucket position of saturation in both the data and MC are set to the first time bucket.
+
+%maybe simple pream fig
+%maybe example from aki paper about times
+%maybe
+
+%Add figure showing saturated time bucket spectra with location of saturation identified and with pulse shape from embedding I would like to add and how it blocks it.
+%Add Figure with 2D pad plane response with and without saturation flag
 
@@
 The rectangular design complicates the calculation of the geometric acceptance, or the efficiency.
-
-\section{Beam Particle Identification}
-Figures of beam contaminants in our beam PID line. 
-Table of beam purity reference to Jon's thesis paper
-
-\section{Solid angle coverage}

### 2020-04-28 chapters/ch_data_analysis_2.tex
Update saturation section

@@
 Table of beam purity reference to Jon's thesis paper
+
+Beam tracking was handled by the BigRIPS spectrometer tracking system.
+Each individual beam particles coming down the spectrometer was measured using a dE-Time-of-Flight method where the charge, Z, and mass to charge ratio, A/Q, of each particle was measured.
+Individual particle species represent a two-dimensional Gaussian in this space; both the $\tin{132}{112}$ and $\tin{108}{124}$ systems are shown in Fig.~\ref{fig:beampid}.
+Clear separation is seen for each particle species and the correct value of Z and A/Q gives us the correct species type.
+Each species of interest is fitted with multivariate normal distribution with two variables defined as,
+
+\begin{equation}
+  f(x,y)=\frac1{2\pi\sigma_x\sigma_y\sqrt{1-\rho^2}}\exp\left\{
+  \frac{-(x - \mu_{x})^2/\sigma_x^2-(y-\mu_{y})^2/\sigma_y^2+2\rho
+  xy/\sigma_x\sigma_y}{2(1-\rho^2)}\right\},
+   \label{multiGauss}
+\end{equation}
+where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths.
+
+The resulting fit parameters are summarized in Table~\ref{beamParameters}. 
+\begin{table}[H]
+  \begin{center}
+    \begin{tabular}{cccccc}
+      \hline 
+      Particle Type & $\mu_\mathrm{A/Q}$ & $\sigma_\mathrm{A/Q}$ & $\mu_\mathrm{Z}$ &
+      $\sigma_\mathrm{Z}$ & $\rho$\\
+      \hline\hline 
+      ${}^{132}$Sn & 2.64 & 0.0014 & 49.95 & 0.209 & -0.052 \\
+    %  $\tin{124}{112}$ & 2.24005 & 0.00150934 & 49.9906 & 0.194804 & -0.0671999 \\
+    %  $\tin{112}{124}$ & 2.47995 & 0.00182327 & 49.961 & 0.214745 & -0.0742297 \\
+      ${}^{108}$Sn & 2.16 & 0.0015 & 49.99 & 0.207 & -0.059 \\
+      \hline
+    \end{tabular}
+    \caption{Multivariate normal distribution fit parameters for four beams.
+    \label{beamParameters}}
+  \end{center}
+\end{table}
+
+For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values; as seen in Fig.~\ref{beampid132,beampid108}.
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{beam-Sn132.png}
+         \caption{$\tin{132}{124}$ Beam particle identification.}
+         \label{fig:beampid132}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{beam-Sn108.png}
+         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \label{fig:beampid108}
+     \end{subfigure}
+        \caption{$\tin{108}{112}$ Beam particle identification. }
+        \label{fig:beampid}
+\end{figure}
 
@@
         \includegraphics[width=\linewidth]{clusterLR.png} 
-        \caption{Generic} \label{fig:mom_S_before}
+        \caption{Generic} \label{fig:clusterLR}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{clusterTop.png} 
-        \caption{Competitors} \label{fig:mom_L_before}
+        \caption{Competitors} \label{fig:clusterTop}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{clusterBottom.png} 
-        \caption{Generic} \label{fig:mom_S_after}
+        \caption{Generic} \label{fig:clusterBottom}
     \end{subfigure}
@@
 \section{Beam angle selection}
-The beam is bent by the magnetic field before hitting the target.
-The BDCs track the beam outside of the magnet.
-The beam is then propagated through the magnetic field until the target position.
-The beam angle on target is categorized by two angles $\theta_{a_{proj}}$ and $\theta$
+The beam is deflected by the magnetic field and impinges on the target at some angle.
+The BDCs track the beam as a straight line right up until the magnet.
+To calculate the incidence angle, the beam is then propagated through the magnetic field until the target position.
+The beam angle on target is categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,
+
+%define angles 
+%define selection
+
+\begin{equation}
+  \theta_\mathrm{a,proj}=\tan^{-1}\frac{p_x}{p_z},\quad
+  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z},
+  \label{beamAngle}
+\end{equation}
+
+where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector.
+
+Figure~\ref{beamAngleFig} shows the beam angle distribution for ${}^{132}$Sn and ${}^{108}$Sn.
+
+\begin{table}[H]
+  \begin{center}
+    \begin{tabular}{ccccc}
+      \hline 
+      System & $\mu_{\theta_\mathrm{a,proj}}$ &
+      $\sigma_{\theta_\mathrm{a,proj}}$ & $\mu_{\theta_\mathrm{b,proj}}$ &
+      $\sigma_{\theta_\mathrm{b,proj}}$ \\
+      \hline\hline 
+      $\tin{132}{124}$ & 0.61 & 2.94 & -44.18 & 1.96 \\
+    %  $\tin{112}{124}$ & -0.41 & 1.58 & -53.52 & 0.90 \\
+      $\tin{108}{112}$ & -0.42 & 1.95 & -55.17 & 0.97 \\
+      \hline
+    \end{tabular}
+    \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
+      $^{108}$Sn beam angle. \label{beamAngleParameters}}
+  \end{center}
+\end{table}
 
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
-        \caption{Generic} \label{fig:mom_S_before}
+        \caption{Generic} \label{fig:beamangle108}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
-        \caption{Competitors} \label{fig:mom_L_before}
+        \caption{Competitors} \label{fig:beamangle112}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
-        \caption{Generic} \label{fig:mom_S_after}
+        \caption{Generic} \label{fig:beamangle124}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
-        \caption{Competitors} \label{fig:mom_L_after}
+        \caption{Competitors} \label{fig:beamangle132}
     \end{subfigure}
-\label{fig:mom_sc}
+\label{fig:beamangle}
 \end{figure}
@@
 \section{Angular Quality Cuts}
+Since the TPC has no spherical symmetry, there are geometric constraints on measuring tracks.
+Most notably tracks going up and down in the TPC where the path of the track would be shortest.
+Tracks traveling up and down in the TPC will have the shortest track length and will correspond to low regions of efficiency.
+The relevant geometries for this discussion are the corners of the field cage.
+Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the field cage were the angles of each corner is given.
+Since the window to the field cage was slightly higher, the upper angles are slightly smaller than the lower angles.
+The $\phi$ angles location of each corner is are $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{-24 \\ +34}$, with a width of about \ang{58}.
+While it depends on polar angle $\theta$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the $\phi$ angles.
+
+The efficiency in these regions are either zero or very small.
+Typically it is not wise to efficiency correct already small numbers with small efficiency values.
+Therefore we have taken the practice to cut away the regions of very low efficiencies.
+The cuts we have made are for $\theta_{C.M.}$ < 90 and for \ang{326} < $\phi$ < \ang{24} and \ang{1} < $\phi$ < \ang{24}.
+
+Maybe add a table of cuts for each system.
+
+We need to correct for the solid angles not measured, or cut away, to arrive at the correct total yields in $4\pi$. The correction is simple given,
+
+\begin{equation}
+\Delta\Omega = \Delta\phi(\cos(\theta_1) - \cos(\theta_2)),
+\end{equation}
+
+where $\Delta\phi$ is the $\phi$ and $\theta_{1,2}$ are the $\theta$ angular coverage after cuts. The multiplicitive correction factor to the data, C, for the missing angular coverage is simply,
+
+\begin{equation}
+C = \frac{4\pi}{\Delta\phi}.
+\end{equation}
+
+%As the spherical angle of the track, $\theta$, increases the track gets even shorter and shorter until there are very few clusters at all.
+We see around $\phi$=\ang{90} and $\phi$=\ang{270}, representing up and down in the TPC respectively, there is near zero efficiency for nearly all angles $\theta$>\ang{30} and for a wide range of $\phi$ angles around these two values.
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{tpcAngle.pdf}
+\caption{Rough Geometry of the TPC field cage.}
+\label{fig:angleEffExplanation}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
+         \caption{$y=x$}
+         \label{fig:pim108angle}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
+         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \label{fig:pim132angle}
+     \end{subfigure}
+        \caption{Three simple graphs}
+        \label{fig:pim}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
+         \caption{$\tin{108}{112}$ $\pi^-$ angular spectra.}
+         \label{fig:pip108angle}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
+         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \label{fig:pip132angle}
+     \end{subfigure}
+        \caption{Three simple graphs}
+        \label{fig:pip}
+\end{figure}

### 2020-04-28 chapters/ch_experiment.tex
Update saturation section

@@
 \section{Pad Response Function}
-The charge distribution on the pad plane resulting from an electron avalanche is two dimensional, as illustrated in Fig.~\ref{fig:prf}, with the projections onto the $x$ and $z$ axis of this distribution labeled as $\rho(x)$ and $\rho(z)$ respectively.
-The charge observed on each pad is the integral of this distribution over the pad's dimensions, which is called the Pad Response Function (PRF).
+\label{sec:prf}
+Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Fig.~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
+If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed on each pad, Q, is given as,
+
+\begin{equation}
+Q(x_o,z_o) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(x,z) dxdz,
+\label{eq:prfpadCharge}
+\end{equation}
+
+where $x_o$ and $z_o$ represent the x and z coordinates of the center of each pad, w is the width, and l is the length.
+This function is commonly refered to as the Pad Respons Funciton (PRF).
+The total charge observed for this track is a superposition of PRFs of each avalanches on all the anode wires.
+It is the use of this PRF that allows for sub-millimeter accuracy in the position resolution of the TPC \cite{blumrol}.
+Typically in a TPC, hits are grouped into clusters and it is practical to cluster in only one direction.
+The PRF can be expressed in a way that is independent of the location of the avalanche.
+If the true position of the avalanche is ($x_o^{'}$,$z_o^{'}$), and the charge distribution can be written in a position independent way %$\rho(\lambda_x,\lambda_z)$, where $\lambda_x = x - x_o^'$ and $\lambda_z = z - z_o^'$.
+The PRF for layer clustering can be written as,
+
+\begin{equation}
+Q(\lambda_x) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(\lambda_x,\lambda_z) dxdz,
+\label{eq:prflayer}
+\end{equation}
+
+and the PRF for row clustering can be written as,
+
+\begin{equation}
+Q(\lambda_z) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(\lambda_x,\lambda_z) dxdz.
+\label{eq:prfrow}
+\end{equation}
+
+The The direction most perpendicular to the track's direction of travel provides the most accurate position resolution.
+Clustering in the \spirit TPC is done for several pads containing charge in the direction most perpendicular to the track's direction of travel.
 
@@
 
-\subsection{Gatti Pad Response Function}
-Gatti \cite{gatti} derived a semi-empirical formula for the charge distribution in a simple multi-wire TPC.
-The function given in Eq.~\ref{eq:gatti}, depends only on the width of a pad ($w$), the distance of the anode plane to the pad plane ($h$), and the distance of the pad center to the avalanche point $\lambda$.
-It is a single parameter equation where the two parameters $K_1 = \frac{K_{2}\sqrt{K_3}}{4 \arctan(\sqrt{K_3})}$ and $K_2 = \frac{\pi}{2}\left(1-\frac{\sqrt{K_{3}}}{2}\right)$ depend on the parameter $K_3$, which is a function of the ratio of the anode wire diameter to the distance of the anode wires to the pad plane.
-$K_3$ can be looked up in a graph in \cite{blumrol} and \cite{gatti}.
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{padsat_Large}
+\caption{A cartoon illustration of the charge distribution resulting from an electron avalanche on one wire and the projections of the distribution onto the two axis $\rho(x)$ onto the x-axis and $\rho(z)$ onto the z-axis.
+The orientation of the wire planes is flipped upside down to display the perspective better.}
+\label{fig:2DPRF}
+\end{figure}
 
+Gatti \cite{gatti} derived a semi-empirical formula for the charge distribution in a simple multi-wire TPC given as, 
 \begin{equation}\label{eq:gatti}
@@
 \end{equation}
+
+where $w$ is the width of the pad, $h$ is the distance of the anode plane to the pad plane, and $\lambda$ is the distance of the pad center to the avalanche point.
+It is a single parameter equation where the two parameters $K_1 = \frac{K_{2}\sqrt{K_3}}{4 \arctan(\sqrt{K_3})}$ and $K_2 = \frac{\pi}{2}\left(1-\frac{\sqrt{K_{3}}}{2}\right)$ depend on the parameter $K_3$, which is a function of the ratio of the anode wire diameter to the distance of the anode wires to the pad plane.
+$K_3$ can be looked up in a graph in \cite{blumrol} and \cite{gatti}.
 
@@
 
-\begin{figure}[ht!]
-\begin{overpic}[width=\linewidth]{expprf}
+\begin{figure}[ht!]s
+\begin{overpic}[width=\linewidth]{fig5.pdf}
 \put(61,55){\contour{white}{ PRF${}_{\mathrm{Gaus}}(\lambda)$ eq. \ref{eq:gaus}  }}
@@
 
-\section{Pad Response Function vs. crossing angle}
 The shape of the PRF depends on the crossing angle of the track \cite{gatti}.
-Plotted in Fig.~\ref{fig:prfangle} is the PRF of $\pi^-$ tracks vs.
+Plotted in Fig.~\ref{fig:prfpimData} is the PRF of $\pi^-$ tracks vs.
 the crossing angle $\theta$.
@@
 
-\begin{figure}[ht!]
-\includegraphics[width=\linewidth]{PRF_vsAngle}
-\caption{Experimental pad response function as a function of the crossing angle $\theta$ with the corresponding fits.}
-\label{fig:prfangle}
+\begin{figure}[!htb]
+     \centering
+	 \includegraphics[width=\textwidth]{PRFs_data_wcut.png}
+     \caption{PRF response from $\pi^-$ data. }
+     \label{fig:prfpimData}
 \end{figure}
@@
 \begin{figure}[ht!]
-\begin{overpic}[width=\linewidth]{normsigma}
-\end{overpic}
+\vspace{5mm}
+\includegraphics[width=\linewidth]{fig7}
 \caption{Parameters $N_{0}$ and $\sigma$ as a function of the crossing angle $\theta$ with the $4^{th}$ order polynomial fits.}

### 2020-04-29 chapters/ch_data_analysis.tex
Update

@@
 \subsection{Drift Task}
+The interaction point in the gas and the energy loss information from Geant4 is converted into total number of electrons by the ionization coefficient of P10 gas given in Table~\ref{tb:gas}.The electric field set up by the field cage is uniform for the area covered by the pad plane.
+Each electron drift is simulated by a straight line trajectory until reaching an anode wire, where the total length drifted is given by $L_{anode}$.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{Electron Transport Gas Properties} \\
+ \midrule
+Drift velocity & 5.53 $\si{\centi\meter\per\micro\second}$\\
+Transverse diffusion & 240 $\si{\micro \meter \centi\meter}^{-1/2}$\\
+Longitudinal diffusion &  340 $\si{\micro \meter \centi\meter}^{-1/2}$\\
+Gas Ionization & 26.2 $\si{\eV}$\\
+\bottomrule
+\end{tabular}
+\caption{An overview of the properties of the TPC}
+\label{tb:gas}
+\end{table*}
+
+The stochastic motion of the electrons can be categorized by the longitudinal, $c_{l}$, and transverse, $c{t}$, diffusion coefficients which are determined by Garfield++ calculations in the presence of a 0.5T magnetic field.
+The diffusion of electrons can be modeled by a Gaussian process where the sigma of the longitudinal and transverse directions are given as a function of the total length, $\sigma_{l}=c_{l}\cdot\sqrt{L_{anode}}$ and $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.
+Where the magnitude of deviation in the transverse direction ($dr$) can be randomly sampled from,
+
+\begin{equation}
+dr = e^{-\frac{r^2}{2\sigma_{t}^2}}
+\end{equation}.
+
+The cartesian directions are given by $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, as there is no preferential angle in the transverse emission.
+The longitudinal diffusion is simulated in the same manner by randomly sampling from,
+
+\begin{equation}
+dt = e^{-\frac{t^2}{2\sigma_{l}^2}}
+\end{equation}.
+
+ The total drift time is calculated as,
+ $t = L_{anode}/v_d + dt + t_{offset}$, 
+  from the drift velocity, longitudinal diffusion, and $t_{offset}$, a parameter added to calibrate the absolute MC time spectrum with the data.
+
+The z-position each electron is terminated onto the nearest anode wire.
+There are three anode wires lying below each pad as shown in FIGURE [add andode wire fig].
+In the near vicinity of an anode wire the electric fields are high enough to cause the electron to be greaty accelerated and create and avalanche of electrons to be produced.
+This distribution of the total number of electrons produced from an avalanche of a single electron was simulated in Garfield++, Fig.~\ref{fig:gaincalib}.
+Two anode wire voltages were applied which correspond to two different anode gains, high and low.
+
+While the overall gain describes the MC data (as compared to the data) quite well, there are small momentum dependent corrections to the dE/dx applied to each particle type as a final calibration of the dE/dx of the MC.
+
 The drift task takes the energy loss points calculated from Geant4 and converts them into electrons.
@@
 \subsection{Pad Response Task}
+
+The signals induced on the pads from an avalanche event create a distribution of charges on the pad plane. A simple two dimensional Gaussian pad response function in both the x and z direction,
+
+\begin{equation}
+PRF(x,z) = \iint e^{-\frac{(x-x_0)^2}{2\sigma_x^2}} e^{-\frac{(z-z_0)^2}{2\sigma_z^2}}dxdz
+\end{equation},
+
+where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the endpoints of each electron.
+The experimentally observed PRF in data depends on the crossing angle of the track due to how the charge is collected on a pad.
+Performing a MC simulation and comparing the MC to data PRF shows that a simple Guassian PRF is all that is needed to simulate the data PRF.
+The PRF does not have any significant dependence on the particle type.
+
 The total charge of the event is split over several pads defined by thePRF (as discussed in section \ref{sec:prf}.
@@
 \subsection{Electronics Task}
+The dynamic range, $d_{range}$, of the TPC electronics was set to 120 $\si{\femto \coulomb}$ over the full 12 bit resolution resulting in 4096 channels, $ADC_{max}$.
+The pedestal ($ADC_{pedestal}$) was set to 300 channels.
+Therefore the conversion from number of electrons to ADC channels is calculated as,
+
+\begin{equation}
+ADC = A_{tot} \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{d_{range}}
+\end{equation},
+
+where $e$ is the typical charge of an electron in $\si{\femto \coulomb}$, and $A_{tot}$ is the total number of avalanche electrons.
+The response of an electronics signal is simulated by measuring the standard pulse response to an input signal in the data.
+It was found the pulse shape did not depend on the momenta, angle, or particle type in any significant way.
+Therefore a standard pulse was extracted from an average over all particle types and signal heights -- as long as it were not saturating-- shown in Fig.~\ref{fig:pulsea}.
+The superposition of the individual pulses from each electron makes up the Monte Carlo (MC) time bucket spectra in ADC.
+This can be directly fed into the software.
+The experimental electronic noise (RMS) was measured to be around 6 ADC; a random number --drawn from a Guassian with a sigma of 6 ADC and 0 mean -- was assigned to each time bucket to simulate the experimental noise.
+
 The electronics task simulates the electronics response to induced charge on a pad, converting electronics into ADC channels.
@@
 \subsection{Simulating Saturation}
+
+Saturation of the electronics manifests itself in two different ways.
+Some channels are dead for the full time bucket range in an event.
+These channels are randomly distributed along the beam path.
+Though the gating grid of the TPC is closed, an un-reacted beam produces several delta electrons.
+The electrons emitted in directions along the magnetic field lines can physically pierce this grid and avalanche on the anode wire causing channels along the beam path to be saturated.
+The recovery time depends on the input charge and can take several ms to recover, killing the pad for future events.
+Tracks depositing enough energy may saturate an electronics channel for the event shadowing tracks that may cross below.
+
+From the time a channel has been saturated, no further signals can be read out.
+To simulate this behavior when embedding signals from Monte Carlo tracks, we must first identify when in a pad a saturating signal has occurred.
+Non-saturating signals have a characteristic long exponential tail where as saturated signals have no long tail and prematurely end.
+By applying a software pole-zero correction technique, the long tail of the non-saturated signal can be removed.
+In the case of saturated pules this creates a negative undershoot which we use to identify when a channel has been saturated.
+
+Before embedding any MC signals into the software, the pole-zero compensation is applied to see when in time the saturation has occurred.
+The point in which MC signals cannot be embedded is set to 30 Tbs prior to this negative undershoot to prevent any MC signals from superimposing upon the saturated signal.
+Another routine identifies if a pad is ``dead" for the event and no MC signals may be embedded.
+
+\begin{equation}
+Corrected(i) = (b_1\cdot Corrected(i-1) + a_0\cdot Pulse(i) - a_1\cdot Pulse(i-1))/b_0
+\end{equation}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{satpulse.png} 
+        \caption{Standard pulse } \label{fig:pulsea}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{satpulsepz.png} 
+        \caption{Pole zero correction} \label{fig:pulseb}
+    \end{subfigure}
+\label{fig:pulse}
+\end{figure}
+
 There are several types of saturation that must be simulated or accounted for to correctly model the TPC response.
@@
         \centering
-        \includegraphics[width=\linewidth]{deltaE_topview} 
+        \includegraphics[width=\linewidth]{deltaE_topview_neg} 
         \caption{Top view} \label{fig:deltaE_topview}
@@
         \centering
-        \includegraphics[width=\linewidth]{deltaE_sideview} 
+        \includegraphics[width=\linewidth]{deltaE_sideview_neg} 
         \caption{Side view} \label{fig:deltaE_sideview}
@@
 \section{Monte Carlo Track Embedding}
+
+The flow of the software implementation of embedding is shown in Fig.~\ref{fig:flow}.
+The raw digitized signals of the MC track are embedded into the experimental data time bucket spectrum pad by pad, we refer to this as the embedded data.
+The Pulse Shape Algorithm (PSA) task reconstructs hits from the experimental, MC, and embedded data set independently.
+By comparing the hits identified in the embedded set against the experimental data set we can identify which embedded hits belong to the original experimental hits; for hits to match they must satisfy two criteria.
+The difference between the two charges be less than 5\% of the charge value of the data hit and the time difference between hits must be less than 3 Tb.
+We then remove the hits in the embedded data set which satisfy the original experimental data hits.
+The surviving candidates are then compared with the MC hit set, where the criteria for a matching hit is the time difference is less than 3 Tb.
+We do not require a matching criteria for the charge values since the pions we wish to study induce small signals relative to most of the particles producing signals in the experimental data.
+Since we embed these small signals on top of larger signals, the PSA reconstruction of the charge value was found to be $>$ 5\% of the charge value of the embedded hit.
+Since we removed almost all of the experimental hits in the first step, requiring only a Tb cut in the second comparison is sufficient to tag the embedded MC hits.
+
+Each hit is tagged as an embedded hit.
+From here the hits are then organized into helix tracks by the tracking algorithm.
+If a track has one hit that is embedded it will be tagged as embedded.
+Once the hits are sorted into independent tracks adjacent hits are clustered together into one ``cluster".
+If any of the hits which constitute a cluster are flagged as embedded, the cluster is also tagged as embedded.
+
+After the clustering algorithm, the final momenta of each track is fit from the collection of clusters and these final tracks are tagged as embedded if they comprise of any embedded clusters.
+The goal of this na\"ive tagging is to preserve all the information of where the embedded hits, clusters have gone.
+Out of all the reconstructed tracks, which are tagged as embedded, must satisfy two criteria to finally be tagged as a true embedded track.
+The embedded track must have more than 5 embedded clusters in the track and the fraction of embedded clusters to the total number of clusters in a track must be $\geq$ 50\%.
+The first criteria is a simple minimum bias cut where the second criteria is the strongest cut on identifying tracks.
+This cuts out tracks which may have picked up a couple of embedded hits from being close in proximity in some regions.
+Each track that passes these two criteria is labeled as the embedded track.
+A track may be split for several reasons, physically introduced by crossing tracks, or by saturation, and each track is saved into an array containing all embedded tracks.
+
+For efficiency analysis
+
+\begin{sidewaysfigure}[H]
+\centering
+\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
+\caption{Flow of embedding implementation in the software.}
+\label{fig:flow}
+\end{sidewaysfigure}
+
+%\clearpage
+%\pagestyle{lscape} % first clear the page and change the pagestyle
+%\begin{landscape}
+% your landscape table(s) or figure(s) here
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
+\caption{Flow of embedding implementation in the software.}
+\label{fig:flow}
+\end{figure}
+
+%\end{landscape}
+%\pagestyle{plain}
+
+Geant4 has considerably simplified the understanding the response and inefficiencies of a particular experimental setup.
+The TPC monte carlo simulation was built upon Geant4, including the dimensions, materials, and gas composition.
+Several classes were built to simulate the operations of a Time Projection Chamber (TPC).
+This included the drift properties of electrons, avalanche process, and electronics response to input signals.
+Combined, these classes simulate the response of the TPC to any input MC track and can be directly input into the \spirit TPC software reconstruction algorithm.
+
+While the goal of these tasks are to simulate important properties of a TPC, there are several effects which cannot be easily simulated.
+Several examples are biases due to triggering detectors (preferential selection of a reaction plane), saturation effects in the electronics, track multiplicity, and several other systematic sources common to all experiments.
+Simulating each issue individually would be impractical and time consuming if not impossible, as in the case of trigger logic bias.
+By embedding monte carlo signals into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all sources, known or unknown.
+Details of the MC simulation and the embedding technique are described in this paper.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{perspective_embed}
+        \caption{Generic} \label{fig:persEmbed}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.3\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{top_embed} 
+        \caption{Competitors} \label{fig:topEmbed}
+    \end{subfigure}
+     \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{side_embed} 
+        \caption{Competitors} \label{fig:sideEmbed}
+    \end{subfigure}
+    \caption{A 200 MeV/c $\pi^-$ embedded into a nuclear collision type event. The embedded track identified by the software is highlighted by the solid green line. }
+
+\label{fig:embedtrack}
+\end{figure}
+
 Add Figure of MC track embedding

### 2020-05-01 MSU-thesis-estee.tex
Regular update
Update saturation section
Update
MC section

@@
 \usepackage{color}
-\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/}}
+\usepackage{amsmath}
+\usepackage{cleveref}
+\usepackage{rotating}
+
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/}}
 
@@
 chapters/ch_data_analysis,
-chapters/ch_data_analysis_2,
+%chapters/ch_data_analysis_2,
 %chapters/ch_results
@@
 
-\include{chapters/ch_intro}
-\include{chapters/ch_theory}
+%\include{chapters/ch_intro}
+%\include{chapters/ch_theory}
 \include{chapters/ch_experiment}
@@
 %\include{chapters/ch_data_analysis_2}
-\include{chapters/ch_results}
+%\include{chapters/ch_results}
 

### 2020-05-01 chapters/ch_data_analysis.tex
MC section

--- 2020-04-28 chapters/ch_data_analysis.tex
@@
 \section{Software}
+\label{sec:software}
 
@@
 \end{figure}
+
+\subsection{Pulse Shape Algorithm}
+\label{sec:psa}
 
@@
 \section{Monte Carlo Simulation}
-We use Geant4 as and event generator for performing Monte Carlo Simulations in the TPC.
-A scale model of field cage, front window, front window frame, pad plane, and aluminum top plate are modeled.
-The correct materials are used as well as the field cage is modeled with P-10 gas at a density of ????.
-By using Geant4 we can also input the the magnetic field map of the SAMURAI dipole magnet (as calculated by the SAMURAI group via a TOSCA simulation).
+The monte carlo (MC) simulation is a to part simulation.
+The first is the modeling of the fundamental interactions of a particle through material which we utilize Geant4 as an event generator.
+A scale model of field cage, front window, front window frame, pad plane, and aluminum top plate are modeled with the correct material type and dimensions assuming P-10 gas at a density of 1 atm.
+We also import the the magnetic field map of the SAMURAI dipole magnet (as calculated by the SAMURAI group via a TOSCA simulation) \cite{magnet}.
 In this way any particle type may be studied and the full interactions (scattering, decay, energy loss, path taken, etc.), are accounted for.
 The output of this simulation is a series of energy loss points which contain the amount of energy lost in $keV/cm$ and the position in $(x,y,z)$.
+The second part of the MC simulation simulates the physical processes involved in the measurement of the TPC such as the drift, avalanche, and signal creation in the electronics, to simulate the detector response of the MC event.
+This is separated into several software tasks; the drift, pad response, and electronics task.
 
-Separate software tasks model the converting energy loss into electrons, the drifting of electrons, the avalanche process, and the electronics response.
+Here we will only discuss in detail, the tasks involved in simulating the TPC response.
+The discussion of how we use Geant4 is not particularly interesting, since we use it only to get the energy loss in a gas and through basic materials.
 
@@
 \subsection{Drift Task}
-The drift task takes the energy loss points calculated from Geant4 and converts them into electrons.
-and then modeling their drift behavior in the field cage.
-As discussed previously the various processes[CITE To formula about drifting electronc]
-
- A full microscopic treatment of the stochastic nature of each electron would be too cumbersome; most of the properties of the electrons are described by macroscopic quantities can be described by 
-The charge of each MC point is converted to the total number of electrons liberated in the P-10 gas.
-This is a well understood property of proportional counters and is stable over a wide range of velocities and particle types [CITE BOOK].
-The conversion factor of P-10 can be calculated by considering the partial volumes of each component of the gas.
-\ref{eq:kev2el}.
+Geant4 returns the primary ionization points and energy loss information in the gas volume. In the drift task this is converted into the total number of electrons, $N_{e^-}$,
 
 \begin{equation}
-Number of e^{-} = 28 keV/cm
+N_{e^{-}} =  \frac{\Delta E}{I},
 \label{eq:kev2el}
@@
 
-\subsection{Pad Response Task}
-The total charge of the event is split over several pads defined by thePRF (as discussed in section \ref{sec:prf}.
-As shown in Fig.~\ref{fig:2DPRF}, there are three wires that lie directly above a given pad.
-The $z$ coordinate of the avalanche is limited to one of these three wires, where as the $x$ coordinate can be any value along the wire.
-A 2-dimensional Gaussian was enough to describe the experimental PRF,
+where the ionization coefficient of P10 gas, $I$, given in Table~\ref{tb:gas} and the energy loss, $\Delta E$, given in \si{\electronvolt}.Since the electric field set up by the field cage is uniform for the area covered by the pad plane, we assume also each electron drifts in a straight line trajectory until reaching an anode wire, where the total length drifted is given by $L_{anode}$.
+In this task the option to put in space charge effects and non-uniform electric fields are allowed but we do not use them for the further discussion of calculating efficiency.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{Electron Transport Gas Properties} \\
+ \midrule
+Drift velocity & 5.53 $\si{\centi\meter\per\micro\second}$\\
+Transverse diffusion & 240 $\si{\micro \meter \centi\meter}^{-1/2}$\\
+Longitudinal diffusion &  340 $\si{\micro \meter \centi\meter}^{-1/2}$\\
+Gas Ionization & 26.2 $\si{\eV}$\\
+\bottomrule
+\end{tabular}
+\caption{An overview of the properties of the TPC}
+\label{tb:gas}
+\end{table*}
+
+The drifting electrons exhibit stochastic motion which can be described by a longitudinal, $c_{l}$, and transverse, $c_{t}$, diffusion coefficients, which are determined by Garfield++ calculations -- in the presence of a 0.5T magnetic field.
+The diffusion of electrons can be modeled by randomly sampling from a Gaussian distribution in both the longitudinal and transverse directions of the electron drift.
+The stochastic deviation in the transverse direction, $dr$, can be randomly sampled from,
 
 \begin{equation}
-PRF(x,z) = \frac{1}{2\pi\sigma_z\sigma_x}\exp^{\frac{-{(x-x_o)}^2}{2\sigma_x^2}}\exp^{\frac{-{(z-z_o)}^2}{2\sigma_z^2}},
-\label{eq:2DPRF}
+dr = e^{-\frac{r^2}{2\sigma_{t}^2}},
 \end{equation}
 
-where the location of the avalanche is ($x_o$,$z_o$).
-The $\sigma$ values were tuned to match the experimental data.
-Figure~\ref{mcdataPRF} shows the PRF response of $\pi^-$ tracks.
-The black line is the PRF fit to the experimental data.
-We can see a good agreement between the MC, Fig.~\ref{fig:prfpimMC}, and experimental data, Fig.~\ref{fig:prfpimData}.
+where $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.The Cartesian directions are given by $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, as there is no preferential angle in the transverse directions.
+The longitudinal diffusion is simulated in the same manner by randomly sampling from,
+
+\begin{equation}
+dt = e^{-\frac{t^2}{2\sigma_{l}^2}},
+\end{equation}.
+
+where $\sigma_{l}=c_{l}\cdot\sqrt{L_{anode}}$.
+
+The total drift time, $t$, is calculated as,
+
+\begin{equation}
+ t = L_{anode}/v_d + dt + t_{offset},
+\end{equation}
+
+ where $v_d$ is the drift velocity and $t_{offset}$ is an offset parameter added to align the MC time bucket spectrum with the data.
+The electron is assumed to terminate on the nearest anode wire which will fix the z-position to be the same as the that anode wire.
+
+The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++, Fig.~\ref{fig:gaincalib}.
+Two anode wire voltages were applied which correspond to two different anode gains, high and low, in the TPC experiment.
+We randomly sample from these two distributions, depending on which anode wire section the electron terminated on and store the number of electrons as an effective gain value.
+We do not assume there is much lateral movement along the wire in the avalanche process.
+
+\subsection{Pad Response Task}
+The total charge of the event is split over several pads defined by the PRF, as discussed in section \ref{sec:prf}.
+We simulate the PRF as the double integral of a 2-dimensional Gaussian with the final output charge on each pad being the superposition of the PRFs of all drifted electrons.
+The MC PRF is expressed as,
+
+\begin{equation}
+PRF(x,z) = \iint e^{-\frac{(x-x_o)^2}{2\sigma_x^2}} e^{-\frac{(z-z_o)^2}{2\sigma_z^2}}dxdz,
+\end{equation}
+
+where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the final position of the drifted electron.
+The $\sigma$ parameters were determined through an iterative comparison between the MC and experimental data set.
+We can see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
 
@@
 \subsection{Electronics Task}
-The electronics task simulates the electronics response to induced charge on a pad, converting electronics into ADC channels.
-The ADC value of a channel, R, is modeled as,
+%The dynamic range, $d_{range}$, of the TPC electronics was set to 120 $\si{\femto \coulomb}$ over the full 12 bit resolution resulting in 4096 channels, $ADC_{max}$.
+The pedestal ($ADC_{pedestal}$) was set to 300 channels.
+The response of an electronics signal is simulated by measuring the standard pulse response to an input signal in the data.
+It was found the pulse shape did not depend on the momenta, angle, or particle type in any significant way.
+Therefore a standard pulse was extracted from an average over all particle types and signal heights -- as long as it were not saturating-- shown in Fig.~\ref{fig:pulsea}.
+The superposition of the individual pulses from each electron makes up the Monte Carlo (MC) time bucket spectra in ADC.
+This can be directly fed into the software.
+The experimental electronic noise (RMS) was measured to be around 6 ADC; a random number --drawn from a Guassian with a sigma of 6 ADC and 0 mean -- was assigned to each time bucket to simulate the experimental noise.
+
+The electronics task simulates the electronics response, converting  the induced charge on a pad into ADC channels.
+The conversion from number of electrons to the response, $R$, in ADC channels is calculated as,
 
 \begin{equation}
-\mathrm{R}= \frac{ADC_{\mathrm{Max}} - ADC_{\mathrm{Pedestal}}}{G*f_c},
+R = A_{tot} \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{f_c}
 \label{eq:etoADC}
-\end{equation}
+\end{equation},
 
-where $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), $\mathrm{ADC_{Pedestal}}$ is the baseline or pedestal value used in the experiment (???), $G$ is the gas gain, and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
-Later a section is dedicated to small particle dependent corrections which account for small changes in the charge calibration which may be due to a whole host of detector effects that are not modeled.
+where $e$ is the typical charge of an electron in $\si{\femto \coulomb}$, $A_{tot}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
+We add random Gaussian noise as measured from the experimental simulating the electronic noise; (RMS) was measured to be around 6 ADC.
 
-The induced current in the pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape read out.
-The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type; for a given shaping constant.
-This allows us to assume the pulse shape is constant with the only two variables being the height of the pulse, Q, and the starting time bucket of the pulse, $t_o$.
-The shaping constant was set to \SI{117}{\nano\second} for most of the data.
-Figure~\ref{fig:pulseshape} shows the pulse shape for a normal pulse that is within the dynamic range as the dashed blue curve, the pulse is normalized so the maximum height is 1.
+The induced current in the pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
+The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, for a given shaping constant.
+This allows us to assume the pulse shape is constant with the only two variables being the height of the pulse, $Q$, and the starting time bucket of the pulse, $t_o$.
+The shaping constant was set to \SI{117}{\nano\second} for the data analyzed here.
+Figure~\ref{fig:pulseshape} shows the pulse shape experimentally extracted from signals that did not saturate the electronics over a wide range of ADC values; it is normalized so the maximum height is 1.
 
-\begin{figure}[!hbt]
-\includegraphics[width=\linewidth]{pulseshape}
-\caption{Pulse shapes for normal and saturated signals. Both are normalized to a max height of 1.}
+\begin{figure}[!htb]
+    \centering       
+    \includegraphics[width=\linewidth]{satpulse.png} 
+    \caption{Standard pulse }
     \label{fig:pulseshape}
@@
 \subsection{Simulating Saturation}
+
+%Saturation of the electronics manifests itself in two different ways.
+Some channels are dead for the full time bucket range in an event.
+These channels are randomly distributed along the beam path.
+Though the gating grid of the TPC is closed, an un-reacted beam produces several delta electrons.
+The electrons emitted in directions along the magnetic field lines can physically pierce this grid and avalanche on the anode wire causing channels along the beam path to be saturated.
+The recovery time depends on the input charge and can take several ms to recover, killing the pad for future events.
+%Tracks depositing enough energy may saturate an electronics channel for the event shadowing tracks that may cross below.
+
+%From the time a channel has been saturated, no further signals can be read out.
+To simulate this behavior when embedding signals from Monte Carlo tracks, we must first identify when in a pad a saturating signal has occurred.
+Non-saturating signals have a characteristic long exponential tail where as saturated signals have no long tail and prematurely end.
+By applying a software pole-zero correction technique, the long tail of the non-saturated signal can be removed.
+In the case of saturated pules this creates a negative undershoot which we use to identify when a channel has been saturated.
+
+%Before embedding any MC signals into the software, the pole-zero compensation is applied to see when in time the saturation has occurred.
+The point in which MC signals cannot be embedded is set to 30 Tbs prior to this negative undershoot to prevent any MC signals from superimposing upon the saturated signal.
+Another routine identifies if a pad is ``dead" for the event and no MC signals may be embedded.
+
 There are several types of saturation that must be simulated or accounted for to correctly model the TPC response.
@@
 
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\linewidth]{satpulsepz.png} 
+\caption{Pole zero correction} 
+\label{fig:pulseSatTag}
+\end{figure}
+
 In the case where the input signal is very large, the electronics can be dead for up to \SI{35}{\milli\second} \cite{akiGET}.
@@
 \section{Monte Carlo Track Embedding}
-Add Figure of MC track embedding
+ The TPC monte carlo simulation was built upon Geant4, including the dimensions, materials, and gas composition.
+Several classes were built to simulate the operations of a Time Projection Chamber (TPC).
+This included the drift properties of electrons, avalanche process, and electronics response to input signals.
+Combined, these classes simulate the response of the TPC to any input MC track and can be directly input into the \spirit TPC software reconstruction algorithm.
+
+While the goal of these tasks are to simulate important properties of a TPC, there are several effects which cannot be easily simulated.
+Several examples are biases due to triggering detectors (preferential selection of a reaction plane), saturation effects in the electronics, track multiplicity, and several other systematic sources common to all experiments.
+Simulating each issue individually would be impractical and time consuming, if not impossible.
+By embedding monte carlo tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all sources these type of sources with a monte carlo type approach.
 
 Track embedding is the process of taking a simulated MC track from Geant4 and embedding its response into a real data event.
-After reconstructing this new embedded event we match the input MC track embedded to its corresponding final reconstructed track.
+After reconstructing this new embedded event we match the input MC track to its corresponding final reconstructed track.
 By doing so we can evaluate the response of the entire TPC system to any given input value.
-The TPC system is composed of three major components (each which can introduce errors and or biases) the software, the detector, and the experimental setup.
+The TPC system is composed of three major components, each which can introduce errors and or biases; the software, the detector, and the experimental setup.
 
-As discussed in [SOFTWARE CHAPTER] the software is composed of several task, each which introduce some error.
-Table {REF TABLE] listing some of the errors each task may introduce illustrating how difficult propagating the errors would be through each system.
+As discussed in Section~\ref{sec:software}, the software is composed of several task, each which introduce errors through assumptions made in the tracking algorithm.
+The detector system itself introduces errors related to the physical measurement itself.
+By modeling the TPC, and its materials, the a Geant4 simulation provides an accurate description of interactions of a particles traversing the materials of the TPC.
 
-The detector system itself introduces errors related to the physical processes of the measurement itself.
-To address this we model the TPC (and its materials) in a Geant4 simulation which provides an accurate description of various interactions of a particle traversing the materials (including the gas volume) of the TPC.
+% Evaluating the errors introduced by the software is the most straight forward exercise; let the software routine process an input and measure the result.
+Understanding the physical measurement requires modeling the physics involved in the theory and operations of TPCs and the electronics.
 
- software is the most straight forward; let the software routine process an input and measure the result.
-Understanding the measurement requires modeling the physics involved in the theory and operations of TPC's and the  electronics.
-The experimental setup itself is quite large and complex, several ancillary detectors such as the Kyoto multiplicity array, Krakow veto array, Active veto array, beam identification detectors, etc.
+ % The experimental setup itself is quite large and complex, several ancillary detectors such as the Kyoto multiplicity array, Krakow veto array, Active veto array, beam identification detectors, etc.
 Even if a full accurate model could be constructed the complex trigger logic of the DAQ system would be impossible to model.
@@
 Therefore, by inputting the MC data into a real experimental event (and measuring the output of that MC track) we can estimate the errors of the experimental setup.
+Before begging the detailed description of the embedding software, it is worth noting that the MC embedding was analyzed as a separate data set from the published data to avoid any issue of MC embedded tracks making it into the final published data.
 
-The software analysis routine and the bias introduced by the trigger settings of the experiment introduce systematic errors in the reconstruction of tracks
+The detailed flow diagram of the software implementation of embedding is shown in Fig.~\ref{fig:flow}.
+
+%\begin{sidewaysfigure}[H]
+%\centering
+%\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
+%\caption{Flow of embedding implementation in the software.}
+%\label{fig:flow}
+%\end{sidewaysfigure}
+
+%\clearpage
+%\pagestyle{lscape} % first clear the page and change the pagestyle
+%\begin{landscape}
+% your landscape table(s) or figure(s) here
+%CONSIDER ROTATING THIS FIGURE IF ALLOWED
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{FlowEmbedding.pdf}
+\caption{Flow of embedding implementation in the software.}
+\label{fig:flow}
+\end{figure}
+
+%\end{landscape}
+%\pagestyle{plain}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{mcresponse.png}
+        \caption{200 MeV/c $\pi^-$ MC digitized track} \label{fig:mcevent}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{event5.png} 
+        \caption{Experimental data event.} \label{fig:dataevent}
+    \end{subfigure}
+    \caption{A 200 MeV/c $\pi^-$ embedded into a nuclear collision type event. The embedded track identified by the software is highlighted by the solid green line. }
+
+\label{fig:embedtrack}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{perspective_embed}
+        \caption{Generic} \label{fig:persEmbed}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.3\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{top_embed} 
+        \caption{Competitors} \label{fig:topEmbed}
+    \end{subfigure}
+     \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{side_embed} 
+        \caption{Competitors} \label{fig:sideEmbed}
+    \end{subfigure}
+    \caption{A 200 MeV/c $\pi^-$ embedded into a nuclear collision type event. The embedded track identified by the software is highlighted by the solid green line. }
+
+\label{fig:embedtrack}
+\end{figure}
+
+Once the MC simulation of a track response in the TPC is created, we directly embed the MC signals into data by superimposing the MC signals into the data time bucket spectrum, pad-by-pad.
+We will refer to this superimposed data as ``embedded data", only the MC signals as ``MC data", and data only containing the experimental signals as ``experimental data".
+The three set of data are each independently analyzed by the PSA algorithm, Section~\ref{sec:psa}, finding the hits in each data set.
+Part of the PSA task is to identify which hits in the embedded data are MC hits and which are from the original data.
+Once identified the embedded hits can be tagged and tracked through the entire software.
+
+ By comparing the hits identified in the embedded set against the experimental data set we can first identify which embedded hits belong to the original experimental hit subset.
+For hits to match, they must satisfy two criteria; $\left|(Q_{\mathrm{Data}} - Q_{\mathrm{Embed} })/Q_{\mathrm{Data}}\right| < .05$ and $\left|t_{\mathrm{Embed} } - t_{\mathrm{Data} }\right| < 3$ where $Q$ and $t$ represent the charge and time of the hit respectively.
+We then remove these hits in the embedded data set which satisfy the conditions of experimental data hits.
+
+The surviving hits are then compared with the MC hit data set, where the criteria for a matching MC hit is, $\left|t_{\mathrm{Embed} } - t_{\mathrm{MC} }\right| < 3$.
+We do not require a matching criteria for the charge values since the pions we wish to study induce small signals relative to most of the particles producing signals in the experimental data as not to bias the charge values in the embedded data.
+Since we removed almost all of the experimental hits in the first step, requiring only a time bucket cut in the second comparison is sufficient to tag the embedded MC hits.
+Each embedded hit that passes both criteria is tagged as an MC embedded hit.
+From here, the embedded data is treated as if it were real data passing through all the software algorithms as we use in the real data analysis.
+If a helix track has one hit that is embedded it will be tagged as embedded.
+Hits in a helix track are then clustered; if a cluster contains one embedded hit it is itself tagged as an embedded cluster.
+If a track that is reconstructed contains any embedded clusters, it is an embedded track.
+
+The goal of this na\"ive tagging is to preserve all the information of where the embedded hits, clusters, and tracks have gone.
+The  job of the embedded correlation task is to identify which of the embedded tracks are candidates for the original input track.
+Several things may happen along the process of embedding.
+A track could break up, lose or share its charge with and adjacent track, or maybe not be identified for other reasons.
+For the embedded track to be a candidate it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track and $N_{\mathrm{Total}}$ is the number of total clusters.
+The first criteria is a simple minimum cut where to ensure there exists an embedded track and the second criteria is the strongest cut ensuring the track has at least half of its clusters coming from embedded MC signals.
+The set of tracks which satisfy both conditions are saved into an array of tracks which the user can use to study efficiency or track splitting.
+
+In the case of efficiency analysis the user must select what they believe to be the track which most represents the original track.
+In this thesis, the track with the minimum distance to vertex is identified as the correct track.
+Though track splitting may occur, typically the split track has a distance to vertex which is very large and cut out from the analysis.
+The details of the efficiency analysis will be described later.
 

### 2020-05-05 MSU-thesis-estee.tex
Update experiment section

@@
 
-\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/}}
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} }
 
@@
 %
-
 %User defined commands
@@
 
+\DeclareSIUnit\torr{Torr}
+\DeclareSIUnit\eVperc{\eV\per\clight}
+\DeclareSIUnit\micron{\micro\metre}
+\DeclareSIUnit\mrad{\milli\rad}
+\DeclareSIUnit\gauss{G}
+
 \includeonly{
@@
 chapters/ch_experiment,
-chapters/ch_data_analysis,
+%chapters/ch_data_analysis,
 %chapters/ch_data_analysis_2,
@@
 \include{chapters/ch_experiment}
-\include{chapters/ch_data_analysis}
+%\include{chapters/ch_data_analysis}
 %\include{chapters/ch_data_analysis_2}

### 2020-05-05 chapters/ch_data_analysis.tex
Update experiment section

@@
 Need to explain pedestal subtraction
-GG noise subtraction
 
@@
 \section{Calibrations and Corrections}
+
+\subsection{Gating grid noise subtraction}
+The gating grid essentially is a fast switch which allows the charges of adjacent wires to reach equilibrium as fast as possible.
+In practice the impedance of both sides was not entirely matched properly and this caused an oscillating current to bounce back and forth in an under-dampened manner.
+This charge bouncing back and forth caused induced signals early in the time bucket spectrum which created and extra source of noise in the data.
+The ground grid shielded some of the noise, but since the path to ground was located at the ends of the TPC, the gating grid noise varied in intensity across the pads.
+
+The gating grid itself was not always stable an only the ${}^{132}$Sn and ${}^{124}$Sn beams were relatively stable.
+Several times during the ${}^{108}$Sn and ${}^{112}$Sn beams the gating grid broke.
+After several data runs were taken, we took what is called gating grid noise runs to get the profile of the gating grid noise throughout the experiment.
+To achieve this the anode wires would be turned off and the trigger was allowed to otherwise operate normally, firing the gating grid to open as usual.
+Since the anode wires were all turned off no signals would be seen except those coming from the gating grid noise.
+Several thousand triggers of the gating grid were taken to get a good noise profile.
+
+The average noise profile was constructed by averaging all the events in the gating grid runs, pad-by-pad; each pad had a slightly different noise profile since the impedance to each pad was slightly different.
+It is worth noting that the small electronics noise would be effectively canceled out by averaging over thousands of events.
+During the first stages of the software, the gating grid noise profile closest to that run would be subtracted from the raw data canceling the gating grid noise in that run.
 
@@
 \label{tb:cocktail300}
-\end{table*}
-
-\begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}rrrrcrrrcrrr@{}}\toprule
-& \multicolumn{3}{c}{$100 MeV$} & \multicolumn{3}{c}{$100 MeV$} & \multicolumn{3}{c}{$300 MeV$}\\
-\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
-& &\multicolumn{2}{c}{Measured} & & \multicolumn{2}{c}{Measured} & & \multicolumn{2}{c}{Measured}\\
-\cmidrule{3-4} \cmidrule{7-8} \cmidrule{11-12}
-Particle &\phantom{abc} & Measured & E$\times$B\\
-\midrule
-p   & 882.8 & 929.5 & 877.3 & 903.5 & 929.5 & 889 &\phantom{abcdef} & f & f \\
-d   & 817.1 & 831.15 & d & 898.5 & e & e & 1621.1 & 1704 & 1612\\
-t   & 589.5 & d & d & 887 & e & e & 1612.4 & f & f  \\
-$^{3}$He  & 1617.3  & d & d & 1795.2 & e & e & 3236.4 & f & f\\
-$^{4}$He  & 1405.6  & d & d & 1782.9 & e & e & 3226.4 & f & f \\
-\bottomrule
-\end{tabular}
-\caption{Summary of expected cocktail. }
-\label{tb:cocktailsummary}
 \end{table*}
@@
 Simulating each issue individually would be impractical and time consuming, if not impossible.
-By embedding monte carlo tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all sources these type of sources with a monte carlo type approach.
+By embedding monte carlo tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all sources with a monte carlo type approach.
 
@@
 
-As discussed in Section~\ref{sec:software}, the software is composed of several task, each which introduce errors through assumptions made in the tracking algorithm.
-The detector system itself introduces errors related to the physical measurement itself.
-By modeling the TPC, and its materials, the a Geant4 simulation provides an accurate description of interactions of a particles traversing the materials of the TPC.
-
-% Evaluating the errors introduced by the software is the most straight forward exercise; let the software routine process an input and measure the result.
-Understanding the physical measurement requires modeling the physics involved in the theory and operations of TPCs and the electronics.
-
- % The experimental setup itself is quite large and complex, several ancillary detectors such as the Kyoto multiplicity array, Krakow veto array, Active veto array, beam identification detectors, etc.
-Even if a full accurate model could be constructed the complex trigger logic of the DAQ system would be impossible to model.
-If we notice that the biases and errors of the entire experimental setup is contained in the measured experimental data.
-Therefore, by inputting the MC data into a real experimental event (and measuring the output of that MC track) we can estimate the errors of the experimental setup.
-Before begging the detailed description of the embedding software, it is worth noting that the MC embedding was analyzed as a separate data set from the published data to avoid any issue of MC embedded tracks making it into the final published data.
-
-The detailed flow diagram of the software implementation of embedding is shown in Fig.~\ref{fig:flow}.
+As discussed in Section~\ref{sec:software}, the software is composed of several tasks, each introduce biases through assumptions made in the tracking algorithm.
+The detector system introduces errors related to the physical measurement itself which can be addressed through modeling the TPC, and its materials.
+The biases involved in the the experimental set up are not always obvious but present in the experimental data.
+All three of these can be simultaneously addressed by embedding of MC tracks.
+Before beginning  the detailed description of the embedding software, it is worth noting that the MC embedding was analyzed as a separate data set from the published data to avoid any issue of MC embedded tracks making it into the final published data.
 
@@
 
+The detailed flow diagram of the software implementation of embedding is shown in Fig.~\ref{fig:flow}.
 Once the MC simulation of a track response in the TPC is created, we directly embed the MC signals into data by superimposing the MC signals into the data time bucket spectrum, pad-by-pad.
-We will refer to this superimposed data as ``embedded data", only the MC signals as ``MC data", and data only containing the experimental signals as ``experimental data".
-The three set of data are each independently analyzed by the PSA algorithm, Section~\ref{sec:psa}, finding the hits in each data set.
-Part of the PSA task is to identify which hits in the embedded data are MC hits and which are from the original data.
-Once identified the embedded hits can be tagged and tracked through the entire software.
-
+Figure~\ref{fig:mcevent} shows the response of a 200 MeV/c $\pi^-$ track in the TPC; we will embed this track into the data event, Fig.~\ref{fig:dataevent}, for an example.
+We will refer to experimental data containing embedded MC signals as ``embedded data", data only containing the MC signals as ``MC data", and data only containing only the experimental signals as ``experimental data".
+The three set of data are each independently analyzed by the PSA algorithm, Section~\ref{sec:psa},  which finds all the hits in each data set.
+
+After this, the PSA algorithm contains the first and most important part of the embedding software.
+Here to the PSA task has the job to identify which hits in the embedded data are originally MC hits and which are from the original experimental data.
+Once identified, the embedded hits can be tagged and tracked through the entire software.
 By comparing the hits identified in the embedded set against the experimental data set we can first identify which embedded hits belong to the original experimental hit subset.
-For hits to match, they must satisfy two criteria; $\left|(Q_{\mathrm{Data}} - Q_{\mathrm{Embed} })/Q_{\mathrm{Data}}\right| < .05$ and $\left|t_{\mathrm{Embed} } - t_{\mathrm{Data} }\right| < 3$ where $Q$ and $t$ represent the charge and time of the hit respectively.
+For hits to match, they must satisfy two criteria, $\left|(Q_{\mathrm{Data}} - Q_{\mathrm{Embed} })/Q_{\mathrm{Data}}\right| < .05$ and $\left|t_{\mathrm{Embed} } - t_{\mathrm{Data} }\right| < 3$ where $Q$ and $t$ represent the charge and time of the hit respectively.
 We then remove these hits in the embedded data set which satisfy the conditions of experimental data hits.
@@
 The surviving hits are then compared with the MC hit data set, where the criteria for a matching MC hit is, $\left|t_{\mathrm{Embed} } - t_{\mathrm{MC} }\right| < 3$.
-We do not require a matching criteria for the charge values since the pions we wish to study induce small signals relative to most of the particles producing signals in the experimental data as not to bias the charge values in the embedded data.
-Since we removed almost all of the experimental hits in the first step, requiring only a time bucket cut in the second comparison is sufficient to tag the embedded MC hits.
-Each embedded hit that passes both criteria is tagged as an MC embedded hit.
-From here, the embedded data is treated as if it were real data passing through all the software algorithms as we use in the real data analysis.
+We do not require a matching criteria for the charge values since as not to bias the charge values selected as identified MC hits.
+This is only accomplished by first removing almost all of the experimental hits in the first step.
+Each embedded hit that passes this criteria is tagged as an MC embedded hit.
+From here, the embedded data is treated as if it were real data passing through the same software analysis.
+
+The goal of this na\"ive tagging is to preserve all the information of where the embedded hits, clusters, and tracks have gone.
 If a helix track has one hit that is embedded it will be tagged as embedded.
-Hits in a helix track are then clustered; if a cluster contains one embedded hit it is itself tagged as an embedded cluster.
-If a track that is reconstructed contains any embedded clusters, it is an embedded track.
-
-The goal of this na\"ive tagging is to preserve all the information of where the embedded hits, clusters, and tracks have gone.
-The  job of the embedded correlation task is to identify which of the embedded tracks are candidates for the original input track.
+When the hits in a helix track are clustered, if a cluster contains one embedded hit it is itself tagged as an embedded cluster.
+Furthermore, if a track that is reconstructed contains any embedded clusters, it is an embedded track.
+It is the job of the embedded correlation task is to identify which of the final embedded tracks are candidates for the original input track.
 Several things may happen along the process of embedding.
-A track could break up, lose or share its charge with and adjacent track, or maybe not be identified for other reasons.
-For the embedded track to be a candidate it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track and $N_{\mathrm{Total}}$ is the number of total clusters.
-The first criteria is a simple minimum cut where to ensure there exists an embedded track and the second criteria is the strongest cut ensuring the track has at least half of its clusters coming from embedded MC signals.
+A track could break up, lose or share its charge with and adjacent track, or maybe not be identified at all for a variety of other reasons.
+For the embedded track to be a candidate of the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track and $N_{\mathrm{Total}}$ is the number of total clusters.
+The first criteria is a simple minimum cut where to ensure there exists an embedded track.
+The second criteria is the strongest cut ensuring the track has at least half of its clusters coming from embedded MC signals.
 The set of tracks which satisfy both conditions are saved into an array of tracks which the user can use to study efficiency or track splitting.
@@
 In this thesis, the track with the minimum distance to vertex is identified as the correct track.
-Though track splitting may occur, typically the split track has a distance to vertex which is very large and cut out from the analysis.
-The details of the efficiency analysis will be described later.
+This is to get rid of the situation of track splitting where typically the split track has a distance to vertex which is very large and cut out from the analysis.
+The details of how the embedding is used to calculate the efficiency will be described later.
 
@@
 
+The arrival time of signals originating from each pad may differ due to timing delays in the electronics and cabling.
+The timing differences will affect the y-position measurement of each track.
+The y-residuals reveal that the timing differences correspond to about $\pm$\SI{2}{\milli\metre} in position differences.
+The timing difference is stable for each pad across several runs.
+Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows , for three different layers.
+Before the correction one can see large deviations in the y-residuals which correspond to timing differences.
+The mean value of the y-residual distribution is fitted and a map is constructed for each pad.
+The data is then reconstructed subtracting the value of that map, $dy$, for each pad.
+The resulting corrected distribution is shown in Fig.~\ref{fig:coboCorr} for the same layer set.
+Figure~\ref{fig:yoff} also show the summary of all the pads before the correction and after the correction.
+There is a significant improvement in the width of the distribution going from \SI{1.5}{\milli\metre} to \SI{0.6}{\milli\metre}.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=0.33\textwidth]{before_row_at_layer10_yoffset.pdf}
+    \includegraphics[width=0.33\textwidth]{before_row_at_layer30_yoffset.pdf}
+    \includegraphics[width=0.33\textwidth]{before_row_at_layer50_yoffset.pdf}
+
+    \includegraphics[width=0.33\textwidth]{after_row_at_layer10_yoffset.pdf}
+    \includegraphics[width=0.33\textwidth]{after_row_at_layer30_yoffset.pdf}
+    \includegraphics[width=0.33\textwidth]{after_row_at_layer50_yoffset.pdf}
+
+    \caption{TBW\label{TBD}}
+  \end{center}
+  \label{fig:coboCorr}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{before_layer_yoffset.pdf}
+        \caption{Y-offset for all layers before correction.} \label{fig:yoff_beforeLayer}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{after_layer_yoffset.pdf} 
+        \caption{Y-offset for all layers after correction.} \label{fig:yoff_afterLayer}
+    \end{subfigure}
+     \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{before_all_yoffset.pdf} 
+        \caption{Distribution for all pads before correction.} \label{fig:yoff_allBefore}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{after_all_yoffset.pdf} 
+        \caption{Distribution for all pads after correction.} \label{fig:yoff_allAfter}
+    \end{subfigure}
+    \caption{ }
+\label{fig:yoff}
+\end{figure}
+
 \section{Efficiency Corrections}

### 2020-05-05 chapters/ch_experiment.tex
Update experiment section

@@
 \section{S$\pi$RIT TPC Overview}
-Add Overview image with labels
-
+%Add Overview image with labels
 The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (S$\pi$RI TPC) is a Multi-Wire Proportional Counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiments.
@@
 
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{exploded.png}
+\caption{Overview of the \spirit TPC}
+\label{fig:tpcExplode}
+\end{figure}
+
 \subsection{Enclosure}
+The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame.
+All walls are constructed of a aluminum frame with think sheet metal.
+All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much.
+This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC.
+The enclosure itself is made to be gas tight with respects to the outside and the field cage.
+This is to allow for the possibility to run a different gas inside the enclosure than the field cage.
+Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.
 
 \subsection{Voltage Step Down}
+The gap between the field cage's cathode and the ground of the enclosure is quite small.
+To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
+There were 7 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt}.
 
@@
 
-The field cage was designed to hang from the top plate and therefore needed to be of a lightweight construction.
+The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
+It was designed to hang from the top plate and therefore needed to be of a lightweight construction.
 Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
@@
 
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{fc_explode}
-\caption{Exploded view of the S$\pi$RIT TPC}
-\label{fig:fc_explode}
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{fc_overview.pdf}
+\label{fig:fc_overview}
+\caption{FC overview}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{fc_overview2.pdf}
+\label{fig:fc_overview2}
+\caption{FC overview 2}
 \end{figure}
@@
 The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires [CITE].
-The halogen free material chosen was Rodgers ????.
+The halogen free material chosen was Cryogenic-G10.
 We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
@@
 
+\begin{equation}
+V_n = V_{cath} \frac{R_p + (50 - n)R}{49\cdot R + R_p}
+\label{eq:FCstrip}
+\end{equation}
+
+\begin{equation}
+R_p = 49 \cdot R  \left(\frac{ y_{g.g.} - y_{cath} }{ y_{TP} - y_{cath} \frac{V_{cath} - V_{gg}}{V_{cath}} }- 1 \right)
+\label{eq:TP_resistor}
+\end{equation}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{FC_schematic.pdf}
+\label{fig:FC_schematic}
+\caption{Schematic of the electric connections relevant to the Field Cage system.}
+\end{figure}
 \subsection{Voltage Step Down}
 
-To lower the risk of a high voltage spark around the cathode region a voltage step down was developed.
-A series of 7 concentric copper rings with 20 MOhm resistors between them steps the cathode voltage down to ground.
-This minimizes sharp transitions of the electric field and therefore sparking.
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{TPC_schematic.pdf}
+\caption{Schematic of the TPC system}
+\label{fig:TPC_schematic}
+\end{figure}
 
-\subsection{Wire Planes}
-Add figure of gating grid transparency closed and open configuration
+Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls.
+The basic idea of the field cage is to set up a constant electric field in the field cage volume to allow for a constant drift, in the presence of gas, for electrons.
+To accomplish this the field cage is composed of a set of copper strips which are set at a defined voltage by a resistor chain for a fixed distance between strips; setting up a constant electric field.
+
+The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
+The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail.
+The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor, $R$, to the next inside strip and the first outside strip.
+The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field.
+The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC.
+Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips.
+The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor with the option to place an additional resistor in parallel to tune the voltage of the top perimeter.
 
@@
 Add figure of close up of pad plane
+
+\subsection{Wire Planes}
+%Add figure of gating grid transparency closed and open configuration 
+There are three wire planes that sit underneath the pad-plane.
+The wire plane closest to the pad-plane (\SI{4}{\milli\metre}), is the anode wire.
+The next plane (\SI{12}{\milli\metre}), is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}), is the gating grid.
+The gating grid is the first plane that electrons meet as they drift upward towards the anode plane.
+The gating grid is operated as a gate either allowing electrons and ions through or blocking them entirely.
+The ground plane shields the inside volume of the TPC from the high voltages of the anode wires where the avalanche process of the electron takes place.
+
+The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
+If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
+
+In the open configuration the gating grid is transparent to electrons coming from the TPC volume and also allows for ions to move from the avalanche region into the TPC volume.
+It functions as the gate of the TPC where we only open the gating grid when an event meets the selection trigger criteria; minimizing the excessive exposure to the un-reacted beams which would quickly build up enormous amounts of space charge.
+We close the TPC after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- of space to prevent the back-flow of ions from the avalanche region of that event.
+Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
+Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations.
+In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency.
+In the off configuration the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field between wires is great enough to block incoming electrons; opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to connect and equilibrate their charges back to the reference voltage.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{wires_open.pdf}
+\caption{Cartoon graphic of the wire plane structures}
+\label{fig:wires_open}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{gg_onoff.pdf}
+\caption{On and off configurations of the gating grid.}
+\label{fig:gg_onoff}
+\end{figure}
+
+The anode wires are typically biased to \SI{e3}{\volt} and are very thin, about \SI{20}{\micro\metre} in diameter.
+This creates a very high electric field very close to the anode wire in which the electron gains significant kinetic energy producing electrons until it terminates on the anode wire.
+The amount of electrons it produces depends on the anode wire voltage and the gas properties.
+The absolute gas gain was not experimentally measured but was simulated by a Garfield simulation.
+During the experiment the anode wires were biased to two different voltages.
+We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
+Only two sections were biased with the lower voltage setting due to concerns of a high current issue.
+Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
+The distribution follows a Polya distribution as expected and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{gain.png}
+\caption{Number of electrons produced in a single avalanche on an anode wire.
+Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
+The expected Polya distribution fit is also given in yellow.}
+\label{fig:anodegain}
+\end{figure}
 
@@
 \subsection{Gas Properties}
-
-\begin{equation}
-\frac{d\vec{x}}{dt} = \frac{\mu}{1+(\omega\tau)^2}\Big(\vec{E} + \omega\tau\frac{\vec{E}\times\vec{B}}{|\vec{B}|}+\omega^2\tau^2\frac{\vec{E}\cdot\vec{B}}{|\vec{B}|^2}\vec{B}\Big)
-\label{eq:elecdrift}
-\end{equation}
+The gas used inside of the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume --so called P10--  gas, operated just under atmospheric pressure (1 atm).
+The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere.
+The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants.
+The water never exceed  ???
+ppm  and the oxygen level never exceeded ???
+ppm.
+Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}.
+Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
+The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.
 
@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{gain.png}
-\caption{Number of electrons produced in a single avalanche on an anode wire.
-Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
-The expected Polya distribution fit is also given in yellow.}
-\label{fig:anodegain}
-\end{figure}
+The general formula for the drift velocity, $d\vec{x}/dt$, of an electron in the presence of electric and magnetic fields, $\vec{E}$ and $\vec{B}$, can be expressed in the Langevin equation as,
+
+\begin{equation}
+\frac{d\vec{x}}{dt} = \frac{\mu}{1+(\omega\tau)^2}\Big(\vec{E} + \omega\tau\frac{\vec{E}\times\vec{B}}{|\vec{B}|}+\omega^2\tau^2\frac{\vec{E}\cdot\vec{B}}{|\vec{B}|^2}\vec{B}\Big),
+\label{eq:elecdrift}
+\end{equation}
+
+where $\mu$ is the signed drift velocity, $\omega$ is the cyclotron frequency, and $\tau$ is the collision parameter for a particular gas \cite{blumrol}.
+
+Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $sigma_l$ and $sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
 
@@
 
-Add table for gas diffusion
+%Add table for gas diffusion
 
 \section{Ancillary Detectors }
+
 \subsection{Kyoto Multiplicity Trigger}
+%kyoto array sets multiplicity trigger
+%scintillator bars
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{kyotoarray}
+\label{fig:kyoto}
+\caption{Kyoto array overview}
+\end{figure}
+
 \subsection{Krakow ?????? (KATANA)}
+%beam veto
+%beam trigger optional
+
+\subsection{Active Veto Array}
 

### 2020-05-05 macrosForThesis/GatingGrid/.#plotTransparency.C

@@ -1,0 +1,1 @@
+justin@localhost.localdomain.23118:1587981660

### 2020-05-07 chapters/ch_experiment.tex
Update

@@
 A full description of the energy loss distribution can be described in CITE HERE.
+
+\section{Time Projection Chambers}
+%explain in breif the basics of a time projection chamber
 
@@
 
-The voltage step down takes the high voltage of the field cage cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.
+The voltage step down takes the high voltage of the  cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.
 
@@
 \subsection{Field Cage}
+
+ADD info how to calculate cathode voltage from HW supply.
 
@@
 \subsection{Pad Plane}
+The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively making 12096 pads in total.
+There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads.
+The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
+There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector CITE HERE.
 Add figure of close up of pad plane
@@
 \subsection{Wire Planes}
+
+ \begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrrrr@{}}\toprule 
+\multicolumn{2}{c}{GET electronics settings}\\
+\midrule
+ Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane & Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
+ Anode  & Au-plated W   &  20  &  4  &  4   &  0.5  &  1460  \\
+ Ground & BeCu          &  75  &  1  &  8   &  1.2  &  0     \\ 
+ Gating & BeCu          &  75  &  1  &  14   &  1.2 &  -110$\pm$ 70\\ 
+ \bottomrule
+\end{tabular}
+\caption{Wire plane properties}
+\label{tb:wireplane}
+\end{table*}
+
 %Add figure of gating grid transparency closed and open configuration 
@@
 
-\begin{figure}
-\includegraphics[width=\textwidth]{wires_open.pdf}
-\caption{Cartoon graphic of the wire plane structures}
-\label{fig:wires_open}
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{wires_open.pdf} 
+        \caption{Wires open} \label{fig:wires_open}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{wires_closed.pdf} 
+        \caption{Wires closed.} \label{fig:wires_closed}
+    \end{subfigure}
+\label{fig:wires}
 \end{figure}
@@
 %scintillator bars
+The Kyoto Multiplicity Array consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
+The entire TPC structure was designed so that light charged particles could easily pass through  the field cage and  side walls of the TPC enclosure.
+In this way the number of tracks passing through the sides of the TPC could be measured by this array.
+In heavy ion collisions the more central a nuclear collision is, the more nucleons participate in the collision, resulting in more measured tracks.
+It is this correlation between the number of tracks and centrality of the collision that makes the Kyoto Array sensitive to the centrality of events.
+It is more likely that in very central collisions more tracks are going to the peripheral angles and measured by the Kyoto array.
+In the experiment the trigger selection criteria was $n_{Kyoto} > 4$, where $n_{Kyoto}$ is the total number of tracks measured by both arrays.
 
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{kyotoarray}
-\label{fig:kyoto}
-\caption{Kyoto array overview}
+\includegraphics[width=\textwidth]{TPCAux.png}
+\label{fig:aux}
+\caption{Exploded views of Kyoto and KATANA arrays.}
 \end{figure}
 
-\subsection{Krakow ?????? (KATANA)}
+\subsection{Krakow KATANA Veto and Multiplicity Array}
 %beam veto
 %beam trigger optional
+The Krakow KATANA array consists of 12 plastic scintillating bars mounted to the downstream wall of the TPC enclosure.
+Three of the 12 bars were thin and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this was a majority of the time.
+The 9 other bars operated as an additional multiplicity array similar to the Kyoto array.
+Since most of the particles are focus forward in a cone in the laboratory frame, it was found the condition on the Kyoto array was sufficient to trigger on central events; thus the KATANA array was used in primarily the beam veto mode.
+This was accomplished by positioning the array so that the expected position of the beam exiting the TPC would be centered on the three thin paddles.
+The threshold of the veto paddles were set so that the charge of a particle, $Z$, was $Z > 20$.
+This allowed the selection of very central events and we did not trigger on very peripheral or no collision events.
 
 \subsection{Active Veto Array}
+The beam was tuned by two sets of quadropole  magnets, STQ 1 and STQ2, so that the beam spot was focused on the TPC target location.
+Because of the inherent angular dispersion of the beam there were incoming beam events which significantly deviated from the target location.
+To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size.
+The threshold was set so that any beam particle which passed through any of the bars it would send a trigger signal to not trigger the system since the beam path would not be on target but on some other material inside the TPC.
 
 \section{Radio Isotope Beam Factory (RIBF) Facility }
-Cyclotron facility overview.
-Samurai line overview.
-Beam line element overview.
-Big rips beam PID. reference
+%Cyclotron facility overview.
+%Samurai line overview.
+%Beam line element overview.
+%Big rips beam PID. reference 
+The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
+The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), reaching a primary beam energy of \SI{345}{\MeVA}.
 
-\section{S$\pi$RIT at RIBF}
-Picture of setup.
-
-SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
-\begin{figure}[H]
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{SAMURAI-beamline.png}
 \caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
-\label{fig:sambl}
+\label{fig:samuraiBeamLine}
 \end{figure}
+
+After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
+These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
+This is accomplished through several dipole magnets, slits, and wedge degraders.
+The resulting secondary beam is not pure and the purity depends on the primary beam and target beam desired.
+
+In these set of experiments several beams were produced with varying intensities and purities.
+Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns.
+
+ \begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrr@{}}\toprule 
+ Primary Beam & Secondary Beam & Energy at mid target \si{\MeVA} & Intensity \si{\kilo\hertz} & Purity (\%) \\ [0.5ex] 
+ \midrule
+ ${}^{238}$U   & ${}^{132}$Sn   &  269.2  &  9.5  &  54   \\
+ ${}^{238}$U   & ${}^{124}$Sn   &  270.3  &  9.1  &  10  \\
+ ${}^{124}$Xe  & ${}^{112}$Sn   &  270.4  &  7.6  &  48  \\
+ ${}^{124}$Xe  & ${}^{108}$Sn   &  269.3  &  7.5  &  52   \\
+ \bottomrule
+\end{tabular}
+\caption{Primary and secondary beam properties produced in the \spirit TPC experimental campaigns. }
+\label{tb:beams}
+\end{table*}
 
@@
 
+The \spirit TPC was designed to fit exactly into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line.
+Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration.
+Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
+This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap.
+The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces.
+These bolts secure the vacuum chamber to the magnet which is not practically removable; though the inside of the magnet was not operated under vacuum.
+This required an extensive rail system and support frame to slowly slide in the TPC over the bolts, finally raising the TPC several \si{\centi\metre} to the final height.
+
+The height of the TPC was roughly aligned with a self-leveling laser system to match the center of target with the center of the beam line.
+Once the TPC was adjusted to the final location, the position of the TPC was measured in fine detail with a photogrametry system CITE HERE.
+Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system.
+Using the commercial software provided the set of different camera perspectives reconstruct a point cloud of all the targets into 3-dimensional coordinates.
+Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC -- and several of its internal components-- relative to the magnet frame.
+The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed CITE HERE or Show data.
+
+Maybe put a position table summary here of the TPC position and definition of the coordinates system in the TPC frame and the Magnet frame
+
+\begin{figure}
+\includegraphics[width=\textwidth]{perspective.png}
+\caption{Drawing of the experimental setup with the TPC inside of the SAMURAI magnet at $\ang{0}$ configuration.}
+\label{fig:experiment}
+\end{figure}
+
+\section{Data Acquisition (DAQ) }
+The Data AcQuisition (DAQ) consisted of three different systems.
+The RIBFDAQ system served as the master DAQ for the BigRIPS beam identificaiton DAQ, the TPC DAQ, the NeuLAND neutron wall DAQ, and the Kyoto Array DAQ systems.
+The TPC DAQ was handled by the NARVAL framework to readout the GET electronics for the \spirit TPC.
+A General Trigger Operator (GTO) trigger was supplied to each DAQ synchronizing the subsystems.
+
 \section{Trigger Condition}
-How it was made using kyoto krakow
+Signals from all of the auxiliary detectors were combined into several logic combinations to form a trigger logic for triggering the data acquisition  (DAQ) to record data.
+An upstream scintillating bar formed the start counter signal, triggering on any beam coming down the beam line.
+The active veto will trigger for any beam that is incident off the target location.
+The KATANA veto produces a signal if the beam passed through the TPC un-reacted, causing no nuclear collision; this produces a veto signal with a width of \SI{4}{\micro\second} which is the approximate time it takes for the beam to drift and clear the field cage volume.
+The Kyoto multiplicity trigger produces a signal when the total number of tracks passing through both Kyoto arrays are greater than 4.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{KatanaLogic.png}
+\caption{KATANA trigger box logic.}
+\label{fig:katanaLogic}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{TriggerLogic.png} 
+\caption{Master trigger logic.}
+\label{fig:trigLogic} 
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{DataTrigger1.png} 
+        \caption{${}^{124}$Xe primary beam trigger.} \label{fig:dataTrigger1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{DataTrigger2.png} 
+        \caption{${}^{238}$U primary beam trigger.} \label{fig:dataTrigger2}
+    \end{subfigure}
+\label{fig:datatrigger}
+\end{figure}
+
+There were several special trigger considerations when we built the trigger for the TPC.
+We required that the gating grid be opened fast to not miss any signal; as soon as there was a condition satisfying the Start Counter, Kyoto Multiplicity, the DAQ was not busy, and there was not a KATANA Veto signal.
+This was referred to as the Fast Trigger.
+If the KATANA trigger box is not satisfied --described later-- this will trigger a Fast Clear signal which will not trigger the DAQ and will quickly close the gating grid.
+Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.
+
+The master trigger for the DAQ was different for each primary beam as the experiment got progressively better.
+During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
+In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating grid trigger CITE HERE JONS THESIS.
+Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.
+
+Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal.
+The KATANA trigger, Kyoto Muliplicity, and Start Counter together trigger the DAQ.
+
+ Where as Fig.~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
+Here the KATANA trigger and the SC SUM???
+togethere trigger the DAQ.
+
+ It is worth mentioning how the busy signals for the experiment were handled.
+The DAQ system itself produces a busy signal which was combined with the busy signals from either opening or closing the gating grid.
+When opening the gating grid it is assumed the full volume of the TPC will be read out and therefore  a \SI{11}{\micro\second} gate is produced; which is slightly more than the time it takes for all the electrons to drift in the field cage.
+In the case were the gating grid should be fast closed, either due to the fast clear circuit or the end of the TPC measurement, a \SI{5}{\micro\second} gate is produced to allow for the gating grid to settle to a closed configuration and clear the drift volume of any residual electrons from the beam.
+Both of these gates are included with the DAQ in an OR configuration which makes the overall busy signal.
 

### 2020-05-07 macrosForThesis/GatingGrid/.#plotTransparency.C

@@ -1,1 +1,0 @@
-justin@localhost.localdomain.23118:1587981660

### 2020-05-08 chapters/ch_experiment.tex
Added GG information

@@
 \chapter{Experiment}
+
+\section{Operational Principles of Time Projection Chambers}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{tpcPrinciple.pdf}
+\label{fig:tpcPrinciple}
+\caption{Operation principle of the TPC}
+\end{figure}
+%explain in breif the basics of a time projection chamber 
+Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions.
+Here we will outline the physical principles involved in the TPC measurement.
+A detector gas is housed in a field cage, which also sets up a constant electric field.
+As charged particles pass through the gas, electron-ion pairs are separated from by passing tracks, the electrons are accelerated by the electric field and the ions move in the opposite direction.
+Since the mean free path of the electrons is very small they quickly collide with a gas molecule slowing down the electron which goes on to repeat the cycle.
+This microscopic behavior manifests into the electron appearing to drift at a constant velocity.
+These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field liberating more electron-ion pairs from the gas creating an avalanche.
+The avalanche electrons finally terminate either on the anode wire while the ions from the avalanche move slowly away from the anode wires creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.
+
+Two of the 3 coordinates are determined from the charge distribution on the pad plane.
+The third dimension comes from projecting the electrons back in time utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics.
+The momentum information is extracted from measuring the radius of curvature of the track and the constant magnetic field value.
+The energy loss (<dE/dx>) is measured by a segmented charge sensitive pads on the pad-plane which are connected to the readout electronics.
+Particle identification is achieve through knowledge of the rigidity and the <dE/dx> of the tracks as will be discussed in latter sections.
+In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.
+
+\section{S$\pi$RIT TPC Overview}
+%Add Overview image with labels
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{exploded.png}
+\caption{Overview of the \spirit TPC}
+\label{fig:tpcExplode}
+\end{figure}
+
+The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiment.
+The TPC is enclosure is an aluminum angle iron skelleton with thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
+The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC.
+More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.
+
+A target mechanism allowed for up to 5 fixed targets to be mounted with the ability for changing the targets during the experiment without needing to open or move the TPC.
+The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate.
+The pad plane and wire plane structures are also mounted to the bottom of the top plate with the electronics are mounted on the top of the top plate.
+Several aluminum ribs were mounted to the top.
+Holes on the top plate allowed the electronics to be connected to the surface mount pads on the pad-plane.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{\spirit TPC Overview} \\
+ \midrule
+Pad plane area & 1.3 m x .9 m\\
+Pad size       & 1.2 cm x .8 cm \\
+Number of pads & 12096 (112 x 108) \\
+Gas composition& 90\% Ar + 10\% CH${}_4$ (1 atm)  \\
+Multiplicity limit & 200  \\
+dE/dx range        & Z=1-3, $\pi$, p, d, t, He, Li \\
+Drift length       & 50 cm \\
+\bottomrule
+\end{tabular}
+\caption{An overview of the properties of the \spirit TPC}
+\label{tb:spiritoverview}
+\end{table*}
+
+\subsection{Enclosure}
+The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame.
+All walls are constructed of a aluminum frame with think sheet metal.
+All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much.
+This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC.
+The enclosure itself is made to be gas tight with respects to the outside and the field cage.
+This is to allow for the possibility to run a different gas inside the enclosure than the field cage.
+Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.
+
+\subsection{Voltage Step Down}
+The gap between the field cage's cathode and the ground of the enclosure is quite small.
+To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
+There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time.
+The first ring is the same voltage as the cathode and the last ring is connected to ground.
+All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.
+
+\subsection{Field Cage}
+
+The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
+It was designed to hang from the top plate and therefore needed to be of a lightweight construction.
+Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
+Therefore instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame with \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips.
+The cathode was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{fc_overview.png}
+\label{fig:fc_overview}
+\caption{FC overview}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{fc_overview2.png}
+\label{fig:fc_overview2}
+\caption{FC overview 2}
+\end{figure}
+
+The field cage was constructed from several panels of printed circuit boards (PCBs).
+The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires CITE HERE.
+The halogen free material chosen was Cryogenic-G10.
+We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
+While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.
+The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.
+
+\begin{equation}
+V_n = V_{cath} \frac{R_p + (50 - n)R}{49\cdot R + R_p}
+\label{eq:FCstrip}
+\end{equation}
+
+\begin{equation}
+R_p = 49 \cdot R  \left(\frac{ y_{g.g.} - y_{cath} }{ y_{TP} - y_{cath} \frac{V_{cath} - V_{gg}}{V_{cath}} }- 1 \right)
+\label{eq:TP_resistor}
+\end{equation}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.6]{FC_schematic.pdf}
+\caption{Schematic of the electric connections relevant to the Field Cage system.}
+\label{fig:FC_schematic}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{TPC_schematic.pdf}
+\caption{Schematic of the TPC system}
+\label{fig:TPC_schematic}
+\end{figure}
+
+The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
+The cathode voltage $V_{cath}$ can be calculated from the schematic of the TPC system in Fig.~\ref{fig:TPC_schematic}, as
+
+\begin{equation}
+V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( \frac{1}{245 + R_{eff}} + \frac{1}{700} \right)^{-1} } },
+\end{equation}
+
+where $R_{eff} = \left( 1/20 + 1/R_{TP2} \right)^{-1}$, and $V_{HV}$ is the high voltage supply; all resistor values are given in \si{\mega\ohm}.
+
+Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls.
+The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail.
+The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the next inside strip and the first outside strip.
+The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field.
+The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC.
+Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips.
+The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter.
+
+\subsection{Wire Planes}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{gain.png}
+\caption{Number of electrons produced in a single avalanche on an anode wire.
+Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
+The expected Polya distribution fit is also given in yellow.}
+\label{fig:anodegain}
+\end{figure}
+
+ \begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rrrrrrr@{}}\toprule 
+\multicolumn{2}{c}{GET electronics settings}\\
+\midrule
+ Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane & Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
+ Anode  & Au-plated W   &  20  &  4  &  4   &  0.5  &  1460  \\
+ Ground & BeCu          &  75  &  1  &  8   &  1.2  &  0     \\ 
+ Gating & BeCu          &  75  &  1  &  14   &  1.2 &  -110$\pm$ 70\\ 
+ \bottomrule
+\end{tabular}
+\caption{Wire plane properties}
+\label{tb:wireplane}
+\end{table*}
+
+%Add figure of gating grid transparency closed and open configuration 
+There are three wire planes that sit underneath the pad-plane.
+The wire plane closest to the pad-plane (\SI{4}{\milli\metre}), is the anode wire.
+The next plane (\SI{12}{\milli\metre}), is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}), is the gating grid.
+The gating grid is the first plane that electrons meet as they drift upward towards the anode plane.
+The gating grid is operated as a gate either allowing electrons and ions through or blocking them entirely.
+The ground plane shields the inside volume of the TPC from the high voltages of the anode wires where the avalanche process of the electron takes place.
+
+The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
+If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
+
+In the open configuration the gating grid is transparent to electrons coming from the TPC volume and also allows for ions to move from the avalanche region into the TPC volume.
+It functions as the gate of the TPC where we only open the gating grid when an event meets the selection trigger criteria; minimizing the excessive exposure to the un-reacted beams which would quickly build up enormous amounts of space charge.
+We close the TPC after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- of space to prevent the back-flow of ions from the avalanche region of that event.
+Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
+Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations.
+In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency.
+In the off configuration the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field between wires is great enough to block incoming electrons; opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to connect and equilibrate their charges back to the reference voltage.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{wires_open.pdf} 
+        \caption{Wires open} \label{fig:wires_open}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.42\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{wires_closed.pdf} 
+        \caption{Wires closed.} \label{fig:wires_closed}
+    \end{subfigure}
+\label{fig:wires}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{gg_onoff.pdf}
+\caption{On and off configurations of the gating grid.}
+\label{fig:gg_onoff}
+\end{figure}
+
+The anode wires are typically biased to \SI{e3}{\volt} and are very thin, about \SI{20}{\micro\metre} in diameter.
+This creates a very high electric field very close to the anode wire in which the electron gains significant kinetic energy producing electrons until it terminates on the anode wire.
+The amount of electrons it produces depends on the anode wire voltage and the gas properties.
+The absolute gas gain was not experimentally measured but was simulated by a Garfield simulation.
+During the experiment the anode wires were biased to two different voltages.
+We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
+Only two sections were biased with the lower voltage setting due to concerns of a high current issue.
+Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
+The distribution follows a Polya distribution as expected and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
+
+\subsection{Pad Plane}
+The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively making 12096 pads in total.
+There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads.
+The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
+There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector CITE HERE.
+Add figure of close up of pad plane
+
+\subsection{Electronics}
+
+Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.
+Short cables transmit the signals from the pads to the inputs of the AGET chips.
+Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz.
+Four AGET chips are mounted on one AsAd (ASIC and ADC) motherboard.
+The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range, and the ADCs on each AsAd board provides 12 bit resolution.
+The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns.
+In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets).
+The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.
+The variations in the electronics were calibrated by measuring the response of each channel to a injected reference pulse, covering the full dynamic range of each channel.
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}rr@{}}\toprule 
+\multicolumn{2}{c}{GET electronics settings}\\
+\midrule
+ADC bit range       & 14 bits \\
+Sampling frequency  & 1-100 MHz \\
+Dynamic range       & .12, .24, 1.0, 10pC \\
+Peaking time        & 69,117,232,501,720,1014 ns \\
+Time bucket range   & 512\\
+\bottomrule
+\end{tabular}
+\caption{Summary of range of GET electronics settings. }
+\label{tb:getoverview}
+\end{table*}
 
@@
 There is a large variation in energy loss around this mean value, with a long high energy loss tail.
+
+The statistical variation of energy loss in a material was described by Landau \cite{landau} and later better described by Shulek \cite{shulek} and Bichsel \cite{bichsel1}.
+In both approximations it is described by a most probable energy loss value with a long, high-energy loss tail.
+Because of this long tail, for a finite set of energy loss measurements along a given track, the fluctuation of the mean value energy loss is a very unreliable observable.
+The most probable energy loss is the most desirable observable either through fitting of the observed distribution or through the truncated mean method.
+The truncated mean is the average mean value calculated after throwing away the top fraction of the highest energy loss entries, where the fraction used in this thesis was 30\%.
+This approximates the most probable value without performing a fit to a known distribution.
+There are
+
+%Suppose $E$ represents the set of $N$ energy loss points measured, sorted from lowest to highest energy loss.
+The $j^{th}$ index marks the position where  $i > j$ entries represent the highest value of energy loss measurements, where their fraction of the total set is expressed as, $f_c = \sum_{i>j} E_i/ \sum_i E_i$.
+The truncated mean is expressed as the mean value of the remaining energy loss measurements, throwing away the top $f_c$ fraction, expressed as,
+
+%\begin{equation}
+%<dE/dx>_t = \frac{\sum_{i < j} E_i}{N}
+%\label{eq:truncatedM}
+%\end{equation}
+
  A full description of the energy loss distribution can be described in CITE HERE.
 
-\section{Time Projection Chambers}
-%explain in breif the basics of a time projection chamber
+\subsection{Gas Properties}
+The gas contained by the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure (1 atm).
+The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere.
+The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants.
+The water never exceed  ???
+ppm  and the oxygen level never exceeded ???
+ppm.
+Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}.
+Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
+The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{driftvel.png}
+\caption{Drift velocity of electrons in P10 gas.}
+\label{fig:driftvel}
+\end{figure}
+
+The general formula for the drift velocity, $d\vec{x}/dt$, of an electron in the presence of electric and magnetic fields, $\vec{E}$ and $\vec{B}$, can be expressed in the Langevin equation as,
+
+\begin{equation}
+\frac{d\vec{x}}{dt} = \frac{\mu}{1+(\omega\tau)^2}\Big(\vec{E} + \omega\tau\frac{\vec{E}\times\vec{B}}{|\vec{B}|}+\omega^2\tau^2\frac{\vec{E}\cdot\vec{B}}{|\vec{B}|^2}\vec{B}\Big),
+\label{eq:elecdrift}
+\end{equation}
+
+where $\mu$ is the signed drift velocity, $\omega$ is the cyclotron frequency, and $\tau$ is the collision parameter for a particular gas \cite{blumrol}.
+
+Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $sigma_l$ and $sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
+
+\begin{table}[!htp] % not just 'h!'
+\centering % not a center environment
+\begin{tabular}{
+  @{}
+  l
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=1.2]
+  S[table-format=5.2]
+  S[table-format=5.2]
+  @{}
+}
+\toprule
+Gas properties &
+ {$\sigma_{t}$} &
+ {$\sigma_{l}$} &
+ {$v_{d}$} &
+ {$v_{i}$}  &
+ {$G_{h}$} &
+ {$G_{l}$} \\
+&
+  {($\si{\centi\meter}^{-1/2}$)} &
+  {($\si{\centi\meter}^{-1/2}$)} &
+  {(\si{\centi\meter\per\micro\second})} &
+ {(\si{\centi\meter\per\micro\second})} \\
+
+\midrule
+\phantom{abc}   &.024   &.034  &5.43  &  \num{2.05e-4} &  903   &150     \\
+\bottomrule
+\end{tabular}
+
+\caption{}
+\label{tb:gasprop}
+\end{table}
+
+%Add table for gas diffusion
 
@@
 
-\section{S$\pi$RIT TPC Overview}
-%Add Overview image with labels
-The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (S$\pi$RI TPC) is a Multi-Wire Proportional Counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiments.
-The TPC is enclosed in a thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach the arrays of scintillating bars detectors on the sides and downstream of the TPC.
-The S$\pi$RI TPC was developed to fit inside the Samurai dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC.
-More detail and specifications of the Samurai dipole magnet are given in \cite{samurai}.
-
-A target ladder allowed for up to 5 fixed targets to be mounted.
-A ACME worm gear allowed for the x-axis motion for changing the targets during the experiment, without needing to open or move the TPC.
-The motion of this worm gear was translated through the target motion feed-through by several brass gears and non-magnetic gear boxes.
-The motion of the target ladder could be controlled by hand or by operation of the drill.
-The targets were mounted on stand-offs on the target ladder and also had z-axis motion.
-This allowed for the targets to be positioned as close as possible to the thin window of the field cage, maximizing the geometric acceptance.
-
-The electronics were mounted to the aluminum top plate.
-Several aluminum ribs were mounted to the top of the plate to bring structural rigidity.
-A flatness within 150$\mu$m was achieved across the whole top plate as measured by a laser position system.
-The charge sensitive pads of the pad plane were etched into several circuit boards were recessed and glued to the bottom portion of the top plate.
-Vias through the pad plane circuit boards brought the signal traces from the pads to surface mount pads on the other side of the boards.
-Several holes cut through the top plate allowed for the interface cables of the electronics to be connected to these surface pads.
-
-Just below the pad plane were a set of three wire planes; the gating grid, ground, and anode wire planes.
-The detailed function of these wires will be explained later but they served to separate the boundary between the drift volume and the avalanche volume.
-
-The front and sides of the field cage were assembled from 8 independent rigid circuit boards.
-The downstream window served as the downstream wall of the field cage all though it was was a large polycarbonate frame in which a removable kapton window could be installed.
-This thin window minimized the scattering of exiting neutrons and charged particles for downstream detectors.
-
-The voltage step down takes the high voltage of the  cathode and steps the voltage down through a set of copper rings and a resistor chain, minimizing the chance for sparking.
-
-\begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}rr@{}}\toprule 
-\multicolumn{2}{c}{\spirit TPC Overview} \\
- \midrule
-Pad plane area & 1.3 m x .9 m\\
-Pad size       & 1.2 cm x .8 cm \\
-Number of pads & 12096 (112 x 108) \\
-Gas composition& 90\% Ar + 10\% CH${}_4$  \\
-Multiplicity limit & 200  \\
-dE/dx range        & Z=1-8, $\pi$, p,d,t,He,Li-O \\
-Drift length       & 50 cm \\
-\bottomrule
-\end{tabular}
-\caption{An overview of the properties of the \spirit TPC}
-\label{tb:spiritoverview}
-\end{table*}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{exploded.png}
-\caption{Overview of the \spirit TPC}
-\label{fig:tpcExplode}
-\end{figure}
-
-\subsection{Enclosure}
-The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame.
-All walls are constructed of a aluminum frame with think sheet metal.
-All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much.
-This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC.
-The enclosure itself is made to be gas tight with respects to the outside and the field cage.
-This is to allow for the possibility to run a different gas inside the enclosure than the field cage.
-Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.
-
-\subsection{Voltage Step Down}
-The gap between the field cage's cathode and the ground of the enclosure is quite small.
-To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
-There were 7 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt}.
-
-\subsection{Field Cage}
-
-ADD info how to calculate cathode voltage from HW supply.
-
-The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
-It was designed to hang from the top plate and therefore needed to be of a lightweight construction.
-Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
-Therefore instead of a downstream wall, a large thin exit window was constructed.
-The cathode was constructed of an aluminum honeycomb laminate.
-Two sheets of ???
-aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{fc_overview.pdf}
-\label{fig:fc_overview}
-\caption{FC overview}
-\end{figure}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{fc_overview2.pdf}
-\label{fig:fc_overview2}
-\caption{FC overview 2}
-\end{figure}
-
-The field cage was constructed from several panels of printed circuit boards (PCBs).
-The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires [CITE].
-The halogen free material chosen was Cryogenic-G10.
-We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
-While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.
-
-The front of the field cage was made of two PCBs and each side was constructed of three PCBs.
-The exit window was a 10$\mu$m thick Kapton window with evaporated aluminum strips on the inside and outside; mounted on a polycarbonate frame.
-
-\begin{equation}
-V_n = V_{cath} \frac{R_p + (50 - n)R}{49\cdot R + R_p}
-\label{eq:FCstrip}
-\end{equation}
-
-\begin{equation}
-R_p = 49 \cdot R  \left(\frac{ y_{g.g.} - y_{cath} }{ y_{TP} - y_{cath} \frac{V_{cath} - V_{gg}}{V_{cath}} }- 1 \right)
-\label{eq:TP_resistor}
-\end{equation}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{FC_schematic.pdf}
-\label{fig:FC_schematic}
-\caption{Schematic of the electric connections relevant to the Field Cage system.}
-\end{figure}
-\subsection{Voltage Step Down}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{TPC_schematic.pdf}
-\caption{Schematic of the TPC system}
-\label{fig:TPC_schematic}
-\end{figure}
-
-Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls.
-The basic idea of the field cage is to set up a constant electric field in the field cage volume to allow for a constant drift, in the presence of gas, for electrons.
-To accomplish this the field cage is composed of a set of copper strips which are set at a defined voltage by a resistor chain for a fixed distance between strips; setting up a constant electric field.
-
-The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
-The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail.
-The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor, $R$, to the next inside strip and the first outside strip.
-The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field.
-The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC.
-Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips.
-The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor with the option to place an additional resistor in parallel to tune the voltage of the top perimeter.
-
-\subsection{Pad Plane}
-The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively making 12096 pads in total.
-There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads.
-The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
-There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector CITE HERE.
-Add figure of close up of pad plane
-
-\subsection{Wire Planes}
-
- \begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}rrrrrrr@{}}\toprule 
-\multicolumn{2}{c}{GET electronics settings}\\
-\midrule
- Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane & Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
- Anode  & Au-plated W   &  20  &  4  &  4   &  0.5  &  1460  \\
- Ground & BeCu          &  75  &  1  &  8   &  1.2  &  0     \\ 
- Gating & BeCu          &  75  &  1  &  14   &  1.2 &  -110$\pm$ 70\\ 
- \bottomrule
-\end{tabular}
-\caption{Wire plane properties}
-\label{tb:wireplane}
-\end{table*}
-
-%Add figure of gating grid transparency closed and open configuration 
-There are three wire planes that sit underneath the pad-plane.
-The wire plane closest to the pad-plane (\SI{4}{\milli\metre}), is the anode wire.
-The next plane (\SI{12}{\milli\metre}), is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}), is the gating grid.
-The gating grid is the first plane that electrons meet as they drift upward towards the anode plane.
-The gating grid is operated as a gate either allowing electrons and ions through or blocking them entirely.
-The ground plane shields the inside volume of the TPC from the high voltages of the anode wires where the avalanche process of the electron takes place.
-
-The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
-If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
-
-In the open configuration the gating grid is transparent to electrons coming from the TPC volume and also allows for ions to move from the avalanche region into the TPC volume.
-It functions as the gate of the TPC where we only open the gating grid when an event meets the selection trigger criteria; minimizing the excessive exposure to the un-reacted beams which would quickly build up enormous amounts of space charge.
-We close the TPC after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- of space to prevent the back-flow of ions from the avalanche region of that event.
-Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
-Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations.
-In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency.
-In the off configuration the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field between wires is great enough to block incoming electrons; opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to connect and equilibrate their charges back to the reference voltage.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{wires_open.pdf} 
-        \caption{Wires open} \label{fig:wires_open}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{wires_closed.pdf} 
-        \caption{Wires closed.} \label{fig:wires_closed}
-    \end{subfigure}
-\label{fig:wires}
-\end{figure}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{gg_onoff.pdf}
-\caption{On and off configurations of the gating grid.}
-\label{fig:gg_onoff}
-\end{figure}
-
-The anode wires are typically biased to \SI{e3}{\volt} and are very thin, about \SI{20}{\micro\metre} in diameter.
-This creates a very high electric field very close to the anode wire in which the electron gains significant kinetic energy producing electrons until it terminates on the anode wire.
-The amount of electrons it produces depends on the anode wire voltage and the gas properties.
-The absolute gas gain was not experimentally measured but was simulated by a Garfield simulation.
-During the experiment the anode wires were biased to two different voltages.
-We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
-Only two sections were biased with the lower voltage setting due to concerns of a high current issue.
-Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
-The distribution follows a Polya distribution as expected and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
-
-\begin{figure}[!htb]
-\includegraphics[width=\linewidth]{gain.png}
-\caption{Number of electrons produced in a single avalanche on an anode wire.
-Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
-The expected Polya distribution fit is also given in yellow.}
-\label{fig:anodegain}
-\end{figure}
-
-\subsection{Electronics}
-
-Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.
-Short cables transmit the signals from the pads to the inputs of the AGET chips.
-Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz.
-Four AGET chips are mounted on one AsAd (ASIC and ADC) motherboard.
-The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range, and the ADCs on each AsAd board provides 12 bit resolution.
-The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns.
-In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets).
-The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.
-The variations in the electronics were calibrated by measuring the response of each channel to a injected reference pulse, covering the full dynamic range of each channel.
-
-\begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}rr@{}}\toprule 
-\multicolumn{2}{c}{GET electronics settings}\\
-\midrule
-ADC bit range       & 14 bits \\
-Sampling frequency  & 1-100 MHz \\
-Dynamic range       & .12, .24, 1.0, 10pC \\
-Peaking time        & 69,117,232,501,720,1014 ns \\
-Time bucket range   & 512\\
-\bottomrule
-\end{tabular}
-\caption{Summary of range of GET electronics settings. }
-\label{tb:getoverview}
-\end{table*}
-
 \subsection{Considerations when constructing a TPC}
@@
 Two volumes of gas.
-
-\subsection{Gas Properties}
-The gas used inside of the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume --so called P10--  gas, operated just under atmospheric pressure (1 atm).
-The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere.
-The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants.
-The water never exceed  ???
-ppm  and the oxygen level never exceeded ???
-ppm.
-Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}.
-Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
-The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.
-
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{driftvel.png}
-\caption{Drift velocity of electrons in P10 gas.}
-\label{fig:driftvel}
-\end{figure}
-
-The general formula for the drift velocity, $d\vec{x}/dt$, of an electron in the presence of electric and magnetic fields, $\vec{E}$ and $\vec{B}$, can be expressed in the Langevin equation as,
-
-\begin{equation}
-\frac{d\vec{x}}{dt} = \frac{\mu}{1+(\omega\tau)^2}\Big(\vec{E} + \omega\tau\frac{\vec{E}\times\vec{B}}{|\vec{B}|}+\omega^2\tau^2\frac{\vec{E}\cdot\vec{B}}{|\vec{B}|^2}\vec{B}\Big),
-\label{eq:elecdrift}
-\end{equation}
-
-where $\mu$ is the signed drift velocity, $\omega$ is the cyclotron frequency, and $\tau$ is the collision parameter for a particular gas \cite{blumrol}.
-
-Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $sigma_l$ and $sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
-
-\begin{table}[!htp] % not just 'h!'
-\centering % not a center environment
-\begin{tabular}{
-  @{}
-  l
-  S[table-format=1.2]
-  S[table-format=1.2]
-  S[table-format=1.2]
-  S[table-format=1.2]
-  S[table-format=5.2]
-  S[table-format=5.2]
-  @{}
-}
-\toprule
-Gas properties &
- {$\sigma_{t}$} &
- {$\sigma_{l}$} &
- {$v_{d}$} &
- {$v_{i}$}  &
- {$G_{h}$} &
- {$G_{l}$} \\
-&
-  {($\si{\centi\meter}^{-1/2}$)} &
-  {($\si{\centi\meter}^{-1/2}$)} &
-  {(\si{\centi\meter\per\micro\second})} &
- {(\si{\centi\meter\per\micro\second})} \\
-
-\midrule
-\phantom{abc}   &.024   &.034  &5.43  &  \num{2.05e-4} &  903   &150     \\
-\bottomrule
-\end{tabular}
-
-\caption{}
-\label{tb:gasprop}
-\end{table}
-
-%Add table for gas diffusion
 

### 2020-05-08 macrosForThesis/GatingGrid/plotTransparency_5T.C

@@ -1,0 +1,41 @@
+
+
+
+
+void plotTransparency_5T()
+{
+
+  //Spring Commision .5T magnetic field
+
+  
+  vector<double> data_v =   {-170.98, -164.98, -145.9, -136.54, -127.16, -119.14, -114.9, -110.06, -90.5, -79.6, -53.86, -39.34};
+  vector<double> data_trans =   {1, 0.999794, 0.998232, 0.995282, 0.980169, 0.954406, 0.942936, 0.932959, 0.910684, 0.901025, 0.873159, 0.844473};
+  vector<double> data_v_err = {0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0};
+  //  vector<double> data_trans_err = {0.033, 0.034, 0.047, 0.025, 0.029, 0.0327, 0.0303, 0.031, 0.029, 0.037, 0.0302, 0.073, 0.0348};
+  vector<double> data_trans_err = {0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0};
+
+
+  vector<double> garf_v =    {-180, -170, -160, -152.33, -142.55, -132.76, -124.38, -114.84, -103.27, -94.48, 0, 10, 20};
+  vector<double> garf_trans =    {0.9975, 0.99525, 0.99375, 0.9945, 0.9855, 0.97975, 0.96725, 0.95275, 0.93275, 0.929, 0.67025, 0.574,0};
+  vector<double> garf_v_err =  { .0, .0, .0, .0, .0, .0, .0, .0, .0, .0, .0, .0, .0};
+  vector<double> garf_trans_err =  {.02, .02, .02, .02, .02, .02, .02, .02, .02, .02, .02, .02, .02};
+
+  TGraphErrors *data_avg = new TGraphErrors(data_v.size(),data_v.data(),data_trans.data(),data_v_err.data(),data_trans_err.data());
+  TGraphErrors *garf_avg = new TGraphErrors(garf_v.size(),garf_v.data(),garf_trans.data(),garf_v_err.data(),garf_trans_err.data());
+
+
+  data_avg->SetMarkerColor(2);
+
+  data_avg->SetMarkerStyle(20);
+  garf_avg->SetMarkerStyle(20);
+
[…]

### 2020-05-14 chapters/ch_experiment.tex
New graphics for gg transparency

@@
 Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions.
-Here we will outline the physical principles involved in the TPC measurement.
-A detector gas is housed in a field cage, which also sets up a constant electric field.
-As charged particles pass through the gas, electron-ion pairs are separated from by passing tracks, the electrons are accelerated by the electric field and the ions move in the opposite direction.
-Since the mean free path of the electrons is very small they quickly collide with a gas molecule slowing down the electron which goes on to repeat the cycle.
-This microscopic behavior manifests into the electron appearing to drift at a constant velocity.
-These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field liberating more electron-ion pairs from the gas creating an avalanche.
-The avalanche electrons finally terminate either on the anode wire while the ions from the avalanche move slowly away from the anode wires creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.
+Here we will outline the physical principles involved in the TPC measurement and discuss the particular details of the TPC used in this thesis.
+Inside of at TPC the detector gas is housed in a field cage, which also sets up a constant electric field.
+As charged particles pass through the gas, electron-ion pairs from neutral gas molecules by passing tracks; the electrons are accelerated opposite to the electric field and the ions move in the opposite direction.
+Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules, slowing down the electron, which is then accelerated again repeating the cycle.
+This microscopic behavior manifests into a constant drift velocity when averaged over several gas collisions.
+These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field, liberating more electron-ion pairs from the gas creating an avalanche.
+The avalanche electrons finally terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires; creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.
 
 Two of the 3 coordinates are determined from the charge distribution on the pad plane.
-The third dimension comes from projecting the electrons back in time utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics.
-The momentum information is extracted from measuring the radius of curvature of the track and the constant magnetic field value.
-The energy loss (<dE/dx>) is measured by a segmented charge sensitive pads on the pad-plane which are connected to the readout electronics.
-Particle identification is achieve through knowledge of the rigidity and the <dE/dx> of the tracks as will be discussed in latter sections.
+The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics.
+The radius of curvature is related to the magnetic rigidity, and therefore the momentum of each track.
+The energy loss  deposited (<dE/dx>) is measured by the segmented charge sensitive pads on the pad-plane, which are connected to the readout electronics.
+Particle identification is achieved through since unique particles exist on unique rigidity and <dE/dx>  lines, which will be discussed in latter sections.
 In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.
@@
 
-The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in a fixed target experiment.
-The TPC is enclosure is an aluminum angle iron skelleton with thin aluminum sheet walls all around in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
-The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC.
+The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in fixed target experiments.
+The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls all around, in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
+The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}.
 More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.
 
-A target mechanism allowed for up to 5 fixed targets to be mounted with the ability for changing the targets during the experiment without needing to open or move the TPC.
-The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate.
-The pad plane and wire plane structures are also mounted to the bottom of the top plate with the electronics are mounted on the top of the top plate.
-Several aluminum ribs were mounted to the top.
-Holes on the top plate allowed the electronics to be connected to the surface mount pads on the pad-plane.
+A target mechanism allowed for up to 5 fixed targets to be mounted at anytime, with the ability to change targets on the outside of the TPC.
+The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal.
+The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate; several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a laser CITE HERE JONS THESIS.
+Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics.
+The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC
 
@@
 The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame.
-All walls are constructed of a aluminum frame with think sheet metal.
+All walls are constructed of a aluminum frame with thin sheet metal.
 All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much.
@@
 
-\subsection{Voltage Step Down}
-The gap between the field cage's cathode and the ground of the enclosure is quite small.
-To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
-There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time.
-The first ring is the same voltage as the cathode and the last ring is connected to ground.
-All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.
-
 \subsection{Field Cage}
 
-The field cage holds the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
+The field cage contains the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
 It was designed to hang from the top plate and therefore needed to be of a lightweight construction.
-Also the materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
-Therefore instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame with \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips.
-The cathode was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure for the cathode.
+The materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
 
@@
 The field cage was constructed from several panels of printed circuit boards (PCBs).
+The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.
 The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires CITE HERE.
-The halogen free material chosen was Cryogenic-G10.
-We built the TPC with the option to run explosive gases such hydrogen, thus we decided to have field cage an isolated volume from the rest of the TPC enclosure.
+The halogen free material used was Cryogenic-G10.
+The field cage is an isolated volume from the enclosure of the TPC for the option to run explosive gases such hydrogen in the field cage, where the enclosure can be filled with an insulating gas.
 While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.
-The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.
-
-\begin{equation}
-V_n = V_{cath} \frac{R_p + (50 - n)R}{49\cdot R + R_p}
-\label{eq:FCstrip}
-\end{equation}
-
-\begin{equation}
-R_p = 49 \cdot R  \left(\frac{ y_{g.g.} - y_{cath} }{ y_{TP} - y_{cath} \frac{V_{cath} - V_{gg}}{V_{cath}} }- 1 \right)
-\label{eq:TP_resistor}
-\end{equation}
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[scale=.6]{FC_schematic.pdf}
-\caption{Schematic of the electric connections relevant to the Field Cage system.}
-\label{fig:FC_schematic}
-\end{figure}
+Instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame bonded to a \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips.
+The PCB boards were epoxied into the cathode which was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure.
+On the other end the boards were epoxied into an aluminum top perimeter which also served as the last ring in the TPC.
+Together with the cathode bottom, the field cage proved to be a rigid lightweight structure.
+A lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC.
+Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate.
+The top plate could be removed and rotated with the field cage on without damaging any internal components.
 
@@
 
-The cathode itself is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
+The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
 The cathode voltage $V_{cath}$ can be calculated from the schematic of the TPC system in Fig.~\ref{fig:TPC_schematic}, as
-
 \begin{equation}
-V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( \frac{1}{245 + R_{eff}} + \frac{1}{700} \right)^{-1} } },
+V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
+\end{equation}
+where 
+\begin{equation}
+R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
+ \label{eq:Reff}
 \end{equation}
 
-where $R_{eff} = \left( 1/20 + 1/R_{TP2} \right)^{-1}$, and $V_{HV}$ is the high voltage supply; all resistor values are given in \si{\mega\ohm}.
+is the effective resistance of the last resistor, and $V_{HV}$ is the high voltage supply; all resistor values are given in \si{\mega\ohm}.
 
-Figure~\ref{fig:FC_schematic} shows a cartoon drawing of the electrode geometry of the field cage walls.
-The field cage is made of 50 inside copper strips and 49 outside copper strips; the strip thickness is exaggerated in the figure to show the detail.
-The first inside strip is connected to the cathode which is connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the next inside strip and the first outside strip.
-The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance setting up a constant electric field.
-The last strip of the field cage is composed of partially a small inner strip on the PCB board and the aluminum top perimeter piece which holds the field cage together with the top plate of the TPC.
-Together their thickness is equal to the \SI{6}{\milli\metre} strip width of the other strips.
-The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter.
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.6]{FC_schematic.pdf}
+\caption{Schematic of the electric connections relevant to the Field Cage system. The strip thickness is exaggerated in the figure to show the detail}
+\label{fig:FC_schematic}
+\end{figure}
+
+Figure~\ref{fig:FC_schematic} shows a cartoon schematic of the field cage walls.
+The field cage contains 50 inside copper strips and 49 outside copper strips.
+The strips on each board was connected to the adjacent strip on the next board; front, back, and side boards were connected by G10 corner pieces with conducting paint strips, creating rings around the whole field cage.
+The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip.
+The first outside strip is connected trough a via to the second inside strip, which continues on in this repeating fashion.
+The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance, setting up a constant electric field.
+The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the aluminum top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths.
+The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.
+
+The voltage on each strip, $V_n$, can be expressed as, 
+\begin{equation}
+V_n = V_{cath} \frac{R_p + (50 - n)R}{49\cdot R + R_p}
+\label{eq:FCstrip}
+\end{equation}
+
+where $n = 1$ represents the index of the first inside strip, and $n= 50$ represents the index of the last inside strip, which is the same as the top perimeter voltage.
+
+%Talk about the uniforminty of the field cage maybe
+%how far the wall was from the pad plane on the sides talk about the front
+
+\subsection{Voltage Step Down}
+The gap between the field cage's cathode and the ground of the enclosure is quite small.
+To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
+There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time.
+The first ring is the same voltage as the cathode and the last ring is connected to ground.
+All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.
 
 \subsection{Wire Planes}
+%Add figure of gating grid transparency closed and open configuration
 
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{gain.png}
-\caption{Number of electrons produced in a single avalanche on an anode wire.
-Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
-The expected Polya distribution fit is also given in yellow.}
-\label{fig:anodegain}
-\end{figure}
+There are three wire planes that are mounted underneath the pad-plane.
+The wire plane closest to the pad-plane (\SI{4}{\milli\metre}) are the anode wires.
+The next plane (\SI{12}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid.
+The gating grid is the first plane that electrons meet as they drift upward from the field cage volume towards the anode plane.
+The gating grid is operated as a gate either allowing electrons and ions through, or blocking them entirely.
+The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires where the avalanche process of the electron takes place.
+The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
+If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
 
- \begin{table*}\centering
+ \begin{table*}[!htb]
+ \centering
 \ra{1.3}
@@
 
-%Add figure of gating grid transparency closed and open configuration 
-There are three wire planes that sit underneath the pad-plane.
-The wire plane closest to the pad-plane (\SI{4}{\milli\metre}), is the anode wire.
-The next plane (\SI{12}{\milli\metre}), is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}), is the gating grid.
-The gating grid is the first plane that electrons meet as they drift upward towards the anode plane.
-The gating grid is operated as a gate either allowing electrons and ions through or blocking them entirely.
-The ground plane shields the inside volume of the TPC from the high voltages of the anode wires where the avalanche process of the electron takes place.
+In the open configuration the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume.
+Typically the gating grid is held in the closed configuration only opening when the data acquisition trigger criteria is met.
+This is to block the electrons which come from the un-reacted beam, which happens frequently, which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions and flood the volume of the field cage with space charge.
+We close the gating grid after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- to also prevent the back-flow of ions from the avalanche region from that event.
+Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
 
-The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
-If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
-
-In the open configuration the gating grid is transparent to electrons coming from the TPC volume and also allows for ions to move from the avalanche region into the TPC volume.
-It functions as the gate of the TPC where we only open the gating grid when an event meets the selection trigger criteria; minimizing the excessive exposure to the un-reacted beams which would quickly build up enormous amounts of space charge.
-We close the TPC after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- of space to prevent the back-flow of ions from the avalanche region of that event.
-Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
 Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations.
 In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency.
-In the off configuration the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field between wires is great enough to block incoming electrons; opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to connect and equilibrate their charges back to the reference voltage.
+In the off configuration, the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field produced by the voltage difference $2\Delta V$ between wires is great enough to block incoming electrons.
+Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the state of the reference voltage.
 
@@
 
+Both configurations of the gating grid were measured and simulated.
+To measure the electron transparency were all wires share the voltage $V_{avg}$, the anode wire was lowered to \SI{500}{\volt} and the beam was allowed to enter the field cage without any target put in.
+By lowering the voltage of the anode we could measure the large charge of the beam without saturating the electronics.
+The average charge deposited in the chamber could be measured as a function of $V_{avg}$; changing the top plane resistor appropriately according to Eq.~\ref{eq:TP_resistor}.
+Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field.
+Theoretically the most negative value represents 100\% electron transparency and was used as the reference run.
+The electron transparency, $T$, was defined as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run.
+Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation.
+The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.
+
+To measure the electron transparency as a function of the difference voltage $\Delta V$, the average voltage was first set to 100\% transparency, $V_{avg}=\SI{-171}{\volt}$, and the difference voltage was added or subtracted from alternating wires.
+Figure~\ref{fig:ggDeltaVTrans} shows the result of the simulation and experiment with and without the magnetic field.
+By introducing the magnetic field the required voltage to close the grid increases.
+In the experiment we selected the value of $\Delta V = \SI{65}{\volt}$ to ensure we were well within the region of 0\% transparency.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{averageTransparency.png} 
+        \caption{Electron transparency for the conditions of all wires are the same voltage.} \label{fig:ggAvgTrans}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{transparencyDeltaV.png} 
+        \caption{Electrons transparency for the mode where adjacent wires have a voltage difference of $2 \Delta V$.} \label{fig:ggDeltaVTrans}
+    \end{subfigure}
+\label{fig:ggTrans}
+\end{figure}
+
 The anode wires are typically biased to \SI{e3}{\volt} and are very thin, about \SI{20}{\micro\metre} in diameter.
-This creates a very high electric field very close to the anode wire in which the electron gains significant kinetic energy producing electrons until it terminates on the anode wire.
+This creates a very high electric field very close to the anode wire in which the electron gains kinetic energy producing electrons until it terminates on the anode wire or the pad plane.
 The amount of electrons it produces depends on the anode wire voltage and the gas properties.
@@
 Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
-The distribution follows a Polya distribution as expected and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
+The distribution follows the expected Polya distribution and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{gain.png}
+\caption{Number of electrons produced in a single avalanche on an anode wire.
+Two different voltages were simulated using Garfield++ at 1470 $V$ and 1214 $V$.
+The expected Polya distribution fit is also given in yellow.}
+\label{fig:anodegain}
+\end{figure}
+
+The wire plane and the cathode define the different electric field regions.
+The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the ground grid and the gating grid (Region 2); the avalanche region is defined as between the ground grid and the anode grid.
+The voltage of the gating grid is set to the point that allows for 100\% transparency for electrons.
+It is possible that the electric field in Region 1 and Region 2 can be matched but not required.
+The voltage of the top perimeter is set by adjusting the value of the last resistor ($R_{TP2}$) to ensure the electric field is constant through Region 1.
+We can imagine Region 1 is split into two virtual volumes, one defined between the cathode and the top perimeter, and one defined between the top perimeter and the gating grid.
+The magnitude of the electric field in the region between the top perimeter and the cathode, $E_1$, is defined as,
+
+\begin{equation}
+E_1 = \frac{V_{g.g.} - V_{tp}}{ y_{g.g.} - y_{tp} },
+\end{equation}
+where  $V_{g.g.}$, $V_{tp}$ , $y_{g.g.}$, and $y_{tp}$ are the voltage and vertical y-position of the gating grid and top perimeter respectively.
+The magnitude of the electric field in the region between the top perimeter and the cathode, $E_2$, is defined as,
+
+\begin{equation}
+E_2 = \frac{V_{tp} - V_{cath}}{ y_{tp} - y_{cath} },
+\end{equation}
+where  $V_{tp}$, $V_{cath}$ , $y_{tp}$, and $y_{cath}$ are the voltage and vertical y-position of the top perimeter and cathode respectively.
+The condition for a smooth electric field across the volume is given by $E_1 = E_2$.
+Substituting Eq.~\ref{eq:FCstrip} for $V_{tp}$ ($n=50$), we can solve for the effective resistance of the top perimeter $R_p$ as,
+
+\begin{equation}
+R_p = 49 \cdot R  \left(\frac{ y_{g.g.} - y_{cath} }{ y_{TP} - y_{cath} \frac{V_{cath} - V_{gg}}{V_{cath}} }- 1 \right),
+\label{eq:TP_resistor}
+\end{equation}
+
+where the relevant vertical dimensions are $y_{g.g.} - y_{cath} = \SI{497.3}{\milli\metre}$ and $y_{tp} - y_{cath} = \SI{490}{\milli\metre}$.
+The value of $R_{TP2}$ can be calculated from Eq.~\ref{eq:Reff} and Eq.~\ref{eq:TP_resistor}.
 

### 2020-05-14 macrosForThesis/GatingGrid/plotTransparency.C

@@ -1,10 +1,11 @@
+#include "/home/justin/mythesis/thesis_style.h"
 
-
-
+using namespace style;
 
 void plotTransparency()
 {
-
+  gStyle->SetOptStat(0);
+  
   //Fall 2015 0T magnetic field
 
   /*
@@ -28,16 +29,130 @@ void plotTransparency()
   TGraphErrors *garf_avg = new TGraphErrors(garf_v.size(),garf_v.data(),garf_trans.data(),garf_v_err.data(),garf_trans_err.data());
 
 
-  data_avg->SetMarkerColor(2);
+  vector<double> data_5_v =   {-170.98, -164.98, -145.9, -136.54, -127.16, -119.14, -114.9, -110.06, -90.5, -79.6, -53.86, -39.34};
+  vector<double> data_5_trans =   {1, 0.999794, 0.998232, 0.995282, 0.980169, 0.954406, 0.942936, 0.932959, 0.910684, 0.901025, 0.873159, 0.844473};
+  vector<double> data_5_v_err = {0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0};
+  //  vector<double> data_5_trans_err = {0.033, 0.034, 0.047, 0.025, 0.029, 0.0327, 0.0303, 0.031, 0.029, 0.037, 0.0302, 0.073, 0.0348};
+  vector<double> data_5_trans_err = {0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0};
 
-  data_avg->SetMarkerStyle(20);
-  garf_avg->SetMarkerStyle(20);
+
+  vector<double> garf_5_v =    {-180,-170,-160,-152.33,-142.55,-132.76,-124.38,-114.84,-103.27,-94.48,-100,-90,-80,-70,-60,-50,-40,-30,0,10,20};
+    vector<double> garf_5_trans =    {0.9975,0.99525,0.99375,0.9945,0.9855,0.97975,0.96725,0.95275,0.93275,0.929,0.929768,0.9205,0.91325,0.90925,0.911,0.893,0.87425,0.825,0.67025,0.574,0};
[…]

### 2020-05-14 macrosForThesis/GatingGrid/plotTransparency_vdiff.C

@@ -30,28 +30,87 @@ void plotTransparency_vdiff()
   TGraphErrors *garf_5 = new TGraphErrors(garf_v_5.size(),garf_v_5.data(),garf_trans_5.data());
 
 
-  data_0->SetMarkerColor(2);
-  data_5->SetMarkerColor(2);
+
+  TCanvas *c1 = new TCanvas("c1","c1",1000,800);
+  c1->SetTopMargin(.05);
+  c1->SetBottomMargin(.12);
+  c1->SetRightMargin(.05);
+  c1->SetLeftMargin(.12);
+
+  gPad->SetFrameLineWidth(10);
+  gStyle->SetLineWidth(10);
+    
+  garf_5->GetXaxis()->SetTitle("#DeltaV (V)");
+  garf_5->GetXaxis()->CenterTitle();
+  garf_5->GetYaxis()->SetTitle("e^{-} Transparency (%)");
+  garf_5->GetYaxis()->CenterTitle();
+
+  garf_5->GetXaxis()->SetTitleSize(.05);
+  garf_5->GetYaxis()->SetTitleSize(.05);
   
-  data_0->SetMarkerStyle(20);
-  garf_0->SetMarkerStyle(20);
+  garf_5->GetXaxis()->SetLabelSize(.05);
+  garf_5->GetYaxis()->SetLabelSize(.05);
+  garf_5->GetXaxis()->SetTitleOffset(1.1);
+  garf_5->GetYaxis()->SetTitleOffset(1.1);
 
-  data_5->SetMarkerStyle(22);
-  garf_5->SetMarkerStyle(22);
+  data_0->SetMarkerColor(kBlue - 4);
+  data_5->SetMarkerColor(kGreen - 4);
+  
+  garf_0->SetMarkerColor(kBlue - 4);
+  garf_5->SetMarkerColor(kGreen - 4);
+  
+  data_0->SetLineColor(kBlue - 4);
+  data_5->SetLineColor(kGreen - 4);
+  garf_0->SetLineColor(kBlue - 4);
+  garf_5->SetLineColor(kGreen - 4);
 
-  data_0->SetMarkerSize(1.);
-  garf_0->SetMarkerSize(1.);
+  data_0->SetLineWidth(6);
+  data_5->SetLineWidth(6);
 
-  data_5->SetMarkerSize(1.);
-  garf_5->SetMarkerSize(1.);
+  garf_0->SetLineWidth(10);
[…]

### 2020-05-14 chapters/ch_data_analysis.tex
Added GG information
quick update before check

@@
 For example, if we are clustering along the $x$ axis for a cluster, the $z$-position is set to the center of the pad in the $z$-direction and vice versa.
+
+\subsection{Experimental Pad Response Function}
+
+The correlations we introduced by only clustering along one direction do not play a significant role in the particle identification, but cause deviations from the expected Gatti distribution.
+Also, analytic PRFs only exist for classical multi-wire TPCs.
+For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
+
+As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as $\lambda_i = x_i-\bar{x}$.
+The PRF is simply defined as the charge fraction of each pad as a function of $\lambda$, as shown in Equation \ref{eq:prf}.
+
+\begin{equation}\label{eq:prf}
+PRF(\lambda_i) = \frac{q_i(\lambda_i)}{Q}
+\end{equation}
+
+Averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
+Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
+
+\begin{equation}\label{eq:gaus}
+PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
+\end{equation}
+
+\begin{figure}[ht!]s
+\begin{overpic}[width=\linewidth]{fig5.pdf}
+\put(61,55){\contour{white}{ PRF${}_{\mathrm{Gaus}}(\lambda)$ eq. \ref{eq:gaus}  }}
+\put(61,49){\contour{white}{ PRF${}_{\mathrm{Gatti}}(\lambda)$ eq. \ref{eq:gatti} }}
+\end{overpic}
+\caption{Experimental pad response function of many events for a crossing angle of $85^{\circ} < \theta \leq 90^{\circ}$.  }
+\label{fig:expprf}
+\end{figure}
+
+The shape of the PRF depends on the crossing angle of the track \cite{gatti}.
+Plotted in Fig.~\ref{fig:prfpimData} is the PRF of $\pi^-$ tracks vs.
+the crossing angle $\theta$.
+The PRF gets wider starting from $90^{\circ}$  and going to $45^{\circ}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
+Since we switch the clustering direction from $x$ to the $z$ direction at $45^{\circ}$, the opposite trend is seen where the PRF becomes narrower as the position resolution gets better going from $45^{\circ}$ to $0^{\circ}$.
+
+\begin{figure}[!htb]
+     \centering
+	 \includegraphics[width=\textwidth]{PRFs_data_wcut.png}
+     \caption{PRF response from $\pi^-$ data. }
+     \label{fig:prfpimData}
+\end{figure}
+
+\begin{figure}[ht!]
+\vspace{5mm}
+\includegraphics[width=\linewidth]{fig7}
+\caption{Parameters $N_{0}$ and $\sigma$ as a function of the crossing angle $\theta$ with the $4^{th}$ order polynomial fits.}
+\label{fig:normsigma}
+\end{figure}
+
+\begin{comment}
+\begin{table}
+\centering
+ \begin{tabular}{||c c c c c c||} 
+ \hline
+ Coefficient & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ [0.5ex] 
+ \hline\hline
+ $0 < \theta < 45$ & & & & &  \\ [.25ex]
+ \hline
+ $N_0$ & .897 & 5.766E-3 & -4.263E-4 & 7.444E-6 & 5.705E-8 \\ 
+ \hline
+ $\sigma$ & 5.496 & -3.920E-2 & 2.693E-3 & -5.208E-5 & 5.334E-7\\
+ \hline
+ $45 < \theta < 90$ & & & &  & \\ [.25ex]
+ \hline	
+ $N_0$ & 1.220 & -6.258E-2 & 1.608E-3 & -1.492E-5  & 4.654E-8 \\
+ \hline
+ $\sigma$ & 31.368 & -1.109 & 1.779E-2 & -1.336E-4 & 3.940E-7\\
+ \hline
+\end{tabular}
+\caption{Coefficients of the $4_th$ order polynomial fit to the Gaussian parameters $N_0$ and $\sigma$. The polynomial form is given as $c_0 + c_1 x + c_2 x^2 + c_3 x^3 + c_4 x^4$}
+\label{tb:coeff}
+\end{table}
+\end{comment}
+
+Fits were performed to the experimental data with  $5^{\circ}$ width bins from $0^{\circ} < \theta \leq 90^{\circ}$.
+The two parameters of the Gaussian fits are plotted versus $\theta$ in Fig.~\ref{fig:normsigma}; a $4^{th}$ order polynomial fit between these points allowed for interpolating between $\theta$.
 
@@
 
-\begin{figure}
-\includegraphics[scale=.01]{place}
-\caption{A summary of all the effects modeled in the TPC MC simulation.}
-\label{fig:place}
-\end{figure}
-
 \subsection{Drift Task}

### 2020-05-14 chapters/ch_experiment.tex
quick update before check

@@
 The radius of curvature is related to the magnetic rigidity, and therefore the momentum of each track.
-The energy loss  deposited (<dE/dx>) is measured by the segmented charge sensitive pads on the pad-plane, which are connected to the readout electronics.
-Particle identification is achieved through since unique particles exist on unique rigidity and <dE/dx>  lines, which will be discussed in latter sections.
+The energy loss  deposited ($\langle dE/dx\rangle$) is measured by the segmented charge sensitive pads on the pad-plane, which are connected to the readout electronics.
+Particle identification is achieved through since unique particles exist on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections.
 In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.
@@
 The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal.
-The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate; several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a laser CITE HERE JONS THESIS.
+The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate; several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a laser \cite{jon}.
 Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics.
@@
 The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.
-The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires CITE HERE.
+The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires \cite{tpcAging}.
 The halogen free material used was Cryogenic-G10.
@@
 
-The anode wires are typically biased to \SI{e3}{\volt} and are very thin, about \SI{20}{\micro\metre} in diameter.
-This creates a very high electric field very close to the anode wire in which the electron gains kinetic energy producing electrons until it terminates on the anode wire or the pad plane.
-The amount of electrons it produces depends on the anode wire voltage and the gas properties.
-The absolute gas gain was not experimentally measured but was simulated by a Garfield simulation.
+The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter.
+They are biased to high voltages around \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
+As the eletron drifts towards the anode wire, it gains kinetic energy knocking out more electron-ion pairs in the gas until it terminates on the anode wire or the pad plane.
+The amount of electrons produced depends on the anode wire voltage and the gas properties.
+The absolute gas gain was not experimentally measured but was calculated in a Garfield simulation.
 During the experiment the anode wires were biased to two different voltages.
@@
 Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
-The distribution follows the expected Polya distribution and the MC data in the simulation was fit with a Polya function \cite{blumrol}.
+The distribution follows the expected Polya distribution and the MC data in the simulation was fitted with a Polya function \cite{blumrol}, described as,
+
+\begin{equation}
+P(x) = A_0^{-1}\cdot \frac{A_1^{A_1}}{\Gamma(A_1)} \left(\frac{x}{A_0}\right)^{A_1-1}e^{\frac{-A_1x}{A_0}}.
+\end{equation}
+
+For the voltage of \SI{1460}{\volt} the parameters of the fit are $A_0=903.9$ and $A_1=1.50$ and for the voltage of \SI{1214}{\volt} the parameters are $A_0=150.0$ and $A_1=1.47$.
 
@@
 The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
-There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector CITE HERE.
-Add figure of close up of pad plane
+There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector.
+Figure~\ref{fig:padplane} shows the pad plane boards being glued to the top plate and the holes which allow for the readout of pads.
+The pads were gold plated for excellent electric conduction properties.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{3boards_ps.jpg}
+\caption{Figure of the pad plane boards being glued to the top plate. }
+\label{fig:padplane}
+\end{figure}
 
@@
 Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.
-Short cables transmit the signals from the pads to the inputs of the AGET chips.
+Short cables transmit the signals from the pads to through a circuit protection board called ZAP, to the inputs of the AGET chips mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}.
 Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz.
@@
 In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets).
-The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.
-The variations in the electronics were calibrated by measuring the response of each channel to a injected reference pulse, covering the full dynamic range of each channel.
 
-\begin{table*}\centering
+\begin{table*}[!htb]
+\centering
 \ra{1.3}
@@
 
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=.5\textwidth]{GET_and_ZAP.pdf}
+\caption{•}
+\label{fig:getzap}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{AGETfncn.jpg}
+\caption{Schematic of the internals of the AGET chip from \cite{get2}}
+\label{fig:aget}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=.5\textwidth]{GETlayout.jpg}
+\caption{Readout structure of the AsAd boards and CoBo board structures. Also the relevant components of the DAQ system.}
+\label{fig:coboDAQ}
+\end{figure}
+
+After each AsAd board has digitized the data it is sent to the Concentration Boards (CoBo).
+Each CoBo board can concentrate the data from 4 AsAd boards.
+The Multiplicity, Trigger, and Time module (MuTanT) \cite{get} provides the common trigger signal for all CoBo boards.
+Each board sends the data to the DAQ server which writes to disk the data from each board, which was handled by two separete DAQ servers; saving to one common analysis server.
+The data could then be analyzed using the RIKEN High Performace Computing cluster or moved to the NSCL or MSU cluster for analysis.
+The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.
+
 \section{Energy loss in material}
@@
 The most probable energy loss is the most desirable observable either through fitting of the observed distribution or through the truncated mean method.
-The truncated mean is the average mean value calculated after throwing away the top fraction of the highest energy loss entries, where the fraction used in this thesis was 30\%.
+The solid curve in Fig.~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc}.
+The dashed line is the distribution under the Landau assumptions.
+The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{bichsel.png}
+\caption{Energy loss of a $\beta\gamma$ = 3.6 particle in Ar gas taken from \cite{bichsel}}
+\label{fig:straggling}
+\end{figure}
+
+The truncated mean is the average mean value calculated after throwing away the top fraction of the highest energy loss entries.
 This approximates the most probable value without performing a fit to a known distribution.
-There are
+To calculate the truncated mean $C$, the $n$ energy loss values in a given track $\Delta_i/x$ are sorted from smallest to largest, and the mean value of a reduced set of points $n_t = f_r n$ is calculated as,
+
+\begin{equation}
+C = \frac{1}{n} \sum\limits_{i}^{n_t} \Delta_i/x,
+\end{equation}
+
+where $f_r$ is the cut off fraction; in this thesis, a value of 0.7 was used.
 
@@
 %\begin{equation}
-%<dE/dx>_t = \frac{\sum_{i < j} E_i}{N}
+%\langledE/dx\rangle_t = \frac{\sum_{i < j} E_i}{N}
 %\label{eq:truncatedM}
@@
 
- A full description of the energy loss distribution can be described in CITE HERE.
+ %A full description of the energy loss distribution can be described in CITE HERE.
 
 \subsection{Gas Properties}
-The gas contained by the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure (1 atm).
+The gas contained by the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure 1 atm.
 The  gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere.
@@
 ppm.
-Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm as a function of electric field value in \si{\volt\per\centi\metre}.
+Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm (\SI{760}{\torr}) as a function of the reduced electric field value given in units of \si{\volt\per\centi\metre\per\torr}.
 Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
@@
 
-Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $sigma_l$ and $sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
+Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $\sigma_l$ and $\sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
 
@@
 Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Fig.~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
-If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed on each pad, Q, is given as,
+If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed on each pad, Q, is expressed as,
 
 \begin{equation}
-Q(x_o,z_o) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(x,z) dxdz,
+Q(x_o,z_o) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(x-x_o\textprime,z - z_o\textprime) dxdz,
 \label{eq:prfpadCharge}
@@
 
-where $x_o$ and $z_o$ represent the x and z coordinates of the center of each pad, w is the width, and l is the length.
-This function is commonly refered to as the Pad Respons Funciton (PRF).
-The total charge observed for this track is a superposition of PRFs of each avalanches on all the anode wires.
-It is the use of this PRF that allows for sub-millimeter accuracy in the position resolution of the TPC \cite{blumrol}.
+where $x_o$ and $z_o$ represent coordinates of the center of each pad, $x_0\textprime$ and $z_o\textprime$ are the coordinates of the avalanche, $w$ is the width, and $l$ is the length.
+The total charge observed for this track is a superposition of each avalanches on all the anode wires.
 Typically in a TPC, hits are grouped into clusters and it is practical to cluster in only one direction.
-The PRF can be expressed in a way that is independent of the location of the avalanche.
-If the true position of the avalanche is ($x_o^{'}$,$z_o^{'}$), and the charge distribution can be written in a position independent way %$\rho(\lambda_x,\lambda_z)$, where $\lambda_x = x - x_o^'$ and $\lambda_z = z - z_o^'$.
-The PRF for layer clustering can be written as,
+The marginal probability distribution over a given z-layer can be written as,
 
 \begin{equation}
-Q(\lambda_x) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(\lambda_x,\lambda_z) dxdz,
+\rho_x(x) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \rho(x,z)dz,
+\end{equation}
+
+and over a given z-layer can be written as,
+
+\begin{equation}
+\rho_z(z) = \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(x,z)dx.
+\end{equation}
+
+By substituting the variables  $\lambda_x = x - x_o\textprime$, and $\lambda_z = z - z_o\textprime$, we can express the charge distribution independent of the avalanche location.
+The Pad Response Function (PRF) along the direction x can be written as,
+
+\begin{equation}
+P_X(\lambda_{x_o}) = \frac{ \int_{\lambda_{x_o}-\frac{w}{2}}^{\lambda_{x_o} + \frac{w}{2}} \rho_x(\lambda_x)d\lambda_x } {\int_{-\infty}^\infty \rho_x(\lambda_x)d\lambda_x     },
 \label{eq:prflayer}
@@
 
-and the PRF for row clustering can be written as,
+where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction the PRF can be written as,
 
 \begin{equation}
-Q(\lambda_z) = \int_{z_o - \frac{l}{2}}^{z_o + \frac{l}{2}} \int_{x_o - \frac{w}{2}}^{x_o + \frac{w}{2}} \rho(\lambda_x,\lambda_z) dxdz.
+P_Z(\lambda_{z_o}) = \frac{ \int_{\lambda_{z_o}-\frac{l}{2}}^{\lambda_{z_o} + \frac{l}{2}} \rho_z(\lambda_z)d\lambda_z }{\int_{-\infty}^\infty \rho_z(\lambda_z)d\lambda_z  },
 \label{eq:prfrow}
@@
 
-The The direction most perpendicular to the track's direction of travel provides the most accurate position resolution.
-Clustering in the \spirit TPC is done for several pads containing charge in the direction most perpendicular to the track's direction of travel.
-
-In this example the clustering direction would be along the $x$ axis, as illustrated by the bolded pads.
-By choosing to cluster only in one direction, we have not included the charge in adjacent pads along the $z$ axis resulting from the tails of $\rho(z)$.
-Therefore the charge not included in the the bolded pads will be incorporated into adjacent clusters, introducing small correlations in charge between neighboring clusters.
+where $\lambda_{z_o} = z_o - z_o\textprime$.
 
@@
 
-\subsection{Experimental Pad Response Function}
-
-The correlations we introduced by only clustering along one direction do not play a significant role in the particle identification, but cause deviations from the expected Gatti distribution.
-Also, analytic PRFs only exist for classical multi-wire TPCs.
-For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
-
-As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as $\lambda_i = x_i-\bar{x}$.
-The PRF is simply defined as the charge fraction of each pad as a function of $\lambda$, as shown in Equation \ref{eq:prf}.
-
-\begin{equation}\label{eq:prf}
-PRF(\lambda_i) = \frac{q_i(\lambda_i)}{Q}
-\end{equation}
-
-Averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
-Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
-
-\begin{equation}\label{eq:gaus}
-PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
-\end{equation}
-
-\begin{figure}[ht!]s
+\begin{figure}[!htb]
 \begin{overpic}[width=\linewidth]{fig5.pdf}
@@
 
-The shape of the PRF depends on the crossing angle of the track \cite{gatti}.
-Plotted in Fig.~\ref{fig:prfpimData} is the PRF of $\pi^-$ tracks vs.
-the crossing angle $\theta$.
-The PRF gets wider starting from $90^{\circ}$  and going to $45^{\circ}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
-Since we switch the clustering direction from $x$ to the $z$ direction at $45^{\circ}$, the opposite trend is seen where the PRF becomes narrower as the position resolution gets better going from $45^{\circ}$ to $0^{\circ}$.
+Since we take the marginal distributions only along one layer or row of pad  correlations are introduced in the PRF from adjacent layers or rows which cause slight deviations from the expected Gatti distribution.
+Also, analytic PRFs only exist for classical multi-wire TPCs.
+For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
+The method for extracting the experimental PRF will be discussed latter, by averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
+Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
 
-\begin{figure}[!htb]
-     \centering
-	 \includegraphics[width=\textwidth]{PRFs_data_wcut.png}
-     \caption{PRF response from $\pi^-$ data. }
-     \label{fig:prfpimData}
-\end{figure}
-
-\begin{figure}[ht!]
-\vspace{5mm}
-\includegraphics[width=\linewidth]{fig7}
-\caption{Parameters $N_{0}$ and $\sigma$ as a function of the crossing angle $\theta$ with the $4^{th}$ order polynomial fits.}
-\label{fig:normsigma}
-\end{figure}
+\begin{equation}\label{eq:gaus}
+PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
+\end{equation}
 
 \begin{comment}
-\begin{table}
-\centering
- \begin{tabular}{||c c c c c c||} 
- \hline
- Coefficient & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ [0.5ex] 
- \hline\hline
- $0 < \theta < 45$ & & & & &  \\ [.25ex]
- \hline
- $N_0$ & .897 & 5.766E-3 & -4.263E-4 & 7.444E-6 & 5.705E-8 \\ 
- \hline
- $\sigma$ & 5.496 & -3.920E-2 & 2.693E-3 & -5.208E-5 & 5.334E-7\\
- \hline
- $45 < \theta < 90$ & & & &  & \\ [.25ex]
- \hline	
- $N_0$ & 1.220 & -6.258E-2 & 1.608E-3 & -1.492E-5  & 4.654E-8 \\
- \hline
- $\sigma$ & 31.368 & -1.109 & 1.779E-2 & -1.336E-4 & 3.940E-7\\
- \hline
-\end{tabular}
-\caption{Coefficients of the $4_th$ order polynomial fit to the Gaussian parameters $N_0$ and $\sigma$. The polynomial form is given as $c_0 + c_1 x + c_2 x^2 + c_3 x^3 + c_4 x^4$}
-\label{tb:coeff}
-\end{table}
-\end{comment}
-
-Fits were performed to the experimental data with  $5^{\circ}$ width bins from $0^{\circ} < \theta \leq 90^{\circ}$.
-The two parameters of the Gaussian fits are plotted versus $\theta$ in Fig.~\ref{fig:normsigma}; a $4^{th}$ order polynomial fit between these points allowed for interpolating between $\theta$.
-
 \subsection{Considerations when constructing a TPC}
@@
 Two volumes of gas. 
+\end{comment}
+
+\section{Radio Isotope Beam Factory (RIBF) Facility }
+%Cyclotron facility overview.
+%Samurai line overview.
+%Beam line element overview.
+%Big rips beam PID. reference 
+The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
+The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), reaching a primary beam energy of \SI{345}{\MeVA}.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{SAMURAI-beamline.png}
+\caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
+\label{fig:samuraiBeamLine}
+\end{figure}
+
+After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
+These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
+This is accomplished through several dipole magnets, slits, and wedge degraders.
+The resulting secondary beam is not pure and the purity depends on the primary beam and target beam desired.
+
+In these set of experiments several beams were produced with varying intensities and purities.
+Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns.
+
+ \begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}ccccc@{}}\toprule 
+ Primary Beam & Secondary Beam & Energy at mid target \si{\MeVA} & Intensity \si{\kilo\hertz} & Purity (\%) \\ [0.5ex] 
+ \midrule
+ ${}^{238}$U   & ${}^{132}$Sn   &  269.2  &  9.5  &  54   \\
+ ${}^{238}$U   & ${}^{124}$Sn   &  270.3  &  9.1  &  10  \\
+ ${}^{124}$Xe  & ${}^{112}$Sn   &  270.4  &  7.6  &  48  \\
+ ${}^{124}$Xe  & ${}^{108}$Sn   &  269.3  &  7.5  &  52   \\
+ \bottomrule
+\end{tabular}
+\caption{Primary and secondary beam properties produced in the \spirit TPC experimental campaigns. }
+\label{tb:beams}
+\end{table*}
+
+\section{Experimental Setup}
+
+The \spirit TPC was designed to fit exactly into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line.
+Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration.
+Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
+This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap.
+The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces.
+These bolts secure the vacuum chamber to the magnet which is not practically removable; though the inside of the magnet was not operated under vacuum.
+This required an extensive rail system and support frame to slowly slide in the TPC over the bolts, finally raising the TPC several \si{\centi\metre} to the final height.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{perspective.png}
+\caption{Drawing of the experimental setup with the TPC inside of the SAMURAI magnet at $\ang{0}$ configuration.}
+\label{fig:experiment}
+\end{figure}
+
+The height of the TPC was roughly aligned with a self-leveling laser system to match the center of target with the center of the beam line.
+Once the TPC was adjusted to the final location, the position of the TPC was measured in fine detail with a the VStars-N photogrametry system \cite{vstars}.
+Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system.
+Using the commercial software provided the set of different camera perspectives reconstruct a point cloud of all the targets into 3-dimensional coordinates.
+Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC -- and several of its internal components-- relative to the magnet frame.
+The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed or Show data.
+
+%Maybe put a position table summary here of the TPC position and definition of the coordinates system in the TPC frame and the Magnet frame
 
 \section{Ancillary Detectors }
+Several ancillary detectors were placed inside and outside of the \spirit TPC to facilitate making the trigger for the experiment.
+Placing detectors around the TPC was one of the important considerations we made when designing the TPC.
+A brief description of each detector system is given here with particular focus on how the experimental trigger was made.
 
@@
 The threshold was set so that any beam particle which passed through any of the bars it would send a trigger signal to not trigger the system since the beam path would not be on target but on some other material inside the TPC.
-
-\section{Radio Isotope Beam Factory (RIBF) Facility }
-%Cyclotron facility overview.
-%Samurai line overview.
-%Beam line element overview.
-%Big rips beam PID. reference 
-The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
-The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), reaching a primary beam energy of \SI{345}{\MeVA}.
-
-\begin{figure}[!htb]
-\includegraphics[width=\linewidth]{SAMURAI-beamline.png}
-\caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
-\label{fig:samuraiBeamLine}
-\end{figure}
-
-After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
-These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
-This is accomplished through several dipole magnets, slits, and wedge degraders.
-The resulting secondary beam is not pure and the purity depends on the primary beam and target beam desired.
-
-In these set of experiments several beams were produced with varying intensities and purities.
-Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns.
-
- \begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}rrrrr@{}}\toprule 
- Primary Beam & Secondary Beam & Energy at mid target \si{\MeVA} & Intensity \si{\kilo\hertz} & Purity (\%) \\ [0.5ex] 
- \midrule
- ${}^{238}$U   & ${}^{132}$Sn   &  269.2  &  9.5  &  54   \\
- ${}^{238}$U   & ${}^{124}$Sn   &  270.3  &  9.1  &  10  \\
- ${}^{124}$Xe  & ${}^{112}$Sn   &  270.4  &  7.6  &  48  \\
- ${}^{124}$Xe  & ${}^{108}$Sn   &  269.3  &  7.5  &  52   \\
- \bottomrule
-\end{tabular}
-\caption{Primary and secondary beam properties produced in the \spirit TPC experimental campaigns. }
-\label{tb:beams}
-\end{table*}
-
-\section{Experimental Setup}
-
-The \spirit TPC was designed to fit exactly into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line.
-Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration.
-Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
-This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap.
-The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces.
-These bolts secure the vacuum chamber to the magnet which is not practically removable; though the inside of the magnet was not operated under vacuum.
-This required an extensive rail system and support frame to slowly slide in the TPC over the bolts, finally raising the TPC several \si{\centi\metre} to the final height.
-
-The height of the TPC was roughly aligned with a self-leveling laser system to match the center of target with the center of the beam line.
-Once the TPC was adjusted to the final location, the position of the TPC was measured in fine detail with a photogrametry system CITE HERE.
-Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system.
-Using the commercial software provided the set of different camera perspectives reconstruct a point cloud of all the targets into 3-dimensional coordinates.
-Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC -- and several of its internal components-- relative to the magnet frame.
-The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed CITE HERE or Show data.
-
-Maybe put a position table summary here of the TPC position and definition of the coordinates system in the TPC frame and the Magnet frame
-
-\begin{figure}
-\includegraphics[width=\textwidth]{perspective.png}
-\caption{Drawing of the experimental setup with the TPC inside of the SAMURAI magnet at $\ang{0}$ configuration.}
-\label{fig:experiment}
-\end{figure}
 
@@
 During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
-In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating grid trigger CITE HERE JONS THESIS.
+In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating grid trigger \cite{jon}.
 Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.

### 2020-05-15 chapters/ch_experiment.tex
First draft before Bill for Experimental section

@@
 Here we will outline the physical principles involved in the TPC measurement and discuss the particular details of the TPC used in this thesis.
-Inside of at TPC the detector gas is housed in a field cage, which also sets up a constant electric field.
-As charged particles pass through the gas, electron-ion pairs from neutral gas molecules by passing tracks; the electrons are accelerated opposite to the electric field and the ions move in the opposite direction.
-Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules, slowing down the electron, which is then accelerated again repeating the cycle.
-This microscopic behavior manifests into a constant drift velocity when averaged over several gas collisions.
-These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field, liberating more electron-ion pairs from the gas creating an avalanche.
-The avalanche electrons finally terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires; creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.
+Figure~\ref{fig:tpcPrinciple} depicts the inside of a TPC, specifically the field cage, which holds the detector gas and also sets up a constant electric field.
+As charged particles resulting from a heavy ion collision pass through the gas, electron-ion pairs are created from neutral gas molecules.
+The electrons are accelerated opposite to the electric field and the ions along with it.
+Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules; slowing down, or even stopping the electron, which is then accelerated again, starting a repeating cycle of stop and go motion.
+This microscopic behavior manifests as a constant drift velocity when averaged over several gas collisions.
+The electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they quickly accelerate in the presence of the high electric field, liberating more electron-ion pairs from the gas, creating an avalanche process to occur.
+The avalanche electrons eventually terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires, creating a large signal which is distributed over the pad-plane.
+The charge and timing  information of these electrons are calculated from the induced signal measured by the  electronics.
 
-Two of the 3 coordinates are determined from the charge distribution on the pad plane.
-The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics.
-The radius of curvature is related to the magnetic rigidity, and therefore the momentum of each track.
-The energy loss  deposited ($\langle dE/dx\rangle$) is measured by the segmented charge sensitive pads on the pad-plane, which are connected to the readout electronics.
-Particle identification is achieved through since unique particles exist on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections.
+Two of the 3 coordinates are determined from the 2-dimensional charge distribution on the pad-plane.
+The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$.
+The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the timing information of the signal in the electronics.
+The radius of curvature of the track is related to the magnetic rigidity, and therefore the momentum of each track.
+The energy loss deposited $\langle dE/dx\rangle$ is measured by the segmented charge sensitive pads on the pad-plane, in which each pad is connected to a channel in the electronics.
+Particle can be uniquely identified since  each particles exists on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections.
 In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.
@@
 The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in fixed target experiments.
-The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls all around, in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
-The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}.
+The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
+The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}.
+The dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}.
 More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.
@@
 The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal.
-The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate; several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a laser \cite{jon}.
+The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate.
+Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurment \cite{jon}.
 Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics.
-The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC
+The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC.
 
@@
 
+\begin{comment}
 \subsection{Enclosure}
@@
 Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.
+\end{comment}
 
@@
 The field cage was constructed from several panels of printed circuit boards (PCBs).
-The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces.
-The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires \cite{tpcAging}.
-The halogen free material used was Cryogenic-G10.
-The field cage is an isolated volume from the enclosure of the TPC for the option to run explosive gases such hydrogen in the field cage, where the enclosure can be filled with an insulating gas.
-While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only.
-Instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame bonded to a \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips.
-The PCB boards were epoxied into the cathode which was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure.
+The front of the field cage was made of two PCBs and each side was constructed from three PCBs, supported by Lexan pieces.
+The common PCB substrate material FR4, contains a bromine epoxy which can outgas and absorb electrons degrating the signal \cite{tpcAging}.
+Therefore a halogen free material, Cryogenic-G10, was used for the board material.
+The field cage volume is isolated from the enclosure volume of the TPC for the option to run two separate  gasses, which may be useful when running explosive gasses such hydrogen.
+Instead of a downstream wall a, large, thin exit window was constructed, which consisted of a  \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips, which was bonded to a removable Lexan frame.
+The PCB boards were epoxied, and screwed, into the cathode which was constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure.
 On the other end the boards were epoxied into an aluminum top perimeter which also served as the last ring in the TPC.
 Together with the cathode bottom, the field cage proved to be a rigid lightweight structure.
-A lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC.
-Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate.
-The top plate could be removed and rotated with the field cage on without damaging any internal components.
+A Lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC.
+Screws with nylon washers, and collars, were used to mount the top perimeter --and therefore the field cage-- to the top plate; it also provided electrical isolation from the top plate.
+In this way the top plate could be removed and rotated with the field cage attached on without damaging any internal components.
 
@@
 
+Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
 The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
-The cathode voltage $V_{cath}$ can be calculated from the schematic of the TPC system in Fig.~\ref{fig:TPC_schematic}, as
+The cathode voltage $V_{cath}$ can be calculated as,
 \begin{equation}
@@
 
-Figure~\ref{fig:FC_schematic} shows a cartoon schematic of the field cage walls.
+Figure~\ref{fig:FC_schematic} shows a schematic detailing the connections involved in the field cage walls.
 The field cage contains 50 inside copper strips and 49 outside copper strips.
-The strips on each board was connected to the adjacent strip on the next board; front, back, and side boards were connected by G10 corner pieces with conducting paint strips, creating rings around the whole field cage.
+The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart.
+The strips on each board was connected to the adjacent strips on the adjacent boards.
+The interfaces were the side pieces met the front boards or back window are connected by G-10 corner pieces with conducting paint strips, which makes the strips into continuous rings around the whole field cage.
 The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip.
-The first outside strip is connected trough a via to the second inside strip, which continues on in this repeating fashion.
-The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance, setting up a constant electric field.
-The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the aluminum top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths.
+The first outside strip is connected through a via to the second inside strip.
+This pattern repeats until the last strip.
+The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field.
+The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths.
 The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.
@@
 \subsection{Voltage Step Down}
-The gap between the field cage's cathode and the ground of the enclosure is quite small.
-To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor.
+The gap between the cathode and the ground of the enclosure is quite small.
+To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground in a controlled manner, reducing the chance of electric breakdown and sparking.
 There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time.
@@
 The wire plane closest to the pad-plane (\SI{4}{\milli\metre}) are the anode wires.
-The next plane (\SI{12}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid.
+The next plane (\SI{8}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid.
 The gating grid is the first plane that electrons meet as they drift upward from the field cage volume towards the anode plane.
-The gating grid is operated as a gate either allowing electrons and ions through, or blocking them entirely.
-The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires where the avalanche process of the electron takes place.
-The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC.
-If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.
+The gating grid is operated as a gate, either allowing electrons and ions through, or blocking them entirely.
+The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires.
+The ground plane is the least interesting plane and is held to ground by shorting the plane to the enclosure through  shorted BNC terminator on the outside of the TPC.
+We also use the ground plane to input a pulser which is used to spread the pulsed signal to all the pads on the TPC in order to calibrate the electronics of the TPC.
+This is done by replacing the shorted BNC with a \SI{50}{\ohm} termination and injecting the pulser on the other end.
 
@@
 
-In the open configuration the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume.
-Typically the gating grid is held in the closed configuration only opening when the data acquisition trigger criteria is met.
-This is to block the electrons which come from the un-reacted beam, which happens frequently, which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions and flood the volume of the field cage with space charge.
-We close the gating grid after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- to also prevent the back-flow of ions from the avalanche region from that event.
-Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
+In the open configuration, the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume.
+Typically the gating grid is held in the closed configuration, only opening it when the data acquisition trigger criteria is met.
+By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, and would flood the volume of the field cage with space charge.
+We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume.
+After this we close the gating grid to prevent the back-flow of ions from the avalanche region from that event.
+Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
 
-Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations.
-In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency.
+Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the drift lines of electrons in both the on and off configurations.
+In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, in which the optimal voltage is selected for the case of 100\% electron transparency.
+Figure~\ref{fig:wires_open} shows the electrons are allowed to drift completely through the gating grid all the way to terminate on the anode wires.
 In the off configuration, the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field produced by the voltage difference $2\Delta V$ between wires is great enough to block incoming electrons.
-Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the state of the reference voltage.
+Figure~\ref{fig:wires_closed} shows this case were the electron drift lines are fully blocked terminating on the more positive wires.
+Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the steady state of the average reference voltage for 100\% transparency.
 
@@
 To measure the electron transparency were all wires share the voltage $V_{avg}$, the anode wire was lowered to \SI{500}{\volt} and the beam was allowed to enter the field cage without any target put in.
-By lowering the voltage of the anode we could measure the large charge of the beam without saturating the electronics.
+By lowering the voltage of the anode wires we could measure the large charge of the beam without saturating the electronics.
 The average charge deposited in the chamber could be measured as a function of $V_{avg}$; changing the top plane resistor appropriately according to Eq.~\ref{eq:TP_resistor}.
@@
 They are biased to high voltages around \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
-As the eletron drifts towards the anode wire, it gains kinetic energy knocking out more electron-ion pairs in the gas until it terminates on the anode wire or the pad plane.
+As the eletron drifts towards the anode wire, it gains enough kinetic energy to knock out more electron-ion pairs in the gas which also go on to do the same; until they terminate on the anode wires or the pad plane.
 The amount of electrons produced depends on the anode wire voltage and the gas properties.
 The absolute gas gain was not experimentally measured but was calculated in a Garfield simulation.
-During the experiment the anode wires were biased to two different voltages.
+For the experimental data pertaining to this thesis, the anode wires were biased to two different voltages.
 We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
-Only two sections were biased with the lower voltage setting due to concerns of a high current issue.
-Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche.
-The distribution follows the expected Polya distribution and the MC data in the simulation was fitted with a Polya function \cite{blumrol}, described as,
+Only two sections were biased with the lower voltage setting due to a leakage around the end of the gating grid \cite{jon}.
+Figure~\ref{fig:anodegain} shows the electron distribution for the total number of electrons produced in a avalanche process created by a single electron.
+The distribution follows the expected Polya distribution, and the MC data in the simulation was fitted with a Polya function \cite{blumrol}, which can be expressed as,
 
@@
 
-The wire plane and the cathode define the different electric field regions.
-The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the ground grid and the gating grid (Region 2); the avalanche region is defined as between the ground grid and the anode grid.
-The voltage of the gating grid is set to the point that allows for 100\% transparency for electrons.
+The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the gating grid and the ground grid (Region 2), and the avalanche region is defined as between the ground grid and the anode grid.
 It is possible that the electric field in Region 1 and Region 2 can be matched but not required.
-The voltage of the top perimeter is set by adjusting the value of the last resistor ($R_{TP2}$) to ensure the electric field is constant through Region 1.
-We can imagine Region 1 is split into two virtual volumes, one defined between the cathode and the top perimeter, and one defined between the top perimeter and the gating grid.
-The magnitude of the electric field in the region between the top perimeter and the cathode, $E_1$, is defined as,
+To ensure the electric field is continuous across Region 1, the voltage of the top perimeter  must be set properly by adjusting the value of the last resistor ($R_{TP2}$).
+We can imagine Region 1 is split into two virtual volumes, one defined as the volume between the cathode and the top-perimeter, and one defined between the top-perimeter and the gating-grid.
+The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_1$, is defined as,
 
@@
 \end{equation}
-where  $V_{g.g.}$, $V_{tp}$ , $y_{g.g.}$, and $y_{tp}$ are the voltage and vertical y-position of the gating grid and top perimeter respectively.
-The magnitude of the electric field in the region between the top perimeter and the cathode, $E_2$, is defined as,
+where  $V_{g.g.}$, $V_{tp}$ , $y_{g.g.}$, and $y_{tp}$ are the voltages and vertical y-positions of the gating-grid and top-perimeter respectively.
+The y-position here refers to the center of the electrodes.
+The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_2$, is defined as,
 
@@
 \end{equation}
-where  $V_{tp}$, $V_{cath}$ , $y_{tp}$, and $y_{cath}$ are the voltage and vertical y-position of the top perimeter and cathode respectively.
-The condition for a smooth electric field across the volume is given by $E_1 = E_2$.
-Substituting Eq.~\ref{eq:FCstrip} for $V_{tp}$ ($n=50$), we can solve for the effective resistance of the top perimeter $R_p$ as,
+where  $V_{tp}$, $V_{cath}$ , $y_{tp}$, and $y_{cath}$ are the voltage and vertical y-position of the top-perimeter and cathode respectively.
+The y-position of the cathode is defined as the face of the cathode.
+The condition for a smooth electric field across these two virtual volumes is defined as the solution to the equation $E_1 = E_2$.
+Substituting Eq.~\ref{eq:FCstrip} for $V_{tp}$ -- $n=50$ -- we can solve for the effective resistance of the top perimeter $R_p$ as,
 
@@
 where the relevant vertical dimensions are $y_{g.g.} - y_{cath} = \SI{497.3}{\milli\metre}$ and $y_{tp} - y_{cath} = \SI{490}{\milli\metre}$.
-The value of $R_{TP2}$ can be calculated from Eq.~\ref{eq:Reff} and Eq.~\ref{eq:TP_resistor}.
+The value of $R_{TP2}$ can then be calculated from Eq.~\ref{eq:Reff}.
 
 \subsection{Pad Plane}
-The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively making 12096 pads in total.
-There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads.
-The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
-There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector.
+The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively, making 12096 pads in total.
+There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads so that the effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
+There is a via and trace coming from each pad, through the board, to the opposite side of the pad plane and is arranged in a surface pads which is readout by a surface mount SAMTEC connector.
 Figure~\ref{fig:padplane} shows the pad plane boards being glued to the top plate and the holes which allow for the readout of pads.
@@
 Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.
-Short cables transmit the signals from the pads to through a circuit protection board called ZAP, to the inputs of the AGET chips mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}.
-Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz.
-Four AGET chips are mounted on one AsAd (ASIC and ADC) motherboard.
-The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range, and the ADCs on each AsAd board provides 12 bit resolution.
-The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns.
-In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets).
+Short cables transmit the signals from the surface mount connectors, through a circuit protection board called ZAP, to the inputs of the AGET chips which are mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}.
+Each AGET chip services 64 pads (63 pads are connected in our case).
+Four AGET chips are mounted on one AsAd  motherboard.
+Figure~\ref{fig:aget} is the schematic of each  AGET chip which contains a charge sensitive pre-amplifier, several other stages of amplifiers, and a Switched Capacitor Array (SCA) with a maximum of 512 time buckets which operates in a circular readout buffer.
+The sampling frequency can be adjusted from 1 to \SI{100}{\mega\hertz}.
+The gain of each AGET can be configured as 0.12, 0.24, 1.0, or \SI{10}{\pico\coulomb} over the whole dynamic range, and the analog-to-digital converters (ADCs) on each AsAd board provides 12 bit resolution.
+The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or \SI{1014}{\nano\second}.
+In this experiment, the gain was set to the highest setting, 0.12 \si{\pico\coulomb}, the peaking time \SI{117}{\nano\second}, and the sampling frequency \SI{25}{\mega\hertz} (resulting in \SI{40}{\nano\second} time buckets).
 
@@
 Each board sends the data to the DAQ server which writes to disk the data from each board, which was handled by two separete DAQ servers; saving to one common analysis server.
-The data could then be analyzed using the RIKEN High Performace Computing cluster or moved to the NSCL or MSU cluster for analysis.
-The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.
+The data could then be analyzed using the RIKEN High Performance Computing (HPC) cluster or moved to the NSCL or MSU cluster for analysis.
+The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used in this analysis.
 
@@
 \begin{equation}\label{eq:bb}
-\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2).
+\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2),
 \end{equation}
-Where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle.
-There is a large variation in energy loss around this mean value, with a long high energy loss tail.
-
+where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle \cite{blumrol}.
+Yet there is a large variation in energy loss around this mean value.
 The statistical variation of energy loss in a material was described by Landau \cite{landau} and later better described by Shulek \cite{shulek} and Bichsel \cite{bichsel1}.
-In both approximations it is described by a most probable energy loss value with a long, high-energy loss tail.
-Because of this long tail, for a finite set of energy loss measurements along a given track, the fluctuation of the mean value energy loss is a very unreliable observable.
-The most probable energy loss is the most desirable observable either through fitting of the observed distribution or through the truncated mean method.
+In both approximations it is described by a most probable energy loss value, with a long, high-energy loss tail.
 The solid curve in Fig.~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc}.
@@
 The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.
+Because of this long tail, for a finite set of energy loss measurements along a given track, the fluctuation of the mean value energy loss as an observable is a very unreliable.
+The most probable energy loss the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.
 
@@
 This approximates the most probable value without performing a fit to a known distribution.
-To calculate the truncated mean $C$, the $n$ energy loss values in a given track $\Delta_i/x$ are sorted from smallest to largest, and the mean value of a reduced set of points $n_t = f_r n$ is calculated as,
+The set of $n$ observed energy loss values, $\Delta_i/x$, in a given track are sorted from smallest to largest value.
+The truncated mean $C$, is calculated from the reduced set of points $n_t = f_r n$ as,
 
@@
 \subsection{Gas Properties}
-The gas contained by the field cage was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure 1 atm.
+The gas used was a mixture of 90\% Ar and 10\% Methane ($\mathrm{CH_4}$) by volume (P10 gas), and operated just under atmospheric pressure 1 atm.
 The gas was continually flowed through the field cage and exited into the enclosure volume, finally passing through a bubbler to atmosphere.
@@
 Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm (\SI{760}{\torr}) as a function of the reduced electric field value given in units of \si{\volt\per\centi\metre\per\torr}.
-Operating near the peak value of the drift velocity curve minimizes the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
+Operating near the peak value of the drift velocity curve minimized the change in the drift velocity as the effective field slightly changes due to slight variations in the pressure.
 The electric field in the experiment was \SI{125}{\volt\per\centi\per\metre} at \SI{760}{\torr}, giving a reduced electric field \SI{0.17}{\volt\per\centi\metre\per\torr}.
@@
 
+ADD THE VALUES OF TAU AND MU HERE
+
 %Add table for gas diffusion
@@
 Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Fig.~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
-If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed on each pad, Q, is expressed as,
+If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed a particular pad, $Q$, is expressed as,
 
@@
 
-where $x_o$ and $z_o$ represent coordinates of the center of each pad, $x_0\textprime$ and $z_o\textprime$ are the coordinates of the avalanche, $w$ is the width, and $l$ is the length.
-The total charge observed for this track is a superposition of each avalanches on all the anode wires.
-Typically in a TPC, hits are grouped into clusters and it is practical to cluster in only one direction.
-The marginal probability distribution over a given z-layer can be written as,
-
+where $x_o$ and $z_o$ represent coordinates of the center of that pad, $x_0\textprime$ and $z_o\textprime$ are the coordinates of the avalanche, $w$ is the width, and $l$ is the length of the pad.
+The total charge observed for a given track is a superposition of all avalanches on all the anode wires.
+Typically in a TPC, the charges on each pad are grouped into clusters, and it is practical to cluster in only one direction.
+Therefore we will speak about the marginal probability distribution over a given layer of pads (x-distribution), or row of pads (z-distribution).
+The marginal distribution for a given layer can be written as,
 \begin{equation}
@@
 
-and over a given z-layer can be written as,
+and over a given row can be written as,
 
@@
 By substituting the variables  $\lambda_x = x - x_o\textprime$, and $\lambda_z = z - z_o\textprime$, we can express the charge distribution independent of the avalanche location.
-The Pad Response Function (PRF) along the direction x can be written as,
+The Pad Response Function (PRF) along the x-direction of a given layer can be written as,
 
@@
 
-where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction the PRF can be written as,
+where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction  of a given roaw the PRF can be written as,
 
@@
 
-Since we take the marginal distributions only along one layer or row of pad  correlations are introduced in the PRF from adjacent layers or rows which cause slight deviations from the expected Gatti distribution.
+Since we take the marginal distributions only along one layer or row of pad, correlations are introduced in the PRF from adjacent layers or rows which cause slight deviations from the expected Gatti distribution.
 Also, analytic PRFs only exist for classical multi-wire TPCs.
 For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
-The method for extracting the experimental PRF will be discussed latter, by averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
+The method for extracting the experimental PRF will be discussed latter, but by averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
 Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
@@
 The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
-The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), reaching a primary beam energy of \SI{345}{\MeVA}.
+The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
+Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.
 
@@
 This is accomplished through several dipole magnets, slits, and wedge degraders.
-The resulting secondary beam is not pure and the purity depends on the primary beam and target beam desired.
+The resulting secondary beam is not pure and the purity depends on the capability of BigRIPS to deliver the secondary beam of choice and the primary beam used.
 
 In these set of experiments several beams were produced with varying intensities and purities.
-Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns.
+Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns where most beams were delivered with an intensity of \SI{10}{\kilo\hertz}.
+The BigRIPS separator had beam tracking elements which allowed for further identification of each individual beam produced.
+Allowing for a selection on the events corresponding to the secondary beam of choice CITE HERE.
 
@@
 Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
-This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap.
+This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap and was operated at \SI{0.5}{\tesla} for these sets of experiments.
 The space between the magnetic pole faces is further complicated by large bolts which protrude from the pole faces.
@@
 Small highly reflective targets were placed all over the TPC both inside and out and pictures were taken with a calibrated lens and camera system.
-Using the commercial software provided the set of different camera perspectives reconstruct a point cloud of all the targets into 3-dimensional coordinates.
-Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC -- and several of its internal components-- relative to the magnet frame.
+Using the commercial software provided the set of different camera perspectives was reconstructed into a 3-dimensional point cloud of all the targets.
+Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC, and its internal components, relative to the magnet frame.
 The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed or Show data.
+
+PUT HERE THE MAGNET FRAME COORDINATE 
+PUT HERE THE TPC FRAME COORDINATE DEFINITION
 
@@
 \section{Ancillary Detectors }
-Several ancillary detectors were placed inside and outside of the \spirit TPC to facilitate making the trigger for the experiment.
+Several ancillary detectors were placed inside and outside of the \spirit TPC to facilitate in making the trigger for the experiment.
 Placing detectors around the TPC was one of the important considerations we made when designing the TPC.
@@
 %scintillator bars
-The Kyoto Multiplicity Array consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
+As seen in Fig.~\ref{fig:aux}, the Kyoto Multiplicity Array consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
 The entire TPC structure was designed so that light charged particles could easily pass through the field cage and side walls of the TPC enclosure.
 In this way the number of tracks passing through the sides of the TPC could be measured by this array.
-In heavy ion collisions the more central a nuclear collision is, the more nucleons participate in the collision, resulting in more measured tracks.
+In heavy ion collisions the more central a nuclear collision is, the more nucleons participate in the collision, resulting in a higher observed track multiplicity.
 It is this correlation between the number of tracks and centrality of the collision that makes the Kyoto Array sensitive to the centrality of events.
@@
 In the experiment the trigger selection criteria was $n_{Kyoto} > 4$, where $n_{Kyoto}$ is the total number of tracks measured by both arrays.
+The Kyoto array proved to be a good trigger for selecting on very central events as will be discussed in later sections.
 
@@
 %beam trigger optional
-The Krakow KATANA array consists of 12 plastic scintillating bars mounted to the downstream wall of the TPC enclosure.
+Also shown in Fig.~\ref{fig:aux} is the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted to the downstream wall of the TPC enclosure.
 Three of the 12 bars were thin and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this was a majority of the time.
 The 9 other bars operated as an additional multiplicity array similar to the Kyoto array.
-Since most of the particles are focus forward in a cone in the laboratory frame, it was found the condition on the Kyoto array was sufficient to trigger on central events; thus the KATANA array was used in primarily the beam veto mode.
+Since most of the particles are focused forward into a cone in the laboratory frame, it was found the condition on the Kyoto array was sufficient to trigger on central events; thus the KATANA array was used in primarily the beam veto mode.
 This was accomplished by positioning the array so that the expected position of the beam exiting the TPC would be centered on the three thin paddles.
-The threshold of the veto paddles were set so that the charge of a particle, $Z$, was $Z > 20$.
-This allowed the selection of very central events and we did not trigger on very peripheral or no collision events.
+The threshold of the veto paddles were set so that the charge of a particle passing through, $Z$, would veto any event that satisfied $Z > 20$, where the charge of the Sn beam is $Z=50$.
 
 \subsection{Active Veto Array}
-The beam was tuned by two sets of quadropole  magnets, STQ 1 and STQ2, so that the beam spot was focused on the TPC target location.
-Because of the inherent angular dispersion of the beam there were incoming beam events which significantly deviated from the target location.
+The beam was tuned by two sets of quadropole  magnets, STQ 1 and STQ2 (as seen in Fig.~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
+Because of the inherent angular dispersion of the beam there still were incoming beam events which significantly deviated from the target location.
 To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size.
@@
 An upstream scintillating bar formed the start counter signal, triggering on any beam coming down the beam line.
-The active veto will trigger for any beam that is incident off the target location.
-The KATANA veto produces a signal if the beam passed through the TPC un-reacted, causing no nuclear collision; this produces a veto signal with a width of \SI{4}{\micro\second} which is the approximate time it takes for the beam to drift and clear the field cage volume.
+The active veto will trigger for any beam that is off the target location.
+The KATANA veto produced a signal if the beam passed through the TPC un-reacted, causing no nuclear collision; this produces a veto signal with a width of \SI{4}{\micro\second} which is the approximate time it takes for the beam to drift and clear the field cage volume.
 The Kyoto multiplicity trigger produces a signal when the total number of tracks passing through both Kyoto arrays are greater than 4.
@@
 There were several special trigger considerations when we built the trigger for the TPC.
-We required that the gating grid be opened fast to not miss any signal; as soon as there was a condition satisfying the Start Counter, Kyoto Multiplicity, the DAQ was not busy, and there was not a KATANA Veto signal.
+We required that the gating grid be opened fast to not miss any signal; as soon as there was a condition satisfying the Start Counter, Kyoto Multiplicity, the DAQ was not busy, and there was not a KATANA Veto signal the gating-grid would be opened.
 This was referred to as the Fast Trigger.
-If the KATANA trigger box is not satisfied --described later-- this will trigger a Fast Clear signal which will not trigger the DAQ and will quickly close the gating grid.
+If the KATANA trigger box is not satisfied --described later-- this will trigger a Fast Clear signal which will not trigger the DAQ, and will quickly close the gating grid.
 Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.
@@
 The master trigger for the DAQ was different for each primary beam as the experiment got progressively better.
-During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
-In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating grid trigger \cite{jon}.
+During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
+In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}.
 Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.
@@
 Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal.
-The KATANA trigger, Kyoto Muliplicity, and Start Counter together trigger the DAQ.
+The KATANA trigger, Kyoto Muliplicity, and Start Counter together triggered the DAQ.
 
@@
 Here the KATANA trigger and the SC SUM???
-togethere trigger the DAQ.
+together triggered the DAQ.
 
  It is worth mentioning how the busy signals for the experiment were handled.
-The DAQ system itself produces a busy signal which was combined with the busy signals from either opening or closing the gating grid.
-When opening the gating grid it is assumed the full volume of the TPC will be read out and therefore  a \SI{11}{\micro\second} gate is produced; which is slightly more than the time it takes for all the electrons to drift in the field cage.
-In the case were the gating grid should be fast closed, either due to the fast clear circuit or the end of the TPC measurement, a \SI{5}{\micro\second} gate is produced to allow for the gating grid to settle to a closed configuration and clear the drift volume of any residual electrons from the beam.
+The DAQ system itself produces a busy signal which was combined with the busy signals from either the opening or closing the gating grid.
+When opening the gating grid, it is assumed the full volume of the TPC will be read out, therefore  a \SI{11}{\micro\second} gate busy signal is produced; which is slightly more than the time it takes for all the electrons to drift in the field cage.
+In the case were the gating grid should be closed, either due to the fast clear circuit or the end of the TPC measurement, a \SI{5}{\micro\second} busy signal is produced to allow for the gating grid to settle to a steady state closed configuration, and to clear the drift volume of any residual electrons from the beam.
 Both of these gates are included with the DAQ in an OR configuration which makes the overall busy signal.
 
-\section{Collision Data Taken}
+%\section{Collision Data Taken}

### 2020-05-21 chapters/ch_data_analysis.tex
Routine upload

@@
 \subsection{Gating grid noise subtraction}
-The gating grid essentially is a fast switch which allows the charges of adjacent wires to reach equilibrium as fast as possible.
+Opening the gating grid essentially short adjacent wires together allowing them to reach equilibrium as fast as possible.
 In practice the impedance of both sides was not entirely matched properly and this caused an oscillating current to bounce back and forth in an under-dampened manner.
-This charge bouncing back and forth caused induced signals early in the time bucket spectrum which created and extra source of noise in the data.
-The ground grid shielded some of the noise, but since the path to ground was located at the ends of the TPC, the gating grid noise varied in intensity across the pads.
+This caused residual induced signals early in the time bucket spectrum which created and extra source of noise in the data.
+The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.
+Figure~\ref{fig:ggNoiseSub} shows in the upper panel the gating grid noise ADC time bucket spectrum for 2000 events in a given pad.
+The signal is stable within a pad and the mean value --shown as the red line-- can be calculated after averaging over several thousands events.
+The raw gating grid noise lasts for 100 tbs extending into the real data with a decreasing amplitude.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{ggNoise.png}
+\caption{Gating grid noise profile and subtraction for a particular pad.}
+\label{fig:ggNoiseSub}
+\end{figure}
 
@@
 
-The average noise profile was constructed by averaging all the events in the gating grid runs, pad-by-pad; each pad had a slightly different noise profile since the impedance to each pad was slightly different.
+Once the mean value response in each pad is calculated, we can use the gating grid noise profile runs to correct the real data and remove the noise of the gating grid.
+To show the best case scenario of this technique, the bottom panel of Fig.~\ref{fig:ggNoiseSub} shows the gating grid noise self correcting itself.
+We can see the pedestal is already subtracted naturally when using this technique, and the subtracted ADC spectra is centered around 0.
+Residual gating grid noise that was not canceled by the subtraction remains but is significantly reduced to a much smaller time bucket region.
+This happens in regions where the gating grid noise is changing rapidly, where the ADC value is dependent upon small jitters in the timing of the signal.
+All time buckets below 30 tb were cut away from the spectrum to get rid of the small region where the gating grid is still opening and to remove the residual noise.
+
+%The average noise profile was constructed by averaging all the events in the gating grid runs, pad-by-pad; each pad had a slightly different noise profile since the impedance to each pad was slightly different.
 It is worth noting that the small electronics noise would be effectively canceled out by averaging over thousands of events.
@@
 
-Light charged particles (p,d,t,${}^{3}$He,${}^{4}$He), beams were produced and measured in the TPC.
-The magnetic and slit settings of the dipoles in the BIGRIPS spectrometer was set so that the measured momentum resolution of the beam was $\frac{\delta p}{p}$ < 1\%.
+A cocktail of several light charged particles were produced (p,d,t,${}^{3}$He,${}^{4}$He), and measured in the TPC to provide a momentum calibration of the TPC.
+The magnetic and slit settings of the dipoles in the BIGRIPS spectrometer was set so that the measured momentum resolution of the beam was $\delta p/p$ < 1\%.
 Two magnetic rigidity settings were studied, with an empty target.
-A thick Aluminum target was used to provide a slightly lower point for part of the lower rigidity setting, effectively creating three calibration points over several particle species.
+A thick Aluminum target was used to provide a slightly lower energy point for part of the lower rigidity setting, effectively creating three calibration points over several particle species.
 The production of certain particles (t,${}^{3}$He), produced too few counts to make a good measurement, and in the high momentum rigidity setting protons could not propagate down the line and there were no counts.
 
-Since the expected momentum resolution resulting from the spectrometer was less than 1\%, the observed momentum resolution measured by the TPC is a good measurement of the combined momentum resolution of the software and TPC (intrinsic detection) system.
-The momentum resolution of the TPC depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
+Since the expected momentum resolution resulting from the spectrometer was less than 1\%, the observed momentum resolution measured by the TPC is a good representation of the combined momentum resolution of the software and TPC (intrinsic detection) system.
+The momentum resolution depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
 This calibration beam represents and ideal situation where the track was parallel to the pad plane and only one particle was measured at one time.
-The energy loss resolution can also be directly inferred from the measurement since each energy setting represents a monochromatic source of each particle species, which has a well defined energy loss distribution.
-An average momentum resolution of ~2\% and the energy loss resolution of ~ 5\% was measured for particles ranging from protons to ${}^{4}$He, over the range of momenta measured in the calibration beam as summarized in Table~\ref{tb:momresolution}.
-
-Since magnetic dipole setting of the BIGRIPS spectrometer define the energy of each particle type we can calculate the expected momenta of each particle species measured.
-Small corrections to the momenta were propagated using LISE++ software which can calculate the energy loss through several materials in the beam line.
+The energy loss resolution can also be directly measured since each energy setting represents a monochromatic source of each particle species, which has a well defined energy loss distribution.
+An average momentum resolution of $dp/p = 2\%$ and the energy loss resolution of $d\langle dE/dx \rangle/\langle dE/dx \rangle = 5\%$ was measured for particles ranging from $H$ to ${}^{4}$He, over the range of momenta measured in the calibration beam as summarized in Table~\ref{tb:momresolution}.
+
+Since the magnetic dipole setting of the BIGRIPS spectrometer defined the energy of each particle type, we can calculate the expected momenta of each particle species measured.
+Small corrections were propagated using LISE++ software which can calculate the energy loss through several materials in the beam line.
 These corrections resulted in a small change in the momenta.
 The measured momenta of the calibration beam differed significantly from the expected values as seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
-This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocity in the direction of $\vec{E}\times\vec{B}$ direction.
-The $\vec{E}\times\vec{B}$ dift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
+This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocity in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
+The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
 The disagreement in measured and expected momenta is upwards of ~5\% difference in the higher momentum calibration settings.
-The details of the correction technique are discussed in the later Section~\ref{sec:spacecharge} are discussed in a more general correction which also includes correction for the space charge; the same correction technique was applied here in the special case of zero space charge which is the special case of only having $\vec{E}\times\vec{B}$ components .
-
-The values under the corrected column of seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data correcting for $\vec{E}\times\vec{B}$.
+The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed in the later Section~\ref{sec:spacecharge} in a more general way which also includes correction for the space charge, which this cocktail beam does not have.
+The same correction technique was applied here in the special case of zero space charge, which is the special case of only having $\vec{E}\times\vec{B}$ components.
+
+The values under the corrected column seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data corrected for the $\vec{E}\times\vec{B}$ effect.
 A significant improvement is seen in the high momentum setting going from around ~5\% disagreement to within ~1\% agreement in the corrected data.
@@
 
-Picture of cocktail before and after ExB effect
-Table of LISE++ expected cocktail energies ridigity setting of dipole magnets (reference big rips line)
+%Apendix of the beam line elements
+%apendix of the settings of the bigrips elments
+%Picture of cocktail before and after ExB effect
+%Table of LISE++ expected cocktail energies ridigity setting of dipole magnets (reference big rips line)
 
 \subsection{Electronics calibration}
-The channel number of the electronics was calibrated by measuring the response of each channel to an input pulse supplied by a pulse generator.
-The pulse was injected into the ground plane of the TPC.
-This distributed the pulse evenly across the entire pad plane over a range of input voltages.
+The ADC channel number of the electronics was calibrated by measuring the response of each channel to an input pulse supplied by a pulse generator.
+The pulse was distributed to all the electronics channels by pulsing the ground plane for a range of input voltages.
+This distributed the pulse evenly across the entire pad plane.
 The input voltage is plotted as a function of the measured ADC channel in Fig.~\ref{fig:gaincalib} for every channel.
@@
 The right panel shows the resulting distribution of channels after calibration.
-This is a relative calibration technique meant to calibrate the varying gains in each channel relative to one another.
+This is a relative calibration technique with the intent to calibrate the varying gains in each channel relative to one another.
 
@@
 By lowering the voltage on these anode wires, the gas gain is lowered as compared with all the other anode wire plane sections which operate at \SI{1460}{\volt}.
-To account for the drop in gain, in the software we increase the gain of the pads which lie above these anode wires.
-To calibrate these sections we perform a relative calibration to the high gain anode sections by comparing the energy loss values of the high and low gain sections.
+To account for the drop in gain, in the software we increase the gain of the pads which lie above these anode wires to match the gain in the other channels.
+To do so, we compare the energy loss in the high gain sections to the low gain sections to get the relative calibration factor.
 
@@
 
+Figure~\ref{fig:highlowcal} shows the correlation plot of $\langle dE/dx\angle$ of the high versus the low gain sections.
+The effects of the high gain channel saturation can around a value of 400 ADC\si{\per\milli\metre} and plateaus, where as the low gain does not.
+A linear fit was performed for the calibration  between the high gain channel $h_c = G\cdot l_c$, where $l_c$ is the value of the low gain channel and $G$ is the gain factor.
+In the case where the low anode voltage was \SI{1214}{\volt}, the gain factor was 9.8 where as in the case of \SI{1085}{\volt}, the gain factor was ???.
+The calibration factor was applied inside of the decoder task.
+A map of the channels which are impacted by the low anode wire voltages and the relevant gain was input into the software.
+In the decoder task the raw ADC spectra was multiplied by the input gain for a given channel.
+The thresholds of those channels also were multiplied by the same factor since the noise levels were also artificially magnified by the same gain factor.
+
 \subsection{Extending the dynamic range of the Electronics}
@@
 Typically in higher energy experiments the charge of particles is Z$e$, where Z=1.
-Also the particles are traveling at higher energies in which the energy losses in a detector range from the minimum of the energy loss curve to the small logarithmic climb CITE FIGURE HERE!!!.
+Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum in the energy loss curve.
 The dynamic range of electronics in such experiments can cover a wide range of particle energies in the energy loss curve.
-In nuclear HICs, we are interested in cluster of nucleons typically from Z=1-3 resulting from the collision, and even higher in some applications.
-As seen in Eq.~\ref{eq:bb}, the energy loss is proportional to $Z^2$; the energy values of the resulting particles are at much lower energies which are in the $1/\beta^2$ region.
-The energy range and particle types covered by the electronics are significantly limited as they quickly reach the dynamic range limitations of the electronics as the charge of a particle increases and the velocity decreases.
+In nuclear HICs, we are interested in measured particles ranging from Z=1-3 resulting from the collision, and even higher in some applications.
+As seen in Eq.~\ref{eq:bb}, the energy loss is proportional to $Z^2$; the energy values in nuclear HIC of intermediate beam energies (around \SI{300}{\MeVA}) in which the resulting particles are at much lower energies which are in the $1/\beta^2$ region.
+The energy range and particle types covered by the electronics are significantly limited by the dynamic range limitations of the electronics as the charge of a particle increases and the velocity decreases.
 
@@
 There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these persistent issues in TPC electronics CITE HERE, by being able to switch to a lower gain value when the maximum range is reached.
-Though, it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of external hardware, espeically in experiments which have already been performed with older electronics technologies.
-
-In TPCs, the effective dynamic range can be very different from the single pad dynamic range.
+Though, it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of external hardware, especially in experiments which have already been performed with preexisting electronics technologies.
+
+In TPCs, the effective dynamic range can be very different from the single channel dynamic range depending on the application.
 Typically TPCs are operated inside of a magnetic field for reconstructing momentum of each particle, which requires sub-millimeter precision in the position determination of the track path.
-This is achieved by averaging the charge distribution over several pads which is discussed in Section CITE HERE.
+This is achieved by averaging the charge distribution over several pads.
 Therefore the effective dynamic range is related to the relative charge values of adjacent pads (between the center pad and outside pads).
-For example, to measure minimum ionizing particles the signal to noise ratio of the pad with the smallest charge in the distribution should be some reasonable value, say at least 6:1.
-From Section CITE HERE we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
-If we require the adjacent pads to have a signal-to-noise ratio of 6:1, then the central pad would have a signal to noise ratio of 50:1.
-Considering this is the signal-to-noise ratio for minimum ionizing particles, and the maximum signal-to-noise ratio is 800:1, this means the effective dynamic range in the TPC is roughly 16 times that of minimum ionizing particles.
+For example, to measure minimum ionizing particles the signal to noise ratio of the pad with the smallest charge in the distribution should be some reasonable value above noise, say at least 6:1 signal-to-noise ratio (SNR).
+From Section \ref{sec:prf}, we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
+If we require the adjacent pads to have a SNR of 6:1, then the central pad would have a SNR of 50:1.
+Considering this is the SNR for minimum ionizing particles, and the maximum SNR in a channel is 800:1, this means the effective dynamic range in the TPC is roughly 16 times that of minimum ionizing particles, if we intend to have adequate position and tracking resolution.
+
 The dashed lines and vertical blue bar in Fig.~ \ref{fig:intro} are separated by a factor of 20, representing the typical effective dynamic range in a TPC.
@@
 Similar problems are encountered when TPCs are used as active targets in direct reaction studies with rare isotope beams \cite{pattpc}.
-
-Several techniques have been employed to increase the observable range of energy losses.
-This can be done by lowering the electronics gain of selected readout channels, or by changing the gas amplification at the readout plane in certain areas of the TPC.
-In the EOS TPC \cite{eos} this was done by decreasing the voltages in select anode wires in the multi-wire readout.
-With the prototype Active Target TPC lowering or increasing the gain was achieved by decreasing or increasing the electric fields on selected pads within a Micromegas \cite{pattpc}.
+Several techniques have been employed to increase the observable range of energy losses; typically by lowering the electronics gain of selected readout channels, or by changing the gas amplification at the readout plane in certain areas of the TPC \cref{eos,pattpc}.
 The results of changing the gas-gain, or the electronics gain, are rather similar in that reducing the gain to sample a range of higher energy loss makes the TPC effectively blind to minimum ionizing particles in  the regions of lower gain.
 
-%DOUBLY DEFINED? IS THIS SUPPOSED TO BE HERE
-%\begin{figure}[ht!]
-%\includegraphics[scale=.5]{top_view_helix_ext.pdf}
-%\caption{Cartoon of a top down view of a fit to a track passing through several pads.
-The bolded pads and the charges $q_i$ represent the hits belonging to that pad and the clusters of the track representing the average position of the track.
-The three clusters at the bottom are clustered in the $x$-direction for the upper three are clustered in the $z$-direction.
-The estimated position of the avalanche is given by the track fit, and the position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
-%\label{fig:topview}
-%\end{figure}
-
-We define the clustering direction depending on the angle of the track, at the point of each cluster, with respects to the $x$ axis, defined as $\theta$.
-For example, the crossing angle is defined as $90^{\circ}$ for a track going along the $z$ axis, and $0^{\circ}$ for a track going along the $x$ axis.
-In the case that the crossing angle is $45^{\circ} < \theta \leq 90^{\circ} $ the clustering direction is along the $x$ axis.
-For $0^{\circ} < \theta \leq 45^{\circ}$ it is along the $z$ axis.
-
- The position along the clustering direction is calculated by weighting the individual hit's positions by their charges $q_i$ and getting the mean value.
-The other direction is set to the center of the pad.
-For example, if we are clustering along the $x$ axis for a cluster, the $z$-position is set to the center of the pad in the $z$-direction and vice versa.
-
 \subsection{Experimental Pad Response Function}
 
-The correlations we introduced by only clustering along one direction do not play a significant role in the particle identification, but cause deviations from the expected Gatti distribution.
-Also, analytic PRFs only exist for classical multi-wire TPCs.
-For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
-
+The pad response function of the TPC was extracted from non-saturating hits and clusters in tracks at various track crossing angles.
 As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as $\lambda_i = x_i-\bar{x}$.
@@
 
+\begin{comment}
 \begin{figure}[ht!]s
@@
 \end{figure}
-
-The shape of the PRF depends on the crossing angle of the track \cite{gatti}.
-Plotted in Fig.~\ref{fig:prfpimData} is the PRF of $\pi^-$ tracks vs.
-the crossing angle $\theta$.
-The PRF gets wider starting from $90^{\circ}$  and going to $45^{\circ}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
-Since we switch the clustering direction from $x$ to the $z$ direction at $45^{\circ}$, the opposite trend is seen where the PRF becomes narrower as the position resolution gets better going from $45^{\circ}$ to $0^{\circ}$.
+\end{comment}
+
+The shape of the PRF depends on the crossing angle of the track \cite{gatti}, which determines how wide the charge is distributed along the wire.
+Figure~\ref{fig:prfpimData} shows the PRF of $\pi^-$ tracks versus the crossing angle $\theta$ of the track.
+The PRF gets wider starting from $90^{\circ}$  and until where we switch clustering directions at $\ang{45}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
+Switching from $x$ to the $z$ direction clustering the opposite trend is seen where the PRF becomes narrower  going from $45^{\circ}$ to $0^{\circ}$, as the position resolution gets better.
 
@@
 
-Fits were performed to the experimental data with  $5^{\circ}$ width bins from $0^{\circ} < \theta \leq 90^{\circ}$.
-The two parameters of the Gaussian fits are plotted versus $\theta$ in Fig.~\ref{fig:normsigma}; a $4^{th}$ order polynomial fit between these points allowed for interpolating between $\theta$.
+A Gaussian fit was performed were performed to the experimental data with  $\ang{5}$ width bins from $\ang{0} < \theta \leq \ang{90}$.
+Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function given in Eq.~\ref{eq:prfgaus} are plotted versus $\theta$.
+A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value as seen by the black line.
 
 \subsection{Method of Desaturation}
-We will use the term ``desaturation'' for our process of correcting the charge values of the saturated pads.
-Figure \ref{fig:satpad} shows a typical situation of saturated signals.
-When an avalanche causes a large induced signal, the pads directly underneath collect the largest charge becoming saturated, denoted as $q_{2'}$ and $q_{3'}$.
-Pads further away experience smaller, non saturating charges, denoted as $q_{1}$ and $q_{4}$.
-Though we do not know the saturated charge values, the distribution of all charges must follow the PRF which we have experimentally measured.
-From the clusters crossing angle, we can get the corresponding parameters for the PRF as described above and in Fig.~\ref{fig:normsigma}.
-
-We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as given by the $PRF(\lambda_i)$ function.
-
-\begin{equation}\label{eq:chi}
-\chi^2 = \sum_i \frac{(q_i^{\mathrm{obs}} - q_i^{\mathrm{expect}})^2}{q_i^{\mathrm{expect}}}
-\end{equation}
-
-To determine the best estimate for the charge values of each saturated pad, a chi squared function is minimized, given in  Equation \ref{eq:chi}, where $q_i^{\mathrm{obs}}$ are the observed, non-saturated charges $q_{1}$ and $q_{4}$, and $q_i^{\mathrm{expect}}$ are the charges we expect to observe, calculated as $q_i^{\mathrm{expected}} = Q\cdot PRF(\lambda_i)$.
-The charges $q_{2'}$ and $q_{3'}$, make up the unknown variable and are allowed to vary in the $\chi^2$ minimization, where they are added to make up the total expected charge $Q$.
 
@@
 
-Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
-
-\begin{figure}[ht!]
-\includegraphics[width=\linewidth]{cocktail.png}
-\caption{Pad plane projection for a cocktail event in the TPC.}
-\label{fig:cocktail}
-\end{figure}
-
-\begin{figure}[ht!]
-\includegraphics[width=\linewidth]{data.png}
-\caption{Pad plane projection for a collision event in the TPC.
-Highlighted by red arrows are two regions of anode wires which had a reduced voltage of 1214 V.
-The voltage of the rest of the TPC anode wires are 1460 V.
-The reduction in voltage reduces the gain by a factor of about 10.
-}
-\label{fig:data}
-\end{figure}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=\linewidth]{dedx_compare}
-\caption{The left panel shows the high gain stopping power vs low gain when the method of desaturation was not applied.
-In the right panel the desaturation technique was applied to the high gain region.
-The low gain does not suffer from saturation and represents the true $dE/dx$ value.}
-\label{fig:lowvshigh}
-\end{figure*}
-
-\begin{figure*}[t]
-\includegraphics[width=\linewidth]{cocktail_combine.png}
-\caption{Uncorrected (left panel) and desaturated (right panel) cocktail data.}
-\label{fig:cocktail_combine}
-\end{figure*}
+We will use the term ``desaturation'' to describe the process of correcting the charge values of the saturated pads.
+Figure~\ref{fig:satpad} shows a typical situation of saturated signals in a given cluster of hits.
+When an avalanche causes a large enough induced signal, it is the pads directly underneath which collect the largest charge becoming saturated; denoted here as $q_{2\textprime}$ and $q_{3\textprime}$.
+Pads further away collect less charge and typically are not saturated; the non-saturated pads are detonated here as $q_{1}$ and $q_{4}$.
+Though the true charge value in the saturated channels are lost, we know that the distribution of all charges must follow the PRF which is a fixed feature of all TPCs, and has been experimentally measured above.
+From the tracking information, we know a given clusters crossing angle and can interpolate to get the corresponding parameters for the PRF described in Fig.~\ref{fig:normsigma}.
+
+We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as given by the $PRF(\lambda_i)$ function.
+To determine the best estimate for the charge values of each saturated pad, a chi squared function is minimized,
+
+\begin{equation}\label{eq:chi}
+\chi^2 = \sum_i \frac{(q_i^{\mathrm{obs}} - q_i^{\mathrm{expect}})^2}{q_i^{\mathrm{expect}}},
+\end{equation}
+
+where $q_i^{\mathrm{obs}}$ are the non-saturated charges and $q_i^{\mathrm{expect}}$ are the charge values expected from the PRF calculated as $q_i^{\mathrm{expected}} = Q\cdot PRF(\lambda_i)$.
+The saturated charge values $q_{i}^{\textprime}$ are treated as unknown variables and are allowed to vary in the $\chi^2$ minimization; the values calculated by the minimizer are added to the total expected charge $Q$.
 
@@
 
-Tracks which saturate pads in the high gain region are not saturated in the low gain region.
-By comparing the dE/dx values of these two sections, we can directly measure the success of the desaturation in the high gain regions using the method described above.
-
-In Fig.~\ref{fig:lowvshigh}, the effect of saturation can be seen in the high gain region for the uncorrected data.
-For signals below 400 ADC/mm \footnote{Un-calibrated ADC channels in arbitrary units.} the electronics are not saturated, and therefore the high and low gain sections agree.
-The data starts to saturate above 400 ADC/mm in the high gain channels eventually reaching a plateau while the low gain sections are not saturated and provide the true $dE/dx$ values.
+Tracks which saturate pads in the high anode wire voltage region are not saturated in the low anode voltage region.
+By comparing the $\langle dE/dx\rangle$ values of these two sections, we can directly measure the success of the desaturation in the high gain regions.
+While this desaturation technique avoids the need to lower the gain of any region, the low anode voltage region proved to be a direct measurement of the success of this technique.
+
+Figure~\ref{fig:lowvshigh} shows the the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
+The effect of saturation can be seen in the high gain region for the uncorrected data above values of 400 ADC/mm where the values plateau, where as the low gain region still returns accurate values\footnote{Un-calibrated ADC channels in arbitrary units.} Below this value the electronics are not saturated, and therefore the high and low gain sections agree.
+
  After applying the desaturation method, the correlation between the high and low gain sections is restored, as seen in Fig.~\ref{fig:lowvshigh}.
-From this comparison, we infer that the correction works up to signals of 2000 ADC/mm, increasing the dynamic range by a factor of at least 5.
-
-Comparing the low to high gain sections directly validates the desaturation technique, but the goal  of this exercise is to improve the particle identification (PID).
+From this comparison, we can approximate that the correction has corrected the high gain sections to agree with low gain sections to values of 2000 ADC/mm, increasing the dynamic range by a factor of at least 5.
+
+The success of the desaturation becomes more clear when looking at the particle identification (PID) lines.
 In the following PID plots the red lines represent the most probable energy loss as given by Geant4 straggling functions.
-A linear calibration was performed to convert keV in Geant4 to ADC in the experiment given by $ADC/mm = 19\;keV/cm$.
+A linear calibration was performed to convert keV in Geant4 to ADC in the experiment given by $ADC/mm = 19;keV/cm$.
 
 There are pronounced PID lines of several particle species in both the uncorrected and corrected cocktail beam PID shown in the subplots of Fig.~\ref{fig:cocktail_combine}.
-Three ovals around a momentum of 1700~MeV/$c$/q and two near 900~MeV/$c$/q correspond to the three $B\rho$ settings injected into the TPC.
-The tails of the PID lines are resulting from the particles passing through the walls and other materials outside the main detector volume, therefore lowering the initial momentum.
-
-The uncorrected data in Fig.~\ref{fig:cocktail_combine} shows the effects of saturation; the PID lines deviate from their theoretical expectations starting at around 400~ADC/mm eventually reaching a plateau.
-After applying the desaturation technique, we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation.
+The uncorrected data in the left panel shows the effects of saturation; the PID lines deviate from their theoretical expectations starting at around 400~ADC/mm eventually reaching a plateau.
+After applying the desaturation technique -- in the right panel-- we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation and even ${}^{6}Li$ and ${}^{7}Li$ can be separated.
 A more subtle improvement of the lighter particles, (p, d, t), can also be seen in the PID lines at lower momenta.
 
-Looking at the collision data, shown in Fig.~\ref{fig:data_combine}, we also see a similar result.
-In the collision data, the PID suffers from more background and inefficiencies than the cocktail beam, nevertheless we can see a similar improvement in the PID lines when comparing before and after applying desaturation.
+A similar effect is seen in the experimental collision data.
+In Figure~\ref{fig:data_combine} though the PID suffers from more background and inefficiencies than the cocktail beam we see a similar improvement in the PID lines when comparing before -- the left panel-- and after applying desaturation -- the right panel.
 Notably the largest improvement is the separation of particle species at lower momenta and the separation of the Li species into ${}^{6}$Li and ${}^{7}$Li.
@@
 The dynamic range was extended by at least a factor of 5, as demonstrated by the improved PID lines, and quantified by direct comparison to low gain sections of the TPC.
-This improved PID will allow for us to extend the momentum distributions of all species to lower momenta and to heavier ions than what was previously available.
+
+%Give some failures of the assumptions made
 
@@
 
-MAYBE ADD A LITTLE NAPKIN CALCULATION OF SPACE CHARGE TO SHOW ORDER OF EFFECT.
+%MAYBE ADD A LITTLE NAPKIN CALCULATION OF SPACE CHARGE TO SHOW ORDER OF EFFECT.
+
+As the beam passes through the field cage it ionizes the gas creating electron-ion pairs.
+The drift velocities of the ions are typically \SI{e4} times slower than the electron drift velocities \cite[blumrol].
+Any source of ions have the potential to build up creating a positive space charge distorting the drifting electrons, biasing the track momentum.
+There are several regions of the TPC in which ion are created.
+The largest source of positive ions is created in the avalanche process near the anode wires.
+These ions slowly drift toward the cathode.
+If the gating grid is closed the ions will terminate there; though it opens occasionally  for \SI{11}{\micro\second} the ions only move \SI{1}{\micro\metre} in this time and will be captured by grid when it closes.
+For this reason, the source of ions resulting from the avalanche region is negligible.
+
+The other source of ions come from the primary ionization produced by the beam and reaction products in the detector gas.
+The energy loss  $\langle dE/dx\rangle \propto Z^2$, where Z is the charge of the particle type.
+Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2.5e3} times that of the light charged particles which mostly are of charge Z~1.
+It is the largest source of positive ions in the TPC.
+
+NEED TO PUT IN ABOUT THE GATING GRID LEAK
+
+ \begin{comment}
 
@@
 \end{table}
-
-\begin{figure}[H]
+\end{comment}
+
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{spacechg_cartoon.png}
@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{beampath.png}
@@
 
-As the beam passes through the field cage it ionizes the gas along the way creating electron-ion pairs.
-The drift velocities of the ions are \num{1e4} times slower than the electron drift velocities.
-Because ions move very slow, the potential build up of positive ions may create a space charge which distorts the drifting electrons of the tracks, impacting the measurement of their momentum.
-There are several regions of the TPC in which electron-ion pairs are created.
-The largest source of positive ions are created in the avalanche process near the anode wires.
-These ions slowly drift toward the cathode and are captured by the gating grid.
-The other source of ions come from the primary ionization produced by the beam and reaction products in the detector gas.
-The energy loss <dE/dx>$\propto Z^2$, where Z is the charge of the particle type.
-Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2.5e3} times that of the light charged particles which mostly are of charge Z~1.
  The beam is positioned about 25\si{\centi\metre} below the anode plane in the TPC.
-It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane.
-It takes the ions \num{5e4}\si{\micro\sec} to travel.
-The beam rate in the experiment varied around a value of about 10\si{\kilo\hertz}, which has an average occurrence of 1 beam every 100\si{\micro\sec}, which is much shorter than the time it takes for the ions created by each beam to terminate on the cathode plane.
-This results in a build up of positive ions in the shape of a sheet charge, carved out by the beam path as shown in Fig.~\ref{fig:spacechg_cartoon}.
-The average distance between sequential ion paths created by each beam is a spacing of about 50\si{\micro\metre} apart, and the average number of beam paths that compose the sheet charge is around 500 tracks.
-Though the arrival time of each beam track is random, the large number of tracks, and small inter beam spacing, allows us to approximate the sheet charge as an uniform sheet charge.
-
-The electric field in the presence of the sheet charge can be calculated by solving Poisson's equation,
+It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane where as it takes the ions \num{5e4}\si{\micro\sec}.
+The beam rate in the experiment varied around a value of about 10\si{\kilo\hertz}, which has an average occurrence of 1 beam every \SI{100}{\micro\sec}; this is shorter than the time it takes for the ions to terminate on the cathode plane, resulting in a build up of positive ions.
+Figure~\ref{fig:spacechg_cartoon} gives an idea of the shape of the sheet of space charge carved out by the beam path.
+The ions from each beam create a line charge which drifts towards the cathode with a constant velocity.
+The average distance between sequential ion paths is about \SI{25}{\micro\metre} apart, therefore we expect the average number of beam paths that compose the sheet charge is around \num{1440} tracks -- where the distance to the cathode is \SI{29.6}{\centi\metre}.
+Though the arrival time of each beam track is random, the large number of tracks and small inter beam spacing allows us to approximate the sheet charge as an uniform sheet charge.
+
+ The secondary beams entering the TPC are also composed of many species of particles distributed in a finite range around the target beam.
+Assuming a beam of ${}^{132}$Sn were the energy loss in gas is \SI{11.2}{\kilo\electronvolt\per\centi\metre}, using the number of beams above, the conversion to electrons given in \ref{tb:gasprop}, and the beam path length of \SI{135}{\centi\metre}, the total charge density expected is \SI{3e-8}{\coulomb\per\metre\squared}.
+
+The electric potential can be calculated by solving Poisson's equation,
 
@@
 
- where $\phi$ is the electric potential and $\rho$ is the free space charge; given the Dirichlet boundary conditions of the field cage.
-The electric field is given as the gradient of the potential,  $\vec{E}= -\nabla \phi$.
-
- To reduce computation time, we notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a reference free charge $\rho_o$ and scale the solution for any other free charge.
+ where $\phi$ is the electric potential and $\rho$ is the free charge.
+Since the Dirichlet boundary conditions of the field cage are defined by the cathode, walls, and pad plane, a numerical solution to the electric potential is solved through the ???
+technique.
+The electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.
+
+ To reduce the need to compute a new electric field for a new value of the space charge, we notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ by the ratio $\rho/\rho_o$.
 The full magnetic field map is provided by the SAMURAI collaboration \cite{magnet}.
-The velocity field map is given by Eq.~\ref{eq:elecdrift} and the electron drift through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration.
-The correction map is calculated by starting from the anode y-position and the measured (x,z) on the pad-plane and stepping backward in time in the Runge-Kutta integration through the velocity field map until the electron reaches the measured y-position.
+The velocity field map can be calculated from Eq.~\ref{eq:elecdrift}.
+The path an electron drifts through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration from a given starting point of the electron.
+We do not add details of the electric field around the anode wires which is only relevant for the small gap between the ground and anode plane.
+Over this small gap the space charge corrections have very little effect.
+
+ The correction map is calculated by starting from the anode y-position and the measured position on the pad-plane (x,z), and stepping backward in time in the Runge-Kutta integration through the velocity field map until the electron reaches the measured y-position, which is split into a grid that fills the 3-dimensional volume of the field cage.
+
+ The correction is handled within the STCorrection task in the software.
+A data base of the sheet charge values for every run is read in and the electric field is calculated for the reference charge for a given beam path.
+The electric field and velocity map is scaled by the ratio of the space charge in the particular run and the reference run $\rho/\rho_o$ and the correction map is calculated.
+The measured clusters value (x,y,z) position is input into the correction map which interpolates and outputs the correction values $dx$ and $dz$ which correspond to that cluster.
+The cluster position is shifted to new positions $x\textprime = x + dx$ and $z\textprime = z + dz$, before going on to the momentum and vertex finding algorithms.
 
  It has been shown before that the amount of space charge present in the chamber is related to observables such as the distance of closest approach of each track to the vertex point \cite{starSC}.
-This is easily understood, as the space charge distorts the electrons as they drift through the chamber eventually distorting the overall shape of the measured track.
-In the presence of no space charge, you would correctly expect the distance of closest approach of each track to the vertex point would be a distribution centered around zero.
-In the presence of space charge the tracks are distorted and the distribution has some mean bias.
-
-Shown in Fig.~\ref{fig:sc_shift} is an example of the distortion map in the TPC for left-going tracks in blue, and right-going tracks in green, with the vectors of distortion magnified.
-The space charge affects left and right going tracks differently, with the right-going tracks going to higher momentum values and the left-going tracks going to lower momentum values.
+In the presence of no space charge, you would correctly expect the distance of closest approach of each track to the vertex point would be a distribution centered around the true vertex location.
+Since the tracks are distorted by the space charge, which affects different regions of the TPC differently, a bias is introduced to the measured vertex location and widens the distribution to vertex of each track.
+
+An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} with left-going tracks shown in blue, and right-going tracks shown in green; the vectors show the direction of distortion and their magnitude have been magnified by ???
+times to show the detail.
+The dashed lines show the effects the track shifted due to the space charge map where left and right-going tracks are affected differently.
+Right-going tracks tend to higher momentum values and the left-going tracks going to lower momentum values, for positively charged particles.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{Effect_SC.pdf}
+\caption{Shift of tracks}
+\label{fig:sc_shift}
+\end{figure}
+
 The inset figure of Fig.~\ref{fig:sc_shift} shows the shift in the x-position distance to vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
-We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR}$, see Fig.~\ref{fig:VLR}.mom
-
-Overview
-Discuss the relevant time scales, drift lengths, magnitudes, and locations of space charge
+We will define the DOCA for left going tracks as $V_x^L$ and for right-going tracks as $V_x^R$.
+We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ as seen in Figure~\ref{fig:VLR}.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{PeakLocation_vs_BR.pdf}
-\caption{$\Delta\mathrm{V}_\mathrm{LR}$ as a function of the beam rate. }
-\label{fig:spacechg_br}
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{VLR_beamrate.png}
+\caption{ $\Delta V_{LR}$ versus the beam rate for all systems with the fitted function.}
+\label{fig:vlr_br}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{sc_beamrate.png}
+\caption{Space charge fit function versus the beam rate for all systems. These are the assumed functions we use for interpolating the space charge value from the beam rate.}
+\label{fig:spacechg_br_all}
 \end{figure}
@@
 The average beam rate was recorded in each experimental run and slightly varied from run to run, due to beam production variations.
-The amount of space charge present in the field cage is directly proportional to the beam rate; therefore $\Delta\mathrm{V}_\mathrm{LR}$ is also proportional to the beam rate as shown in Fig.~\ref{fig:spacechg_br}.
+The amount of space charge present in the field cage is directly proportional to the beam rate; therefore $\Delta\mathrm{V}_\mathrm{LR}$ is also proportional to the beam rate as shown in Fig.~\ref{fig:VLR_br}.
 The only parameter in the space charge correction algorithm is the surface charge density $\sigma_{\mathrm{SC}}$.
-By varying $\sigma_{\mathrm{SC}}$ for a wide range of values the $\Delta\mathrm{V}_\mathrm{LR}$ observable is measured and plotted in the left panel of Fig.~\ref{fig:spacechg_relation}.
+Varying $\sigma_{\mathrm{SC}}$ for a wide range of values the $\Delta\mathrm{V}_\mathrm{LR}$ observable is measured and plotted in the left panel of Fig.~\ref{fig:spacechg_relation}.
 The surface charge density which gives $\Delta\mathrm{V}_\mathrm{LR} = 0$ is taken to be the estimate for the average amount of space charge present.
-This is done for several runs which vary in beam intensity.
-Since the relation is linear between the surface charge density and the beam rate, a linear fit gives good agreement for interpolating the surface charge values as a function of beam rate, as seen in the right figure of Fig.~\ref{fig:spacechg_relation}.
-
-This is done for all the systems. I NEED TO PUT THE ASSUMPTIONS of SPACE CHARGE SHEET DENSITY OF 132 108 HERE!!!
-
-From the BDC tracking, the precise interaction point on target is known.
-Using this point greatly improves the momentum resolution of the track fitting, but there is a systematic shift when comparing the momentum value without using the vertex BDC point in the fit.
-The space charge affects right-going and left-going tracks differently making right-going tracks more rigid (higher momentum) and left-going tracks less rigid (lower momentum).
-For polar angles of $\theta < 40 \deg$, the disagreement between momentum values fitted with the BDC, versus without, are much less because the projection of the track does not disagree as much with the BDC point as tracks with polar angles of $\theta > 40 \deg$.
-This is seen in
-
-\begin{figure}[H]
+
+This is done for several runs which vary in beam intensity though the solution for the estimated space charge value will be different.
+Since the surface charge density is proportional to the beam rate, a linear fit gives good agreement for interpolating the surface charge values as a function of beam rate.
+Figure~\ref{fig:spacechg_relation} shows the relation of the dependence of the space charge as a function of beam rate for the ${}^{132}$Sn system.
+
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{SC_Relation.pdf}
@@
 
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{Effect_SC.pdf}
-\caption{Shift of tracks}
-\label{fig:sc_shift}
-\end{figure}
+Following this algorithm the space charge is estimate for each run and system.
+Figure~\ref{fig:scDensity} shows the summary of the extracted space charge values for each secondary beam.
+The linear fits relating the space charge density to the beam rate for each system is shown in Fig.~\ref{fig:spacechg_br_all} where the space charge value is inferred from the measured beam rate.
+Notice that in the $\tin{112}{124}$ system a constant value of was assumed since the beam rate did not span a large enough range to warrant a more detailed analysis.
+
+%What is the best figure to put in? VLR vs Charge density? 
+%estimate of the error of VLR and how it corresponds to charge density
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{spaceChgDensity.png}
+\caption{Distribution of estimated space charge densities for each beam type.}
+\label{fig:scDensity}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{VLRresidual.png}
+\caption{Residuals in the fitted line of $\Delta V_{LR}$ observable for all systems.}
+\label{fig:vlrResidual}
+\end{figure}
+
+Adding the BDC vertex greatly improves the momentum resolution of the track fitting, but there is a systematic shift when as compared with the momentum value without using a vertex as an extra constraint.
+Since the space charge affects right-going and left-going tracks differently making right-going tracks differently than left-going tracks, it appears they no longer originate from the BDC which is not affected by the space charge.
+For tracks at polar angles of $\theta < 40 \deg$, the disagreement between momentum values with and without the BDC are much less because the projection of the track does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
+Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ are more sensitive to small changes in the BDC when including it as an extra constraining point.
+After correcting for the space charge, adding the BDC point as an extra constraint on the track does not bias the reconstructed momentum value for either $\theta < 40 \deg$ or $\theta > 40 \deg$, which has the greatest improvement, as seen in Fig.~\ref{fig:mom_S_after} and Fig.~\ref{fig:mom_L_after} respectively.
+Recall that only the relative distance between left and right-going tracks was minimized; there was no guarantee that the corrected tracks coincide with the absolute BDC position at the target.
+This is one of the evidences of the correction's success.
+The others being the agreement with the expected cocktail calibration beam as described above.
+Later we will show how the space charge corrects the momentum distributions of left and right-going pions in the TPC.
 
@@
 
+\begin{comment}
+
 \begin{figure}[!htb]
@@
 \end{figure}
-
-Reference appendix for poisson solver 
-Tables for ion drift velocity in P-10 Gas reference Sauli
-Figure of sDAC or POCA 
-Figure of cartoon of what is happening to tracks
-Figure of correction map in TPC and MC map 
-Figure of before and after correction BDC vs reco momentum
-Figure of track residuals before and after?
-
-In theory, a TPC functions with the electric and magnetic fields parallel to each other.
-In this way the electrons move opposite to the electric field winding tightly around the magnetic field lines, reducing transverse diffusion in the process.
-In practice, due to the finite size of the dipole magnet and field cage, there are traverse components to both the electric and magnetic fields.
-These transverse components introduce drift velocities in the transverse directions, causing a shift in the measured cluster positions of the track.
-Thereby introducing systematic shifts in the calculation of the momentum and the vertex calculation.
-
-Most of the time the beam does not undergo a nuclear collision with the target and passes through the TPC drift volume.
-The KATANA array threshold was set to veto such events, ensuring the gating grid remain closed to prevent the large amount of charge deposited by the beam into the avalanche region.
-While the electron drift velocity is fast enough for the electrons produced by the unreacted beam to terminate on the closed gating grid, the drift velocities of the positive ions produced are of order $10^{-4}$ times slower.
-At higher beam rates the positive charge is allowed to pile up producing a region of positive space charge, introducing perturbations to the nominal electric field.
-
-Since the beam comes in along the z-axis, and the drift direction of the ions is in the -y direction, we can estimate the charge density as an uniform sheet charge.
-The surface charge density is related to the beam rate, ion drift length, ion drift velocity, and the energy loss of the beam.
-Though the incoming beam comes randomly, the slow drift velocity combined with the high beam rate makes the uniform approximation valid.
-Tracking or estimating the beam rate as a function of time with in a given run would provide a better estimate of the space charge.
-Or experimentally a laser system could be pulsed after each event (throughout the drift volume), giving the experimental correction map for the drift.
-While the potential for a laser system was implemented in the field cage design, a laser system ultimately was not developed for the \spirit tpc.
-
-The beam rate within a run is roughly constant, therefore we can estimate the space charge and provide a first order correction for the space charge effect.
-
-As mentioned in CHAPTER ???, the cocktail beam momentum was well known to within 1\% as set by the BIGRips spectrometer.
-Also time of flight analysis of upstream and downstream scintilators also independently confirmed the beam rigidity setting.
-The expected momentum is given in TABLE ???
-as calculated by from the beam rigidity setting of the magnet (WHICH MAGNET).
-The measured momenta as determined by the TPC software (given in FIG TABLE ), shows a disagreement on the level of 5\% too high.
-We noticed that if one only uses the first 90 layers (out of 112) of the TPC, the momenta is lower; one should expect the momentum to go higher as the track length is shorter (short tracks effectively are straight lines).
-In the cocktail beam there is no un-reacted beam causing any significant space charge.
-The magnetic field map of the SAMURAI magnet has been calculated by TOSCA simulation CITE ???, and several points have been verified experimentally with a hall probe to be within XXX \%.
-We assume that the electric field (to first order) is uniform in the y-direction.
-Using Garfield++ CITE ???, we can model the transverse drift of the electrons in the presence of such fields.
-
-A grid of electrons uniformly distributed in the TPC model space were drifted to the gating grid position. The final shifts in the x and z positions was measured.
+\end{comment}
+
+%Reference appendix for poisson solver 
+%Tables for ion drift velocity in P-10 Gas reference Sauli
+%Figure of sDAC or POCA 
+%Figure of cartoon of what is happening to tracks
+%Figure of correction map in TPC and MC map 
+%Figure of before and after correction BDC vs reco momentum
+%Figure of track residuals before and after?
 
 \section{Monte Carlo Simulation}
-The monte carlo (MC) simulation is a to part simulation.
-The first is the modeling of the fundamental interactions of a particle through material which we utilize Geant4 as an event generator.
+The monte carlo (MC) simulation is composed of two separate simulations.
+The first simulation is an accurate model of the fundamental interactions of a particle passing through the various materials in the TPC; to accomplish this we have modled everything in Geant4 which also provies a particle generator.
 A scale model of field cage, front window, front window frame, pad plane, and aluminum top plate are modeled with the correct material type and dimensions assuming P-10 gas at a density of 1 atm.
-We also import the the magnetic field map of the SAMURAI dipole magnet (as calculated by the SAMURAI group via a TOSCA simulation) \cite{magnet}.
-In this way any particle type may be studied and the full interactions (scattering, decay, energy loss, path taken, etc.), are accounted for.
-The output of this simulation is a series of energy loss points which contain the amount of energy lost in $keV/cm$ and the position in $(x,y,z)$.
-The second part of the MC simulation simulates the physical processes involved in the measurement of the TPC such as the drift, avalanche, and signal creation in the electronics, to simulate the detector response of the MC event.
-This is separated into several software tasks; the drift, pad response, and electronics task.
-
-Here we will only discuss in detail, the tasks involved in simulating the TPC response.
-The discussion of how we use Geant4 is not particularly interesting, since we use it only to get the energy loss in a gas and through basic materials.
+The magnetic field map of the SAMURAI dipole magnet was imported into Geant4 as well after assuming rotational symmetry along the axis perpendicular to the pole face.
+Along with the energy loss and particle transport, Geant4 also handles multiple scattering, and particle decay.
+The output of the particle generator is a series of energy loss points which contain the amount of energy lost in $\si{\kilo\electronvolt\per\centi\metre}$ and the location of the energy loss in Cartesian space $(x,y,z)$.
+
+In the second part of the MC simulation the physical processes involved in the operation and measurement of the TPC are simulated, such as the electron drift, avalanche process, and all processes involved in the signal creation in the electronics; to simulate the detector response of the MC event.
+This is separated into several software tasks; the drift, pad response, and electronics tasks.
+In the following section we will discuss in detail the tasks involved in simulating the TPC response.
+The discussion of Geant4 is not covered here and is left to the reader to explore.
 
 \subsection{Drift Task}
-Geant4 returns the primary ionization points and energy loss information in the gas volume. In the drift task this is converted into the total number of electrons, $N_{e^-}$,
-
+ The first step of the drift task is to convert the primary ionization points of the MC track, provided by Geant4, into the total number of electrons, $N_{e^-}$.
+The average number of electrons created in a gas can be described as,
 \begin{equation}
@@
 
-where the ionization coefficient of P10 gas, $I$, given in Table~\ref{tb:gas} and the energy loss, $\Delta E$, given in \si{\electronvolt}.Since the electric field set up by the field cage is uniform for the area covered by the pad plane, we assume also each electron drifts in a straight line trajectory until reaching an anode wire, where the total length drifted is given by $L_{anode}$.
-In this task the option to put in space charge effects and non-uniform electric fields are allowed but we do not use them for the further discussion of calculating efficiency.
+where the ionization coefficient of P10 gas, $I$, (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited in units of \si{\electronvolt}.
+Since the electric field set up by the field cage is uniform for the most of the area covered by the pad plane, we assume a straight line trajectory for each electron until it reaches an anode wire.
+The total length drifted from the initial primary ionization point to the final anode wire is given by $L_{anode}$.
 
@@
 
-The drifting electrons exhibit stochastic motion which can be described by a longitudinal, $c_{l}$, and transverse, $c_{t}$, diffusion coefficients, which are determined by Garfield++ calculations -- in the presence of a 0.5T magnetic field.
-The diffusion of electrons can be modeled by randomly sampling from a Gaussian distribution in both the longitudinal and transverse directions of the electron drift.
-The stochastic deviation in the transverse direction, $dr$, can be randomly sampled from,
+Drifting electrons exhibit stochastic motion which can be described by a a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
+The longitudinal ,$c_{l}$, and transverse ,$c_{t}$, diffusion coefficients are determined by Garfield++ calculations in the presence of a \SI{0.5}{\tesla} magnetic field, given in Tb.~\ref{tb:gas}.
+The diffusion process is modeled by randomly sampling the from a Gaussian distributions describing both directions and adding the shifts to the electron final position.
+The stochastic deviation in the transverse direction, $dr$, is randomly sampled from,
 
@@
 
-where $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.The Cartesian directions are given by $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, as there is no preferential angle in the transverse directions.
-The longitudinal diffusion is simulated in the same manner by randomly sampling from,
+where $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.The Cartesian directions are given by $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, since there is no preferential angle  of emission in the transverse plane.
+The shift associated with the longitudinal diffusion, $dl$, is randomly sampled from,
 
 \begin{equation}
-dt = e^{-\frac{t^2}{2\sigma_{l}^2}},
-\end{equation}.
+dl = e^{-\frac{t^2}{2\sigma_{l}^2}},
+\end{equation}
 
@@
 \begin{equation}
- t = L_{anode}/v_d + dt + t_{offset},
+ t = \frac{L_{anode} + dl}{v_d} + t_{offset},
 \end{equation}
 
- where $v_d$ is the drift velocity and $t_{offset}$ is an offset parameter added to align the MC time bucket spectrum with the data.
-The electron is assumed to terminate on the nearest anode wire which will fix the z-position to be the same as the that anode wire.
-
-The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++, Fig.~\ref{fig:gaincalib}.
-Two anode wire voltages were applied which correspond to two different anode gains, high and low, in the TPC experiment.
-We randomly sample from these two distributions, depending on which anode wire section the electron terminated on and store the number of electrons as an effective gain value.
-We do not assume there is much lateral movement along the wire in the avalanche process.
+ where $v_d$ is the drift velocity.
+The parameter $t_{offset}$ aligns the MC time bucket spectrum with the data; the data time bucket zero does not correspond to the same y-position in the MC for the same time bucket position.
+The electron is assumed to terminate on the nearest anode wire which fixes the z-position that of the anode wire.
+
+The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++ as seen in Fig.~\ref{fig:gaincalib} for the anode wire voltages used in the experiment.
+The number of electrons produced in the avalanche is randomly sampled from the corresponding distribution depending on which anode wire section the electron terminated; the number of electrons produced is stored as a number instead of multiple instances of the original electron to save space and computation time.
+We assume that the lateral movement along the wire arising form $\vec{E}\times\vec{B}$ effect near the wire is negligible as compared with diffusion and other effects described later.
+
+It is worth mentioning there is the possibility to simulate the space charge effects in this class where the inverse to the correction map described in \ref{sec:spacecharge} is applied to the drifting electrons.
+This is not used for calculating the response of the TPC for efficiency calculations since it is a trivial exercise to input the space charge map and correct for it using the same inverse map.
 
 \subsection{Pad Response Task}
-The total charge of the event is split over several pads defined by the PRF, as discussed in section \ref{sec:prf}.
-We simulate the PRF as the double integral of a 2-dimensional Gaussian with the final output charge on each pad being the superposition of the PRFs of all drifted electrons.
+The total charge of each avalanche must be distributed according to the pad response function (PRF) described in section \ref{sec:prf}.
+The PRF is simulated as the double integral of a 2-dimensional Gaussian.
+The final output charge on all the pads are the superposition of the PRFs of all drifted electrons.
 The MC PRF is expressed as,
@@
 where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the final position of the drifted electron.
-The $\sigma$ parameters were determined through an iterative comparison between the MC and experimental data set.
+The $\sigma$ parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
 We can see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
+It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.
+Therefore these effects must arise from geometric effects related to the track angle, the amount of charge distributed over an anode wire, and the superposition of the PRF from neighboring anode wires which contribute to the appearance of a changing PRF.
 
@@
 \subsection{Electronics Task}
-%The dynamic range, $d_{range}$, of the TPC electronics was set to 120 $\si{\femto \coulomb}$ over the full 12 bit resolution resulting in 4096 channels, $ADC_{max}$.
-The pedestal ($ADC_{pedestal}$) was set to 300 channels.
-The response of an electronics signal is simulated by measuring the standard pulse response to an input signal in the data.
-It was found the pulse shape did not depend on the momenta, angle, or particle type in any significant way.
-Therefore a standard pulse was extracted from an average over all particle types and signal heights -- as long as it were not saturating-- shown in Fig.~\ref{fig:pulsea}.
-The superposition of the individual pulses from each electron makes up the Monte Carlo (MC) time bucket spectra in ADC.
-This can be directly fed into the software.
-The experimental electronic noise (RMS) was measured to be around 6 ADC; a random number --drawn from a Guassian with a sigma of 6 ADC and 0 mean -- was assigned to each time bucket to simulate the experimental noise.
 

### 2020-05-21 chapters/ch_experiment.tex
Routine upload

@@
 
-\begin{equation}\label{eq:gaus}
+\begin{equation}\label{eq:prfgaus}
 PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
@@
 
-%\section{Collision Data Taken}
+\section{Collision Data Taken}
+
+Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
+tb
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{cocktail.png}
+\caption{Pad plane projection for a cocktail event in the TPC.}
+\label{fig:cocktail}
+\end{figure}
+
+\begin{figure}[ht!]
+\includegraphics[width=\linewidth]{data.png}
+\caption{Pad plane projection for a collision event in the TPC.
+Highlighted by red arrows are two regions of anode wires which had a reduced voltage of 1214 V.
+The voltage of the rest of the TPC anode wires are 1460 V.
+The reduction in voltage reduces the gain by a factor of about 10.
+}
+\label{fig:data}
+\end{figure}

### 2020-05-21 macrosForThesis/GatingGrid/Noise/plotGGNoise.C

@@ -1,0 +1,82 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotGGNoise()
+{
+
+  TFile *raw = TFile::Open("adc_raw.root");
+  TFile *corr = TFile::Open("adc_corr.root");  
+  TFile *mean = TFile::Open("meannoise.root");  
+
+
+  TH1D *adcraw = (TH1D *)raw->Get("rawdist");
+  TH1D *adcorr = (TH1D *)corr->Get("adcdist");
+  TH1D *meanfcn = (TH1D *)mean->Get("mean");
+
+  meanfcn->SetLineWidth(3);
+  meanfcn->SetLineColor(kRed - 4);
+
+  adcraw->SetTitle("");
+  adcorr->SetTitle("");
+
+  TLegend *leg = new TLegend(.5,.65,.8,.8);
+  leg->AddEntry(meanfcn,"Mean Value","l");
+  leg->SetBorderSize(0);
+  
+  TCanvas *c1 = new TCanvas("c1","c1",1000,800);
+  c1->SetTopMargin(.15);
+  c1->SetBottomMargin(.18);
+  c1->SetRightMargin(.05);
+  c1->SetLeftMargin(.12);
+
+  c1->Divide(1,2,0,0);
+
+
+   
+  //  adcraw->GetXaxis()->SetTitle("V_{avg} (V)");
+  //  adcraw->GetXaxis()->CenterTitle();
+  adcraw->GetYaxis()->SetTitle("Raw ADC");
+  adcraw->GetYaxis()->CenterTitle();
+
+  adcraw->GetXaxis()->SetTitleSize(.08);
+  adcraw->GetYaxis()->SetTitleSize(.08);
+  
+  adcraw->GetXaxis()->SetLabelSize(.05);
+  adcraw->GetYaxis()->SetLabelSize(.05);
+  adcraw->GetXaxis()->SetTitleOffset(.6);
+  adcraw->GetYaxis()->SetTitleOffset(.6);
+
+  adcorr->GetXaxis()->SetTitle("Time Bucket");
+  adcorr->GetXaxis()->CenterTitle();
+
+  adcorr->GetYaxis()->SetTitle("Subtracted ADC");
+  adcorr->GetYaxis()->CenterTitle();
+
+  adcorr->GetXaxis()->SetTitleSize(.08);
[…]

### 2020-05-21 macrosForThesis/SpaceCharge/SpaceChargeDensity/plotBRVLR.C

@@ -1,0 +1,157 @@
+#include "/home/justin/mythesis/thesis_style.h"
+using namespace style;
+
+void plotBRVLR()
+{
+  gStyle->SetOptStat(0);
+
+
+
+
+  vector<double> sn108_br = {11702.11,10798.15,10776.19,11415.63,10983.03,11735.92,10494.24,10323.1,12436.16,11470.08,11206.14,11189.46,10495.89,11417.21,11004.11,11155.06,11430.47,10300.64,9517.32,9567.52,10020.06,9319.23,10340.03,10792.4,11314.69,9600.99,10128.85,10005.5,9889.38,9670.79,9793.15,10048.46,9566.91,8927.96,9006.96,8942.42,9500.53,8812.99,9285.82,9037.49,7092.41,6132.07,5961.7,6766.93,7872.51,7349.74,6970.62,7900.94,7122.43,5909.41,7949.15,7430.78,8725.24,11026.43,9923.22,10511.85,9591.99,9181.09};
+
+  vector<double> sn108_vlr = {9.48,9.45,9.38,9.31,9.46,9.33,8.86,9.28,9.68,9.45,9.17,9.54,9.62,9.18,9.55,9.3,9.17,9.08,8.9,8.93,8.79,8.74,8.45,9.14,9.12,9.22,9.12,9.13,8.82,9.06,9.04,8.91,8.79,8.84,8.76,8.74,8.56,8.72,8.53,8.63,8.01,7.59,7.48,7.46,8.21,7.89,7.95,7.98,7.89,7.46,7.89,7.86,7.91,9.24,8.84,8.71,8.66,8.63};
+
[…]

### 2020-05-21 macrosForThesis/SpaceCharge/SpaceChargeDensity/plotSCRelation.C

@@ -1,0 +1,85 @@
+#include "/home/justin/mythesis/thesis_style.h"
+using namespace style;
+
+void plotSCRelation()
+{
+  gStyle->SetOptStat(0);
+  int sn132_c = kRed - 4;
+  int sn108_c = kBlue - 4;
+  int sn112_c = kGreen - 4;
+  int sn124_c = kCyan - 4;
+
+
+  TF1 *l_132 = new TF1("l_132","[0] + [1]*x",0,13);
+  TF1 *l_124 = new TF1("l_124","[0] + [1]*x",0,12);
+  TF1 *l_112 = new TF1("l_112","[0] + [1]*x",0,13);
+  TF1 *l_108 = new TF1("l_108","[0] + [1]*x",0,13);  
+
+
+ TLegend *leg = new TLegend(.7,.2,.9,.4);
+  leg->AddEntry(l_132,"{}^{132}Sn","l");
+  leg->AddEntry(l_124,"{}^{124}Sn","l");
+  leg->AddEntry(l_112,"{}^{112}Sn","l");
+  leg->AddEntry(l_108,"{}^{108}Sn","l");
+    
+  TCanvas *c2 = new TCanvas("c2","c2",1500,900);
+  c2->SetRightMargin(.1);
+  c2->SetLeftMargin(.15);
+  c2->SetTopMargin(.1);
+  c2->SetBottomMargin(.15);
+  c2->SetFrameLineWidth(5);
+  gStyle->SetLineWidth(5);
+  gPad->SetLineWidth(5);
+
+  l_132->SetLineColor(sn132_c);
+  l_108->SetLineColor(sn108_c);
+  l_112->SetLineColor(sn112_c);
+  l_124->SetLineColor(sn124_c);
+
+  l_132->SetLineWidth(10);
+  l_108->SetLineWidth(10);
+  l_112->SetLineWidth(10);
+  l_124->SetLineWidth(10);
+
+  l_132->SetLineStyle(10);
+  l_108->SetLineStyle(10);
+  l_112->SetLineStyle(10);
+  l_124->SetLineStyle(10);
+
+  l_108->SetParameters(-1.066,2.939e-1);
+  l_132->SetParameters(1.344,2.79e-1);
+  l_124->SetParameters(2.546,0);
+  l_112->SetParameters(-.4394,2.9655e-1);
+
+  l_132->Draw();
+  l_132->GetXaxis()->CenterTitle();
[…]

### 2020-05-30 chapters/ch_data_analysis.tex
Update analysis section

@@
 \subsection{Method of Desaturation}
+\label{sec:desat}
 
@@
 
-In the second part of the MC simulation the physical processes involved in the operation and measurement of the TPC are simulated, such as the electron drift, avalanche process, and all processes involved in the signal creation in the electronics; to simulate the detector response of the MC event.
-This is separated into several software tasks; the drift, pad response, and electronics tasks.
-In the following section we will discuss in detail the tasks involved in simulating the TPC response.
-The discussion of Geant4 is not covered here and is left to the reader to explore.
+In the second part of the MC simulation, the physical processes involved in the operation and measurement of the TPC are simulated, such as the electron drift, avalanche process, and all processes involved in the signal creation in the electronics, simulating the detector's response to a MC track.
+This is separated into three software tasks, the drift, pad response, and electronics tasks.
+In the following section we will discuss these tasks in more detail, the discussion of Geant4 is not covered here and is left to the reader to explore.
 
 \subsection{Drift Task}
- The first step of the drift task is to convert the primary ionization points of the MC track, provided by Geant4, into the total number of electrons, $N_{e^-}$.
-The average number of electrons created in a gas can be described as,
+ The first step of the drift task is to convert the primary ionization points of the MC track provided by Geant4 into electrons.
+The average number of electrons created in a gas,$N_{e^-}$, can be described as,
 \begin{equation}
@@
 
-where the ionization coefficient of P10 gas, $I$, (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited in units of \si{\electronvolt}.
+where $I$ is the ionization coefficient of P10 gas (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited in units of \si{\electronvolt}.
 Since the electric field set up by the field cage is uniform for the most of the area covered by the pad plane, we assume a straight line trajectory for each electron until it reaches an anode wire.
@@
 
-Drifting electrons exhibit stochastic motion which can be described by a a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
-The longitudinal ,$c_{l}$, and transverse ,$c_{t}$, diffusion coefficients are determined by Garfield++ calculations in the presence of a \SI{0.5}{\tesla} magnetic field, given in Tb.~\ref{tb:gas}.
-The diffusion process is modeled by randomly sampling the from a Gaussian distributions describing both directions and adding the shifts to the electron final position.
-The stochastic deviation in the transverse direction, $dr$, is randomly sampled from,
+Drifting electrons frequently collide with the detector gas causing them to change direction.
+This stochastic motion is  described by a a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
+The longitudinal ($c_{l}$) and transverse ($c_{t}$) diffusion coefficients are determined by Garfield++ calculation in the presence of a \SI{0.5}{\tesla} magnetic field, listed in Tb.~\ref{tb:gas}.
+The diffusion process is modeled by randomly sampling from a Gaussian distributions describing both of the directions.
+The random shift is then added to the final position of the electron.
+The deviation in the transverse direction, $dr$, is randomly sampled from,
 
@@
 
-The total drift time, $t$, is calculated as,
+The final position of the electron along the wire is calculated  $x' = x + dx$.
+Since the electron will terminate on the closest anode wire the anode wire that is closest to the shifted z-position $z\textprime = z + dz$, the final electron z position is then updated to the anode wire it terminated on.
+The total drift time of the electron $t$ is calculated as,
 
@@
  t = \frac{L_{anode} + dl}{v_d} + t_{offset},
+ \label{eq:electronTime}
 \end{equation}
@@
  where $v_d$ is the drift velocity.
-The parameter $t_{offset}$ aligns the MC time bucket spectrum with the data; the data time bucket zero does not correspond to the same y-position in the MC for the same time bucket position.
-The electron is assumed to terminate on the nearest anode wire which fixes the z-position that of the anode wire.
+The parameter $t_{offset}$ aligns the MC time bucket spectrum with the data.
+This is because the y-position corresponding to time zero in the data time bucket spectra does not correspond to the same position in the MC for time zero.
+This value was found to be \SI{23}{\nano\second} by aligning the MC vertex to the data vertex.
 
@@
 The $\sigma$ parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
-We can see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
-It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.
-Therefore these effects must arise from geometric effects related to the track angle, the amount of charge distributed over an anode wire, and the superposition of the PRF from neighboring anode wires which contribute to the appearance of a changing PRF.
-
-\begin{figure}[!htb]
-         \centering
-         \includegraphics[width=\textwidth]{PRFs_MC_wcut.png}
-         \caption{PRF response of Monte Carlo simulation of $\pi^-$}
-         \label{fig:prfpimMC}
-\end{figure}
 
@@
 
-The electronics task simulates the electronics response, converting  the induced charge on a pad into ADC channels.
-The conversion from number of electrons to the response, $R$, in ADC channels is calculated as,
-
-\begin{equation}
-R = A_{tot} \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{f_c}
-\label{eq:etoADC}
-\end{equation},
-
-where $e$ is the typical charge of an electron in $\si{\femto \coulomb}$, $A_{tot}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
-We add random Gaussian noise as measured from the experimental simulating the electronic noise; (RMS) was measured to be around 6 ADC.
-
+The purpose of the electronics task is to simulates the electronics response to the charge induced on each pad, converting charge into ADC channels.
 The induced current in the pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
-The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, for a given shaping constant.
-This allows us to assume the pulse shape is constant with the only two variables being the height of the pulse, $Q$, and the starting time bucket of the pulse, $t_o$.
-The shaping constant was set to \SI{117}{\nano\second} for the data analyzed here.
-Figure~\ref{fig:pulseshape} shows the pulse shape experimentally extracted from signals that did not saturate the electronics over a wide range of ADC values; it is normalized so the maximum height is 1.
+The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, as long as the electronics settings are fixed.
+This allows us to assume the pulse shape is constant and can be described by two variables, the height of the pulse, $Q$, and the starting time bucket of the pulse, $t_o$.
+The starting time of the pulse is defined as the time at 10\% of the rising edge.
+The shape of the pulse depends on the shaping time constant which was set to \SI{117}{\nano\second} for the data analyzed here.
+Figure~\ref{fig:pulseshape} shows the pulse shape which was extracted from the experimental signals in the data.
+These were signals that did not saturate the electronics and was averaged over a wide range of ADC values.
+Here it is normalized so the maximum height is 1.
 
@@
 
+Converting the charge in each pad into the height of the ADC response in the electronics is calculated as,
+
+\begin{equation}
+Q = f_G  A_{tot} \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{f_c}
+\label{eq:etoADC}
+\end{equation},
+
+where $e$ is the charge of an electron in $\si{\femto \coulomb}$, $A_{tot}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
+The pulse shape given in Fig.~\ref{fig:pulseshape} is multiplied by the pulse height $Q$ giving the full time bucket estimate of the TPC response.
+Random Gaussian noise is added to each time bucket where the root-mean-squared value of the electronics noise was measured to be around 6 ADC.
+The timing information of the pulse is calculated from Eq.~\ref{eq:electronTime}.
+The coefficient $f_G$ corresponds to a factor which allows for fine tuning of the ADC calibration.
+This factor is calculated by
+
+\begin{equation}
+f_G^p = \frac{\langle dE/dx\rangle_{Data}^p}{\langle dE/dx\rangle_{MC}^p}
+\label{eq:dedxcalibration}
+\end{equation}
+
+where $\langle dE/dx\rangle_{MC}^p$ and $\langle dE/dx\rangle_{Data}^p$  are the energy loss values for a particle in a given momentum bin $p$. The calibration will be discussed later in more detail.
+
 \subsection{Simulating Saturation}
-
-%Saturation of the electronics manifests itself in two different ways.
-Some channels are dead for the full time bucket range in an event.
-These channels are randomly distributed along the beam path.
-Though the gating grid of the TPC is closed, an un-reacted beam produces several delta electrons.
-The electrons emitted in directions along the magnetic field lines can physically pierce this grid and avalanche on the anode wire causing channels along the beam path to be saturated.
-The recovery time depends on the input charge and can take several ms to recover, killing the pad for future events.
-%Tracks depositing enough energy may saturate an electronics channel for the event shadowing tracks that may cross below.
-
-%From the time a channel has been saturated, no further signals can be read out.
-To simulate this behavior when embedding signals from Monte Carlo tracks, we must first identify when in a pad a saturating signal has occurred.
-Non-saturating signals have a characteristic long exponential tail where as saturated signals have no long tail and prematurely end.
-By applying a software pole-zero correction technique, the long tail of the non-saturated signal can be removed.
-In the case of saturated pules this creates a negative undershoot which we use to identify when a channel has been saturated.
-
-%Before embedding any MC signals into the software, the pole-zero compensation is applied to see when in time the saturation has occurred.
-The point in which MC signals cannot be embedded is set to 30 Tbs prior to this negative undershoot to prevent any MC signals from superimposing upon the saturated signal.
-Another routine identifies if a pad is ``dead" for the event and no MC signals may be embedded.
+\label{sec:simSat}
 
@@
 The picture is complicated by the fact that there is usually an RC feedback loop which dissipates the input charge.
-The time in which a pre-amplifier returns to its typical linear behavior depends on the input charge and how quickly it can be dissipated \cite{akiGET}.
-The pad is otherwise considered dead and no further charge can be measured until the electronics recover.
-
-Due to the long high energy tail in energy loss distributions of a particle traversing matter, it is common to see pads along the track which have very high energy loss values and saturate the electronics of that pad, even for minimum ionizing tracks.
-Saturation in this case is infrequent and not an issue as many other clusters are not saturated.
-As the particle's charge gets higher, or the momentum gets lower, the mean energy loss value gets much higher and a significant fraction of pads in a track may be saturated.
-While earlier we outlined and algorithm to correct the saturation for the track measurement itself, saturation also affects surrounding tracks.
-
-As mentioned earlier the pad is dead for some time, which depends on the total input charge.
-For charges that are relatively at or above the threshold for saturation the pad is certainly dead for the remainder of time of the measurement; i.e.
-$T_{dead}$ > \SI{10}{\micro\second}.
-This will cause signals from other tracks passing underneath the offending track to not be measured; the saturated pad ``shadows" any future signal.
-In the case where two tracks are separated in time due to different polar angles, but have significant overlap in a pad-plane projection, a significant portion of the later track may be shadowed and the track itself will suffer in quality significantly.
-For events with high track multiplicity this causes a significant shadowing effects, in that the upward going tracks recorded at earlier times shadow the downward going tracks recorded at later times.
+The time a pre-amplifier returns to linear behavior depends on the input charge and how quickly it can be dissipated \cite{akiGET}.
+The pad is otherwise considered dead and no further charge can be measured until the channel can recover.
+
+Due to the long, high energy tail in  the energy loss distribution of a particle traversing matter, it is common to see pads along the track which have collected large energy loss values saturating the electronics of that pad.
+This occurs for all tracks, even for minimum ionizing tracks; saturation in this case is infrequent and not an issue as there are many other clusters are not saturated.
+Yet as the charge of the particle gets higher, or the momentum gets lower, the most probable energy loss value also becomes higher, and a significant fraction of pads in a track may be saturated.
+In Section~\ref{sec:desat}, we outlined an algorithm to correct for the saturation for a track, where as here we are interested in the effects saturation has on surrounding tracks to properly simulate the effect in the MC simulation.
+
+Pads that saturate have no signal or ''dead" for a given amount of time which depends on the amount of input charge.
+For charges that are relatively at or above the level of saturation, the pad will certainly be dead for the remainder of time after the saturating event.
+Signals that would have arrived later in time from other tracks passing underneath the saturating track will therefore not be measured; the saturated pad is ``shaddowing" any future signals.
+As the track multiplicity of an event increases, the probability earlier tracks shaddow later tracks increases.
 This will change event by event and is very difficult to simulate except through a MC track embedding approach which will be discussed later.
@@
 In the case where the input signal is very large, the electronics can be dead for up to \SI{35}{\milli\second} \cite{akiGET}.
-The beam rate in the experiment was about \SI{10}{\kilo\hertz}, which corresponds to an average of \SI{10}{\micro\second} between subsequent beams.
-Pads can be effectively dead for several events before recovering.
-Due to the large charge of the beam, a large amount of very high energy electrons are produced via scattering from the beam passing through the gas.
-In the presence of the magnetic field the radius of curvature of most of the electrons are within several \si{cm}.
-Figure~\ref{fig:deltaE} shows the horizontal extent of the electrons in a top down view of the TPC.
-While some of the electrons can stop in the gas, a significant fraction can travel to the top and bottom of the TPC.
-These electron can pass through the gating grid without being blocked and either terminate on the pad or possibly deposit their charge directly in the pad.
-The charge induced on the pad was large enough to kill the pad for a time long enough to last until at least the next event.
-These pads are dead for the whole event and are randomly distributed around the beam path projection on the pad plane.
+The beam rate in the experiment was about \SI{10}{\kilo\hertz}, which corresponds to an average of \SI{100}{\micro\second} between subsequent beams.
+In this case channels can be effectively dead for several events before recovering.
+Due to the large charge of the beam, a large amount of very high energy electrons --delta rays-- are produced from the beam passing through the gas.
+In the presence of the magnetic field the radius of curvature of most of the electrons are within several \si{cm} in the xz-plane.
+Figure~\ref{fig:deltaE} shows the horizontal extent of the delta rays in a top down view of the TPC.
+While some of the electrons can stop in the gas, some electrons have a high enough kinetic energy going in the vertical direction, penetrating the gating grid without being blocked.
+They then either terminate on the pad or possibly deposit their charge directly in the pad.
+The charge induced on the pad is large enough to kill the pad for a time long enough to last until at least the next event.
+This type of saturation is randomly distributed for pads sitting directly in the beam path and are dead for that event and the next event.
 
@@
 
-Simulation of saturation is handled naturally by embedding MC tracks into real data, by correctly identifying when in the time bucket spectrum saturation has occurred in each pad.
-Once the time of saturation has been identified, no further signals are embedded as the pad is assumed to be dead.
-In the case a pad is dead for the whole event the time of saturation is set to the first time bucket and no signal is embedded.
-The characteristic signature of a saturated signal is the fast fall time of the falling edge of the pulse going quickly to zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
+The simulation of these saturation effects is handled naturally by embedding MC tracks into real data.
+Once the time bucket of the saturating signal has been identified, no further signals are embedded since the pad is assumed to be dead.
+In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded at all.
+
+The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
 The long exponential tail can be effectively removed by a simple software technique which is similar to the electronics concept of ``pole-zero compensation".
@@
 where $a_o$ = .9723, $a_1$=-.9453, $b_o$=.9545, and $b_1$=-.9203.
-These numbers are experimentally fit to give the fastest, non-zero, correction to the tail of a normal pulse.
-This correction produces a negative undershoot for the saturated pulse, since there was no long exponential tail to begin with.
-Figure~\ref{fig:pulseshape} shows before the correction technique as the red curves, and after the technique as the blue curves for both the normal and saturated pulse.
-As the program is looking for and identifying positive peaks in the spectrum, we also calculate the corrected ADC value of that particular time bucket according to Eq.~\ref{eq:satpolez}.
-A peak is identified as a saturated signal if it is less than -20$\cdot G$ for more than 8 Tbs, where $G$ is the gain calibration of the pad, and the max ADC value is > $G\cdot$500 to eliminate false negative pulses which come from the gating grid subtraction in dead pads.
-The saturated flag is set to true and the data time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket of the negative peak.
-The MC time bucket position is set to $t_{peak} - 30$ to ensure that the MC signals do not overlap the saturating data signal.
-
-Saturation can also occur if the sum of all the pulse heights are over the max ADC threshold of 3500 ADC.
-This is because the fall time of the pre-amplifier circuit is much longer than the time bucket window and pulses are allowed to pile up.
-While pads that saturate via this method still are identified through the same algorithm described above, nothing is done in the MC class to take this into account when embedding signals.
-We assume this method of saturation is a higher order correction and the current algorithm approximates most of the saturation effects.
-Figure~\ref{fig:satTag} shows the tagging algorithm for an event in the TPC.
-The max ADC values in a given pad are colored and the black dots are pads tagged as saturated.
-Some pads appear to be below 3500 ADC, but upon further inspection they satisfy the sum of all the signals causes saturation though the max ADC never goes over 3500.
-
-Dead pads are identified earlier in the software in the STCore.cc class.
-A dead pad is classified as having a r.m.s.
-value  < 50 ADC and a max ADC < 50.
-The saturated flag is set to true, and the time bucket position of saturation in both the data and MC are set to the first time bucket.
-
+These coefficients correspond to the case where the long tail of the standard non-saturating goes to zero, without producing any negative undershoots.
+In the case of applying this technique to the saturated pulse, the correction produces a negative undershoot since there was no long exponential tail to begin with.
+Figure~\ref{fig:pulseshape} shows the pulses as the red curves, and after the pole-zero compensation as the blue curves for both the normal and saturated pulses.
+
+By applying this technique to the real data we can identify whether a pad has a saturating pulse in it and what is the time bucket location by looking for this negative undershoot.
+A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ for more than 8 time buckets and the max ADC value is > $G\cdot$500 where $G$ is the gain calibration of the pad.
+The max ADC condition eliminates false positives which come from  dead pads where the gating grid subtraction is still applied introducing false negative peaks.
+The pad is then flagged as saturated and the data time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket of the negative peak, since the falling edge is around 5 time buckets after the beginning of the saturation.
+A separate time is also stored called the MC time bucket position which sets the point in time in which a signal cannot be embedded any further.
+This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible.
+
+Saturation can also occur if of all the pulse heights within the time bucket spectrum add to a value over the max ADC threshold of 3500 ADC.
+This arises because the fall time of the pre-amplifier circuit is much longer than the time bucket measurement window and pulses are allowed to pile up in the pre-amplifier.
+In this case no single pulse will reach over the max ADC value yet the last pulse to create saturation will have a pulse shape that has no long tail but otherwise look like a normal pulse.
+These pulses are also identified by the algorithm described above since there is still no long tail.
+When embedding tracks into pads, we do not consider the previous amount of total charge in the pad.
+We assume this method of saturation is a higher order correction and the current algorithm approximates most of the saturation effects which progress through the first process.
+Figure~\ref{fig:satTag} shows all of the pads tagged by the algorithm as saturated for an event in the TPC.
+The max ADC values in a given pad are colored and the black dots represent saturated pads.
+Some pads appear to be below 3500 ADC, but upon further inspection they satisfy the second process of saturation where the sum is greater than the max ADC value.
+
+Dead pads are identified earlier in the software in the STCore class.
+A dead pad is simply a pad which only contains electronic noise and no signals.
+A dead pad is identified by having a total ADC r.m.s.
+value  of less than 50 ADC and a max ADC < 50 over the full time bucket spectrum.
+Here the dead pad is also tagged as saturated and the time bucket position of saturation is set to the first time bucket.
+
+NEED TO ADDRESS WHY THE MAX ADC THRESHOLD IS 3500
 %maybe simple pream fig
@@
 \section{Monte Carlo Track Embedding}
- The TPC monte carlo simulation was built upon Geant4, including the dimensions, materials, and gas composition.
-Several classes were built to simulate the operations of a Time Projection Chamber (TPC).
-This included the drift properties of electrons, avalanche process, and electronics response to input signals.
-Combined, these classes simulate the response of the TPC to any input MC track and can be directly input into the \spirit TPC software reconstruction algorithm.
-
-While the goal of these tasks are to simulate important properties of a TPC, there are several effects which cannot be easily simulated.
-Several examples are biases due to triggering detectors (preferential selection of a reaction plane), saturation effects in the electronics, track multiplicity, and several other systematic sources common to all experiments.
-Simulating each issue individually would be impractical and time consuming, if not impossible.
-By embedding monte carlo tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all sources with a monte carlo type approach.
-
-Track embedding is the process of taking a simulated MC track from Geant4 and embedding its response into a real data event.
-After reconstructing this new embedded event we match the input MC track to its corresponding final reconstructed track.
-By doing so we can evaluate the response of the entire TPC system to any given input value.
-The TPC system is composed of three major components, each which can introduce errors and or biases; the software, the detector, and the experimental setup.
-
-As discussed in Section~\ref{sec:software}, the software is composed of several tasks, each introduce biases through assumptions made in the tracking algorithm.
+\label{sec:embedding}
+Track embedding is the process of taking a MC simulated track and embedding its response into an experimental data event.
+After reconstructing this new embedded event we match the input MC track to the corresponding final reconstructed track.
+By doing so we can evaluate the response of the entire TPC system to any given input.
+The TPC measurement can be thought of three different systems each which introduce errors and or biases into the final reconstructed tracks; the software, the detector, and the experimental setup.
+The
+
+There are several effects in a complex experimental set up that influence the data in either unknown or un-quantifiable ways.
+One such effect is the bias of the triggering system, here the Kyoto and Katana multiplicity and veto arrays preferentially select data that is emitted in a particular reaction plane.
+As previously discussed saturation effects in the electronics, notably the shadowing of other tracks.
+The effects of track multiplicity and the distribution of heavy ion and residues from the breakup of the target and projectile.
+One could propose to make a simulation or investigation into each individual effect in order to make a full MC simulation of data and study the response of the detector this way.
+Though great care was taken in developing an accurate MC simulation, like all models there are assumptions and simplifications that were made.
+Simply put, there is no better substitute for experimental data than itself.
+By embedding MC tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all these sources of biases which are contained in the experimental data, while at the same time measuring the response of the TPC.
+
+As discussed in Section~\ref{sec:software}, the software is composed of several tasks, each of which have the possibility of introducing errors or biases through assumptions made in the tracking algorithm.
 The detector system introduces errors related to the physical measurement itself which can be addressed through modeling the TPC, and its materials.
-The biases involved in the the experimental set up are not always obvious but present in the experimental data.
-All three of these can be simultaneously addressed by embedding of MC tracks.
-Before beginning  the detailed description of the embedding software, it is worth noting that the MC embedding was analyzed as a separate data set from the published data to avoid any issue of MC embedded tracks making it into the final published data.
+
+For all of the sources of biases in the TPC, it is clear that a MC type approach to solving for the response of the TPC would be the most practical method.
+Also by embedding MC events into real data we can take into account the effects of the experimental setup as well as the software and the detector system all together at once.
 
@@
 
-The detailed flow diagram of the software implementation of embedding is shown in Fig.~\ref{fig:flow}.
-Once the MC simulation of a track response in the TPC is created, we directly embed the MC signals into data by superimposing the MC signals into the data time bucket spectrum, pad-by-pad.
-Figure~\ref{fig:mcevent} shows the response of a 200 MeV/c $\pi^-$ track in the TPC; we will embed this track into the data event, Fig.~\ref{fig:dataevent}, for an example.
-We will refer to experimental data containing embedded MC signals as ``embedded data", data only containing the MC signals as ``MC data", and data only containing only the experimental signals as ``experimental data".
-The three set of data are each independently analyzed by the PSA algorithm, Section~\ref{sec:psa},  which finds all the hits in each data set.
-
-After this, the PSA algorithm contains the first and most important part of the embedding software.
-Here to the PSA task has the job to identify which hits in the embedded data are originally MC hits and which are from the original experimental data.
-Once identified, the embedded hits can be tagged and tracked through the entire software.
-By comparing the hits identified in the embedded set against the experimental data set we can first identify which embedded hits belong to the original experimental hit subset.
-For hits to match, they must satisfy two criteria, $\left|(Q_{\mathrm{Data}} - Q_{\mathrm{Embed} })/Q_{\mathrm{Data}}\right| < .05$ and $\left|t_{\mathrm{Embed} } - t_{\mathrm{Data} }\right| < 3$ where $Q$ and $t$ represent the charge and time of the hit respectively.
-We then remove these hits in the embedded data set which satisfy the conditions of experimental data hits.
+The detailed flow diagram of the software implementation of embedding algorithm is shown in Fig.~\ref{fig:flow}.
+Starting from Geant4 and passing through the 3 MC tasks described above, the output response of the MC track in the TPC is created.
+From here we directly embed the MC signals into data by adding the MC signals into the data time bucket spectrum, pad-by-pad.
+As described in Sec.~\ref{sec:simSat}, prior to embedding the MC signals, all pads which are saturated in the experimental data are identified and flagged.
+The MC signals are forbidden to be embedded after the MC saturation time bucket value.
+Figure~\ref{fig:mcevent} shows the response of a 200 MeV/c $\pi^-$ track in the TPC; we have embed this track into the data event, Fig.~\ref{fig:dataevent}, just as an example.
+We will refer to experimental data with embedded MC signals as ``embedded data" for short, the MC generated response as ``MC data", and data only containing only the experimental signals as ``experimental data".
+The three set of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
+From this the three independent hit data sets are temporarily stored.
+
+Here within the PSA algorithm, is the first and most important part of the embedding software.
+The PSA task has the job to identify which hits in the embedded data set originate from the MC hits set and which are from the experimental data.
+Once the MC hits are identified, these embedded hits can be tagged and tracked through the entire software.
+
+First, the hits in the embedded set are matched against the experimental data, identifying which of the embedded hits originate from the experimental hit set.
+For two hits to match, they must satisfy two criteria, $\left|(Q_{\mathrm{Data}} - Q_{\mathrm{Embed} })/Q_{\mathrm{Data}}\right| < .05$ and $\left|t_{\mathrm{Embed} } - t_{\mathrm{Data} }\right| < 3$ where $Q$ and $t$ represent the charge and time of the hit respectively.
+Hits that satisfy these criteria are then removed from the embedded data set.
 
 The surviving hits are then compared with the MC hit data set, where the criteria for a matching MC hit is, $\left|t_{\mathrm{Embed} } - t_{\mathrm{MC} }\right| < 3$.
-We do not require a matching criteria for the charge values since as not to bias the charge values selected as identified MC hits.
-This is only accomplished by first removing almost all of the experimental hits in the first step.
-Each embedded hit that passes this criteria is tagged as an MC embedded hit.
+There is no requirement for a matching criteria for the charge values which would artificially bias the charge values selected depending on our cut; which is critical for minimum ionizing particles such as pions.
+This can only be accomplished by first removing almost all of the experimental hits as was done in the first step.
+Each embedded hit that passes this step of criteria is tagged as a originating from a MC hit.
 From here, the embedded data is treated as if it were real data passing through the same software analysis.
 
-The goal of this na\"ive tagging is to preserve all the information of where the embedded hits, clusters, and tracks have gone.
-If a helix track has one hit that is embedded it will be tagged as embedded.
+ If a helix track has one hit that is embedded it will also be tagged as embedded.
 When the hits in a helix track are clustered, if a cluster contains one embedded hit it is itself tagged as an embedded cluster.
-Furthermore, if a track that is reconstructed contains any embedded clusters, it is an embedded track.
-It is the job of the embedded correlation task is to identify which of the final embedded tracks are candidates for the original input track.
-Several things may happen along the process of embedding.
-A track could break up, lose or share its charge with and adjacent track, or maybe not be identified at all for a variety of other reasons.
-For the embedded track to be a candidate of the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track and $N_{\mathrm{Total}}$ is the number of total clusters.
-The first criteria is a simple minimum cut where to ensure there exists an embedded track.
-The second criteria is the strongest cut ensuring the track has at least half of its clusters coming from embedded MC signals.
-The set of tracks which satisfy both conditions are saved into an array of tracks which the user can use to study efficiency or track splitting.
-
-In the case of efficiency analysis the user must select what they believe to be the track which most represents the original track.
-In this thesis, the track with the minimum distance to vertex is identified as the correct track.
-This is to get rid of the situation of track splitting where typically the split track has a distance to vertex which is very large and cut out from the analysis.
-The details of how the embedding is used to calculate the efficiency will be described later.
+Furthermore, if a track that is reconstructed contains any embedded clusters, it is tagged as an embedded track.
+The goal of this na\"ive tagging approach is to preserve all the information of where the embedded hits, clusters, and tracks have gone, preserving as much information until the end without introducing a bias.
+
+  It is the job of the embedded correlation task to identify which of the final embedded tracks are candidates for the original input track.
+For example, several things may happen along the process of embedding that may disqualify a track as a candidate in this na\"ive tagging.
+A track could break up, lose or share its charge with and adjacent track; it may be not be identified at all for a variety of other reasons.
+For the embedded track to be a candidate of the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track, and $N_{\mathrm{Total}}$ is the number of total clusters.
+The first criteria is a simple minimum cut where to ensure the minimum condition of a embedded track is met at least having 5 clusters.
+The second criteria is the strongest cut, ensuring the track has at least half of its clusters coming from embedded MC signals.
+The set of tracks which satisfy both conditions are saved into an array of candidate tracks.
+In this way the user can decide how to sort these tracks for studying either efficiency or track splitting.
+
+If the embedded track is split into multiple tracks, the user must select what they believe to be the track which most represents the original track, for the purposes of calculating efficiency.
+Typically only one of the tracks in the set will satisfy one of the reasonable quality cut conditions in an analysis.
+For example the most reasonable assumption is the track originating from the vertex is the most correct in this case, where as the split fraction of the track wont originate from the vertex but instead have a large distance to vertex value.
+In this thesis, the track with the minimum distance to vertex is identified as the track in which the efficiency analysis will be performed.
 
 \subsection{MC and Data Comparison}
+
+Here we compare several important observables to ensure the MC is simulating the data.
+These observables will be relevant for the quality cuts and efficiency analysis discussed in Section~\ref{sec:efficiency}.
+The most important observable is the total number of clusters.
+The number of clusters in a track depends on the geometry of the TPC, the spherical angles of the track, and the clustering algorithm.
+MC tracks were embedded into data for several particle types; the input angular and momentum distribution was uniform and the final distributions were weighted such that the angular and momentum distribution matched that of the data for a fair comparison.
+Figure~\ref{fig:clustcomp} shows the normalized cluster distributions for the MC and data for several particle types.
+There is no data below 20 clusters because this is the minimum cut to ensure quality PID lines in the data.
+The normalization is changed for each particle type to display them on the same scale.
+Very good agreement is seen between the MC and data over a wide range of particle species.
 
@@
 
+The second most important observable is the distance to vertex.
+Figure~\ref{fig:pocacomp} shows the distance to vertex distributions for the MC embedded tracks and the data for the same range of particle species.
+The distributions were again normalized to different values to display on the same scale.
+Both data sets here are track number of clusters > 20.
+It is also worth mentioning the data plotted is also after the space charge correction.
+Before the space charge correction the distributions would be much wider.
+Since we correct for the space charge in the data we compare here to the embedded MC without any space charge effects, making a fair comparison.
+
 \begin{figure}[!hbt]
@@
 
-Add Figure of Pad response function for pion,proton.... for MC vs Data vs angles...
-Add Figure of Number of clusters of MC vs Data
+We can see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
+It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.
+Therefore these effects must arise from geometric effects related to the track angle, the amount of charge distributed over an anode wire, and the superposition of the PRF from neighboring anode wires which contribute to the appearance of a changing PRF.
+
+\begin{figure}[!htb]
+         \centering
+         \includegraphics[width=\textwidth]{PRFs_MC_wcut.png}
+         \caption{PRF response of Monte Carlo simulation of $\pi^-$}
+         \label{fig:prfpimMC}
+\end{figure}
+
+%Add Figure of Pad response function for pion,proton.... for MC vs Data vs angles...
+%Add Figure of Number of clusters of MC vs Data
 Add Figure of dEdx MC vs Data
@@
 \section{Aligning TPC}
+
+%explain the different shifts in the TPC
+%BDC shift
+%Hit offshet shift
 
@@
 \section{Efficiency Corrections}
+\label{sec:efficiency}
 Add Figures of efficiency vs angles in TPC polar angle plot for pions
@@
 The rectangular design complicates the calculation of the geometric acceptance, or the efficiency.
+Because of this there are regions of the TPC where it is impossible to reconstruct a track due to the geometric in these regions the efficiency is exactly 0.
+In the regions of non-zero efficiency, the efficiency is a function of at least three main parameters that define the phase space of a track, momentum $p$ and the two spherical lab angles $\theta_{Lab}$ and $\phi_{Lab}$.
+
+The track efficiency $\epsilon$ is calculated bin-by-bin and is simply written as, 
+\begin{equation}
+\epsilon = \frac{n_{reco}}{N},
+\end{equation}
+
+where $n_{reco}$ is the number of embedded tracks which are successfully reconstructed, and $N$ is the total number of input embedded tracks for that given bin.
+A track is defined to be successfully reconstructed if it exists in the correlation task described in \ref{sec:embedding}, and therefore meets the minimum criteria for an embedded track, and if it passes the various quality cuts performed on the experimental data set; this could be angular cuts, PID cuts, etc.
+The TPC response will introduce a finite resolution in both the angles and the momentum.
+Because of the final measured track could have migrated out of the input bin into another bin.
+The bin-by-bin correction method is only valid if this effect is small as compared with the size of the bin.
+If not a more complicated un-folding procedure is needed.
+The efficiency is defined as the number of all tracks that were reconstructed and passed the cuts successfully, even if some had migrated to other bins, the efficiency value is assumed to represent the center of the input MC bin.
+%TALK ABOUT BIN SIZES HEREE?????
+
+ Single MC tracks are embedded into a set of \num{e4} on target events and the measured output tracks are correlated to the input track from the embedding software.
+If the track split in the analysis software, and there are several tracks identified by the software as an embedded track, or part of one, we certainly cannot double count the track as the efficiency is  defined strictly as $\epsilon < 0$.
+In this case, the track with the minimum distance to vertex is taken for the purposes of calculating efficiency.
+This is because in the case the track splits, it is very rare for the later segments of the track to have a distance-to-vertex which originates from the target region.
+It's reasonable to assume the track with the minimum distance to vertex is the track that best represents the ``real" track.
+The embedded MC tracks are generated as a uniform distribution in $\theta$,$\phi$, and $p$, to ensure each bin has as similar number of input tracks $N$, giving approximately the same statistical error.
+Here the efficiency distribution is independent of the initial input distribution since we were careful to not assign tracks that migrated out of the input bin, into another bin, which would make the efficiency biased by the input distribution.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_200_pol.png}
+        \caption{Polar efficiency plot where $\theta$ is represented by the radius of the circle and $\phi$ is measured as the polar angle made between the y and x axis; this view is  best understood as looking at the TPC from the downstream point of view.} \label{fig:pim_pol_eff_ex}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_200_efficiency.png} 
+        \caption{Flattened angle plot} \label{fig:pim_flat_eff_ex}
+    \end{subfigure}
+
+\caption{Efficiency calculations for $\pi^-$ particle traveling at \SI{200}{\mega\eVperc}.}
+\label{fig:pim_eff_ex}
+\end{figure}
+
+Figure~\ref{fig:pim_eff_ex} shows the efficiency calculated for $\pi^-$ tracks traveling with momentum \SI{200}{\mega\eVperc} as a function of the two laboratory angles.
+It is easiest to visualize the efficiency distribution first in a polar representation which is most like viewing the actual TPC progressing to a visualization that is easier to see the values.
+Figure~\ref{fig:pim_pol_eff_ex} shows the efficiency values plotted in a polar representation where $\theta_{Lab}$ is represented as the radial dimension and $\phi$ is represented as the polar angle.
+This way is best understood as if one looked at the TPC from the downstream point of view.
+Of course the vertical direction of the TPC,$\phi$=\ang{90}, is limited in vertical space, therefore tracks are not reconstructed well.
+The same is the case in downward going tracks, the region centered around $\phi$=\ang{270}.
+The best regions are seen by the high efficiency values which are left and right in the TPC where the rectangular TPC has the highest acceptance.
+Though this view is the most intuitive to think about in a physical geometry sense, the values of the efficiency in each bin are difficult to see.
+Figure~\ref{fig:pim_flat_eff_ex} shows the same efficiency plot but plotted in a more standard two dimensional plot.
+Here we can see even better the acceptance of downward going tracks , $\phi$=\ang{270}, is greater than the dip in upward going tracks, $\phi$=\ang{90}, due to the field cage being centered in the magnet and not in the TPC.
+
+The area enclosed by the red lines represent the region of the highest efficiency over the full region in $\theta_{Lab}$.
+Since we take cuts similar to these in the real data, we assume the efficiency is independent of $\phi$ over these small regions.
+The efficiency then can be represented as just a function of the two remaining variables.
+Figure~\ref{fig:132_eff} shows the efficiency of $\pi^-$ and $\pi^+$ particles in the $\tin{132}{124}$ system and Fig.~\ref{fig:108_eff} for the $\tin{108}{112}$ system as a function of $\theta_{Lab}$ and momentum $p_{Lab}$.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_132_efficiency.png}
+        \caption{$\pi^-$ efficiency for $\tin{132}{124}$} \label{fig:pim_132_eff}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pip_132_efficiency.png} 
+        \caption{$\pi^+$ efficiency for $\tin{132}{124}$} \label{fig:pip_132_eff}
+    \end{subfigure}
+
+    \caption{ }
+\label{fig:132_eff}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_132_efficiency.png}
+        \caption{$\pi^-$ efficiency for $\tin{108}{112}$} \label{fig:pim_108_eff}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pip_108_efficiency.png} 
+        \caption{$\pi^+$ efficiency for $\tin{108}{112}$} \label{fig:pip_108_eff}
+    \end{subfigure}
+
+    \caption{ }
+\label{fig:108_eff}
+\end{figure}
+
+The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database. The correction factor $C_i$ is defined as,
+
+\begin{equation}
+C_i = \epsilon_i^{-1}.
+\end{equation}
+
+The track is then weighted by the factor $C_i$ when plotting in any type of graph for any observable.

### 2020-05-30 chapters/ch_data_analysis_2.tex
Update analysis section

@@
 \section{Beam Particle Identification}
-Figures of beam contaminants in our beam PID line. 
-Table of beam purity reference to Jon's thesis paper
+%Figures of beam contaminants in our beam PID line. 
+%Table of beam purity reference to Jon's thesis paper
 
-Beam tracking was handled by the BigRIPS spectrometer tracking system.
-Each individual beam particles coming down the spectrometer was measured using a dE-Time-of-Flight method where the charge, Z, and mass to charge ratio, A/Q, of each particle was measured.
+Each beam  was tracked by the BigRIPS spectrometer tracking system as it traveled down the beam pipe using the dE-Time-of-Flight method where the charge, Z, and mass to charge ratio, A/Q, of each particle was measured.
 Individual particle species represent a two-dimensional Gaussian in this space; both the $\tin{132}{112}$ and $\tin{108}{124}$ systems are shown in Fig.~\ref{fig:beampid}.
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
-        \caption{Generic} \label{fig:beamangle108}
+        \caption{Beam angle of the $\tin{108}{124}$ system} \label{fig:beamangle108}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
-        \caption{Competitors} \label{fig:beamangle112}
+        \caption{Beam angle of the $\tin{112}{124}$ system} \label{fig:beamangle112}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
-        \caption{Generic} \label{fig:beamangle124}
+        \caption{Beam angle of the $\tin{124}{112}$ system} \label{fig:beamangle124}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
-        \caption{Competitors} \label{fig:beamangle132}
+        \caption{Beam angle of the $\tin{132}{124}$ system} \label{fig:beamangle132}
     \end{subfigure}

### 2020-05-30 macrosForThesis/EffiencyCorr/plotEfficiency.C

@@ -1,0 +1,68 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotEfficiency()
+{
+
+  TFile *f = TFile::Open("EfficiencyPlot.root");
+  TH2D *pim_132 = (TH2D* )f->Get("efficiency_132_pim");
+  TH2D *pip_132 = (TH2D* )f->Get("efficiency_132_pip");
+  TH2D *pim_108 = (TH2D* )f->Get("efficiency_108_pim");
+  TH2D *pip_108 = (TH2D* )f->Get("efficiency_108_pip");
+
+
+  make(pim_132);
+  make(pip_132);
+  make(pim_108);
+  make(pip_108);
+
+  TCanvas *cvs = stdcvs();
+  cvs->SetRightMargin(.2);
+  pim_132->SetTitle("");
+  pim_132->GetYaxis()->SetTitle("#theta_{Lab}");
+  pim_132->GetXaxis()->SetTitle("p_{Lab}");
+  pim_132->GetZaxis()->SetTitle("Efficiency");
+  pim_132->GetZaxis()->CenterTitle();
+  pim_132->GetXaxis()->SetRangeUser(50,550);
+  pim_132->GetYaxis()->SetRangeUser(0,70);
+  pim_132->GetZaxis()->SetRangeUser(0,1.);
+  pim_132->Draw("colz");
+  cvs->SaveAs("pim_132_efficiency.png");
+
+  pim_108->SetTitle("");
+  pim_108->GetYaxis()->SetTitle("#theta_{Lab}");
+  pim_108->GetXaxis()->SetTitle("p_{Lab}");
+  pim_108->GetZaxis()->SetTitle("Efficiency");
+  pim_108->GetZaxis()->CenterTitle();
+  pim_108->GetXaxis()->SetRangeUser(50,550);
+  pim_108->GetYaxis()->SetRangeUser(0,70);
+  pim_108->GetZaxis()->SetRangeUser(0,1.);
+  pim_108->Draw("colz");
+  cvs->SaveAs("pim_108_efficiency.png");
+
+
+  pip_132->SetTitle("");
+  pip_132->GetYaxis()->SetTitle("#theta_{Lab}");
+  pip_132->GetXaxis()->SetTitle("p_{Lab}");
+  pip_132->GetZaxis()->SetTitle("Efficiency");
+  pip_132->GetZaxis()->CenterTitle();
[…]

### 2020-05-30 macrosForThesis/EffiencyCorr/plotFullEfficiencyEx.C

@@ -1,0 +1,70 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotFullEfficiencyEx()
+{
+
+  double phi[4] = {25, 158, 212, 320}; // for pim cuts
+
+  TLine *l1 = new TLine(0,phi[0],90,phi[0]);
+  l1->SetLineColor(2);
+  l1->SetLineStyle(10);
+  l1->SetLineWidth(4);
+
+  TLine *l2 = new TLine(0,phi[1],90,phi[1]);
+  l2->SetLineColor(2);
+  l2->SetLineStyle(10);
+  l2->SetLineWidth(4);
+
+  TLine *l3 = new TLine(0,phi[2],90,phi[2]);
+  l3->SetLineColor(2);
+  l3->SetLineStyle(10);
+  l3->SetLineWidth(4);  
+
+  TLine *l4 = new TLine(0,phi[3],90,phi[3]);
+  l4->SetLineColor(2);
+  l4->SetLineStyle(10);
+  l4->SetLineWidth(4);
+
+  TFile *f = TFile::Open("EfficiencyPlot.root");
+  TH2D *mom_200 = (TH2D* )f->Get("mom_200");
+  TH2D *mom_400 = (TH2D* )f->Get("mom_400");
+
+  make(mom_200);
+  make(mom_400);
+
+
+  TCanvas *cvs = stdcvs();
+  cvs->SetRightMargin(.2);
+  mom_200->SetTitle("");
+  mom_200->GetXaxis()->SetTitle("#theta_{Lab}");
+  mom_200->GetYaxis()->SetTitle("#phi");
+  mom_200->GetZaxis()->SetTitle("Efficiency");
+  mom_200->GetZaxis()->CenterTitle();
+  mom_200->GetXaxis()->SetRangeUser(0,90);
+  mom_200->GetZaxis()->SetRangeUser(0,1.);
+  mom_200->Draw("colz");
+  l1->Draw();
+  l2->Draw();
+  l3->Draw();
+  l4->Draw();
+  cvs->SaveAs("mom_200_efficiency.png");
+
+
+  mom_400->SetTitle("");
+  mom_400->GetXaxis()->SetTitle("#theta_{Lab}");
+  mom_400->GetYaxis()->SetTitle("#phi");
+  mom_400->GetZaxis()->SetTitle("Efficiency");
[…]

### 2020-05-30 macrosForThesis/EffiencyCorr/plotPolEfficiency.C

@@ -1,0 +1,41 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotPolEfficiency()
+{
+  gStyle->SetOptStat(0);
+  
+  TGraphPolargram *toppol = new TGraphPolargram("toppol",0,90,0,2*TMath::Pi());
+
+  toppol->SetLineColorAlpha(0,0);
+
+  /*
+  toppol->SetPolarLabel(0,"x");
+  toppol->SetPolarLabel(2,"y");
+  toppol->SetPolarLabel(4,"180");
+  toppol->SetPolarLabel(6,"270");
+*/
+
+  toppol->SetRadialOffset(.01);
+
+  TFile *f = TFile::Open("EfficiencyPlot.root");
+  TH2D *mom_200_pol = (TH2D *)f->Get("mom_200_pol");
+
+  mom_200_pol->SetTitle("");
+  //  make(mom_200_pol);
+
+  TCanvas *cvs = new TCanvas("c1","c1",1000,800);
+  //  TCanvas *cvs = stdcvs();
+  //  cvs->SetRightMargin(.2);
+  //  cvs->SetLeftMargin(.2);
+
+  mom_200_pol->Draw("same pol lego2");
+  toppol->Draw("same");
+
+  gPad->SetTheta(90);
+  gPad->SetPhi(180);
+
+  cvs->SaveAs("pim_200_pol.png");
+
+}

### 2020-05-30 macrosForThesis/PionMultiplicity/GenieAnalysis/create_submit_Cocktail100.C

@@ -1,0 +1,8 @@
+void create_submit_Cocktail100() {
+  Int_t runNo[] = {3204,3205};
+  Int_t numRuns = sizeof(runNo)/sizeof(Int_t);
+
+  gInterpreter -> ProcessLine(".L create_submit.C");
+  for (auto i = 0; i < numRuns; i++)
+    gInterpreter -> ProcessLine(Form("create_submit(%d,%d);", runNo[i], runNo[i]));
+}

### 2020-05-30 macrosForThesis/PionMultiplicity/GenieAnalysis/create_submit_Sn132.C

@@ -1,0 +1,8 @@
+void create_submit_Sn132() {
+  Int_t runNo[] = {2841,2843,2844,2845,2846,2848,2849,2850,2851,2852,2855,2856,2857,2858,2859,2860,2861,2875,2877,2878,2879,2880,2881,2882,2883,2884,2887,2888,2889,2890,2891,2892,2893,2894,2896,2898,2899,2900,2901,2902,2903,2904,2905,2907,2914,2916,2917,2919,2920,2921,2922,2924,2925,2926,2927,2929,2930,2931,2932,2933,2934,2935,2936,2939,2940,2941,2942,2943,2944,2945,2946,2948,2955,2956,2958,2959,2960,2961,2962,2964,2965,2966,2968,2969,2970,2971,2972,2973,2975,2976,2977,2978,2979,2980,2981,2982,2983,2984,2985,2986,2988,2989,2990,2991,2992,2993,2997,2999,3000,3002,3003,3007,3039};
+  Int_t numRuns = sizeof(runNo)/sizeof(Int_t);
+
+  gInterpreter -> ProcessLine(".L create_submit.C");
+  for (auto i = 0; i < numRuns; i++)
+    gInterpreter -> ProcessLine(Form("create_submit(%d,%d);", runNo[i], runNo[i]));
+}

### 2020-05-30 macrosForThesis/PionMultiplicity/GenieAnalysis/plotPionMultTheoryData.py

@@ -1,0 +1,568 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+# In[7]:
+
+
+import plotly.graph_objects as go
+from plotly.subplots import make_subplots
+import pandas as pd
+import numpy as np
+from plotly.offline import iplot, init_notebook_mode
+import matplotlib.pyplot as plt
+
+
+# In[8]:
+
+
+#codes with all L values
+sn132_theory = pd.read_csv('Pion_Multiplicity_Theory - 132.csv', sep=',')
+sn108_theory = pd.read_csv('Pion_Multiplicity_Theory - 108.csv', sep=',')
+sn124_theory = pd.read_csv('Pion_Multiplicity_Theory - 124.csv', sep=',')
+
+#codes with only min max L values
+#sn132_theory = pd.read_csv('Pion_Multiplicity_Theory - Copy of 132.csv', sep=',')
+#sn108_theory = pd.read_csv('Pion_Multiplicity_Theory - Copy of 108.csv', sep=',')
+#sn124_theory = pd.read_csv('Pion_Multiplicity_Theory - Copy of 124.csv', sep=',')
+
+sn132_theory['ratio'] = sn132_theory['pim']/sn132_theory['pip']
+sn124_theory['ratio'] = sn124_theory['pim']/sn124_theory['pip']
+sn108_theory['ratio'] = sn108_theory['pim']/sn108_theory['pip']
+
+
+# In[9]:
+
+
+
+sn132_theory_min = sn132_theory.reset_index()
+sn132_theory_min = sn132_theory_min.loc[sn132_theory.groupby('name')['L'].idxmin()]
+
+sn132_theory_max = sn132_theory.reset_index()
+sn132_theory_max = sn132_theory_max.loc[sn132_theory.groupby('name')['L'].idxmax()]
+
+
+sn124_theory_min = sn124_theory.reset_index()
+sn124_theory_min = sn124_theory_min.loc[sn124_theory.groupby('name')['L'].idxmin()]
+
+sn124_theory_max = sn124_theory.reset_index()
+sn124_theory_max = sn124_theory_max.loc[sn124_theory.groupby('name')['L'].idxmax()]
+
+sn108_theory_min = sn108_theory.reset_index()
+sn108_theory_min = sn108_theory_min.loc[sn108_theory.groupby('name')['L'].idxmin()]
+
+sn108_theory_max = sn108_theory.reset_index()
[…]

### 2020-05-30 macrosForThesis/PionMultiplicity/GenieAnalysis/plotTotalDoubleRatio.C

@@ -1,0 +1,76 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotTotalDoubleRatio()
+{
+  double x[2] = {-1,10};
+  double y[2] = {2.42299,2.42299};
+  double yerr[2] = {.046775,.046775};
+  double xerr[2] = {0,0};
+  //                          0          1         2       3           4      5       6
+  vector<TString> label = {"AMD+JAM","#chi BUU","UrQMD","IQMD-BNU","SMASH","TuQMD","pBUU"};
+
+  vector<double> stiff_dr = {2.010429448,2.3261826,1.831156419,2.014091353,1.846693731,2.139241691,2.053547455};//soft
+
+ vector<double> soft_dr = {1.905637255,2.132652096,1.811179408,1.77183966,1.89199499,2.209500946,1.919319357};//stiff
+
+ vector<int> order = {1,5,6,0,4,3,2};
+
+ TGraph *stiff_g = new TGraph(label.size()+1);
+ TGraph *soft_g = new TGraph(label.size()+1);
+
+ TGraphErrors *data = new TGraphErrors(2,x,y,xerr,yerr);
+
+ make(stiff_g);
+ make(soft_g);
+
+ for(int i = 0; i < label.size(); i++)
+   {
+     stiff_g->SetPoint(i,i+1,stiff_dr.at(order.at(i)) );
+     soft_g->SetPoint(i,i+1,soft_dr.at(order.at(i)) );
+   }
+
+ stiff_g->SetMarkerStyle(20);
+ soft_g->SetMarkerStyle(20);
+
+ stiff_g->SetMarkerSize(4);
+ soft_g->SetMarkerSize(4);
+
+ stiff_g->SetMarkerColor(kRed - 4);
+ soft_g->SetMarkerColor(kBlue - 4);
+
+ data->SetFillColorAlpha(kCyan - 4, .35);
+ data->SetLineStyle(10);
+ data->SetLineWidth(4);
+ 
+ TAxis *ax = stiff_g->GetXaxis();
+ ax->SetNdivisions(10);
+ 
+ ax->ChangeLabel(1,-1,0,1,-1,-1,""); //0 erase
+
+ for(int i = 0; i < label.size(); i++)
[…]

### 2020-05-30 macrosForThesis/PionMultiplicity/GenieAnalysis/plotTotalDoubleRatio.py

@@ -1,0 +1,123 @@
+import plotly.graph_objects as go
+from plotly.subplots import make_subplots
+import pandas as pd
+import numpy as np
+from plotly.offline import iplot, init_notebook_mode
+import matplotlib.pyplot as plt
+
+#codes with all L values
+theory = pd.read_csv('Pion_Multiplicity_Theory - Double Ratio.csv', sep=',')
+
+theory_min = theory.reset_index()
+theory_min = theory_min.loc[theory.groupby('name')['L'].idxmin()]
+
+theory_max = theory.reset_index()
+theory_max = theory_max.loc[theory.groupby('name')['L'].idxmax()]
+
+
+print(theory_min)
+custom_dict = {'AMD+JAM':3, "r'$\chi \text{BUU}$'":0, 'UrQMD':6, 'IQMD-BNU':5, 'SMASH':4, 'pBUU':2, 'TuQMD':1}
+
+theory_min['rank'] = theory_min['name'].map(custom_dict)
+
+theory_min.fillna(0, inplace=True)
+
+theory_min.sort_values(by=['rank'], inplace=True)
+
+print(theory_min)
+
+
+fig = go.Figure()
+
+#plot the soft double ratio
+fig.add_trace(go.Scatter(
+        x = theory_min.name,
+        y = theory_min.doubleratio,
+        name = 'Soft',
+        marker=dict(
+        #color='rgba(156, 165, 196, 0.95)',
+        color='lightblue',
+        line_color='lightgray',
+        ))
+)
+
+#plot the stiff double ratio
+fig.add_trace(go.Scatter(
+        x = theory_max.name,
+        y = theory_max.doubleratio,
+        name = 'Stiff',
+        marker=dict(
+        #color='rgba(156, 165, 196, 0.95)',
+        color='lightpink',
+        line_color='lightgray',
+        ))
+)
+
+
+
+# Add mean line for data measurment
+
+doubleratio = 2.42299
+
+#error bars
+err_doubleratio = .046775
+
+y0 = -.5
+y1 = 6.5
+
+
+#rectangle represents error bar
+fig.add_shape(type="rect", fillcolor='blue', opacity=.3, xref="x1", yref="y1",
+            y0=doubleratio-err_doubleratio, x0=y0, y1=doubleratio+err_doubleratio, x1=y1)
+        
+        #lines represent mean value of measurment
+fig.add_shape(type="line", xref="x1", yref="y1",
[…]

### 2020-05-30 thesis_style.h

@@ -120,6 +120,11 @@ TH2 *style::make(TH2 *h) {
   h->GetYaxis()->SetLabelSize(fAxisLabelSize);
   h->GetZaxis()->SetLabelSize(fAxisLabelSize);
 
+  h->GetZaxis()->CenterTitle();
+  h->GetZaxis()->SetTitleOffset(fXTitleOffset);
+  h->GetZaxis()->SetTitleSize(fAxisTitleSize);
+  h->GetZaxis()->SetLabelSize(fAxisLabelSize);
+
   
   return h;
 }

### 2020-06-03 chapters/ch_data_analysis.tex
Still working on analysis section. backup

@@
 
-Need to explain pedestal subtraction
-
-\section{Software}
+\section{S$\pi$RITROOT Software}
 \label{sec:software}
@@
 
-The decoder task converts the binary data file into a container class which maps the electronics channels into the corresponding pads and (x,z) coordinates.
-
-There may be several pulses in a pad coming from two tracks passing under the same pad separated  by arrival time.
-Using an expected pulse shape the PSA task fits the signal pulses within a pad, giving the arrival time of the drifted electrons from each particular track.
-The height of the fitted pulse is proportional to the total charge of that event, Q and the y-coordinate is calculated as $y = v\cdot t_0$ where $v$ is the drift velocity and $t_0$ the arrival time.
-Combining the information from these first two tasks, (x,y,z,Q), we construct what is called a "hit".
-
- The Helix Track Finding Algorithm finds the collection of hits belonging to one track out of all the hits in an event.
-The hits within a track are then reduced into clusters.
-A cluster's position is the average position of the hits within a cluster, with the total charge of the cluster being the sum of the hits charges.
-
- A tracks average position is estimated by the cluster's average position.
+\subsection{Decoder task}
+The decoder task reads in the raw the binary data output from the GET electronics and separates the data from the headers using the GETDecoder CITE HERE.
+The raw data is mapped from the cobo layout on the TPC to each individual channel and the information is stored in the STPad class.
+This is a container class which stores the row and layer information, and the raw ADC time bucket spectrum.
+All of the pads combined form the STRawEvent class which represents event information.
+
+Three basic calibrations are performed at the raw data level within this class and will be described in more detail later.
+The gain calibration calibrates the electronics described in Section\ref{sec:elecCalib} and the anode gain calibration discussed in Section~\ref{sec:anodeCalib}.
+The gating grid noise subtraction and time bucket cut which is described in Section~\ref{sec:ggnoisesub}.
+%gain calibration 
+%ggnoise subtraction
+%GG time bucket cut
+
+\subsection{Pulse Shape Algorithm (PSA) Task}
+\label{sec:psa}
+%Yoffset match the TPC-Vertex_y with BDC_Y
+%YPedestalOffset what is it for?
+%cobo gain matching
+An algorithm loops over the raw time bucket spectrum to find the unique pules in each pad.
+It is very likely there are several pulses in a pad coming from multiple tracks passing under the same pad separated only in arrival time.
+Using an expected pulse shape the Pulse Shape Algorithm (PSA) task finds the first peak in the spectrum, fits the peak with the standard pulse shape.
+If the fit is successful the fitted pulse is subtracted from the raw ADC spectrum and the height of the pulse $Q$, which is related to the total charge, and the timing information $t_o$ is saved into a new container class which represent a ``hit", called STHit.
+The algorithm continues through the subtracted spectrum finding all the pules and assigning the fitted values to new hits until there are no more pules in a spectrum.
+Figure~\ref{fig:psaTask} shows the PSA task working on one particular pad in real experimental data.
+The shaded histogram in every sub-figure represents the ADC raw spectrum of the pad.
+In Fig.~\ref{fig:psa0} we see the PSA has identified the first pulse and fitted it with the standard pulse, the result of the fit is shown in the red line.
+The fitted red line is then subtracted from the ADC spectrum which is shown in the next Fig.~\ref{fig:psa1}.
+Here the second pulse is found and fitted again with the same procedure repeating were the red line will be subtracted from the raw ADC spectrum.
+Here the black line shows the sum of all the fitted pules.
+This continues for every pules in the spectrum until all pules are found or until it reaches the last bin in the spectrum.
+Figure~\ref{fig:psa4} shows the final result of all 4 fitted pules and their sum which is given in the black line.
+The agreement between the total summed line and the raw ADC spectrum is very good.
+
+After the PSA processes each pad in the raw event, an entire hit cloud of the event is produced.
+The x and z-coordinates come from the center of the pad which is already stored in the map of STPad.
+Since the STHit class inherits from the STPad class this information is also carried over.
+The y-coordinate is calculated using the known drift velocity $v$ and time of the signal $t_o$ as $y = v\cdot t_0$, $t_0$.
+The hit class forms the fundamental structure from which all tracking is performed since it contains the x,y,z spacial coordinates, and charge information $Q$.
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{0.24\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{psa0.png}
+         \caption{Fist step of the PSA}
+         \label{fig:psa0}
+     \end{subfigure}
+     \hfill
+     \centering
+     \begin{subfigure}[b]{0.24\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{psa1.png}
+         \caption{Second iteration of the PSA}
+         \label{fig:psa1}
+     \end{subfigure}
+     \hfill
+      \centering
+     \begin{subfigure}[b]{0.24\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{psa2.png}
+         \caption{Third iteration of the PSA}
+         \label{fig:psa2}
+     \end{subfigure}
+     \hfill
+      \centering
+     \begin{subfigure}[b]{0.24\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{psa3.png}
+         \caption{Fourth iteration of the PSA}
+         \label{fig:psa3}
+     \end{subfigure}
+     \hfill
+     \centering
+     \begin{subfigure}[b]{.5\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{psa4.png}
+         \caption{Final output of the PSA}
+         \label{fig:psa4}
+     \end{subfigure}
+     \hfill
+
+\caption{Example of the pulse shape algorithm through each step of in processing a given pad. }
+\label{fig:psaTask}
+\end{figure}
+
+\subsection{Helix Tracking}
+%Cluster cut
+%ellipsoid cut
+
+ The function of the  helix track finding algorithm is to associate the collection of hits belonging to one track out of all the entire hit cloud produced by the PSA.
+This is performed by a Reimann track finding algorithm where the cartesian space is mapped onto a Reimann sphere.
+This is a particular transformation in which unique helices in the Cartesian space form unique circles on the Riemann sphere.
+This is particularly useful since hits which correspond to a track form a helix the Cartesian space of the TPC coordinates.
+It is quite difficult to search the hit cloud space fitting a general helix with different combinations of hits.
+It is much easier to search for collection of hits by fitting a plane to the unique great circles formed on the Riemann sphere.
+From here the hits which belong to unique circles are set into a new helix track container called the STHelixTrack.
+
+\paragraph{Definition of clustering}
+The hits within a helix track are dense and the position resolution of an individual hit is not precise.
+As mentioned in Section~\ref{sec:padResponse}, due to the PRF, the collection of hits perpendicular to the track gives the TPC sub millimeter position resolution.
+In the helix track finding task, the hits are further reduced into clusters.
+A brief description of the method of clustering is illustrated in Figure \ref{fig:topview}.
+It is impractical to cluster in both the x and z-axis at the same time and instead we only cluster in one direction.
+For example, the three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis and the upper three are along the z-axis.
+The bolded pads highlight an example from each set.
+The direction perpendicular to the track is the direction of the highest position resolution for a track.
+Since we decide to cluster only in one direction, there is a natural inflection point in a track at which the direction of clustering switches.
+This  depends on the crossing angle $\theta$, which is defined as the angle the tangent makes with respects to the x-axis.
+In this definition $\theta = \ang{90}$ when the track is going along the z-axis, and $\theta = \ang{0}$ for a track going along the x-axis.
+In the case that the crossing angle is $\ang{45} < \theta \leq \ang{90} $ the we cluster along the x-direction.
+For $\ang{0} < \theta \leq \ang{45}$ we cluster along the z-axis.
+
+%A cluster's position is the average, charge-weighted position of the hits within a cluster and the total charge of the cluster is the sum of all the hits charges.
+
+ %A tracks average position is estimated by the cluster's average position.
 The clusters are then fitted in the GENFIT track fitting package \cite{genfit}, giving the final momentum of the track.
 A final vertex of the event is fitted from all tracks using the package RAVE \cite{rave}.
-
-\paragraph{Definition of clustering}
-
-A brief description of the method of clustering is illustrated in Figure \ref{fig:topview}.
-It is impractical to cluster in both the x and z-axis and we only cluster the hits along one axis.
-The three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis and the upper three are along the z-axis, as shown by the bolded pads for one of the clusters in each direction.
-
- The clustering direction depends on the angle  of the track with respects to the x-axis, defined as $\theta$.
-For example, a track going along the z-axis the crossing angle is defined as $90^{\circ}$, and a track going along the x-axis defined as $0^{\circ}$.
-In the case that the crossing angle is $45^{\circ} < \theta \leq 90^{\circ} $ the clustering direction is along the x-axis.
-For $0^{\circ} < \theta \leq 45^{\circ}$ it is along the z-axis.
-
- The position along the clustering direction is calculated by weighting the individual hit's positions by their charges $q_i$ and getting the mean value.
+ The position along the clustering direction, $\overline{X}$, is calculated as the charge-weighted average, $\overline{X} = \sum_i q_ix_i$, where $q_i$ and $x_i$ are the charge and position of the i-th hit.
 The other direction is set to the center of the pad.
-For example if we are clustering along the x-axis for a cluster, the z-position is set to the center of the pad in the z-direction and vice versa.
-
+For example, if we are clustering hits along the x-axis, the z-position is set to the center of the pad in the z-direction and vice versa.
 Clustering in this way gives us better position resolution for calculating the position of each cluster.
-You could imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
-By clustering in the direction most perpendicular to the track, we get a better position resolution.
+You can imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
 
@@
 
-\subsection{Pulse Shape Algorithm}
-\label{sec:psa}
+\subsection{Correction Task}
+
+\subsection{Space charge}
+
+\subsection{Pad response function cut}
+
+\subsection{Momentum and track reconstruction (GENFIT)}
+
+\subsection{Vertex tracking (RAVE)}
+\label{sec:vertex
+}
 
@@
 \subsection{Gating grid noise subtraction}
+\label{sec:ggnoisesub}
 Opening the gating grid essentially short adjacent wires together allowing them to reach equilibrium as fast as possible.
@@
 \subsection{Electronics calibration}
+\label{sec:elecCalib}
 The ADC channel number of the electronics was calibrated by measuring the response of each channel to an input pulse supplied by a pulse generator.
@@
 \subsection{Anode gain calibration}
-
+\label{sec:anodeCalib}
 As mentioned earlier, the anode wires were separated into 14 independently biased sections.
@@
 \section{CoBo timing correction}
-
 The arrival time of signals originating from each pad may differ due to timing delays in the electronics and cabling.
@@
 The timing difference is stable for each pad across several runs.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset6before_row_at_layer10.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset5before_row_at_layer30.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset4before_row_at_layer50.png}
+
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset6after_row_at_layer10.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset5after_row_at_layer30.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset4after_row_at_layer50.png}
+
+    \caption{TBW\label{TBD}}
+  \end{center}
+  \label{fig:coboCorr}
+\end{figure}
+
+\begin{figure}[!htb]
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
+        \caption{Distribution for all pads before correction.} \label{fig:yoff_allBefore}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
+        \caption{Distribution for all pads after correction.} \label{fig:yoff_allAfter}
+    \end{subfigure}
+    \caption{ }
+\label{fig:yoff}
+\end{figure}
+
  Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows , for three different layers.
@@
 There is a significant improvement in the width of the distribution going from \SI{1.5}{\milli\metre} to \SI{0.6}{\milli\metre}.
-
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=0.33\textwidth]{before_row_at_layer10_yoffset.pdf}
-    \includegraphics[width=0.33\textwidth]{before_row_at_layer30_yoffset.pdf}
-    \includegraphics[width=0.33\textwidth]{before_row_at_layer50_yoffset.pdf}
-
-    \includegraphics[width=0.33\textwidth]{after_row_at_layer10_yoffset.pdf}
-    \includegraphics[width=0.33\textwidth]{after_row_at_layer30_yoffset.pdf}
-    \includegraphics[width=0.33\textwidth]{after_row_at_layer50_yoffset.pdf}
-
-    \caption{TBW\label{TBD}}
-  \end{center}
-  \label{fig:coboCorr}
-\end{figure}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{before_layer_yoffset.pdf}
-        \caption{Y-offset for all layers before correction.} \label{fig:yoff_beforeLayer}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{after_layer_yoffset.pdf} 
-        \caption{Y-offset for all layers after correction.} \label{fig:yoff_afterLayer}
-    \end{subfigure}
-     \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{before_all_yoffset.pdf} 
-        \caption{Distribution for all pads before correction.} \label{fig:yoff_allBefore}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{after_all_yoffset.pdf} 
-        \caption{Distribution for all pads after correction.} \label{fig:yoff_allAfter}
-    \end{subfigure}
-    \caption{ }
-\label{fig:yoff}
-\end{figure}
 

### 2020-06-03 chapters/ch_data_analysis_2.tex
Still working on analysis section. backup

@@
 
+\section{Radio Isotope Beam Factory (RIBF) Facility }
+%Cyclotron facility overview.
+%Samurai line overview.
+%Beam line element overview.
+%Big rips beam PID. reference 
+The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
+The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
+Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{SAMURAI-beamline.png}
+\caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
+\label{fig:samuraiBeamLine}
+\end{figure}
+
+After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
+These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
+This is accomplished through several dipole magnets, slits, and wedge degraders.
+The resulting secondary beam is not pure and the purity depends on the capability of BigRIPS to deliver the secondary beam of choice and the primary beam used.
+
+In these set of experiments several beams were produced with varying intensities and purities.
+Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns where most beams were delivered with an intensity of \SI{10}{\kilo\hertz}.
+
+ \begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}ccccc@{}}\toprule 
+ Primary Beam & Secondary Beam & Energy at mid target \si{\MeVA} & Intensity \si{\kilo\hertz} & Purity (\%) \\ [0.5ex] 
+ \midrule
+ ${}^{238}$U   & ${}^{132}$Sn   &  269.2  &  9.5  &  54   \\
+ ${}^{238}$U   & ${}^{124}$Sn   &  270.3  &  9.1  &  10  \\
+ ${}^{124}$Xe  & ${}^{112}$Sn   &  270.4  &  7.6  &  48  \\
+ ${}^{124}$Xe  & ${}^{108}$Sn   &  269.3  &  7.5  &  52   \\
+ \bottomrule
+\end{tabular}
+\caption{Primary and secondary beam properties produced in the \spirit TPC experimental campaigns. }
+\label{tb:beams}
+\end{table*}
+
 \section{Beam Particle Identification}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{beamPID.png}
+\label{fig:beampid}
+\caption{Overall beam PID for all the systems. Several contaminants other than the desired secondary beam can be seen.}
+\end{figure}
+
+\begin{figure}[!htb]
+
+     \centering
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{beam-Sn132.png}
+         \caption{$\tin{132}{124}$ Beam particle identification.}
+         \label{fig:beampid132}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{beam-Sn108.png}
+         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \label{fig:beampid108}
+     \end{subfigure}
+        \caption{$\tin{108}{112}$ Beam particle identification. }
+        \label{fig:beampidTwo}
+
+\end{figure}
+
 %Figures of beam contaminants in our beam PID line. 
 %Table of beam purity reference to Jon's thesis paper
+The secondary beam is produced through the projectile fragmentation of the primary beam off of a \SI{3}{\milli\metre} thick, rotating Be target \cite{inflightsep}.
+The resulting fragments are filtered in-flight to the desired seconary beam.
+The in-flight separation is handled by the BigRIPS fragment separator which is shown in Fig.~\ref{fig:samuraiBeamLine}.
+The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$.
+Several sets of slits further purify the secondary beam quality by throwing away particles which do not focus on the right focal planes.
+These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.
+Each beam is tracked with the remaining part of the BigRIPS spectrometer tracking system.
+The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}.
+From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussians in this space.
 
-Each beam  was tracked by the BigRIPS spectrometer tracking system as it traveled down the beam pipe using the dE-Time-of-Flight method where the charge, Z, and mass to charge ratio, A/Q, of each particle was measured.
-Individual particle species represent a two-dimensional Gaussian in this space; both the $\tin{132}{112}$ and $\tin{108}{124}$ systems are shown in Fig.~\ref{fig:beampid}.
-Clear separation is seen for each particle species and the correct value of Z and A/Q gives us the correct species type.
-Each species of interest is fitted with multivariate normal distribution with two variables defined as,
+Figure~\ref{fig:beampid} shows the beam PID for all the systems for events which satisfied the trigger, several contaminants other than the desired secondary beams still passed through the BigRIPS spectrometer and made it into the TPC.
+The beam purity of each desired secondary beam is listed in Table~\ref{tb:beams}.
+The desired secondary beam of interest can be selected by using an appropriate gate around the corresponding group, since the BigRIPS PID resolution is good enough to separate separate beam types.
+Each particle gate is selected by fitting a multivariate normal distribution with two variables defined as,
 
@@
 \end{equation}
-where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths.
+where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Fig.~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.
 
-The resulting fit parameters are summarized in Table~\ref{beamParameters}. 
 \begin{table}[H]
@@
 \end{table}
+Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam.
+The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events.
+For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values.
 
-For both ${}^{132}$Sn an ${}^{108}$Sn a 2.83$\sigma$ cut is taken around the mean values; as seen in Fig.~\ref{beampid132,beampid108}.
+\section{Edge Cuts}
 
 \begin{figure}[!htb]
+
     \centering
-     \begin{subfigure}[b]{0.49\textwidth}
+    \begin{subfigure}[t]{0.45\textwidth}
         \centering
-         \includegraphics[width=\textwidth]{beam-Sn132.png}
-         \caption{$\tin{132}{124}$ Beam particle identification.}
-         \label{fig:beampid132}
+        \includegraphics[width=\linewidth]{clusterLR.png} 
+        \caption{Top view of the tracks showing the edge effect on the left and right sides of the TPC.} 	   \label{fig:clusterLR}
     \end{subfigure}
     \hfill
-     \begin{subfigure}[b]{0.49\textwidth}
+    \begin{subfigure}[t]{0.45\textwidth}
         \centering
-         \includegraphics[width=\textwidth]{beam-Sn108.png}
-         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
-         \label{fig:beampid108}
+        \includegraphics[width=\linewidth]{clusterTop.png} 
+        \caption{Side view of the TPC showing the edge effect near the top.} \label{fig:clusterTop}
     \end{subfigure}
-        \caption{$\tin{108}{112}$ Beam particle identification. }
-        \label{fig:beampid}
+
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{clusterBottom.png} 
+        \caption{Side view of the TPC showing the edge effect near the bottom.} \label{fig:clusterBottom}
+    \end{subfigure}
+\caption{}    
+\label{fig:edge}
 \end{figure}
 
-\section{Edge Cuts}
-As seen in Fig.~\ref{fig:edge}, edges of the measurement volume of the TPC we can see several edge effects distortion the clusters positions.
-The last pads on the edges of the pad-plane naturally have no neighbors on the outsides, therefore their cluster position is biased towards the inside of the TPC.
-Also since we measure a finite window in time for the vertical direction, pulses near the edge of the time-bucket window are clipped and the reconstructed times are biased towards inside the TPC volume.
-
-The number of affected clusters is small, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
-Simple cuts around the left,right, top, and bottom of the TPC remove these clusters from the analysis,
+Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.
+This effect comes from an edge effects where the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad, or in the case of the vertical time bucket spectrum, represents the last known position of the collected charge.
+In this case the cluster position is biased towards the inside of the TPC.
+While the number of affected clusters is small as compared with the total number in the track, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
+Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC.
+The hits that were cut out satisfied the following conditions,
 
@@
 
+\section{High Density Cut}
+
 \begin{figure}[!htb]
+
      \centering
-    \begin{subfigure}[t]{0.45\textwidth}
+     \begin{subfigure}[b]{0.49\textwidth}
          \centering
-        \includegraphics[width=\linewidth]{clusterLR.png} 
-        \caption{Generic} \label{fig:clusterLR}
+         \includegraphics[width=\textwidth]{sideviewHighDenCut.jpg}
+         \caption{Side view.}
+         \label{fig:sideHigh}
      \end{subfigure}
      \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
+     \begin{subfigure}[b]{0.49\textwidth}
          \centering
-        \includegraphics[width=\linewidth]{clusterTop.png} 
-        \caption{Competitors} \label{fig:clusterTop}
+         \includegraphics[width=\textwidth]{topviewHighDenCut.jpg}
+         \caption{Top view.}
+         \label{fig:topHigh}
      \end{subfigure}
-
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{clusterBottom.png} 
-        \caption{Generic} \label{fig:clusterBottom}
-    \end{subfigure}
-
-\label{fig:edge}
+        \label{fig:highcut}
+        \caption{Views showing the high density and edge cuts.}
+        \label{fig:elipsecut}
 \end{figure}
-
-\section{High Density Cut}
-The density of tracks near the target region is very high and most of electronics information from this region is inaccurate in both timing and charge value.
-Hits lying within an semi-ellipsoidal cut region around the target are removed from the software and not included in the track and momentum reconstruction.
-
-Add picture of ellipsoid
+The high track multiplicity in each event is quite high and since the tracks all originate from a common vertex, the density of tracks near the target region is very high.
+In this region the track separation is too small for the software to correctly determine and sort the charge information into the appropriate track.
+The information provided by extra vertex point from external BDC tracking described in \ref{sec:bdc}, provides all the information about the vertex location.
+The bad quality of hit information near the target region only hurt in the tracking and PID of a track.
+Hits lying within an semi-ellipsoidal cut around the target are removed from the software and not included in the track and momentum reconstruction.
+Figure~\ref{fig:elipsecut} shows the extent of the ellipsoidal cut in the high density region, along with the edge cuts as shaded red regions in both views of the TPC.
 
 \section{Beam angle selection}
-The beam is deflected by the magnetic field and impinges on the target at some angle.
-The BDCs track the beam as a straight line right up until the magnet.
-To calculate the incidence angle, the beam is then propagated through the magnetic field until the target position.
-The beam angle on target is categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,
+The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle.
+From the BDCs tracking information, the beam is projected as a straight line right up until entering the magnet.
+The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position.
+The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,
 
@@
 where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector.
-
-Figure~\ref{beamAngleFig} shows the beam angle distribution for ${}^{132}$Sn and ${}^{108}$Sn.
 
@@
 
+Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems.
+Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}.
+Helping to eliminate some of the poorly reconstructed beam events.
+
 \section{Vertex Cut}
-A vertex cut on verticies originating from the target region ensure we are selecting on nuclear collisions with the target.
-When looking at the z-component of the verticies of all events, we can see several features which correspond to different elements within the beam line and TPC.
-We also note that near the expected position of the target, WHAT POSITION??, we see a large enhancement of counts which corresponds to nuclear collisions on the target.
-The z-component of all runs in each beam type are plotted in the range of the target region in Fig.~\ref{fig:vertexdist}.
-Since the target thickness was less than 1 mm, the vertex resolution achieved can be experimentally measured by fitting the vertex distribution, assuming a Gaussian distribution.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale =.5]{vertexDistribution.png}
+\caption{Projection of the vertex onto the z-axis.}
+\label{fig:overviewVertex}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{vertex-Sn108.png} 
+        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{vertex-Sn112.png} 
+        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{vertex-Sn124.png} 
+        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{vertex-Sn132.png} 
+        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
+    \end{subfigure}
+\caption{}
+\label{fig:vertexdist}
+\end{figure}
+
+The vertex information of an event, described in \ref{sec:vertex}, is estimated from reconstructing the tracks of an event to one common source.
+The secondary beam encounters several solid and gaseous materials were a nuclear reaction can occur anywhere along the beam line.
+Figure~\ref{fig:overviewVertex} shows the z component of the reconstructed vertex for all events in the ${}^{132}$Sn system.
+Several peaks are seen in the spectrum which correspond to several dense materials in the beam line such as the entrance windows, target frame, Active Veto, with the largest peak representing the target.
+Also collisions happen with the detector gas inside the TPC volume which are typically called active target events.
+To ensure that the secondary beam is really on the target of choice we perform a vertex cut around where we believe the vertex location of the target to be.
+
+ The target position was measured to be \SI{-13.2}{\milli\metre}.
+The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}.
+The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position.
+The mean position for each system is listed in Table~\ref{tb:vertexresol}.
+Since the target thickness was less than \SI{1}{\milli\metre} for all targets.
+Neglecting the small distributions in the thin target, the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC.
 The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}.
+The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is really and insignificant difference.
 
@@
 
-\begin{figure}[!htb]
+\clearpage
+
+\section{Impact Parameter Selection}
+
+\begin{figure}[htb]
 \centering
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{vertex-Sn108.png} 
-        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{vertex-Sn112.png} 
-        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{vertex-Sn124.png} 
-        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{vertex-Sn132.png} 
-        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
-    \end{subfigure}
-\label{fig:vertexdist}
+\includegraphics[scale=.5]{error_b_full.pdf}
+\caption{Estimate of the impact parameter from the track multiplicity.}
+\label{fig:impactPar}
 \end{figure}
 
-\section{Impact Parameter Selection}
-The impact parameter which is defined as the distance (center-to-center), between two nuclei in a collision is a pure theoretical concept.
-Though experimentally there is no way in experiment to measure the impact parameter directly, several quantities have been indirectly related to the centrality of the collision CITE HERE.
-As discussed in Section CITE HERE, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events.
-This is under the reasonable assumption that the number of charged particles in a collision grows as the impact parameter decreases.
+The impact parameter is a theoretical quantity defined as the distance, between two centers of colliding nuclei.
+Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity can be indirectly related to the estimated impact parameter of a collision \cite{impactpar}.
+As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events.
+This is under the assumption that as the number of charged particles produced in collision is related to the overlap region of the two nuclei.
 This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons.
-In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters (peripheral collisions), less nucleons participate.
+In the case where the impact parameter is zero, all the nucleons in both nuclei participate, were as larger impact parameters -- more peripheral collisions -- less nucleons participate.
 
-Certainly the Kyoto array introduces a bias in our measurement of the multiplicity of all the events.
+%Certainly the Kyoto array introduces a bias in our measurement of the multiplicity of all the events.
 At low track multiplicities, the orientation of the event starts to play a major role.
@@
 Therefore events with low track multiplicity (therefore more peripheral), that emit preferentially up and down in the TPC will not be triggered on.
-This becomes less of an issue as even in mid-peripheral collisions were the number of particles emitted becomes much greater than the Kyoto multiplicity selection.
+This becomes less of an issue as even in mid-peripheral collisions were the number of particles emitted becomes much greater than the Kyoto multiplicity selection
 
-Assuming a geometric cross section $\sigma$ and the impact parameter of the collision $b$,
+The geometric cross section $\sigma$ can be described as,
 
@@
 
-we can estimate the maximum impact parameter $b_{\mathrm{max}}$, by measuring the total cross section of each system.
-A detailed analysis on determining the cross section is given in CITE HERE.
-I have summarized the finding of the cross section and $b_{\mathrm{max}}$ for all the systems.
-Assuming $b_{\mathrm{max}}$ represents the most peripheral collisions we can define an estimate for the impact parameter $\hat{b}$, which we can define from the cumulative integral of the measured multiplicity distribution,
+where $b$ is the impact parameter of the collision.
+If the track multiplicity of an event $N_C$ is monotonically related to the cross section, the impact parameter can be written as a reduced impact parameter $\hat{b}$ as,
+\begin{equation}
+\hat{b} =  \frac{b}{b_{max}} = \int^{\infty}_{N_C} \frac{dP(N_C)}{dN_C} dN_C,
+\end{equation}
+
+where $b_{\mathrm{max}}$ represents the maximum impact parameter detected by the TPC, $b$ is the impact parameter of the vent,  and $dP(N_C/dN_C$ is the normalized multiplicity distribution \cite{reducedimpact}.
+The reduced impact parameter ranges from $\hat{b}=0$ for the most central collisions and to $\hat{b}=1$ for the most peripheral.
+A  detailed analysis was performed, determining the maximum cross section $\sigma_{max}$, for each system, in which $b_{\mathrm{max} = \sqrt{\sigma_{max}/\pi}}$.
+The detailed analysis is given in \cite{jon}.
+Figure~\ref{fig:impactPar} shows the estimated impact parameter for a given track multiplicity with the estimated error bands for each system.
+
+In the $\tin{132}{124}$ system the multiplicity cut was $N_C > 50$ corresponding to $\hat{b} = 0.4$ and $b = \SI{3.1}{\femto\metre}$.
+For the $\tin{108}{112}$ system the multiplicity cut was $N_C > 49$ corresponding to $\hat{b} = 0.4$ and $b = \SI{3.1}{\femto\metre}$.
+These values were averaged over the multiplicity distribution as,
 
 \begin{equation}
-\hat{b} = b_{\mathrm{max}} \frac{\int_{\infty}^{M_x} f(M) dM}{\int_{\infty}^{0} f(M) dM}.
+ \overline{X} = \int_{N_C}^{\infty} X\frac{dP(N_C)}{dN_C} dN_C
 \end{equation}
+
+where X can be the variables $b$, $\hat{b}$.
+In the $\tin{132}{124}$ system the average $\overline{\hat{b}} = \num{.1(1)}$ and $\overline{b} = \SI{3}{\femto\metre}$.
+In the $\tin{108}{112}$ system the average $\overline{\hat{b}} = .1$ and $\overline{b} = \SI{3}{\femto\metre}$.
+These quantities are useful when comparing to theory.
 
@@
 %distance to vertex
-In this section we will discuss the quality of the reconstructed track and will simply refer to it as the ``track quality".
-Certainly tracks with more clusters are of higher quality than those with few clusters.
-Occasionally the software will associate clusters and construct a track whose vertex does not coincide with the event vertex.
-This can occur for a variety of reasons.
-For a low momentum pion the radius of curvature is such that several rotations can occur in the event which the software can interpret as several tracks; while only one of the reconstructed track appears to originate from the event vertex.
-It could also happen that false tracks are created by the software from clusters which originate from several different real tracks.
-These are fairly obvious as they usually traverse the detector horizontally and the number of clusters is very low.
-In this situation the false track does not originate from anywhere near the vertex.
-
-To reduce the number of fake tracks, or multiple counting of tracks, a simple cut on the measurement of a track's distance to vertex can be used.
-The vector representing the minimum distance to vertex for each track can be calculated and its corresponding magnitude.
-We define the distance to vertex cut as a small sphere centered on the vertex location with a radius R.
-If the distance to vertex of a particular track $d_i$ < R the track is assumed to have originated from the vertex; tracks that are outside are thrown out of the analysis.
-
-Assuming tracks are mostly continuous in clusters, i.e. only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track.
+In this section we will discuss cuts which address the quality of the reconstructed track, in the following we will simply refer this as the ``track quality".
+Assuming tracks are mostly continuous in clusters, i.e.
+only randomly missing a few clusters, the number of clusters is directly related to the momentum resolution of a track.
+Tracks with more clusters correspond to better PID and momentum resolution.
+Upward-going and downward-going tracks, at larger values of $\theta_{Lab}$ are limited by the vertical space of the TPC.
+In these regions the track length is short and there are few clusters which are reconstructed.
+This leads to the low efficiency in these regions as discussed earlier in Sec.~\ref{sec:efficiency}.
+A cut where tracks with the number of clusters $N_{cl} > 20$ are considered quality tracks.
+Later in Section~\ref{sec:cutvar} the exact choice of the number 20 is discussed in more detail.
 
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{distancetovertex.pdf}
+\centering
+\includegraphics[scale=1.]{distancetovertex.pdf}
 \label{fig:poca}
@@
 
+For very long tracks, such as low momentum pions, which can make a complete circular path in the TPC, it can happen that the software incorrectly identifies the track as several separate tracks.
+This can happen due to discontinuities in the track due to missing hits from low energy loss, shadowing due to saturation, or possibly the software algorithm itself.
+Of course including all of the tracks would be counting the track several times over, and would lead to incorrect particle yields.
+The software will occasionally associate random disassociated clusters and construct a false track, or a so called ``ghost track".
+These type of tracks contribute to the background in the PID.
+In both of the situation of the false track or in the case of a split track the false track origin cannot be traced back to anywhere near the vertex of the event, and the distance-Of-Closest-Approach (dOCA) between the track and vertex is very large.
+
+To reduce the number of false tracks, a simple cut on the dOCA can be used.
+The vector representing the minimum distance to vertex for each track can be calculated and its corresponding magnitude.
+We define the distance to vertex cut as a small sphere centered around the vertex location with a some radius $R$.
+If the dOCA of a particular track $d_i < R$, the track is assumed to have originated from the vertex; tracks outside of the cut are thrown out of the analysis.
+
 \section{Angular Quality Cuts}
-Since the TPC has no spherical symmetry, there are geometric constraints on measuring tracks.
-Most notably tracks going up and down in the TPC where the path of the track would be shortest.
-Tracks traveling up and down in the TPC will have the shortest track length and will correspond to low regions of efficiency.
-The relevant geometries for this discussion are the corners of the field cage.
-Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the field cage were the angles of each corner is given.
+
+Since the TPC does not have spherical symmetry, there are geometric constraints and considerations that effect the efficiency of track reconstruction.
+Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}.
+The relevant geometries which determine these regions are the corners of the field cage.
+Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given.
 Since the window to the field cage was slightly higher, the upper angles are slightly smaller than the lower angles.
-The $\phi$ angles location of each corner is are $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{-24 \\ +34}$, with a width of about \ang{58}.
-While it depends on polar angle $\theta$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the $\phi$ angles.
-
-The efficiency in these regions are either zero or very small.
-Typically it is not wise to efficiency correct already small numbers with small efficiency values.
-Therefore we have taken the practice to cut away the regions of very low efficiencies.
-The cuts we have made are for $\theta_{C.M.}$ < 90 and for \ang{326} < $\phi$ < \ang{24} and \ang{1} < $\phi$ < \ang{24}.
-
-Maybe add a table of cuts for each system.
-
-We need to correct for the solid angles not measured, or cut away, to arrive at the correct total yields in $4\pi$. The correction is simple given,
-
-\begin{equation}
-\Delta\Omega = \Delta\phi(\cos(\theta_1) - \cos(\theta_2)),
-\end{equation}
-
-where $\Delta\phi$ is the $\phi$ and $\theta_{1,2}$ are the $\theta$ angular coverage after cuts. The multiplicitive correction factor to the data, C, for the missing angular coverage is simply,
-
-\begin{equation}
-C = \frac{4\pi}{\Delta\phi}.
-\end{equation}
-
-%As the spherical angle of the track, $\theta$, increases the track gets even shorter and shorter until there are very few clusters at all.
-We see around $\phi$=\ang{90} and $\phi$=\ang{270}, representing up and down in the TPC respectively, there is near zero efficiency for nearly all angles $\theta$>\ang{30} and for a wide range of $\phi$ angles around these two values.
+These angles are actually the same as the spherical $\phi$ angle, where the left and right, most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.
+While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks, where the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly.
+The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}.
+The values are very similar to the simple calculation that was outlined above.
 
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{tpcAngle.pdf}
+\centering
+\includegraphics[scale=.75]{tpcAngle.pdf}
 \caption{Rough Geometry of the TPC field cage.}
@@
 \end{figure}
+
+ \begin{table*}[!htb]
+ \centering
+\ra{1.3}
+\begin{tabular}{@{}ccc@{}}\toprule 
+Particle Type & $\theta_{CM}$ cut & $\phi$ cut  \\ [0.5ex] 
+ \midrule
+$\pi^+$  & $\theta_{CM}$ < \ang{90}   &  \ang{-35} < $\phi$ < \ang{20} $\cap$ \ang{157} < $\phi$ < \ang{219}  \\
+$\pi^-$  & $\theta_{CM}$ < \ang{90}   &  \ang{-40} < $\phi$ < \ang{25} $\cap$ \ang{158} < $\phi$ < \ang{212}   \\
+ \bottomrule
+\end{tabular}
+\caption{Angular cuts for each system and particle type}
+\label{tb:anglecuts}
+\end{table*}
+
+These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.
+We could of course include a larger region where the efficiency is small but not zero but it is good practice not to efficiency correct already small numbers with small efficiency values.
+The solid angle covered by the cut is written as,
+
+\begin{equation}
+\Delta\Omega = \Delta\phi(\cos(\theta_1) - \cos(\theta_2)),
+\end{equation}
+
+where $\Delta\phi = (\phi_2 - \phi_1)$  and $\theta_{1,2}$ are the $\theta$ angle cuts, in units of \si{\radian}.
+The solid angle covered by the $\pi^-$ cuts is \SI{2.077}{\steradian} and for $\pi^+$ cuts \SI{2.042}{\steradian}.
+Assuming the pion emission is isotropic, we multiply each observed pion by a correction factor, $C_a$, to correct for the full 4$\pi$ angular coverage,
+
+\begin{equation}
+C_a = \frac{4\pi}{\Delta\Omega},
+\end{equation}
+
+for $\pi^+$ that is \num{6.15} and $\pi^-$ \num{6.05}.
+We assume the pion emission is isotropic for a couple of reasons.
+The symmetry of very central collisions is invariant with respects to any rotations.
+Also the mass of the pion is smaller than the nucleons mass and therefore collective motion leading to anisotropies would be small for the pion, i.e.
+most of the motion would be thermal.
+We also measured two systems $\tin{112}{124}$ and its inverse $\tin{124}{112}$.
+The forward emission in the $\tin{124}{112}$ system is really the same as the backward emission of the $\tin{112}{124}$ system and vice versa.
+It was shown for pions emitted in central collisions the forward and backward emission was the same \cite{jon}, proving that at least for central collisions pions are emitted isotropically to a good approximation.

### 2020-06-03 macrosForThesis/COBOcalb_images_fromJW/draw_calibrated.C

@@ -1,0 +1,59 @@
+void draw_calibrated()
+{
+  auto file1 = new TFile("overview_before_fix.root");
+  auto file2 = new TFile("overview_after_fix.root");
+
+  gStyle -> SetOptStat(0);
+
+  for (auto ba : {0,1}) {
+    TFile *file = file1;
+    TString header = "before_";
+    if (ba==1) {
+      file = file2;
+      header = "after_";
+    }
+
+    //for (auto ii : {1,2,4,5,6})
+    for (auto ii : {1})
+    {
+      TString gname = Form("run2905_yoffset%d",ii);
+
+      TString cname;
+      TString title;
+      if (ii==1) { cname = "all";   title = ";dy (mm);"; }
+      if (ii==2) { cname = "layer"; title = ";dy (mm);layer"; }
+      if (ii==4) { cname = "row_at_layer50";  title = "layer 50;dy (mm);row"; }
+      if (ii==5) { cname = "row_at_layer30";  title = "layer 30;dy (mm);row"; }
+      if (ii==6) { cname = "row_at_layer10";  title = "layer 10;dy (mm);row"; }
+
+      auto cvs = new TCanvas(header+cname,header+cname,1000,1000);
+      cvs->SetBottomMargin(.15);
+      cvs->SetLeftMargin(.15);      
+      cvs->SetRightMargin(.05);
+      
+      auto hist = (TH2D *) file -> Get(gname);
+      hist -> SetNameTitle(header+cname,title);
+      hist -> GetXaxis()->SetTitle("dy (mm)");
+
+      if(ii >= 4)
+	{
+	  hist -> GetYaxis()->SetTitle("Row");
+	  hist -> GetYaxis()->SetTitleOffset(1.1);
+	}
+      else if( ii == 1)
+	{
+	  hist -> GetYaxis()->SetTitle("# Clusters");
+	  hist -> GetYaxis()->SetTitleOffset(1.1);
+	  hist->SetLineWidth(4);
+	}
+      
+      hist -> Draw("col");
+
+      //      cvs->SetRightMargin(.15);
+      //      hist -> Draw("colz");
+
+      cvs -> Draw();
+      cvs->SaveAs(gname + header+cname + ".png");
+    }
+  }
+}

### 2020-06-10 MSU-thesis-estee.tex
Update
Added GG information
New graphics for gg transparency
quick update before check
Routine upload
Update analysis section
Still working on analysis section. backup
Backup

@@
 \usepackage{rotating}
+\usepackage{mathtools}
+\usepackage{comment}\renewcommand{\arraystretch}{1.5}
 
-\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} }
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/} }
 
@@
 \usepackage{newtxtext,newtxmath,booktabs,siunitx,caption,subcaption}
+
+%\sisetup{separate-uncertainty}
+\sisetup{input-uncertainty-signs}
+
 %
@@
 \fieldofstudy{Physics} % This should be in sentence case
-\date{2019}
+\date{2020}
 
@@
 \newcommand{\spirit}{S$\pi$RIT }
+\newcommand{\numerr}[3]{ \num[parse-numbers=false]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
+%\newcommand{\numerr}[3]{ \num[input-close-uncertainty]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
+
 \newcommand{\tin}[2]{{}^{#1}\mathrm{Sn} + {}^{#2}\mathrm{Sn}}
@@
 
+\makeatletter
+\DeclareTextCommand{\textprime}{\encodingdefault}{%
+  \mbox{$\m@th'\kern-\scriptspace$}% 
+}
+\makeatother
+
 \DeclareSIUnit\torr{Torr}
 \DeclareSIUnit\eVperc{\eV\per\clight}
+\DeclareSIUnit\MeVA{\mega\eV A}
+\DeclareSIUnit\clight{\text{\ensuremath{c}}}
 \DeclareSIUnit\micron{\micro\metre}
@@
 %chapters/ch_intro,
-%chapters/ch_theory,
-chapters/ch_experiment,
-%chapters/ch_data_analysis,
+%/ch_theory,
+%chapters/ch_experiment,
+chapters/ch_data_analysis,
 %chapters/ch_data_analysis_2,
-%chapters/ch_results
+chapters/ch_results
 }
@@
 %\include{chapters/ch_theory}
-\include{chapters/ch_experiment}
-%\include{chapters/ch_data_analysis}
+%\include{chapters/ch_experiment}
+\include{chapters/ch_data_analysis}
 %\include{chapters/ch_data_analysis_2}
-%\include{chapters/ch_results}
+\include{chapters/ch_results}
 

### 2020-06-10 chapters/ch_data_analysis.tex
Backup

@@
 
-The S$\pi$RITROOT software is modular tasked based code based on the FAIRROOT package written in C++ \cite{fairroot}. The main tasks in the S$\pi$RITROOT software reconstruction are:
+The S$\pi$RITROOT software is modular code based on the FAIRROOT package written in C++ \cite{fairroot}.
+The code is separated into modules which can be turned on or off depending on the users needs.
+The base classes and frame work is provided by the FAIRROOT package, while the specifics of each detector need to be added into the base classes.
+In the following an outline of the main tasks in the S$\pi$RITROOT software will be briefly discussed with the details of each task discussed later.
+The discussion of the software is really a practical, base understanding, appropriate for this thesis.
+For more details about the tracking algorithm the reader is refered to \cite{spiritroot_paper}, and to actual software which is published on GitHub \cite{spiritroot_git}.
+
+The main tasks which make up the reconstruction algorithm are:
+
 \begin{itemize}
@@
   \item Helix Track Finding Algorithm
-  \item Clustering Algorithm
-  \item Track Fitting (GENFIT package)
-  \item Vertex Fitting (RAVE package)
+  \item Track Fitting and momentum reconstruction(GENFIT package)
+  \item Vertex Reconstruction (RAVE package).
 \end{itemize}
@@
 \subsection{Decoder task}
-The decoder task reads in the raw the binary data output from the GET electronics and separates the data from the headers using the GETDecoder CITE HERE.
-The raw data is mapped from the cobo layout on the TPC to each individual channel and the information is stored in the STPad class.
-This is a container class which stores the row and layer information, and the raw ADC time bucket spectrum.
-All of the pads combined form the STRawEvent class which represents event information.
-
-Three basic calibrations are performed at the raw data level within this class and will be described in more detail later.
-The gain calibration calibrates the electronics described in Section\ref{sec:elecCalib} and the anode gain calibration discussed in Section~\ref{sec:anodeCalib}.
+The decoder task reads in the raw data output from the GET electronics and separates the data from the headers using the GETDecoder \cite{getdecoder}.
+A map is stored mapping each channel in the CoBo board onto the physical pad plane, the information is stored in a pad container called the STPad class.
+This container stores the row, layer, and the raw ADC time bucket spectrum.
+All of the pads combined form the event container, STRawEvent class, which represents the raw event information.
+
+Three basic calibrations are performed at the raw data level within this task.
+The gain calibration calibrates the electronics will be described in Section\ref{sec:elecCalib} and the anode gain calibration is discussed in Section~\ref{sec:anodeCalib}.
 The gating grid noise subtraction and time bucket cut which is described in Section~\ref{sec:ggnoisesub}.
@@
 %cobo gain matching
-An algorithm loops over the raw time bucket spectrum to find the unique pules in each pad.
-It is very likely there are several pulses in a pad coming from multiple tracks passing under the same pad separated only in arrival time.
-Using an expected pulse shape the Pulse Shape Algorithm (PSA) task finds the first peak in the spectrum, fits the peak with the standard pulse shape.
-If the fit is successful the fitted pulse is subtracted from the raw ADC spectrum and the height of the pulse $Q$, which is related to the total charge, and the timing information $t_o$ is saved into a new container class which represent a ``hit", called STHit.
-The algorithm continues through the subtracted spectrum finding all the pules and assigning the fitted values to new hits until there are no more pules in a spectrum.
-Figure~\ref{fig:psaTask} shows the PSA task working on one particular pad in real experimental data.
+An algorithm loops over the raw time bucket spectrum to find all the unique pules in each pad.
+It is very likely that there are several pulses in a pad coming from multiple tracks passing under the same pad which are separated only in arrival time.
+Figure~\ref{fig:psaTask} shows the PSA task working on one particular pad in experimental data.
 The shaded histogram in every sub-figure represents the ADC raw spectrum of the pad.
-In Fig.~\ref{fig:psa0} we see the PSA has identified the first pulse and fitted it with the standard pulse, the result of the fit is shown in the red line.
-The fitted red line is then subtracted from the ADC spectrum which is shown in the next Fig.~\ref{fig:psa1}.
-Here the second pulse is found and fitted again with the same procedure repeating were the red line will be subtracted from the raw ADC spectrum.
-Here the black line shows the sum of all the fitted pules.
-This continues for every pules in the spectrum until all pules are found or until it reaches the last bin in the spectrum.
-Figure~\ref{fig:psa4} shows the final result of all 4 fitted pules and their sum which is given in the black line.
-The agreement between the total summed line and the raw ADC spectrum is very good.
-
+In Fig.~\ref{fig:psa0} we see the PSA has identified the first pulse and fitted it with the standard pulse, the result of the fit is shown by the red line.
+The fitted red line is then subtracted from the ADC spectrum, which is shown in the shaded histogram in Fig.~\ref{fig:psa1}.
+Here, the second pulse is found and fitted again with the same procedure, the red line will be subtracted from the raw ADC spectrum.
+The black line represents the sum of all the fitted pules in each step.
+This continues until all pules are found or until the PSA reaches the last bin in the spectrum.
+Figure~\ref{fig:psa4} shows the final result where 4 fitted pulses were found.
+The agreement between the total summed line and the raw ADC spectrum exemplifies how successful the PSA is in this particular set of data.
+
+%Mention about the minimum distance between hits?
+
+The fit of each individual pulse contains the height of the pulse and the timing information of the start of the pulse, defined to be the location in time of 10\% of the rising edge.
+The height of the pulse, which is related to the charge induced on the pad, $Q$, and the timing information $t_o$, is saved into a new container representing a ``hit", called the STHit class.
 After the PSA processes each pad in the raw event, an entire hit cloud of the event is produced.
@@
          \includegraphics[width=\textwidth]{psa0.png}
-         \caption{Fist step of the PSA}
+         \caption{$\mathrm{1^{st}}$ step of the PSA}
          \label{fig:psa0}
@@
          \includegraphics[width=\textwidth]{psa1.png}
-         \caption{Second iteration of the PSA}
+         \caption{$\mathrm{2^{nd}}$ iteration}
          \label{fig:psa1}
@@
          \includegraphics[width=\textwidth]{psa2.png}
-         \caption{Third iteration of the PSA}
+         \caption{$\mathrm{3^{rd}}$ iteration}
          \label{fig:psa2}
@@
          \includegraphics[width=\textwidth]{psa3.png}
-         \caption{Fourth iteration of the PSA}
+         \caption{$\mathrm{4^{th}}$ iteration}
          \label{fig:psa3}
@@
 %ellipsoid cut
-
- The function of the  helix track finding algorithm is to associate the collection of hits belonging to one track out of all the entire hit cloud produced by the PSA.
-This is performed by a Reimann track finding algorithm where the cartesian space is mapped onto a Reimann sphere.
-This is a particular transformation in which unique helices in the Cartesian space form unique circles on the Riemann sphere.
-This is particularly useful since hits which correspond to a track form a helix the Cartesian space of the TPC coordinates.
-It is quite difficult to search the hit cloud space fitting a general helix with different combinations of hits.
-It is much easier to search for collection of hits by fitting a plane to the unique great circles formed on the Riemann sphere.
-From here the hits which belong to unique circles are set into a new helix track container called the STHelixTrack.
+\label{sec:helixtrack}
+
+ The function of the  helix track finding algorithm is to sort hit cloud into unique sets of hits representing tracks.
+This is performed by a Riemann track finding algorithm where the Cartesian space is mapped onto a Riemann sphere.
+This is a particular transformation in which unique helices in the Cartesian space form great circles on the Riemann sphere.
+This is particularly useful since hits which correspond to a track form a helices the Cartesian space of the TPC coordinates.
+It is difficult to search for tracks in the hit cloud by fitting a general helix to different combinations of hits.
+It is much easier to search for collection of hits by fitting a plane to the great circles which are formed on the Riemann sphere.
+From here the set of hits which form a unique track are stored into a new helix track container called the STHelixTrack class.
 
 \paragraph{Definition of clustering}
-The hits within a helix track are dense and the position resolution of an individual hit is not precise.
-As mentioned in Section~\ref{sec:padResponse}, due to the PRF, the collection of hits perpendicular to the track gives the TPC sub millimeter position resolution.
-In the helix track finding task, the hits are further reduced into clusters.
+The hits within a helix track are dense and the position resolution of an individual hit is not very precise.
+As mentioned in Section~\ref{sec:padResponse}, the TPC can achieve sub-millimeter position resolution by utilizing the PRF.
+In the helix track finding task, the hits are further reduced into more precise clusters.
 A brief description of the method of clustering is illustrated in Figure \ref{fig:topview}.
-It is impractical to cluster in both the x and z-axis at the same time and instead we only cluster in one direction.
-For example, the three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis and the upper three are along the z-axis.
-The bolded pads highlight an example from each set.
-The direction perpendicular to the track is the direction of the highest position resolution for a track.
-Since we decide to cluster only in one direction, there is a natural inflection point in a track at which the direction of clustering switches.
-This  depends on the crossing angle $\theta$, which is defined as the angle the tangent makes with respects to the x-axis.
-In this definition $\theta = \ang{90}$ when the track is going along the z-axis, and $\theta = \ang{0}$ for a track going along the x-axis.
-In the case that the crossing angle is $\ang{45} < \theta \leq \ang{90} $ the we cluster along the x-direction.
+As will be seen, it is sufficient to simply cluster hits along one direction, either along the x or z-direction, which ever is most perpendicular to the local track region.
+For example, the three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis, and the upper three are along the z-axis.
+The bold pads highlight the hits belonging to an example cluster from each set.
+Since we decide to cluster only in one direction, there is a natural inflection point in a track at which the direction of clustering must switch.
+This depends on the crossing angle of the track $\theta$, which is defined as the angle between the tangent of the track  at that cluster and the x-axis.
+In this definition $\theta = \ang{90}$ when the track is going along the z-axis and $\theta = \ang{0}$ for a track going along the x-axis.
+The clustering direction in the case $\ang{45} < \theta \leq \ang{90} $ is along the x-direction.
 For $\ang{0} < \theta \leq \ang{45}$ we cluster along the z-axis.
 
-%A cluster's position is the average, charge-weighted position of the hits within a cluster and the total charge of the cluster is the sum of all the hits charges.
-
- %A tracks average position is estimated by the cluster's average position.
-The clusters are then fitted in the GENFIT track fitting package \cite{genfit}, giving the final momentum of the track.
-A final vertex of the event is fitted from all tracks using the package RAVE \cite{rave}.
- The position along the clustering direction, $\overline{X}$, is calculated as the charge-weighted average, $\overline{X} = \sum_i q_ix_i$, where $q_i$ and $x_i$ are the charge and position of the i-th hit.
-The other direction is set to the center of the pad.
-For example, if we are clustering hits along the x-axis, the z-position is set to the center of the pad in the z-direction and vice versa.
-Clustering in this way gives us better position resolution for calculating the position of each cluster.
-You can imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
-
-\begin{figure}[H]
+\begin{figure}[!htb]
+\centering
 \includegraphics[scale=.5]{top_view_helix_ext.pdf}
@@
 
+ The position along the clustering direction, $\overline{X}$, is calculated as the charge-weighted average, $\overline{X} = \sum_i q_ix_i$, where $q_i$ and $x_i$ are the charge and position of the i-th hit in the cluster.
+The other direction is set to the center of the pad.
+For example, if we are clustering hits along the x-axis, the z-position is set to the center of the pad in the z-direction and vice versa.
+Switching the clustering direction gives better position resolution than if we clustered only along one direction.
+You can imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
+The cluster position and charge is stored into a new container represented the cluster called the STHitCluster class.
+All the clusters which belong to a particular helix track is also stored in the corresponding STHelixTrack class.
+
 \subsection{Correction Task}
-
-\subsection{Space charge}
-
-\subsection{Pad response function cut}
+%extend dynamic range
+%PRF cut
+%space charge correciton
+In this task the hit distribution, within a given cluster, is checked against the measured PRF discussed in Section~\ref{sec:prf}.
+Figure~\ref{fig:prfpimData} shows black lines around the measured PRF in the data, which form gates defining the acceptable PRF region.
+If less than 50\% of the hits within a cluster lie within this gate, the cluster is thrown out from the track.
+A majority of the clusters contains either 2 or 3 hits in total; in the case of 2 total hits, at least one of them needs to lie in the gate, in the case of 3 total hits, at least 2 hits need to lie within the gate.
+Clusters which do not follow this PRF gate usually have been corrupted.
+Typically charges from other nearby tracks is the typical way clusters become corrupt.
+
+%CONSIDER SHOWING SOME PROOF OF IT IN ACTION
+
+This correction task also handles several other important corrections.
+The first is a novel algorithm which extends the dynamic range of the electronics and is discussed in detail in Section~\ref{sec:extendDynamicRange}.
+The other correction task that is handled here is the space charge correction which will be discussed in more detail in Section~\ref{sec:spacecharge}.
 
 \subsection{Momentum and track reconstruction (GENFIT)}
+After all of the corrections are applied to the clusters, the cluster's position and charge values are passed to a software package which reconstructs the momentum of the track, called GENFIT \cite{genfit}.
+The details of Kalman fitting algorithm used in the GENFIT reconstruction are left for the reader here \cite{genfit}.
+GENFIT returns a momentum and charge value for the track which is stored into a new track container called the STRecoTrack class.
+
+The task is repeated a second time, storing a second copy of the track class but this time with the BDC vertex point added as a constraining cluster.
+The momentum resolution is greatly improved by adding the vertex as a constraining point; since the BDC detectors, described in Section~\ref{sec:bdc}, give such an accurate vertex location.
+The energy loss of the track is calculated by the truncated mean method described in Eq.~\ref{eq:truncmean}.
+This is the final part of the analysis which pertains to tracks, and the PID can be constructed using the energy loss, momentum, and charge information.
 
 \subsection{Vertex tracking (RAVE)}
-\label{sec:vertex
-}
+\label{sec:vertex}
+After all tracks are reconstructed by GENFIT, the tracks are then passed to the RAVE software package which reconstructs the event vertex from a collection of tracks \cite{rave}.
+Only tracks which are THIS CONDITION are used in making the event vertex.
+The vertex location along the beam axis determines whether or not the beam collided with the target.
 
 \section{Calibrations and Corrections}
+In the following subsections we will discuss in detail the calibrations and corrections happening within the software framework that was outlined above.
 
@@
 \label{sec:ggnoisesub}
-Opening the gating grid essentially short adjacent wires together allowing them to reach equilibrium as fast as possible.
-In practice the impedance of both sides was not entirely matched properly and this caused an oscillating current to bounce back and forth in an under-dampened manner.
-This caused residual induced signals early in the time bucket spectrum which created and extra source of noise in the data.
-The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.
+The opening of the gating grid is essentially shorting adjacent wires in the grid together, allowing them to reach equilibrium as fast as possible.
+Since the impedance of both sides was not entirely matched properly this caused an imbalance of charge which bounce back and forth in the grid in an under-dampened manner.
+This oscillating charged induced  residual induced signals early in the time bucket spectrum creating extra source of noise in the data.
 Figure~\ref{fig:ggNoiseSub} shows in the upper panel the gating grid noise ADC time bucket spectrum for 2000 events in a given pad.
-The signal is stable within a pad and the mean value --shown as the red line-- can be calculated after averaging over several thousands events.
-The raw gating grid noise lasts for 100 tbs extending into the real data with a decreasing amplitude.
+The gating grid noise is stable for a given pad and the mean value --shown as the red line-- can be calculated after averaging over several thousands events.
+The raw gating grid noise lasts for 100 Tbs extending into the real data with a exponentially decreasing ampli
+
+ %The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.
 
@@
 
-The gating grid itself was not always stable an only the ${}^{132}$Sn and ${}^{124}$Sn beams were relatively stable.
-Several times during the ${}^{108}$Sn and ${}^{112}$Sn beams the gating grid broke.
-After several data runs were taken, we took what is called gating grid noise runs to get the profile of the gating grid noise throughout the experiment.
-To achieve this the anode wires would be turned off and the trigger was allowed to otherwise operate normally, firing the gating grid to open as usual.
-Since the anode wires were all turned off no signals would be seen except those coming from the gating grid noise.
-Several thousand triggers of the gating grid were taken to get a good noise profile.
-
-Once the mean value response in each pad is calculated, we can use the gating grid noise profile runs to correct the real data and remove the noise of the gating grid.
-To show the best case scenario of this technique, the bottom panel of Fig.~\ref{fig:ggNoiseSub} shows the gating grid noise self correcting itself.
-We can see the pedestal is already subtracted naturally when using this technique, and the subtracted ADC spectra is centered around 0.
-Residual gating grid noise that was not canceled by the subtraction remains but is significantly reduced to a much smaller time bucket region.
-This happens in regions where the gating grid noise is changing rapidly, where the ADC value is dependent upon small jitters in the timing of the signal.
-All time buckets below 30 tb were cut away from the spectrum to get rid of the small region where the gating grid is still opening and to remove the residual noise.
+The gating grid noise profile is measured by recording experimental data, only turning off the anode wires.
+By doing so there are no signals originating from tracks in the TPC and there is only the gating grid noise, since the grid is still being triggered.
+The mean value of each pad is calculated from several thousand events of the gating grid noise profile.
+Once the mean value response in each pad is calculated, we can subtract the mean profile from the real data and remove the noise of the gating grid.
+
+The bottom panel of Fig.~\ref{fig:ggNoiseSub} shows the gating grid noise self correcting itself; a best case scenario.
+We can see the pedestal, which is around 3800 ADC in the upper pannel,  is automatically taken into account when subtracting the mean profile, and the subtracted ADC spectra is centered around 0.
+The residual gating grid noise that was not canceled by the subtraction remains, but is significantly reduced to a much smaller time bucket region.
+Most of the residual noise happens in regions where the derivative of the gating grid noise is large.
+In these areas the ADC value is dependent upon small jitters in the timing of the signal and as a result varies wildly.
+The residual noise in the subtracted spectra is removed by cutting away all the bins below 30 tb.
+
+In the experiment the gating grid itself was quite stable in the  ${}^{132}$Sn and ${}^{124}$Sn systems and was not very stable in the ${}^{108}$Sn and ${}^{112}$Sn beams, where the gating grid broke several times.
+The gating grid noise was monitored and we regularly took gating grid noise profiles, especially after a gating grid board was replaced.
 
@@
 \subsection{Cocktail calibration}
+\label{sec:cocktail}
 
@@
 \end{tabular}
-\caption{Summary of expected cocktail. }
+\caption{Summary of expected cocktail for the calibration run taken with the Al target. }
 \label{tb:cocktail100tar}
@@
 \midrule
-p   & 882.8 & 903.5 & 889   &  2.0   & -1.6  \\
+p   & 882.8 & 903.5 & 889.0 &  2.0   & -1.6  \\
 d   & 817.1 & 898.5 & 874.5 &  2.1   & -2.7\\
@@
 \end{tabular}
-\caption{Summary of expected cocktail. }
+\caption{Summary of expected cocktail from the lower beam energy.}
 \label{tb:cocktail100}
@@
 \end{tabular}
-\caption{Summary of expected cocktail. }
+\caption{Summary of expected cocktail from the higher beam energy.}
 \label{tb:cocktail300}
@@
 
-A cocktail of several light charged particles were produced (p,d,t,${}^{3}$He,${}^{4}$He), and measured in the TPC to provide a momentum calibration of the TPC.
-The magnetic and slit settings of the dipoles in the BIGRIPS spectrometer was set so that the measured momentum resolution of the beam was $\delta p/p$ < 1\%.
+A cocktail of several light charged particles was produced (p,d,t,${}^{3}$He,${}^{4}$He), and measured in the TPC to provide a momentum calibration of the TPC.
+The magnetic and slit settings of the dipoles in the BigRIPS spectrometer was set so that the momentum resolution of the beam was $\delta p/p$ < 1\%.
 Two magnetic rigidity settings were studied, with an empty target.
-A thick Aluminum target was used to provide a slightly lower energy point for part of the lower rigidity setting, effectively creating three calibration points over several particle species.
-The production of certain particles (t,${}^{3}$He), produced too few counts to make a good measurement, and in the high momentum rigidity setting protons could not propagate down the line and there were no counts.
-
-Since the expected momentum resolution resulting from the spectrometer was less than 1\%, the observed momentum resolution measured by the TPC is a good representation of the combined momentum resolution of the software and TPC (intrinsic detection) system.
-The momentum resolution depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
+A thick Aluminum target was also inserted to provide a slightly lower energy point for part of the lower rigidity setting, effectively creating three energy calibration points over several particle species.
+In some cases the certain particles were not able to propagate or were strongly attenuated.
+
+Since the cocktail beam was a mono-energetic source of particles and the actual momentum resolution was less than 1\%, the observed momentum resolution can be interpreted as the intrinsic TPC momentum resolution.
+The momentum resolution in general depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
 This calibration beam represents and ideal situation where the track was parallel to the pad plane and only one particle was measured at one time.
-The energy loss resolution can also be directly measured since each energy setting represents a monochromatic source of each particle species, which has a well defined energy loss distribution.
-An average momentum resolution of $dp/p = 2\%$ and the energy loss resolution of $d\langle dE/dx \rangle/\langle dE/dx \rangle = 5\%$ was measured for particles ranging from $H$ to ${}^{4}$He, over the range of momenta measured in the calibration beam as summarized in Table~\ref{tb:momresolution}.
-
-Since the magnetic dipole setting of the BIGRIPS spectrometer defined the energy of each particle type, we can calculate the expected momenta of each particle species measured.
-Small corrections were propagated using LISE++ software which can calculate the energy loss through several materials in the beam line.
-These corrections resulted in a small change in the momenta.
-The measured momenta of the calibration beam differed significantly from the expected values as seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
-This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocity in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
+The average momentum resolution was measured to be $dp/p = 2\%$ over the full range of particle species for all settings.
+
+The energy loss resolution can also be directly measured since each energy setting represents a monochromatic source of each particle species, which have a well defined energy loss distribution.
+The energy loss resolution was measured to be 5\% for all particles over the energy setting measured and is summarized in Table~\ref{tb:momresolution}.
+
+Since the magnetic dipole setting of the BigRIPS spectrometer is well defined, we can calculate the expected momenta of each particle species measured.
+The enegy losses through various materials in the beam line were propagated using LISE++ software \cite{lise++}.
+The measured momenta of the calibration beam differed significantly from the expected values -- up to 5\% in the high momentum calibration -- as seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
+This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocities in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
 The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
-The disagreement in measured and expected momenta is upwards of ~5\% difference in the higher momentum calibration settings.
-The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed in the later Section~\ref{sec:spacecharge} in a more general way which also includes correction for the space charge, which this cocktail beam does not have.
-The same correction technique was applied here in the special case of zero space charge, which is the special case of only having $\vec{E}\times\vec{B}$ components.
-
-The values under the corrected column seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data corrected for the $\vec{E}\times\vec{B}$ effect.
+The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed later in Section~\ref{sec:spacecharge} in a  general way which also includes the correction for the positive space charge.
+The same correction technique was applied here where the cocktail beam is the special case of zero space charge, only having $\vec{E}\times\vec{B}$ components.
+
+The values under the corrected column in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the corrected data.
 A significant improvement is seen in the high momentum setting going from around ~5\% disagreement to within ~1\% agreement in the corrected data.
 For the lower momentum settings (Tables~\cref{tb:cocktail100tar,tb:cocktail100}), protons see a slight improvement of about ~1\% where as the deuterons are over corrected in both settings.
-The level of agreement of the all corrected values is still within the estimated momentum resolution of the TPC.
+The level of agreement of the all corrected values is still well within the estimated momentum resolution of the TPC.
 
@@
 \ra{1.3}
-\begin{tabular}{@{}rr@{}}\toprule
+\begin{tabular}{@{}cc@{}}\toprule
 Momentum Resolution \% & <dE/dx> Resolution \% \\
@@
 \end{tabular}
-\caption{Summary of expected cocktail. }
+\caption{Summary of the estimated momentum and energy loss resolutions.}
 \label{tb:momresolution}
@@
 \label{sec:elecCalib}
-The ADC channel number of the electronics was calibrated by measuring the response of each channel to an input pulse supplied by a pulse generator.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{gaincalib.png}
+\caption{Repose of each channel in the electronics to various input voltages of an input signal supplied by a pulse generator.}
+\label{fig:gaincalib}
+\end{figure}
+
+The electronics were calibrated by measuring the response of each channel to an input signal supplied by a pulse generator.
+This is a relative calibration technique with the intent to calibrate the varying gains in each channel relative to one another, and not an absolute calibration.
 The pulse was distributed to all the electronics channels by pulsing the ground plane for a range of input voltages.
 This distributed the pulse evenly across the entire pad plane.
-The input voltage is plotted as a function of the measured ADC channel in Fig.~\ref{fig:gaincalib} for every channel.
+The input voltage is plotted as a function of the measured height of the pulse in the electronics channel given in units of ADC in Fig.~\ref{fig:gaincalib}, for every channel.
 The small variation in each channel can be seen as the wide band around each measurement point.
-A linear fit is performed to get the best fit line which provides a reference line which each channel is calibrated to.
+A linear fit is performed for each channel to get the calibration line for each channel.
+A channel is chosen which provides a reference calibration to which all other channels are calibrated too.
+This is done by matching the slopes of each channel to the reference calibration slope.
 The right panel shows the resulting distribution of channels after calibration.
-This is a relative calibration technique with the intent to calibrate the varying gains in each channel relative to one another.
-
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{gaincalib.png}
-\caption{Calibration of electronics}
-\label{fig:gaincalib}
-\end{figure}
 
@@
 \label{sec:anodeCalib}
-As mentioned earlier, the anode wires were separated into 14 independently biased sections.
-The high voltage of sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires.
-The wire section voltages were lowered once and adjusted once again.
+As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires.
 Out of all the runs used in the analysis in this thesis \ref{tb:runList}, the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all the other runs.
 By lowering the voltage on these anode wires, the gas gain is lowered as compared with all the other anode wire plane sections which operate at \SI{1460}{\volt}.
-To account for the drop in gain, in the software we increase the gain of the pads which lie above these anode wires to match the gain in the other channels.
-To do so, we compare the energy loss in the high gain sections to the low gain sections to get the relative calibration factor.
+To account for the drop in gain, we increase the gain of the pads which lie above these anode wires  in the software gain of the other channels.
+The gain factor is found by plotting the energy loss in the high gain sections to the low gain sections.
 
@@
 \includegraphics[width=\linewidth]{highlowcal.png}
-\caption{Calibration of low high.}
+\caption{Calibration of low and high gain sections of the anode wires.}
 \label{fig:highlowcal}
@@
 Figure~\ref{fig:highlowcal} shows the correlation plot of $\langle dE/dx\angle$ of the high versus the low gain sections.
-The effects of the high gain channel saturation can around a value of 400 ADC\si{\per\milli\metre} and plateaus, where as the low gain does not.
+The effects of the high gain channel saturation can around a value of 400 ADC\si{\per\milli\metre} and plateaus, where as the low gain does not; this region is left out of the fit.
 A linear fit was performed for the calibration between the high gain channel $h_c = G\cdot l_c$, where $l_c$ is the value of the low gain channel and $G$ is the gain factor.
-In the case where the low anode voltage was \SI{1214}{\volt}, the gain factor was 9.8 where as in the case of \SI{1085}{\volt}, the gain factor was ???.
-The calibration factor was applied inside of the decoder task.
-A map of the channels which are impacted by the low anode wire voltages and the relevant gain was input into the software.
+In the case where the low anode voltage was \SI{1214}{\volt}, the $G=9.8$, where in the case of \SI{1085}{\volt}, the $G = $.
+
+ A map of the channels which are affected by the low gain and the relevant gain factor was input into the software.
 In the decoder task the raw ADC spectra was multiplied by the input gain for a given channel.
-The thresholds of those channels also were multiplied by the same factor since the noise levels were also artificially magnified by the same gain factor.
+The thresholds of those channels also were multiplied by the same gain factor since the noise levels were also artificially magnified by this method.
 
 \subsection{Extending the dynamic range of the Electronics}
+\label{sec:extendDynamicRange}
+
 Using a TPC for measurements of HIC in nuclear physics presents a different set of challenges as opposed to higher energy experiments.
-Typically in higher energy experiments the charge of particles is Z$e$, where Z=1.
+Typically in higher energy experiments fundamental particles are produced with charge $\pm e$.
 Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum in the energy loss curve.
-The dynamic range of electronics in such experiments can cover a wide range of particle energies in the energy loss curve.
-In nuclear HICs, we are interested in measured particles ranging from Z=1-3 resulting from the collision, and even higher in some applications.
-As seen in Eq.~\ref{eq:bb}, the energy loss is proportional to $Z^2$; the energy values in nuclear HIC of intermediate beam energies (around \SI{300}{\MeVA}) in which the resulting particles are at much lower energies which are in the $1/\beta^2$ region.
-The energy range and particle types covered by the electronics are significantly limited by the dynamic range limitations of the electronics as the charge of a particle increases and the velocity decreases.
+The dynamic range of most TPC electronics can cover a wide range of particles in these types of experiments.
+In nuclear HICs, we are interested in measured particles with charges $Ze$ where Z=1-3, and even higher in some applications.
+The energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$.
+Even more, low energy nuclear HIC of intermediate beam energies (around \SI{300}{\MeVA}) produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
+The dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases where the electronics will be saturated.
 
 Several TPCs have tried to address this issue by having regions of low and high gain, either in amplification gain or in electronics gain.
-Here we run into the issue that high velocity particles will have little to no signal in the low gain regions.
-At lower velocities, particles will deposit much charge over the low gain regions but will saturate the high gain regions, whose charge values are usually lost.
-Only within the dynamic range will tracks have the best momentum resolution, outside the dynamic range clusters will be missing in the sections which depend on the track velocity, lowering the momentum resolution of tracking.
-There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these persistent issues in TPC electronics CITE HERE, by being able to switch to a lower gain value when the maximum range is reached.
-Though, it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of external hardware, especially in experiments which have already been performed with preexisting electronics technologies.
-
-In TPCs, the effective dynamic range can be very different from the single channel dynamic range depending on the application.
-Typically TPCs are operated inside of a magnetic field for reconstructing momentum of each particle, which requires sub-millimeter precision in the position determination of the track path.
-This is achieved by averaging the charge distribution over several pads.
-Therefore the effective dynamic range is related to the relative charge values of adjacent pads (between the center pad and outside pads).
-For example, to measure minimum ionizing particles the signal to noise ratio of the pad with the smallest charge in the distribution should be some reasonable value above noise, say at least 6:1 signal-to-noise ratio (SNR).
-From Section \ref{sec:prf}, we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
-If we require the adjacent pads to have a SNR of 6:1, then the central pad would have a SNR of 50:1.
-Considering this is the SNR for minimum ionizing particles, and the maximum SNR in a channel is 800:1, this means the effective dynamic range in the TPC is roughly 16 times that of minimum ionizing particles, if we intend to have adequate position and tracking resolution.
-
-The dashed lines and vertical blue bar in Fig.~ \ref{fig:intro} are separated by a factor of 20, representing the typical effective dynamic range in a TPC.
-This dynamic range estimate should be regarded as approximate because the energy loss fluctuates significantly about the most probable energy loss, with a long ``Landau'' like tail, as described by Bichsel \cite{bichsel}.
+This mitigates the complete loss of information but introduces a new problem.
+Though particles which deposit large amounts of charge will now have good measurements in the low gain areas, particles which deposit minimal energy losses will lose information in the same low gain areas.
+The reconstruction of such tracks will suffer.
+There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics CITE HERE.
+Never the less it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of new external hardware, especially for experiments which have already been performed with older electronics technologies.
+In this section we will outline a novel software technique that takes advantage of the PRF described in Section~\ref{sec:prf} and can effectively extend the dynamic range of the TPC electronics.
+
+In TPCs, the effective dynamic range is very different from the single channel dynamic range depending on how the TPC measurement is used.
+Typically TPCs are operated inside of a magnetic field for the purpose of reconstructing the momentum of a track, which requires sub-millimeter precision in the position determination of the track path.
+This is achieved by clustering several pads together as discussed in Section~\ref{sec:helixtrack}.
+To achieve this at least 2 adjacent pads must be measured, and the precision increases as the number of adjacent pads increases.
+
+The effective dynamic range is then not the single channel dynamic range, but the dynamic range between central pad --holding the largest charge-- and the adjacent pads --holding the smaller charges in the PRF distribution.
+For example, to measure minimum ionizing particles, the Signal-to-Noise-Ratio (SNR) of the pad with the smallest charge in the distribution should be some reasonable value above noise, so that the signals can be measured.
+From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
+In the \spirit TPC the electronics gain was set so that the the pads have a SNR of 6:1 for minimum ionizing tracks, and therefore the central pad had a SNR of 50:1.
+In the case of the maximum charge is collected in the central pad, before the electronics saturate, the SNR is 800:1.
+Therefore the maximum SNR is roughly 16 times that  --800 over 50 -- of minimum ionizing particles.
+This is order to have a complete measurement, where adjacent pads are all measured and position resolution is not compromised.
+
+Figure~\ref{fig:intro} shows the theoretical energy loss curves for several particles as a function of the momentum over charge.
+Minimum ionization can be seen to take place around \SI{10}{\kilo\electronvolt\per\centi\metre}.
+The dashed lines and vertical blue bar in are separated by a factor of 16, representing the effective dynamic range in the \spirit TPC.
+This dynamic range should be regarded as approximate because the energy loss fluctuates significantly about the most probable energy loss as described in Section~\ref{sec:energyloss}.
 Nevertheless, the blue dashed lines and vertical blue bar illustrate that the range of energy losses sampled in a fixed gain readout system is limited.
@@
 
-The rapid increase in the stopping powers at low momentum illustrates the degree to which the effective dynamic range can be exceeded and highlights the problems encountered in studies of intermediate HICs, in which light particles with low momenta are abundantly produced along with highly charged particles.
-Similar problems are encountered when TPCs are used as active targets in direct reaction studies with rare isotope beams \cite{pattpc}.
-Several techniques have been employed to increase the observable range of energy losses; typically by lowering the electronics gain of selected readout channels, or by changing the gas amplification at the readout plane in certain areas of the TPC \cref{eos,pattpc}.
-The results of changing the gas-gain, or the electronics gain, are rather similar in that reducing the gain to sample a range of higher energy loss makes the TPC effectively blind to minimum ionizing particles in  the regions of lower gain.
-
 \subsection{Experimental Pad Response Function}
-
+\label{sec:expPRF}
 The pad response function of the TPC was extracted from non-saturating hits and clusters in tracks at various track crossing angles.
@@
 Averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
-Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
+Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$.
 
@@
 
-The shape of the PRF depends on the crossing angle of the track \cite{gatti}, which determines how wide the charge is distributed along the wire.
-Figure~\ref{fig:prfpimData} shows the PRF of $\pi^-$ tracks versus the crossing angle $\theta$ of the track.
-The PRF gets wider starting from $90^{\circ}$  and until where we switch clustering directions at $\ang{45}$; if we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
+The shape of the PRF depends on the crossing angle of the track, which determines how wide the charge is distributed along the wire \cite{gatti}.
+Figure~\ref{fig:prfpimData} shows the PRF for $\pi^-$ tracks versus the crossing angle $\theta$ of the track.
+The PRF gets wider starting from $90^{\circ}$  and until where we switch clustering directions at $\ang{45}$.
+If we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
 Switching from $x$ to the $z$ direction clustering the opposite trend is seen where the PRF becomes narrower  going from $45^{\circ}$ to $0^{\circ}$, as the position resolution gets better.
@@
 	 \includegraphics[width=\textwidth]{PRFs_data_wcut.png}
-     \caption{PRF response from $\pi^-$ data. }
+     \caption{PRF from $\pi^-$ tracks in experimental data. }
      \label{fig:prfpimData}
@@
 
-A Gaussian fit was performed were performed to the experimental data with  $\ang{5}$ width bins from $\ang{0} < \theta \leq \ang{90}$.
-Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function given in Eq.~\ref{eq:prfgaus} are plotted versus $\theta$.
-A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value as seen by the black line.
+A Gaussian fit was performed to the experimental PRF in $\ang{5}$ width bins ranging from $\ang{0} < \theta \leq \ang{90}$.
+Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function given in Eq.~\ref{eq:prfgaus} which are are plotted versus $\theta$.
+A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value, which is shown as the black line.
 
@@
 
-We will use the term ``desaturation'' to describe the process of correcting the charge values of the saturated pads.
-Figure~\ref{fig:satpad} shows a typical situation of saturated signals in a given cluster of hits.
-When an avalanche causes a large enough induced signal, it is the pads directly underneath which collect the largest charge becoming saturated; denoted here as $q_{2\textprime}$ and $q_{3\textprime}$.
-Pads further away collect less charge and typically are not saturated; the non-saturated pads are detonated here as $q_{1}$ and $q_{4}$.
-Though the true charge value in the saturated channels are lost, we know that the distribution of all charges must follow the PRF which is a fixed feature of all TPCs, and has been experimentally measured above.
-From the tracking information, we know a given clusters crossing angle and can interpolate to get the corresponding parameters for the PRF described in Fig.~\ref{fig:normsigma}.
-
-We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as given by the $PRF(\lambda_i)$ function.
+In the following, we will use the term ``desaturation'' to describe the technique described here of correcting the charge values of the saturated pads.
+Figure~\ref{fig:satpad} shows a typical situation of saturated hits in a cluster.
+When an avalanche causes a large enough induced signal, the pads directly underneath the avalanche collect the largest charge becoming saturated, denoted here as $q_{2\textprime}$ and $q_{3\textprime}$.
+Pads further away collect less charge and typically are not saturated, detonated here as $q_{1}$ and $q_{4}$.
+Though the charge values in the saturated channels are lost, we know that the distribution of all charges must follow the PRF which is a fixed function, and a fundamental operating principle of all TPCs.
+We have already measured the PRF as a function of crossing angle in Section~\ref{sec:expPRF}, and from the tracking information, we know the crossing angle of the track at a given cluster and therefore the PRF corresponding to that cluster.
+
+We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as defined by the $PRF(\lambda_i)$ function.
 To determine the best estimate for the charge values of each saturated pad, a chi squared function is minimized,
@@
 
-where $q_i^{\mathrm{obs}}$ are the non-saturated charges and $q_i^{\mathrm{expect}}$ are the charge values expected from the PRF calculated as $q_i^{\mathrm{expected}} = Q\cdot PRF(\lambda_i)$.
-The saturated charge values $q_{i}^{\textprime}$ are treated as unknown variables and are allowed to vary in the $\chi^2$ minimization; the values calculated by the minimizer are added to the total expected charge $Q$.
-
-\begin{figure*}[t]
+where $q_i^{\mathrm{obs}}$ are the non-saturated charges and $q_i^{\mathrm{expect}}$ are the charge values expected for that pad as calculated from the PRF; $q_i^{\mathrm{expect}} = Q\cdot PRF(\lambda_i)$.
+The saturated charge values $q_{i}^{\textprime}$ are treated as unknown variables and are allowed to vary in the $\chi^2$ minimization, they enter the minimization routine when they are added to get the total charge $Q = \sum q_i + \sum q_i^{'}$.
+
+\begin{figure*}[!htb]
+\centering
 \includegraphics[width=\linewidth]{data_combine}
@@
 
+\begin{figure*}[!htb]
+\centering
+\includegraphics[width=\linewidth]{lowvshigh.png}
+\caption{Uncorrected (left panel) and desaturated (right panel) collision data comparing the low gain region to the high gain anode regions of the TPC.}
+\label{fig:lowvshigh}
+\end{figure*}
+
 Tracks which saturate pads in the high anode wire voltage region are not saturated in the low anode voltage region.
-By comparing the $\langle dE/dx\rangle$ values of these two sections, we can directly measure the success of the desaturation in the high gain regions.
+By comparing the $\langle dE/dx\rangle$ values of these two sections.
 While this desaturation technique avoids the need to lower the gain of any region, the low anode voltage region proved to be a direct measurement of the success of this technique.
@@
 Figure~\ref{fig:lowvshigh} shows the the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
-The effect of saturation can be seen in the high gain region for the uncorrected data above values of 400 ADC/mm where the values plateau, where as the low gain region still returns accurate values\footnote{Un-calibrated ADC channels in arbitrary units.} Below this value the electronics are not saturated, and therefore the high and low gain sections agree.
-
+The effect of saturation can be seen in the high gain region for the uncorrected data above values of 400 ADC/mm where the values plateau, where as the low gain region still returns accurate values.
+Below this value the electronics are not saturated, and therefore the high and low gain sections agree.
 After applying the desaturation method, the correlation between the high and low gain sections is restored, as seen in Fig.~\ref{fig:lowvshigh}.
@@
 
-The success of the desaturation becomes more clear when looking at the particle identification (PID) lines.
+The success of the desaturation becomes more clear when looking at the particle identification (PID) lines from the experimental data in Fig.~\ref{fig:data_combine}.
 In the following PID plots the red lines represent the most probable energy loss as given by Geant4 straggling functions.
-A linear calibration was performed to convert keV in Geant4 to ADC in the experiment given by $ADC/mm = 19;keV/cm$.
-
-There are pronounced PID lines of several particle species in both the uncorrected and corrected cocktail beam PID shown in the subplots of Fig.~\ref{fig:cocktail_combine}.
-The uncorrected data in the left panel shows the effects of saturation; the PID lines deviate from their theoretical expectations starting at around 400~ADC/mm eventually reaching a plateau.
-After applying the desaturation technique -- in the right panel-- we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation and even ${}^{6}Li$ and ${}^{7}Li$ can be separated.
-A more subtle improvement of the lighter particles, (p, d, t), can also be seen in the PID lines at lower momenta.
-
-A similar effect is seen in the experimental collision data.
-In Figure~\ref{fig:data_combine} though the PID suffers from more background and inefficiencies than the cocktail beam we see a similar improvement in the PID lines when comparing before -- the left panel-- and after applying desaturation -- the right panel.
-Notably the largest improvement is the separation of particle species at lower momenta and the separation of the Li species into ${}^{6}$Li and ${}^{7}$Li.
-In these regions, there was little to no PID resolution before desaturation.
-
-The dynamic range was extended by at least a factor of 5, as demonstrated by the improved PID lines, and quantified by direct comparison to low gain sections of the TPC.
+The uncorrected data in the left panel shows the effects of saturation, where the PID lines deviate significantly from their theoretical expectations starting at around 400~ADC/mm.
+After applying the desaturation technique -- in the right panel-- we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation.
+Even the ${}^{6}Li$ and ${}^{7}Li$ particles can be separated and a more subtle improvement of the lighter particles, (p, d, t), can also be seen in the PID lines at lower momenta.
+In these regions, there was little to no PID resolution before desaturation technique was applied.
 
@@
 As the beam passes through the field cage it ionizes the gas creating electron-ion pairs.
-The drift velocities of the ions are typically \SI{e4} times slower than the electron drift velocities \cite[blumrol].
-Any source of ions have the potential to build up creating a positive space charge distorting the drifting electrons, biasing the track momentum.
-There are several regions of the TPC in which ion are created.
+The drift velocities of the ions are typically \SI{e4} times slower than electron drift velocities \cite[blumrol].
+Any source of ions have the potential to build up in the drift volume, creating a positive space charge, distorting the drifting electrons and  biasing the track momentum measurement.
+There are several regions of the TPC in which ions are created.
 The largest source of positive ions is created in the avalanche process near the anode wires.
-These ions slowly drift toward the cathode.
-If the gating grid is closed the ions will terminate there; though it opens occasionally  for \SI{11}{\micro\second} the ions only move \SI{1}{\micro\metre} in this time and will be captured by grid when it closes.
-For this reason, the source of ions resulting from the avalanche region is negligible.
-
+But as discussed in Section~\ref{sec:wireplanes}, the gating grid captures all of the ions from this region.
 The other source of ions come from the primary ionization produced by the beam and reaction products in the detector gas.
 The energy loss  $\langle dE/dx\rangle \propto Z^2$, where Z is the charge of the particle type.
-Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2.5e3} times that of the light charged particles which mostly are of charge Z~1.
-It is the largest source of positive ions in the TPC.
+Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2e3} times that of the light charged particles which mostly are of charge Z~1.
+Therefore the ions resulting from the un-reacted beam is the largest source of positive ions in the TPC.
 
@@
 
- The beam is positioned about 25\si{\centi\metre} below the anode plane in the TPC.
-It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane where as it takes the ions \num{5e4}\si{\micro\sec}.
-The beam rate in the experiment varied around a value of about 10\si{\kilo\hertz}, which has an average occurrence of 1 beam every \SI{100}{\micro\sec}; this is shorter than the time it takes for the ions to terminate on the cathode plane, resulting in a build up of positive ions.
+ The beam is positioned about \SI{25}{\centi\metre} below the anode plane and \SI{29.6}{\centi\metre} above the cathode in the TPC.
+It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane where as it takes the ions \num{5e4}\si{\micro\sec} to drift to the cathode.
+The beam rate in the experiment was approximately \SI{10}{\kilo\hertz}, which has an average occurrence of 1 beam every \SI{100}{\micro\sec}.
+This is  much shorter than the time it takes for the ions to terminate on the cathode plane, resulting in a build up of positive ions.
 Figure~\ref{fig:spacechg_cartoon} gives an idea of the shape of the sheet of space charge carved out by the beam path.
 The ions from each beam create a line charge which drifts towards the cathode with a constant velocity.
-The average distance between sequential ion paths is about \SI{25}{\micro\metre} apart, therefore we expect the average number of beam paths that compose the sheet charge is around \num{1440} tracks -- where the distance to the cathode is \SI{29.6}{\centi\metre}.
-Though the arrival time of each beam track is random, the large number of tracks and small inter beam spacing allows us to approximate the sheet charge as an uniform sheet charge.
-
- The secondary beams entering the TPC are also composed of many species of particles distributed in a finite range around the target beam.
-Assuming a beam of ${}^{132}$Sn were the energy loss in gas is \SI{11.2}{\kilo\electronvolt\per\centi\metre}, using the number of beams above, the conversion to electrons given in \ref{tb:gasprop}, and the beam path length of \SI{135}{\centi\metre}, the total charge density expected is \SI{3e-8}{\coulomb\per\metre\squared}.
+The average distance between sequential ion paths is about \SI{25}{\micro\metre} apart, therefore we expect the average number of beam paths that make up the sheet charge is around \num{1440} tracks.
+Since the number of tracks in the sheet charge is large, and inter beam spacing is very small, this allows us to approximate the sheet charge as an uniform sheet charge density.
+
+ The secondary beams entering the TPC are also composed of many species of particles distributed in a finite range around the target beam as will be discussed in Section~\ref{sec:beam}.
+Let us assume a beam of ${}^{132}$Sn where the energy loss in P10 gas, at 270 AMeV, is \SI{11.2}{\kilo\electronvolt\per\centi\metre}.
+From the total number of beams above, and the beam path length of \SI{135}{\centi\metre}, the estimated charge density would be on the order of \SI{3e-8}{\coulomb\per\metre\squared}.
 
@@
  where $\phi$ is the electric potential and $\rho$ is the free charge.
-Since the Dirichlet boundary conditions of the field cage are defined by the cathode, walls, and pad plane, a numerical solution to the electric potential is solved through the ???
-technique.
-The electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.
-
- To reduce the need to compute a new electric field for a new value of the space charge, we notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ by the ratio $\rho/\rho_o$.
-The full magnetic field map is provided by the SAMURAI collaboration \cite{magnet}.
-The velocity field map can be calculated from Eq.~\ref{eq:elecdrift}.
-The path an electron drifts through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration from a given starting point of the electron.
-We do not add details of the electric field around the anode wires which is only relevant for the small gap between the ground and anode plane.
-Over this small gap the space charge corrections have very little effect.
-
- The correction map is calculated by starting from the anode y-position and the measured position on the pad-plane (x,z), and stepping backward in time in the Runge-Kutta integration through the velocity field map until the electron reaches the measured y-position, which is split into a grid that fills the 3-dimensional volume of the field cage.
-
- The correction is handled within the STCorrection task in the software.
-A data base of the sheet charge values for every run is read in and the electric field is calculated for the reference charge for a given beam path.
-The electric field and velocity map is scaled by the ratio of the space charge in the particular run and the reference run $\rho/\rho_o$ and the correction map is calculated.
-The measured clusters value (x,y,z) position is input into the correction map which interpolates and outputs the correction values $dx$ and $dz$ which correspond to that cluster.
-The cluster position is shifted to new positions $x\textprime = x + dx$ and $z\textprime = z + dz$, before going on to the momentum and vertex finding algorithms.
-
- It has been shown before that the amount of space charge present in the chamber is related to observables such as the distance of closest approach of each track to the vertex point \cite{starSC}.
-In the presence of no space charge, you would correctly expect the distance of closest approach of each track to the vertex point would be a distribution centered around the true vertex location.
+The numerical solution is provided by the Jacobi method \cite{poisson} provided all 6 sides have defined Dirchlet boundary conditions for the potentials.
+We neglect the wire plane region since the drift details are not important for the space charge effects we are discussing here.
+The pad plane and cathode are trivial, where the side walls of the field cage are given linearly varying potentials on the surface such that the electric field is the same strength as the real TPC with no space charge present.
+Once the electric potential is solved, the electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.
+
+It has been shown before that the amount of space charge present in the chamber is related to observables such as the distance-of-closest-approach (dOCA) of each track to the vertex point \cite{starSC}.
+In the presence of no space charge, the dOCA distribution of each track would be a centered around the true vertex location.
 Since the tracks are distorted by the space charge, which affects different regions of the TPC differently, a bias is introduced to the measured vertex location and widens the distribution to vertex of each track.
 
-An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} with left-going tracks shown in blue, and right-going tracks shown in green; the vectors show the direction of distortion and their magnitude have been magnified by ???
-times to show the detail.
-The dashed lines show the effects the track shifted due to the space charge map where left and right-going tracks are affected differently.
+An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} where left-going tracks shown in blue, and right-going tracks shown in green.
+The vectors show the direction of distortion, and their magnitude have been magnified by 10 times to show the detail.
+The dashed lines show the shifted track due to the space charge where left and right-going tracks are affected differently.
 Right-going tracks tend to higher momentum values and the left-going tracks going to lower momentum values, for positively charged particles.
@@
 \includegraphics[width=\linewidth]{Effect_SC.pdf}
-\caption{Shift of tracks}
+\caption{Example of the map of the electron shift.
+The effect is also shown on left-going (blue) and right-going (green) tracks.
+The original tracks are shown in the solid line and the shifted track in the dotted line.}
 \label{fig:sc_shift}
@@
 
-The inset figure of Fig.~\ref{fig:sc_shift} shows the shift in the x-position distance to vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
-We will define the DOCA for left going tracks as $V_x^L$ and for right-going tracks as $V_x^R$.
-We are able to measure the amount of distortion the space charge creates by measuring the difference between the most probable values of the left-going and right-going tracks which we define as $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ as seen in Figure~\ref{fig:VLR}.
-
-\begin{figure}[H]
+The inset figure of Fig.~\ref{fig:sc_shift} shows the effects of the space charge have on the x-component of the dOCA to the vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
+Figure~\ref{fig:VLR} shows the amount of distortion caused by the space charge effect as a new observable $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ , where  $V_x^L$ and $V_x^R$ are the most probable value of the dOCA distribution for left and right-going tracks respectively.
+
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{DVTP_raw.pdf}
-\caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks in the TPC for the }
+\caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks. }
 \label{fig:VLR}
@@
 
+To reduce the need to compute a new electric field for each space charge value, notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ by the ratio $\rho/\rho_o$.
+The full magnetic field map is provided by the SAMURAI collaboration \cite{magnet}.
+The velocity field map is calculated following Eq.~\ref{eq:elecdrift}.
+The electron drift through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration from a certain starting point.
+
+ The correction map is calculated by starting from the anode y-position and the measured position on the pad-plane (x,z), and stepping backward in time in the Runge-Kutta integration, through the velocity field map until the electron reaches the measured y-position.
+This is done over a 3-dimensional grid where points in between are interpolated by a trilinear interpolation.
+The measured clusters value (x,y,z) position is input into the correction map which outputs the interpolated correction values $dx$ and $dz$ which correspond to that cluster.
+The cluster position is then shifted to new positions $x\textprime = x + dx$ and $z\textprime = z + dz$.
+
 %What is the best figure to put in? VLR vs Charge density? 
@@
 Since the space charge affects right-going and left-going tracks differently making right-going tracks differently than left-going tracks, it appears they no longer originate from the BDC which is not affected by the space charge.
-For tracks at polar angles of $\theta < 40 \deg$, the disagreement between momentum values with and without the BDC are much less because the projection of the track does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
-Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ are more sensitive to small changes in the BDC when including it as an extra constraining point.
-After correcting for the space charge, adding the BDC point as an extra constraint on the track does not bias the reconstructed momentum value for either $\theta < 40 \deg$ or $\theta > 40 \deg$, which has the greatest improvement, as seen in Fig.~\ref{fig:mom_S_after} and Fig.~\ref{fig:mom_L_after} respectively.
-Recall that only the relative distance between left and right-going tracks was minimized; there was no guarantee that the corrected tracks coincide with the absolute BDC position at the target.
-This is one of the evidences of the correction's success.
-The others being the agreement with the expected cocktail calibration beam as described above.
-Later we will show how the space charge corrects the momentum distributions of left and right-going pions in the TPC.
+For tracks at polar angles of $\theta_{Lab} < \ang{40}$, the disagreement between momentum values with and without the BDC are much less because the projection of the track does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
+Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ which are more sensitive to small changes in the BDC when including it as an extra constraining point.
+After correcting for the space charge and adding the BDC point, the reconstructed momentum value agrees with or without the BDC included for both  $\theta_{Lab} < \ang{40}$ or $\theta_{Lab} > \ang{40}$, as seen in Fig.~\ref{fig:mom_S_after} and Fig.~\ref{fig:mom_L_after} respectively.
 
@@
         \includegraphics[width=\linewidth]{BDC_P_small_angle.pdf} 
-        \caption{Generic} \label{fig:mom_S_before}
+        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, before the space charge correciton.} \label{fig:mom_S_before}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{BDC_P_large_angle.pdf} 
-        \caption{Competitors} \label{fig:mom_L_before}
+        \caption{Comparison of the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} > \ang{40}$, before the space charge correction.} \label{fig:mom_L_before}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{BDC_P_aftercor_small_angle.pdf} 
-        \caption{Generic} \label{fig:mom_S_after}
+        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction. } \label{fig:mom_S_after}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{BDC_P_aftercor_large_angle.pdf} 
-        \caption{Competitors} \label{fig:mom_L_after}
+        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction.} \label{fig:mom_L_after}
     \end{subfigure}
@@
 \end{comment}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{sn132_SC} 
+        \caption{$\tin{132}{123}$ system.} \label{fig:132momdist_sc}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{sn108_SC} 
+        \caption{$\tin{108}{112}$ system.} \label{fig:108momdist_sc}
+    \end{subfigure}
+\label{fig:sc_momdist}
+\end{figure}
+
+Recall that only the relative distance between left and right-going tracks was minimized, and there was no guarantee that the corrected tracks coincide with the absolute BDC position at the target.
+This is one of the evidences of the correction's success.
+The others being the agreement with the expected cocktail calibration beam as described in Section~\ref{sec:cocktail}.
+
+Here we will discuss some of the final results of the pion kinetic energy spectrum as it pertains to the verification of the space charge analysis,  the details of the pion PID analysis is discussed in detail in Section~\ref{sec:pid}.
+Figure~\ref{fig:sc_momdist} shows the $\pi^+$ and $\pi^-$ kinetic energy distributions in the center of mass system for the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
+The momentum distributions are split into the beam-left and beam-right side of the TPC.
+In central collisions, one would expect no difference between the two momentum distributions due to the symmetry of the emission.
+Though as seen in the left panels of both figures, the data without the space charge correction is shown, where as in the right panel the data after the space charge correction is shown.
+There is a significant improvement in the matching of the left and right side momentum distributions for both particle species, in both systems.
+It is remarkable that the dOCA observable which was used to calculate the amount of space charge was able to also correct the pion momentum distributions.
+There is no reason to assume correcting the dOCA distribution would be related to the momentum distribution, and is the strongest evidence to the success of the space charge correction.
+
+\section{CoBo timing correction}
+The arrival time of signals originating from each pad may differ due to timing delays in the electronics and cabling.
+The timing differences will affect the y-position measurement of each track.
+The y-direction track residuals reveal that the timing differences correspond to about $\pm$\SI{2}{\milli\metre} in position differences.
+The timing difference is stable for each pad across several runs.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset6before_row_at_layer10.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset5before_row_at_layer30.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset4before_row_at_layer50.png}
+
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset6after_row_at_layer10.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset5after_row_at_layer30.png}
+    \includegraphics[width=0.33\textwidth]{run2907_yoffset4after_row_at_layer50.png}
+
+    \caption{Cobo timing correction.}
+  \end{center}
+  \label{fig:coboCorr}
+\end{figure}
+
+\begin{figure}[!htb]
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
+        \caption{Y-residual distribution for all pads before correction.} \label{fig:yoff_allBefore}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
+        \caption{Y-residual distribution for all pads after correction.} \label{fig:yoff_allAfter}
+    \end{subfigure}
+    \caption{ }
+\label{fig:yoff}
+\end{figure}
+
+ Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows, for three different layers.
+Before the correction, one can see large deviations in the y-residuals which correspond to timing differences.
+The mean value of the y-residual distribution is fitted and an inverse correction map is constructed for each pad.
+The data is then reconstructed subtracting value from the inverse map, $dy$, for each pad.
+The resulting corrected distribution is shown in Fig.~\ref{fig:coboCorr} for the same layer set.
+Figure~\ref{fig:yoff} also show the summary of all the pads before the correction and after the correction.
+There is a significant improvement in the width of the distribution going from \SI{1.5}{\milli\metre} to \SI{0.6}{\milli\metre}.
 
@@
 \section{Monte Carlo Simulation}
-The monte carlo (MC) simulation is composed of two separate simulations.
-The first simulation is an accurate model of the fundamental interactions of a particle passing through the various materials in the TPC; to accomplish this we have modled everything in Geant4 which also provies a particle generator.
-A scale model of field cage, front window, front window frame, pad plane, and aluminum top plate are modeled with the correct material type and dimensions assuming P-10 gas at a density of 1 atm.
+The MC simulation is composed of two separate simulations.
+The first simulation models the interactions of a particle passing through the various materials in the TPC using Geant4.
+A scale model of the field cage, and its materials, where the gas mixture is P-10 gas at a density of 1 atm.
 The magnetic field map of the SAMURAI dipole magnet was imported into Geant4 as well after assuming rotational symmetry along the axis perpendicular to the pole face.
-Along with the energy loss and particle transport, Geant4 also handles multiple scattering, and particle decay.
-The output of the particle generator is a series of energy loss points which contain the amount of energy lost in $\si{\kilo\electronvolt\per\centi\metre}$ and the location of the energy loss in Cartesian space $(x,y,z)$.
-
-In the second part of the MC simulation, the physical processes involved in the operation and measurement of the TPC are simulated, such as the electron drift, avalanche process, and all processes involved in the signal creation in the electronics, simulating the detector's response to a MC track.
+Along with the energy loss and particle transport, Geant4 also handles multiple scattering, and particle decay, which are all important effects for pions.
+The output of Geant4 is a series of energy loss points which contain the amount of energy lost in $\si{\kilo\electronvolt\per\centi\metre}$ and the location of the energy loss in Cartesian space $(x,y,z)$.
+
+In the second part of the MC simulation, the physical processes of the TPC measurement are simulated, such as the electron drift, avalanche process, and all processes involved in the signal creation in the electronics.
 This is separated into three software tasks, the drift, pad response, and electronics tasks.
-In the following section we will discuss these tasks in more detail, the discussion of Geant4 is not covered here and is left to the reader to explore.
+In the following section we will discuss these tasks in more detail, the discussion of Geant4 is not covered here and the reader is refered to \cite{geant4}.
 
 \subsection{Drift Task}
+
  The first step of the drift task is to convert the primary ionization points of the MC track provided by Geant4 into electrons.
-The average number of electrons created in a gas,$N_{e^-}$, can be described as,
+The average number of electrons created in a gaseous detector,$N_{e^-}$, can be described as,
 \begin{equation}
@@
 
-where $I$ is the ionization coefficient of P10 gas (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited in units of \si{\electronvolt}.
-Since the electric field set up by the field cage is uniform for the most of the area covered by the pad plane, we assume a straight line trajectory for each electron until it reaches an anode wire.
+where $I$ is the ionization coefficient of P10 gas (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited.
+Each electron is propagated along the electric field lines assuming the electric field is uniform, which is true for mos tof the pad plane region.
 The total length drifted from the initial primary ionization point to the final anode wire is given by $L_{anode}$.
@@
 \end{tabular}
-\caption{An overview of the properties of the TPC}
+\caption{An overview of electron drift properties in P10 gas.}
 \label{tb:gas}
@@
 This stochastic motion is  described by a a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
-The longitudinal ($c_{l}$) and transverse ($c_{t}$) diffusion coefficients are determined by Garfield++ calculation in the presence of a \SI{0.5}{\tesla} magnetic field, listed in Tb.~\ref{tb:gas}.
+The longitudinal ($c_{l}$) and transverse ($c_{t}$) diffusion coefficients are determined by Garfield++ calculation  \cite{garfield++} in the presence of a \SI{0.5}{\tesla} magnetic field, listed in Tb.~\ref{tb:gas}.
 The diffusion process is modeled by randomly sampling from a Gaussian distributions describing both of the directions.
-The random shift is then added to the final position of the electron.
+The random displacement vector is then added to the final position of the electron.
 The deviation in the transverse direction, $dr$, is randomly sampled from,
@@
 The final position of the electron along the wire is calculated  $x' = x + dx$.
-Since the electron will terminate on the closest anode wire the anode wire that is closest to the shifted z-position $z\textprime = z + dz$, the final electron z position is then updated to the anode wire it terminated on.
+The electron will terminate on the closest anode wire, that is the anode wire that is closest to the shifted z-position $z\textprime = z + dz$, and the final electron z position is then updated to be the same as the anode wire it terminated on.
 The total drift time of the electron $t$ is calculated as,
@@
 where $v_d$ is the drift velocity.
-The parameter $t_{offset}$ aligns the MC time bucket spectrum with the data.
-This is because the y-position corresponding to time zero in the data time bucket spectra does not correspond to the same position in the MC for time zero.
+The parameter $t_{offset} = \SI{49.92}{\nano\second}$ is to allow for an alignment of the MC time bucket spectrum with the data.
+This is because the y-position corresponding to $t=0$ in the data time bucket spectra, does not correspond to the same position in the MC.
 This value was found to be \SI{23}{\nano\second} by aligning the MC vertex to the data vertex.
 
-The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++ as seen in Fig.~\ref{fig:gaincalib} for the anode wire voltages used in the experiment.
-The number of electrons produced in the avalanche is randomly sampled from the corresponding distribution depending on which anode wire section the electron terminated; the number of electrons produced is stored as a number instead of multiple instances of the original electron to save space and computation time.
-We assume that the lateral movement along the wire arising form $\vec{E}\times\vec{B}$ effect near the wire is negligible as compared with diffusion and other effects described later.
-
-It is worth mentioning there is the possibility to simulate the space charge effects in this class where the inverse to the correction map described in \ref{sec:spacecharge} is applied to the drifting electrons.
-This is not used for calculating the response of the TPC for efficiency calculations since it is a trivial exercise to input the space charge map and correct for it using the same inverse map.
+The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++ and discussed in Section~\ref{sec:wireplanes} for the anode wire voltages used in the experiment.
+The number of electrons produced in the avalanche is randomly sampled from the distribution depending on which anode wire section the electron terminated.
+The number of electrons produced is stored as a gain factor instead of multiple instances of the original electron to save space and computation time.
+
+%We assume that the lateral movement along the wire arising form $\vec{E}\times\vec{B}$ effect near the wire is negligible as compared with diffusion and other effects described later.
+
+It is worth mentioning there is the possibility to simulate the space charge effects in this task where as the inverse to the correction map described in Section~\ref{sec:spacecharge}.
+This is not used for calculating the response of the TPC for efficiency calculations since it is a trivial exercise to input the space charge map only to then correct for it using the inverse map.
 
 \subsection{Pad Response Task}
-The total charge of each avalanche must be distributed according to the pad response function (PRF) described in section \ref{sec:prf}.
+The total charge of each avalanche is distributed according to the pad response function (PRF) described in Section~\ref{sec:prf}.
 The PRF is simulated as the double integral of a 2-dimensional Gaussian.
@@
 where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the final position of the drifted electron.
-The $\sigma$ parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
+The parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
 
@@
 
-The purpose of the electronics task is to simulates the electronics response to the charge induced on each pad, converting charge into ADC channels.
-The induced current in the pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
+The purpose of the electronics task is to simulate the electronics response to charge induced on each pad, converting charge into ADC channels.
+The induced charge on each pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
 The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, as long as the electronics settings are fixed.
 This allows us to assume the pulse shape is constant and can be described by two variables, the height of the pulse, $Q$, and the starting time bucket of the pulse, $t_o$.
-The starting time of the pulse is defined as the time at 10\% of the rising edge.
+The starting time of the pulse is defined as the time at .1$\cdot Q$ of  the rising edge.
 The shape of the pulse depends on the shaping time constant which was set to \SI{117}{\nano\second} for the data analyzed here.
@@
     \includegraphics[width=\linewidth]{satpulse.png} 
-    \caption{Standard pulse }
+    \caption{The standard pulse shape and saturated pulse shape extracted from the experimental data, and used in the MC software.}
     \label{fig:pulseshape}
@@
 \begin{equation}
-Q = f_G  A_{tot} \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{f_c}
+Q = f_G  N_{e}  \cdot e \cdot\frac{ADC_{max} - ADC_{pedestal}}{f_c}
 \label{eq:etoADC}
@@
 
-where $e$ is the charge of an electron in $\si{\femto \coulomb}$, $A_{tot}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
+where $e$ is the charge of the electron in $\si{\femto \coulomb}$, $N_{e}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
 The pulse shape given in Fig.~\ref{fig:pulseshape} is multiplied by the pulse height $Q$ giving the full time bucket estimate of the TPC response.
-Random Gaussian noise is added to each time bucket where the root-mean-squared value of the electronics noise was measured to be around 6 ADC.
+Random Gaussian noise is added to each time bucket, where the root-mean-squared value of the electronics noise was measured to be around 6 ADC.
 The timing information of the pulse is calculated from Eq.~\ref{eq:electronTime}.
@@
 where $\langle dE/dx\rangle_{MC}^p$ and $\langle dE/dx\rangle_{Data}^p$  are the energy loss values for a particle in a given momentum bin $p$. The calibration will be discussed later in more detail.
-
-\subsection{Simulating Saturation}
-\label{sec:simSat}
-
-There are several types of saturation that must be simulated or accounted for to correctly model the TPC response.
-They all are varying degrees of the same effect which manifest in different ways in the detector.
-In general all amplifiers have a finite range of output values set by a positive and negative rail (typically -12V and +12V).
-If the input charge into a charge sensitive pre-amplifier causes the output voltage to reach the max (or min) output voltage, the electronics is considered saturated and the response may be non-linear.
-The picture is complicated by the fact that there is usually an RC feedback loop which dissipates the input charge.
-The time a pre-amplifier returns to linear behavior depends on the input charge and how quickly it can be dissipated \cite{akiGET}.
-The pad is otherwise considered dead and no further charge can be measured until the channel can recover.
-
-Due to the long, high energy tail in  the energy loss distribution of a particle traversing matter, it is common to see pads along the track which have collected large energy loss values saturating the electronics of that pad.
-This occurs for all tracks, even for minimum ionizing tracks; saturation in this case is infrequent and not an issue as there are many other clusters are not saturated.
-Yet as the charge of the particle gets higher, or the momentum gets lower, the most probable energy loss value also becomes higher, and a significant fraction of pads in a track may be saturated.
-In Section~\ref{sec:desat}, we outlined an algorithm to correct for the saturation for a track, where as here we are interested in the effects saturation has on surrounding tracks to properly simulate the effect in the MC simulation.
-
-Pads that saturate have no signal or ''dead" for a given amount of time which depends on the amount of input charge.
-For charges that are relatively at or above the level of saturation, the pad will certainly be dead for the remainder of time after the saturating event.
-Signals that would have arrived later in time from other tracks passing underneath the saturating track will therefore not be measured; the saturated pad is ``shaddowing" any future signals.
-As the track multiplicity of an event increases, the probability earlier tracks shaddow later tracks increases.
-This will change event by event and is very difficult to simulate except through a MC track embedding approach which will be discussed later.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\linewidth]{satpulsepz.png} 
-\caption{Pole zero correction} 
-\label{fig:pulseSatTag}
-\end{figure}
-
-In the case where the input signal is very large, the electronics can be dead for up to \SI{35}{\milli\second} \cite{akiGET}.
-The beam rate in the experiment was about \SI{10}{\kilo\hertz}, which corresponds to an average of \SI{100}{\micro\second} between subsequent beams.
-In this case channels can be effectively dead for several events before recovering.
-Due to the large charge of the beam, a large amount of very high energy electrons --delta rays-- are produced from the beam passing through the gas.
-In the presence of the magnetic field the radius of curvature of most of the electrons are within several \si{cm} in the xz-plane.
-Figure~\ref{fig:deltaE} shows the horizontal extent of the delta rays in a top down view of the TPC.
-While some of the electrons can stop in the gas, some electrons have a high enough kinetic energy going in the vertical direction, penetrating the gating grid without being blocked.
-They then either terminate on the pad or possibly deposit their charge directly in the pad.
-The charge induced on the pad is large enough to kill the pad for a time long enough to last until at least the next event.
-This type of saturation is randomly distributed for pads sitting directly in the beam path and are dead for that event and the next event.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{deltaE_topview_neg} 
-        \caption{Top view} \label{fig:deltaE_topview}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{deltaE_sideview_neg} 
-        \caption{Side view} \label{fig:deltaE_sideview}
-    \end{subfigure}
-    \caption{Geant4 simulation of ${}^{108}$Sn beam at 270 MeV/A in P10 gas. Notice the extent of the delta electrons in the vertical direction as compared to the horizontal extent. }
-\label{fig:deltaE}
-\end{figure}
-
-\begin{figure}[!htb]
-\includegraphics[width=\textwidth]{padplane_sat.png}
-\caption{Tagging the saturated pads.}
-\label{fig:satTag}
-\end{figure}
-
-The simulation of these saturation effects is handled naturally by embedding MC tracks into real data.
-Once the time bucket of the saturating signal has been identified, no further signals are embedded since the pad is assumed to be dead.
-In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded at all.
-
-The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
-The long exponential tail can be effectively removed by a simple software technique which is similar to the electronics concept of ``pole-zero compensation".
-If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,
-
-\begin{equation}
-f_i^{'} = \frac{-b_1\cdot f_{i-1}^{'} + a_o\cdot f_i + a_1 \cdot f_{i-1}}{b_o}
-\label{eq:satpolez}
-\end{equation}
-
-where $a_o$ = .9723, $a_1$=-.9453, $b_o$=.9545, and $b_1$=-.9203.
-These coefficients correspond to the case where the long tail of the standard non-saturating goes to zero, without producing any negative undershoots.
-In the case of applying this technique to the saturated pulse, the correction produces a negative undershoot since there was no long exponential tail to begin with.
-Figure~\ref{fig:pulseshape} shows the pulses as the red curves, and after the pole-zero compensation as the blue curves for both the normal and saturated pulses.
-
-By applying this technique to the real data we can identify whether a pad has a saturating pulse in it and what is the time bucket location by looking for this negative undershoot.
-A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ for more than 8 time buckets and the max ADC value is > $G\cdot$500 where $G$ is the gain calibration of the pad.
-The max ADC condition eliminates false positives which come from  dead pads where the gating grid subtraction is still applied introducing false negative peaks.
-The pad is then flagged as saturated and the data time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket of the negative peak, since the falling edge is around 5 time buckets after the beginning of the saturation.
-A separate time is also stored called the MC time bucket position which sets the point in time in which a signal cannot be embedded any further.
-This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible.
-
-Saturation can also occur if of all the pulse heights within the time bucket spectrum add to a value over the max ADC threshold of 3500 ADC.
-This arises because the fall time of the pre-amplifier circuit is much longer than the time bucket measurement window and pulses are allowed to pile up in the pre-amplifier.
-In this case no single pulse will reach over the max ADC value yet the last pulse to create saturation will have a pulse shape that has no long tail but otherwise look like a normal pulse.
-These pulses are also identified by the algorithm described above since there is still no long tail.
-When embedding tracks into pads, we do not consider the previous amount of total charge in the pad.
-We assume this method of saturation is a higher order correction and the current algorithm approximates most of the saturation effects which progress through the first process.
-Figure~\ref{fig:satTag} shows all of the pads tagged by the algorithm as saturated for an event in the TPC.
-The max ADC values in a given pad are colored and the black dots represent saturated pads.
-Some pads appear to be below 3500 ADC, but upon further inspection they satisfy the second process of saturation where the sum is greater than the max ADC value.
-
-Dead pads are identified earlier in the software in the STCore class.
-A dead pad is simply a pad which only contains electronic noise and no signals.
-A dead pad is identified by having a total ADC r.m.s.
-value  of less than 50 ADC and a max ADC < 50 over the full time bucket spectrum.
-Here the dead pad is also tagged as saturated and the time bucket position of saturation is set to the first time bucket.
-
-NEED TO ADDRESS WHY THE MAX ADC THRESHOLD IS 3500
-%maybe simple pream fig
-%maybe example from aki paper about times
-%maybe
-
-%Add figure showing saturated time bucket spectra with location of saturation identified and with pulse shape from embedding I would like to add and how it blocks it.
-%Add Figure with 2D pad plane response with and without saturation flag
 
@@
 \includegraphics[width=\textwidth]{FlowEmbedding.pdf}
-\caption{Flow of embedding implementation in the software.}
+\caption{Flow diagram of the embedding software.}
 \label{fig:flow}
@@
     \end{subfigure}
-    \caption{A 200 MeV/c $\pi^-$ embedded into a nuclear collision type event. The embedded track identified by the software is highlighted by the solid green line. }
-
+    \caption{}
 \label{fig:embedtrack}
@@
         \includegraphics[width=\linewidth]{perspective_embed}
-        \caption{Generic} \label{fig:persEmbed}
+        \caption{Perspective view.} \label{fig:persEmbed}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{top_embed} 
-        \caption{Competitors} \label{fig:topEmbed}
+        \caption{Top down view.} \label{fig:topEmbed}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{side_embed} 
-        \caption{Competitors} \label{fig:sideEmbed}
+        \caption{Side view.} \label{fig:sideEmbed}
     \end{subfigure}
@@
 
+\subsection{Simulating Saturation}
+\label{sec:simSat}
+
+In general all amplifiers have a finite range of output values set by a positive and negative rail (typically -12V and +12V).
+If the input charge into a charge sensitive pre-amplifier causes the output voltage to reach the max (or min) output voltage, the electronics is considered saturated and the response may be non-linear.
+The picture is complicated by the fact that there is usually an RC feedback loop which dissipates the input charge.
+The time a pre-amplifier returns to linear behavior depends on the input charge and how quickly it can be dissipated \cite{akiGET}.
+The pad is otherwise considered dead and no further charge can be measured until the channel can recover.
+There are several types of saturation that must be simulated or accounted for to correctly model the TPC response.
+They all are varying degrees of the same effect which manifest in different ways in the detector.
+
+Due to the long, high energy tail in  the energy loss distribution of a particle traversing matter, it is common to see pads along the track which have collected large energy loss values saturating the electronics of a couple of pads.
+This occurs for all tracks, even for minimum ionizing tracks.
+Saturation in this case is infrequent and not an issue as there are many other clusters are not saturated.
+As the charge of the particle gets higher, or the momentum gets lower, the  energy loss value also becomes higher, and a significant fraction of pads in a track may be saturated.
+In Section~\ref{sec:desat}, we outlined an algorithm to correct for the saturation for a  particular track, but this saturation also affects surrounding tracks.
+In order to make an accurate MC simulation, we must understand how saturation affects surrounding tracks.
+
+Pads that saturate have no signal or ''dead" for a given amount of time  after the saturating signal, which depends on the amount of input charge.
+For charges that are at or above the level of the dynamic range, the pad will certainly be dead for the remaining time of the measurement.
+Signals from tracks passing directly underneath this pad, with signals that would arrive later in time,  will therefore not be observed.
+In a sense the saturated pad is ``shadowing" any future tracks.
+As the track multiplicity of an event increases, the probability of track shadowing increases.
+This will change event by event and is very difficult to simulate except through a MC track embedding approach, which will be discussed in Section~\ref{sec:embedding}.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\linewidth]{satpulsepz.png} 
+\caption{Example of the pole zero correction when applied to the normal pulse and saturated pulse.} 
+\label{fig:pulseSatTag}
+\end{figure}
+
+When the input signal is very large, the electronics can be dead for up to \SI{35}{\milli\second} \cite{akiGET}.
+The beam rate in the experiment was about \SI{10}{\kilo\hertz}, which corresponds to an average of \SI{100}{\micro\second} between subsequent beams.
+In this case, channels can be effectively dead for several events before recovering.
+There is a large amount of high energy electrons, commonly referred to as ``delta rays",  which are produced from the beam passing through the gas, due to the high atomic charge of the beam.
+Horizontally almost all of the electrons cannot travel very far, even for very high energy electrons, due to the vertical magnetic field causing them to curl in tight circles.
+Figure~\ref{fig:deltaE} shows the horizontal and vertical extent of the delta rays in a top down and side views of the TPC.
+While many electrons can stop in the gas, some electrons have a high enough kinetic energy in the vertical direction where they can penetrate the gating grid without being blocked.
+Then they either terminate on the anode wire or possibly deposit their charge directly in the pad.
+In either case, the charge induced on the pad is large enough to kill the pad for a time long enough to last until at least the next event.
+This manifests as random dead pads in the experiment, that follow the beam path.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{deltaE_topview_neg} 
+        \caption{Top view} \label{fig:deltaE_topview}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{deltaE_sideview_neg} 
+        \caption{Side view} \label{fig:deltaE_sideview}
+    \end{subfigure}
+    \caption{Geant4 simulation of ${}^{108}$Sn beam at 270 MeV/A in P10 gas. Notice the extent of the delta electrons in the vertical direction as compared to the horizontal extent. }
+\label{fig:deltaE}
+\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=\textwidth]{padplane_sat.png}
+\caption{Results of the algorithm to tag saturated pads, where the black dot denotes a pad was saturated. The charge values plotted is the maximum charge in the pad plane for the entire event.}
+\label{fig:satTag}
+\end{figure}
+
+The simulation saturation effects is handled naturally by embedding MC tracks into real data; once the time bucket of the saturating signal has been identified, simply no further signals are embedded since the pad is assumed to be dead.
+In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded for any time bucket.
+We only need to identify where the time of saturation occurs.
+The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
+The long exponential tail can be effectively removed by a simple software technique which is similar to the electronics concept of ``pole-zero compensation".
+If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,
+
+\begin{equation}
+f_i^{'} = \frac{-b_1\cdot f_{i-1}^{'} + a_o\cdot f_i + a_1 \cdot f_{i-1}}{b_o}
+\label{eq:satpolez}
+\end{equation}
+
+where $a_o$ = .9723, $a_1$=-.9453, $b_o$=.9545, and $b_1$=-.9203.
+These coefficients were tuned to remove the long tail of the standard non-saturating pulse, without producing any negative undershoots.
+If the same procedure is applied to the saturated pulse, the correction produces a negative undershoot, since there was no long exponential tail to begin with.
+Figure~\ref{fig:pulseshape} shows the pulses as the red curves, and after the pole-zero correction as the blue curves for both the normal and saturated pulses.
+
+By applying this technique to the real data we can identify whether a pad has a saturating pulse in it by looking for this negative undershoot.
+A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ ADC for more than 8 time buckets and the max ADC value is > $G\cdot$500; $G$ is the gain calibration coming from low gain sections discussed in Section~\ref{sec:anodeCalib}.
+The max ADC condition eliminates false positives which come from  dead pads where the gating grid subtraction is still applied, introducing large false negative peaks, but have no positive signals.
+After meeting these two conditions, the pad is flagged as saturated and the time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket position of the negative peak.
+This is because the falling edge since the falling edge is around 5 time buckets after the beginning of the saturation.
+A separate time is also stored called the MC time bucket position, which sets the point in time in which a signal cannot be embedded any further.
+This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible.
+These two positions are also shown in Fig.~\ref{fig:pulseSatTag}.
+
+There is another way a pad can saturate which occurs when all the pulse heights within the time bucket spectrum add to more than the max ADC threshold of 3500 ADC.
+This arises because the fall time of the pre-amplifier circuit is much longer than the time bucket measurement window, and pulses are allowed to pile up in the pre-amplifier.
+In this case no single pulse will reach over the max ADC value yet the last pulse will have a pulse shape that is missing the long characteristic tail, but otherwise looks like a normal pulse.
+These pulses are also identified by the algorithm described above, since this type of saturated pulse is still no long tail.
+When embedding signals into pads, we do not consider the previous amount of total charge in the pad.
+We assume this type of saturation is a higher order correction, where and the current algorithm approximates most of the saturation effects.
+Figure~\ref{fig:satTag} shows all of the pads in an event which are tagged by the algorithm as saturated, denoted by the black dots.
+The max ADC values in a pad are shown in the z-axis color scale, to give a sense of the success of the algorithm.
+It may appear that the ADC values of some pads appear to be lower than 3500 ADC, but upon further inspection they all satisfy the mode of saturation where the total sum of heights is greater than the max ADC value as described earlier.
+
+Dead pads are identified earlier in the software when reading in the raw data (STCore class).
+A dead pad is simply a pad which only contains electronic noise and no signals.
+A dead pad is identified by having a total ADC r.m.s.
+value  of less than 50 ADC and a max ADC < 50 over the full time bucket spectrum.
+Here the dead pad is also tagged as saturated and the time bucket position of saturation is set to the first time bucket, or time zero.
+
+%NEED TO ADDRESS WHY THE MAX ADC THRESHOLD IS 3500
+%maybe simple pream fig
+%maybe example from aki paper about times
+%maybe
+
+%Add figure showing saturated time bucket spectra with location of saturation identified and with pulse shape from embedding I would like to add and how it blocks it.
+%Add Figure with 2D pad plane response with and without saturation flag
+
 \subsection{MC and Data Comparison}
 
-Here we compare several important observables to ensure the MC is simulating the data.
-These observables will be relevant for the quality cuts and efficiency analysis discussed in Section~\ref{sec:efficiency}.
+Here we compare several important observables to ensure the MC is simulating the data sufficiently.
+These observables will be relevant later in the discussion of quality cuts and efficiency analysis discussed in Section~\ref{sec:efficiency}.
 The most important observable is the total number of clusters.
 The number of clusters in a track depends on the geometry of the TPC, the spherical angles of the track, and the clustering algorithm.
-MC tracks were embedded into data for several particle types; the input angular and momentum distribution was uniform and the final distributions were weighted such that the angular and momentum distribution matched that of the data for a fair comparison.
+MC tracks were embedded into data for several particle types.
+The input angular and momentum distribution was uniform where the final distributions were weighted such that the angular and momentum distribution matched that of the data for a fair comparison.
 Figure~\ref{fig:clustcomp} shows the normalized cluster distributions for the MC and data for several particle types.
@@
 \includegraphics[width=\linewidth]{numcluster.png}
-\caption{Comparison of MC to data.}
+\caption{Comparing the distribution of the number of clusters in MC embedded tracks to experimental data.}
 \label{fig:clustcomp}
@@
 
-The second most important observable is the distance to vertex.
-Figure~\ref{fig:pocacomp} shows the distance to vertex distributions for the MC embedded tracks and the data for the same range of particle species.
+The second most important observable is the distance-of-closest-approach (dOCA) to the vertex.
+Figure~\ref{fig:pocacomp} shows the dOCA distributions for the MC embedded tracks and the data, for the same range of particle species.
 The distributions were again normalized to different values to display on the same scale.
-Both data sets here are track number of clusters > 20.
+Here the only other cut was the  track number of clusters cut > 20, which cleaned up the background spectrum.
 It is also worth mentioning the data plotted is also after the space charge correction.
@@
 \includegraphics[width=\linewidth]{poca.png}
-\caption{Comparison of MC to data.}
+\caption{Comparison of the dOCA distribution of MC embedded tracks to data.}
 \label{fig:pocacomp}
@@
 
-We can see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
+We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
 It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.
@@
          \includegraphics[width=\textwidth]{PRFs_MC_wcut.png}
-         \caption{PRF response of Monte Carlo simulation of $\pi^-$}
+         \caption{PRF response of the MC $\pi^-$ tracks.}
          \label{fig:prfpimMC}
@@
 %Add Figure of Number of clusters of MC vs Data
-Add Figure of dEdx MC vs Data
-Add Figure of Momentum resolution MC vs Data
-Add Figure of track residuals? MC vs data?
-
-\section{Aligning TPC}
+%Add Figure of dEdx MC vs Data
+%Add Figure of Momentum resolution MC vs Data
+%Add Figure of track residuals? MC vs data?
+
+%\section{Aligning TPC}
 
@@
 
-\section{CoBo timing correction}
-The arrival time of signals originating from each pad may differ due to timing delays in the electronics and cabling.
-The timing differences will affect the y-position measurement of each track.
-The y-residuals reveal that the timing differences correspond to about $\pm$\SI{2}{\milli\metre} in position differences.
-The timing difference is stable for each pad across several runs.
-
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset6before_row_at_layer10.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset5before_row_at_layer30.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset4before_row_at_layer50.png}
-
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset6after_row_at_layer10.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset5after_row_at_layer30.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset4after_row_at_layer50.png}
-
-    \caption{TBW\label{TBD}}
-  \end{center}
-  \label{fig:coboCorr}
-\end{figure}
-
-\begin{figure}[!htb]
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
-        \caption{Distribution for all pads before correction.} \label{fig:yoff_allBefore}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
-        \caption{Distribution for all pads after correction.} \label{fig:yoff_allAfter}
-    \end{subfigure}
-    \caption{ }
-\label{fig:yoff}
-\end{figure}
-
- Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows , for three different layers.
-Before the correction one can see large deviations in the y-residuals which correspond to timing differences.
-The mean value of the y-residual distribution is fitted and a map is constructed for each pad.
-The data is then reconstructed subtracting the value of that map, $dy$, for each pad.
-The resulting corrected distribution is shown in Fig.~\ref{fig:coboCorr} for the same layer set.
-Figure~\ref{fig:yoff} also show the summary of all the pads before the correction and after the correction.
-There is a significant improvement in the width of the distribution going from \SI{1.5}{\milli\metre} to \SI{0.6}{\milli\metre}.
-
 \section{Efficiency Corrections}
 \label{sec:efficiency}
-Add Figures of efficiency vs angles in TPC polar angle plot for pions
 
@@
         \includegraphics[width=\linewidth]{pim_200_efficiency.png} 
-        \caption{Flattened angle plot} \label{fig:pim_flat_eff_ex}
+        \caption{Cartesian representation of the same efficiency plot  as in the polar plot.} \label{fig:pim_flat_eff_ex}
     \end{subfigure}
@@
 
-The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database. The correction factor $C_i$ is defined as,
-
-\begin{equation}
-C_i = \epsilon_i^{-1}.
-\end{equation}
-
-The track is then weighted by the factor $C_i$ when plotting in any type of graph for any observable.
+\section{PID Analysis}
+\label{sec:pid}
+%PID lines over all
+%PID of pion cut
+%background estimation

### 2020-06-10 chapters/ch_data_analysis_2.tex
Backup

@@
 \section{Radio Isotope Beam Factory (RIBF) Facility }
+\label{sec:beam}
 %Cyclotron facility overview.
@@
 \end{figure}
-The high track multiplicity in each event is quite high and since the tracks all originate from a common vertex, the density of tracks near the target region is very high.
+The track multiplicity in each event is quite high and since the tracks all originate from a common vertex, the density of tracks near the target region is very high.
 In this region the track separation is too small for the software to correctly determine and sort the charge information into the appropriate track.

### 2020-06-10 chapters/ch_experiment.tex
Still working on analysis section. backup
Backup

@@
 \subsection{Wire Planes}
+\label{sec:wireplanes}
 %Add figure of gating grid transparency closed and open configuration
@@
 \section{Energy loss in material}
+\label{sec:energyloss}
+
 The average energy loss in a material can be described by the Bethe-Bloch equation,
@@
 C = \frac{1}{n} \sum\limits_{i}^{n_t} \Delta_i/x,
+\label{eq:truncmean}
 \end{equation}
@@
 
-where $\mu$ is the signed drift velocity, $\omega$ is the cyclotron frequency, and $\tau$ is the collision parameter for a particular gas \cite{blumrol}.
+where $\mu=\SI{5.43}{\centi\metre\per\second}$ is the signed drift velocity, $\omega=\SI{8.79e10}{\radian\per\sec}$ is the cyclotron frequency, and $\tau=\SI{2.48e-11}{\second}$ is the collision parameter for a particular gas \cite{blumrol}.
 
@@
 \end{table}
-
-ADD THE VALUES OF TAU AND MU HERE
 
@@
 
+\section{Beam Drift Chambers (BDC)}
+\label{sec:bdc}
+
 \section{Ancillary Detectors }
@@
 \subsection{Kyoto Multiplicity Trigger}
+\label{sec:kyoto}
 %kyoto array sets multiplicity trigger

### 2020-06-10 chapters/ch_results.tex
Routine upload
Still working on analysis section. backup
Backup

--- 2019-06-09 chapters/ch_results.tex
@@
 \chapter{Results}
+\cite{jon}
+
+\section{Pion Spectra}
+
+The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database. The correction factor $C_i$ is defined as,
+
+\begin{equation}
+C_i = \epsilon_i^{-1}.
+\end{equation}
+
+The track is then weighted by the factor $C_i$ when plotting in any type of graph for any observable.
+
+%need to add error analysis
+
+%Efficiency correction
+%space charge before and after
+%left and right of space charge 
+%error analysis
+%systematic error analysis
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{pionSpectra.png}
+\caption{Pion spectra. }
+\label{fig:pionspectra}
+\end{figure}
+
+%Add figure on Glauber model 
+%Add figure on comparison to FOPI data
+
+\section{Pion Yield}
+
+\begin{table*}\centering
+\ra{1.3}
+\begin{tabular}{@{}cccc@{}}\toprule
+System & $\pi^-$ & $\pi^+$ & $Y(\pi^-)/Y(\pi^+)$  \\
+\midrule
+$\tin{132}{124}$ & 0.717(24)(4) & 0.148(5)(2) & 4.84(10)(6)  \\
+$\tin{108}{112}$ & 0.399(14)(3) & 0.200(8)(2) & 1.99(4)(3)  \\
+%$\tin{132}{124}$ & \numerr{0.717}{0.024}{0.004} & \numerr{0.148}{0.005}{0.002} & \numerr{4.84}{0.10}{0.06}  \\
+%$\tin{108}{112}$ & \numerr{0.399}{0.014}{0.003} & \numerr{0.200}{0.008}{0.002} & \numerr{1.99}{0.04}{0.03}  \\
+\bottomrule
+\end{tabular}
+\caption{Total pion yield.}
+\label{tb:pionyield}
+\end{table*}
+
+\section{$\pi^-$/$\pi^+$ Ratio}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{singleRatio.png}
+\caption{Single ratio spectra}
+\label{fig:SRspectra}
+\end{figure}
+
+%Add figure of Theory for pion ratios
+
+\section{Pion Double Ratio}
+%Add figure of Theory for pion ratios
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{doubleRatio.png}
+\caption{}
+\label{fig:spectraDR}
+\end{figure}
+
+\section{Comparison to Previous Data Sets (FOPI)}
+asdfasdfasd
 \section{Comparison to Theory}
-\section{$\pi$ Yields}
-\section{$\pi^-$/$\pi^+$ Ratio}
+%Add figure of Pion spectra for different theory 
+%Reference paper for multiple theories
+asdfasdf
+
+\section{Systematic Errors and Cut Variations}
+\label{sec:cutvar}
+dfsadfsaf
 \section{Constraint on the Symmetry Energy}
+%maybe not...

### 2020-06-10 macrosForThesis/PionMultiplicity/AuAuData/SpiRITStyle.C

@@ -1,0 +1,160 @@
+namespace style2
+{
+  TCanvas *square    (TString name="",double w=0,double h=0);
+  TCanvas *cc   (TString name="",double w=0,double h=0);
+
+   TGraph *make (TGraph  *graph);
+      TH1 *make (TH1     *h);
+      TF1 *make (TF1     *f);
+  TLegend *make (TLegend *legend);
+  TCanvas *make (TCanvas *cvs);
+
+     void save  (TCanvas *cvs,TString format="png");
+
+  /********************************************************/
+
+  void init();
+
+     int fICvs=0;
+     int fWCvs = 1600;
+     int fHCvs = 900;
+
+     int fWCvs_Sq = 600;
+     int fHCvs_Sq = 600;
+
+  double fWUnit=0.006;
+  double fHUnit=0.035;
+
+  double fWStat=0.25;
+  double fHStat=0.18;
+  double fXStat=0.95;
+  double fYStat=0.88;
+
+  double fTMargin=0.12;
+  double fBMargin=0.10;
+  double fLMargin=0.12;
+  double fRMargin=0.055;
+  double fRMarginH1=0.03;
+  double fRMarginH2=0.155;
+
+  double fMainTitleSize=0.05;
+  double fAxisLabelSize=0.03;
+  double fAxisTitleSize=0.04;
+  double fXTitleOffset=1.15;
+  double fYTitleOffset=1.35;
+};
+
+void style2::init() {
+  fXStat=1.-fRMargin;
+  fYStat=1.-fTMargin;
+  gStyle->SetStatStyle(0);
+  gStyle->SetStatW(fWStat);
+  gStyle->SetStatH(fHStat);
+  gStyle->SetStatX(fXStat);
+  gStyle->SetStatY(fYStat);
+  gStyle->SetTitleFontSize(fMainTitleSize);
+
+  gStyle -> SetTitleFont(22);
+  gStyle -> SetLabelFont(22,"X");
+  gStyle -> SetLabelFont(22,"Y");
+  gStyle -> SetOptStat(0);
+  gStyle -> SetCanvasColor(10);
+  gStyle -> SetFillColor(10);
+  gStyle -> SetTitleColor(10);
+
+}
+
+TCanvas *style2::square(TString name,double w,double h) {
+  fRMargin=fRMarginH1;
+  init();
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/AuAuData/plotLitData.C

@@ -1,0 +1,126 @@
+
+
+void plotLitData()
+{
+  double au_au_energy[6] = {.409,.600,.799,1.00,1.2,1.5};
+  double au_au_ma[6]     = {.0157,.0430,.0795,.11848,.1631,.22834};
+  
+  double au_au_x[6]     = {0,0,0,0,0,0};
+  double au_au_y[6]     = {.002,.004,.007,.01,.015,.03};
+
+  double pim = 1.1738;
+  double pip = .2188;
+
+  //  double apart = 230;  // 3.45 fm b value MC estimated glauber model
+  double apart = 256;  //0 fm b value MC estimated glauber model
+  //  double apart = 245.48;  //2.42 fm b value MC estimated glauber model
+  double t_pion = 1.5*(pim + pip)/apart;
+  cout<<t_pion<<endl;
+  double error_pion = .1;
+  double error_apart = 13;
+
+  double error = 1.5 * (pim + pip)/apart * sqrt( pow(error_pion/(pim+pip),2) + pow( error_apart/apart,2));
+  
+  cout<<"Error is "<<error<<endl;
+  
+  TGraphErrors *tpc = new TGraphErrors(1);
+  tpc->SetPoint(1,.27,t_pion);
+  tpc->SetPointError(1,0,error);
+  tpc->SetMarkerColor(3);
+  tpc->SetMarkerStyle(20);
+  tpc->SetMarkerSize(2.);
+  tpc->SetLineWidth(2);
+  tpc->SetLineColor(1);
+  
+  TGraphErrors *litdata = new TGraphErrors(6,au_au_energy,au_au_ma,au_au_x,au_au_y);
+  litdata->SetLineColor(1);
+  
+  TF1 *fit = new TF1("fit","TMath::Exp([0] + [1]/x + [2]/pow(x,2))",0,2);
+  fit->SetParameters(-.654307,-1.9530607,.20315227);
+
+  //  TF1 *fit_2 = new TF1("fit_2","TMath::Exp([0] + [1]/x + [2]/pow(x,2))",0,2);
+  //  fit_2->SetParameters(-.5405,-2.1372,.265309);
+  /*
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/AuAuData/plotLitData_Scaled.C

@@ -1,0 +1,164 @@
+
+
+void plotLitData_Scaled()
+{
+  gStyle->SetEndErrorSize(10);
+
+  /*
+  double au_au_energy[8] = {.27,.409,.600,.799,1.00,1.2,1.5};
+  double au_au_ma[8]     = {.004195,.0157,.0430,.0795,.11848,.1631,.22834};
+  
+  double au_au_x[8]     = {0,0,0,0,0,0,0};
+  double au_au_y[8]     = {.002,.002,.004,.007,.01,.015,.03};
+  */
+
+  double au_au_energy[7] = {.409,.600,.799,1.00,1.2,1.5};
+  double au_au_ma[7]     = {.0157,.0430,.0795,.11848,.1631,.22834};
+  
+  double au_au_x[7]     = {0,0,0,0,0,0};
+  double au_au_y[7]     = {.002,.004,.007,.01,.015,.03};
+
+  //  double pim = 1.1738;
+  //  double pip = .2188;
+
+  //  double pim = .626; //as of 06/02/19
+  //  double pip = .1235;  //as of 06/02/19
+
+  //  double pim = .59; //as of 06/10/19
+  //  double pip = .12;  //as of 06/10/19
+
+  double pim = .6968; //as of 08/30/19
+  double pip = .1486;  //as of 08/30/19
+
+  //  double apart = 230;  // 3.45 fm b value MC estimated glauber model
+  //  double apart = 244.7;  //0 fm b value MC estimated glauber model
+  double apart = 245.48;  //2.42 fm b value MC estimated glauber model
+  //  double t_pion = 1.5*(pim + pip)/apart;//we use geometric mean now
+
+  double error_apart = 13;
+
+  auto pimData132 = pim;
+  auto pimData132Err = 0.00364318;
+  auto pipData132 = pip;
+  auto pipData132Err = 0.00170804;
+
+  auto piTotalData132 = pimData132 + pipData132 + TMath::Sqrt(pimData132*pipData132); 
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/AuAuData/plotNZData.C

@@ -1,0 +1,394 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+
+void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
+                     Float_t lMargin, Float_t rMargin,
+                     Float_t bMargin, Float_t tMargin)
+{
+  if (!C) return;
+
+  // Setup Pad layout:
+  Float_t vSpacing = 0.0;
+  Float_t vStep  = (1.- bMargin - tMargin - (Ny-1) * vSpacing) / Ny;
+
+  Float_t hSpacing = 0.0;
+  Float_t hStep  = (1.- lMargin - rMargin - (Nx-1) * hSpacing) / Nx;
+
+  Float_t vposd,vposu,vmard,vmaru,vfactor;
+  Float_t hposl,hposr,hmarl,hmarr,hfactor;
+
+  for (Int_t i=0;i<Nx;i++) {
+
+    if (i==0) {
+      hposl = 0.0;
+      hposr = lMargin + hStep;
+      hfactor = hposr-hposl;
+      hmarl = lMargin / hfactor;
+      hmarr = 0.0;
+    } else if (i == Nx-1) {
+      hposl = hposr + hSpacing;
+      hposr = hposl + hStep + rMargin;
+      hfactor = hposr-hposl;
+      hmarl = 0.0;
+      hmarr = rMargin / (hposr-hposl);
+    } else {
+      hposl = hposr + hSpacing;
+      hposr = hposl + hStep;
+      hfactor = hposr-hposl;
+      hmarl = 0.0;
+      hmarr = 0.0;
+    }
+
+    for (Int_t j=0;j<Ny;j++) {
+
+      if (j==0) {
+	vposd = 0.0;
+	vposu = bMargin + vStep;
+	vfactor = vposu-vposd;
+	vmard = bMargin / vfactor;
+	vmaru = 0.0;
+      } else if (j == Ny-1) {
+	vposd = vposu + vSpacing;
+	vposu = vposd + vStep + tMargin;
+	vfactor = vposu-vposd;
+	vmard = 0.0;
+	vmaru = tMargin / (vposu-vposd);
+      } else {
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/CoulombPotential/fitCoulombPotential.C

@@ -1,0 +1,135 @@
+
+
+void fitCoulombPotential()
+{
+ TFile *f = TFile::Open("/home/justin/mythesis/macrosForThesis/PionSpectra/rootfiles/publishPions_sm.root");
+  TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
+  TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
+
+ TFile *g = TFile::Open("/home/justin/mythesis/macrosForThesis/PionSpectra/rootfiles/publishPions.root");
+  TH1D *pip_132 = (TH1D *)g->Get("pip_sn132");
+  TH1D *pim_132 = (TH1D *)g->Get("pim_sn132");
+  TH1D *pip_108 = (TH1D *)g->Get("pip_sn108");
+  TH1D *pim_108 = (TH1D *)g->Get("pim_sn108");
+
+  double s132_n = 132 + 124 - 100;
+  double s132_p = 100;
+
+  double s108_n = 108 + 112 - 100;
+  double s108_p = 100;
+
+  /*
+  double r132 = (5*pow(s132_n,2) + s132_n*s132_p)/(5*pow(s132_p,2) + s132_n*s132_p);
+  double r108 = (5*pow(s108_n,2) + s108_n*s108_p)/(5*pow(s108_p,2) + s108_n*s108_p);
+  */
+
+  //r132 from N/Z  2.09
+  //r108 from N/Z  1.36
+
+  //Data driven
+  double r132 = 4.83;
+  double r108 = 2.04;
+
+  cout<<"Sn132 ratio "<<r132<<endl;
+  cout<<"Sn108 ratio "<<r108<<endl;
+
+  double coul_factor = 6 * 100 * 1.44/5; //fm * MeV
+  double pi_mass = 139.57;
+  double sn132_temp = 0;
+  double sn108_temp = 0;
+  
+  /*
+  TF1 *blz = new TF1("blz","[0]*(x + [3] - [1])*TMath::Exp(-(x + [3])/[2])",0,300);
+  blz->SetParameters(100,10,30);
+  blz->FixParameter(3,pi_mass);
+  */
+  TF1 *blz = new TF1("blz","[0]*TMath::Exp(-(x + [1])/[2])",0,300);
+  blz->SetParameter(0,1000);
+  blz->SetParameter(2,20);
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/PlotDataTheory/calculateMultiplicity_Sn132.C

@@ -1,0 +1,57 @@
+
+
+/*
+TH1D * calculateUpError(TH1D *h)
+{
+  TH1D *copy = (TH1D *)h->Clone();
+
+  for(int i = 1; i <= h->GetNbinsX(); i++)
+    copy->SetBinContent(i,h->GetBinErrorUp(i) + copy->GetBinContent(i));
+
+  return copy;
+}
+*/
+
+double UpError(TH1D *h)
+{
+  TH1D *copy = (TH1D *)h->Clone();
+
+  for(int i = 1; i <= h->GetNbinsX(); i++)
+    copy->SetBinContent(i,h->GetBinErrorUp(i) + copy->GetBinContent(i));
+
+
+  return (copy->Integral("width") - h->Integral("width"));
+}
+
+double LowError(TH1D *h)
+{
+  TH1D *copy = (TH1D *)h->Clone();
+
+  for(int i = 1; i <= h->GetNbinsX(); i++)
+    copy->SetBinContent(i,copy->GetBinContent(i) - h->GetBinErrorUp(i));
+
+  return ( h->Integral("width") - copy->Integral("width") );
+}
+
+
+void calculateMultiplicity_Sn132()
+{
+  TFile *f = TFile::Open("../rootfiles/publishPions.root");
+  TH1D * pim_sn132 = (TH1D *)f->Get("pim_sn132");
+  TH1D * pip_sn132 = (TH1D *)f->Get("pip_sn132");
+
+  TH1D * pim_sn108 = (TH1D *)f->Get("pim_sn108");
+  TH1D * pip_sn108 = (TH1D *)f->Get("pip_sn108");
+
+
+  cout<<"Sn132 "<<endl;
+  cout << "pi- "<< ","<< pim_sn132->Integral("width") << ","<< UpError(pim_sn132) << "," << LowError(pim_sn132)<<endl;
+  cout << "pi+ "<< ","<< pip_sn132->Integral("width") << ","<< UpError(pip_sn132) << "," << LowError(pip_sn132)<<endl;
+
+  cout<<"Sn108 "<<endl;
+  cout << "pi- "<< ","<< pim_sn108->Integral("width") << ","<< UpError(pim_sn108) << "," << LowError(pim_sn108)<<endl;
+  cout << "pi+ "<< ","<< pip_sn108->Integral("width") << ","<< UpError(pip_sn108) << "," << LowError(pip_sn108)<<endl;
+
+
+
+}

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/PlotDataTheory/publishPionsSpectra108Sn.C

@@ -1,0 +1,382 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+
+
+void publishPionsSpectra108Sn()
+{
+
+ TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft.root");
+  TGraphErrors *pBUU_soft_no_pim = (TGraphErrors *) pBUU_soft_no_f->Get("pim108");
+  TGraphErrors *pBUU_soft_no_pip = (TGraphErrors *) pBUU_soft_no_f->Get("pip108");
+  
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff.root");
+  TGraphErrors *pBUU_stiff_no_pim = (TGraphErrors *) pBUU_stiff_no_f->Get("pim108");
+  TGraphErrors *pBUU_stiff_no_pip = (TGraphErrors *) pBUU_stiff_no_f->Get("pip108");
+  
+  TFile *amdJAM55_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML55new.root");
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/PlotDataTheory/publishPionsSpectra132Sn.C

@@ -1,0 +1,408 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+
+
+void publishPionsSpectra132Sn()
+{
+
+ TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft.root");
+  TGraphErrors *pBUU_soft_no_pim = (TGraphErrors *) pBUU_soft_no_f->Get("pim132");
+  TGraphErrors *pBUU_soft_no_pip = (TGraphErrors *) pBUU_soft_no_f->Get("pip132");
+  
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff.root");
+  TGraphErrors *pBUU_stiff_no_pim = (TGraphErrors *) pBUU_stiff_no_f->Get("pim132");
+  TGraphErrors *pBUU_stiff_no_pip = (TGraphErrors *) pBUU_stiff_no_f->Get("pip132");
+  
+  TFile *amdJAM55_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML55new.root");
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/SpaceChargeVariation/plotLeftRight.C

@@ -1,0 +1,223 @@
+
+
+void plotLeftRight()
+{
+  gStyle->SetOptStat(0);
+  
+  TFile *sn108_sc = TFile::Open("analysis-Sn108-nominal_LR.root");
+  //TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
+  TFile *sn132_sc = TFile::Open("analysis-Sn132-sn132_1_2_LR.root");
+  
+  TFile *sn108_nosc = TFile::Open("analysis-Sn108-noSC.root");
+  TFile *sn132_nosc = TFile::Open("analysis-Sn132-noSC.root");
+
+  TH1D *pim_108_L = (TH1D *)sn108_sc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R = (TH1D *)sn108_sc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L = (TH1D *)sn108_sc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R = (TH1D *)sn108_sc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L = (TH1D *)sn132_sc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R = (TH1D *)sn132_sc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L = (TH1D *)sn132_sc->Get("pipkecm_ec_L");
+  TH1D *pip_132_R = (TH1D *)sn132_sc->Get("pipkecm_ec_R");
+
+  TH1D *pim_108_L_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_L");
+  TH1D *pip_132_R_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_R");
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/TheoryBootstrap/Weight/MomDep/theoryBootStrapBCA.C

@@ -1,0 +1,165 @@
+TRandom3 *ran = new TRandom3(12345);
+
+void loadVec(vector< vector<double>> &d, TGraphErrors *g)
+{
+  int num_bins = g->GetN();
+   for(int iBin = 0; iBin < num_bins; iBin++)
+    {
+      double x,y;
+      g->GetPoint(iBin,x,y);
+      d.at(iBin).push_back(y);
+    }
+}
+
+
+vector<double> SampleBoot(vector< vector<double>> data)
+{
+  vector<double> boot(data.size(),0);
+
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      int num_th = data.at(iBin).size(); //# of theories
+
+      for(int i = 0; i < num_th; i++)
+	boot.at(iBin) += data.at(iBin).at(ran->Integer(num_th));
+
+      boot.at(iBin) /= num_th;
+    }
+  
+  return boot;
+}
+
+
+vector<TH1D *> BootStrap(int n_throws, vector< vector<double>> data)
+{
+  vector<TH1D *> hist_vec;
+  for(int iBin = 0; iBin < data.size(); iBin++)
+    {
+      TH1D *h = new TH1D(Form("bin_%d",iBin),Form("bin_%d",iBin),75,0,4);
+      hist_vec.push_back(h);
+    }
+
+  for(int iBoot = 0; iBoot < n_throws; iBoot++)
+    {
+      auto vec = SampleBoot(data);
+      for(int iBin = 0; iBin < vec.size(); iBin++)
+	hist_vec.at(iBin)->Fill(vec.at(iBin));
+    }
+
+  return hist_vec;
+}
+
+
+void theoryBootStrapBCA()
+{
+
+    TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
+  TGraphErrors *pBUU_soft_no_DR = (TGraphErrors *) pBUU_soft_no_f->Get("doubleRatio");
+
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff_r-1.root");
+  TGraphErrors *pBUU_stiff_no_DR = (TGraphErrors *) pBUU_stiff_no_f->Get("doubleRatio");
+
[…]

### 2020-06-10 macrosForThesis/PionMultiplicity/PionData/publishPionRatios.C

@@ -1,0 +1,102 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void publishPionRatios()
+{
+
+  TFile *f = TFile::Open("publishPions_sm.root");
+  TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
+  TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
+
+  TH1D *doubleRatio = (TH1D *)f->Get("doubleRatio");
+
+  make(singleRatio_sn108);
+  make(singleRatio_sn132);
+  make(doubleRatio);
+
+  gStyle->SetEndErrorSize(0);
+
+  //style
+  int lineS132 = 1; 
+  int lineS108 = 7;
+
+  int markS132 = 21;
+  int markS108 = 25;  
+
+  //color
+  int lineC132 = kRed ; 
+  int lineC108 = kRed - 7;
+
+  int markC132 = kRed ;
+  int markC108 = kRed - 7;  
+
+  //size
+  int markSz132 = 2; //marker size
+  int markSz108 = 2; //marker size 
+
+  //  TLegend *leg = new TLegend(.2,.2,.55,.5);
+  TLegend *leg = new TLegend(.3,.5,.8,.8);
+  leg->AddEntry(singleRatio_sn132,"{}^{132}Sn +{}^{124}Sn","lpe");
+  leg->AddEntry(singleRatio_sn108,"{}^{108}Sn +{}^{112}Sn","lpe");
+
+  /*  TLegend *leg_2 = new TLegend(.75,.75,.95,.9);
+  leg_2->SetHeader("{}^{108}Sn +{}^{112}Sn");
+  leg_2->AddEntry(pim_sn108,"#pi^{-}","lpe");
+  leg_2->AddEntry(pip_sn108,"#pi^{+} (x10^{-1} )","lpe");
+  */
+
+  TCanvas *cvs = style::stdcvs();
+  //  cvs->SetLogy();
+  //  singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
+  
+  singleRatio_sn132->GetYaxis()->SetRangeUser(0,15);
+  singleRatio_sn132->GetYaxis()->SetTitle("#pi^{-}/#pi^{+}");
+  singleRatio_sn132->GetYaxis()->CenterTitle();
+  singleRatio_sn132->GetXaxis()->SetTitle("T_{#scale[.6]{COM}} (MeV)");
[…]

### 2020-06-10 macrosForThesis/PionSpectra/CutVariation/publishCutVariations_sn132.C

@@ -90,7 +90,7 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132(int this_var = 1,TString filename = "blank")
+void publishCutVariations_sn132(int this_var = 3,TString filename = "blank")
 {
   //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 

### 2020-06-10 macrosForThesis/PionSpectra/SpaceChargeVariation/plotLeftRight.C

@@ -1,12 +1,16 @@
+#include "/home/justin/mythesis/thesis_style.h"
 
+using namespace style;
 
 void plotLeftRight()
 {
   gStyle->SetOptStat(0);
+  gStyle->SetFrameLineWidth(5);
+
   
   TFile *sn108_sc = TFile::Open("analysis-Sn108-nominal_LR.root");
-  //TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
-  TFile *sn132_sc = TFile::Open("analysis-Sn132-sn132_1_2_LR.root");
+  TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
+  //  TFile *sn132_sc = TFile::Open("analysis-Sn132-sn132_1_2_LR.root");
   
   TFile *sn108_nosc = TFile::Open("analysis-Sn108-noSC.root");
   TFile *sn132_nosc = TFile::Open("analysis-Sn132-noSC.root");
@@ -76,6 +80,59 @@ void plotLeftRight()
   */
 
 
+  pim_132_L->GetYaxis()->SetTitle("Yield");
+  pim_132_L->GetXaxis()->SetTitle("T_{CM}");
+  make(pim_132_L);
+  make(pim_132_R);
+
+  pip_132_L->GetYaxis()->SetTitle("Yield");
+  pip_132_L->GetXaxis()->SetTitle("T_{CM}");
+  make(pip_132_L);
+  make(pip_132_R);
+
+  pim_108_L->GetYaxis()->SetTitle("Yield");
+  pim_108_L->GetXaxis()->SetTitle("T_{CM}");
+  make(pim_108_L);
+  make(pim_108_R);
+
+  pip_108_L->GetYaxis()->SetTitle("Yield");
+  pip_108_L->GetXaxis()->SetTitle("T_{CM}");
+  make(pip_108_L);
+  make(pip_108_R);
+
+  pim_132_L_no->GetYaxis()->SetTitle("Yield");
+  pim_132_L_no->GetXaxis()->SetTitle("T_{CM}");
+  make(pim_132_L_no);
+  make(pim_132_R_no);
+
+  pip_132_L_no->GetYaxis()->SetTitle("Yield");
+  pip_132_L_no->GetXaxis()->SetTitle("T_{CM}");
[…]

### 2020-06-10 macrosForThesis/PionSpectra/SpaceChargeVariation/plotLeftRightRatio.C

@@ -1,0 +1,309 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void plotLeftRightRatio()
+{
+  gStyle->SetOptStat(0);
+  
+  TFile *sn108_sc = TFile::Open("analysis-Sn108-nominal_LR.root");
+  TFile *sn132_sc = TFile::Open("analysis-Sn132-nominal_LR.root");
+  //  TFile *sn132_sc = TFile::Open("analysis-Sn132-sn132_1_2_LR.root");
+  
+  TFile *sn108_nosc = TFile::Open("analysis-Sn108-noSC.root");
+  TFile *sn132_nosc = TFile::Open("analysis-Sn132-noSC.root");
+
+  TH1D *pim_108_L = (TH1D *)sn108_sc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R = (TH1D *)sn108_sc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L = (TH1D *)sn108_sc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R = (TH1D *)sn108_sc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L = (TH1D *)sn132_sc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R = (TH1D *)sn132_sc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L = (TH1D *)sn132_sc->Get("pipkecm_ec_L");
+  TH1D *pip_132_R = (TH1D *)sn132_sc->Get("pipkecm_ec_R");
+
+  TH1D *pim_108_L_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_108_R_no = (TH1D *)sn108_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_108_L_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_L");
+  TH1D *pip_108_R_no = (TH1D *)sn108_nosc->Get("pipkecm_ec_R");
+  
+  TH1D *pim_132_L_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_L");
+  TH1D *pim_132_R_no = (TH1D *)sn132_nosc->Get("pimkecm_ec_R");
+  TH1D *pip_132_L_no = (TH1D *)sn132_nosc->Get("pipkecm_ec_L");
[…]

### 2020-06-10 macrosForThesis/PionSpectra/publishPionRatios.C

@@ -5,7 +5,7 @@ using namespace style;
 void publishPionRatios()
 {
 
-  TFile *f = TFile::Open("publishPions_sm.root");
+  TFile *f = TFile::Open("./rootfiles/publishPions_sm.root");
   TH1D *singleRatio_sn132 = (TH1D *)f->Get("singleRatio_sn132");
   TH1D *singleRatio_sn108 = (TH1D *)f->Get("singleRatio_sn108");
 
@@ -46,7 +46,23 @@ void publishPionRatios()
   leg_2->AddEntry(pip_sn108,"#pi^{+} (x10^{-1} )","lpe");
   */
 
+  doubleRatio->GetYaxis()->SetTitle("Double Ratio");
+  doubleRatio->GetYaxis()->CenterTitle();
+  doubleRatio->GetXaxis()->SetTitle("T_{#scale[.6]{COM}} (MeV)");
+  doubleRatio->GetXaxis()->CenterTitle();
+
+  doubleRatio->GetYaxis()->SetRangeUser(1.8,3.5);
+  doubleRatio->SetLineWidth(5);
+  doubleRatio->SetLineStyle(1);
+  doubleRatio->SetLineColor(4);
+  doubleRatio->SetMarkerStyle(21);
+  doubleRatio->SetMarkerSize(3);
+  doubleRatio->SetMarkerColor(4);
+
+
   TCanvas *cvs = style::stdcvs();
+
+
   //  cvs->SetLogy();
   //  singleRatio_sn132->GetYaxis()->SetRangeUser(.5,20);
   
@@ -72,25 +88,36 @@ void publishPionRatios()
   singleRatio_sn108->SetMarkerSize(markSz108);
   singleRatio_sn108->SetMarkerColor(markC108);
   singleRatio_sn108->Draw("same E1");
+  cvs->Update();
+
+
+
+  double rightmax = 1.1*singleRatio_sn132->GetMaximum();
+  cout<<rightmax<<endl;
+  double rightlow    = singleRatio_sn132->GetMinimum();
+  cout<<gPad->GetUymin()<<endl;
+  double scale    = gPad->GetUymax()/(rightmax-rightlow);
+  cout<<scale<<endl;
+  doubleRatio->Scale(scale);
+  doubleRatio->Draw("same E1");
+
   //  singleRatio_sn108->DrawCopy("same hist L");
   leg->SetBorderSize(0);
[…]

### 2020-06-10 thesis_style.h

@@ -54,7 +54,7 @@ void style::init() {
   gStyle->SetStatX(fXStat);
   gStyle->SetStatY(fYStat);
   gStyle->SetTitleFontSize(fMainTitleSize);
-  gStyle->SetLineWidth(10);
+  //  gStyle->SetLineWidth(10);
 
 }
 

### 2020-06-14 chapters/ch_data_analysis.tex
Backup

@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{highlowcal.png}
@@
 
- \begin{table}[!htp] % not just 'h!'
+ \begin{table}[!htb] % not just 'h!'
 \centering % not a center environment
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{spacechg_cartoon.png}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{beampath.png}
@@
 \begin{figure}[!htb]
-\includegraphics[width=\linewidth]{Effect_SC.pdf}
+\centering
+\includegraphics[scale=.5]{Effect_SC.pdf}
 \caption{Example of the map of the electron shift.
@@
 \begin{figure}[!htb]
-\includegraphics[width=\linewidth]{DVTP_raw.pdf}
+\centering
+\includegraphics[scale=.5]{DVTP_raw.pdf}
 \caption{$\Delta\mathrm{V}_\mathrm{x}$ distribution for left-going and right-going tracks. }
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{VLR_beamrate.png}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{sc_beamrate.png}
@@
 
-%\begin{sidewaysfigure}[H]
+%\begin{sidewaysfigure}[!htb]
 %\centering
@@
 %CONSIDER ROTATING THIS FIGURE IF ALLOWED
-\begin{figure}[H]
+\begin{figure}[!htb]
 \centering
@@
 \label{sec:pid}
-%PID lines over all
+%PID lines over all picture
+
 %PID of pion cut
 %background estimation
+
+The energy loss curve in Eq.~\ref{eq:bb} can be described as a 5-parameter general function as,
+
+\begin{equation}
+\frac{dE}{dx} = \frac{p_0}{\beta^{p_3}}(p_1 - \beta^{p_3} + \ln(p_2 + {\beta\gamma}^{-p_4})),
+\label{eq:5par_bb}
+\end{equation}
+
+where the parameters $p_0-p_4$ are free parameters, $\beta= p/\sqrt{p^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum.
+The PID lines vary slightly as a function of the emission angle of each track.
+The PID was subdivided into 6 pitch angles, $\theta_P = \arctan(p_y/p_z)$, and 6 yaw angles, $\theta_Y = \arctan(p_x/p_z)$, ranging from \ang{-90} to \ang{90} for both.
+The pion spectra of each yaw-pitch bin was fit with Eq.~\ref{eq:5par_bb} to get the PID line which best describes the data.
+The distribution around this mean value is not a Gaussian distribution.
+The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Guassian variable, defined as
+
+\begin{equation}
+z_i = \ln\Big(\frac{\langle dE/dx\rangle}{\langle dE/dx\rangle_i}\Big),
+\end{equation}
+
+where $\langle dE/dx\rangle_i$ is the mean energy loss curve fit from Eq.~\ref{eq:5par_bb} for a given particle type $i$.
+The $z_i$ distribution of the particle of interest will be centered around 0 in this new variable.
+Figure~\ref{fig:pidpion_raw} shows the typical PID lines zoomed in on the pions.
+Figure~\ref{fig:pidpion} shows the corresponding $z_i$ distribution after all the pitch-yaw bins have been merged.
+Both charged pions lie on $z_i=0$, where other particle species rapidly diverge.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pidPion.png}
+        \caption{Output PID pions.} \label{fig:pidpion_raw}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pidPion_zi.png} 
+        \caption{PID plot for $\pi^-$ and $\pi^+$ species in $z_i$. } \label{fig:pipion_zi}
+    \end{subfigure}
+
+    \caption{$\langle dE/dx\rangle$ versus p/q for $\tin{132}{124}$ system.}
+\label{fig:pidpion}
+\end{figure}
+
+Now that the pion distribution is flattened, we take bin slices along $p/q$ to determine the particle yield in a certain bin and also estimate the background contribution.
+The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}.
+Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Fig.~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.
+A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_mom2.png}
+        \caption{Bin -54 MeV/c < p/q < -52. } \label{fig:pimmom2}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pim_mom1.png} 
+        \caption{Bin -68 MeV/c < p/q < -66.} \label{fig:pimmom1}
+    \end{subfigure}
+
+    \caption{$z_i$ projection for binned slices in p/q around the $\pi^-$.}
+\label{fig:pimmom_flat}
+\end{figure}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[t]{0.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pip_mom1.png}
+        \caption{Bin 74 MeV/c < p/q < 76. } \label{fig:pipmom1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{pip_mom2.png} 
+        \caption{Bin 94 MeV/c < p/q < 86.} \label{fig:pipmom2}
+    \end{subfigure}
+
+    \caption{$z_i$ projection for binned slices in p/q around the $\pi^+$.}
+\label{fig:pipmom_flat}
+\end{figure}

### 2020-06-14 chapters/ch_data_analysis_2.tex
Backup

@@
 
-\begin{table}[H]
+\begin{table}[!htb]
   \begin{center}
@@
 \section{Beam angle selection}
+\label{sec:beamangle}
 The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle.

### 2020-06-14 chapters/ch_results.tex
Backup

@@
 
-The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database. The correction factor $C_i$ is defined as,
+The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
+The correction factor $C_i$ is defined as,
 
@@
 
-The track is then weighted by the factor $C_i$ when plotting in any type of graph for any observable.
+Each track that is identified as a pion, is then weighted by the correction factor $C_i$, when filling any histogram.
+By using the track-by-track method, we can transform the track into any observable of interest, notably we can transform the  track from the lab frame into the center-of-mass (CM) frame.
+Before performing the transformation to the CM system recall the beam is at a small, but non-negligible, angle in the Lab frame, see Section~\ref{sec:beamangle}.
+The beam angle for each even is measured and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis.
+Doing so makes the transformation into the CM system much simpler to describe.
+If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.
+
+Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation.
+Where the 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
+Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,
+
+\begin{equation}
+A = \begin{pmatrix}
+1 & 0 & 0 & 0\\
+0 & 1 & 0 & 0\\
+0 & 0 & \gamma & -\beta \gamma\\
+0 & 0 & -\beta \gamma & \gamma
+\end{pmatrix},
+\end{equation}
+
+where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$.
+The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 - M_{T}^2}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ as $\beta = -P/E$, where the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
+The CM transformed track is defined as $p^{CM} = \textbf{A}p^{Lab}$.
 
 %need to add error analysis
+%systematic error analysis
 
-%Efficiency correction
-%space charge before and after
-%left and right of space charge 
-%error analysis
-%systematic error analysis
+Figure~\ref{fig:pionspectra} shows the pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
+This data marks the first pion spectral data at sub-threshold energies, extending to very low pion energies.
+The coulomb potential works to accelerate $\pi^+$ and decelerate $\pi^-$ particles due to the positive charge of the nuclear medium.
+The coulomb barrier prevents the production of low energy $\pi^+$, which is why there is a reduction in the phase space around zero kinetic energy.
 
@@
 \section{Pion Yield}
+
+The integrated pion yield for both systems and the $\pi^-/\pi^+$ ratio is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
+It is remarkable that the pion ratio is significantly greater than the N/Z of the system which is what is expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ CITE HERE; 2x in the $\tin{132}{124}$ system and 1.4x in the $\tin{108}{112}$ system.
+The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect and lower the ratio as the system tends toward isospin equilibrium.
+A na\"ive interpretation would be the high density N/Z fraction is much higher for some reason than the two systems, though, the system spends very little time in this state which would require a very soft Symmetry Energy.
+A more likely conclusion is the mechanisms involved in pion productions via the $\Delta(1232)$ resonance is more complicated than the simple $(N/Z)^2$ relation.
+For example the $\Delta$ resonance potential has an iso-scalar component (independent of iso-spin) and an iso-vector component.
+The nature of these two potentials is still unknown CITE HERE, and the role it plays in pion production has shown to be very important CITE HERE.
 
@@
 
-\section{$\pi^-$/$\pi^+$ Ratio}
+We have compared the total pion yields and ratios to the 6 common transport codes for the systems measured.
+The table of the transport codes are listed in Appendix~\ref{tb:pionyieldTheory}.
+Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the codes.
+The codes plotted here are only the soft symmetry energy since the variation in code is much larger than the intra-code variation between different Symmetry Energies.
+While some codes make a reasonable approximation of a particular charge pion, no code reasonably predict both.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=.8\textwidth]{totalpiYield.png}
+\caption{Total pion yields as compared with 6 common transport codes.}
+\label{fig:totalpiYield}
+\end{figure}
+
+Figure~\ref{fig:totalpiRatio} shows the total single pion ratio and the double ratio of the $\tin{132}{124}$ and $\tin{108}{112}$ system.
+Here, the variation between codes is much larger than the variation between the Symmetry Energy within a particular code.
+The Symmetry Energy variation of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and all codes in the double ratio; the data band represents the data error bar.
+Certainly the variation between Symmetry Energy in the $\tin{132}{124}$, though small, still exists in many of the codes as predicted initially CITE HERE BAO ANN.
+The small error bars in the data also would allow for a detailed analysis to extract a constraint of the Symmetry Energy, if the variation between codes could be solved.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{totalpiRatio.png}
+\caption{Total pion ratio and double ratio compared with 6 common transport codes.}
+\label{fig:totalpiRatio}
+\end{figure}
+
+\section{Pion Spectral Ratio}
 

### 2020-06-14 macrosForThesis/PID/plotPILine.C

@@ -1,0 +1,53 @@
+
+auto mpi = 139.57018;
+auto me = 0.5109989461;
+auto mp = 938.2720813;
+auto mn = 939.565346;
+auto md = 1875.612762;
+auto mt = 2808.921112;
+auto mhe3 = 2808.39132;
+auto mal = 3727.379378;
+
+Double_t particleMass = mpi; // Select particle you'll going to analyze                            
+
+Double_t dedxModelFunc(Double_t *x, Double_t *p) {
+  if (x[0] < 0)
+    x[0] = -x[0];
+
+  const Double_t m = particleMass;
+  Double_t totalE = TMath::Sqrt(m*m + x[0]*x[0]);
+  Double_t beta = x[0]/totalE;
+  Double_t gamma = 1./TMath::Sqrt(1 - beta*beta);
+
+  Double_t betaP3 = TMath::Power(beta, p[3]);
+
+  return p[0]/betaP3*(p[1] - betaP3 + TMath::Log(p[2] + TMath::Power(1./(beta*gamma), p[4])));
+}
+
+void plotPILine()
+{
+
+  TFile *f = TFile::Open("histPID-merged_sn132.root");
+
+  auto c1 = (TCanvas *)f->Get("c5");
+
+  auto pip = new TF1("pip", dedxModelFunc, 25., 550., 5, 1);
+  pip -> SetParameters( 0.0305075     , 513.643  ,6.70773e-46    ,  2.34166  ,   -101.122);
+
+  auto pim = new TF1("pim", dedxModelFunc, -650., -25., 5, 1);
+  pim -> SetParameters( 0.0305075     , 513.643  ,6.70773e-46    ,  2.34166  ,   -101.122);
+
+  /*
+  auto hist = (TH2D *)f->Get("histPIDCharged");
+
+  TCanvas *c1 = new TCanvas("c1","c1",1);
+  c1->SetLeftMargin(.15);
+  c1->SetLogz();
+   hist -> Draw("col");
+
+*/
+  c1->Draw();
+  pim->Draw("same");
+  pip->Draw("same");
+
+}

### 2020-06-14 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_NOEff.C

@@ -1,0 +1,301 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+
+void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
+                     Float_t lMargin, Float_t rMargin,
+                     Float_t bMargin, Float_t tMargin)
+{
+  if (!C) return;
+
+  // Setup Pad layout:
+  Float_t vSpacing = 0.0;
+  Float_t vStep  = (1.- bMargin - tMargin - (Ny-1) * vSpacing) / Ny;
+
+  Float_t hSpacing = 0.0;
+  Float_t hStep  = (1.- lMargin - rMargin - (Nx-1) * hSpacing) / Nx;
+
+  Float_t vposd,vposu,vmard,vmaru,vfactor;
+  Float_t hposl,hposr,hmarl,hmarr,hfactor;
+
+  for (Int_t i=0;i<Nx;i++) {
+
+    if (i==0) {
+      hposl = 0.0;
+      hposr = lMargin + hStep;
+      hfactor = hposr-hposl;
+      hmarl = lMargin / hfactor;
+      hmarr = 0.0;
+    } else if (i == Nx-1) {
+      hposl = hposr + hSpacing;
+      hposr = hposl + hStep + rMargin;
+      hfactor = hposr-hposl;
+      hmarl = 0.0;
+      hmarr = rMargin / (hposr-hposl);
+    } else {
+      hposl = hposr + hSpacing;
+      hposr = hposl + hStep;
+      hfactor = hposr-hposl;
+      hmarl = 0.0;
+      hmarr = 0.0;
+    }
+
+    for (Int_t j=0;j<Ny;j++) {
+
+      if (j==0) {
+	vposd = 0.0;
+	vposu = bMargin + vStep;
+	vfactor = vposu-vposd;
+	vmard = bMargin / vfactor;
+	vmaru = 0.0;
+      } else if (j == Ny-1) {
+	vposd = vposu + vSpacing;
+	vposu = vposd + vStep + tMargin;
+	vfactor = vposu-vposd;
+	vmard = 0.0;
+	vmaru = tMargin / (vposu-vposd);
+      } else {
+	vposd = vposu + vSpacing;
+	vposu = vposd + vStep;
+	vfactor = vposu-vposd;
+	vmard = 0.0;
+	vmaru = 0.0;
+      }
+
[…]

### 2020-06-18 MSU-thesis-estee.tex
Backup
Something broken. Saftey backup

@@
 \usepackage{comment}\renewcommand{\arraystretch}{1.5}
+\usepackage[version=3]{mhchem}
 
-\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/} }
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/}  {images/intro/} }
 
@@
 \includeonly{
-%chapters/ch_intro,
-%/ch_theory,
+chapters/ch_intro,
+%ch_theory,
 %chapters/ch_experiment,
-chapters/ch_data_analysis,
+%chapters/ch_data_analysis,
 %chapters/ch_data_analysis_2,
@@
 
-%\include{chapters/ch_intro}
+\include{chapters/ch_intro}
 %\include{chapters/ch_theory}
 %\include{chapters/ch_experiment}
-\include{chapters/ch_data_analysis}
+%\include{chapters/ch_data_analysis}
 %\include{chapters/ch_data_analysis_2}

### 2020-06-18 chapters/ch_appendix.tex
Backup
Something broken. Saftey backup

@@
-\chapter{Runs analyzed in this data}
+\subsection{$\Delta$ Resonance Production and Decay Channels}
+\label{appen:deltadecay}
+
+The Clebsh-Gordon coefficients for nucleon-nucleon leading to certain $\Delta$ resonances and the respective nucleon: 
+\begin{equation}
+\begin{split}
+& p + p \rightarrow \sqrt{\frac{3}{4}}(\Delta^{++} + n) - \sqrt{\frac{1}{4}}(\Delta^{+} + p) \\
+& n + p \rightarrow \sqrt{\frac{1}{2}}(\Delta^{+} + n) - \sqrt{\frac{1}{2}}(\Delta^{0} + p)  \\
+& n + n \rightarrow \sqrt{\frac{1}{4}}(\Delta^{0} + n) - \sqrt{\frac{3}{4}}(\Delta^{-} + p) 
+\end{split}
+\label{eq:deltaProduction}
+\end{equation}
+
+A corresponding Clebsh-Gordon coefficients for the decay of $\Delta$ pions into the constituent $\pi$ and corresponding nucleon:
+
+\begin{equation}
+\begin{split}
+& p + p \rightarrow \sqrt{\frac{5}{6}}(\pi^+ + p + n) - \sqrt{\frac{1}{6}}(\pi^0 + p + p) \\
+& n + p \rightarrow \sqrt{\frac{1}{6}}(\pi^+ + n + n) + \sqrt{\frac{2}{3}}(\pi^0 + n + p) + \sqrt{\frac{1}{6}}(\pi^- + p + p)  \\
+& n + n \rightarrow \sqrt{\frac{1}{6}}(\pi^0 + n + n) - \sqrt{\frac{5}{6}}(\pi^- + n + p) 
+\end{split}
+\label{eq:deltaProduction}
+\end{equation}
+
+Here we can see that that proton-proton collisions are connected with $\pi^+$ and neutron-neutron collisions are connected with $\pi^-$.
+
+\subsection{Runs analyzed in this data}
+
 \newcommand{\hsn}{$^{132}$Sn+$^{124}$Sn}
@@
 \end{table}
+
+\clearpage
+
+\subsection{Pion Yield Theory}
+\label{tb:pionyieldTheory}
+
+\begin{table*}[!htb]
+\centering
+\ra{1.1}
+\begin{tabular}{@{}ccccccccccc@{}}
+\toprule
+        &  & \multicolumn{3}{c}{(a) $^{132}$Sn+$^{124}$Sn} &  & \multicolumn{3}{c}{(b) $^{108}$Sn+$^{112}$Sn} &  \\
+        \cmidrule{3-5} \cmidrule{7-9} 
+        Code name & $\mathrm{L~(MeV)}$ &  $\mathrm{Y}(\pi^-)$ & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  & $\mathrm{Y}(\pi^-)$  & $\mathrm{Y}(\pi^+)$ & $\mathrm{SR}(\pi^-/\pi^+)$ &  $\mathrm{DR}(\pi^-/\pi^+)$ \\
+    \midrule
+        $\chi$BUU& 45.6 & 0.509 & 0.109 & 4.67 & & 0.269 & 0.134 & 2.01 & 2.33 \\
+        & 120 & 0.483 & 0.117 & 4.13 &  & 0.271 & 0.140 & 1.94 & 2.13 \\
+        \addlinespace[0.2cm]
+        TuQMD & 54.6 & 0.779 & 0.145 & 5.37 &  & 0.442 & 0.176 & 2.51 & 2.14 \\
+        & 145 & 0.839 & 0.145 & 5.79  &  & 0.474 & 0.181 & 2.62 & 2.21 \\
+        \addlinespace[0.2cm]
+        pBUU & 56.1 & 0.698 & 0.181 & 3.86 &  & 0.401 & 0.213 & 1.88 & 2.05 \\
+        & 135 & 0.649 & 0.185 & 3.51  &   & 0.392 & 0.214 & 1.83 & 1.92 \\
+        \addlinespace[0.2cm]
+        AMD+JAM& 55 & 0.339 & 0.0978 & 3.47 &  & 0.200 & 0.116 & 1.72 & 2.02 \\
+        & 152 & 0.311 & 0.0986 & 3.15 &  & 0.192 & 0.116 & 1.66 & 1.90 \\
+        \addlinespace[0.2cm]
+        IQMD-BNU & 54.6 & 0.542 & 0.148 & 3.67 &  & 0.319 & 0.175 & 1.82 & 2.01 \\
+        & 145 & 0.452 & 0.153 & 2.95 &  & 0.278 & 0.167 & 1.67 & 1.77 \\
+        \addlinespace[0.2cm]
+        SMASH  & 55 & 0.468 & 0.168 & 2.79  &  & 0.287 & 0.190 & 1.51 & 1.85 \\
+        & 152 & 0.479 & 0.163 & 2.93 &  & 0.292 & 0.188 & 1.55 & 1.89 \\
+        \addlinespace[0.2cm]
+        UrQMD & 46 & 0.479 & 0.129 & 3.71 &  & 0.292 & 0.144 & 2.03 & 1.83 \\
+        & 104 & 0.449 & 0.133 & 3.38 &  & 0.274 & 0.147 & 1.86 & 1.81 \\
+        \addlinespace[0.2cm]
+        \bottomrule
+    \end{tabular}
+    \caption{Pion multiplicities, $Y(\pi^{\pm})$, single ratios $SR(\pi^-/\pi^+)$, and double multiplicity ratios, $DR(\pi^-/\pi^+)$ from seven transport codes.
+Each code uses two different symmetry energy functions, with all other parameters identical in the codes.}
+    \label{tab:pionyieldTheory}
+\end{table*}
+
+\clearpage
+
+\section{Dalitz Decay of the $\pi^0$}
+\label{appen:dalitz}
+
 \chapter{Second appendix}

### 2020-06-18 chapters/ch_intro.tex
Something broken. Saftey backup

@@
 \chapter{Introduction}
+
+Rutherford discovered in 1911 that the atomic nuclei are composed of a dense nucleus surrounded by an electron cloud, and eventually that the nucleus if composed of positive and neutral spin 1/2 particles called protons and neutrons.
+The proton and neutron are actually composed of 3 fundamental particles called quarks.
+Besides for the charge differences between protons and neutrons, they exhibit a fundamental symmetry which explains their almost identical mass.
+Typically neutrons and protons are treated as two different isospin projections of a ``nucleon", (-1/2 and 1/2) respectively.
+
+The nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it incredibly dense.
+Without a balancing force the Coulomb force between protons would render the nucleus unstable.
+The nuclear strong interaction is the fundamental force governing the interactions between the constituent quarks in the nucleons, deriving its name from the large force which acts only over small distances.
+The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances.
+It is for this reason that the distance between nuclei is a near constant value, and therefore the density is remarkably constant over a wide range of nuclei CITE HERE.
+This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$.
+
+ In this way nuclei can be thought of as an in-compressible liquid.
+This picture was remarkably successful at describing the binding energies of nuclei at saturation density.
+The Bethe-Weizsacker semi-empirical formula predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:
+
+\begin{equation}
+\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + \mathcal{O}
+\label{eq:semiEmp}
+\end{equation}
+
+Since the number of nucleons is related to the volume of the nucleus, the volume term ,$a_V$, originates from the the saturation of the strong force and its short range nature.
+There are correction terms accounting for the surface, $a_s$, since nucleons near the surface do not have as many neighbors as nucleons inside, and the coulomb term which is related to the size of the nucleus which scales as $A^{1/3}$.
+The asymmetry term, $a_A$, is related to the cost in energy one pays to become more neutron or proton rich; it is typically referred to as the \emph{Symmetry Energy}.
+This originates from pauli blocking where it is more energetically favorable to form neutron-proton pairs since their isospin numbers are different.
+The di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, where as the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more energetically favorable.
+
+Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter CITE HERE, which is normally not stable.
+The large extent of the star allows for the gravitational force to balance the strong force creating a compact dense star.
+The pressures vary in the neutron star from low densities near the crust to the dense interior which can reach up to 9$\rho_o$ CITE HERE.
+To understand these exotic systems, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp} to describe matter over a wide range of densities.
+
+Guided by Eq.~\ref{eq:semiEmp}, we can separate the energy density $E$ of a system into two components,
+
+\begin{equation}
+E(\rho,\delta) = E(\rho	) + S(\rho)\delta^2,
+\label{eq:energyEos}
+\end{equation}
+
+where $E(\rho)$ describes the symmetric term (i.e.
+independent of isospin), and the symmetry energy $S(\rho)$ which depends on the asymmetry of the system, written now in terms of the neutron and proton densities,
+
+\begin{equation}
+\delta = \frac{\rho_n - \rho_p}{\rho}.
+\label{eq:asym}
+\end{equation}
+
+The Equation of State (EoS) of nuclear matter can be calculated by,
+
+\begin{equation}
+P = \Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N} = -\rho^2 \frac{\delta E}{\delta \rho}\Big)\vert_{T=0,N}, 
+\label{eq:pressEos}
+\end{equation}
+
+for a fixed number of particles $N$ and zero temperature.
+One can extend to higher temperatures by adding the Boltzman dependence CITE HERE.
+The partial derivative with respects to volume can be rewritten in terms of density:
+
+\begin{equation}
+P = -\rho^2 \frac{\delta E}{\delta \rho}\vert_{T=0,N}.
+\label{eq:densEos}
+\end{equation}
+
+To understand macroscopic pressures in the neutron star, which balance the gravitational force, we must understand the density dependence of the symmetry energy.
+
+\section{Density Dependence of the Symmetry Energy}
+In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ CITE HERE.
+In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$.
+Figure~\ref{fig:symDen} shows some of the experimental constraints which have been performed by a series of independent measurements and observables CITE HERE ALL.
+Typically an effective interaction is used to describe the phenomenological observations of nucleon-nucleon interactions observed in nuclei.
+One of such interactions is described by the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations CITE HERE.
+Several Skyrme parameterizations are shown as lines in Fig.~\ref{fig:symDen}.
+Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars.
+
+\section{Heavy Ion Collisions}
+Besides observing neutron stars directly, heavy-ion collisions (HIC) provide the only way we can probe the density dependence of the symmetry energy in the laboratory setting.
+When two nuclei collide in a collision, in the very early stages they compress to form a high density region where the nuclei overlap.
+This momentary density can reach up to $3\rho_0$ depending on the incident beam energy.
+HICs provide the only way we can probe the isospin asymmetry dependence of the nuclear EoS.
+This is accomplished by using radioactive neutron-rich beams to collide on stable targets.
+
+The pressure arising from the symmetry energy depends on the curvature of the symmetry energy at a given density.
+If the density dependence of the symmetry energy is positive at high densities the symmetry energy would work to force neutrons out of the system.
+Where as if the derivative was negative the symmetry energy would attract neutrons.
+It is this pressure that is driving the dynamics of neutrons and protons.
+By measuring protons and neutrons we can see signatures of the effects of the symmetry energy in the final spectra of these particles.
+Measuring neutrons experimentally can be quite challenging and space is limited by large neutron wall arrays CITE HERE.
+Also, though the overlap region temporarily reaches a high density, the neutrons which participated in this region also evolve through regions of lower densities until they reach their final state, diluting the signal from the dense region.
+
+\begin{figure}
+\centering
+\includegraphics[scale=.01]{place}
+\caption{Experimental constraints of the density dependence of the symmetry energy}
+\label{fig:symDen}
+\end{figure}
+
+\section{Pion Observable}
+It is preferable to find an observable that is easier to measure experimentally than neutrons, and is more sensitive to the high density region.
+Pions are produced though and intermediate process where nucleon-nucleon collisions form an excited $\Delta$ baryon resonance from one of the nucleons, which then decay decay shortly after into a pion:
+
+\begin{equation}
+\ce{ NN <=> \Delta N <=> \pi NN}.
+\end{equation}
+
+It has been shown in \cite{mingzhang} that most of the $\Delta$'s are produced in the early dense regions of the collision.
+Figure~\ref{fig:deltaProduction} shows the average local density (c) which $\Delta$'s are produced and the number in the system (b), as a function of time in the simulation of Au + Au collisions at \SI{400}{\MeVA}.
+Panel (a) shows the density distribution of the density at the moment of creation for $\Delta$'s.
+Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon.
+Thus the outgoing $\pi$ is contains information on the high density region of the collision.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.2]{deltaProduction.png}
+\caption{Figure taken from \cite{mingzhang} for Au + Au collisions at \SI{400}{\MeVA}.
+Panel (a) shows the density in the region of the $\Delta$ resonance creation for two different symmetry energies (x=0 soft) and (x=1 stiff).
+Panel (b) and (c) show the evolution of collision in time steps, where (b) shows the number of deltas in the system as a function of time and (c) shows the mean local baryon density in the region where $\Delta$ resonances are produced.
+The blue line in (c) represents the average baryon density in the most central region of the collision.
+This evidence shows that a majority of $\Delta$'s are produced in the high density region of the early collision.}
+\label{fig:deltaProduction}
+\end{figure}
+
+The branching ratio of the various flavors of $\Delta$'s is given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}.
+Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$.
+In this $\Delta$ resonance model the charged pion ratio can be described as,
+
+\begin{equation}
+\frac{\pi^-}{\pi^+} = \frac{ 5pp + pn }{5nn + pn}.
+\label{eq:deltaModel}
+\end{equation}
+
+In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
+Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the froward and backwards process $\ce{\Delta <=> \pi N}$.
+This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission would change the pion dynamics or even the charge of the pion which would reflect the asymmetry at the point of creation and eventual re-emission.
+Total pion absorption back into two nucleons requires a three body process where a pion is absorbed creating a $\Delta$ resonance, then another nucleon must collide with the resonance to create two nucleons.
+Because of this, the total pion absorption (removing pions totally) is a less frequent effect than the absorption re-emission process.
+Yet in general by measuring the $\pi^-$ and $\pi^+$ is connected with n-n and p-p collisions.
+Instead of measuring neutrons we can measure the $\pi^-$ which is much easier to measure experimentally.
+
+%nuclear chart
+%Dense nuclear matter 
+%Liquid drop model explain density around saturation
+%explain symmetry energy 
+%extend to systems of larger pressure, isospin assymetry 
+%symmetry energy how to calculate 
+%Previous constraints on symmetry energy
+%heavy ion collisions only way to probe density and 
+%dynamics of proton neutrons
+
+%pion production
 

### 2020-06-18 chapters/ch_results.tex
Something broken. Saftey backup

@@
 \section{Pion Spectra}
-
+\label{sec:pionSpectra}
 The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
@@
 
-\begin{figure}
+\begin{figure}[!htb]
 \centering
@@
 \end{figure}
+
+The pion spectral ratio is a promising observable in particular it may be more sensitive to the high density regions of the early collision.
+In theory the high energy pions are more likely to exit the nuclear medium earlier, and therefore be less prone to effects such as pion absorption and re-emission which dilute the sensitivity of the pion observable to the high density behavior.
+Also low energy pions are more likely to be affected by other effects such as the $\Delta$ potential in medium CITE HERE BAO AN.
+If we integrate the total pion yield we combine these two regions of low and high energy pions.
+Instead we can construct the $Y(\pi^-)/Y(\pi^+)$ ratio as a function of the kinetic energy in the CM system.
+
+%Mention and cite appendix for systematics
+
+Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of accuracy.
+The general hyperbolic shape comes from the Coulomb force on the two charged pion spectra mentioned in Section~\ref{sec:pionSpectra}.
+Naturally the pion ratio is smaller for less neutron poor system, i.e.
+less neutron-neutron collisions produce less $\pi^-$.
+The bin size of the last bin was increased to reduce the statistical error bars since the number of pions, especially the $\pi^+$, is reaching the limits of the measured distribution.
 
@@
 
+Another promising observable is the spectral double ratio.
+In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory, to cancel out.
+For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
+
 \section{Comparison to Previous Data Sets (FOPI)}
-asdfasdfasd
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.4]{fopi_pionratio_comp}
+\caption{Comparing the total $\pi^-/\pi^+$ ratio of the $\tin{132}{124}$ system to the ${}^{197}$Au + ${}^{197}$Au data from the FOPI collaboration.
+The \spirit TPC data was scaled by a factor to compare to the lower N/Z of the Au + Au system.
+This was extracted from measuring the N/Z dependence measured in the experiment.
+}
+\label{fig:fopiPionRatio}
+\end{figure}
+
+The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher beam energies.
+In the $\tin{132}{124}$ data  $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$, $N/Z=1.493$.
+The pion ratio is proportional to the $N/Z$ of the system.
+Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}.
+Here the dependence is fitted with a 2-nd order polynomial fit.
+The pion ratio in the $\tin{132}{124}$ data  was scaled by the ratio between $N/Z=1.56$ and 1.493.
+Figure~\ref{fig:fopiPionRatio} shows the scaled pion ratio as compared with the FOPI pion ratio from literature CITE HERE.
+The fitted function has the functional form of $p_1(E - p_2)^{-2}$ where $p_1$ and $p_2$ are free parameters, and is meant to guide the eye.
+It is worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model CITE HERE.
+Scaling the other systems in the \spirit TPC data leads to almost the exact same value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$.
+
 \section{Comparison to Theory}
@@
 %Reference paper for multiple theories
-asdfasdf
+
+Here we will compare with 7 different commonly used transport theoretical codes.
+These codes took part in a large collaboration effort to compare the codes to each other.
+The goal of these comparison projects was to standardize the input into each code to systematically go through the details in each code comparing things such as initialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc CITE HERE.
+These codes were taken in their best state, without any prior knowledge of the experimental data, and simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
+Each code differs in its treatment of pion and $\Delta$ dynamics.
+Some codes contain modifications to the pion in nuclear matter, usually by including a pion optical potential.
+Some codes include the iso-scalar and the iso-vector delta potential which are not very well constrained but are important in the production of pions CITE HERE.
 
@@
 \label{sec:cutvar}
-dfsadfsaf
-\section{Constraint on the Symmetry Energy}
-%maybe not...
+The particular track quality cuts described in Section~\ref{sec:qualitycut} are there to reduced the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra.
+The best set of values were found for all the data sets which include the charged particle multiplicity of the event, dOCA, and the number of clusters cut.
+By varying the cuts in both the data and the MC efficiency, we can evaluate the analysis to see if there are any systematic dependencies on the observable we plot.
+But varying the cuts undoutably means as the cut get tighter, and less data is included, the statistical error bar will increase making it difficult to compare to more loose cuts.
+This is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}.
+Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts and the uncorrelated error bars, then there is some systematic trend that either exists due to physics or some error or miscalculation in the analysis.
+It is usually not recommended to estimated systematic error bars using this method, but sometimes is the only way.
+
+Here we will discuss the particular analysis using the total pion ratio as the example.
+Figure~\ref{fig:totalRatioError} shows the total pion as a function of several cut variations, where one cut is varied and all the others are held constant.
+Three cuts are of particular interest, the event multiplicity, dOCA, and the number of cluster cut.
+Figure~\ref{fig:clustTotalRatio} shows the variation as the number of cluster cut is varied.
+The default cut used is > 20 clusters which is represented by the middle point.
+The red bars represent the statistical error of this default value.
+The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}.
+Here it is clear that there is a systematic bias in the pion ratio as a function
+
+%show the pim total yield 
+%discuss others in the apendix
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
+         \caption{No efficiency correction.}
+         \label{fig:multCutVar}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:multCutVar}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
+\label{fig:cutVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_NoEC_pimTotal}
+         \caption{Before efficiency correction.}
+         \label{fig:pim_clustVar_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_pimTotal}
+         \caption{After efficiency correction.}
+         \label{fig:pim_clustVar_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the number of cluster cut of the tracks.}
+\label{fig:pim_clustVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_NoEC_pimTotal}
+         \caption{No efficiency correction.}
+         \label{fig:pim_cutVarPOCA_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_pimTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:pim_cutVarPOCA_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the dOCA cut.}
+\label{fig:pim_cutVarPOCA}
+\end{figure}

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/oldScripts/publishCutVariations_sn132.C

@@ -87,13 +87,13 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank")
+void publishCutVariations_sn132(int this_var = 3,TString filename = "blank")
 {
   //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};    
+    
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -106,36 +106,32 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields_NOEff.root");
+  TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
-  TArrow *arrowR = new TArrow(5.3,.03,5.7,.03,.02,"|>");
+  TArrow *arrowL = new TArrow(.22,.04,.28,.04,.02,"<|");
+  TArrow *arrowR = new TArrow(.78,.04,.84,.04,.02,"|>");
 
-  arrowL->SetNDC();
-  arrowR->SetNDC();
-  
   gStyle->SetOptStat(0);
 
-    TCanvas *C = new TCanvas("c1","",1000,300);
-
-  // Margins
-  Float_t lMargin = 0.1;
-  Float_t rMargin = 0.05;
-  Float_t bMargin = 0.3;
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/oldScripts/publishCutVariations_sn132_NOEff.C

@@ -10,7 +10,7 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
   if (!C) return;
 
   // Setup Pad layout:
-  Float_t vSpacing = 0.0;
+  Float_t vSpacing = 0.05;
   Float_t vStep  = (1.- bMargin - tMargin - (Ny-1) * vSpacing) / Ny;
 
   Float_t hSpacing = 0.0;
@@ -117,12 +117,12 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
 
   TCanvas *C = (TCanvas*) gROOT->FindObject("C");
   if (C) delete C;
-  C = new TCanvas("C","canvas",1024,640);
+  C = new TCanvas("C","canvas",1200,1000);
   C->SetFillStyle(4000);
 
   // Number of PADS
   const Int_t Nx = 1;
-  const Int_t Ny = 4;
+  const Int_t Ny = 3;
 
   // Margins
   Float_t lMargin = 0.12;
@@ -166,7 +166,6 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
 	{
 	  for (Int_t iBin = 0 ;iBin < Ny; iBin++)
 	    {
-	      iVar = iBin;
 	      C->cd(0);
 	      // Get the pads previously created.
 	      char pname[16];
@@ -203,11 +202,16 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
 	      box[i ][iBin] = new TBox(-.5,y-ye,6.5,y+ye);
 	      line[i ][iBin] = new TLine(-.5,y,6.5,y);
 	      singleRatio[i ][iBin]->GetYaxis()->SetRangeUser(y-ye*frac,y+ye*frac);
+	      if(iBin == 2)
+		singleRatio[i ][iBin]->GetYaxis()->SetRangeUser(y-ye*frac*8,y+ye*frac*8);
+	      if(iBin == 0)
+		singleRatio[i ][iBin]->GetYaxis()->SetRangeUser(y-ye*frac*5,y+ye*frac*5);
+
 	      singleRatio[i ][iBin]->GetXaxis()->SetLimits(-.5,6.5);
 
 	      singleRatio[i ][iBin]->SetLineWidth(4);
 	      singleRatio[i ][iBin]->GetYaxis()->SetNdivisions(ndiv);
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/oldScripts/publishCutVariations_sn132_NOEff_pip.C

@@ -3,9 +3,6 @@
 using namespace style;
 
 
-
-
-
 void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
                      Float_t lMargin, Float_t rMargin,
                      Float_t bMargin, Float_t tMargin)
@@ -13,7 +10,7 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
   if (!C) return;
 
   // Setup Pad layout:
-  Float_t vSpacing = 0.0;
+  Float_t vSpacing = 0.05;
   Float_t vStep  = (1.- bMargin - tMargin - (Ny-1) * vSpacing) / Ny;
 
   Float_t hSpacing = 0.0;
@@ -90,13 +87,17 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn108(int this_var = 0)
+void publishCutVariations_sn132_NOEff_pip(int this_var = 3,TString filename = "blank")
 {
+  //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
-  int bins = 6;
+  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
+  vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
+    
+  int bins = 1;
   int var  = 4;
   int default_p = 3;
-  double frac = 1.2;
+  double frac = 4.;
 
   int lineSt = 6;
   int markCl = 1;
@@ -105,23 +106,23 @@ void publishCutVariations_sn108(int this_var = 0)
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("cutVariaiton_singleRatio.root");
+  TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields_NOEff.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132.C

@@ -10,7 +10,7 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
   if (!C) return;
 
   // Setup Pad layout:
-  Float_t vSpacing = 0.0;
+  Float_t vSpacing = 0.1;
   Float_t vStep  = (1.- bMargin - tMargin - (Ny-1) * vSpacing) / Ny;
 
   Float_t hSpacing = 0.0;
@@ -93,7 +93,8 @@ void publishCutVariations_sn132(int this_var = 3,TString filename = "blank")
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-    
+  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};
+  
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -106,33 +107,36 @@ void publishCutVariations_sn132(int this_var = 3,TString filename = "blank")
   int linew = 4;
   int ndiv = 405;
 
+  int Ny = 4;
+  int Nx = 1;
+
   TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields.root");
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.22,.04,.28,.04,.02,"<|");
-  TArrow *arrowR = new TArrow(.78,.04,.84,.04,.02,"|>");
+   TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
+  TArrow *arrowR = new TArrow(5.3,.03,5.7,.03,.02,"|>");
+
+  arrowL->SetNDC(true);
+  arrowR->SetNDC(true);
+
 
   gStyle->SetOptStat(0);
 
-  TCanvas *C = (TCanvas*) gROOT->FindObject("C");
-  if (C) delete C;
-  C = new TCanvas("C","canvas",1024,640);
-  C->SetFillStyle(4000);
-
-  // Number of PADS
-  const Int_t Nx = 1;
-  const Int_t Ny = 4;
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_NOEff_pip.C

@@ -87,7 +87,7 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank")
+void publishCutVariations_sn132_NOEff_pip(int this_var = 3,TString filename = "blank")
 {
   //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
@@ -110,15 +110,15 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
   TGraphErrors *singleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
-  TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
+   TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
   TArrow *arrowR = new TArrow(5.3,.03,5.7,.03,.02,"|>");
 
   arrowL->SetNDC();
   arrowR->SetNDC();
-  
+ 
   gStyle->SetOptStat(0);
 
-    TCanvas *C = new TCanvas("c1","",1000,300);
+  TCanvas *C = new TCanvas("c1","",1000,300);
 
   // Margins
   Float_t lMargin = 0.1;
@@ -157,7 +157,7 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
       for(int iVar = 0; iVar < var; iVar++)
 	{
 
-	  singleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpim_%d_%d",iBin,iVar));
+	  singleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpip_%d_%d",iBin,iVar));
 	}
     }
   
@@ -170,7 +170,8 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
 	{
 	  for (Int_t iBin = 0 ;iBin < Ny; iBin++)
 	    {
-	     
+	   
+
 	      singleRatio[i][iBin]->GetXaxis()->ChangeLabel(1,-1,-1,-1,-1,-1,Form("%d",labels.at(iBin).at(0)));
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_NOEff_ratio.C

@@ -87,13 +87,13 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank")
+void publishCutVariations_sn132_NOEff_ratio(int this_var = 3,TString filename = "blank")
 {
   //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};    
+    
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -108,6 +108,7 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
 
   TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields_NOEff.root");
   TGraphErrors *singleRatio[bins][var];
+  TGraphErrors *pipsingleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
   TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
@@ -156,11 +157,33 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
     {
       for(int iVar = 0; iVar < var; iVar++)
 	{
-
 	  singleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpim_%d_%d",iBin,iVar));
+	  pipsingleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpip_%d_%d",iBin,iVar));
+	  //calculate ratio
+	  for(int i = 0; i < singleRatio[iBin][iVar]->GetN(); i++)
+	    {
[…]

### 2020-06-18 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_ratio.C

@@ -87,13 +87,13 @@ void CanvasPartition(TCanvas *C,const Int_t Nx,const Int_t Ny,
 
 
 
-void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank")
+void publishCutVariations_sn132_ratio(int this_var = 3,TString filename = "blank")
 {
   //  vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {48,49,50,51,52,53,54}, {47,48,49,50,51,52,53} };
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};    
+    
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -106,8 +106,9 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
   int linew = 4;
   int ndiv = 405;
 
-  TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields_NOEff.root");
+  TFile *f = TFile::Open("../rootfiles/cutVariaiton_totalYields.root");
   TGraphErrors *singleRatio[bins][var];
+  TGraphErrors *pipsingleRatio[bins][var];
   TBox *box[bins][var];
   TLine *line[bins][var];
   TArrow *arrowL = new TArrow(.22,.03,.7,.03,.02,"<|");
@@ -156,11 +157,33 @@ void publishCutVariations_sn132_NOEff(int this_var = 3,TString filename = "blank
     {
       for(int iVar = 0; iVar < var; iVar++)
 	{
-
 	  singleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpim_%d_%d",iBin,iVar));
+	  pipsingleRatio[iBin][iVar] = (TGraphErrors *)f->Get(Form("sn132_totalpip_%d_%d",iBin,iVar));
[…]

### 2020-06-19 chapters/ch_appendix.tex
Backup

@@
 
+\section{Cut Variation Analysis}
+\label{sec:cutvar}
+The particular track quality cuts described in Section~\ref{sec:qualitycut} are there to reduced the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra.
+The best set of values were found for all the data sets which include the charged particle multiplicity of the event, dOCA, and the number of clusters cut.
+By varying the cuts in both the data and the MC efficiency, we can evaluate the analysis to see if there are any systematic dependencies on the observable we plot.
+But varying the cuts undoutably means as the cut get tighter, and less data is included, the statistical error bar will increase making it difficult to compare to more loose cuts.
+This is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}.
+Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts and the uncorrelated error bars, then there is some systematic trend that either exists due to physics or some error or miscalculation in the analysis.
+It is usually not recommended to estimated systematic error bars using this method, but sometimes is the only way.
+
+Here we will discuss the particular analysis using the total pion ratio as the example.
+Figure~\ref{fig:totalRatioError} shows the total pion as a function of several cut variations, where one cut is varied and all the others are held constant.
+Three cuts are of particular interest, the event multiplicity, dOCA, and the number of cluster cut.
+Figure~\ref{fig:clustTotalRatio} shows the variation as the number of cluster cut is varied.
+The default cut used is > 20 clusters which is represented by the middle point.
+The red bars represent the statistical error of this default value.
+The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}.
+Here it is clear that there is a systematic bias in the pion ratio as a function
+
+%show the pim total yield 
+%discuss others in the apendix
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
+         \caption{No efficiency correction.}
+         \label{fig:multCutVar}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:multCutVar}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
+\label{fig:cutVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_NoEC_pimTotal}
+         \caption{Before efficiency correction.}
+         \label{fig:pim_clustVar_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_pimTotal}
+         \caption{After efficiency correction.}
+         \label{fig:pim_clustVar_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the number of cluster cut of the tracks.}
+\label{fig:pim_clustVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_NoEC_pimTotal}
+         \caption{No efficiency correction.}
+         \label{fig:pim_cutVarPOCA_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_pimTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:pim_cutVarPOCA_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^-$) when varying the dOCA cut.}
+\label{fig:pim_cutVarPOCA}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pipTotal}
+         \caption{No efficiency correction.}
+         \label{fig:pip_multCutVar}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pipTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:pip_multCutVar}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
+\label{fig:cutVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_NoEC_pipTotal}
+         \caption{Before efficiency correction.}
+         \label{fig:pip_clustVar_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_pipTotal}
+         \caption{After efficiency correction.}
+         \label{fig:pip_clustVar_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
+\label{fig:pim_clustVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_NoEC_pipTotal}
+         \caption{No efficiency correction.}
+         \label{fig:pip_cutVarPOCA_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_pipTotal}
+         \caption{Efficiency corrected.}
+         \label{fig:pip_cutVarPOCA_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$) when varying the dOCA cut.}
+\label{fig:pip_cutVarPOCA}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_TotalRatio}
+         \caption{No efficiency correction.}
+         \label{fig:ratio_multCutVar}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{132Sn_Multiplicity_TotalRatio}
+         \caption{Efficiency corrected.}
+         \label{fig:ratio_multCutVar}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
+\label{fig:cutVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_NoEC_TotalRatio}
+         \caption{Before efficiency correction.}
+         \label{fig:ratio_clustVar_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{Clusters_TotalRatio}
+         \caption{After efficiency correction.}
+         \label{fig:ratio_clustVar_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
+\label{fig:pim_clustVar}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering
+     \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_NoEC_TotalRatio}
+         \caption{No efficiency correction.}
+         \label{fig:ratio_cutVarPOCA_NoEC}
+     \end{subfigure}
+     \hfill
+    \begin{subfigure}[b]{\textwidth}
+         \centering
+         \includegraphics[width=\textwidth]{POCA_TotalRatio}
+         \caption{Efficiency corrected.}
+         \label{fig:ratio_cutVarPOCA_EC}
+     \end{subfigure}
+     \hfill
+\caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}
+\label{fig:pip_cutVarPOCA}
+\end{figure}
+
 \chapter{Second appendix}

### 2020-06-19 chapters/ch_intro.tex
Backup

@@
 It is for this reason that the distance between nuclei is a near constant value, and therefore the density is remarkably constant over a wide range of nuclei CITE HERE.
-This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$.
+This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \SI{0.16}{\per\femto\metre\cubed}.
 
@@
 This picture was remarkably successful at describing the binding energies of nuclei at saturation density.
-The Bethe-Weizsacker semi-empirical formula predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:
+The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:
 
 \begin{equation}
-\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + \mathcal{O}
+\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
 \label{eq:semiEmp}
@@
 There are correction terms accounting for the surface, $a_s$, since nucleons near the surface do not have as many neighbors as nucleons inside, and the coulomb term which is related to the size of the nucleus which scales as $A^{1/3}$.
-The asymmetry term, $a_A$, is related to the cost in energy one pays to become more neutron or proton rich; it is typically referred to as the \emph{Symmetry Energy}.
+The asymmetry term, $a_A$, is related to the cost in energy one pays to become more neutron or proton rich; it is typically referred to as the \emph{symmetry energy}.
 This originates from pauli blocking where it is more energetically favorable to form neutron-proton pairs since their isospin numbers are different.
@@
 for a fixed number of particles $N$ and zero temperature.
-One can extend to higher temperatures by adding the Boltzman dependence CITE HERE.
+One can extend to higher temperatures by adding the Boltzman dependence.
 The partial derivative with respects to volume can be rewritten in terms of density:
@@
 Typically an effective interaction is used to describe the phenomenological observations of nucleon-nucleon interactions observed in nuclei.
-One of such interactions is described by the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations CITE HERE.
+One of such interactions is described by the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations \cite{skyrme}.
 Several Skyrme parameterizations are shown as lines in Fig.~\ref{fig:symDen}.
@@
 By measuring protons and neutrons we can see signatures of the effects of the symmetry energy in the final spectra of these particles.
-Measuring neutrons experimentally can be quite challenging and space is limited by large neutron wall arrays CITE HERE.
+Measuring neutrons experimentally can be quite challenging and space is limited by large neutron wall arrays.
 Also, though the overlap region temporarily reaches a high density, the neutrons which participated in this region also evolve through regions of lower densities until they reach their final state, diluting the signal from the dense region.
 
-\begin{figure}
+\begin{figure}[!htb]
 \centering
-\includegraphics[scale=.01]{place}
-\caption{Experimental constraints of the density dependence of the symmetry energy}
+\includegraphics[scale=.5]{constraints}
+\caption{Experimental constraints of the density dependence of the symmetry energy taken from \cite{awayforward}}
 \label{fig:symDen}
@@
 It is preferable to find an observable that is easier to measure experimentally than neutrons, and is more sensitive to the high density region.
-Pions are produced though and intermediate process where nucleon-nucleon collisions form an excited $\Delta$ baryon resonance from one of the nucleons, which then decay decay shortly after into a pion:
+Pions are produced though and intermediate process where nucleon-nucleon collisions form an excited $\Delta$(1232) baryon resonance from one of the nucleons, which then decay decay shortly after into a pion:
 
@@
 \end{equation}
+
+The threshold for $\Delta$ resonance production, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} kinetic energy.
+In large nuclei the nucleons exist in orbitals which have large amounts of energy (Fermi energy), allowing for $\Delta$ production even at sub-threshold beam energies \cite{fermiEnergy}.
 
@@
 \centering
-\includegraphics[scale=.2]{deltaProduction.png}
+\includegraphics[width=.6\textwidth]{deltaProduction.png}
 \caption{Figure taken from \cite{mingzhang} for Au + Au collisions at \SI{400}{\MeVA}.
@@
 In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
-Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the froward and backwards process $\ce{\Delta <=> \pi N}$.
+Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the froward and backwards process $\Delta <=> \pi + N$.
 This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission would change the pion dynamics or even the charge of the pion which would reflect the asymmetry at the point of creation and eventual re-emission.
@@
 Instead of measuring neutrons we can measure the $\pi^-$ which is much easier to measure experimentally.
+
+\section{Motivation for Thesis}
+In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collision utilizing inverse kinematics where the beam is made of a radio-active neutron rich beam.
+This allows for the neutron-proton asymmetry of the system to be changed depending on the beam.
+Pion production has been studied in stable beams for beam energies of \SI{400}{\MeVA} and above CITE HERE.
+Here only total pion yields were published and no pion spectra were published.
+The goals of this Thesis were to measure the pion spectra efficiently, to low pion energies.
+To do this a new Time Projection Chamber (TPC) was made, where we measured pions and light charged particles (up to Li) in neutron-rich heavy ion collisions at \SI{270}{\MeVA}
 
@@
 %pion production
+
+\begin{comment}
 
@@
 Effort to reduce the experimental error bars on such spectra which theory could use
+
+\end{comment}

### 2020-06-19 chapters/ch_results.tex
Backup

--- 2020-06-14 chapters/ch_results.tex
@@
 \chapter{Results}
-\cite{jon}
 
 \section{Pion Spectra}
+\label{sec:pionSpectra}
 
@@
 
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{pionSpectra.png}
+\caption{Pion spectra. }
+\label{fig:pionspectra}
+\end{figure}
 %need to add error analysis
@@
 
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{pionSpectra.png}
-\caption{Pion spectra. }
-\label{fig:pionspectra}
-\end{figure}
-
 %Add figure on Glauber model 
 %Add figure on comparison to FOPI data
+
+\section{Comparison to Theory}
+%Add figure of Pion spectra for different theory 
+%Reference paper for multiple theories
+
+Here we will compare the data with 7 different commonly used transport theoretical codes.
+These codes took part in a large collaboration effort to compare the codes to each other.
+The goal of these comparison projects was to standardize the input into each code to systematically go through the details in each code comparing things such as initialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc \cite{theoryComp1,theoryComp2} CITE MORE HERE.
+These codes were taken in their best state, without any prior knowledge of the experimental data, and simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
+Each code differs in its treatment of pion and $\Delta$ dynamics.
+Some codes contain modifications to the pion in nuclear matter, usually by including a pion optical potential.
+Some codes include the iso-scalar and the iso-vector delta potential which are not very well constrained but are important in the production of pions CITE HERE.
+What will be the consistent theme will be the large variation between theoretical codes.
+Though many of the numerical and theoretical treatments have been addressed, many of the details of pion production and dynamics have not been addressed.
+Though several codes show sensitivity to different symmetry energies within a particular code, the variation between codes  was much larger.
 
@@
 The integrated pion yield for both systems and the $\pi^-/\pi^+$ ratio is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
-It is remarkable that the pion ratio is significantly greater than the N/Z of the system which is what is expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ CITE HERE; 2x in the $\tin{132}{124}$ system and 1.4x in the $\tin{108}{112}$ system.
+It is remarkable that the pion ratio is significantly greater than the N/Z of the system which is what is expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}; 2x in the $\tin{132}{124}$ system and 1.4x in the $\tin{108}{112}$ system.
 The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect and lower the ratio as the system tends toward isospin equilibrium.
@@
 For example the $\Delta$ resonance potential has an iso-scalar component (independent of iso-spin) and an iso-vector component.
-The nature of these two potentials is still unknown CITE HERE, and the role it plays in pion production has shown to be very important CITE HERE.
+The nature of these two potentials is still unknown, and the role it plays in pion production has shown to be very important \cite{baoan_deltapotential}.
 
@@
 The Symmetry Energy variation of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and all codes in the double ratio; the data band represents the data error bar.
-Certainly the variation between Symmetry Energy in the $\tin{132}{124}$, though small, still exists in many of the codes as predicted initially CITE HERE BAO ANN.
+Certainly the variation between Symmetry Energy in the $\tin{132}{124}$, though small, still exists in many of the codes as predicted initially \cite{baoan_piprod1,baoan_piprod2}.
 The small error bars in the data also would allow for a detailed analysis to extract a constraint of the Symmetry Energy, if the variation between codes could be solved.
@@
 \end{figure}
+
+The pion spectral ratio is a promising observable in particular it may be more sensitive to the high density regions of the early collision.
+In theory the high energy pions are more likely to exit the nuclear medium earlier, and therefore be less prone to effects such as pion absorption and re-emission which dilute the sensitivity of the pion observable to the high density behavior.
+Also low energy pions are more likely to be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
+If we integrate the total pion yield we combine these two regions of low and high energy pions.
+Instead we can construct the $Y(\pi^-)/Y(\pi^+)$ ratio as a function of the kinetic energy in the CM system.
+
+%Mention and cite appendix for systematics
+
+Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of accuracy.
+The general hyperbolic shape comes from the Coulomb force on the two charged pion spectra mentioned in Section~\ref{sec:pionSpectra}.
+Naturally the pion ratio is smaller for less neutron poor system, i.e.
+less neutron-neutron collisions produce less $\pi^-$.
+The bin size of the last bin was increased to reduce the statistical error bars since the number of pions, especially the $\pi^+$, is reaching the limits of the measured distribution.
 
@@
 
+Another promising observable is the spectral double ratio.
+In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory, to cancel out.
+For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
+
 \section{Comparison to Previous Data Sets (FOPI)}
-asdfasdfasd
-\section{Comparison to Theory}
-%Add figure of Pion spectra for different theory 
-%Reference paper for multiple theories
-asdfasdf
 
-\section{Systematic Errors and Cut Variations}
-\label{sec:cutvar}
-dfsadfsaf
-\section{Constraint on the Symmetry Energy}
-%maybe not...
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.4]{fopi_pionratio_comp}
+\caption{Comparing the total $\pi^-/\pi^+$ ratio of the $\tin{132}{124}$ system to the ${}^{197}$Au + ${}^{197}$Au data from the FOPI collaboration.
+The \spirit TPC data was scaled by a factor to compare to the lower N/Z of the Au + Au system.
+This was extracted from measuring the N/Z dependence measured in the experiment.
+}
+\label{fig:fopiPionRatio}
+\end{figure}
+
+The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher beam energies.
+In the $\tin{132}{124}$ data  $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$, $N/Z=1.493$.
+The pion ratio is proportional to the $N/Z$ of the system.
+Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}.
+Here the dependence is fitted with a 2-nd order polynomial fit.
+The pion ratio in the $\tin{132}{124}$ data  was scaled by the ratio between $N/Z=1.56$ and 1.493.
+Figure~\ref{fig:fopiPionRatio} shows the scaled pion ratio as compared with the FOPI pion ratio \cite{fopi}.
+The fitted function has the functional form of $p_1(E - p_2)^{-2}$ where $p_1$ and $p_2$ are free parameters, and is meant to guide the eye.
+It is worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}.
+Scaling the other systems in the \spirit TPC data leads to almost the exact same value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$.

### 2020-06-19 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_NOEff_ratio.C

@@ -93,7 +93,7 @@ void publishCutVariations_sn132_NOEff_ratio(int this_var = 3,TString filename = 
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-    
+  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};    
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -258,7 +258,7 @@ void publishCutVariations_sn132_NOEff_ratio(int this_var = 3,TString filename = 
 	      //   arrowR->SetLineWidth(linew);
 	      // arrowL->Draw();
 	      // arrowR->Draw();    
-	      C -> SaveAs(labels_n.at(iBin) + "_NoEC_TotalRatio.png");
+	      C -> SaveAs(labels_f.at(iBin) + "_NoEC_TotalRatio.png");
 	    }
 	}
     

### 2020-06-19 macrosForThesis/PionMultiplicity/PionData/CutVariation/publishCutVariations_sn132_ratio.C

@@ -93,7 +93,8 @@ void publishCutVariations_sn132_ratio(int this_var = 3,TString filename = "blank
 
   vector< vector<int>> labels = { {14,16,18,20,22,24,26}, {14,16,18,20,22,24,26}, {35,40,45,51,55,60,65}, {35,40,45,50,55,60,65} };
   vector<TString> labels_n = {"# Clusters","POCA","{}^{132}Sn Multiplicity","{}^{108}Sn Multiplicity"};
-    
+  vector<TString> labels_f = {"Clusters","POCA","132Sn_Multiplicity","108Sn_Multiplicity"};
+  
   int bins = 1;
   int var  = 4;
   int default_p = 3;
@@ -261,7 +262,7 @@ void publishCutVariations_sn132_ratio(int this_var = 3,TString filename = "blank
 	      //   arrowR->SetLineWidth(linew);
 	      // arrowL->Draw();
 	      // arrowR->Draw();    
-	      C -> SaveAs(labels_n.at(iBin) + "_TotalRatio.png");
+	      C -> SaveAs(labels_f.at(iBin) + "_TotalRatio.png");
 	    }
 	}
     

### 2020-06-23 chapters/ch_data_analysis.tex
backup

@@
 The hits within a helix track are dense and the position resolution of an individual hit is not very precise.
-As mentioned in Section~\ref{sec:padResponse}, the TPC can achieve sub-millimeter position resolution by utilizing the PRF.
+As mentioned in Section~\ref{sec:prf}, the TPC can achieve sub-millimeter position resolution by utilizing the PRF.
 In the helix track finding task, the hits are further reduced into more precise clusters.
@@
 As the beam passes through the field cage it ionizes the gas creating electron-ion pairs.
-The drift velocities of the ions are typically \SI{e4} times slower than electron drift velocities \cite[blumrol].
+The drift velocities of the ions are typically \SI{e4} times slower than electron drift velocities \cite{blumrol}.
 Any source of ions have the potential to build up in the drift volume, creating a positive space charge, distorting the drifting electrons and  biasing the track momentum measurement.
@@
 \caption{ $\Delta V_{LR}$ versus the beam rate for all systems with the fitted function.}
-\label{fig:vlr_br}
+\label{fig:VLR_br}
 \end{figure}
@@
     \end{subfigure}
+\caption{Momentum distributions of the Left and Right sides of the TPC for $\pi^+$ and $\pi^-$ particles.}
 \label{fig:sc_momdist}
@@
 
+  \end{center}
   \caption{Cobo timing correction.}
-  \end{center}
   \label{fig:coboCorr}
@@
     \caption{}
-\label{fig:embedtrack}
+\label{fig:mcDataEmbedtrack}
 \end{figure}
@@
 
-We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{mcdataPRF}, and the data PRF, Fig.~\ref{prfpimData}, where the black line is the PRF fit to the experimental data.
+We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Fig.~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data.
 It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.

### 2020-06-23 chapters/ch_intro.tex
backup

--- 2020-04-04 chapters/ch_intro.tex
@@
 \chapter{Introduction}
+Mankind has been interested in what is the nature of the matter which composes the visible universe, and what are the constituents that make it up.
+The most significant discovery on the study of fundamental sub-atomic particles was Rutherford's discovery in 1911 in which he concluded that atomic nuclei are composed of a dense nucleus surrounded by an electron cloud.
+Later it was discovered that the nucleus is composed of nucleons, a positive charge (protons) and a neutral charge (neutrons).
+The nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it incredibly dense.
+
+Without a balancing force the Coulomb force between protons would render the nucleus unstable.
+This balancing force is called the strong nuclear force which exerts a large attractive force over a very short range.
+The strong force itself is the fundamental interaction between the fundamental quarks which make up the nucleons.
+At low energies the quark structure of nucleons is less important and nucleons can be thought of as fundamental particles.
+The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress.
+It is for this reason that the distance between nuclei is a near constant value, and therefore the density distribution over a wide range of nuclei is remarkably constant CITE HERE.
+This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.
+
+Due to the nature of the strong force, nuclei can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility.
+This picture was remarkably successful at describing the binding energies of normal nuclea matter at saturation density.
+The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:
+
+\begin{equation}
+\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
+\label{eq:semiEmp}
+\end{equation}
+
+Since the strong force makes the inter-nucleon distance approximately constant, the volume of the nucleus is related to the number of total nucleons, therefore the volume term $a_V$ is proportional to $A$.
+There are several correction terms accounting for the surface, $a_s$, since nucleons near the surface have fewer surrounding neighbors as nucleons inside, and the coulomb term which is related to the typical coulomb force which scales like the radius of the nuclei $R^{-1}$; which scales as $A^{1/3}$ (since volume scales like A).
+The asymmetry term, $a_A$, is related to the cost in energy to become more neutron or proton rich; it is typically referred to as the \emph{symmetry energy}.
+This originates from Pauli blocking where it is more energetically favorable to form neutron-proton pairs since their isospin numbers are different.
+The di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, where as the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more energetically favorable.
+
+Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter CITE HERE with a typical size of \SI{11}{\kilo\metre} in radius.
+The di-neutron system is typically not stable in such an extreme system except for the the large gravitational force, which balances the strong forces.
+In this extreme example of dense nuclear matter, the pressures generates in the neutron star range from low densities near the crust to the dense interior which can reach up to 9$\rho_o$ CITE HERE.
+To understand these exotic forms of nuclear matter, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}, where we must describe nuclear matter over a wide range of densities.
+
+Guided by Eq.~\ref{eq:semiEmp}, we can separate the energy density $E$ of a system into two components,
+
+\begin{equation}
+E(\rho,\delta) = E(\rho	) + S(\rho)\delta^2,
+\label{eq:energyEos}
+\end{equation}
+
+where $E(\rho)$ describes the symmetric term (i.e.
+independent of isospin of the nucleons), and the symmetry energy $S(\rho)$ which depends on the asymmetry of the system, written now in terms of the neutron and proton densities,
+
+\begin{equation}
+\delta = \frac{\rho_n - \rho_p}{\rho}.
+\label{eq:asym}
+\end{equation}
+
+The Equation of State (EoS) of nuclear matter can be calculated by fundamental thermodynamic relations,
+
+\begin{equation}
+P = \Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N}
+\label{eq:pressEos}
+\end{equation}
+
+for a fixed number of particles $N$ and zero temperature.
+One can always extend to higher temperatures by adding the typical Boltzman dependence if needed, but here the simplification will suffice.
+The partial derivative with respects to volume can be rewritten in terms of density:
+
+\begin{equation}
+P = -\rho^2 \frac{dE}{d\rho}\vert_{T=0,N}.
+\label{eq:densEos}
+\end{equation}
+
+The picture that is built up here is the gravitational force is attempting to compress and in-compressible matter, and the fundamental forces of the strong force build up a competing pressure which is related to the derivative w.r.t density of the symmetry energy \cite{tovEq}.
+
+\section{Density Dependence of the Symmetry Energy}
+In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ CITE HERE.
+In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$.
+Figure~\ref{fig:symDen} shows some of the experimental constraints which have been performed by a series of independent measurements and observables \cite{awayforward}.
+Typically the density dependence of the symmetry energy can be described through an effective interaction used to describe the phenomenological observations of nucleon-nucleon interactions observed in nuclei.
+One of such interactions is the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations \cite{skyrme}.
+Several Skyrme parameterizations are shown as lines in Fig.~\ref{fig:symDen}.
+Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars.
+The goal of the nuclear EoS community over the last decade have been particularly focused on constraining the symmetry energy at these high densities.
+
+\section{Heavy Ion Collisions}
+Besides observing neutron stars directly, heavy-ion collisions (HIC) provide the only way we can probe the density dependence of the symmetry energy in the laboratory setting.
+When two nuclei collide in a collision, in the very early stages they compress to form a high density region where the nuclei overlap.
+This momentary density can reach up to $3\rho_0$ depending on the incident beam energy.
+HICs also provide the only way we can probe the isospin asymmetry dependence of the nuclear EoS.
+This is accomplished by using radioactive neutron-rich beams to collide on stable targets.
+
+The pressure arising from the symmetry energy depends on the curvature of the symmetry energy at a given density as described in Eq.~\ref{eq:densEos}.
+If the density dependence of the symmetry energy is positive at high densities the symmetry energy would in general force neutrons out of the high density system.
+Where as if the derivative was negative the symmetry energy would attract neutrons.
+The curvature of the symmetry energy at a particular density is commonly referred to as \emph{soft} in the case the curvatures turns over, and \emph{stiff} in the case of a strong positive curvature.
+It is the symmetry energy pressure that is driving the relative dynamics of neutrons and protons, where a strong symmetry energy expels neutrons from dense matter and vice versa.
+By measuring protons and neutrons, we can see signatures of the symmetry energy in the final spectra of these particles.
+There are two experimental challenges.
+First,  measuring neutrons can be quite challenging and space and acceptance is limited by large neutron wall arrays.
+Lastly, though the overlap region of the two nuclei temporarily reaches a high density, the nucleons which participated in this region continue to evolve throughout the collision, even into regions of lower densities until they reach their final state.
+They are continuously interacting with the rest of the nuclear matter, covering a wide range of densities, diluting any symmetry energy effects and complicating the high density constraint.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.5]{constraints}
+\caption{Experimental constraints of the density dependence of the symmetry energy taken from \cite{awayforward}}
+\label{fig:symDen}
+\end{figure}
+
+\section{Pion Observable}
+It is preferable to find an observable that is easier to measure experimentally than neutrons, and is more sensitive to the just the high density region.
+Pions are produced though an intermediate step via $ NN \leftrightarrow \Delta$, where nucleon-nucleon collisions form an excited $\Delta$(1232) baryon resonance from one of the nucleons, which then decay decay shortly after into a pion $\Delta \leftrightarrow \pi N$.
+The threshold for $\Delta$ resonance production, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
+In large nuclei, the internal motion of the nucleons is substantial and non-negligible.
+In the Fermi gas model, nucleons are arranged filling up higher energy levels up until they reach the Fermi energy.
+This extra energy allows for $\Delta$ production even at sub-threshold beam energies \cite{fermiEnergy}.
+
+It has been shown that most of the $\Delta$'s are produced in the early dense regions of the collision \cite{mingzhang}.
+Figure~\ref{fig:deltaProduction} shows the average local density (c) which $\Delta$'s are produced and the number in the system (b), as a function of time in the simulation of Au + Au collisions at \SI{400}{\MeVA}.
+Panel (a) shows the density distribution of the density at the moment of creation for $\Delta$'s.
+Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon.
+Thus the outgoing $\pi$ is contains information on the high density region of the collision.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=.6\textwidth]{deltaProduction.png}
+\caption{Figure taken from \cite{mingzhang} for Au + Au collisions at \SI{400}{\MeVA}.
+Panel (a) shows the density in the region of the $\Delta$ resonance creation for two different symmetry energies (x=0 soft) and (x=1 stiff).
+Panel (b) and (c) show the evolution of collision in time steps, where (b) shows the number of deltas in the system as a function of time and (c) shows the mean local baryon density in the region where $\Delta$ resonances are produced.
+The blue line in (c) represents the average baryon density in the most central region of the collision.
+This evidence shows that a majority of $\Delta$'s are produced in the high density region of the early collision.}
+\label{fig:deltaProduction}
+\end{figure}
+
+The branching ratio of the various flavors of $\Delta$'s is given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}.
+Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$.
+In this $\Delta$ resonance model the charged pion ratio can be described as,
+
+\begin{equation}
+\frac{\pi^-}{\pi^+} = \frac{ 5pp + pn }{5nn + pn}.
+\label{eq:deltaModel}
+\end{equation}
+
+In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
+Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the backward process of $\Delta \leftrightarrow \pi + N$.
+This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the asymmetry at the point of creation and re-emission.
+Total pion absorption back into two nucleons requires a three body process, where a pion is absorbed creating a $\Delta$ resonance, then another nucleon must collide with the resonance to create two nucleons.
+Because of this, the total pion absorption (removing pions from the system) is a smaller effect than the absorption re-emission process.
+In general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision, effectively turning the neutrons  measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally.
+
+\section{Motivation for Thesis}
+In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}.
+Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam impinging on a stable Sn target.
+We can probe the isospin asymmetry of the symmetry energy by changing the neutron-proton asymmetry of the incoming beam.
+Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}.
+In these previous experiments, only total pion yields were published and no pion spectra were published, and the efficiency of the detector at detecting low energy pions was difficult to estimate.
+Also though the pion production increases with the beam energy, the symmetry energy affects the dynamics of the system less.
+This is because the time which the two nuclei coalesce in a HIC becomes shorter, and the time in which the symmetry energy can affect nucleons is shortened, thus the effect will becomes less dominant at higher beam energies.
+The goals of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs.
+To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) Time Projection Chamber (TPC).
+
+%nuclear chart
+%Dense nuclear matter 
+%Liquid drop model explain density around saturation
+%explain symmetry energy 
+%extend to systems of larger pressure, isospin assymetry 
+%symmetry energy how to calculate 
+%Previous constraints on symmetry energy
+%heavy ion collisions only way to probe density and 
+%dynamics of proton neutrons
+
+%pion production
+
+\begin{comment}
 

### 2020-06-23 chapters/ch_results.tex
backup

--- 2020-06-14 chapters/ch_results.tex
@@
 \chapter{Results}
-\cite{jon}
 
 \section{Pion Spectra}
+\label{sec:pionSpectra}
 
-The efficiency correction is applied track-by-track, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
-The correction factor $C_i$ is defined as,
+To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
+The correction factor $C_i$ is then defined as,
 
@@
 
-Each track that is identified as a pion, is then weighted by the correction factor $C_i$, when filling any histogram.
-By using the track-by-track method, we can transform the track into any observable of interest, notably we can transform the  track from the lab frame into the center-of-mass (CM) frame.
-Before performing the transformation to the CM system recall the beam is at a small, but non-negligible, angle in the Lab frame, see Section~\ref{sec:beamangle}.
-The beam angle for each even is measured and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis.
+Each track that is identified as a pion as describe in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable.
+In this way, we can correct for the efficiency of the track into any observable of interest, notably transforming from the lab frame into the center-of-mass (CM) frame.
+Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame; see Section~\ref{sec:beamangle}.
+The beam angle for each event is measured by the beam tracking software and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis.
 Doing so makes the transformation into the CM system much simpler to describe.
 If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.
+A rotation is then applied around the vector $\hat{v}$.
 
 Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation.
-Where the 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
+The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
 Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,
@@
 
-%need to add error analysis
-%systematic error analysis
-
-Figure~\ref{fig:pionspectra} shows the pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
-This data marks the first pion spectral data at sub-threshold energies, extending to very low pion energies.
-The coulomb potential works to accelerate $\pi^+$ and decelerate $\pi^-$ particles due to the positive charge of the nuclear medium.
-The coulomb barrier prevents the production of low energy $\pi^+$, which is why there is a reduction in the phase space around zero kinetic energy.
-
 \begin{figure}[!htb]
@@
 \end{figure}
+%need to add error analysis
+%systematic error analysis
+
+Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$.
+This data marks the first time pion spectral data  has been measured at sub-threshold energies.
+In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerate $\pi^-$ particles, due to the positive charge of the nuclear medium.
+The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons.
+The effects of the Coulomb force are the largest effects on the spectra of the pions and will play an important role when looking at pion spectral ratios.
 
@@
 
+\section{Comparison to Theory}
+%Add figure of Pion spectra for different theory 
+%Reference paper for multiple theories
+
+The data taken has been compared with 7 commonly used transport theoretical codes.
+These codes took part in a large collaborative effort in order to standarize certain common elements of each codes.
+Each comparison project simulating nuclear matter in a box type simulation with fixed, well defined, initial conditions.
+Though not physical, this provided a benchmark for codes to compare to; the solutions of the total number of pions and $\Delta$ produced had an analytical solution one would expect from statistical equalibrium, providing a good benchmark.
+This allowed for each code to systematically go through the numerical treatment and details in each code such as intialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc.
+\cite{theoryComp1,theoryComp2}.
+These codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data.
+We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
+While numerical treatments of each code are resonably similar, each code differs in its treatment of pion and $\Delta$ dynamics.
+Some codes contain modifications to how the pion behaves in nuclear matter, i.e.
+\emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorptions.
+Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but are important in the production of pions CITE HERE.
+The set of 7 codes represent the best codes which can simulate pions at these energies.
+As will be the re-occuring theme, there is a large variation between theoretical codes; greater than the variation between different symmetry energy in a particular code.
+Though many of the numerical and theoretical treatments have been addressed, many of the details of pion production and dynamics have not been addressed.
+This is an extremely complex task requiring major theoretical efforts and new data.
+We expect the data in this thesis and in \cite{jon} to provide much needed experimental data in which each code may compare too.
+The total theoretical uncertainty is too large relative to the experimental error bars to make a constraint on the density dependence of the symmetry energy.
+Any constraint would depend on the particular code one is using and would result in conflicting conclusions when using another theory.
+There is no good consensus at this time which codes should be favored over other codes, as the magnitude of the effects included in each code are still debated in order of importance.
+Here I will present the current level of agreement with the codes and remark on some of the implications it has moving forward.
+
 \section{Pion Yield}
 
-The integrated pion yield for both systems and the $\pi^-/\pi^+$ ratio is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
-It is remarkable that the pion ratio is significantly greater than the N/Z of the system which is what is expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ CITE HERE; 2x in the $\tin{132}{124}$ system and 1.4x in the $\tin{108}{112}$ system.
-The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect and lower the ratio as the system tends toward isospin equilibrium.
-A na\"ive interpretation would be the high density N/Z fraction is much higher for some reason than the two systems, though, the system spends very little time in this state which would require a very soft Symmetry Energy.
-A more likely conclusion is the mechanisms involved in pion productions via the $\Delta(1232)$ resonance is more complicated than the simple $(N/Z)^2$ relation.
-For example the $\Delta$ resonance potential has an iso-scalar component (independent of iso-spin) and an iso-vector component.
-The nature of these two potentials is still unknown CITE HERE, and the role it plays in pion production has shown to be very important CITE HERE.
+The integrated pion yield for both systems, and the $\pi^-/\pi^+$ ratio, is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
+It is remarkable that the pion ratio is significantly greater than the $N/Z$ of the system, as expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}.
+In the $\tin{132}{124}$ system this is a factor of 2 times greater and in $\tin{108}{112}$ the system a factor of 1.4 times greater.
+The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect, lowering the pion ratio as the system tends toward isospin equilibrium.
+A na\"ive interpretation would be the high density N/Z fraction is much higher than the average of the target and projectile, though, the system spends very little time in this early high density phase and there is really no time for the system to evole such that it could be enhanced in neutron fraction.
+A more likely conclusion is the mechanisms involved in pion productions via the $\Delta(1232)$ resonance is more complicated than the simple $(N/Z)^2$ relation, which was also seen in \cite{fopi}.
+Some codes assume the potential of the delta is just the same as the corresponding nucleon potential, as given by the iso-spin.
+Other codes introduce $\Delta$ resonance in-medium potential, which has an iso-scalar component (independent of iso-spin) and an iso-vector component.
+The nature of these two potentials is still unknown, and the role it plays in pion production has shown to be very important \cite{baoan_deltapotential}.
 
@@
 
-We have compared the total pion yields and ratios to the 6 common transport codes for the systems measured.
+We have compared the total pion yields and ratios to the 7 common transport codes for the systems measured.
 The table of the transport codes are listed in Appendix~\ref{tb:pionyieldTheory}.
 Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the codes.
-The codes plotted here are only the soft symmetry energy since the variation in code is much larger than the intra-code variation between different Symmetry Energies.
-While some codes make a reasonable approximation of a particular charge pion, no code reasonably predict both.
+The codes plotted here are only the soft symmetry energy since the variation in code is much larger than the variation within a code between different symmetry energies.
+While some codes make a reasonable approximation of a particular charge pion yields, no code reasonably predicts both.
 
@@
 Figure~\ref{fig:totalpiRatio} shows the total single pion ratio and the double ratio of the $\tin{132}{124}$ and $\tin{108}{112}$ system.
-Here, the variation between codes is much larger than the variation between the Symmetry Energy within a particular code.
-The Symmetry Energy variation of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and all codes in the double ratio; the data band represents the data error bar.
-Certainly the variation between Symmetry Energy in the $\tin{132}{124}$, though small, still exists in many of the codes as predicted initially CITE HERE BAO ANN.
-The small error bars in the data also would allow for a detailed analysis to extract a constraint of the Symmetry Energy, if the variation between codes could be solved.
+Here, the variation between codes is much larger than the variation between the symmetry energy within a particular code.
+The symmetry energy variation (soft and stiff extremes) of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and in all codes in the double ratio.
+The data ciruclar markers and band represents the total error bar.
+Certainly the variation between symmetry energy extremes in the codes, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}.
+The small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large variation between codes.
 
@@
 \includegraphics[width=\textwidth]{totalpiRatio.png}
-\caption{Total pion ratio and double ratio compared with 6 common transport codes.}
+\caption{Total pion ratio and double ratio compared with 7 common transport codes.}
 \label{fig:totalpiRatio}
@@
 \end{figure}
+
+The pion spectral ratio is a rather promising observable.
+In particular it may be more sensitive to the high density regions of the early collision.
+In theory the high energy pions are more likely to exit the nuclear medium earlier, and therefore be less prone to effects such as pion absorption and re-emission which dilute the sensitivity of the pion observable to the high density behavior.
+Also low energy pions are more likely to be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
+If we integrate the total pion yield we would combine these two regions diluting the observable.
+Instead we can construct the $Y(\pi^-)/Y(\pi^+)$ ratio as a function of the kinetic energy in the CM system with a particular focus on high energy pions.
+
+%Mention and cite appendix for systematics
+
+Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of efficiency and accuracy.
+The general hyperbolic shape arises from the Coulomb force on the two charged pion spectra mentioned in Section~\ref{sec:pionSpectra}.
+Also notice the pion ratio is smaller for less neutron poor system, as we would expect since less neutron-neutron collisions produce less $\pi^-$.
+The bin size of the last bin was increased to reduce the statistical error bars since the number of pions, especially the $\pi^+$, reaches the limits of the measured distribution.
 
@@
 
+Another promising observable is the spectral double ratio.
+In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory, to cancel out.
+For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
+
 \section{Comparison to Previous Data Sets (FOPI)}
-asdfasdfasd
-\section{Comparison to Theory}
-%Add figure of Pion spectra for different theory 
-%Reference paper for multiple theories
-asdfasdf
 
-\section{Systematic Errors and Cut Variations}
-\label{sec:cutvar}
-dfsadfsaf
-\section{Constraint on the Symmetry Energy}
-%maybe not...
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.4]{fopi_pionratio_comp}
+\caption{Comparing the total $\pi^-/\pi^+$ ratio of the $\tin{132}{124}$ system to the ${}^{197}$Au + ${}^{197}$Au data from the FOPI collaboration.
+The \spirit TPC data was scaled by a factor to compare to the lower N/Z of the Au + Au system.
+This was extracted from measuring the N/Z dependence measured in the experiment.
+}
+\label{fig:fopiPionRatio}
+\end{figure}
+
+The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies.
+In the $\tin{132}{124}$ data the $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
+We also expect the pion ratio is proportional to the $N/Z$ of the system.
+Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}.
+Here the dependence is fitted with a 2-nd order polynomial fit.
+To compare the pion ratio in the $\tin{132}{124}$ data with that of the lower N/Z in the FOPI experiment, we scaled by the pion ratio between $N/Z=1.56$ and 1.493 as given from the fitted polynomial line.
+Figure~\ref{fig:fopiPionRatio} shows the scaled pion ratio as compared with the FOPI pion ratio \cite{fopi}.
+The fitted function has the functional form of $p_1(E - p_2)^{-2}$ where $p_1$ and $p_2$ are free parameters, and only is meant to guide the eye.
+It is also worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}.
+Scaling the other systems in the \spirit TPC data leads to almost the exact same value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.

### 2020-06-24 chapters/ch_experiment.tex
backup
Bill edits added Experiment section

@@
 
-\begin{figure}
-\includegraphics[width=\textwidth]{tpcPrinciple.pdf}
-\label{fig:tpcPrinciple}
-\caption{Operation principle of the TPC}
-\end{figure}
 %explain in breif the basics of a time projection chamber 
@@
 The charge and timing  information of these electrons are calculated from the induced signal measured by the  electronics.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.5]{tpcPrinciple.pdf}
+\caption{Operation principle of the TPC}
+\label{fig:tpcPrinciple}
+\end{figure}
 
@@
 
-Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
+Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
 The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
@@
 The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter.
-They are biased to high voltages around \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
-As the eletron drifts towards the anode wire, it gains enough kinetic energy to knock out more electron-ion pairs in the gas which also go on to do the same; until they terminate on the anode wires or the pad plane.
-The amount of electrons produced depends on the anode wire voltage and the gas properties.
-The absolute gas gain was not experimentally measured but was calculated in a Garfield simulation.
+They are biased to high voltages of about  \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
+As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gain between collisions exceeds the ionization potential and ion atom collisions can therefore  knock out more electron-ion pairs in the gas.
+This amplifies the electron current.
+The final number of electrons produced depends on the anode wire voltage and the gas properties.
+The absolute gas gain was not experimentally determined but it was calculated in a Garfield simulation.
 For the experimental data pertaining to this thesis, the anode wires were biased to two different voltages.
 We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
-Only two sections were biased with the lower voltage setting due to a leakage around the end of the gating grid \cite{jon}.
+Two sections were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.
+
  Figure~\ref{fig:anodegain} shows the electron distribution for the total number of electrons produced in a avalanche process created by a single electron.
@@
 For the voltage of \SI{1460}{\volt} the parameters of the fit are $A_0=903.9$ and $A_1=1.50$ and for the voltage of \SI{1214}{\volt} the parameters are $A_0=150.0$ and $A_1=1.47$.
+These Polya distribution fits will be important later as the input to the Monte Carlo simulaitons of the detector response in Section~\ref{sec:monteCarlo}.
 
@@
 
+Care must be taken to choose the resister ($R_{TP2}$) that connects the field cage divider chain to the top plate.
+The end of the field cage nearest to the top plate is at the same potential as the rectangular metal top perimeter piece to which the field cage is glued.
+This top perimeter piece forms a horizontal surface that is an equipotential.
+It is separated from the top plate by an annular poly-carbonate insulator.
+The voltage difference between the average voltage on  the gating grid and the voltage of the top perimeter influences the electric field just below the gating grid.
+If $R_{TP2}$ is incorrectly chosen, the electric field near the gating grid will be different for electrons that pass near the field cage walls than it is for electrons in the center of the TPC,  causing distortions of the tracks in the TPC.
+This requires the value for $R_{TP2}$ to be chosen correctly.
+Since this simple point can be easily forgotten, we now discuss this choice in greater detail.
+
 The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the gating grid and the ground grid (Region 2), and the avalanche region is defined as between the ground grid and the anode grid.
-It is possible that the electric field in Region 1 and Region 2 can be matched but not required.
-To ensure the electric field is continuous across Region 1, the voltage of the top perimeter  must be set properly by adjusting the value of the last resistor ($R_{TP2}$).
-We can imagine Region 1 is split into two virtual volumes, one defined as the volume between the cathode and the top-perimeter, and one defined between the top-perimeter and the gating-grid.
+It is possible that the electric field in Region 1 and Region 2 can be matched but that is not essential.
+To ensure the electric field is the same across Region 1 for every point x and y value across the pad plane, the voltage of the top perimeter  must be set properly by adjusting the value of the last resistor $R_{TP2}$.
+This is not automatically correct because the gating grid average voltage is set independently of the cathode voltage.
+
+To understand how to calculate $R_{TP2}$ we imagine Region 1 to be split into two virtual volumes, one defined as the volume between the cathode and the top-perimeter, and one defined between the top-perimeter and the gating-grid.
 The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_1$, is defined as,
@@
 
-Several properties of the gas were also simulated in Garfield such as the longitudinal and transverse diffusion, $\sigma_l$ and $\sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}; summarized in Table~\ref{tb:gasprop}.
+Several properties of the gas were simulated in Garfield including the longitudinal and transverse diffusion, $\sigma_l$ and $\sigma_t$ respectively, and the electron and ion drift velocities, $v_d$ and $v_i$ respectively for the experimental electric field of \SI{125}{\volt\per\centi\metre}.
+The results are summarized in Table~\ref{tb:gasprop}.
 
@@
 
-Since we take the marginal distributions only along one layer or row of pad, correlations are introduced in the PRF from adjacent layers or rows which cause slight deviations from the expected Gatti distribution.
+Since we take the marginal distributions only along one layer or row of pads, correlations are introduced in the PRF from adjacent layers or rows which cause slight deviations from the expected Gatti distribution.
 Also, analytic PRFs only exist for classical multi-wire TPCs.
@@
 \end{equation}
+
+While the differences between the Gaussian distribution and the Gatti distribution are small for this TPC geometry, we use the Gaussian distribution because of its superior reproduction of the tails of the pad response function.
 
@@
 The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
-Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.
+Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines to which these beams can be sent.
+For this dissertation, the beam went through the BigRIPS spectrometer, where the specific rare isotope beams of interest were produced and on to the SAMURAI dipole where the  \spirit TPC was placed.
+The red box indicates the location of the experimental setup in Figure~\ref{fig:samuraiBeamLine} .
 
@@
 
-After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
+After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which fragments the beam into many different species.
 These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
@@
 The resulting secondary beam is not pure and the purity depends on the capability of BigRIPS to deliver the secondary beam of choice and the primary beam used.
+The BigRIPS separator had many scintillators for timing, an ion chamber for Z identification and beam tracking elements used to determine the magnetic rigidity of the beam.
+Information from these detectors allow  allowed  identification  of the mass, charge and momentum of each isotope in the secondary beam.
+This information was determined beam particle by beam particle and recorded along with the TPC data on each event.
+With this information, it was possible to select with precision the reactions to be included in any subsequent data analysis.
 
-In these set of experiments several beams were produced with varying intensities and purities.
-Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns where most beams were delivered with an intensity of \SI{10}{\kilo\hertz}.
-The BigRIPS separator had beam tracking elements which allowed for further identification of each individual beam produced.
-Allowing for a selection on the events corresponding to the secondary beam of choice CITE HERE.
+In the experimental campaign,   several beams were utilized that have  different intensities and purities.
+Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams that were used in our experimental campaign.
+Most beams were delivered with an intensity of \SI{10}{\kilo\hertz}.
 
@@
 Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration.
-Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
+Typically the SAMURAI (Superconducting Analyzer for Multi-particles from Radioisotope beams) dipole magnet is operated under vacuum as a large-acceptance multi-particle spectrometer for radioactive-beam experiments.
 This magnet can reach magnetic fields up to \SI{3}{\tesla} at the center of the pole gap and was operated at \SI{0.5}{\tesla} for these sets of experiments.
@@
 
-PUT HERE THE MAGNET FRAME COORDINATE 
-PUT HERE THE TPC FRAME COORDINATE DEFINITION
+The origin of the \spirit TPC was defined to be at the center of the TPC pad plane in the x-direction and at the edge of the last upstream pad.
+The center of the magnet is defined as the center of the magnet in x and z-directions and the middle of the dipole gap in the y-direction.
+The origin of the \spirit TPC in the magnet frame is (1.794,205.502,-580.526) for x,y,z-coordinates in units of \si{\milli\metre}.
+The error in each coordinate estimated to be (.2,.1,.4) \si{\milli\metre} respectively.
 
@@
 
+There were two beam drift chambers (BDC) located along the beam pipe upstream of the magnet and after the last focusing quadropole magnet.
+These beam drift chambers contained two sets of parallel plate avalanche counters (PPAC) which could get the (x,y) coordinates of the beam.
+Since they were places approximately \SI{1}{\metre} apart, we were able to track the direction of the beam in the beam line using a linear extrapolation.
+The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}.
+The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found.
+The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Sections~\ref{sec:beamangle,sec:pionSpectra}.
+
 \section{Ancillary Detectors }
 Several ancillary detectors were placed inside and outside of the \spirit TPC to facilitate in making the trigger for the experiment.
-Placing detectors around the TPC was one of the important considerations we made when designing the TPC.
-A brief description of each detector system is given here with particular focus on how the experimental trigger was made.
+The ability of ancillary detectors to work effectively when placed outside the TPC  was one of the more important considerations we made when designing the TPC.
+A brief description of each ancillary detector system is given here with particular focus on how the experimental trigger was made.
 
@@
 %scintillator bars
-As seen in Fig.~\ref{fig:aux}, the Kyoto Multiplicity Array consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
-The entire TPC structure was designed so that light charged particles could easily pass through the field cage and side walls of the TPC enclosure.
-In this way the number of tracks passing through the sides of the TPC could be measured by this array.
+ The Kyoto Multiplicity Array, shown in Fig.~\ref{fig:aux}, consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
+The entire TPC structure was designed so that light charged particles could pass through the field cage and side walls of the TPC enclosure without excessive energy loss and scattering.
+This allowed the number of tracks passing through the sides of the TPC could be measured by in an external array.
 In heavy ion collisions the more central a nuclear collision is, the more nucleons participate in the collision, resulting in a higher observed track multiplicity.
-It is this correlation between the number of tracks and centrality of the collision that makes the Kyoto Array sensitive to the centrality of events.
-It is more likely that in very central collisions more tracks are going to the peripheral angles and measured by the Kyoto array.
+It is this correlation between the number of tracks and centrality of the collision that makes the number of hits in the Kyoto Array sensitive to the centrality of events.
+It is more likely that in very central collisions more tracks are going to the large angles and measured by the Kyoto array.
 In the experiment the trigger selection criteria was $n_{Kyoto} > 4$, where $n_{Kyoto}$ is the total number of tracks measured by both arrays.
-The Kyoto array proved to be a good trigger for selecting on very central events as will be discussed in later sections.
+The Kyoto array proved to be a good trigger that suppressed peripheral collisions; this will be discussed in later sections.
 
@@
 \includegraphics[width=\textwidth]{TPCAux.png}
+\caption{Exploded views of Kyoto and KATANA arrays.The Kyoto array elements are in tan and are located on the left an right side of the TPC.
+When they are in place, they are nested against the thin aluminum side windows of the TPC.
+The KATANA array is in green and is located adjacent to the downstream window of the TPC.}
 \label{fig:aux}
-\caption{Exploded views of Kyoto and KATANA arrays.}
 \end{figure}
@@
 %beam trigger optional
-Also shown in Fig.~\ref{fig:aux} is the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted to the downstream wall of the TPC enclosure.
-Three of the 12 bars were thin and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this was a majority of the time.
-The 9 other bars operated as an additional multiplicity array similar to the Kyoto array.
-Since most of the particles are focused forward into a cone in the laboratory frame, it was found the condition on the Kyoto array was sufficient to trigger on central events; thus the KATANA array was used in primarily the beam veto mode.
+ Fig.~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure.
+Three of the 12 bars 1 mm in thickness and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time.
+The 9 other bars were 1 cm in thickness and provided additional information about the light particle multiplicity array similar to the Kyoto array.
+We found that a large number of events involved peripheral collisions and active target events in which the beam interactions with the counter gas.
+Active target events contributed to a large multiplicity in the KATANA but not to the Kyoto array.
+So multiplicity $M_{Kyoto} > 4$ in the Kyoto array combined with a veto on events with a forward going projectile remnant with Z>20 provided the most effective suppression of peripheral collisions and active target events.
+Thus the KATANA array was used in primarily the beam veto mode.
 This was accomplished by positioning the array so that the expected position of the beam exiting the TPC would be centered on the three thin paddles.
@@
 \subsection{Active Veto Array}
-The beam was tuned by two sets of quadropole  magnets, STQ 1 and STQ2 (as seen in Fig.~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
+The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Fig.~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
 Because of the inherent angular dispersion of the beam there still were incoming beam events which significantly deviated from the target location.
+These could be suppressed in the analyses by a gate on the reaction vertex; however, there remained a concern about the dead time such target frame events would contribute and how this would limit the amount of good data we could obtain.
 To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size.
 The threshold was set so that any beam particle which passed through any of the bars it would send a trigger signal to not trigger the system since the beam path would not be on target but on some other material inside the TPC.
+Minimizing these active veto rates also proved to be useful when tuning the beam so as to be centered on the target.
 
@@
 The Data AcQuisition (DAQ) consisted of three different systems.
-The RIBFDAQ system served as the master DAQ for the BigRIPS beam identificaiton DAQ, the TPC DAQ, the NeuLAND neutron wall DAQ, and the Kyoto Array DAQ systems.
+The RIBFDAQ system served as the master DAQ for the BigRIPS beam identification DAQ, the TPC DAQ, the NeuLAND neutron wall DAQ, and the Kyoto Array DAQ systems.
 The TPC DAQ was handled by the NARVAL framework to readout the GET electronics for the \spirit TPC.
@@
 
-There were several special trigger considerations when we built the trigger for the TPC.
-We required that the gating grid be opened fast to not miss any signal; as soon as there was a condition satisfying the Start Counter, Kyoto Multiplicity, the DAQ was not busy, and there was not a KATANA Veto signal the gating-grid would be opened.
-This was referred to as the Fast Trigger.
-If the KATANA trigger box is not satisfied --described later-- this will trigger a Fast Clear signal which will not trigger the DAQ, and will quickly close the gating grid.
+There were several special trigger considerations incorporated into the trigger for the TPC.
+We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid.
+If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicty >4 signal, and neither a Katana Veto signal nor an active target signal.
+This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns.
+Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC.
+If the last beam particle came within xx us after the fast trigger, electrons from this beam particle could pass through the gating grid.
+Such data would be useless, so we closed the gating grid and stopped the data equisition with a fast clear signal.
+Additional fast clears could also initiated if the Katana trigger box did not generate a trigger.
 Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.
 
+The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first.
+In later runs, most of the trigger functions were performed the in Katana trigger box.
 The master trigger for the DAQ was different for each primary beam as the experiment got progressively better.
 During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
-In either case the differences in the trigger were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}.
+In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}.
 Figure~\ref{fig:katanaLogic} summarizes the KATANA trigger box logic.
@@
 
- Where as Fig.~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
+ Fig.~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
 Here the KATANA trigger and the SC SUM???
@@
 
+ %I don't think this is true.
+
  It is worth mentioning how the busy signals for the experiment were handled.
-The DAQ system itself produces a busy signal which was combined with the busy signals from either the opening or closing the gating grid.
+The DAQ system itself produces a busy signal which was combined in an OR with the busy signals from either the opening or closing the gating grid.
 When opening the gating grid, it is assumed the full volume of the TPC will be read out, therefore  a \SI{11}{\micro\second} gate busy signal is produced; which is slightly more than the time it takes for all the electrons to drift in the field cage.
@@
 
-Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
-tb
+Figure~\ref{fig:data} shows the example pad plane readout of the TPC from a central nuclear collision of Sn + Sn.
+Here we can see several tracks resulting from a single vertex location which is near the target region.
+The color values represent the max ADC observed in a given pad.
+The large amount of saturation that occurs in the electronics which is represented by the brightest yellow value.
+Also the lower voltage anode sections described in Section~\ref{sec:wireplanes} are denoted by red arrows.
+The pads directly over these wire planes have an effective gain reduction due to the lower anode voltage on the wires.
+Tracks with a high enough energy loss continue through these sections, but tracks with low energy loss values do not deposit enough energy and disappear, only to reappear in the higher gain section in between.
 
-\begin{figure}[ht!]
-\includegraphics[width=\linewidth]{cocktail.png}
-\caption{Pad plane projection for a cocktail event in the TPC.}
-\label{fig:cocktail}
-\end{figure}
-
-\begin{figure}[ht!]
+\begin{figure}[!htb]
 \includegraphics[width=\linewidth]{data.png}
@@
 \end{figure}
+
+Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
+The primary goal of this data set was to calibrate the TPC in both momentum and dE/dx as will be discussed in Section~\ref{sec:cocktail}.
+
+\begin{figure}[!htb]
+\includegraphics[width=\linewidth]{cocktail.png}
+\caption{Pad plane projection for a cocktail event in the TPC.}
+\label{fig:cocktail}
+\end{figure}

### 2020-06-24 chapters/ch_intro.tex
Bill edits added Experiment section

@@
 Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon.
-Thus the outgoing $\pi$ is contains information on the high density region of the collision.
+Thus the outgoing $\pi$ contains information on the high density region of the collision.
 
@@
 
-The branching ratio of the various flavors of $\Delta$'s is given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}.
-Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$.
+The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}.
+Here we see that in general p-p collisions give rise primarily to  $\pi^+$ and n-n collisions give rise to primarily $\pi^-$.
 In this $\Delta$ resonance model the charged pion ratio can be described as,
@@
 \begin{equation}
-\frac{\pi^-}{\pi^+} = \frac{ 5pp + pn }{5nn + pn}.
+\frac{\pi^-}{\pi^+} = \frac{ 5N^2 + ZN }{5Z^2 + ZN}.
 \label{eq:deltaModel}
@@
 In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
-Pions can be reabsorbed into a $\Delta$ resonance after colliding with the another nucleon in the backward process of $\Delta \leftrightarrow \pi + N$.
-This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the asymmetry at the point of creation and re-emission.
-Total pion absorption back into two nucleons requires a three body process, where a pion is absorbed creating a $\Delta$ resonance, then another nucleon must collide with the resonance to create two nucleons.
-Because of this, the total pion absorption (removing pions from the system) is a smaller effect than the absorption re-emission process.
-In general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision, effectively turning the neutrons  measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally.
+Pions can be reabsorbed into a $\Delta$ resonance after colliding with another nucleon in the backward process of $\Delta \leftrightarrow \pi N$.
+This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the neutron-proton asymmetry at the point of creation and re-emission.
+Total pion absorption, back into two nucleons, requires a three body process; the pion collides with a nucleon creating a $\Delta$ resonance, then another nucleon must quickly collide with the resonance to create two nucleons.
+Since this is a three body process,h total pion absorption (removing them from the system) is a smaller effect than the absorption and re-emission process.
+Aside from these two effects which degrade the observable, in general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision; effectively turning the neutron measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally, and probing the high density region more than any other observable.
 
@@
 Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}.
-In these previous experiments, only total pion yields were published and no pion spectra were published, and the efficiency of the detector at detecting low energy pions was difficult to estimate.
-Also though the pion production increases with the beam energy, the symmetry energy affects the dynamics of the system less.
-This is because the time which the two nuclei coalesce in a HIC becomes shorter, and the time in which the symmetry energy can affect nucleons is shortened, thus the effect will becomes less dominant at higher beam energies.
-The goals of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs.
+Yet in these previous experiments, only total pion yields were published and no pion spectra was not published.
+It was also known that the efficiency analysis of the FOPI detector was particularly difficult especially for low energy pions.
+Although the pion production increases with the beam energy, the effects of the symmetry energy decrease.
+This is because higher beam velocities mean the time two nuclei spend together becomes shorter, shortening the time in which the symmetry energy can affect nucleons.
+The goal of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs.
 To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) Time Projection Chamber (TPC).

### 2020-06-26 chapters/ch_data_analysis.tex
Bill edits added Experiment section
Halfway through data analysis 1 backup

@@
 
-The S$\pi$RITROOT software is modular code based on the FAIRROOT package written in C++ \cite{fairroot}.
-The code is separated into modules which can be turned on or off depending on the users needs.
-The base classes and frame work is provided by the FAIRROOT package, while the specifics of each detector need to be added into the base classes.
-In the following an outline of the main tasks in the S$\pi$RITROOT software will be briefly discussed with the details of each task discussed later.
-The discussion of the software is really a practical, base understanding, appropriate for this thesis.
+The S$\pi$RITROOT software is modular code based on the FAIRROOT package, and is written in c++ \cite{fairroot}.
+The code is separated into modules which can be turned on or off depending on the users particular analysis.
+The base classes and frame work is provided by the FAIRROOT package, while the specifics of each detector need to be added.
+Here we will breifly describe the outline of the main tasks in the S$\pi$RITROOT software and then will discuss of each task later.
+The level of discussion concerning the software is really a practical, base understanding, appropriate for the discussion of this dissertation.
 For more details about the tracking algorithm the reader is refered to \cite{spiritroot_paper}, and to actual software which is published on GitHub \cite{spiritroot_git}.
@@
 \subsection{Decoder task}
-The decoder task reads in the raw data output from the GET electronics and separates the data from the headers using the GETDecoder \cite{getdecoder}.
-A map is stored mapping each channel in the CoBo board onto the physical pad plane, the information is stored in a pad container called the STPad class.
-This container stores the row, layer, and the raw ADC time bucket spectrum.
-All of the pads combined form the event container, STRawEvent class, which represents the raw event information.
-
-Three basic calibrations are performed at the raw data level within this task.
-The gain calibration calibrates the electronics will be described in Section\ref{sec:elecCalib} and the anode gain calibration is discussed in Section~\ref{sec:anodeCalib}.
-The gating grid noise subtraction and time bucket cut which is described in Section~\ref{sec:ggnoisesub}.
+The decoder task reads in the raw data output from the GET electronics converting the binary data into \spirit ROOT classes using the GETDecoder \cite{getdecoder}.
+When the software is loaded, a map is loaded which maps each channel in the CoBo board onto the physical pad location on the pad plane.
+The information is stored in a pad container called the \emph{STPad} class.
+This container stores the row, layer, and the raw ADC time bucket spectrum; where the row refers to the pad numbering along the x-direction and the layer refers to the numbering along the z-direction.
+All of the pads combined form the event container called the \emph{STRawEvent} class which represents the raw event information.
+
+Three basic calibrations are performed on the raw data level within this task.
+The gain calibration calibrates the electronics will be described in Section\ref{sec:elecCalib}.
+The anode gain calibration which will be discussed in Section~\ref{sec:anodeCalib}, and the gating grid noise subtraction and time bucket cut which is described in Section~\ref{sec:ggnoisesub}.
 %gain calibration 
@@
 The shaded histogram in every sub-figure represents the ADC raw spectrum of the pad.
-In Fig.~\ref{fig:psa0} we see the PSA has identified the first pulse and fitted it with the standard pulse, the result of the fit is shown by the red line.
+In Fig.~\ref{fig:psa0} we see the PSA has identified the first pulse and fitted it with the standard pulse; the result of the fit is shown by the red line.
 The fitted red line is then subtracted from the ADC spectrum, which is shown in the shaded histogram in Fig.~\ref{fig:psa1}.
-Here, the second pulse is found and fitted again with the same procedure, the red line will be subtracted from the raw ADC spectrum.
+Here, the second pulse is found and fitted again with the same procedure; the red line will be subtracted from the raw ADC spectrum.
 The black line represents the sum of all the fitted pules in each step.
@@
 
-The fit of each individual pulse contains the height of the pulse and the timing information of the start of the pulse, defined to be the location in time of 10\% of the rising edge.
-The height of the pulse, which is related to the charge induced on the pad, $Q$, and the timing information $t_o$, is saved into a new container representing a ``hit", called the STHit class.
+The fit of each individual pulse contains the height of the pulse and the timing information of the start of the pulse.
+We define the start of the pulse to be the location in time of 10\% of the rising edge.
+The height of the pulse is related to the charge induced on the pad, $Q$.
+Combined with the start time of the pulse $t_o$  it is saved into a new container representing a \emph{hit}, called the \emph{STHit} class.
 After the PSA processes each pad in the raw event, an entire hit cloud of the event is produced.
@@
 It is much easier to search for collection of hits by fitting a plane to the great circles which are formed on the Riemann sphere.
+More details on the Riemann tracking can be found here \cite{spiritroot_paper}.
 From here the set of hits which form a unique track are stored into a new helix track container called the STHelixTrack class.
 
-\paragraph{Definition of clustering}
+%\paragraph{Definition of clustering}
 The hits within a helix track are dense and the position resolution of an individual hit is not very precise.
@@
 The bold pads highlight the hits belonging to an example cluster from each set.
-Since we decide to cluster only in one direction, there is a natural inflection point in a track at which the direction of clustering must switch.
-This depends on the crossing angle of the track $\theta$, which is defined as the angle between the tangent of the track  at that cluster and the x-axis.
-In this definition $\theta = \ang{90}$ when the track is going along the z-axis and $\theta = \ang{0}$ for a track going along the x-axis.
+Since we decide to cluster only in one direction, there is a natural inflection point in a track where the direction of clustering must switch.
+This depends on the crossing angle of the track $\theta$, which is defined as the angle between the tangent of the track -- at that cluster -- and the x-axis.
+In this definition, $\theta = \ang{90}$ corresponds to the case when the track is going along the z-axis and $\theta = \ang{0}$ for a track going along the x-axis.
 The clustering direction in the case $\ang{45} < \theta \leq \ang{90} $ is along the x-direction.
-For $\ang{0} < \theta \leq \ang{45}$ we cluster along the z-axis.
+For tracks with angles $\ang{0} < \theta \leq \ang{45}$ we cluster along the z-axis.
 
@@
 Switching the clustering direction gives better position resolution than if we clustered only along one direction.
-You can imagine if we calculated the clusters only along the x-axis for tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
-The cluster position and charge is stored into a new container represented the cluster called the STHitCluster class.
+You can imagine if we calculated the clusters only along the x-axis, tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
+The cluster position and charge is stored into a new container representing the cluster called the STHitCluster class.
 All the clusters which belong to a particular helix track is also stored in the corresponding STHelixTrack class.
@@
 In this task the hit distribution, within a given cluster, is checked against the measured PRF discussed in Section~\ref{sec:prf}.
-Figure~\ref{fig:prfpimData} shows black lines around the measured PRF in the data, which form gates defining the acceptable PRF region.
+Figure~\ref{fig:prfpimData} shows black lines around the measured PRF in the data representing gates defining the acceptable PRF region.
 If less than 50\% of the hits within a cluster lie within this gate, the cluster is thrown out from the track.
-A majority of the clusters contains either 2 or 3 hits in total; in the case of 2 total hits, at least one of them needs to lie in the gate, in the case of 3 total hits, at least 2 hits need to lie within the gate.
-Clusters which do not follow this PRF gate usually have been corrupted.
-Typically charges from other nearby tracks is the typical way clusters become corrupt.
+A majority of the clusters contains either 2 or 3 hits in total.
+Clusters which do not follow this PRF gate usually have been corrupted in some way.
+Typically this occurs when charges from other nearby tracks mix distorting the hit distribution which corrupts the energy loss and position of the cluster.
+This tends to occur most frequently in the dense region surrounding the target, where the track density is high.
+But it may occur along the track for any other reason.
+The width of the gate is to allow for a reasonable range of PRF values to be accepted, without biasing the data too much.
 
@@
 The other correction task that is handled here is the space charge correction which will be discussed in more detail in Section~\ref{sec:spacecharge}.
+These two corrections represent the most significant corrections that greatly improve the data.
 
 \subsection{Momentum and track reconstruction (GENFIT)}
-After all of the corrections are applied to the clusters, the cluster's position and charge values are passed to a software package which reconstructs the momentum of the track, called GENFIT \cite{genfit}.
-The details of Kalman fitting algorithm used in the GENFIT reconstruction are left for the reader here \cite{genfit}.
-GENFIT returns a momentum and charge value for the track which is stored into a new track container called the STRecoTrack class.
-
-The task is repeated a second time, storing a second copy of the track class but this time with the BDC vertex point added as a constraining cluster.
-The momentum resolution is greatly improved by adding the vertex as a constraining point; since the BDC detectors, described in Section~\ref{sec:bdc}, give such an accurate vertex location.
+After all of the corrections are applied to the clusters, the cluster's position and charge values are then passed to a software package which reconstructs the momentum of the track called GENFIT.
+GENFIT utilizes a Kalman fitting algorithm which returns the estimated momentum and charge values for a given track \cite{genfit}.
+These values are then stored in into a new track container called the STRecoTrack class.
+
+This task is repeated a second time, but this time including the vertex as a constraining point in which all tracks are refitted.
+In this way, the momentum resolution is greatly improved by adding the vertex as a constraining point; since the BDC detectors, described in Section~\ref{sec:bdc}, give such an accurate vertex location.
 The energy loss of the track is calculated by the truncated mean method described in Eq.~\ref{eq:truncmean}.
-This is the final part of the analysis which pertains to tracks, and the PID can be constructed using the energy loss, momentum, and charge information.
+The GENFIT task represents the final part of the analysis which pertaining to track reconstruction.
+From here the PID can be constructed using the energy loss, momentum, and charge information.
 
@@
 After all tracks are reconstructed by GENFIT, the tracks are then passed to the RAVE software package which reconstructs the event vertex from a collection of tracks \cite{rave}.
-Only tracks which are THIS CONDITION are used in making the event vertex.
-The vertex location along the beam axis determines whether or not the beam collided with the target.
+The RAVE vertex is typically referred to as the TPC vertex in this dissertation.
+It is used primarily to find if an event is on target by observing the z-distribution as will be discussed in Section~\ref{sec:vertexcut}.
+The vertex which is used in refitting the tracks as mentioned in the above section comes from the BDC interpolated vertex which is more accurate than the TPC generated vertex.
 
@@
 \label{sec:ggnoisesub}
-The opening of the gating grid is essentially shorting adjacent wires in the grid together, allowing them to reach equilibrium as fast as possible.
-Since the impedance of both sides was not entirely matched properly this caused an imbalance of charge which bounce back and forth in the grid in an under-dampened manner.
-This oscillating charged induced  residual induced signals early in the time bucket spectrum creating extra source of noise in the data.
-Figure~\ref{fig:ggNoiseSub} shows in the upper panel the gating grid noise ADC time bucket spectrum for 2000 events in a given pad.
+Opening the gating grid is essentially done by shorting the adjacent wires in the grid together, and allowing them to reach equilibrium as fast as possible.
+Since the impedance of both sides was not entirely matched, this caused an imbalance of charge where one side discharged faster than the other side, causing charge to bounce back and forth in the grid in an under-dampened system.
+This create oscillating residual noise early in the time bucket spectrum.
+Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum of the gating grid noise for 2000 events in a given pad.
 The gating grid noise is stable for a given pad and the mean value --shown as the red line-- can be calculated after averaging over several thousands events.
-The raw gating grid noise lasts for 100 Tbs extending into the real data with a exponentially decreasing ampli
-
  %The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.
@@
 
-The bottom panel of Fig.~\ref{fig:ggNoiseSub} shows the gating grid noise self correcting itself; a best case scenario.
-We can see the pedestal, which is around 3800 ADC in the upper pannel,  is automatically taken into account when subtracting the mean profile, and the subtracted ADC spectra is centered around 0.
-The residual gating grid noise that was not canceled by the subtraction remains, but is significantly reduced to a much smaller time bucket region.
-Most of the residual noise happens in regions where the derivative of the gating grid noise is large.
+The bottom panel of Fig.~\ref{fig:ggNoiseSub} shows the gating grid noise self correcting itself, representing a best case scenario.
+The pedestal -- or base line ADC-- is around 3800 ADC in the upper panel, is automatically subtracted when subtracting the gating grid noise profile since the subtracted ADC spectra is centered around 0.
+Still some residual gating grid noise remains even after the subtraction, but it is significantly reduced and its extent is over a much smaller time bucket region.
+This failure happens in regions where the derivative of the gating grid noise is large.
 In these areas the ADC value is dependent upon small jitters in the timing of the signal and as a result varies wildly.
-The residual noise in the subtracted spectra is removed by cutting away all the bins below 30 tb.
-
-In the experiment the gating grid itself was quite stable in the  ${}^{132}$Sn and ${}^{124}$Sn systems and was not very stable in the ${}^{108}$Sn and ${}^{112}$Sn beams, where the gating grid broke several times.
-The gating grid noise was monitored and we regularly took gating grid noise profiles, especially after a gating grid board was replaced.
+To take care of the residual noise in the subtracted spectra we remove all bins below 30 Tb.
+
+In the experiment the gating grid itself was quite stable in the  ${}^{132}$Sn and ${}^{124}$Sn systems and was not stable in the ${}^{108}$Sn and ${}^{112}$Sn beams, where the it broke many times.
+Because of this, the gating grid noise was monitored regularly and we took gating grid noise profiles frequently, especially after a gating grid board was replaced.
 
@@
 
-A cocktail of several light charged particles was produced (p,d,t,${}^{3}$He,${}^{4}$He), and measured in the TPC to provide a momentum calibration of the TPC.
-The magnetic and slit settings of the dipoles in the BigRIPS spectrometer was set so that the momentum resolution of the beam was $\delta p/p$ < 1\%.
-Two magnetic rigidity settings were studied, with an empty target.
-A thick Aluminum target was also inserted to provide a slightly lower energy point for part of the lower rigidity setting, effectively creating three energy calibration points over several particle species.
-In some cases the certain particles were not able to propagate or were strongly attenuated.
+A cocktail of several light charged particles --at a known energy-- was produced (p,d,t,${}^{3}$He,${}^{4}$He) and measured in the TPC to provide a momentum calibration of the TPC.
+The magnetic and slit settings of the dipoles in the BigRIPS spectrometer were set so that the momentum resolution of the beam was known to within $\delta p/p$ < 1\%.
+Two magnetic rigidity settings were studied with no target in place.
+A thick Aluminum target was also inserted to provide a slightly lower energy calibration point for part of the lower rigidity setting, effectively creating three energy calibration points over several particle species.
+In some cases the certain particles were not able to propagate or were strongly attenuated and provided poor statistical measurements.
 
@@
 The momentum resolution in general depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
-This calibration beam represents and ideal situation where the track was parallel to the pad plane and only one particle was measured at one time.
-The average momentum resolution was measured to be $dp/p = 2\%$ over the full range of particle species for all settings.
-
-The energy loss resolution can also be directly measured since each energy setting represents a monochromatic source of each particle species, which have a well defined energy loss distribution.
+This calibration beam represents an ideal situation where the track was parallel to the pad plane and the track is not affected by the presence of other tracks.
+The average momentum resolution was measured to be $dp/p = 2\%$ over the full range of particle species for all settings as measured by the sigma of the Gaussian momentum distribution.
+
+The energy loss resolution can also be directly measured a monochromatic source of each particle species have a well defined energy loss distribution.
 The energy loss resolution was measured to be 5\% for all particles over the energy setting measured and is summarized in Table~\ref{tb:momresolution}.
@@
 The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
-The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed later in Section~\ref{sec:spacecharge} in a  general way which also includes the correction for the positive space charge.
-The same correction technique was applied here where the cocktail beam is the special case of zero space charge, only having $\vec{E}\times\vec{B}$ components.
-
-The values under the corrected column in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the corrected data.
+The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed later in Section~\ref{sec:spacecharge} in a more general way which also includes the correction for the presence of positive space charge in the field cage volume.
+The same correction technique was applied here, where the cocktail beam is the special case of zero space charge.
+
+The values under the corrected column in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data correcting for the $\vec{E}\times\vec{B}$ effect.
 A significant improvement is seen in the high momentum setting going from around ~5\% disagreement to within ~1\% agreement in the corrected data.
@@
 The electronics were calibrated by measuring the response of each channel to an input signal supplied by a pulse generator.
-This is a relative calibration technique with the intent to calibrate the varying gains in each channel relative to one another, and not an absolute calibration.
+This is a relative calibration technique with the intent to calibrate the varying gains in each channel relative to one another, not an absolute calibration.
 The pulse was distributed to all the electronics channels by pulsing the ground plane for a range of input voltages.
@@
 The small variation in each channel can be seen as the wide band around each measurement point.
-A linear fit is performed for each channel to get the calibration line for each channel.
-A channel is chosen which provides a reference calibration to which all other channels are calibrated too.
-This is done by matching the slopes of each channel to the reference calibration slope.
-The right panel shows the resulting distribution of channels after calibration.
+A linear fit is performed to get the calibration line for each channel.
+A particular channel is chosen which provides a reference calibration to which the slopes of all other channels are calibrated too.
+The right panel shows the resulting distribution of channels after calibration, in which the channel variation is reduced significantly.
 
@@
 As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires.
-Out of all the runs used in the analysis in this thesis \ref{tb:runList}, the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all the other runs.
+Out of all the runs used in the analysis in this thesis -- see \ref{tb:runList} -- the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all other runs.
 By lowering the voltage on these anode wires, the gas gain is lowered as compared with all the other anode wire plane sections which operate at \SI{1460}{\volt}.
-To account for the drop in gain, we increase the gain of the pads which lie above these anode wires  in the software gain of the other channels.
-The gain factor is found by plotting the energy loss in the high gain sections to the low gain sections.
+To account for the drop in gain, we artificially increase the gain of the pads which lie above these anode wires in the software.
+The multiplicative factor is estimated by plotting the energy loss in the high gain sections relative to the low gain sections.
 
@@
 
- A map of the channels which are affected by the low gain and the relevant gain factor was input into the software.
-In the decoder task the raw ADC spectra was multiplied by the input gain for a given channel.
-The thresholds of those channels also were multiplied by the same gain factor since the noise levels were also artificially magnified by this method.
+ A map defining the gain calibration of each pad is pre-loaded into the software.
+In the decoder task the raw ADC spectra was multiplied by this gain factor for each channel.
+The PSA thresholds of those channels also were multiplied by the same gain factor since the noise levels were also artificially magnified as well.
 
@@
 
-Using a TPC for measurements of HIC in nuclear physics presents a different set of challenges as opposed to higher energy experiments.
+Using a TPC for measurements of HICs in nuclear physics presents a different set of challenges as opposed to higher energy experiments.
 Typically in higher energy experiments fundamental particles are produced with charge $\pm e$.
-Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum in the energy loss curve.
-The dynamic range of most TPC electronics can cover a wide range of particles in these types of experiments.
+Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum ionization in the energy loss.
+In these cases the dynamic range of most TPC electronics can cover a wide range of particles.
 In nuclear HICs, we are interested in measured particles with charges $Ze$ where Z=1-3, and even higher in some applications.
-The energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$.
-Even more, low energy nuclear HIC of intermediate beam energies (around \SI{300}{\MeVA}) produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
-The dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases where the electronics will be saturated.
+Recall the energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$.
+Even more, in low energy nuclear HIC of intermediate beam energies -- around \SI{300}{\MeVA} -- produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
+In this case, the limited dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases, where they saturate the electronics.
 
@@
 This mitigates the complete loss of information but introduces a new problem.
-Though particles which deposit large amounts of charge will now have good measurements in the low gain areas, particles which deposit minimal energy losses will lose information in the same low gain areas.
+Particles which deposit large amounts of charge will have good measurements in the low gain areas, where as particles depositing minimal energy losses will lose information in the same low gain areas.
 The reconstruction of such tracks will suffer.
 There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics CITE HERE.
-Never the less it is quite useful to develop a software technique which may extend the dynamic range of TPC electronics without the use of new external hardware, especially for experiments which have already been performed with older electronics technologies.
+Never the less, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware.
+Which can be particularly useful for experiments which have already taken data with older electronics technologies.
 In this section we will outline a novel software technique that takes advantage of the PRF described in Section~\ref{sec:prf} and can effectively extend the dynamic range of the TPC electronics.
@@
 In TPCs, the effective dynamic range is very different from the single channel dynamic range depending on how the TPC measurement is used.
-Typically TPCs are operated inside of a magnetic field for the purpose of reconstructing the momentum of a track, which requires sub-millimeter precision in the position determination of the track path.
+Typically TPCs are operated inside of a magnetic field for the purpose of reconstructing the momentum of a track, which requires sub-millimeter precision in the position determination along the track path.
 This is achieved by clustering several pads together as discussed in Section~\ref{sec:helixtrack}.
@@
 
-The effective dynamic range is then not the single channel dynamic range, but the dynamic range between central pad --holding the largest charge-- and the adjacent pads --holding the smaller charges in the PRF distribution.
-For example, to measure minimum ionizing particles, the Signal-to-Noise-Ratio (SNR) of the pad with the smallest charge in the distribution should be some reasonable value above noise, so that the signals can be measured.
-From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
+In this case the effective dynamic range is not the single channel dynamic range, but the relative dynamic range between central pad --holding the largest charge-- and the adjacent pads --holding the smaller charges in the PRF distribution.
+For example, to measure minimum ionizing particles, the Signal-to-Noise-Ratio (SNR) of the pad with the smallest charge in the distribution should be some reasonable value above noise, so that the full charge distribution can be measured.
+From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds about 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
 In the \spirit TPC the electronics gain was set so that the the pads have a SNR of 6:1 for minimum ionizing tracks, and therefore the central pad had a SNR of 50:1.
-In the case of the maximum charge is collected in the central pad, before the electronics saturate, the SNR is 800:1.
+The maximum SNR in the central pad -- before the electronics saturate -- the SNR is 800:1.
 Therefore the maximum SNR is roughly 16 times that  --800 over 50 -- of minimum ionizing particles.
-This is order to have a complete measurement, where adjacent pads are all measured and position resolution is not compromised.
-
-Figure~\ref{fig:intro} shows the theoretical energy loss curves for several particles as a function of the momentum over charge.
+
+Figure~\ref{fig:intro} shows the theoretical energy loss curves for several particles as a function of rigidity $p/q$.
 Minimum ionization can be seen to take place around \SI{10}{\kilo\electronvolt\per\centi\metre}.
@@
 \label{sec:expPRF}
-The pad response function of the TPC was extracted from non-saturating hits and clusters in tracks at various track crossing angles.
-As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as $\lambda_i = x_i-\bar{x}$.
+As mentioned in Section~\ref{sec:prf}, the PRF depends on the angle of the track as it crosses the wires and pads.
+The PRF of the TPC was experimentally determined from non-saturating hits and clusters in tracks at various track crossing angles.
+As in Fig.~\ref{fig:topview}, we postulate that the PRF is a function of the total charge deposited in a cluster $Q = \sum_i q_i$, and the difference in position of the center of the $i^{th}$ pad, $x_i$, to the mean position $\bar{x} = \sum_i x_i q_i/Q$, defined as,
+
+\begin{equation}
+\lambda_i = x_i-\bar{x}. 
+\label{eq:lambda}
+\end{equation}
+
 The PRF is simply defined as the charge fraction of each pad as a function of $\lambda$, as shown in Equation \ref{eq:prf}.
@@
 
-Averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
-Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$.
+Averaging over many events in the experimental data, the resulting PRF is particularly well behaved for the S$\pi$RIT TPC as seen in Fig.~\ref{fig:expprf}.
+Here we see the deviations from the expected analytic Gatti distribution (black curve).
+Fitting with a two parameter Gaussian function -- the red curve -- describes the data better.
+The PRF was fit with a two parameters Gaussian function, where $N_)$ is the normalization coefficient and $\sigma$ the corresponding width:
 
 \begin{equation}\label{eq:gaus}
-PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}
+PRF_{\mathrm{Gaus}}(\lambda) = N_0 e^\frac{-\lambda^2}{2\sigma^2}.
 \end{equation}
@@
 Figure~\ref{fig:prfpimData} shows the PRF for $\pi^-$ tracks versus the crossing angle $\theta$ of the track.
-The PRF gets wider starting from $90^{\circ}$  and until where we switch clustering directions at $\ang{45}$.
+The PRF gets wider starting from $90^{\circ}$  until where the clustering switches directions at $\ang{45}$.
 If we did not switch clustering directions the PRF would become wider until it was a uniform distribution and there was no position resolution.
-Switching from $x$ to the $z$ direction clustering the opposite trend is seen where the PRF becomes narrower  going from $45^{\circ}$ to $0^{\circ}$, as the position resolution gets better.
+Switching shows the opposite trend where the PRF becomes narrower going from $45^{\circ}$ to $0^{\circ}$, as the position resolution gets better.
 
@@
 
-A Gaussian fit was performed to the experimental PRF in $\ang{5}$ width bins ranging from $\ang{0} < \theta \leq \ang{90}$.
-Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function given in Eq.~\ref{eq:prfgaus} which are are plotted versus $\theta$.
+The Gaussian fit was performed on each PRF for a crossing angle steps of  $\ang{5}$ ranging from $\ang{0} < \theta \leq \ang{90}$.
+Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function -- Eq.~\ref{eq:prfgaus} -- which are are plotted versus $\theta$.
 A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value, which is shown as the black line.
+Once we have this interpolation, we can predict the PRF for any given value of $\theta$.
 
@@
 
-In the following, we will use the term ``desaturation'' to describe the technique described here of correcting the charge values of the saturated pads.
+In the following, we will use the term \emph{desaturation} to describe the technique of correcting saturated pads.
 Figure~\ref{fig:satpad} shows a typical situation of saturated hits in a cluster.
@@
 Pads further away collect less charge and typically are not saturated, detonated here as $q_{1}$ and $q_{4}$.
-Though the charge values in the saturated channels are lost, we know that the distribution of all charges must follow the PRF which is a fixed function, and a fundamental operating principle of all TPCs.
+Although the charge values in the saturated channels are lost, we know that the distribution of all charges must follow the PRF; this is fundamental operating principle of all TPCs.
 We have already measured the PRF as a function of crossing angle in Section~\ref{sec:expPRF}, and from the tracking information, we know the crossing angle of the track at a given cluster and therefore the PRF corresponding to that cluster.
 
-We assume the distance of each pad to the track, $\lambda_i$, is fixed, defining the fraction of charge each pad receives as defined by the $PRF(\lambda_i)$ function.
+We assume the distance of each pad to fitted track, $\lambda_i$ , Eq.~\ref{eq:lambda}, is fixed, defining the fraction of charge each pad receives as defined by the $PRF(\lambda_i)$ function.
 To determine the best estimate for the charge values of each saturated pad, a chi squared function is minimized,
@@
 
-where $q_i^{\mathrm{obs}}$ are the non-saturated charges and $q_i^{\mathrm{expect}}$ are the charge values expected for that pad as calculated from the PRF; $q_i^{\mathrm{expect}} = Q\cdot PRF(\lambda_i)$.
-The saturated charge values $q_{i}^{\textprime}$ are treated as unknown variables and are allowed to vary in the $\chi^2$ minimization, they enter the minimization routine when they are added to get the total charge $Q = \sum q_i + \sum q_i^{'}$.
+where $q_i^{\mathrm{obs}}$ are the non-saturated charges and $q_i^{\mathrm{expect}}$ are the charge values expected for that pad as calculated from the PRF, $q_i^{\mathrm{expect}} = Q\cdot PRF(\lambda_i)$.
+The saturated charge values $q_{i}^{\textprime}$ are treated as unknown variables and are allowed to vary in the $\chi^2$ minimization routine, they enter the calculation when they are added to get the total charge $Q = \sum q_i + \sum q_i^{'}$.
+The minimum $\chi^2$ value returns the best estimate for the unknown saturated charge values.
+The saturated hit charges are updated to reflect these estimates and the cluster position and charge is updated accordingly.
 
@@
 
-Tracks which saturate pads in the high anode wire voltage region are not saturated in the low anode voltage region.
-By comparing the $\langle dE/dx\rangle$ values of these two sections.
+To provide evidence of the success of this technique, we observe tracks which saturate pads in the high anode wire voltage region are not saturated in the low anode voltage region.
 While this desaturation technique avoids the need to lower the gain of any region, the low anode voltage region proved to be a direct measurement of the success of this technique.
@@
 Figure~\ref{fig:lowvshigh} shows the the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
-The effect of saturation can be seen in the high gain region for the uncorrected data above values of 400 ADC/mm where the values plateau, where as the low gain region still returns accurate values.
+The effect of saturation can be seen in the high gain region for the uncorrected data above values of \SI{400}{ADC \per \milli\metre} where the values plateau, where as the low gain region still returns accurate values.
 Below this value the electronics are not saturated, and therefore the high and low gain sections agree.
 After applying the desaturation method, the correlation between the high and low gain sections is restored, as seen in Fig.~\ref{fig:lowvshigh}.
-From this comparison, we can approximate that the correction has corrected the high gain sections to agree with low gain sections to values of 2000 ADC/mm, increasing the dynamic range by a factor of at least 5.
-
-The success of the desaturation becomes more clear when looking at the particle identification (PID) lines from the experimental data in Fig.~\ref{fig:data_combine}.
+From this comparison, we can approximate that the correction has corrected the high gain sections to agree with low gain sections to values of \SI{2000}{ADC \per \milli\metre}, increasing the dynamic range by a factor of a factor of at least 5.
+
+The success of the desaturation becomes more clear when looking at the PID lines from the experimental data in Fig.~\ref{fig:data_combine}.
 In the following PID plots the red lines represent the most probable energy loss as given by Geant4 straggling functions.
-The uncorrected data in the left panel shows the effects of saturation, where the PID lines deviate significantly from their theoretical expectations starting at around 400~ADC/mm.
+The uncorrected data in the left panel shows the effects of saturation, where the PID lines deviate significantly from their theoretical expectations starting at around \SI{400}{ADC \per \milli\metre}.
 After applying the desaturation technique -- in the right panel-- we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation.
-Even the ${}^{6}Li$ and ${}^{7}Li$ particles can be separated and a more subtle improvement of the lighter particles, (p, d, t), can also be seen in the PID lines at lower momenta.
+Even the ${}^{6}Li$ and ${}^{7}Li$ particles can be separated and a more subtle improvement of the lighter particles (p, d, t), can also be seen at lower momenta.
 In these regions, there was little to no PID resolution before desaturation technique was applied.
@@
 
-As the beam passes through the field cage it ionizes the gas creating electron-ion pairs.
+As the particles passes through the gas inside of the field cage, they ionize the gas creating electron-ion pairs.
 The drift velocities of the ions are typically \SI{e4} times slower than electron drift velocities \cite{blumrol}.
-Any source of ions have the potential to build up in the drift volume, creating a positive space charge, distorting the drifting electrons and  biasing the track momentum measurement.
+Because of this, any source of ions have the potential to build up in the drift volume, creating a positive space charge.
+If the space charge build up is large enough, it would distort drifting electrons therefore biasing the track momentum measurement.
 There are several regions of the TPC in which ions are created.
@@
 
-NEED TO PUT IN ABOUT THE GATING GRID LEAK
+%NEED TO PUT IN ABOUT THE GATING GRID LEAK
 
@@
 The average distance between sequential ion paths is about \SI{25}{\micro\metre} apart, therefore we expect the average number of beam paths that make up the sheet charge is around \num{1440} tracks.
-Since the number of tracks in the sheet charge is large, and inter beam spacing is very small, this allows us to approximate the sheet charge as an uniform sheet charge density.
-
- The secondary beams entering the TPC are also composed of many species of particles distributed in a finite range around the target beam as will be discussed in Section~\ref{sec:beam}.
-Let us assume a beam of ${}^{132}$Sn where the energy loss in P10 gas, at 270 AMeV, is \SI{11.2}{\kilo\electronvolt\per\centi\metre}.
+Since the number of tracks in the sheet charge is large, and inter beam spacing is very small, we approximate the sheet charge as a uniform charge density.
+
+ The secondary beams entering the TPC are also composed of many secondary species which are impurities as will be discussed in Section~\ref{sec:beam}.
+Their charge values are mostly distributed around Z~50.
+Let us assume a beam of ${}^{132}$Sn where the energy loss in P10 gas, at \SI{270}{\AMeV}, is \SI{11.2}{\kilo\electronvolt\per\centi\metre}.
 From the total number of beams above, and the beam path length of \SI{135}{\centi\metre}, the estimated charge density would be on the order of \SI{3e-8}{\coulomb\per\metre\squared}.
-
-The electric potential can be calculated by solving Poisson's equation,
+This gives us an understanding of the amount of space charge we should expect to see in the chamber.
+
+Once the amount of charge is known, the electric potential can be calculated by solving Poisson's equation,
 
@@
  where $\phi$ is the electric potential and $\rho$ is the free charge.
-The numerical solution is provided by the Jacobi method \cite{poisson} provided all 6 sides have defined Dirchlet boundary conditions for the potentials.
-We neglect the wire plane region since the drift details are not important for the space charge effects we are discussing here.
-The pad plane and cathode are trivial, where the side walls of the field cage are given linearly varying potentials on the surface such that the electric field is the same strength as the real TPC with no space charge present.
+The numerical solution is provided by the Jacobi method \cite{poisson}, provided all 6 sides have defined Dirchlet boundary conditions for the potentials.
+We neglect the wire plane region since the drift details around it are not significant for the space charge effects we are discussing here.
+The pad plane and cathode are trivial, where the side walls of the field cage are given linearly varying potentials on the surface, such that the electric field is the same strength as the real TPC with no space charge present.
 Once the electric potential is solved, the electric field is simply the gradient of the potential  $\vec{E}= -\nabla \phi$.
@@
 In the presence of no space charge, the dOCA distribution of each track would be a centered around the true vertex location.
-Since the tracks are distorted by the space charge, which affects different regions of the TPC differently, a bias is introduced to the measured vertex location and widens the distribution to vertex of each track.
-
-An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} where left-going tracks shown in blue, and right-going tracks shown in green.
-The vectors show the direction of distortion, and their magnitude have been magnified by 10 times to show the detail.
-The dashed lines show the shifted track due to the space charge where left and right-going tracks are affected differently.
+Because of the left right symmetry in the TPC, the space charge affects different regions of the TPC differently, and a bias is introduced to the measured vertex location and widening the distribution to vertex of each track.
+
+An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} where left-going tracks are shown in blue, and right-going tracks shown in green.
+The vectors show the direction of electron shift, and their magnitude have been magnified by 10 CITE HERE times to show the detail.
+The dashed lines show the tracks under the influence of the space charge where left and right-going tracks are affected differently.
 Right-going tracks tend to higher momentum values and the left-going tracks going to lower momentum values, for positively charged particles.
@@
 
-The inset figure of Fig.~\ref{fig:sc_shift} shows the effects of the space charge have on the x-component of the dOCA to the vertex for the displaced left-going track given by $\Delta\mathrm{V}_\mathrm{x}$, with the opposite direction for right-going tracks.
-Figure~\ref{fig:VLR} shows the amount of distortion caused by the space charge effect as a new observable $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ , where  $V_x^L$ and $V_x^R$ are the most probable value of the dOCA distribution for left and right-going tracks respectively.
+The inset figure of Fig.~\ref{fig:sc_shift} shows the effects of the space charge have on the x-component of the dOCA to the vertex.
+The dOCA distribution is displaced left-going track labeled by $\Delta\mathrm{V}_\mathrm{x}$, and in the opposite direction for right-going tracks.
+Figure~\ref{fig:VLR} shows the difference between right and left tracks as $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ , where  $V_x^L$ and $V_x^R$ are the most probable values of the dOCA distribution for left and right-going tracks respectively.
+As the space charge increases this value increases, where in the presence of no space charge, we would expect exactly $\Delta\mathrm{V}_\mathrm{LR} = 0$.
 
@@
 
-The average beam rate was recorded in each experimental run and slightly varied from run to run, due to beam production variations.
-The amount of space charge present in the field cage is directly proportional to the beam rate; therefore $\Delta\mathrm{V}_\mathrm{LR}$ is also proportional to the beam rate as shown in Fig.~\ref{fig:VLR_br}.
+The average beam rate was recorded in each experimental run and slightly varied from run to run due to beam production variations.
+The amount of space charge present in the field cage is directly proportional to the beam rate and therefore $\Delta\mathrm{V}_\mathrm{LR}$.
+Fig.~\ref{fig:VLR_br} shows the relation between the measured $\Delta\mathrm{V}_\mathrm{LR}$ values and the estimated beam rate.
+
 The only parameter in the space charge correction algorithm is the surface charge density $\sigma_{\mathrm{SC}}$.
-Varying $\sigma_{\mathrm{SC}}$ for a wide range of values the $\Delta\mathrm{V}_\mathrm{LR}$ observable is measured and plotted in the left panel of Fig.~\ref{fig:spacechg_relation}.
-The surface charge density which gives $\Delta\mathrm{V}_\mathrm{LR} = 0$ is taken to be the estimate for the average amount of space charge present.
-
-This is done for several runs which vary in beam intensity though the solution for the estimated space charge value will be different.
+By varying $\sigma_{\mathrm{SC}}$ for a wide range of values, the $\Delta\mathrm{V}_\mathrm{LR}$ observable is measured and plotted in the left panel of Fig.~\ref{fig:spacechg_relation}.
+The values are linearly fit to find the solution of the surface charge density which gives $\Delta\mathrm{V}_\mathrm{LR} = 0$, which is then taken to be the estimate for the average amount of space charge present in a given run.
+
+This is done for several runs which vary in beam intensity.
 Since the surface charge density is proportional to the beam rate, a linear fit gives good agreement for interpolating the surface charge values as a function of beam rate.
@@
 
-Following this algorithm the space charge is estimate for each run and system.
+Following this algorithm, the space charge can be estimated for each run and system.
 Figure~\ref{fig:scDensity} shows the summary of the extracted space charge values for each secondary beam.
 The linear fits relating the space charge density to the beam rate for each system is shown in Fig.~\ref{fig:spacechg_br_all} where the space charge value is inferred from the measured beam rate.
-Notice that in the $\tin{112}{124}$ system a constant value of was assumed since the beam rate did not span a large enough range to warrant a more detailed analysis.
-
-To reduce the need to compute a new electric field for each space charge value, notice that $\vec{E}\propto \rho$, we can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ by the ratio $\rho/\rho_o$.
+Notice that in the $\tin{112}{124}$ system is a constant value, since the variation in $\Delta\mathrm{V}_\mathrm{LR}$ rate did not span a large enough range to warrant a more detailed analysis.
+
+To reduce the need to compute a new electric field for each space charge value, notice that $\vec{E}\propto \rho$.
+We can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ linearly by the ratio $\rho/\rho_o$.
 The full magnetic field map is provided by the SAMURAI collaboration \cite{magnet}.
-The velocity field map is calculated following Eq.~\ref{eq:elecdrift}.
-The electron drift through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration from a certain starting point.
-
- The correction map is calculated by starting from the anode y-position and the measured position on the pad-plane (x,z), and stepping backward in time in the Runge-Kutta integration, through the velocity field map until the electron reaches the measured y-position.
-This is done over a 3-dimensional grid where points in between are interpolated by a trilinear interpolation.
+The velocity field map is calculated following Eq.~\ref{eq:elecdrift}, where the electron drift through this velocity map is propagated by using a time stepped $\mathrm{4}^{\mathrm{th}}$-Order Runge-Kutta integration from a certain starting point.
+
+ The correction map is calculated in the inverse method, starting from the anode wires and location on the pad-plane (x,z), and stepping backward in time in the Runge-Kutta integration through the velocity field map until the electron reaches the measured y-position.
+This is done over a 3-dimensional grid where points in between are interpolated by a tri-linear interpolation.
 The measured clusters value (x,y,z) position is input into the correction map which outputs the interpolated correction values $dx$ and $dz$ which correspond to that cluster.
@@
 \section{Monte Carlo Simulation}
+\label{sec:monteCarlo}
+
 The MC simulation is composed of two separate simulations.

### 2020-06-26 chapters/ch_experiment.tex
Halfway through data analysis 1 backup

@@
 %explain in breif the basics of a time projection chamber 
-Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions.
-Here we will outline the physical principles involved in the TPC measurement and discuss the particular details of the TPC used in this thesis.
-Figure~\ref{fig:tpcPrinciple} depicts the inside of a TPC, specifically the field cage, which holds the detector gas and also sets up a constant electric field.
-As charged particles resulting from a heavy ion collision pass through the gas, electron-ion pairs are created from neutral gas molecules.
-The electrons are accelerated opposite to the electric field and the ions along with it.
-Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules; slowing down, or even stopping the electron, which is then accelerated again, starting a repeating cycle of stop and go motion.
-This microscopic behavior manifests as a constant drift velocity when averaged over several gas collisions.
-The electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they quickly accelerate in the presence of the high electric field, liberating more electron-ion pairs from the gas, creating an avalanche process to occur.
-The avalanche electrons eventually terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires, creating a large signal which is distributed over the pad-plane.
-The charge and timing  information of these electrons are calculated from the induced signal measured by the  electronics.
-
 \begin{figure}[!htb]
@@
 
-Two of the 3 coordinates are determined from the 2-dimensional charge distribution on the pad-plane.
-The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$.
-The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the timing information of the signal in the electronics.
-The radius of curvature of the track is related to the magnetic rigidity, and therefore the momentum of each track.
-The energy loss deposited $\langle dE/dx\rangle$ is measured by the segmented charge sensitive pads on the pad-plane, in which each pad is connected to a channel in the electronics.
-Particle can be uniquely identified since  each particles exists on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections.
-In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.
+%explain in breif the basics of a time projection chamber 
+A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector.
+TPCs have been employed in gases, liquid, and even solid detectors CITE HERE.
+In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis.
+Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage.
+The field cage holds the detector gas and also defines a constant electric field that is typically aligned with the magnetic field of a magnet that surrounds the TPC.
+In this Spirit TPC, both these electric and magnetic fields are vertical.
+
+As a charged particle resulting from a heavy ion collision passes through the field cage, energetic \emph{primary} electrons are knocked out of neutral gas molecules and atoms in the detector gas.
+These primary electrons further ionize the detector gas until there is a distribution of \emph{secondary} electrons and positive ions that lie along the track's path.
+The electric field of the field cage separates these electrons and ions, where the electrons are accelerated upward by the electric field in the direction against the electric field, and the positive ions are accelerated along the electric field.
+Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules, which slow down, or even stop the electron; only to be re-accelerated by the electric field, continuing a repeating cycle of stop and go motion.
+This microscopic behavior results in a constant drift velocity when averaged over multiple gas collisions.
+
+The electrons drift with this drift velocity through a set of wire planes, eventually reaching a set of high voltage anode wires, where they quickly accelerate in the presence of the high electric field emanating from the anode wires.
+This high electric field imparts enough energy between collisions to ionize the gas, creating additional electron-ion pairs from the gas, culminating in an avalanche process near the anode wires.
+The avalanche electrons typically terminate on the anode wire,  while the ions created in the avalanche are accelerated  away from the anode wires.
+This ionic motion induces image charges in nearby pads within the pad plane that is located close to the anode wires.
+The corresponding image currents in these pads are read out by electronics that are attached to these pads.
+This electronics signal encodes the energy loss information and arrival time of the signal from the electrons drifting from the ion track, and can be analyzed to determine the position of the ion track within the TPC.
+
+The pad plane is perpendicular to the electric field of the field cage and is divided into a two dimensional grid of pads.
+Two of the 3 coordinates are determined from the 2-dimensional charge distribution corresponding to the projection of the track onto the pad-plane.
+The third dimension comes from extrapolating the electrons back in time, utilizing the known constant drift velocity $v_d$.
+The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the drift time information obtained from combining the time of the event trigger with arrival time of the signal from the  electrons in the electronics.
+The radius of curvature of the track on the pad plane is related to the magnetic rigidity, and therefore the momentum of each track.
+The energy loss deposited $\langle dE/dx\rangle$ is proportional to the image charge induced on the pads on the pad-plane.
+With these two observations, particles can be uniquely identified by their mass and charge since there is a unique correlation between the rigidity and $\langle dE/dx\rangle$  lines.
+This correlation allows the particle type to be identified as discussed later in Section~\ref{sec:pid}.
+In this chapter we will provide more details about the particle detection within the \spirit TPC  that is relevant to the reactions studied in this thesis.
 
@@
 
-The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in fixed target experiments.
+The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) was developed to measure pions and other light charge particles produced in collisions of radioactive rare isotope beams with stable targets in fixed target experiments.
 The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
@@
 
-A target mechanism allowed for up to 5 fixed targets to be mounted at anytime, with the ability to change targets on the outside of the TPC.
-The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal.
-The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate.
-Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurment \cite{jon}.
-Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics.
-The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC.
+The exploded drawing shown in Fig.~\ref{fig:tpcExplode} illustrates these major internal components of the  of the \spirit TPC.
+A target mechanism allowed for up to 5 fixed targets to be mounted on the target ladder situated just upstream of the field cage, where changing the targets could be performed outside of the magnet.
+Reaction products from nuclear collisions exit the target and enter the field cage of the TPC.
+The field cage contains the detector gas and defines the constant electric field.
+It is rigidly mounted to the large aluminum top plate of the TPC gas containent vessel.
+The field cage is electrically isolated from the top plate by a lexan top perimeter ring with o-rings to provide as gas seal.
+The pad plane and wire plane structures are  mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate.
+Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurement \cite{jon}.
+Holes in the top plate allowed signals from the individual charge sensitive pads on the pad plane to be transmitted via cables to the electronics.
 
@@
 \subsection{Enclosure}
-The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame.
-All walls are constructed of a aluminum frame with thin sheet metal.
-All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much.
-This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC.
-The enclosure itself is made to be gas tight with respects to the outside and the field cage.
-This is to allow for the possibility to run a different gas inside the enclosure than the field cage.
-Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.
+The skeleton of the enclosure is composed of a rigid rectangular aluminum angle-iron frame.
+To this frame, six rectangular sides are attached.
+The downstream window and the two side windows are constructed of a aluminum window frames with a 1 mm  aluminum sheet metal window.
+These thin sheet metal windows allow charged particles and neutrons to exit the TPC without significant energy loss or scattering.
+This allows for a trigger to be created by placing detectors outside of  the side windows and downstream window of the TPC.
+The enclosure itself is made to be gas tight with no leakage from the outside or from the interior of the field cage.
+This is to allow for the possibility to choosing one insulation gas outside of the field cage and a different counter gas inside the field cage.
+In the first campaign of experiments, however, we used the same gas in the field cage and enclosure volume.
 \end{comment}
@@
 \subsection{Field Cage}
-
-The field cage contains the detector gas and sets up a uniform electric field in which electrons can drift upwards toward the anode wires.
-It was designed to hang from the top plate and therefore needed to be of a lightweight construction.
-The materials needed to be thin to allow for light charged particle and neutrons to pass through without significant scattering for ancillary detectors.
+%The field cage contains the detector gas and defines a uniform electric field in which electrons can drift upwards toward the anode wires.
+It was designed to hang from the top plate and therefore needed to be of a lightweight construction and the materials needed to be thin to allow for light charged particle and neutrons to pass through to ancillary trigger detectors without significant energy loss or scattering .
 
@@
 
-The field cage was constructed from several panels of printed circuit boards (PCBs).
-The front of the field cage was made of two PCBs and each side was constructed from three PCBs, supported by Lexan pieces.
-The common PCB substrate material FR4, contains a bromine epoxy which can outgas and absorb electrons degrating the signal \cite{tpcAging}.
-Therefore a halogen free material, Cryogenic-G10, was used for the board material.
-The field cage volume is isolated from the enclosure volume of the TPC for the option to run two separate  gasses, which may be useful when running explosive gasses such hydrogen.
-Instead of a downstream wall a, large, thin exit window was constructed, which consisted of a  \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips, which was bonded to a removable Lexan frame.
-The PCB boards were epoxied, and screwed, into the cathode which was constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure.
-On the other end the boards were epoxied into an aluminum top perimeter which also served as the last ring in the TPC.
+The field cage was constructed from several panels of two-layer printed circuit board (PCBs).
+The front of the field cage was made of two PCBs and each side was constructed from three PCBs.
+These boards had copper strips etched onto the inside and outside of the board which when all the boards were assembed, formed copper rings which were used to set up the uniform electric field, as will be described later.
+The array of circuit boards were glued and supported by Lexan pieces.
+We did not use the common PCB substrate material FR4, which contains a Bromine impregnated epoxy which can outgas Bromine, which can absorb the secondary electrons from the track, thereby degrading the signal \cite{tpcAging}.
+Instead, a halogen free material, Cryogenic-G10, was used for the board material.
+The downstream wall mainly consists of a large, thin exit window.
+The window  is constructed of   \SI{10}{\micro\metre} Kapton upon which are  evaporated aluminum strips that define the electric field.
+At the bottom of field cage is the cathode that is  constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure.
+On the other end the boards were epoxied into an aluminum top perimeter which also served as the last electrode ring in the TPC.
 Together with the cathode bottom, the field cage proved to be a rigid lightweight structure.
 A Lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC.
-Screws with nylon washers, and collars, were used to mount the top perimeter --and therefore the field cage-- to the top plate; it also provided electrical isolation from the top plate.
-In this way the top plate could be removed and rotated with the field cage attached on without damaging any internal components.
+Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate; it also provided electrical isolation from the top plate.
+In this way the top plate could be removed and rotated with the field cage attached, without damaging any internal components.
 
@@
 
-Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
-The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
+Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
+The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
 The cathode voltage $V_{cath}$ can be calculated as,
@@
 The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart.
-The strips on each board was connected to the adjacent strips on the adjacent boards.
-The interfaces were the side pieces met the front boards or back window are connected by G-10 corner pieces with conducting paint strips, which makes the strips into continuous rings around the whole field cage.
+Since the field cage was build with an array of PCB boards, connections from adjacent strips were connected all the way around until all the strips formed a continuous ring.
+The interfaces were the side pieces met were connected by G-10 corner pieces where conductive paint strips were painted on.
 The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip.
 The first outside strip is connected through a via to the second inside strip.
-This pattern repeats until the last strip.
-The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field.
-The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths.
+These outside strips formed a small electric field which repelled electrons away from the PCB substrate between inside strips, preventing build up on the insulating material.
+This pattern repeats until the very last strip.
+A resistor chain, connected to the outside strips, creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field.
+The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, which is the same as the other strip widths.
 The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.
+Allowing for a tunable resistor allowed for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.
 
@@
 The gap between the cathode and the ground of the enclosure is quite small.
-To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground in a controlled manner, reducing the chance of electric breakdown and sparking.
+To prevent electric breakdown in the gas between this gap, a series of concentric copper rings safely stepped down the voltage to ground in a controlled manner, reducing the chance of electric breakdown and sparking.
 There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time.
@@
 The wire plane closest to the pad-plane (\SI{4}{\milli\metre}) are the anode wires.
-The next plane (\SI{8}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid.
+The next plane (\SI{8}{\milli\metre}) is the ground plane or Frisch grid.
+The last plane (\SI{14}{\milli\metre}) is the gating grid.
 The gating grid is the first plane that electrons meet as they drift upward from the field cage volume towards the anode plane.
-The gating grid is operated as a gate, either allowing electrons and ions through, or blocking them entirely.
+The gating grid functions as a gate, either allowing electrons and ions through, or blocking them entirely.
 The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires.
-The ground plane is the least interesting plane and is held to ground by shorting the plane to the enclosure through  shorted BNC terminator on the outside of the TPC.
-We also use the ground plane to input a pulser which is used to spread the pulsed signal to all the pads on the TPC in order to calibrate the electronics of the TPC.
+The ground plane is the least interesting plane and is held to ground by shorting the plane to the enclosure ground through shorted BNC terminator on the outside of the TPC.
+We also use the ground plane to input a pulser, which is used to spread the pulsed signal to all the pads on the TPC in order to calibrate the electronics of the TPC.
 This is done by replacing the shorted BNC with a \SI{50}{\ohm} termination and injecting the pulser on the other end.
@@
 In the open configuration, the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume.
-Typically the gating grid is held in the closed configuration, only opening it when the data acquisition trigger criteria is met.
-By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, and would flood the volume of the field cage with space charge.
+In the closed configuration it is opaque to electrons and ions, when set to the right voltages.
+Typically the gating grid is always in the closed configuration, only opening it momentarily when the data acquisition trigger criteria is met to take data.
+By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, which would flood the field cage with space charge, distorting the electrons represnting the track paths.
 We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume.
-After this we close the gating grid to prevent the back-flow of ions from the avalanche region from that event.
-Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.
+After this, we close the gating grid to prevent the back-flow of ions from the avalanche region from that event.
+Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open.
+This allows for electrons to pass through while preventing the back-flow of ions into the FC volume, reducing or even eliminating the space charge resulting from the avalanche region.
 
 Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the drift lines of electrons in both the on and off configurations.
-In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, in which the optimal voltage is selected for the case of 100\% electron transparency.
+In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is set to the optimal voltage that represents of 100\% electron transparency.
 Figure~\ref{fig:wires_open} shows the electrons are allowed to drift completely through the gating grid all the way to terminate on the anode wires.
 In the off configuration, the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field produced by the voltage difference $2\Delta V$ between wires is great enough to block incoming electrons.
-Figure~\ref{fig:wires_closed} shows this case were the electron drift lines are fully blocked terminating on the more positive wires.
-Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the steady state of the average reference voltage for 100\% transparency.
+Figure~\ref{fig:wires_closed} shows this case were the electron drift lines are fully blocked, as they terminate on the more positive wires.
+Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, coming back to the average voltage setting in the on configuration.
 
@@
 Both configurations of the gating grid were measured and simulated.
-To measure the electron transparency were all wires share the voltage $V_{avg}$, the anode wire was lowered to \SI{500}{\volt} and the beam was allowed to enter the field cage without any target put in.
+To measure the electron transparency in the on configuration all wires were set to the same voltage $V_{avg}$, and the anode wires were lowered to \SI{500}{\volt}.
 By lowering the voltage of the anode wires we could measure the large charge of the beam without saturating the electronics.
-The average charge deposited in the chamber could be measured as a function of $V_{avg}$; changing the top plane resistor appropriately according to Eq.~\ref{eq:TP_resistor}.
+The average charge deposited in the chamber was measured as a function of $V_{avg}$.
+Changing the average voltage required a different top perimeter resistor value according to Eq.~\ref{eq:TP_resistor}.
 Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field.
-Theoretically the most negative value represents 100\% electron transparency and was used as the reference run.
-The electron transparency, $T$, was defined as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run.
-Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation.
+Theoretically the most negative value represents 100\% electron transparency and was used as the reference run in which we scaled all measurements to.
+The measured electron transparency, $T$, was estimated as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run.
+Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation \cite{garfield++}.
 The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.
 
-To measure the electron transparency as a function of the difference voltage $\Delta V$, the average voltage was first set to 100\% transparency, $V_{avg}=\SI{-171}{\volt}$, and the difference voltage was added or subtracted from alternating wires.
+To measure the electron transparency as a function of the difference voltage $\Delta V$, the average voltage was first set to 100\% transparency, $V_{avg}=\SI{-171}{\volt}$, and the difference voltage was added and subtracted from alternating wires.
 Figure~\ref{fig:ggDeltaVTrans} shows the result of the simulation and experiment with and without the magnetic field.
-By introducing the magnetic field the required voltage to close the grid increases.
+By introducing the magnetic field the required voltage to close the grid increased as expected.
 In the experiment we selected the value of $\Delta V = \SI{65}{\volt}$ to ensure we were well within the region of 0\% transparency.
+In both the on and off configurations good agreement was seen with the corresponding Garfield++ simulations when accounting for statistical error in the MC simulations and in the data.
 
@@
 They are biased to high voltages of about  \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
-As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gain between collisions exceeds the ionization potential and ion atom collisions can therefore  knock out more electron-ion pairs in the gas.
-This amplifies the electron current.
+As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gained between collisions exceeds the ionization potential of the gas nocking out more electron-ion pairs in the gas.
+This happens over several collisions creating an avalanche effect amplifying the electron current.
 The final number of electrons produced depends on the anode wire voltage and the gas properties.
 The absolute gas gain was not experimentally determined but it was calculated in a Garfield simulation.
-For the experimental data pertaining to this thesis, the anode wires were biased to two different voltages.
-We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage".
+For the experimental data pertaining to this disertation, the anode wires were biased to two different voltages.
+We will refer to the voltage \SI{1460}{\volt} as the \emph{high voltage} and \SI{1214}{\volt} as the \emph{low voltage}.
 Two sections were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.
@@
 For the voltage of \SI{1460}{\volt} the parameters of the fit are $A_0=903.9$ and $A_1=1.50$ and for the voltage of \SI{1214}{\volt} the parameters are $A_0=150.0$ and $A_1=1.47$.
-These Polya distribution fits will be important later as the input to the Monte Carlo simulaitons of the detector response in Section~\ref{sec:monteCarlo}.
+These Polya distribution fits will be important later as the input to the Monte Carlo simulations of the detector response in Section~\ref{sec:monteCarlo}.
 
@@
 Care must be taken to choose the resister ($R_{TP2}$) that connects the field cage divider chain to the top plate.
-The end of the field cage nearest to the top plate is at the same potential as the rectangular metal top perimeter piece to which the field cage is glued.
-This top perimeter piece forms a horizontal surface that is an equipotential.
-It is separated from the top plate by an annular poly-carbonate insulator.
+The end of the field cage nearest to the top plate is at the same potential as the aluminum top perimeter piece to which the field cage is glued.
+This top perimeter piece forms a horizontal surface that is an equipotential representing the last ring.
 The voltage difference between the average voltage on  the gating grid and the voltage of the top perimeter influences the electric field just below the gating grid.
@@
 
-The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the gating grid and the ground grid (Region 2), and the avalanche region is defined as between the ground grid and the anode grid.
-It is possible that the electric field in Region 1 and Region 2 can be matched but that is not essential.
-To ensure the electric field is the same across Region 1 for every point x and y value across the pad plane, the voltage of the top perimeter  must be set properly by adjusting the value of the last resistor $R_{TP2}$.
-This is not automatically correct because the gating grid average voltage is set independently of the cathode voltage.
-
-To understand how to calculate $R_{TP2}$ we imagine Region 1 to be split into two virtual volumes, one defined as the volume between the cathode and the top-perimeter, and one defined between the top-perimeter and the gating-grid.
+To understand how to calculate $R_{TP2}$ we imagine the space between the cathode and the gating grid can be split into two virtual volumes, Region 1 defined as the space between the cathode and the top-perimeter, and Region 2 defined between the top-perimeter and the gating-grid.
 The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_1$, is defined as,
@@
 \subsection{Pad Plane}
-The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively, making 12096 pads in total.
+The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads, arranged in an array of 108 x 112 pads in the x and z-directions respectively; making 12096 pads in total.
 There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads so that the effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}.
-There is a via and trace coming from each pad, through the board, to the opposite side of the pad plane and is arranged in a surface pads which is readout by a surface mount SAMTEC connector.
+There is a via and trace coming from each pad, through the board, to the opposite side of the pad plane, and is arranged in a surface pads which is readout by a surface mount SAMTEC connector.
 Figure~\ref{fig:padplane} shows the pad plane boards being glued to the top plate and the holes which allow for the readout of pads.
@@
 In both approximations it is described by a most probable energy loss value, with a long, high-energy loss tail.
-The solid curve in Fig.~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc}.
+The solid curve in Fig.~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc} \cite{bichsel}.
 The dashed line is the distribution under the Landau assumptions.
 The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.
-Because of this long tail, for a finite set of energy loss measurements along a given track, the fluctuation of the mean value energy loss as an observable is a very unreliable.
-The most probable energy loss the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.
+Because of this long tail, for a finite set of energy loss measurements along a given track, the mean value is a very unreliable.
+The most probable energy loss is the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.
 
@@
 This approximates the most probable value without performing a fit to a known distribution.
-The set of $n$ observed energy loss values, $\Delta_i/x$, in a given track are sorted from smallest to largest value.
+If a set of $n$ observed energy loss values, $\Delta_i/x$, in a given track are sorted from smallest to largest value.
 The truncated mean $C$, is calculated from the reduced set of points $n_t = f_r n$ as,
@@
 
-where $f_r$ is the cut off fraction; in this thesis, a value of 0.7 was used.
+where $f_r$ is the cut off fraction. In this thesis, a value of 0.7 was used and it was also used in the STAR TPC collaboration \cite{starsyst}.
 

### 2020-07-04 chapters/ch_data_analysis.tex
Second draft. Intro needs work

@@
 \subsection{Correction Task}
+\label{sec:corrTask}
 %extend dynamic range
@@
 Their charge values are mostly distributed around Z~50.
-Let us assume a beam of ${}^{132}$Sn where the energy loss in P10 gas, at \SI{270}{\AMeV}, is \SI{11.2}{\kilo\electronvolt\per\centi\metre}.
+Let us assume a beam of ${}^{132}$Sn where the energy loss in P10 gas, at \SI{270}{\MeVA}, is \SI{11.2}{\kilo\electronvolt\per\centi\metre}.
 From the total number of beams above, and the beam path length of \SI{135}{\centi\metre}, the estimated charge density would be on the order of \SI{3e-8}{\coulomb\per\metre\squared}.
@@
 
-Adding the BDC vertex greatly improves the momentum resolution of the track fitting, but there is a systematic shift when as compared with the momentum value without using a vertex as an extra constraint.
-Since the space charge affects right-going and left-going tracks differently making right-going tracks differently than left-going tracks, it appears they no longer originate from the BDC which is not affected by the space charge.
+Adding the BDC vertex greatly improves the momentum resolution of the track fitting, but there is a systematic shift when  compared with the momentum value before using the vertex as an extra constraint.
+Since the space charge affects right and left-going tracks differently, it appears they no longer originate from the BDC vertex which is independent of the space charge effects coming from the BDC trackers.
 For tracks at polar angles of $\theta_{Lab} < \ang{40}$, the disagreement between momentum values with and without the BDC are much less because the projection of the track does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
@@
 
+This is one of the arguments for the success of the space charge correction task.
 Recall that only the relative distance between left and right-going tracks was minimized, and there was no guarantee that the corrected tracks coincide with the absolute BDC position at the target.
-This is one of the evidences of the correction's success.
 The others being the agreement with the expected cocktail calibration beam as described in Section~\ref{sec:cocktail}.
 
-Here we will discuss some of the final results of the pion kinetic energy spectrum as it pertains to the verification of the space charge analysis,  the details of the pion PID analysis is discussed in detail in Section~\ref{sec:pid}.
+Here we will discuss some of the final results of the pion kinetic energy spectrum as it pertains to the verification of the space charge analysis.
+The details of the pion PID analysis will be discussed in detail in Section~\ref{sec:pid}.
 Figure~\ref{fig:sc_momdist} shows the $\pi^+$ and $\pi^-$ kinetic energy distributions in the center of mass system for the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
 The momentum distributions are split into the beam-left and beam-right side of the TPC.
-In central collisions, one would expect no difference between the two momentum distributions due to the symmetry of the emission.
-Though as seen in the left panels of both figures, the data without the space charge correction is shown, where as in the right panel the data after the space charge correction is shown.
-There is a significant improvement in the matching of the left and right side momentum distributions for both particle species, in both systems.
-It is remarkable that the dOCA observable which was used to calculate the amount of space charge was able to also correct the pion momentum distributions.
-There is no reason to assume correcting the dOCA distribution would be related to the momentum distribution, and is the strongest evidence to the success of the space charge correction.
+For central collisions, one would expect no difference between the two momentum distributions due to the symmetry of the emission.
+As seen in the left panels of both figures, the data before the space charge correction is shown, where as in the right panels the data after the space charge correction is shown.
+There is a significant improvement in the matching of the left and right side momentum distributions for both charged pion species, in both systems, especially for the high energy tail of the pions.
+It is remarkable that the dOCA observable provided a good enough measure of the average space charge in the chamber.
+There is no reason to assume correcting the dOCA distribution would correct the momentum distribution.
+This is the strongest evidence to the success of the space charge correction.
 
@@
 The arrival time of signals originating from each pad may differ due to timing delays in the electronics and cabling.
-The timing differences will affect the y-position measurement of each track.
-The y-direction track residuals reveal that the timing differences correspond to about $\pm$\SI{2}{\milli\metre} in position differences.
+These timing differences affect the y-position measurement of each track.
+By measuring the y-direction track residuals, we find the timing differences correspond to about $\pm$\SI{2}{\milli\metre} in position differences.
 The timing difference is stable for each pad across several runs.
@@
 
- Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows, for three different layers.
+ Figure~\ref{fig:coboCorr} shows the y-residuals, across all the rows, for three different layers; where each row here represents a unique pad with a row and layer ID.
 Before the correction, one can see large deviations in the y-residuals which correspond to timing differences.
@@
 The MC simulation is composed of two separate simulations.
-The first simulation models the interactions of a particle passing through the various materials in the TPC using Geant4.
-A scale model of the field cage, and its materials, where the gas mixture is P-10 gas at a density of 1 atm.
-The magnetic field map of the SAMURAI dipole magnet was imported into Geant4 as well after assuming rotational symmetry along the axis perpendicular to the pole face.
-Along with the energy loss and particle transport, Geant4 also handles multiple scattering, and particle decay, which are all important effects for pions.
+The first simulation utilizes Geant4 to simulate the interactions of a particle passing through the various materials in the TPC.
+A scale model of the field cage was made, and the correct materials types were put in, with the correct gas mixture and density of P-10 gas at a density of 1 atm.
+The magnetic field map of the SAMURAI dipole magnet was also imported into Geant4 as well after assuming rotational symmetry along the axis perpendicular to the pole face -- since only 1/4 of the magnet was simulated.
+Along with the energy loss and particle transport, Geant4 also handles multiple scattering, and particle decays, which are all important effects especially for calculating the inefficiencies of pions.
 The output of Geant4 is a series of energy loss points which contain the amount of energy lost in $\si{\kilo\electronvolt\per\centi\metre}$ and the location of the energy loss in Cartesian space $(x,y,z)$.
@@
 
- The first step of the drift task is to convert the primary ionization points of the MC track provided by Geant4 into electrons.
+ The first step of the drift task is to convert the primary and secondary ionization points of the MC track provided by Geant4 into electrons.
 The average number of electrons created in a gaseous detector,$N_{e^-}$, can be described as,
@@
 where $I$ is the ionization coefficient of P10 gas (Table~\ref{tb:gas}) and $\Delta E$ is the energy loss deposited.
-Each electron is propagated along the electric field lines assuming the electric field is uniform, which is true for mos tof the pad plane region.
-The total length drifted from the initial primary ionization point to the final anode wire is given by $L_{anode}$.
+Each electron is then drifted along the electric field lines, under the assumption  that the electric field is uniform, which is true for most of the pad plane region.
+The total length drifted from the initial primary ionization point to the final anode wire is $L_{anode}$.
 
@@
 Drifting electrons frequently collide with the detector gas causing them to change direction.
-This stochastic motion is  described by a a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
+This stochastic motion is  described by a diffusion process occurring along the direction of travel (longitudinal) and transverse to the motion.
 The longitudinal ($c_{l}$) and transverse ($c_{t}$) diffusion coefficients are determined by Garfield++ calculation  \cite{garfield++} in the presence of a \SI{0.5}{\tesla} magnetic field, listed in Tb.~\ref{tb:gas}.
-The diffusion process is modeled by randomly sampling from a Gaussian distributions describing both of the directions.
+The diffusion process is modeled by randomly sampling from a Gaussian distributions describing the diffusion in both of the directions.
 The random displacement vector is then added to the final position of the electron.
-The deviation in the transverse direction, $dr$, is randomly sampled from,
+The deviation in the transverse direction, $dr$, is randomly sampled from the distribution,
 
@@
 
-where $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.The Cartesian directions are given by $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, since there is no preferential angle  of emission in the transverse plane.
+where $\sigma_{t}=c_{t}\cdot\sqrt{L_{anode}}$.The transverse Cartesian directions can be written as $dx = dr \cdot \cos(\alpha)$ and $dz = dr \cdot \sin(\alpha)$, where $\alpha$ is a random angle from 0 to 2$\pi$, since there is no preferential angle  of emission in the transverse plane.
 The shift associated with the longitudinal diffusion, $dl$, is randomly sampled from,
@@
 This is because the y-position corresponding to $t=0$ in the data time bucket spectra, does not correspond to the same position in the MC.
-This value was found to be \SI{23}{\nano\second} by aligning the MC vertex to the data vertex.
-
+
+Once the electrons have terminated on a particular wire, the avalanche process is simulated.
 The total number of electrons produced in the avalanche process of a single electron was simulated in Garfield++ and discussed in Section~\ref{sec:wireplanes} for the anode wire voltages used in the experiment.
-The number of electrons produced in the avalanche is randomly sampled from the distribution depending on which anode wire section the electron terminated.
-The number of electrons produced is stored as a gain factor instead of multiple instances of the original electron to save space and computation time.
+The number of electrons produced in the avalanche is randomly sampled from the Polya distribution  corresponding to the anode wire sections the electron terminated.
+The number of electrons produced is stored as a gain factor in that particular electron, instead of multiple instances of the original electron, to save storage space and computation time.
 
@@
 
-It is worth mentioning there is the possibility to simulate the space charge effects in this task where as the inverse to the correction map described in Section~\ref{sec:spacecharge}.
+It is worth mentioning there is the possibility to simulate the space charge effects in this task.
 This is not used for calculating the response of the TPC for efficiency calculations since it is a trivial exercise to input the space charge map only to then correct for it using the inverse map.
@@
 \subsection{Pad Response Task}
-The total charge of each avalanche is distributed according to the pad response function (PRF) described in Section~\ref{sec:prf}.
+The total charge of each avalanche is then distributed according to the pad response function (PRF) described in Section~\ref{sec:prf}.
 The PRF is simulated as the double integral of a 2-dimensional Gaussian.
@@
 where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the final position of the drifted electron.
-The parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
+The Gaussian width parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
 
@@
 
-The purpose of the electronics task is to simulate the electronics response to charge induced on each pad, converting charge into ADC channels.
-The induced charge on each pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
-The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, as long as the electronics settings are fixed.
+The purpose of the electronics task is to simulate the electronics response to a particular charge induced on each pad, converting charge into ADC channels.
+In a simplified picture, the induced charge on each pad goes through a pre-amplifier and shaping amplifier which determine the final pulse shape that is read out.
+The pulse shape did not change significantly in any circumstance, such as pulse height, data type, or particle type, as long as the electronics settings were fixed.
 This allows us to assume the pulse shape is constant and can be described by two variables, the height of the pulse, $Q$, and the starting time bucket of the pulse, $t_o$.
-The starting time of the pulse is defined as the time at .1$\cdot Q$ of  the rising edge.
+The starting time of the pulse is defined as the time at 10\% pulse height, on the rising edge.
 The shape of the pulse depends on the shaping time constant which was set to \SI{117}{\nano\second} for the data analyzed here.
@@
 
-where $e$ is the charge of the electron in $\si{\femto \coulomb}$, $N_{e}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
+where $e$ is the fundamental charge of the electron in $\si{\femto \coulomb}$, $N_{e}$ is the total number of avalanche electrons, $ADC_{pedestal}$) is the pedestal (300 ADC, $\mathrm{ADC_{Max}}$ is the maximum allowed ADC value (4096), and $f_c$ is the dynamic range setting (\SI{120}{\femto\coulomb}).
 The pulse shape given in Fig.~\ref{fig:pulseshape} is multiplied by the pulse height $Q$ giving the full time bucket estimate of the TPC response.
@@
 
-where $\langle dE/dx\rangle_{MC}^p$ and $\langle dE/dx\rangle_{Data}^p$  are the energy loss values for a particle in a given momentum bin $p$. The calibration will be discussed later in more detail.
+where $\langle dE/dx\rangle_{MC}^p$ and $\langle dE/dx\rangle_{Data}^p$  are the energy loss values for a particle in a given momentum bin $p$.
+The calibration will be discussed later in Section~\ref{sec:mccomparison}.
 
@@
 \label{sec:embedding}
+
+There are several effects in a complex experimental set up that influence the data in either unknown or un-quantifiable ways.
+One such effect is the bias of the triggering system, here the Kyoto and Katana multiplicity arrays preferentially select data that are emitted in a particular reaction plane.
+As previously discussed, the saturation effects of the electronics, notably the shadowing of other tracks.
+The effects of track multiplicity and the distribution of heavy ion and residues from the breakup of the target and projectile which can cause tracking failures.
+In these cases, a full MC simulation would be cumbersome if not impossible to do.
+Simply put, there is no better substitute for simulating experimental data, than the experimental data itself.
+By embedding MC tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all these sources of biases which are contained in the data, while at the same time measuring the response of the TPC.
+
+%As discussed in Section~\ref{sec:software}, the software is composed of several tasks, each of which have the possibility of introducing errors or biases through assumptions made in the tracking algorithm.
+The TPC system itself introduces errors related to the physical measurement which can be addressed through modeling the TPC, and its materials.
+
+%For all of the sources of biases in the TPC, it is clear that a MC type approach to solving for the response of the TPC would be the most practical method.
+Also by embedding MC events into real data we can take into account the effects of the experimental setup as well as the software and the detector system all together at once.
 Track embedding is the process of taking a MC simulated track and embedding its response into an experimental data event.
 After reconstructing this new embedded event we match the input MC track to the corresponding final reconstructed track.
-By doing so we can evaluate the response of the entire TPC system to any given input.
-The TPC measurement can be thought of three different systems each which introduce errors and or biases into the final reconstructed tracks; the software, the detector, and the experimental setup.
-The
-
-There are several effects in a complex experimental set up that influence the data in either unknown or un-quantifiable ways.
-One such effect is the bias of the triggering system, here the Kyoto and Katana multiplicity and veto arrays preferentially select data that is emitted in a particular reaction plane.
-As previously discussed saturation effects in the electronics, notably the shadowing of other tracks.
-The effects of track multiplicity and the distribution of heavy ion and residues from the breakup of the target and projectile.
-One could propose to make a simulation or investigation into each individual effect in order to make a full MC simulation of data and study the response of the detector this way.
-Though great care was taken in developing an accurate MC simulation, like all models there are assumptions and simplifications that were made.
-Simply put, there is no better substitute for experimental data than itself.
-By embedding MC tracks into experimental data -- and propagating through the tracking and reconstruction algorithm -- one can account for all these sources of biases which are contained in the experimental data, while at the same time measuring the response of the TPC.
-
-As discussed in Section~\ref{sec:software}, the software is composed of several tasks, each of which have the possibility of introducing errors or biases through assumptions made in the tracking algorithm.
-The detector system introduces errors related to the physical measurement itself which can be addressed through modeling the TPC, and its materials.
-
-For all of the sources of biases in the TPC, it is clear that a MC type approach to solving for the response of the TPC would be the most practical method.
-Also by embedding MC events into real data we can take into account the effects of the experimental setup as well as the software and the detector system all together at once.
+By doing so, we can evaluate the response of the entire TPC system to any given input.
 
@@
 
-The detailed flow diagram of the software implementation of embedding algorithm is shown in Fig.~\ref{fig:flow}.
-Starting from Geant4 and passing through the 3 MC tasks described above, the output response of the MC track in the TPC is created.
-From here we directly embed the MC signals into data by adding the MC signals into the data time bucket spectrum, pad-by-pad.
-As described in Sec.~\ref{sec:simSat}, prior to embedding the MC signals, all pads which are saturated in the experimental data are identified and flagged.
-The MC signals are forbidden to be embedded after the MC saturation time bucket value.
-Figure~\ref{fig:mcevent} shows the response of a 200 MeV/c $\pi^-$ track in the TPC; we have embed this track into the data event, Fig.~\ref{fig:dataevent}, just as an example.
-We will refer to experimental data with embedded MC signals as ``embedded data" for short, the MC generated response as ``MC data", and data only containing only the experimental signals as ``experimental data".
-The three set of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
-From this the three independent hit data sets are temporarily stored.
+The detailed software flow diagram of the embedding algorithm is shown in Fig.~\ref{fig:flow}.
+Starting from Geant4 and passing through the 3 MC digitization tasks described above, the output response of the MC track in the TPC is created.
+From here we directly embed the MC signals into experimental data by adding the MC signals directly into the experimental time bucket spectrum, pad-by-pad.
+As will be described in Sec.~\ref{sec:simSat}, prior to embedding the MC signals, all pads which are saturated in the experimental data are identified and flagged.
+MC signals are forbidden to be embedded after the (MC) saturation time bucket position identified in the procedure outlined in Section~\ref{sec:simSat}.
+Figure~\ref{fig:mcevent} shows the response of a \SI{200}{\mega\electronvolt\per\clight} $\pi^-$ track in the TPC.
+Figure~\ref{fig:dataevent} shows this track embedded into an experimental event.
+In the following, here I will refer to experimental data with embedded MC signals as ``embedded data" for short, the MC generated response as ``MC data", and data only containing only the experimental signals as ``experimental data".
+The three sets of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
+These three independent hit data sets are temporarily stored.
 
 Here within the PSA algorithm, is the first and most important part of the embedding software.
-The PSA task has the job to identify which hits in the embedded data set originate from the MC hits set and which are from the experimental data.
+If the embedding portion of the PSA task is turned on,  it has the job to identify which hits in the embedded data set originate from the MC hits set and which are from the experimental data.
 Once the MC hits are identified, these embedded hits can be tagged and tracked through the entire software.
@@
 The surviving hits are then compared with the MC hit data set, where the criteria for a matching MC hit is, $\left|t_{\mathrm{Embed} } - t_{\mathrm{MC} }\right| < 3$.
-There is no requirement for a matching criteria for the charge values which would artificially bias the charge values selected depending on our cut; which is critical for minimum ionizing particles such as pions.
-This can only be accomplished by first removing almost all of the experimental hits as was done in the first step.
+There is no requirement for a matching criteria for the charge values which would artificially bias the charge values selected depending on our cut.
+This is critical for minimum ionizing particles such as pions; this can only be accomplished by first removing almost all of the experimental hits as was done in the first step.
 Each embedded hit that passes this step of criteria is tagged as a originating from a MC hit.
@@
 When the hits in a helix track are clustered, if a cluster contains one embedded hit it is itself tagged as an embedded cluster.
-Furthermore, if a track that is reconstructed contains any embedded clusters, it is tagged as an embedded track.
+Furthermore, if a track that is reconstructed contains any embedded clusters, it is tagged as an embedded track also.
 The goal of this na\"ive tagging approach is to preserve all the information of where the embedded hits, clusters, and tracks have gone, preserving as much information until the end without introducing a bias.
 
-  It is the job of the embedded correlation task to identify which of the final embedded tracks are candidates for the original input track.
-For example, several things may happen along the process of embedding that may disqualify a track as a candidate in this na\"ive tagging.
-A track could break up, lose or share its charge with and adjacent track; it may be not be identified at all for a variety of other reasons.
-For the embedded track to be a candidate of the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track, and $N_{\mathrm{Total}}$ is the number of total clusters.
+  When MC embedding is performed, another task is added to the end of the analysis routine, namely the embedded correlation task.
+It is the job of this task to identify which of the final embedded tracks are candidates for the original input track.
+For example, several things may happen along the process of embedding that may disqualify a track as a candidate in this na\"ive tagging approach.
+A track could break up, lose or share its charge with an adjacent track, or it may be not be identified at all for a variety of other reasons.
+For the final reconstructed track to be a candidate for the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track, and $N_{\mathrm{Total}}$ is the number of total clusters.
 The first criteria is a simple minimum cut where to ensure the minimum condition of a embedded track is met at least having 5 clusters.
 The second criteria is the strongest cut, ensuring the track has at least half of its clusters coming from embedded MC signals.
-The set of tracks which satisfy both conditions are saved into an array of candidate tracks.
-In this way the user can decide how to sort these tracks for studying either efficiency or track splitting.
+The set of tracks which satisfy both conditions are saved into an array of candidate tracks, which may represent the original MC track.
+This is done so that if any track splitting has occurred, all of the split tracks will be saved into the vector.
+it is then left to the user to decide how to further study these tracks.
 
 If the embedded track is split into multiple tracks, the user must select what they believe to be the track which most represents the original track, for the purposes of calculating efficiency.
-Typically only one of the tracks in the set will satisfy one of the reasonable quality cut conditions in an analysis.
-For example the most reasonable assumption is the track originating from the vertex is the most correct in this case, where as the split fraction of the track wont originate from the vertex but instead have a large distance to vertex value.
+This is to preserve the definition of efficiency is strictly $\leq 1$.
+Typically only one of the tracks in the set will satisfy one of the reasonable quality cut conditions in an analysis which will be discussed further in Section~\ref{sec:qualitycut}.
+For example, the most reasonable assumption is the track originating from the vertex is the most correct in this case, where as the split fraction of the track wont originate from the vertex but instead have a large distance to vertex value.
 In this thesis, the track with the minimum distance to vertex is identified as the track in which the efficiency analysis will be performed.
@@
 
-In general all amplifiers have a finite range of output values set by a positive and negative rail (typically -12V and +12V).
-If the input charge into a charge sensitive pre-amplifier causes the output voltage to reach the max (or min) output voltage, the electronics is considered saturated and the response may be non-linear.
+Since saturation is one of the largest and most important effects in the data, simulating it correctly in MC embedding becomes paramount.
+In general all preamplifiers have a finite range of output values set by a positive and negative rail (typically -12V and +12V).
+If the input charge into a charge sensitive preamplifier causes the output voltage to reach the max (or min) output voltage, the preamplifier saturates and the response may be non-linear.
 The picture is complicated by the fact that there is usually an RC feedback loop which dissipates the input charge.
@@
 This occurs for all tracks, even for minimum ionizing tracks.
-Saturation in this case is infrequent and not an issue as there are many other clusters are not saturated.
+In this case, the saturation is infrequent and typically not an issue as there are many other clusters are not saturated.
 As the charge of the particle gets higher, or the momentum gets lower, the  energy loss value also becomes higher, and a significant fraction of pads in a track may be saturated.
-In Section~\ref{sec:desat}, we outlined an algorithm to correct for the saturation for a  particular track, but this saturation also affects surrounding tracks.
+In Section~\ref{sec:desat}, we outlined an algorithm to correct for the saturation for a  particular track, but saturation also affects the measurement of surrounding tracks.
 In order to make an accurate MC simulation, we must understand how saturation affects surrounding tracks.
 
-Pads that saturate have no signal or ''dead" for a given amount of time  after the saturating signal, which depends on the amount of input charge.
-For charges that are at or above the level of the dynamic range, the pad will certainly be dead for the remaining time of the measurement.
-Signals from tracks passing directly underneath this pad, with signals that would arrive later in time,  will therefore not be observed.
-In a sense the saturated pad is ``shadowing" any future tracks.
+Pads that saturate have no signal or are \emph{dead} for a given amount of time after the saturating signal, depending on the input charge.
+For charges that are at or above the level of the dynamic range, the pad will certainly be dead from that moment on for the remaining time of the measurement.
+Signals from tracks passing directly underneath this pad, arriving later in time,  will therefore not be observed in that pad.
+In a sense the saturating signal is \emph{shadowing} any future tracks.
 As the track multiplicity of an event increases, the probability of track shadowing increases.
@@
 In this case, channels can be effectively dead for several events before recovering.
-There is a large amount of high energy electrons, commonly referred to as ``delta rays",  which are produced from the beam passing through the gas, due to the high atomic charge of the beam.
-Horizontally almost all of the electrons cannot travel very far, even for very high energy electrons, due to the vertical magnetic field causing them to curl in tight circles.
+There is a large amount of high energy electrons, commonly referred to as \emph{delta rays},  which are produced from the beam passing through the gas.
+These electrons scale with the charge of a particle, in which the Sn beam produces several high energy electrons \cite{pdg}.
+In the presence of the magnetic field, almost all of the electrons traveling perpendicular to the field curl up and cannot travel very far, even for very high energy electrons.
 Figure~\ref{fig:deltaE} shows the horizontal and vertical extent of the delta rays in a top down and side views of the TPC.
@@
 
+%Example picture of dead pads?
+
 \begin{figure}[!htb]
@@
 
-The simulation saturation effects is handled naturally by embedding MC tracks into real data; once the time bucket of the saturating signal has been identified, simply no further signals are embedded since the pad is assumed to be dead.
+Simulating the saturation effects is handled naturally by embedding.
+Once the time bucket of the saturating signal has been identified, no further signals are embedded since the pad is assumed to be dead.
 In the case a pad is dead for the whole event, the time of saturation is set to the first time bucket and no signal is embedded for any time bucket.
@@
 The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
-The long exponential tail can be effectively removed by a simple software technique which is similar to the electronics concept of ``pole-zero compensation".
+The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation CITE HERE.
 If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,
@@
 
-By applying this technique to the real data we can identify whether a pad has a saturating pulse in it by looking for this negative undershoot.
-A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ ADC for more than 8 time buckets and the max ADC value is > $G\cdot$500; $G$ is the gain calibration coming from low gain sections discussed in Section~\ref{sec:anodeCalib}.
-The max ADC condition eliminates false positives which come from  dead pads where the gating grid subtraction is still applied, introducing large false negative peaks, but have no positive signals.
+By applying this technique to the real data, we can identify whether a pad has a saturating pulse by looking for this negative undershoot.
+A negative peak is identified as a saturated signal if it exceeds the threshold of less than -20$\cdot G$ ADC for more than 8 time buckets, and if the max ADC value is > $G\cdot$500, where $G$ is the gain calibration coming from low gain sections discussed in Section~\ref{sec:anodeCalib}.
+The max ADC condition eliminates false positives which come from  the dead pads where the gating grid subtraction is still applied, introducing large false negative peaks, but have no positive signals.
 After meeting these two conditions, the pad is flagged as saturated and the time bucket position of saturation is set to $t_{peak} - 5$, where $t_{peak}$ is the time bucket position of the negative peak.
-This is because the falling edge since the falling edge is around 5 time buckets after the beginning of the saturation.
+This is because the falling edge is around 5 time buckets after the rising edge of the saturation.
 A separate time is also stored called the MC time bucket position, which sets the point in time in which a signal cannot be embedded any further.
@@
 There is another way a pad can saturate which occurs when all the pulse heights within the time bucket spectrum add to more than the max ADC threshold of 3500 ADC.
-This arises because the fall time of the pre-amplifier circuit is much longer than the time bucket measurement window, and pulses are allowed to pile up in the pre-amplifier.
-In this case no single pulse will reach over the max ADC value yet the last pulse will have a pulse shape that is missing the long characteristic tail, but otherwise looks like a normal pulse.
-These pulses are also identified by the algorithm described above, since this type of saturated pulse is still no long tail.
+This arises since the fall time of the preamplifier circuit is much longer than the time bucket measurement window, and pulses are allowed to pile up in the preamplifier.
+In this case, no single pulse will reach over the max ADC value yet the last pulse will have a pulse shape that is missing the long characteristic tail, but otherwise looking like a normal pulse.
+These pulses are also identified by the algorithm described above, since this type of saturated pulse  still has no long tail.
+
  When embedding signals into pads, we do not consider the previous amount of total charge in the pad.
-We assume this type of saturation is a higher order correction, where and the current algorithm approximates most of the saturation effects.
+We assume this type of saturation is a higher order correction, and the current algorithm approximates most of the saturation effects.
 Figure~\ref{fig:satTag} shows all of the pads in an event which are tagged by the algorithm as saturated, denoted by the black dots.
 The max ADC values in a pad are shown in the z-axis color scale, to give a sense of the success of the algorithm.
-It may appear that the ADC values of some pads appear to be lower than 3500 ADC, but upon further inspection they all satisfy the mode of saturation where the total sum of heights is greater than the max ADC value as described earlier.
+It may appear that the ADC values of some pads appear to be lower than 3500 ADC, but upon further inspection they all satisfy the mode of saturation where the total sum of heights is greater than the max ADC value as mentioned earlier.
 
@@
 value  of less than 50 ADC and a max ADC < 50 over the full time bucket spectrum.
-Here the dead pad is also tagged as saturated and the time bucket position of saturation is set to the first time bucket, or time zero.
+Here the dead pad is also tagged as saturated and the time bucket position of saturation is set to the first time bucket.
 
@@
 \subsection{MC and Data Comparison}
-
+\label{sec:mccomparison}
 Here we compare several important observables to ensure the MC is simulating the data sufficiently.
 These observables will be relevant later in the discussion of quality cuts and efficiency analysis discussed in Section~\ref{sec:efficiency}.
-The most important observable is the total number of clusters.
+The most important observable is the total number of clusters, $N_{clust}$.
 The number of clusters in a track depends on the geometry of the TPC, the spherical angles of the track, and the clustering algorithm.
 MC tracks were embedded into data for several particle types.
-The input angular and momentum distribution was uniform where the final distributions were weighted such that the angular and momentum distribution matched that of the data for a fair comparison.
+The input angular and momentum distribution was uniform.
+The final measured distributions were weighted such that the angular and momentum distribution matched that of the data for a fair comparison.
 Figure~\ref{fig:clustcomp} shows the normalized cluster distributions for the MC and data for several particle types.
-There is no data below 20 clusters because this is the minimum cut to ensure quality PID lines in the data.
+There is no data below $N_{clust} < 20$ because this is the minimum cut to ensure quality PID lines in the data.
 The normalization is changed for each particle type to display them on the same scale.
@@
 The distributions were again normalized to different values to display on the same scale.
-Here the only other cut was the  track number of clusters cut > 20, which cleaned up the background spectrum.
+Here the only other cut was the $N_{clust} > 20,$ which cleaned up the background spectrum.
 It is also worth mentioning the data plotted is also after the space charge correction.
 Before the space charge correction the distributions would be much wider.
-Since we correct for the space charge in the data we compare here to the embedded MC without any space charge effects, making a fair comparison.
+It is a fair comparison to compare the space charge corrected data distributions to the MC without space charge effects.
 
@@
 
+The other important cut which is made inside of the software is on the PRF distribution of each cluster discussed in Section~\ref{sec:corrTask}.
 We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Fig.~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data.
@@
 Because the target is several cm away from the widow of the field cage the geometric acceptance is not even 2$\pi$.
-The rectangular design complicates the calculation of the geometric acceptance, or the efficiency.
-Because of this there are regions of the TPC where it is impossible to reconstruct a track due to the geometric in these regions the efficiency is exactly 0.
-In the regions of non-zero efficiency, the efficiency is a function of at least three main parameters that define the phase space of a track, momentum $p$ and the two spherical lab angles $\theta_{Lab}$ and $\phi_{Lab}$.
+The rectangular design complicates the calculation of the geometric acceptance of the TPC.
+Because of this, there are regions of the TPC where it is impossible to reconstruct a track due to the geometric constraints; in these regions the efficiency is exactly 0.
+In general the efficiency is a function of at least four main parameters.
+Three define the phase space of a track, momentum $p$ and the two spherical lab angles $\theta_{Lab}$ and $\phi_{Lab}$, and one is the multiplicity of the event.
+These are the main contributing factors to the efficiency calculations.
 
@@
 where $n_{reco}$ is the number of embedded tracks which are successfully reconstructed, and $N$ is the total number of input embedded tracks for that given bin.
-A track is defined to be successfully reconstructed if it exists in the correlation task described in \ref{sec:embedding}, and therefore meets the minimum criteria for an embedded track, and if it passes the various quality cuts performed on the experimental data set; this could be angular cuts, PID cuts, etc.
-The TPC response will introduce a finite resolution in both the angles and the momentum.
+A track is defined to be successfully reconstructed if it exists in the correlation task described in \ref{sec:embedding}, and therefore meets the minimum criteria for an embedded track, and if it passes the various quality cuts performed on the experimental data set; this could be angular cuts, multiplicity cuts, PID cuts, etc.
+The TPC's response introduces a finite resolution in both the angles and the momentum.
 Because of the final measured track could have migrated out of the input bin into another bin.
@@
 If not a more complicated un-folding procedure is needed.
-The efficiency is defined as the number of all tracks that were reconstructed and passed the cuts successfully, even if some had migrated to other bins, the efficiency value is assumed to represent the center of the input MC bin.
+A track is defined as successfully reconstructed even it it has migrated out of the bin, and has successfully passed all cuts.
+The efficiency value is assumed to represent the center of the input MC bin.
 %TALK ABOUT BIN SIZES HEREE?????
 
- Single MC tracks are embedded into a set of \num{e4} on target events and the measured output tracks are correlated to the input track from the embedding software.
+ MC tracks are embedded into a set of \num{e4} on target events and the measured output tracks are correlated to the input track using the embedding software.
 If the track split in the analysis software, and there are several tracks identified by the software as an embedded track, or part of one, we certainly cannot double count the track as the efficiency is  defined strictly as $\epsilon < 0$.
 In this case, the track with the minimum distance to vertex is taken for the purposes of calculating efficiency.
-This is because in the case the track splits, it is very rare for the later segments of the track to have a distance-to-vertex which originates from the target region.
-It's reasonable to assume the track with the minimum distance to vertex is the track that best represents the ``real" track.
+It is very rare for the later segments of the track to have a distance-to-vertex which originates from the target region.
+It's reasonable to assume the track with the minimum distance to vertex is the track that best represents the MC input track.
 The embedded MC tracks are generated as a uniform distribution in $\theta$,$\phi$, and $p$, to ensure each bin has as similar number of input tracks $N$, giving approximately the same statistical error.
-Here the efficiency distribution is independent of the initial input distribution since we were careful to not assign tracks that migrated out of the input bin, into another bin, which would make the efficiency biased by the input distribution.
+Here the efficiency distribution is independent of the initial input distribution since we were careful to define reconstructed tracks, as those which may have migrated out of the bin.
+Neglecting them would make the efficiency biased by the input distribution.
+Embedding into \num{e4}  events ensures a good sample of the multiplicity distribution, which the efficiency is also a function of.
 
@@
 Figure~\ref{fig:pim_eff_ex} shows the efficiency calculated for $\pi^-$ tracks traveling with momentum \SI{200}{\mega\eVperc} as a function of the two laboratory angles.
-It is easiest to visualize the efficiency distribution first in a polar representation which is most like viewing the actual TPC progressing to a visualization that is easier to see the values.
+It is easiest to visualize the efficiency distribution first in a polar representation which is most like viewing the actual TPC and progressing to a visualization that is easier to see the values.
 Figure~\ref{fig:pim_pol_eff_ex} shows the efficiency values plotted in a polar representation where $\theta_{Lab}$ is represented as the radial dimension and $\phi$ is represented as the polar angle.
-This way is best understood as if one looked at the TPC from the downstream point of view.
+This representation is best understood as if one looked at the tracks in the TPC from the downstream point of view.
 Of course the vertical direction of the TPC,$\phi$=\ang{90}, is limited in vertical space, therefore tracks are not reconstructed well.
 The same is the case in downward going tracks, the region centered around $\phi$=\ang{270}.
-The best regions are seen by the high efficiency values which are left and right in the TPC where the rectangular TPC has the highest acceptance.
-Though this view is the most intuitive to think about in a physical geometry sense, the values of the efficiency in each bin are difficult to see.
-Figure~\ref{fig:pim_flat_eff_ex} shows the same efficiency plot but plotted in a more standard two dimensional plot.
-Here we can see even better the acceptance of downward going tracks , $\phi$=\ang{270}, is greater than the dip in upward going tracks, $\phi$=\ang{90}, due to the field cage being centered in the magnet and not in the TPC.
+The most efficient regions left and right in the TPC where it has the most acceptance.
+Though this view is the most intuitive to think about in geometric sense, the values of the efficiency in each bin are difficult to see.
+Figure~\ref{fig:pim_flat_eff_ex} shows the same efficiency plot but in a Cartesian plot.
+Here we can see even better the acceptance of downward going tracks , $\phi$=\ang{270}, is greater than the dip in upward going tracks, $\phi$=\ang{90}, resulting from the field cage being shorter in the top than in the bottom, this was due to the space the electronics took up on the top of the TPC, making the field cage asymmetric up and down.
 

### 2020-07-04 chapters/ch_data_analysis_2.tex
Something broken. Saftey backup
backup
Halfway through data analysis 1 backup
Second draft. Intro needs work

@@
 The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
-The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by 1 linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
+The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by a linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
 Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.
@@
 \includegraphics[width=\textwidth]{beamPID.png}
+\caption{Overall beam PID for all the systems. Several contaminants other than the desired secondary beam can be seen.}
 \label{fig:beampid}
-\caption{Overall beam PID for all the systems. Several contaminants other than the desired secondary beam can be seen.}
 \end{figure}
@@
 The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$.
-Several sets of slits further purify the secondary beam quality by throwing away particles which do not focus on the right focal planes.
+Several sets of slits further purify the secondary beam quality by discarding particles which do not focus on the right focal planes.
 These are the areas where the particles with different velocities focus to different locations in space, which occur at F3,F5, and F7 positions.
@@
 The particle identification of each beam is achived by the TOF-B$\rho$-$\Delta$E method described in \cite{bigrips}, where the Time of Flight (TOF) information is given by the time it takes to cross two plastic scintilators at F3 and F7 focal planes, and the $\Delta$E information is given by the MUlti-Sampling Ionization Chamber (MUSIC) \cite{music}.
-From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussians in this space.
+From this method the atomic charge, Z, and mass to charge ratio, A/Q, of each particle was measured and separate species represent two-dimensional Gaussian in this space.
 
@@
 The beam purity of each desired secondary beam is listed in Table~\ref{tb:beams}.
-The desired secondary beam of interest can be selected by using an appropriate gate around the corresponding group, since the BigRIPS PID resolution is good enough to separate separate beam types.
+The desired secondary beam of interest can be selected by using an appropriate gate around the corresponding particle.
 Each particle gate is selected by fitting a multivariate normal distribution with two variables defined as,
@@
 \end{table}
+
 Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam.
@@
 Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.
-This effect comes from an edge effects where the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad, or in the case of the vertical time bucket spectrum, represents the last known position of the collected charge.
+This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge.
 In this case the cluster position is biased towards the inside of the TPC.
-While the number of affected clusters is small as compared with the total number in the track, but the deviation at the end of the track is enough to start causing issues in the momentum reconstruction.
+This also occurs for the first and last time bucket in the vertical direction.
+While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction.
 Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC.
@@
 \end{figure}
-The track multiplicity in each event is quite high and since the tracks all originate from a common vertex, the density of tracks near the target region is very high.
-In this region the track separation is too small for the software to correctly determine and sort the charge information into the appropriate track.
-The information provided by extra vertex point from external BDC tracking described in \ref{sec:bdc}, provides all the information about the vertex location.
+
+The track multiplicity in the TPC ranges from 20-80 tracks.
+Since all the tracks originate from a common vertex, the density of tracks near the target region is very high.
+In this region the separation between tracks is too small for the software to correctly determine which clusters belong to which tracks, making the information useless or even incorrect.
+Also, the information provided by extra vertex point from external BDC tracking described in \ref{sec:bdc}, provides all the information about the vertex location.
 The bad quality of hit information near the target region only hurt in the tracking and PID of a track.
@@
 \label{sec:beamangle}
-The incoming secondary beam is deflected by the magnetic field and impinges on the target at some small but significant angle.
-From the BDCs tracking information, the beam is projected as a straight line right up until entering the magnet.
-The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position.
-The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ defined as,
+The incoming secondary beam is deflected by the magnetic field and impinges on the target at some angle.
+From the BDC tracking information, the beam is projected as a straight line right up until entering the magnet.
+The beam is then propagated through the magnetic field using a Runge-Kutta integration until reaching the target position \cite{jon}.
+The beam angle on target can be categorized by two angles $\theta_{a_{proj}}$ and $\theta_{b_{proj}}$ where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector:
 
@@
   \theta_\mathrm{a,proj}=\tan^{-1}\frac{p_x}{p_z},\quad
-  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z},
+  \theta_\mathrm{b,proj}=\tan^{-1}\frac{p_y}{p_z}.
   \label{beamAngle}
 \end{equation}
-
-where $p_x$, $p_y$, and $p_z$ are the components of beam momentum vector.
 
@@
     \end{subfigure}
+\caption{Beam angle at the face of the target in the TPC for all 4 systems.}
 \label{fig:beamangle}
@@
 Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}.
-Helping to eliminate some of the poorly reconstructed beam events.
+This helped to eliminate some of the poorly reconstructed beam events.
+The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionspectra}.
 
 \section{Vertex Cut}
+\label{sec:vertexcut}
 
@@
 
-The vertex information of an event, described in \ref{sec:vertex}, is estimated from reconstructing the tracks of an event to one common source.
-The secondary beam encounters several solid and gaseous materials were a nuclear reaction can occur anywhere along the beam line.
-Figure~\ref{fig:overviewVertex} shows the z component of the reconstructed vertex for all events in the ${}^{132}$Sn system.
+The vertex information of an event, described in \ref{sec:vertex}, is estimated from reconstructing the tracks of an event to one common point.
+The secondary beam encounters several solid and gaseous materials along the beam line where a nuclear reaction can occur.
+Figure~\ref{fig:overviewVertex} shows the z-component of the reconstructed vertex for all events in the ${}^{132}$Sn system.
 Several peaks are seen in the spectrum which correspond to several dense materials in the beam line such as the entrance windows, target frame, Active Veto, with the largest peak representing the target.
-Also collisions happen with the detector gas inside the TPC volume which are typically called active target events.
+Collisions also happen with the detector gas inside the TPC volume where the gas acts as a target.
 To ensure that the secondary beam is really on the target of choice we perform a vertex cut around where we believe the vertex location of the target to be.
 
- The target position was measured to be \SI{-13.2}{\milli\metre}.
+ The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame.
 The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}.
@@
 The mean position for each system is listed in Table~\ref{tb:vertexresol}.
-Since the target thickness was less than \SI{1}{\milli\metre} for all targets.
-Neglecting the small distributions in the thin target, the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC.
+Since the target thickness was less than \SI{1}{\milli\metre} for all targets, we can assume the thickness of the target has a minimal effect on the vertex resolution of the TPC.
+Therefore the measured width of the vertex distribution can be interpreted directly as the vertex resolution of the TPC.
 The extracted vertex resolutions of each system are summarized in Table~\ref{tb:vertexresol}, with an average vertex resolution of \SI{1.2}{\centi\metre}.
-The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is really and insignificant difference.
+The difference between the measured and actual target location is 10 times smaller than the intrinsic vertex resolution of the detector and is an insignificant difference.
 
@@
 
-The impact parameter is a theoretical quantity defined as the distance, between two centers of colliding nuclei.
-Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity can be indirectly related to the estimated impact parameter of a collision \cite{impactpar}.
+The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei.
+Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity is indirectly related to the estimated impact parameter of a collision \cite{impactpar}.
 As discussed in Section \ref{sec:kyoto}, the Kyoto multiplicity array was used to experimentally trigger on central nuclear collision events.
-This is under the assumption that as the number of charged particles produced in collision is related to the overlap region of the two nuclei.
+This is under the assumption that number of charged particles produced in collision is related to the overlap region of the two nuclei.
 This is best described by a spectator-participant model of nuclear collisions, where a large fraction of the participating nucleons in the overlap region of the two colliding nuclei fragment into individual and clusters of nucleons.
@@
 \section{Track Quality Cuts}
+\label{sec:qualitycut}
 %graphic showing distance to vertex, number of clusters
@@
 %distance to vertex
-In this section we will discuss cuts which address the quality of the reconstructed track, in the following we will simply refer this as the ``track quality".
+In this section we will discuss cuts which address the quality of the reconstructed track, in the following we will simply refer this as the \emph{track quality}.
 Assuming tracks are mostly continuous in clusters, i.e.
@@
 This can happen due to discontinuities in the track due to missing hits from low energy loss, shadowing due to saturation, or possibly the software algorithm itself.
-Of course including all of the tracks would be counting the track several times over, and would lead to incorrect particle yields.
-The software will occasionally associate random disassociated clusters and construct a false track, or a so called ``ghost track".
+Including all of the tracks would be counting the track several times over, and would lead to incorrect particle yields.
+The software can also occasionally associate random disassociated clusters and construct a false track, or a so called \emph{ghost track}.
 These type of tracks contribute to the background in the PID.
-In both of the situation of the false track or in the case of a split track the false track origin cannot be traced back to anywhere near the vertex of the event, and the distance-Of-Closest-Approach (dOCA) between the track and vertex is very large.
+In both of these situations the track origin cannot be traced back to anywhere near the vertex of the event, and the distance-Of-Closest-Approach (dOCA) between the track and vertex is very large.
 
@@
 
-Since the TPC does not have spherical symmetry, there are geometric constraints and considerations that effect the efficiency of track reconstruction.
+Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that effect the efficiency of track reconstruction.
 Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}.
@@
 Figure~\ref{fig:angleEffExplanation} shows a cartoon drawing of the front of the field cage were the angles of each corner is given.
-Since the window to the field cage was slightly higher, the upper angles are slightly smaller than the lower angles.
-These angles are actually the same as the spherical $\phi$ angle, where the left and right, most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.
-While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks, where the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly.
+Since the window to the field cage was slightly higher, the upper angles are smaller than the lower angles.
+These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.
+While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks.
+Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly.
 The angular cuts applied to the data is listed by particle type in Table~\ref{tb:anglecuts}.
-The values are very similar to the simple calculation that was outlined above.
+The values are very close to the simple calculation outlined previously.
 
@@
 These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.
-We could of course include a larger region where the efficiency is small but not zero but it is good practice not to efficiency correct already small numbers with small efficiency values.
+We could of course include a larger region where the efficiency is small, but not zero, but it is not good practice to  correct using small efficiency values.
+Instead we apply a reasonable $\phi$ cut in the good areas of reasonable efficiency and correct for the regions excluded from the cut.
 The solid angle covered by the cut is written as,

### 2020-07-04 chapters/ch_results.tex
Second draft. Intro needs work

@@
 
-To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
-The correction factor $C_i$ is then defined as,
+To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}.
+The correction factor $C_i$ is defined as,
 
@@
 
-Each track that is identified as a pion as describe in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable.
-In this way, we can correct for the efficiency of the track into any observable of interest, notably transforming from the lab frame into the center-of-mass (CM) frame.
-Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame; see Section~\ref{sec:beamangle}.
-The beam angle for each event is measured by the beam tracking software and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis.
+Each track that is identified as a pion as described in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable.
+In this way, we can correct for the efficiency on a track-by-track basis, allowing for any transformation of the track of interest into any observable, notably transforming from the lab frame into the center-of-mass (CM) frame.
+Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame discussed in Section~\ref{sec:beamangle}.
+First we rotate each event so that the beam will align with the z-axis.
 Doing so makes the transformation into the CM system much simpler to describe.
 If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.
-A rotation is then applied around the vector $\hat{v}$.
+A rotation with the angle $\theta$ is then applied around the vector $\hat{v}$.
 
@@
 The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
-Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,
+Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as the Lorentz transformation matrix $A$,
 
@@
 where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$.
-The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 - M_{T}^2}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ as $\beta = -P/E$, where the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
-The CM transformed track is defined as $p^{CM} = \textbf{A}p^{Lab}$.
+The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
+The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.
 
@@
 In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerate $\pi^-$ particles, due to the positive charge of the nuclear medium.
-The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons.
-The effects of the Coulomb force are the largest effects on the spectra of the pions and will play an important role when looking at pion spectral ratios.
+The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons which limits $\pi^+$ production at low energies.
+The Coulomb force has the largest effect on the spectra of the pions and will play an important role when looking at pion spectral ratios.
+It has almost no role in the total pion yield.
 
@@
 These codes took part in a large collaborative effort in order to standarize certain common elements of each codes.
-Each comparison project simulating nuclear matter in a box type simulation with fixed, well defined, initial conditions.
+Each comparison project simulated nuclear matter in a box type simulation with fixed, well defined, initial conditions.
 Though not physical, this provided a benchmark for codes to compare to; the solutions of the total number of pions and $\Delta$ produced had an analytical solution one would expect from statistical equalibrium, providing a good benchmark.
@@
 We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
-While numerical treatments of each code are resonably similar, each code differs in its treatment of pion and $\Delta$ dynamics.
+While the numerical treatments of each code are resonably similar, each code differs considerably in their treatment of pion and $\Delta$ dynamics.
 Some codes contain modifications to how the pion behaves in nuclear matter, i.e.
@@
 The set of 7 codes represent the best codes which can simulate pions at these energies.
+
 As will be the re-occuring theme, there is a large variation between theoretical codes; greater than the variation between different symmetry energy in a particular code.

### 2020-07-14 chapters/ch_data_analysis.tex
Bill's edits and mine to results

@@
 
-The S$\pi$RITROOT software is modular code based on the FAIRROOT package, and is written in c++ \cite{fairroot}.
-The code is separated into modules which can be turned on or off depending on the users particular analysis.
-The base classes and frame work is provided by the FAIRROOT package, while the specifics of each detector need to be added.
-Here we will breifly describe the outline of the main tasks in the S$\pi$RITROOT software and then will discuss of each task later.
+The S$\pi$RITROOT software is written in c++ and is based on the FAIRROOT package \cite{fairroot}.
+It is separated into modules which can be turned on or off depending on the particular analysis.
+While the base classes and frame work are provided by the FAIRROOT package, we added the specifics of the \spirit TPC and the algorithms for track reconstrucition.
+Here we will briefly describe the outline of the main tasks in the S$\pi$RITROOT software and then will discuss of each task later.
 The level of discussion concerning the software is really a practical, base understanding, appropriate for the discussion of this dissertation.
@@
 \subsection{Decoder task}
-The decoder task reads in the raw data output from the GET electronics converting the binary data into \spirit ROOT classes using the GETDecoder \cite{getdecoder}.
-When the software is loaded, a map is loaded which maps each channel in the CoBo board onto the physical pad location on the pad plane.
+The decoder task reads in the raw data files output from the GET electronics converting the binary data into \spirit ROOT classes using the GETDecoder \cite{getdecoder}.
+When the software is loaded, a map is also loaded which maps each channel in the CoBo board onto the physical pad location on the pad plane.
 The information is stored in a pad container called the \emph{STPad} class.
 This container stores the row, layer, and the raw ADC time bucket spectrum; where the row refers to the pad numbering along the x-direction and the layer refers to the numbering along the z-direction.
-All of the pads combined form the event container called the \emph{STRawEvent} class which represents the raw event information.
+Then the information from all of the pads is combined to form the event container called the \emph{STRawEvent} class which represents the raw event information.
 
@@
 The fit of each individual pulse contains the height of the pulse and the timing information of the start of the pulse.
-We define the start of the pulse to be the location in time of 10\% of the rising edge.
+We define the start of the pulse to be the location in time when the rising edge of the pulse reaches of 10\% of its peak value.
 The height of the pulse is related to the charge induced on the pad, $Q$.
@@
 
- The function of the  helix track finding algorithm is to sort hit cloud into unique sets of hits representing tracks.
+ The function of the  helix track finding algorithm is to sort the hit cloud into unique sets of hits representing tracks.
 This is performed by a Riemann track finding algorithm where the Cartesian space is mapped onto a Riemann sphere.
@@
 The hits within a helix track are dense and the position resolution of an individual hit is not very precise.
-As mentioned in Section~\ref{sec:prf}, the TPC can achieve sub-millimeter position resolution by utilizing the PRF.
+As discussed in Section~\ref{sec:prf}, the TPC achieves sub-millimeter position resolution by combining neighboring pads into clusters using the the Pad Response Function (PRF).
 In the helix track finding task, the hits are further reduced into more precise clusters.
 A brief description of the method of clustering is illustrated in Figure \ref{fig:topview}.
-As will be seen, it is sufficient to simply cluster hits along one direction, either along the x or z-direction, which ever is most perpendicular to the local track region.
+We have found that it is sufficient to simply cluster hits along one direction, either along the x or z-direction, depending on which direction is most perpendicular to the track.
 For example, the three clusters at the bottom of Figure \ref{fig:topview} are clustered along the x-axis, and the upper three are along the z-axis.
@@
 Since we decide to cluster only in one direction, there is a natural inflection point in a track where the direction of clustering must switch.
-This depends on the crossing angle of the track $\theta$, which is defined as the angle between the tangent of the track -- at that cluster -- and the x-axis.
+This switching point is determined by the crossing angle of the track $\theta$, which is defined as the angle between the tangent of the track, projected onto the pad plane, and the x-axis -- at a particular cluster location.
 In this definition, $\theta = \ang{90}$ corresponds to the case when the track is going along the z-axis and $\theta = \ang{0}$ for a track going along the x-axis.
@@
 \includegraphics[scale=.5]{top_view_helix_ext.pdf}
-\caption{Cartoon graphic of a top down view of a fit to a track passing through several pads.
-The bolded pads and the charges $q_i$ represent the hits belonging to that pad and the clusters of the track representing the average position of the track.
-The three clusters at the bottom are clustered in the x-direction and for the upper three clustered in the z-direction.
-The estimate of the position of the avalanche is given by the track fit and the position from the center to each pad to the $\bar{x}$ position is given as $\lambda_i$.}
+\caption{Top down view of a fit to a track passing through several pads.
+Here we outline the pads in bold, where the charges $q_i$ represent the charged deposited on each pad.
+In this example, three clusters at the bottom are clustered in the x-direction and for the upper three clustered in the z-direction.
+The cluster representing each group of hits represent the average position of the track.
+The track fit represents the position of the avalanche, $\bar{x}$,  where the position from the center to each pad to the $\bar{x}$ position is defined as $\lambda_i$.}
 \label{fig:topview}
@@
 Figure~\ref{fig:prfpimData} shows black lines around the measured PRF in the data representing gates defining the acceptable PRF region.
-If less than 50\% of the hits within a cluster lie within this gate, the cluster is thrown out from the track.
-A majority of the clusters contains either 2 or 3 hits in total.
+If the charge distribution of the hits in a particular cluster is such that less than 50\% of the hits lie within this PRF gate, the cluster is thrown out from the track.
+Later, in Section~\ref{sec:expPRF}, we will describe how to define the PRF for a cluster.
 Clusters which do not follow this PRF gate usually have been corrupted in some way.
@@
 This task is repeated a second time, but this time including the vertex as a constraining point in which all tracks are refitted.
-In this way, the momentum resolution is greatly improved by adding the vertex as a constraining point; since the BDC detectors, described in Section~\ref{sec:bdc}, give such an accurate vertex location.
+In this way, the momentum resolution is greatly improved by adding the vertex as a constraining point; since the BDC detectors, described in Section~\ref{sec:bdc}, provide very accurate x and y values for the vertex location on the target.
 The energy loss of the track is calculated by the truncated mean method described in Eq.~\ref{eq:truncmean}.
@@
 It is used primarily to find if an event is on target by observing the z-distribution as will be discussed in Section~\ref{sec:vertexcut}.
-The vertex which is used in refitting the tracks as mentioned in the above section comes from the BDC interpolated vertex which is more accurate than the TPC generated vertex.
+Off target events stemming from reactions in the active collimator or with the counter gas are rejected this way.
+After such off-target events are rejected,  we use the BDC vertex and refit the tracks as mentioned in the above section; this leads to more accurate momentum values than the TPC generated vertex provided by the RAVE routine.
 
@@
 \label{sec:ggnoisesub}
-Opening the gating grid is essentially done by shorting the adjacent wires in the grid together, and allowing them to reach equilibrium as fast as possible.
-Since the impedance of both sides was not entirely matched, this caused an imbalance of charge where one side discharged faster than the other side, causing charge to bounce back and forth in the grid in an under-dampened system.
-This create oscillating residual noise early in the time bucket spectrum.
-Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum of the gating grid noise for 2000 events in a given pad.
-The gating grid noise is stable for a given pad and the mean value --shown as the red line-- can be calculated after averaging over several thousands events.
+Opening the gating grid is essentially done by shorting the adjacent wires in the gating grid together, and allowing them to reach equilibrium as fast as possible.
+Due to coupling between the gating grid, the ground plane, and other effects, the impedance of the supply cables was not perfectly matched, causing an imbalance in the current where one side discharged faster than the other side.
+This caused an underdamped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum.
+Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum after averaging over 2000 events in a given pad.
  %The ground grid shielded some of the noise, but the path to ground was on two ends and was not sufficiency for this frequency.
@@
 
-The gating grid noise profile is measured by recording experimental data, only turning off the anode wires.
+The gating grid noise profile was measured by recording experimental data, only turning off the anode wires.
 By doing so there are no signals originating from tracks in the TPC and there is only the gating grid noise, since the grid is still being triggered.
-The mean value of each pad is calculated from several thousand events of the gating grid noise profile.
-Once the mean value response in each pad is calculated, we can subtract the mean profile from the real data and remove the noise of the gating grid.
+The gating grid noise was stable for a given pad and the average value --shown as the red line-- can be calculated after averaging over several thousands of events and then removed by subtracting the profile from the real data for each pad.
 
@@
 This failure happens in regions where the derivative of the gating grid noise is large.
-In these areas the ADC value is dependent upon small jitters in the timing of the signal and as a result varies wildly.
-To take care of the residual noise in the subtracted spectra we remove all bins below 30 Tb.
-
-In the experiment the gating grid itself was quite stable in the  ${}^{132}$Sn and ${}^{124}$Sn systems and was not stable in the ${}^{108}$Sn and ${}^{112}$Sn beams, where the it broke many times.
+In these areas the ADC value is dependent upon small jitters in the timing of the signal and as a result can vary significantly from event to event.
+To remove this residual noise in the subtracted spectra we ignored the first 30 time buckets of the readout.
+
+In the experiment the gating grid itself was quite stable in the  ${}^{132}$Sn and ${}^{124}$Sn systems and was not stable in the ${}^{108}$Sn and ${}^{112}$Sn beams, where gating grid resistors broke in these earlier runs many times.
 Because of this, the gating grid noise was monitored regularly and we took gating grid noise profiles frequently, especially after a gating grid board was replaced.
@@
 \cmidrule{2-4}
-Particle &\phantom{abc} & Measured & Corrected & \% Difference & \% Difference\\
+Particle &\phantom{abc} & Measured & Corrected & \% Difference Raw & \% Difference Corrected\\
 \midrule
@@
 \cmidrule{2-4}
-Particle & Expected & Measured & Corrected & \% Difference & \% Difference\\
+Particle & Expected & Measured & Corrected & \% Difference Raw & \% Difference Corrected\\
 \midrule
@@
 t   & 1612 & 1691 & 1596   &  4.9  & -1.0\\
-${}^{4}$He   & 1613 & 1698 &  5.3 & 1595  & -1.1\\
+${}^{4}$He   & 1613 & 1698 &  1595 & 5.3 & -1.1\\
 
@@
 
-A cocktail of several light charged particles --at a known energy-- was produced (p,d,t,${}^{3}$He,${}^{4}$He) and measured in the TPC to provide a momentum calibration of the TPC.
-The magnetic and slit settings of the dipoles in the BigRIPS spectrometer were set so that the momentum resolution of the beam was known to within $\delta p/p$ < 1\%.
+A cocktail of several light charged particles (p,d,t,${}^{3}$He,${}^{4}$He) of well defined magnetic rigidity was produced was produced and measured in the TPC to provide a momentum calibration of the TPC.
+The magnetic and slit settings of the dipoles in the BigRIPS spectrometer were set so that the momentum resolution of the beam was defined to better than $\delta p/p$ < 1\%.
 Two magnetic rigidity settings were studied with no target in place.
 A thick Aluminum target was also inserted to provide a slightly lower energy calibration point for part of the lower rigidity setting, effectively creating three energy calibration points over several particle species.
-In some cases the certain particles were not able to propagate or were strongly attenuated and provided poor statistical measurements.
-
-Since the cocktail beam was a mono-energetic source of particles and the actual momentum resolution was less than 1\%, the observed momentum resolution can be interpreted as the intrinsic TPC momentum resolution.
+In some cases the certain particles were not able to propagate or were strongly attenuated and resulted in measurements of poor statistical precision.
+
+Since the cocktail beam was a mono-energetic source of particles and rigidity selection in the BigRIPS corresponding to an actual momentum resolution of better than 1\%, we can interpret the  observed momentum resolution as a measurement of the intrinsic TPC momentum resolution of the TPC for these particles.
 The momentum resolution in general depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
@@
 
-The energy loss resolution can also be directly measured a monochromatic source of each particle species have a well defined energy loss distribution.
-The energy loss resolution was measured to be 5\% for all particles over the energy setting measured and is summarized in Table~\ref{tb:momresolution}.
+The energy loss resolution can also be directly measured from the monochromatic source provided by the cocktail beam.
+Each particle species in the cocktail beam has a well defined most probable energy loss;  the energy loss resolution was measured to be approximately 5\% for all particles.
+These measurements are summarized in Table~\ref{tb:momresolution}.
 
 Since the magnetic dipole setting of the BigRIPS spectrometer is well defined, we can calculate the expected momenta of each particle species measured.
-The enegy losses through various materials in the beam line were propagated using LISE++ software \cite{lise++}.
-The measured momenta of the calibration beam differed significantly from the expected values -- up to 5\% in the high momentum calibration -- as seen in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
-This effect is attributed to inhomogenatities in the magnetic field which introduces electron drift velocities in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
+The energy losses through various materials in the beam line were propagated using LISE++ software \cite{lise++}.
+We found that the momenta of light particles in the  cocktail calibration beam differed significantly (up to 5\% ) from the expected values.
+This is shown in Tables~cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
+We could address this discrepancy by incorporating the smaller horizontal magnetic field components at the edge of the dipole and taking into account the electron drift velocities in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
 The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
@@
 
-The values under the corrected column in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}, represent the data correcting for the $\vec{E}\times\vec{B}$ effect.
-A significant improvement is seen in the high momentum setting going from around ~5\% disagreement to within ~1\% agreement in the corrected data.
-For the lower momentum settings (Tables~\cref{tb:cocktail100tar,tb:cocktail100}), protons see a slight improvement of about ~1\% where as the deuterons are over corrected in both settings.
-The level of agreement of the all corrected values is still well within the estimated momentum resolution of the TPC.
+(Tables~\cref{tb:cocktail100tar,tb:cocktail100}), protons see a slight improvement of about ~1\% where as the deuterons are over corrected in both settings.
+In general, all these light particles corrected momentum are of the order of 1\% smaller than value given by the rigidity selection in the BigRIPS separator.
+This discrepancy could reflect the accuracy of our knowledge of the magnetic field of the SAMURAI dipole.
+Without a definitive explanation for this discrepancy, we have not attempted further correction of the calculated momenta.
+We note that it is of the same order as the resolution of the measurement of these momenta.
 
@@
 \label{sec:anodeCalib}
-As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires.
+As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires due to a leak in the gating grid that has since been corrected.
 Out of all the runs used in the analysis in this thesis -- see \ref{tb:runList} -- the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all other runs.
@@
 
-Using a TPC for measurements of HICs in nuclear physics presents a different set of challenges as opposed to higher energy experiments.
+Using a TPC for measurements of HICs in intermediate energy heavy ion collisions presents a different set of challenges as opposed to higher energy experiments.
 Typically in higher energy experiments fundamental particles are produced with charge $\pm e$.
@@
 In nuclear HICs, we are interested in measured particles with charges $Ze$ where Z=1-3, and even higher in some applications.
-Recall the energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$.
+Since the energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$, the range of energy losses reflects the possible range of z values.
 Even more, in low energy nuclear HIC of intermediate beam energies -- around \SI{300}{\MeVA} -- produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
-In this case, the limited dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases, where they saturate the electronics.
+In this case, the limited dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases, leading to very large and even saturated pulses in the electronics.
 
@@
 This mitigates the complete loss of information but introduces a new problem.
-Particles which deposit large amounts of charge will have good measurements in the low gain areas, where as particles depositing minimal energy losses will lose information in the same low gain areas.
+Particles which deposit large amounts of charge will have good measurements in the low gain areas, whereas particles depositing minimal energy losses will lose information in the same low gain areas.
 The reconstruction of such tracks will suffer.
-There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics CITE HERE.
+There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics \cite{feanics,feanics2}.
 Never the less, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware.
@@
 The maximum SNR in the central pad -- before the electronics saturate -- the SNR is 800:1.
-Therefore the maximum SNR is roughly 16 times that  --800 over 50 -- of minimum ionizing particles.
+Therefore the maximum SNR is roughly 16 times larger than that of minimum ionizing particles.
 
@@
 	 \includegraphics[width=\textwidth]{PRFs_data_wcut.png}
-     \caption{PRF from $\pi^-$ tracks in experimental data. }
+     \caption{Pad Response Function of $\pi^-$ tracks in experimental data. }
      \label{fig:prfpimData}
@@
 
-It has been shown before that the amount of space charge present in the chamber is related to observables such as the distance-of-closest-approach (dOCA) of each track to the vertex point \cite{starSC}.
-In the presence of no space charge, the dOCA distribution of each track would be a centered around the true vertex location.
+It has been shown before that the amount of space charge present in the chamber is related to observables such as the Distance-Of-Closest-Approach (DOCA) of each track to the vertex point \cite{starSC}.
+In the presence of no space charge, the DOCA distribution of each track would be a centered around the true vertex location.
 Because of the left right symmetry in the TPC, the space charge affects different regions of the TPC differently, and a bias is introduced to the measured vertex location and widening the distribution to vertex of each track.
@@
 An example of the distortion map in the TPC is shown in Fig.~\ref{fig:sc_shift} where left-going tracks are shown in blue, and right-going tracks shown in green.
-The vectors show the direction of electron shift, and their magnitude have been magnified by 10 CITE HERE times to show the detail.
+The vectors show the direction of electron shift, and their magnitude have been magnified to show the detail.
 The dashed lines show the tracks under the influence of the space charge where left and right-going tracks are affected differently.
@@
 
-The inset figure of Fig.~\ref{fig:sc_shift} shows the effects of the space charge have on the x-component of the dOCA to the vertex.
-The dOCA distribution is displaced left-going track labeled by $\Delta\mathrm{V}_\mathrm{x}$, and in the opposite direction for right-going tracks.
-Figure~\ref{fig:VLR} shows the difference between right and left tracks as $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ , where  $V_x^L$ and $V_x^R$ are the most probable values of the dOCA distribution for left and right-going tracks respectively.
+The inset figure of Fig.~\ref{fig:sc_shift} shows the effects of the space charge have on the x-component of the DOCA to the vertex.
+The DOCA distribution is displaced left-going track labeled by $\Delta\mathrm{V}_\mathrm{x}$, and in the opposite direction for right-going tracks.
+Figure~\ref{fig:VLR} shows the difference between right and left tracks as $\Delta\mathrm{V}_\mathrm{LR} = \Delta\mathrm{V}_\mathrm{x}^L - \Delta\mathrm{V}_\mathrm{x}^R$ , where  $V_x^L$ and $V_x^R$ are the most probable values of the DOCA distribution for left and right-going tracks respectively.
 As the space charge increases this value increases, where in the presence of no space charge, we would expect exactly $\Delta\mathrm{V}_\mathrm{LR} = 0$.
@@
 The linear fits relating the space charge density to the beam rate for each system is shown in Fig.~\ref{fig:spacechg_br_all} where the space charge value is inferred from the measured beam rate.
-Notice that in the $\tin{112}{124}$ system is a constant value, since the variation in $\Delta\mathrm{V}_\mathrm{LR}$ rate did not span a large enough range to warrant a more detailed analysis.
+Notice that in the $\tin{112}{124}$ system we assume a constant the space charge correction, since the variation in $\Delta\mathrm{V}_\mathrm{LR}$ rate did not span a large enough range to warrant a more detailed analysis.
 
@@
 
-Adding the BDC vertex greatly improves the momentum resolution of the track fitting, but there is a systematic shift when  compared with the momentum value before using the vertex as an extra constraint.
-Since the space charge affects right and left-going tracks differently, it appears they no longer originate from the BDC vertex which is independent of the space charge effects coming from the BDC trackers.
-For tracks at polar angles of $\theta_{Lab} < \ang{40}$, the disagreement between momentum values with and without the BDC are much less because the projection of the track does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
-Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ which are more sensitive to small changes in the BDC when including it as an extra constraining point.
-After correcting for the space charge and adding the BDC point, the reconstructed momentum value agrees with or without the BDC included for both  $\theta_{Lab} < \ang{40}$ or $\theta_{Lab} > \ang{40}$, as seen in Fig.~\ref{fig:mom_S_after} and Fig.~\ref{fig:mom_L_after} respectively.
+Adding the BDC vertex greatly improves the momentum resolution of the track fitting.
+Since we know that the space charge affects right and left-going tracks differently, both tracks appear to diverge from the original vertex location appearing to no longer originate from the BDC vertex which is independent of the space charge effects.
+Without correcting for the space charge effects adding the BDC introduces a systematic shift in the momentum when comparing right and left-going tracks as compared with the momentum value before adding the BDC.
+For tracks at polar angles of $\theta_{Lab} < \ang{40}$, the disagreement between momentum values with and without the BDC are much less obvious.
+This is because for small deviations in the track -- at smaller polar angles -- leads to small deviations at the target location.
+Therefore the momentum without does not disagree as much as with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
+Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ which are more sensitive to small changes in the track -- corresponding to larger deviations at the target location.
+Therefore we get large systematic shifts in the momentum after including the BDC.
+After correcting for the space charge effects, the reconstructed momentum value agrees with or without the BDC included, for both  polar angles $\theta_{Lab} < \ang{40}$ and $\theta_{Lab} > \ang{40}$, as seen in Fig.~\ref{fig:mom_S_after} and Fig.~\ref{fig:mom_L_after} respectively.
 
@@
 There is a significant improvement in the matching of the left and right side momentum distributions for both charged pion species, in both systems, especially for the high energy tail of the pions.
-It is remarkable that the dOCA observable provided a good enough measure of the average space charge in the chamber.
-There is no reason to assume correcting the dOCA distribution would correct the momentum distribution.
+It is remarkable that the DOCA observable provided a good enough measure of the average space charge in the chamber.
+There is no reason to assume correcting the DOCA distribution would correct the momentum distribution.
 This is the strongest evidence to the success of the space charge correction.
@@
 The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
-The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation CITE HERE.
+The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation \cite{polezero}.
 If the raw ADC value at a particular time, $i$, is represented as $f_i$, the corrected pulse which differentiates out the exponential tail, $f_i^{'}$ can be expressed as,
@@
 
-The second most important observable is the distance-of-closest-approach (dOCA) to the vertex.
-Figure~\ref{fig:pocacomp} shows the dOCA distributions for the MC embedded tracks and the data, for the same range of particle species.
+The second most important observable is the distance-of-closest-approach (DOCA) to the vertex.
+Figure~\ref{fig:pocacomp} shows the DOCA distributions for the MC embedded tracks and the data, for the same range of particle species.
 The distributions were again normalized to different values to display on the same scale.
@@
 \includegraphics[width=\linewidth]{poca.png}
-\caption{Comparison of the dOCA distribution of MC embedded tracks to data.}
+\caption{Comparison of the DOCA distribution of MC embedded tracks to data.}
 \label{fig:pocacomp}

### 2020-07-14 chapters/ch_data_analysis_2.tex
Bill's edits and mine to results

@@
 \chapter{Data Analysis II}
-
-\section{Radio Isotope Beam Factory (RIBF) Facility }
-\label{sec:beam}
-%Cyclotron facility overview.
-%Samurai line overview.
-%Beam line element overview.
-%Big rips beam PID. reference 
-The primary and secondary beams were produced at the Radioactive Isotope Beam Factory (RIFB) facility at RIKEN, in Wako-shi, Japan.
-The RIBF facility starts with two primary beam types, ${}^{132}$Xe and ${}^{238}$U, which are produced by an ion-source and accelerated to progressively higher kinetic energies by a linear accelerator (RILAC), and 4 different cyclotrons (RRC, fRC, IRC, and SRC), until they reach a beam energy of \SI{345}{\MeVA}.
-Figure~\ref{fig:samuraiBeamLine} shows the later stages of the cyclotrons and the following beam lines they feed into.
-
-\begin{figure}[!htb]
-\includegraphics[width=\linewidth]{SAMURAI-beamline.png}
-\caption{Overview of the RIBF, BigRIPS, and SAMURAI beamline.}
-\label{fig:samuraiBeamLine}
-\end{figure}
-
-After the SCR, the primary beams impinge on a rotating \SI{3}{\milli\metre} Be target which produces many different species by fragmentation.
-These fragments are then separated by the BigRIPS spectrometer which is tuned to the particular secondary fragment of interest.
-This is accomplished through several dipole magnets, slits, and wedge degraders.
-The resulting secondary beam is not pure and the purity depends on the capability of BigRIPS to deliver the secondary beam of choice and the primary beam used.
-
-In these set of experiments several beams were produced with varying intensities and purities.
-Table~\ref{tb:beams} summarizes the average qualities of the 4 secondary beams produced in the two experimental campaigns where most beams were delivered with an intensity of \SI{10}{\kilo\hertz}.
-
- \begin{table*}\centering
-\ra{1.3}
-\begin{tabular}{@{}ccccc@{}}\toprule 
- Primary Beam & Secondary Beam & Energy at mid target \si{\MeVA} & Intensity \si{\kilo\hertz} & Purity (\%) \\ [0.5ex] 
- \midrule
- ${}^{238}$U   & ${}^{132}$Sn   &  269.2  &  9.5  &  54   \\
- ${}^{238}$U   & ${}^{124}$Sn   &  270.3  &  9.1  &  10  \\
- ${}^{124}$Xe  & ${}^{112}$Sn   &  270.4  &  7.6  &  48  \\
- ${}^{124}$Xe  & ${}^{108}$Sn   &  269.3  &  7.5  &  52   \\
- \bottomrule
-\end{tabular}
-\caption{Primary and secondary beam properties produced in the \spirit TPC experimental campaigns. }
-\label{tb:beams}
-\end{table*}
+As Chapter~\ref{chap:data1} covered the corrections and various calibrations to the data, this Chapter will cover the various cuts taken to improve the signal in the experimental data.
+Some cuts which identify the incoming beam and the event on target are the minimum cut in order to perform an analysis.
+The cuts pertaining to the track reconstruction quality are also required to suppress as much of the background noise as possible, most of which comes from reconstruction inefficiencies which the cuts presented here address.
 
@@
 It was shown for pions emitted in central collisions the forward and backward emission was the same \cite{jon}, proving that at least for central collisions pions are emitted isotropically to a good approximation.
+
+\subsection{Correcting Pion Spectra (Efficiency and Acceptance)}
+
+To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}.
+The correction factor $C_i$ is defined as,
+
+\begin{equation}
+C_i = \epsilon_i^{-1}.
+\end{equation}
+
+Each track that is identified as a pion as described in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable.
+In this way, we can correct for the efficiency on a track-by-track basis, allowing for any transformation of the track of interest into any observable, notably transforming from the lab frame into the center-of-mass (CM) frame.
+Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame discussed in Section~\ref{sec:beamangle}.
+First we rotate each event so that the beam will align with the z-axis.
+Doing so makes the transformation into the CM system much simpler to describe.
+If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.
+A rotation with the angle $\theta$ is then applied around the vector $\hat{v}$.
+
+Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation.
+The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
+Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as the Lorentz transformation matrix $A$,
+
+\begin{equation}
+A = \begin{pmatrix}
+1 & 0 & 0 & 0\\
+0 & 1 & 0 & 0\\
+0 & 0 & \gamma & -\beta \gamma\\
+0 & 0 & -\beta \gamma & \gamma
+\end{pmatrix},
+\end{equation}
+
+where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$.
+The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
+The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.

### 2020-07-14 chapters/ch_experiment.tex
Bill's edits and mine to results

@@
 \includegraphics[scale=.5]{tpcPrinciple.pdf}
-\caption{Operation principle of the TPC}
+\caption{Cartoon depicting the operating principles of the TPC is.}
 \label{fig:tpcPrinciple}
@@
 \includegraphics[width=\textwidth]{exploded.png}
-\caption{Overview of the \spirit TPC}
+\caption{Exploded overview of the of the \spirit TPC}
 \label{fig:tpcExplode}
@@
 \end{tabular}
-\caption{An overview of the properties of the \spirit TPC}
+\caption{Summary of general properties of the \spirit TPC.}
 \label{tb:spiritoverview}
@@
 \label{fig:fc_overview}
-\caption{FC overview}
+\caption{Overview of the field cage mechanical components.}
 \end{figure}
@@
 \label{fig:fc_overview2}
-\caption{FC overview 2}
+\caption{Schematic of the field cage.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{TPC_schematic.pdf}
-\caption{Schematic of the TPC system}
+\caption{Schematic of the resistive elements in the TPC system.}
 \label{fig:TPC_schematic}
@@
 \includegraphics[scale=.6]{FC_schematic.pdf}
-\caption{Schematic of the electric connections relevant to the Field Cage system. The strip thickness is exaggerated in the figure to show the detail}
+\caption{Schematic of the electric connections relevant to the field cage system. The strip thickness is exaggerated in the figure to show the detail.}
 \label{fig:FC_schematic}
@@
 \end{tabular}
-\caption{Wire plane properties}
+\caption{Wire plane properties.}
 \label{tb:wireplane}
@@
     \end{subfigure}
+\caption{Cartoon depiction of electron drift lines going through the various wire planes.}
 \label{fig:wires}
@@
 \includegraphics[width=\textwidth]{gg_onoff.pdf}
-\caption{On and off configurations of the gating grid.}
+\caption{Garfield simulations of the on and off configuration of the gating grid.}
 \label{fig:gg_onoff}
@@
         \includegraphics[width=\linewidth]{averageTransparency.png} 
-        \caption{Electron transparency for the conditions of all wires are the same voltage.} \label{fig:ggAvgTrans}
+        \caption{On configurations where all wires are the same voltage.} \label{fig:ggAvgTrans}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{transparencyDeltaV.png} 
-        \caption{Electrons transparency for the mode where adjacent wires have a voltage difference of $2 \Delta V$.} \label{fig:ggDeltaVTrans}
+        \caption{Off configuration where adjacent wires have a voltage difference of $2 \Delta V$.} \label{fig:ggDeltaVTrans}
     \end{subfigure}
+
+\caption{Electron transparency as measured and simulated in the on and off configurations of the gating grid.}    
 \label{fig:ggTrans}
@@
 \includegraphics[width=.5\textwidth]{GET_and_ZAP.pdf}
-\caption{•}
+\caption{Drawing of the ZAP and AsAd board which is connected to the surface pads on the backside of the pad-plane.}
 \label{fig:getzap}
@@
 \begin{equation}\label{eq:bb}
-\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2),
+\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (\ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2),
 \end{equation}
@@
 \includegraphics[width=\textwidth]{bichsel.png}
-\caption{Energy loss of a $\beta\gamma$ = 3.6 particle in Ar gas taken from \cite{bichsel}}
+\caption{Energy loss distribution of a $\beta\gamma$ = 3.6 particle in Ar gas taken from \cite{bichsel}}
 \label{fig:straggling}
@@
 \includegraphics[width=\linewidth]{driftvel.png}
-\caption{Drift velocity of electrons in P10 gas.}
+\caption{Drift velocity of electrons in P10 gas as a function of the reduced electric field.}
 \label{fig:driftvel}
@@
     \end{subfigure}
+\caption{Master triggers in each primary beam setting.}    
 \label{fig:datatrigger}

### 2020-07-14 chapters/ch_intro.tex
Bill's edits and mine to results

@@
 \chapter{Introduction}
+On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations, reinvigorating the question about the nature of dense matter in the universe.
+
 Mankind has been interested in what is the nature of the matter which composes the visible universe, and what are the constituents that make it up.
@@
 
-Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter CITE HERE with a typical size of \SI{11}{\kilo\metre} in radius.
+Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of \SI{11}{\kilo\metre} in radius.
 The di-neutron system is typically not stable in such an extreme system except for the the large gravitational force, which balances the strong forces.
-In this extreme example of dense nuclear matter, the pressures generates in the neutron star range from low densities near the crust to the dense interior which can reach up to 9$\rho_o$ CITE HERE.
+In this extreme example of dense nuclear matter, the pressures generates in the neutron star range from low densities near the crust to the dense interior which can reach up to 9$\rho_o$ \cite{neutronstar}.
 To understand these exotic forms of nuclear matter, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}, where we must describe nuclear matter over a wide range of densities.
@@
 \section{Pion Observable}
+\label{sec:pionObs}
 It is preferable to find an observable that is easier to measure experimentally than neutrons, and is more sensitive to the just the high density region.

### 2020-07-14 chapters/ch_results.tex
Bill's edits and mine to results

--- 2020-06-23 chapters/ch_results.tex
@@
 \chapter{Results}
-
-\section{Pion Spectra}
-\label{sec:pionSpectra}
-
-To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the database, from the parameters discussed in Section~\ref{sec:efficiency}.
-The correction factor $C_i$ is then defined as,
-
-\begin{equation}
-C_i = \epsilon_i^{-1}.
-\end{equation}
-
-Each track that is identified as a pion as describe in Section~\ref{sec:pid}, is then weighted by the correction factor $C_i$, when filling the histogram of any observable.
-In this way, we can correct for the efficiency of the track into any observable of interest, notably transforming from the lab frame into the center-of-mass (CM) frame.
-Before performing the transformation to the CM system, recall there is a small, but non-negligible, beam angle in the Lab frame; see Section~\ref{sec:beamangle}.
-The beam angle for each event is measured by the beam tracking software and a rotation is applied to all the tracks in an event, to align the beam angle along the z-axis.
-Doing so makes the transformation into the CM system much simpler to describe.
-If the beam direction is defined by a unit vector $\hat{b}$, we can define the rotation that rotates the beam into the z-axis as a rotation about an arbitrary vector $\hat{v} = \hat{b}\times\hat{z}$ where the angle between the two is given by $\cos \theta = \hat{b}\cdot\hat{z}$.
-A rotation is then applied around the vector $\hat{v}$.
-
-Once all the events have been rotated to align with the z-axis, transforming from the Lab to the CM frame is done by a Lorentz transformation.
-The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
-Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as,
-
-\begin{equation}
-A = \begin{pmatrix}
-1 & 0 & 0 & 0\\
-0 & 1 & 0 & 0\\
-0 & 0 & \gamma & -\beta \gamma\\
-0 & 0 & -\beta \gamma & \gamma
-\end{pmatrix},
-\end{equation}
-
-where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$.
-The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 - M_{T}^2}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ as $\beta = -P/E$, where the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
-The CM transformed track is defined as $p^{CM} = \textbf{A}p^{Lab}$.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{pionSpectra.png}
-\caption{Pion spectra. }
-\label{fig:pionspectra}
-\end{figure}
-%need to add error analysis
-%systematic error analysis
-
-Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$.
-This data marks the first time pion spectral data  has been measured at sub-threshold energies.
-In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerate $\pi^-$ particles, due to the positive charge of the nuclear medium.
-The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons.
-The effects of the Coulomb force are the largest effects on the spectra of the pions and will play an important role when looking at pion spectral ratios.
-
-%Add figure on Glauber model 
-%Add figure on comparison to FOPI data
-
-\section{Comparison to Theory}
+%\section{Comparison to Theory}
 %Add figure of Pion spectra for different theory 
@@
  The data taken has been compared with 7 commonly used transport theoretical codes.
-These codes took part in a large collaborative effort in order to standarize certain common elements of each codes.
-Each comparison project simulating nuclear matter in a box type simulation with fixed, well defined, initial conditions.
-Though not physical, this provided a benchmark for codes to compare to; the solutions of the total number of pions and $\Delta$ produced had an analytical solution one would expect from statistical equalibrium, providing a good benchmark.
-This allowed for each code to systematically go through the numerical treatment and details in each code such as intialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc.
-\cite{theoryComp1,theoryComp2}.
-These codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data.
+These codes took part in a large collaborative effort in order to standardize certain common elements of each codes \cite{theoryComp1,theoryComp2}.
+To reduce uncertainties in the pion production mechanism, each model used comparison project simulated nuclear matter in a box type simulation with fixed, well defined, initial conditions.
+Though these box calculations model an equilibrated system at finite density, which is  not replicated by experiment, the limit is theoretically clearly defined and provides a benchmark comparison between models where the solutions  yield values for the total number of produced pions and $\Delta$ that can be analytically calculated from the ansatz of  statistical equilibrium, providing a good benchmark.
+This allowed for each code to systematically go through the numerical treatment and details in each code such as initialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc.
+After satisfying these benchmark calculations, each model was used with what the authors considered the most reliable input and model assumption.
+codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data.
 We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
-While numerical treatments of each code are resonably similar, each code differs in its treatment of pion and $\Delta$ dynamics.
+While the numerical treatments of each code are reasonably similar, each code differs considerably in their treatment of pion and $\Delta$ dynamics.
 Some codes contain modifications to how the pion behaves in nuclear matter, i.e.
-\emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorptions.
-Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but are important in the production of pions CITE HERE.
-The set of 7 codes represent the best codes which can simulate pions at these energies.
-As will be the re-occuring theme, there is a large variation between theoretical codes; greater than the variation between different symmetry energy in a particular code.
-Though many of the numerical and theoretical treatments have been addressed, many of the details of pion production and dynamics have not been addressed.
-This is an extremely complex task requiring major theoretical efforts and new data.
-We expect the data in this thesis and in \cite{jon} to provide much needed experimental data in which each code may compare too.
+\emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorption.
+Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but can be important in the production of pions \cite{baoan_deltapotential,inmedPionKo,inmedPionFeng}.
+The set of 7 codes represent the best codes at the moment, which can simulate pions at these energies.
+
+It is important to note that the interaction of pions and deltas with dense matter should not be viewed as a nasty detail that a "better" probe like nucleonic elliptical flow would avoid.
+One of the central questions regarding matter at twice saturation density concerns the importance of deltas and pions as important constituents of the matter at those densities \cite{pionNS,deltaNS,awayforward}.
+Questions of the role of delta's and pion at 2$\rho_o$ are strongly connect to these potentials.
+Calculations of pion production indicate that selected pionic observables such as the energy spectra and transverse momentum distributions can be used to constrain these densities \cite{cozmaPC}.
+Obtaining such constraints are a major objective of our current work, but such constraints are ready for discussion at the present time.
+
+As will be the re-occurring theme, there is a large variation in the predicted observables, between theoretical codes; greater than the variation between different symmetry energy in a particular code.
+However, the remedy for this discrepancy is for the authors of these models to include the other physics that many of them have neglected.
+This includes a better description of delta interactions of matter, which is critical to reproducing the total pion yields, and also the momentum dependence of the isovector mean field potentials, which have a large impact on the shapes of the energy spectra.
+Though some of the elementary issues relevant to such transport theories, such as the Pauli principal and the incorporation of inelastic processes, have been addressed.
+The relevant physics include the mean field potentials for nucleons, pions and deltas.
+A primary focus is on the isovector potentials of which only the momentum independent potentials for nucleons have been varied in the calculations shown below.
+The momentum dependent nucleonic mean field potentials and the mean field potentials for delta's and pions are only included in a subset of calculations.
+Calculations have already identified that the overall pion yield are influenced by the mean field potentials for the delta's \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of for delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
+
+%Similar to the experience with the symmetric matter EoS, the next steps will be to connect these mean field to to observables and begin constraining them [cite danielewics science article and Cozma private communication]
+With the \spirit TPC, we were able to measure the pion yield without resorting to extrapolations, and were able to measure the pion energy spectra of low to high energy pions with high inefficiencies.
+In the following pages we will present the total pion yield and energy spectra of this dissertation, which complements the dissertaion of Jon Barney \cite{jon}.
+
+%to provide much needed experimental data in which each code may compare too.
 The total theoretical uncertainty is too large relative to the experimental error bars to make a constraint on the density dependence of the symmetry energy.
@@
 
+We begin with the total charged pion multiplicities for the two systems of extreme isospin asymmetry:  the neutron rich $\tin{132}{124}$ system and the neutron deficient  $\tin{108}{112}$ system.
+Though the previous FOPI data set provided significant constraints pertaining the the symmetry energy, they relied on extrapolations to low transverse momenta, $P_T = \sqrt{p_x^2 + p_y^2}$, below \SI{160}{\mega\electronvolt\per\clight} \cite{fopi}.
+In the \spirit TPC, we were able to measure pions to very low $P_T$ for rapidities $y > y{CM}$; where the rapidity is expressed along the beam direction (z) as,
+
+\begin{equation}
+y = \frac{1}{2} \ln\Big( \frac{E + p_zc}{E - p_zc}\Big).
+\end{equation}
+
+The pion rapidity in the COM system $y_{\pi CM}$ is normalized,
+
+\begin{equation}
+y_o = \frac{y_{\pi CM}}{y_{beam} - y_{CM}},
+\end{equation}
+
+where $y_{beam}$ is the beam rapidity in the lab frame and $y_{CM}$ is the CM system rapidity.
+This is keeping with the same notation of $y_o$ in Reference \cite{fopi}.
+Here $y_o = 1$ corresponds to a particle moving at beam rapidity.
+The   $P_T - y_o$ phase space of the \spirit TPC is plotted for both charged pions in the $\tin{132}{124}$ system , in Figure~\ref{ptrap_sn132}, and the $\tin{108}{112}$ system , in Figure~\ref{ptrap_sn108}.
+Good efficiency down to very low $P_T$ avoided any need for extrapolations.
+There was good acceptance of pions $y_o > 0$ -- corresponding to $\theta_{CM} < \ang{90}$ -- where we cut off the lower rapidities.
+This marks the first time that pions have been measured to such low $P_T$.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{ypt_sn132.png}
+\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{132}{124}$ system.}
+\label{fig:ptrap_sn132}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{ypt_sn108.png}
+\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system.}
+\label{fig:ptrap_sn132}
+\end{figure}
+
 The integrated pion yield for both systems, and the $\pi^-/\pi^+$ ratio, is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
-It is remarkable that the pion ratio is significantly greater than the $N/Z$ of the system, as expected in the Delta Resonance model or under the assumption of chemical equilibrium which is $\pi^-/\pi^+ = (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}.
-In the $\tin{132}{124}$ system this is a factor of 2 times greater and in $\tin{108}{112}$ the system a factor of 1.4 times greater.
-The pion ratio was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect, lowering the pion ratio as the system tends toward isospin equilibrium.
-A na\"ive interpretation would be the high density N/Z fraction is much higher than the average of the target and projectile, though, the system spends very little time in this early high density phase and there is really no time for the system to evole such that it could be enhanced in neutron fraction.
-A more likely conclusion is the mechanisms involved in pion productions via the $\Delta(1232)$ resonance is more complicated than the simple $(N/Z)^2$ relation, which was also seen in \cite{fopi}.
+Notice that the pion ratio is significantly greater than the $N/Z$ of the system.
+This is not completely unexpected; in the Delta Resonance model where $\pi^-$ come primarily from neutron-neutron collisions and $\pi^+$ from proton-proton collisions, we expect  $\pi^-/\pi^+ = (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
+For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this naive scaling assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.
+
+Due to this connection between pion ratio and the isospin asymmetry at high density, the pion ratio was proposed to be a good probe of the isovector dynamics at high density.
+was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect, lowering the pion ratio as the system tends toward isospin equilibrium.
+It is likely however, that the dynamics plays are more significant role in pion productions via the $\Delta(1232)$ resonance than the simple $(N/Z)^2$ relation, which was also seen in \cite{fopi}.
 Some codes assume the potential of the delta is just the same as the corresponding nucleon potential, as given by the iso-spin.
-Other codes introduce $\Delta$ resonance in-medium potential, which has an iso-scalar component (independent of iso-spin) and an iso-vector component.
-The nature of these two potentials is still unknown, and the role it plays in pion production has shown to be very important \cite{baoan_deltapotential}.
+While other codes -- namly TuQMD -- introduce $\Delta$ resonance in-medium potential, which has an iso-scalar component (independent of iso-spin) and an iso-vector component.
+The nature of these two potentials is still unknown, and the role it plays in pion production has shown to be very important \cite{baoan_deltapotential, cozmaPC}.
 
@@
 \includegraphics[width=.8\textwidth]{totalpiYield.png}
-\caption{Total pion yields as compared with 6 common transport codes.}
+\caption{Total pion yields as compared with 7 common transport codes.}
 \label{fig:totalpiYield}
@@
 
-Figure~\ref{fig:totalpiRatio} shows the total single pion ratio and the double ratio of the $\tin{132}{124}$ and $\tin{108}{112}$ system.
-Here, the variation between codes is much larger than the variation between the symmetry energy within a particular code.
+Figure~\ref{fig:totalpiRatio} shows the total single pion ratio, $R = Y(\pi^-)/Y(\pi^+)$, and the double ratio (DR), of the $\tin{132}{124}$ and $\tin{108}{112}$ system's single ratio, i.e.
+$DR = R_{\tin{132}{124}}/R_{\tin{124}{112}}$.
+Here, the variation between codes is much larger than the variation between the momentum dependence of the symmetry energy within a particular code.
 The symmetry energy variation (soft and stiff extremes) of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and in all codes in the double ratio.
-The data ciruclar markers and band represents the total error bar.
-Certainly the variation between symmetry energy extremes in the codes, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}.
-The small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large variation between codes.
+The ciruclar markers and band represents the total error bar in the data.
+Certainly it can be seen that the variation between symmetry energy extremes in the codes, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}, and the small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large disagreement between codes.
 
@@
 \end{figure}
-
-\section{Pion Spectral Ratio}
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{singleRatio.png}
-\caption{Single ratio spectra}
-\label{fig:SRspectra}
-\end{figure}
-
-The pion spectral ratio is a rather promising observable.
-In particular it may be more sensitive to the high density regions of the early collision.
-In theory the high energy pions are more likely to exit the nuclear medium earlier, and therefore be less prone to effects such as pion absorption and re-emission which dilute the sensitivity of the pion observable to the high density behavior.
-Also low energy pions are more likely to be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
-If we integrate the total pion yield we would combine these two regions diluting the observable.
-Instead we can construct the $Y(\pi^-)/Y(\pi^+)$ ratio as a function of the kinetic energy in the CM system with a particular focus on high energy pions.
-
-%Mention and cite appendix for systematics
-
-Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of efficiency and accuracy.
-The general hyperbolic shape arises from the Coulomb force on the two charged pion spectra mentioned in Section~\ref{sec:pionSpectra}.
-Also notice the pion ratio is smaller for less neutron poor system, as we would expect since less neutron-neutron collisions produce less $\pi^-$.
-The bin size of the last bin was increased to reduce the statistical error bars since the number of pions, especially the $\pi^+$, reaches the limits of the measured distribution.
-
-%Add figure of Theory for pion ratios
-
-\section{Pion Double Ratio}
-%Add figure of Theory for pion ratios
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{doubleRatio.png}
-\caption{}
-\label{fig:spectraDR}
-\end{figure}
-
-Another promising observable is the spectral double ratio.
-In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory, to cancel out.
-For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
 
@@
 Scaling the other systems in the \spirit TPC data leads to almost the exact same value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.
+
+\section{Pion Spectra}
+\label{sec:pionSpectra}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{pionSpectra.png}
+\caption{Pion spectra. }
+\label{fig:pionspectra}
+\end{figure}
+%need to add error analysis
+%systematic error analysis
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{pionSpectra_sn108_sum.png}
+\caption{Pion spectra $\tin{108}{112}$ system. }
+\label{fig:pionspectraSn108}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{pionSpectra_sn132_sum.png}
+\caption{Pion spectra $\tin{132}{124}$ system. }
+\label{fig:pionspectraSn132}
+\end{figure}
+
+Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$.
+This data marks the first time pion spectral data has been measured at sub-threshold energies.
+In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerates $\pi^-$ particles, due to the positive charge of the nuclear medium.
+The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons which limits $\pi^+$ production at low energies.
+The Coulomb force has the largest effect on the spectra of the pions and will play an important role when observing the general shape of the pion spectral ratio.
+Besides the initial coulomb barrier for $\pi^+$ production, the coulomb force affects only the final energy of produced pions, and has almost no role in the total pion yield.
+
+%Add figure on Glauber model 
+%Add figure on comparison to FOPI data
+
+\section{Pion Spectral Ratio}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{singleRatio.png}
+\caption{Single ratio spectra}
+\label{fig:SRspectra}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{singleRatio108_log.png}
+\caption{Single ratio $\tin{108}{112}$ system.}
+\label{fig:SRsn108}
+\end{figure}
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{singleRatio132_sum.png}
+\caption{Single ratio $\tin{132}{124}$ system.}
+\label{fig:SRsn132}
+\end{figure}
+
+The pion spectral ratio is a rather promising observable.
+In particular, it may be more sensitive to the high density regions of the early collision.
+It is generally true, that high energy pions are more likely to exit the nuclear medium  earlier, and therefore be less prone to effects such as pion absorption and re-emission; all of which dilute the sensitivity of the pion observable to the high density behavior.
+Conversly, low energy pions are more likely to stay in the nuclear medium, and be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
+Integrating these two regions to get the total pion yield dilutes the possible sensitivity the high energy pions have.
+Instead, we can construct the $Y(\pi^-)/Y(\pi^+)$ ratio as a function of the kinetic energy in the CM system with a particular focus on high energy pions.
+
+%Mention and cite appendix for systematics
+
+Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of efficiency and accuracy.
+The general hyperbolic shape arises from the Coulomb force on the two charged pion spectra mentioned earlier in Section~\ref{sec:pionSpectra}.
+Also notice the pion ratio is smaller for less neutron poor system, as we would expect since less neutron-neutron collisions produce less $\pi^-$.
+
+%Add figure of Theory for pion ratios
+
+\section{Pion Double Ratio}
+%Add figure of Theory for pion ratios
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[width=\textwidth]{doubleRatio_sum.png}
+\caption{}
+\label{fig:spectraDR}
+\end{figure}
+
+In a similar way to the total pion double ratio, we can construct the spectral double ratio, the ratio between the two system's single ratios.
+In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory --between the two systems-- to cancel out.
+For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
+Here, just as in the total double ratio, the double ratio is significantly higher than any theory.
+Here as before, the bands in the theory represent the two extremes of the symmetry energy.

### 2020-07-14 chapters/ch_summary.tex
Bill's edits and mine to results

\chapter{Summary}
The need for precise experimental pion data is paramount to understanding exotic, dense matter, such as neutron stars. Besides direct observation of these objects, the lab is the best way to probe the density dependence of the symmetry energy, and is the only way to probe different asymmetries. Pions have been proposed to be sensitive to the high density regions of these collisions, making it an ideal probe to constrain the symmetry energy. Though simulating pion production has proved to be quite challenging, it has motivated the design of high efficiency detectors such as the \spirit TPC to measure pion emission from neutron rich heavy ion collisions. 
We preformed a campaign of experiments measuring collisions of neutron rich, radioactive beams, on stable targets in 4 different configurations -- $\tin{132}{124}, \tin{108}{112}, \tin{112}{124},$ and $\tin{124}{112}$. All measured with the \spirit TPC which was able to measure pions to very low energies without the need for any extrapolations and very little assumptions.

\subsection{Results}
The pion yield, ratio, and some new observables -- namely the double ratio-- were measured to within <4\% accuracy. We observed a deviation from the  naive $\pi^-/\pi^+ ~ (N/Z)^2$ scaling expected from the $\Delta$-resonance model discussed in Section~\ref{sec:pionObs}.

\subsection{Outlook}

### 2020-07-14 macrosForThesis/PionPtRap/plotRap.C

@@ -1,0 +1,136 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void reflect(TH2D *h)
+{
+  int zerobin = h->GetXaxis()->FindBin(0.);
+  cout<<zerobin<<" "<<h->GetNbinsX()<<endl;
+  for(int i = 1; i < zerobin; i++)
+    {
+      for(int j = 1; j < h->GetNbinsY(); j++)
+	h->SetBinContent(i + 1,j,h->GetBinContent(h->GetNbinsX()-i,j));
+    }
+
+}
+
+
+void plotRap()
+{
+ gStyle->SetFrameLineWidth(6);
+ gStyle->SetLineWidth(6);
+ 
+ double events_sn132 = 807197;
+
+
+  TFile *f = TFile::Open("analysis-Sn132-merge.root");
+  TFile *g = TFile::Open("analysis-Sn108-merge.root");
+
+
+  auto pim132_ypt_ec = (TH2D *)f->Get("pim_ypt_ec");
+  auto pip132_ypt_ec = (TH2D *)f->Get("pip_ypt_ec");
+
+  auto pim108_ypt_ec = (TH2D *)g->Get("pim_ypt_ec");
+  auto pip108_ypt_ec = (TH2D *)g->Get("pip_ypt_ec");
+
+  pim132_ypt_ec->SetTitle("");
+  pim132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip132_ypt_ec->SetTitle("");
+  pip132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pim108_ypt_ec->SetTitle("");
+  pim108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip108_ypt_ec->SetTitle("");
+  pip108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+
+  make(pim132_ypt_ec);
[…]

### 2020-07-14 macrosForThesis/PionPtRap/plotRap_sn108.C

@@ -1,0 +1,138 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void reflect(TH2D *h)
+{
+  int zerobin = h->GetXaxis()->FindBin(0.);
+  cout<<zerobin<<" "<<h->GetNbinsX()<<endl;
+  for(int i = 1; i < zerobin; i++)
+    {
+      for(int j = 1; j < h->GetNbinsY(); j++)
+	h->SetBinContent(i + 1,j,h->GetBinContent(h->GetNbinsX()-i,j));
+    }
+
+}
+
+
+void plotRap_sn108()
+{
+ gStyle->SetFrameLineWidth(6);
+ gStyle->SetLineWidth(6);
+ 
+ double events_sn132 = 807197;
+
+
+  TFile *f = TFile::Open("analysis-Sn108-merge.root");
+  TFile *g = TFile::Open("analysis-Sn108-merge.root");
+
+
+  auto pim132_ypt_ec = (TH2D *)f->Get("pim_ypt_ec");
+  auto pip132_ypt_ec = (TH2D *)f->Get("pip_ypt_ec");
+
+  auto pim108_ypt_ec = (TH2D *)g->Get("pim_ypt_ec");
+  auto pip108_ypt_ec = (TH2D *)g->Get("pip_ypt_ec");
+
+  pim132_ypt_ec->SetTitle("");
+  pim132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip132_ypt_ec->SetTitle("");
+  pip132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pim108_ypt_ec->SetTitle("");
+  pim108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip108_ypt_ec->SetTitle("");
+  pip108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+
+  make(pim132_ypt_ec);
[…]

### 2020-07-14 macrosForThesis/PionPtRap/plotRap_sn132.C

@@ -1,0 +1,137 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+void reflect(TH2D *h)
+{
+  int zerobin = h->GetXaxis()->FindBin(0.);
+  cout<<zerobin<<" "<<h->GetNbinsX()<<endl;
+  for(int i = 1; i < zerobin; i++)
+    {
+      for(int j = 1; j < h->GetNbinsY(); j++)
+	h->SetBinContent(i + 1,j,h->GetBinContent(h->GetNbinsX()-i,j));
+    }
+
+}
+
+
+void plotRap_sn132()
+{
+ gStyle->SetFrameLineWidth(6);
+ gStyle->SetLineWidth(6);
+ 
+ double events_sn132 = 807197;
+
+
+  TFile *f = TFile::Open("analysis-Sn132-merge.root");
+  TFile *g = TFile::Open("analysis-Sn108-merge.root");
+
+
+  auto pim132_ypt_ec = (TH2D *)f->Get("pim_ypt_ec");
+  auto pip132_ypt_ec = (TH2D *)f->Get("pip_ypt_ec");
+
+  auto pim108_ypt_ec = (TH2D *)g->Get("pim_ypt_ec");
+  auto pip108_ypt_ec = (TH2D *)g->Get("pip_ypt_ec");
+
+  pim132_ypt_ec->SetTitle("");
+  pim132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip132_ypt_ec->SetTitle("");
+  pip132_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip132_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pim108_ypt_ec->SetTitle("");
+  pim108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pim108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+  pip108_ypt_ec->SetTitle("");
+  pip108_ypt_ec->GetXaxis()->SetTitle("y_{o}");
+  pip108_ypt_ec->GetYaxis()->SetTitle("P_{T}  (MeV c^{-1})");
+
+
+  make(pim132_ypt_ec);
[…]

### 2020-07-15 macrosForThesis/PionMultiplicity/AuAuData/pionRatioData.C
Fixed Appendix

@@ -5,7 +5,8 @@ void pionRatioData()
   gStyle->SetEndErrorSize(10);
   
   double global_e[7] = {.27,.4,.6,.8,1,1.2,1.5};
-  double global_r[7] = {4.22,2.96,2.3,2.04,1.92,1.78,1.65};
+  //  double global_r[7] = {4.22,2.96,2.3,2.04,1.92,1.78,1.65}; //b < 3fm
+  double global_r[7] = {4.74,2.96,2.3,2.04,1.92,1.78,1.65};
   double global_x[7] = {0,0,0,0,0,0,0};
   double global_y[7] = {.1,.15,.11,.11,.1,.09,.08};
   
@@ -23,11 +24,15 @@ void pionRatioData()
   //  double pip_m = .12; //as of 06/10/19
   //  double pim_m = .59; //as of 06/10/19
 
-  double pim_m = .6968; //as of 08/30/19
-  double pip_m = .1486;  //as of 08/30/19
+  //  double pim_m = .6968; //as of 08/30/19
+  //  double pip_m = .1486;  //as of 08/30/19
+
+  //  double pim_m = .6968; //as of 08/30/19
+  //  double pip_m = .1486;  //as of 08/30/19
 
   TGraphErrors *tpc = new TGraphErrors(1);
-  tpc->SetPoint(1,.27,4.22);
+  //  tpc->SetPoint(1,.27,4.22); //previous b < 3fm
+  tpc->SetPoint(1,.27,4.74); //b_o < .15
   tpc->SetPointError(1,0,.1);
   tpc->SetMarkerColor(kRed - 4);
   tpc->SetMarkerStyle(20);
@@ -35,10 +40,11 @@ void pionRatioData()
   tpc->SetMarkerSize(3);
 
   TF1 *fit_global = new TF1("fit","[0] + [1]/(x - [2])",0,3);
-  fit_global->SetParameters(1.318,5.0268e-1,9.56644e-2);
+  //  fit_global->SetParameters(1.318,5.0268e-1,9.56644e-2); //b < 3fm
+  fit_global->SetParameters(1.386,4.109e-1,1.473e-1); //b_o < .15
 
   TF1 *fit_my = new TF1("fit","[0] + [1]/x",0,2);
-  fit_my->SetParameters(1.1973,6.8879e-1);
+  //  fit_my->SetParameters(1.1973,6.8879e-1);// b < 3fm
[…]

### 2020-07-16 MSU-thesis-estee.tex
backup
Bill edits added Experiment section
Bill's edits and mine to results
Fixed Appendix
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 \documentclass[]{msu-thesis}
+
 %
@@
 \usepackage{mathtools}
+\usepackage{appendix}
 \usepackage{comment}\renewcommand{\arraystretch}{1.5}
@@
 \title{Constraining the High Density Nuclear Symmetry Energy with Pions}
-\author{Justin Estee}
+\author{Justin Brian Estee}
 \fieldofstudy{Physics} % This should be in sentence case
@@
 %
-\dedication{This thesis is dedicated to someone.}
+\dedication{}
 %
@@
 \DeclareSIUnit\eVperc{\eV\per\clight}
-\DeclareSIUnit\MeVA{\mega\eV A}
+\DeclareSIUnit\MeVA{A \mega\eV }
 \DeclareSIUnit\clight{\text{\ensuremath{c}}}
@@
 %ch_theory,
-%chapters/ch_experiment,
-%chapters/ch_data_analysis,
-%chapters/ch_data_analysis_2,
-chapters/ch_results
+chapters/ch_experiment,
+chapters/ch_data_analysis,
+chapters/ch_data_analysis_2,
+chapters/ch_results,
+chapters/ch_summary,
+chapters/ch_appendix
 }
@@
 % Your abstract goes here.  Master's 1 page max. PhD 2 page max.
+
+Recent astronomical measurements of neutron star mergers have reinvigorated questions about the nature of dense matter in the universe.
+The densities reached in these stars range from  normal nuclear densities $\rho_o$ to upwards of $9\rho_o$, where $\rho_o = \SI{1.7e14}{\gram\per\centi\metre\squared}$, extremely dense matter, which is primarily composed of neutrons.
+The symmetry energy describes the difference in binding energy between pure neutron matter and symmetric matter with equal number of neutrons and protons.
+The density dependence of the symmetry energy is related to the pressure, which supports the star preventing it from collapsing under its own gravitation force.
+For years, theory and experiment  have combined to progressively constrain the equation of state of symmetric matter, and it was in the last decade that the density dependence of pure neutron matter (the symmetry energy), had been constrained at normal nuclear densities and lower, $\rho < \rho_o$.
+Still large uncertainties still remain at densities around 2$\rho_o$ which are important to the understanding the properties of neutron stars.
+
+Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities, and to probe different proton-neutron asymmetries.
+Charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonance, in the early, dense parts of heavy pion collisions.
+Yet, existing experimental data sensitive to the symmetry energy was lacking, prompting the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pion spectra ($\pi^\pm$) with a high efficiency produced in neutron-rich heavy-ion collisions.
+This dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems.
+Also we will compare the results with the most recent transport  simulations.
+We will highlight some of the successes and hope to motivate a discussion in the theoretical community on how to better reproduce the pion phenomena in neutron-rich heavy-ion collisions, which is important for not only constraining the density dependence of the symmetry energy, but also for understanding the $\Delta$ and pion roles in neutron stars.
+
 \end{abstract}
@@
 %\addtocontents{toc}{\protect\renewcommand{\protect\cftchapterdotsep} {\cftnodots}}
+\chapterstyle{default}
 
@@
 %\include{chapters/ch_theory}
-%\include{chapters/ch_experiment}
-%\include{chapters/ch_data_analysis}
-%\include{chapters/ch_data_analysis_2}
+\include{chapters/ch_experiment}
+\include{chapters/ch_data_analysis}
+\include{chapters/ch_data_analysis_2}
 \include{chapters/ch_results}
+\include{chapters/ch_summary}
 
@@
 \appendix
-\input{chapters/ch_appendix}
+%\addappheadtotoc
+\include{chapters/ch_appendix}
 

### 2020-07-16 chapters/ch_data_analysis.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 \chapter{Data Analysis I: Calibration and Corrections}
+\label{chap:data1}
 
@@
         \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
-        \caption{Y-residual distribution for all pads before correction.} \label{fig:yoff_allBefore}
+        \caption{Before applying the timing correction.} \label{fig:yoff_allBefore}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
-        \caption{Y-residual distribution for all pads after correction.} \label{fig:yoff_allAfter}
-    \end{subfigure}
-    \caption{ }
+        \caption{After applying the timing correction.} \label{fig:yoff_allAfter}
+    \end{subfigure}
+    \caption{Y-residual distributions for all pads.}
 \label{fig:yoff}
@@
     \end{subfigure}
-    \caption{}
+    \caption{Charge readout of the pad plane, showing the max ADC in a given pad, for both the MC simulation and experimental data.}
 \label{fig:mcDataEmbedtrack}
@@
 
-    \caption{ }
+    \caption{Pion efficiencies in the $\tin{132}{124}$ system. }
 \label{fig:132_eff}
@@
 
-    \caption{ }
+    \caption{ Pion efficiencies in the $\tin{108}{112}$ system. }
 \label{fig:108_eff}

### 2020-07-16 chapters/ch_data_analysis_2.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 \section{Beam Particle Identification}
+\label{sec:beam}
 
@@
     \end{subfigure}
-\caption{}    
+\caption{Example showing the edge effect has on the cluster positions near the extreme edges of the TPC volume. }    
 \label{fig:edge}
@@
 This helped to eliminate some of the poorly reconstructed beam events.
-The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionspectra}.
+The angle rotation becomes important in the discussion of transforming back to the center-of-mass system as will be discussed in Section~\ref{sec:pionSpectra}.
 
@@
         \includegraphics[width=\textwidth]{vertex-Sn108.png} 
-        \caption{$\tin{108}{124}$ vertex distributon for all runs.} \label{fig:vertex108}
+        \caption{$\tin{108}{124}$ system.} \label{fig:vertex108}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{vertex-Sn112.png} 
-        \caption{$\tin{112}{124}$ vertex distributon for all runs.} \label{fig:vertex112}
+        \caption{$\tin{112}{124}$ system.} \label{fig:vertex112}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{vertex-Sn124.png} 
-        \caption{$\tin{124}{112}$ vertex distributon for all runs.} \label{fig:vertex124}
+        \caption{$\tin{124}{112}$ system.} \label{fig:vertex124}
     \end{subfigure}
@@
         \includegraphics[width=\textwidth]{vertex-Sn132.png} 
-        \caption{$\tin{132}{124}$ vertex distributon for all runs.} \label{fig:vertex132}
+        \caption{$\tin{132}{124}$ system.} \label{fig:vertex132}
     \end{subfigure}
-\caption{}
+\caption{Vertex distributions, around the target region, of all experimental runs for each secondary beam system. }
 \label{fig:vertexdist}
@@
          \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
-         \caption{$y=x$}
+         \caption{$\pi^-$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
          \label{fig:pim108angle}
@@
          \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
-         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \caption{$\pi^-$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
          \label{fig:pim132angle}
      \end{subfigure}
-        \caption{Three simple graphs}
+        \caption{Measured $\pi^-$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
         \label{fig:pim}
@@
          \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
-         \caption{$\tin{108}{112}$ $\pi^-$ angular spectra.}
+         \caption{$\pi^+$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
          \label{fig:pip108angle}
@@
          \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
-         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \caption{$\pi^+$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
          \label{fig:pip132angle}
      \end{subfigure}
-        \caption{Three simple graphs}
+        \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
         \label{fig:pip}

### 2020-07-16 chapters/ch_experiment.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector.
-TPCs have been employed in gases, liquid, and even solid detectors CITE HERE.
+TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}.
 In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis.
@@
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{fc_overview.png}
+\includegraphics[width=.9\textwidth]{fc_overview.png}
 \label{fig:fc_overview}
@@
 
-Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
+Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
 The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
@@
 
-\caption{}
+\caption{Gas properties of P-10 gas at 1 atm pressure.}
 \label{tb:gasprop}
@@
 The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found.
-The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Sections~\ref{sec:beamangle,sec:pionSpectra}.
+The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Section~\ref{sec:beamangle} and Section~\ref{sec:pionSpectra}.
 

### 2020-07-16 chapters/ch_intro.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 \chapter{Introduction}
-On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations, reinvigorating the question about the nature of dense matter in the universe.
+On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations \cite{ns_2017}, reinvigorating the question about the nature of dense matter in the universe.
 
@@
 The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress.
-It is for this reason that the distance between nuclei is a near constant value, and therefore the density distribution over a wide range of nuclei is remarkably constant CITE HERE.
+It is for this reason that the distance between nuclei is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the heavist to the light \cite{krane}.
 This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.
@@
 This picture was remarkably successful at describing the binding energies of normal nuclea matter at saturation density.
-The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $\epsilon/A$:
+The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $E/A$:
 
 \begin{equation}
-\frac{\epsilon}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
+\frac{E}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
 \label{eq:semiEmp}
@@
 \section{Density Dependence of the Symmetry Energy}
-In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ CITE HERE.
+
+\begin{figure}[!htb]
+\centering
+\includegraphics[scale=.3]{densityDep.png}
+\caption{Current experimental constraints on the density dependence of the symmetry energy.
+The measured data points are discussed in the text.
+The 4 polygonal and shaded regions are constraints from: 1.) pion analysis from Ref.
+\cite{xia2009,xie2013}, 2.) a separate pion analysis from Ref.
+\cite{feng2010}, 3.) ASYEOS experiment \cite{russo2011, russo2016, cozma2013} and 4.) isobaric analog states \cite{dan2014} }
+\label{fig:symDen}
+\end{figure}
+
+In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ \cite{neutronstar}.
 In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$.
-Figure~\ref{fig:symDen} shows some of the experimental constraints which have been performed by a series of independent measurements and observables \cite{awayforward}.
-Typically the density dependence of the symmetry energy can be described through an effective interaction used to describe the phenomenological observations of nucleon-nucleon interactions observed in nuclei.
-One of such interactions is the Skyrme interaction, which typically is described by a multi-parameter function which takes into account momentum dependence (through an effective mass), 2-body interactions, and correlations \cite{skyrme}.
-Several Skyrme parameterizations are shown as lines in Fig.~\ref{fig:symDen}.
+Figure~\ref{fig:symDen} shows the current status of laboratory constrains on on the density dependence of the symmetry energy.
+The red region labeled 4.) for $\rho < \rho_o$, corresponds to constraints from isobaric analog states \cite{dan2014}.
+Notice the symmetry energy is consistent across  a series of independent measurements and observables, but not at higher densities \cite{awayforward}.
+At higher densities, the two green sections correspond to constriants obtained from total pion yields, and pion ratio $\pi^-/\pi^+$, from Au + Au collisions at various beam energies \cite{fopi}.
+Region 1.) is obtained from analysis performed in References~\cite{xia2009,xie2013}.
+Region 2.) corresponds to Reference~\cite{feng2010}.
+The blue shaded region (labeled 3.) corresponds to constraints obtained through proton and neutron elliptical flow reported in References~\cite{russo2016,cozma2016,cozma2017}.
+
  Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars.
@@
 They are continuously interacting with the rest of the nuclear matter, covering a wide range of densities, diluting any symmetry energy effects and complicating the high density constraint.
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[scale=.5]{constraints}
-\caption{Experimental constraints of the density dependence of the symmetry energy taken from \cite{awayforward}}
-\label{fig:symDen}
-\end{figure}
 

### 2020-07-16 chapters/ch_results.tex
Fixed Appendix
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 Here $y_o = 1$ corresponds to a particle moving at beam rapidity.
-The   $P_T - y_o$ phase space of the \spirit TPC is plotted for both charged pions in the $\tin{132}{124}$ system , in Figure~\ref{ptrap_sn132}, and the $\tin{108}{112}$ system , in Figure~\ref{ptrap_sn108}.
+The   $P_T - y_o$ phase space of the \spirit TPC is plotted for both charged pions in the $\tin{132}{124}$ system , in Figure~\ref{fig:ptrap_sn132}, and the $\tin{108}{112}$ system , in Figure~\ref{fig:ptrap_sn108}.
 Good efficiency down to very low $P_T$ avoided any need for extrapolations.
@@
 \caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system.}
-\label{fig:ptrap_sn132}
+\label{fig:ptrap_sn108}
 \end{figure}
@@
 
+\begin{comment}
 \begin{figure}[!htb]
@@
 %need to add error analysis
-%systematic error analysis
+\end{comment}
 
@@
 
-Figure~\ref{fig:pionspectra} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems; corrected for efficiency and accounting for the solid angle of 4$\pi$.
+Figures~\ref{fig:pionspectraSn132} and \ref{fig:pionspectraSn108} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems respectively; corrected for efficiency and accounting for the solid angle of 4$\pi$.
 This data marks the first time pion spectral data has been measured at sub-threshold energies.
@@
 \end{figure}
+\end{comment}
 
@@
 
-Figure~\ref{fig:SRspectra} shows the pion spectral ratio for both systems, which was measured with a high degree of efficiency and accuracy.
-The general hyperbolic shape arises from the Coulomb force on the two charged pion spectra mentioned earlier in Section~\ref{sec:pionSpectra}.
+Figure~\ref{fig:SRsn132} and \ref{fig:SRsn108} shows the pion spectral ratio for the $\tin{132}{124}$ and $\tin{108}{112}$ system respectively, which was measured with a high degree of efficiency and accuracy.
+The general hyperbolic shape seen in both figures arises from the Coulomb force on the two charged pion spectra mentioned earlier in Section~\ref{sec:pionSpectra}.
 Also notice the pion ratio is smaller for less neutron poor system, as we would expect since less neutron-neutron collisions produce less $\pi^-$.
@@
 \section{Pion Double Ratio}
+\label{sec:doubleRatio}
 %Add figure of Theory for pion ratios
@@
 \includegraphics[width=\textwidth]{doubleRatio_sum.png}
-\caption{}
+\caption{Comparing the double spectral ratio, $R_{\tin{132}{124}}/R_{\tin{108}{112}}$, to the 7 theoretical transport simulations.}
 \label{fig:spectraDR}

### 2020-07-16 chapters/ch_summary.tex
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.

@@
 Though simulating pion production has proved to be quite challenging, it has motivated the design of high efficiency detectors such as the \spirit TPC to measure pion emission from neutron rich heavy ion collisions.
-We preformed a campaign of experiments measuring collisions of neutron rich, radioactive beams, on stable targets in 4 different configurations -- $\tin{132}{124}, \tin{108}{112}, \tin{112}{124},$ and $\tin{124}{112}$.
-All measured with the \spirit TPC which was able to measure pions to very low energies without the need for any extrapolations and very little assumptions.
+We have preformed a campaign of experiments measuring collisions of neutron rich, radioactive beams, on stable targets in 4 different configurations -- $\tin{132}{124}, \tin{108}{112}, \tin{112}{124},$ and $\tin{124}{112}$, all measured with the \spirit TPC.
 
@@
 The pion yield, ratio, and some new observables -- namely the double ratio-- were measured to within <4\% accuracy.
-We observed a deviation from the  naive $\pi^-/\pi^+ ~ (N/Z)^2$ scaling expected from the $\Delta$-resonance model discussed in Section~\ref{sec:pionObs}.
+Marking the first time pions have been measured in the sub-threshold region.
+We observed a deviation from the  naive expectation in a simple $\Delta$ resonance model, where $\pi^-/\pi^+ ~ (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
+This is somewhat expected, as the mean field potential contributes to the chemical potential of pions in a thermal statistical equilibrium simple model.
+Although it is clear, that other effects such as the relatively unknown isoscalar and isovector $\Delta$ potential plays an important role.
+The pion spectral ratio was also measured with a high degree of accuracy, also providing much needed information about the dynamics of pions, even to very low pion energies, where previous experiments failed to measure.
+This is a promissing observable since high energy pions are more likely to exit the nuclear matter sooner and unaffected by re-scattering and adsorption-re-emission processes, which dilute the observable.
+In this way high energy pions may be more sensitive to the earlier high density timescales of the collision, whereas the low energy pions suffer from other influences, namely the Coulomb potential, pion optical potentials, and the $\Delta$ potential.
 
 \subsection{Outlook}
+Theoretical collaboration, which has made considerable progress improving details of the calculation, will further improve in light of this new data.
+It is clear that a serious effort must be made to include effects such as the symmetry potentials of pions and deltas.
+Which most codes do not.
+It also appears to be possible to make first constraints on the generally unknown $\Delta$ potential by fitting the pion yield data \cite{cozmaPC}.
+Making such a constraint would allow for the possibility to start making constraints on the density dependence of the symmetry energy.

### 2020-07-17 chapters/ch_appendix.tex
backup
Fixed Appendix
Almost final draft. Before Bill's edits for Intro, Results, abstract, summary.
Added bills edits working on appendix figures

@@
-\subsection{$\Delta$ Resonance Production and Decay Channels}
+%\setcounter{figure}{0} % Perhaps
+\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\thetable}{A\arabic{figure}}
+
+\section{$\Delta$ Resonance Production and Decay Channels}
 \label{appen:deltadecay}
@@
 \end{split}
-\label{eq:deltaProduction}
+\label{eq:pionProduction}
 \end{equation}
@@
 
-\subsection{Runs analyzed in this data}
+\section{Runs analyzed in this data}
 
@@
 
-\subsection{Pion Yield Theory}
+\section{Pion Yield Theory}
 \label{tb:pionyieldTheory}
@@
 \label{sec:cutvar}
-The particular track quality cuts described in Section~\ref{sec:qualitycut} are there to reduced the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra.
-The best set of values were found for all the data sets which include the charged particle multiplicity of the event, dOCA, and the number of clusters cut.
-By varying the cuts in both the data and the MC efficiency, we can evaluate the analysis to see if there are any systematic dependencies on the observable we plot.
-But varying the cuts undoutably means as the cut get tighter, and less data is included, the statistical error bar will increase making it difficult to compare to more loose cuts.
-This is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}.
-Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts and the uncorrelated error bars, then there is some systematic trend that either exists due to physics or some error or miscalculation in the analysis.
-It is usually not recommended to estimated systematic error bars using this method, but sometimes is the only way.
+
+We employ a variety of track quality cuts described in Section~\ref{sec:qualitycut} to reduce the contamination from poorly reconstructed tracks which contribute to the background in the PID spectra.
+The best set of values were found for all the data sets which include the charged particle multiplicity of the event, the distance of closest approach to the vertex (DOCA), and the minimum number of clusters required for a track to be analyzed.
+By varying the cuts in both the data and in our MC efficiency calculations, we can evaluate the analysis to see how stable are the results with respect to these cuts.
+It is intuitive that while varying the cuts we will find that as the cut get tighter, and less data is included, the statistical error bar will increase.
+This problem is mitigated by looking at the uncorrelated error described in \cite{dataAnalysis}.
+Here if the systematic trend of the observable is much larger than the statistical error bars on the default cuts, with respects to the uncorrelated error bars, then there truly is some systematic trend that either exists due to physics or some error or miscalculation in the analysis.
+It is usually not recommended to estimated systematic error bars using this method, but sometimes it is the only way.
 
 Here we will discuss the particular analysis using the total pion ratio as the example.
-Figure~\ref{fig:totalRatioError} shows the total pion as a function of several cut variations, where one cut is varied and all the others are held constant.
-Three cuts are of particular interest, the event multiplicity, dOCA, and the number of cluster cut.
-Figure~\ref{fig:clustTotalRatio} shows the variation as the number of cluster cut is varied.
-The default cut used is > 20 clusters which is represented by the middle point.
+Figure~\ref{fig:totalRatioError} shows the total pion ratio as a function of several cut variations, where one cut is varied and all the others are held constant.
+Three cuts are of particular interest, the event multiplicity, DOCA, and the number of cluster cut.
+
+ Figure~\ref{fig:pim_clustVar_NoEC} shows the variation as the number of clusters cut is varied before the efficiency analysis.
+The default cut for the number of clusters in a track is $N_c$ > 20 clusters which is represented by the middle point.
 The red bars represent the statistical error of this default value.
 The error bars on the other points are the uncorrelated error bars as described by the prescription in \cite{dataAnalysis}.
-Here it is clear that there is a systematic bias in the pion ratio as a function
+Here it is clear that there is a systematic bias in the $\pi^-$ yield as we increase the number of cluster cut.
+This expected since cutting on the number of clusters eventually throws away short tracks that may be in the large angular regions of acceptance, thus biasing the yield.
+As seen in Figure~\ref{fig:pim_clustVar_EC}, this is corrected mostly be the efficiency analysis, though there seems to be some residual over correction for above 20 clusters and under correction under it, but still approximately within the statistical error bar of the default cut value of 20.
+
+When looking at the variation of the DOCA cut, we see an even better story.
+Before the efficiency correction in Figure~\ref{fig:pim_cutVarPOCA_NoEC}, there is very little systematic dependence, except maybe for higher values of DOCA.
+After efficiency correction we really see no significant variation and can safely claim there are no systematic dependencies on this particular cut.
+
+In Figure~\ref{fig:pim_multCutVar}, we see  a significant systematic variation of the $\pi^-$ yield as a function of the event track multiplicity.
+Since the event multiplicity is related to how central of a collision we had, this is purely a physics effect that is expected.
+This is because as the collision becomes more central, more nucleons participate in the collision due to geometric overlap.
+As the participating nucleons increases, there is a higher chance to make more pions, therefore the yield scales with the track multiplicity, which is ultimately related to the impact parameter as discussed in Section~\ref{sec:impact}.
+Therefore we should not associate any systematic error due to this effect.
+
+There is a similar discussion that is made for the $\pi^+$ yield in the same cut variations; for the multiplicity cut in Figure~\ref{fig:pip_multCutVar}, the DOCA cut in Figure~\ref{fig:pip_cutVarPOCA}, and the cluster cut in Figure~\ref{fig:pip_clustVar}.
+
+The analysis becomes more interesting when analyzing the total pion ratio for the same cut variation analysis performed above.
+Where the total pion ratio $\pi^-/\pi^+$ is plotted as a function of the multiplicity cut in Figure~\ref{fig:ratio_multCutVar}, the DOCA cut in Figure~\ref{fig:ratio_cutVarPOCA}, and the cluster cut in Figure~\ref{fig:ratio_clustVar}.
+Neglecting the multiplicity cut variation, which we know is derived from physical effects, it appears the pion ratio cancels out most of the systematic variations we saw previously in the single particle yields.
+This was expected, as the ratio tends to cancel systematic effects in the detector system.
+This is definitive proof of this in action.
+The systematic trends are less significant than the statistical error bars given in the red region for the default cuts, except maybe for a cluster cut of $N_c > 26$ which is already at the extreme end of the cut, since we know that it cuts into the particle acceptance.
 
@@
          \caption{No efficiency correction.}
-         \label{fig:multCutVar}
+         \label{fig:pim_multCutVar_NoEC}
      \end{subfigure}
@@
          \caption{Efficiency corrected.}
-         \label{fig:multCutVar}
+         \label{fig:pim_multCutVar}
      \end{subfigure}
@@
 \caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
-\label{fig:cutVar}
+\label{fig:pim_multCutVar}
 \end{figure}
@@
          \caption{No efficiency correction.}
-         \label{fig:pip_multCutVar}
+         \label{fig:pip_multCutVar_NoEC}
      \end{subfigure}
@@
 \caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
-\label{fig:cutVar}
+\label{fig:pip_multCutVar}
 \end{figure}
@@
 \caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
-\label{fig:pim_clustVar}
+\label{fig:pip_clustVar}
 \end{figure}
@@
          \caption{No efficiency correction.}
-         \label{fig:ratio_multCutVar}
+         \label{fig:ratio_multCutVar_NoEC}
      \end{subfigure}
@@
 \caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
-\label{fig:cutVar}
+\label{fig:ratio_multCutVar}
 \end{figure}
@@
 \caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
-\label{fig:pim_clustVar}
+\label{fig:ratio_clustVar}
 \end{figure}
@@
 \caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}
-\label{fig:pip_cutVarPOCA}
+\label{fig:ratio_cutVarPOCA}
 \end{figure}
 
-\chapter{Second appendix}
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{clusterSpectraSn132}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
+	\label{fig:clusterSpectralSn132}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{pocaSpectraSn132}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
+	\label{fig:docaSpectralSn132}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{multSpectraSn132}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}Sn$ cut.}
+	\label{fig:multSpectralSn132}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{clusterSpectraSn108}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
+	\label{fig:clusterSpectralSn132}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{pocaSpectraSn108}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
+	\label{fig:docaSpectralSn132}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{multSpectraSn108}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{108}Sn$ cut.}
+	\label{fig:multSpectralSn132}
+\end{figure}
+
+%\chapter{Second appendix}

### 2020-07-17 chapters/ch_experiment.tex
Added bills edits working on appendix figures

@@
 \chapter{Experiment}
+A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector.
+TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}.
+In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Thesis, along with the rest of the experimental setup.
 
@@
 %explain in breif the basics of a time projection chamber 
-A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector.
-TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}.
-In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this thesis.
 Figure~\ref{fig:tpcPrinciple} depicts the interior of a gaseous TPC and its field cage.
@@
 
-where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction  of a given roaw the PRF can be written as,
+where $\lambda_{x_o} = x_o - x_o\textprime$; in a similar manner for the situation we cluster along the z-direction  of a given row the PRF can be written as,
 
@@
 All materials and glues of the TPC were selected as low out-gassing materials.
-Several materials (that are common place in nuclear labs), such as vacuum grease, viton o-rings, all out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time.
+Several materials (that are common place in nuclear labs), such as vacuum grease, Viton o-rings, all out-gas organic chemicals into the counter gas which damage the TPC by permanently lowering the gain over time.
 The organic molecules responsible are difficult to identify exactly, but lists of good and bad materials are well known in the literature from experiments.
@@
 The resolution of the initial angle the beam enters the SAMURAI dipole was estimated to be \SI{0.64}{\milli\radian} and \SI{0.24}{\milli\radian} for the two angles defining the beam vector \cite{jon}.
-The initial angle, energy, and charge of the beam is then propogated through the magnetic field map until the angle on target is found.
+The initial angle, energy, and charge of the beam is then propagated through the magnetic field map until the angle on target is found.
 The angle on target though small will be important later for transforming back into the center of mass system as will be seen in Section~\ref{sec:beamangle} and Section~\ref{sec:pionSpectra}.
@@
 We required that the gating grid be opened quickly to not lose electrons from portions of the tracks that pass close to the gating grid.
-If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicty >4 signal, and neither a Katana Veto signal nor an active target signal.
+If the DAQ was not busy, we therefore generated a trigger once we had a start detector signal, a Kyoto multiplicity >4 signal, and neither a Katana Veto signal nor an active target signal.
 This was referred to as the Fast Trigger, and it triggered the opening of the gating grid, which took about 350 ns.
@@
 If the last beam particle came within xx us after the fast trigger, electrons from this beam particle could pass through the gating grid.
-Such data would be useless, so we closed the gating grid and stopped the data equisition with a fast clear signal.
+Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal.
 Additional fast clears could also initiated if the Katana trigger box did not generate a trigger.
@@
 Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal.
-The KATANA trigger, Kyoto Muliplicity, and Start Counter together triggered the DAQ.
+The KATANA trigger, Kyoto Multiplicity, and Start Counter together triggered the DAQ.
 

### 2020-07-17 chapters/ch_intro.tex
Added bills edits working on appendix figures

@@
 \chapter{Introduction}
-On August 17, 2017, a collision of two presumptive neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with several other Astronomical telescopes, marking a new era of advanced multi-messenger observations \cite{ns_2017}, reinvigorating the question about the nature of dense matter in the universe.
+On August 17, 2017, a collision of two  neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with many other astronomical observatories, marking a new era of advanced multi-messenger observations \cite{ns_2017}, dramatically increasing already very strong interesting in  questions concerning the nature of dense matter in the universe.
 
-Mankind has been interested in what is the nature of the matter which composes the visible universe, and what are the constituents that make it up.
-The most significant discovery on the study of fundamental sub-atomic particles was Rutherford's discovery in 1911 in which he concluded that atomic nuclei are composed of a dense nucleus surrounded by an electron cloud.
-Later it was discovered that the nucleus is composed of nucleons, a positive charge (protons) and a neutral charge (neutrons).
-The nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it incredibly dense.
+Mankind has been interested in the nature of the matter which composes the visible universe, and what are its "constituents" or building blocks.
+The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie.
+His studies led him to conclude that atoms consist of a dense positively charged atomic nucleus surrounded by a dilute electron cloud.
+Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to positively charged protons as the basic constituents of nuclear matter.
+Remarkably, the nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.
 
-Without a balancing force the Coulomb force between protons would render the nucleus unstable.
-This balancing force is called the strong nuclear force which exerts a large attractive force over a very short range.
-The strong force itself is the fundamental interaction between the fundamental quarks which make up the nucleons.
-At low energies the quark structure of nucleons is less important and nucleons can be thought of as fundamental particles.
-The strong force is attractive only for a small regions approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress.
-It is for this reason that the distance between nuclei is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the heavist to the light \cite{krane}.
-This density is referred to as the  \emph{saturation density}, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.
+Without an attractive force between nucleons, the Coulomb force between protons would render the nucleus unstable.
+It is the strong nuclear force which exerts a large attractive force  between nucleons over a very short range.
+This strong force is a manifestation of the fundamental interactions between the quarks which constitute each nucleon.
+Three quarks of two types or \emph{flavors} labeled \emph{up} and \emph{down} with nearly equal masses are needed to form each nucleon.
+At low energies the internal quark structure of nucleons is less important and nucleons can be thought of as fundamental particles.
+The strong force between nucleons is attractive only for a small distance of approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress.
+It is for this reason that the distance between any pair of nucleons is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the lightest to the heaviest nuclei \cite{krane}.
+This density is referred to as the  \emph{saturation density}, since the central density seems to saturate at a common value, namely, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.
 
-Due to the nature of the strong force, nuclei can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility.
-This picture was remarkably successful at describing the binding energies of normal nuclea matter at saturation density.
-The Bethe-Weizsacker semi-empirical formula \cite{awayforward}, predicts the binding energy as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$, where the binding energy per nucleon is $E/A$:
+Due to the nature of the strong force, nuclear matter can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility.
+This analogous picture treating nuclei as liquid droplets is remarkably successful at describing the nuclear binding energy $E_{B}$, which is the energy it would take to disassemble a nucleus into its constituents.
+The Bethe-Weizsacker semi-empirical mass formula \cite{awayforward}, predicts that $E_{B}$ can be expressed as a function of the number of neutrons $N$, protons $Z$, and total nucleons $A = Z + N$:
 
 \begin{equation}
-\frac{E}{A} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A\frac{(N - Z)^2}{A} + ...
+E_{B} = a_vA - a_s A^{2/3} - a_c \frac{Z^2}{A^{1/3}} - a_A \frac{(N - Z)^2}{A}
 \label{eq:semiEmp}
@@
 
-Since the strong force makes the inter-nucleon distance approximately constant, the volume of the nucleus is related to the number of total nucleons, therefore the volume term $a_V$ is proportional to $A$.
-There are several correction terms accounting for the surface, $a_s$, since nucleons near the surface have fewer surrounding neighbors as nucleons inside, and the coulomb term which is related to the typical coulomb force which scales like the radius of the nuclei $R^{-1}$; which scales as $A^{1/3}$ (since volume scales like A).
-The asymmetry term, $a_A$, is related to the cost in energy to become more neutron or proton rich; it is typically referred to as the \emph{symmetry energy}.
-This originates from Pauli blocking where it is more energetically favorable to form neutron-proton pairs since their isospin numbers are different.
-The di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, where as the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more energetically favorable.
+Since the strong force makes the inter-nucleon distance approximately constant, the volume the nucleus is related to the total number of nucleons and the volume, $A$.
+Therefore the first \emph{volume} term $a_VA$ is proportional to $A$.
+Since the  constituent quarks that make of the nucleons are of nearly equal mass, there is a fundamental symmetry of the strong interaction which the strong interaction between any pair of nuclei is very similar, regardless of the particular flavor of quark composition.
+This symmetry is called \emph{isospin symmetry}.
+The term isospin was chosen because this symmetry was similar to the spin symmetry of electrons, which have spin up and down projections but are otherwise indistinguishable, unless in the presence of a symmetry breaking field.
+The isospin mathematics are is identical to the spin mathematics describing where nucleons have an isospin of 1/2 and with projections of  +1/2 for protons or -1/2 for neutrons.
 
-Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of \SI{11}{\kilo\metre} in radius.
-The di-neutron system is typically not stable in such an extreme system except for the the large gravitational force, which balances the strong forces.
-In this extreme example of dense nuclear matter, the pressures generates in the neutron star range from low densities near the crust to the dense interior which can reach up to 9$\rho_o$ \cite{neutronstar}.
-To understand these exotic forms of nuclear matter, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}, where we must describe nuclear matter over a wide range of densities.
+Continuing on with the Bethe-Weizacher formula, we see that there are several correction terms accounting for such effects like the surface, $a_s$, where nucleons near the surface have fewer surrounding neighbors as nucleons inside.
+Since the volume of the nuclei scales as $A$, radius of the nuclei would scale as $A^{1/3}$.
+Also there is a correction accounting for the Coulomb force, $a_c$, which is proportional to the radial extent of the nuclei $R^{-1}$; which scales as $A^{1/3}$.
+The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}.
+Although the strong force does not distinguish between protons and neutrons the overall wavefunction is made up of a combination of the spacial, spin, and isospin wavefunctions.
+Since nucleons are fermions,the total wavefunction must be antisymmetric under particle exchange.
+Because of this, each spacial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin asymmetry.
+Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich.
+Thus the asymmetry term above is proportional to the factor $\frac{(N - Z)^2}{A}$ which comes directly from Pauli blocking.
+While Eq.~\ref{eq:semiEmp} describes normal nuclear matter, at saturation density, quite well, a more general description must be made when going to higher or lower densities.
+
+% This state cannot be occupied by two neutrons or two protons because nucleons are spin 1/2 fermions and the Pauli exclusion principal i.e.
+"Pauli blocking" only allows one fermion of a specific type (i.e.
+neutron or proton) per state.
+The asymmetry term above is proportional to A and the factor $\frac{(N - Z)^2}{A^2}$ comes directly from Pauli blocking.
+This term is present even when one neglects the nuclear force and approximates the energy of nucleus by that of a Fermi gas.
+If you add a neutron to a Fermi gas in its ground state, it goes into the lowest energy vacant neutron state.
+If there is a vacant proton state at lower energy it does not matter because the corresponding neutron state is already filled.
+This simple consideration accounts for roughly 40 percent of $a_A$.
+The remaining 60 percent primarily occurs because the nuclear force is most attractive between nucleons in the same state, a configuration that is only allowed if one of the nueleons is a proton and the other is a neutron.
+In other words, the di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, where as the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more strongly bound.
+
+Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately \SI{11(2)}{\kilo\metre}, though they weigh about 1.5 times the mass of our sun.
+Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the balancing outward force arising from the strong force which exerts a large outward pressure arising from the symmetry energy at high densities.
+The densities reached in a neutron star's interior is a matter of debate.
+In nucleonic models it is typically 2-4$\rho_o$, but in some quark models for the interior, it can range up to  9$\rho_o$ \cite{neutronstar}.
+To describe matter at such high densities, the energy density of the system must be described in a more general way than Eq.~\ref{eq:semiEmp}.
+
+%Unlike nuclei, symmetric nuclear matter with equal proton and nucleon densities is not the ground state.
+This occurs because each proton must be accompanied by an electron in order to locally maintain charge neutrality.
+This means that the electron density must equal the proton density, which for symmetric matter at normal density puts the electron Fermi energy at xxx MeV.
+Because the proton fraction in the interior of a neutron star is typically predicted to be less than 10 percent, the symmetry energy in a neutron star is huge and it gets even larger as one goes into the interior where the density is higher.
 
@@
 
-The Equation of State (EoS) of nuclear matter can be calculated by fundamental thermodynamic relations,
+The Equation of State (EoS) of nuclear matter can be calculated from fundamental thermodynamic relations,
 
@@
 
-The picture that is built up here is the gravitational force is attempting to compress and in-compressible matter, and the fundamental forces of the strong force build up a competing pressure which is related to the derivative w.r.t density of the symmetry energy \cite{tovEq}.
+The outward pressure generated  by the dense neutron rich matter is related to the derivative -- with respects to density  -- of the symmetry energy \cite{tovEq}.
+By studying the density dependence of the symmetry energy we can better understand the properties of neutron stars and the dynamics of heavy ion collisions.
 
@@
 
-In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 9\rho_0$ \cite{neutronstar}.
-In contrast, the symmetry energy has only been experimentally constrained for densities at or below $\rho_0$.
+In the last couple decades, the symmetric term of Eq.~\ref{eq:energyEos} has been determined for a wide range of densities ranging from $\rho_0 - 4.5\rho_0$ \cite{scienceEos}.
+In contrast, the symmetry energy has only been accurately constrained for densities at or below $\rho_0$ and only preliminary constraints have been obtained at higher densities.
 Figure~\ref{fig:symDen} shows the current status of laboratory constrains on on the density dependence of the symmetry energy.
 The red region labeled 4.) for $\rho < \rho_o$, corresponds to constraints from isobaric analog states \cite{dan2014}.
-Notice the symmetry energy is consistent across  a series of independent measurements and observables, but not at higher densities \cite{awayforward}.
-At higher densities, the two green sections correspond to constriants obtained from total pion yields, and pion ratio $\pi^-/\pi^+$, from Au + Au collisions at various beam energies \cite{fopi}.
-Region 1.) is obtained from analysis performed in References~\cite{xia2009,xie2013}.
-Region 2.) corresponds to Reference~\cite{feng2010}.
+Notice that the symmetry energy is consistent across  a series of independent measurements and observables, but not at higher densities \cite{awayforward}.
+At higher densities, the two green sections correspond to constriants obtained from total pion yields, and the total pion ratio $\pi^-/\pi^+$, from Au + Au collisions at various beam energies \cite{fopi}.
+Region 1.) is obtained from analysis performed in References~\cite{xia2009,xie2013} while Region 2.) corresponds to Reference~\cite{feng2010}, giving conflicting results from the same data set.
 The blue shaded region (labeled 3.) corresponds to constraints obtained through proton and neutron elliptical flow reported in References~\cite{russo2016,cozma2016,cozma2017}.
-
- Though most of the functional forms satisfy the experimental constraints at low densities there is a considerable uncertainty at high densities, which are more relevant to neutron stars.
-The goal of the nuclear EoS community over the last decade have been particularly focused on constraining the symmetry energy at these high densities.
+There is still a considerable uncertainty in the symmetry energy at high densities, which are more relevant to neutron stars.
+It has been the goal of the nuclear EoS community over the last decade to constrain the symmetry energy at these high densities.
 
@@
 When two nuclei collide in a collision, in the very early stages they compress to form a high density region where the nuclei overlap.
-This momentary density can reach up to $3\rho_0$ depending on the incident beam energy.
+This momentary density can reach up to $3\rho_0$ depending on the incident beam energy, but only a small volume attains such densities.
 HICs also provide the only way we can probe the isospin asymmetry dependence of the nuclear EoS.
@@
 If the density dependence of the symmetry energy is positive at high densities the symmetry energy would in general force neutrons out of the high density system.
-Where as if the derivative was negative the symmetry energy would attract neutrons.
-The curvature of the symmetry energy at a particular density is commonly referred to as \emph{soft} in the case the curvatures turns over, and \emph{stiff} in the case of a strong positive curvature.
-It is the symmetry energy pressure that is driving the relative dynamics of neutrons and protons, where a strong symmetry energy expels neutrons from dense matter and vice versa.
+If the symmetry energy decreases at high densities, making the derivative  negative, then the symmetry energy term would tend to attract neutrons.
+The symmetry energy is referred to as \emph{soft} if the second order term in the Taylor expansion -- at saturation density-- is negative, and \emph{stiff} in the case of a curvature that is close to zero or positive.
+It is the pressure coming from the symmetry energy that is driving the relative dynamics of neutrons and protons, where a stiff symmetry energy expels neutrons from dense matter and vice versa for as soft symmetry energy.
+
 By measuring protons and neutrons, we can see signatures of the symmetry energy in the final spectra of these particles.
 There are two experimental challenges.
-First,  measuring neutrons can be quite challenging and space and acceptance is limited by large neutron wall arrays.
+First,  measuring neutrons precisely can be quite challenging.
+The existing technology of neutron time of flight arrays requires long flight paths and large, inefficient detectors.
+Inevitably, the acceptance is limited by cost and the space required for large neutron wall arrays.
 Lastly, though the overlap region of the two nuclei temporarily reaches a high density, the nucleons which participated in this region continue to evolve throughout the collision, even into regions of lower densities until they reach their final state.
-They are continuously interacting with the rest of the nuclear matter, covering a wide range of densities, diluting any symmetry energy effects and complicating the high density constraint.
+They are continuously interacting with the rest of the nuclear matter, covering a wide range of densities, which dilutes any information about the symmetry energy at at high densities.
 
@@
 \label{sec:pionObs}
-It is preferable to find an observable that is easier to measure experimentally than neutrons, and is more sensitive to the just the high density region.
-Pions are produced though an intermediate step via $ NN \leftrightarrow \Delta$, where nucleon-nucleon collisions form an excited $\Delta$(1232) baryon resonance from one of the nucleons, which then decay decay shortly after into a pion $\Delta \leftrightarrow \pi N$.
-The threshold for $\Delta$ resonance production, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
-In large nuclei, the internal motion of the nucleons is substantial and non-negligible.
+Another approach is to find an observable that can be detected with high efficiencies, unlike the neutron, and is more sensitive to the just the high density region.
+For these reasons pions are a very interesting obserabvle, as they are produced though an intermediate step via production of $\Delta$ resonances.
+Here the reaction $ NN \leftrightarrow \Delta$ takes place, where nucleon-nucleon collisions form an excited $\Delta$(1232) baryon resonance from one of the nucleons, which then decays shortly thereafter into a pion $\Delta \leftrightarrow \pi N$.
+The threshold for $\Delta$ resonance production in nucleon-nucleon collisions, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
+In large nuclei, the internal motion of the nucleons is substantial and this modifies the production rates.
 In the Fermi gas model, nucleons are arranged filling up higher energy levels up until they reach the Fermi energy.
-This extra energy allows for $\Delta$ production even at sub-threshold beam energies \cite{fermiEnergy}.
+The associated Fermi velocity can add to the beam velocity and actually be above the threshold condition, even when the beam velocity is below the threshold \cite{fermiEnergy}.
 
-It has been shown that most of the $\Delta$'s are produced in the early dense regions of the collision \cite{mingzhang}.
-Figure~\ref{fig:deltaProduction} shows the average local density (c) which $\Delta$'s are produced and the number in the system (b), as a function of time in the simulation of Au + Au collisions at \SI{400}{\MeVA}.
-Panel (a) shows the density distribution of the density at the moment of creation for $\Delta$'s.
+Nuclear collisions are typically calculated via semi-classical theories based on a semi-classical extension of the Boltzmann equation or by semi-classical modifications of molecular dynamics \cite{bertsch}.
+These theories predict that most of the $\Delta$'s are produced in the early, dense regions of the collision \cite{mingzhang}.
+Figure~\ref{fig:deltaProduction} shows the average local density in panel (c)  in which $\Delta$'s are produced and the total  number of $\Delta$'s produced in the system (b), as a function of time in the simulation, for Au + Au collisions at \SI{400}{\MeVA}.
+Panel (a) shows the distribution of density at moment of creation for each $\Delta$.
 Since the average lifetime of the $\Delta$  is $\tau_{\Delta} = \SI{1.7}{\femto\metre\per\clight}$, the $\Delta$ resonance has very little time to be affected by the medium before decaying into a $\pi$ and nucleon.
-Thus the outgoing $\pi$ contains information on the high density region of the collision.
+Thus the outgoing $\pi$ can contains information on the high density region of the collision.
 
@@
 
-The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordon coefficients as shown in \ref{appen:deltadecay}.
-Here we see that in general p-p collisions give rise primarily to  $\pi^+$ and n-n collisions give rise to primarily $\pi^-$.
-In this $\Delta$ resonance model the charged pion ratio can be described as,
+The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordon coefficients as shown in Appendix~\ref{appen:deltadecay}.
+Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$.
+From these branching ratios the charged pion ratio can be described as,
 
@@
 
-In this $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
-Pions can be reabsorbed into a $\Delta$ resonance after colliding with another nucleon in the backward process of $\Delta \leftrightarrow \pi N$.
+In this simple $\Delta$ resonance model, $\pi^-/\pi^+ \approx (N/Z)^2$ where $N/Z$ is the neutron-proton ratio of the dense central collision where they are produced.
+Pions can be reabsorbed into a $\Delta$ resonances after colliding with another nucleon in the backward process of $\Delta \leftrightarrow \pi N$, only to be re-emitted again into a pion.
 This process generally dilutes the pion sensitivity to the high density region, since with each absorption and re-emission changes the pion dynamics or even the charge of the pion reflecting the neutron-proton asymmetry at the point of creation and re-emission.
-Total pion absorption, back into two nucleons, requires a three body process; the pion collides with a nucleon creating a $\Delta$ resonance, then another nucleon must quickly collide with the resonance to create two nucleons.
-Since this is a three body process,h total pion absorption (removing them from the system) is a smaller effect than the absorption and re-emission process.
-Aside from these two effects which degrade the observable, in general the $\pi^-$ and $\pi^+$ are connected with n-n and p-p behavior in the high density, early collision; effectively turning the neutron measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally, and probing the high density region more than any other observable.
+Total pion absorption, going back into two nucleons, requires a three body process; i.e.
+the pion collides with a nucleon creating a $\Delta$ resonance, then another nucleon must quickly collide with the resonance to create two nucleons.
+Since this is a three body process, the total pion absorption (removing them from the system) is a smaller effect than the absorption and re-emission process.
+Aside from these two effects which degrade the observable, in general the $\pi^-$ are connected with neutrons and the $\pi^+$ production with proton behavior in the high density, early collision; effectively turning the problematic neutron measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally.
+It is also directly related to mostly the high density region where the $\Delta$'s are produced.
 
@@
 In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}.
-Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam impinging on a stable Sn target.
+Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam, impinging on a stable Sn target.
 We can probe the isospin asymmetry of the symmetry energy by changing the neutron-proton asymmetry of the incoming beam.
 Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}.
-Yet in these previous experiments, only total pion yields were published and no pion spectra was not published.
+Yet in these previous experiments, only the total pion yields were published and no pion spectra were published.
 It was also known that the efficiency analysis of the FOPI detector was particularly difficult especially for low energy pions.
-Although the pion production increases with the beam energy, the effects of the symmetry energy decrease.
-This is because higher beam velocities mean the time two nuclei spend together becomes shorter, shortening the time in which the symmetry energy can affect nucleons.
-The goal of this Thesis was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs.
-To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) Time Projection Chamber (TPC).
-
-%nuclear chart
-%Dense nuclear matter 
-%Liquid drop model explain density around saturation
-%explain symmetry energy 
-%extend to systems of larger pressure, isospin assymetry 
-%symmetry energy how to calculate 
-%Previous constraints on symmetry energy
-%heavy ion collisions only way to probe density and 
-%dynamics of proton neutrons
-
-%pion production
-
-\begin{comment}
-
-\section{From Nuclear Forces to the Equation of State}
-Isospin as a good quantum number at low energies 
-Figure showing the nuclear forces for the pp, pn, nn 
-Pauli exclusion 
-inter particle distance , saturation density (maybe figure of all densities?)
-Building the infinite matter
-Statistical model 
-Forces manifest in Energy/particle
-
-\section{The Nuclear Equation of State}
-Figure showing the binding energy vs p-n asym
-Liquid drop model, mass equation, move to higher densities
-Symmetric EoS asymmetric EoS(symmetry energy)
-Density dependence of the symmetry energy
-
-\section{Phases of Nuclear Matter}
-gas liquid phase, gas, where are we
-Progression through the heavy ion collision 
-liquid, liquid gas, gas
-
-\section{Studying EoS through Heavy Ion Collisions}
-going from infinite matter to finite matter 
-approximate nuclear matter with neutron rich radioactive ion beams
-probe different asymmetries 
-probe different densities with different beam energies 
-Symmetry energy goes down at higher beam energies
-
-\section{Boltzmann Ulong Uhlenbeck (BUU) Transport Code}
-Does not reach equilibrium for all observable (hadrons) what about pions?
-Build up a mean field picture (momentum dependent) unknown quantities here
-Non equilibrium can be solved with Boltzman equation with collision term and solved by MC
-Figure of transport simulation code showing collision progression 
-Clustering is an issue 
-Meson, resonance production
-
-\section{Observables of interest to the EoS}
-Particle yields of isospin opposites 
-Flow??
-Figure showing p-n observable lessens at high energy and from all densities
-
-\section{Pion Production }
-Figure showing delta resonance 
-Figure showing pion's produced at 2po 
-Figure showing fermi motion effect (not really sub threshold but ok)
-(chemical potential model, delta isobar) prediction for pion ratio
-pion mass is small momentum shifted by coulomb affecting shape of pion spectra
-
-\section{Previous Constraints}
-Figure showing GW constraint and other previous constraints 
-FOPI data at 400 A MeV not as sensitive to symmetry energy 
-Saturation density and lower constrained but issues at high density
-conflicting analysis on FOPI data which was not intended to be used for Symm Energy
-Can you group all the constraints to one nice plot???
-
-\section{Motivation for building the S$\pi$RIT TPC}
-Based on EOS TPC
-High efficiency for detecting pions and other light charged particles 
-Effort to reduce the experimental error bars on such spectra which theory could use
-
-\end{comment}
+In fact the published pion multiplicities are not the measured values but rather extrapolated to low pion energies.
+We know now that below the cut off cited in \cite{fopi}, approximately 35\% percent of the $\pi^+$ multiplicity and 55\% percent of the $\pi^-$ multiplicity, showing the importance of measuring down to low pion energies.
+While the pion production increases with the beam energy, this increase is countered by the fact that the effects of the symmetry energy decrease with energy \cite{fopi}.
+This is because available energy to produce pions is larger and therefore less sensitive to the relatively small symmetry energy.
+The goal of this Dissertation was to design and build a high efficiency detector in order to measure pion and light charged particle spectra resulting from HICs.
+To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) TPC.

### 2020-07-17 chapters/ch_results.tex
Added bills edits working on appendix figures

@@
 
- The data taken has been compared with 7 commonly used transport theoretical codes.
-These codes took part in a large collaborative effort in order to standardize certain common elements of each codes \cite{theoryComp1,theoryComp2}.
-To reduce uncertainties in the pion production mechanism, each model used comparison project simulated nuclear matter in a box type simulation with fixed, well defined, initial conditions.
-Though these box calculations model an equilibrated system at finite density, which is  not replicated by experiment, the limit is theoretically clearly defined and provides a benchmark comparison between models where the solutions  yield values for the total number of produced pions and $\Delta$ that can be analytically calculated from the ansatz of  statistical equilibrium, providing a good benchmark.
-This allowed for each code to systematically go through the numerical treatment and details in each code such as initialization of nuclei, stability of the code, numerical handing of pauli-blocking, etc.
-After satisfying these benchmark calculations, each model was used with what the authors considered the most reliable input and model assumption.
-codes were taken in their best configuration, as a result of these comparison projects, without any prior knowledge of the experimental data.
+ The data taken has been compared with 7 commonly used transport theoretical models.
+These models took part in a large collaborative effort in order to standardize certain common elements of each models \cite{theoryComp1,theoryComp2}.
+To reduce uncertainties in the pion production mechanism, each model simulated nuclear matter in a box type simulation with fixed, well defined, initial conditions.
+Though these box calculations model an equilibrated system at finite density, which is  not replicated by experiment, the solutions  of  pion and $\Delta$ yields can be analytically calculated from the ansatz of statistical equilibrium, providing a good benchmark.
+This allowed for each model to systematically go through the numerical treatment and details in each model such as initialization of nuclei, stability of the model, numerical handing of pauli-blocking, etc.
+After satisfying these benchmark calculations, each model was used with what each author considered the most reliable input and model assumption.
+These models were taken in their best configuration, without any prior knowledge of the experimental data.
 We then simulated the 4 systems measured in the \spirit TPC at an impact parameter of \SI{3}{\femto\metre} at \SI{270}{\MeVA} beam energy.
-While the numerical treatments of each code are reasonably similar, each code differs considerably in their treatment of pion and $\Delta$ dynamics.
-Some codes contain modifications to how the pion behaves in nuclear matter, i.e.
-\emph{in-medium effects}, typically introduced by including a pion optical potential which describes the pion scattering and absorption.
-Some codes include the iso-scalar and the iso-vector delta potential, which are not very well constrained but can be important in the production of pions \cite{baoan_deltapotential,inmedPionKo,inmedPionFeng}.
-The set of 7 codes represent the best codes at the moment, which can simulate pions at these energies.
+While the numerical treatments of each model are reasonably similar, each model differs considerably in their treatment of pion and $\Delta$ dynamics.
+Some models contain modifications to how the pion behaves in nuclear matter, i.e.
+\emph{in-medium effects}, typically introduced by including a pion optical potential, which describes the pion scattering and absorption.
+Some models include the iso-scalar and the iso-vector delta potential, which are not very well constrained, but can be important in the production of pions \cite{baoan_deltapotential,inmedPionKo,inmedPionFeng}.
 
-It is important to note that the interaction of pions and deltas with dense matter should not be viewed as a nasty detail that a "better" probe like nucleonic elliptical flow would avoid.
+%It is important to note that the interaction of pions and deltas with dense matter should not be viewed as a nasty detail that a "better" probe like nucleonic elliptical flow would avoid. 
 One of the central questions regarding matter at twice saturation density concerns the importance of deltas and pions as important constituents of the matter at those densities \cite{pionNS,deltaNS,awayforward}.
-Questions of the role of delta's and pion at 2$\rho_o$ are strongly connect to these potentials.
-Calculations of pion production indicate that selected pionic observables such as the energy spectra and transverse momentum distributions can be used to constrain these densities \cite{cozmaPC}.
-Obtaining such constraints are a major objective of our current work, but such constraints are ready for discussion at the present time.
+Questions of the role of deltas and pions at 2$\rho_o$ are strongly connected to these potentials.
+Calculations of pion production indicate that selected pionic observables such as the energy spectra and transverse momentum distributions can be used to constrain these potentials \cite{cozmaPC}.
+Obtaining such constraints are a major objective of our current work, but most of these constraints have not yet been obtained.
 
-As will be the re-occurring theme, there is a large variation in the predicted observables, between theoretical codes; greater than the variation between different symmetry energy in a particular code.
-However, the remedy for this discrepancy is for the authors of these models to include the other physics that many of them have neglected.
-This includes a better description of delta interactions of matter, which is critical to reproducing the total pion yields, and also the momentum dependence of the isovector mean field potentials, which have a large impact on the shapes of the energy spectra.
-Though some of the elementary issues relevant to such transport theories, such as the Pauli principal and the incorporation of inelastic processes, have been addressed.
-The relevant physics include the mean field potentials for nucleons, pions and deltas.
-A primary focus is on the isovector potentials of which only the momentum independent potentials for nucleons have been varied in the calculations shown below.
-The momentum dependent nucleonic mean field potentials and the mean field potentials for delta's and pions are only included in a subset of calculations.
-Calculations have already identified that the overall pion yield are influenced by the mean field potentials for the delta's \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of for delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
+As will be the re-occurring theme, there is a large variation in the predicted pion observables between theoretical models; this is greater than the variation between the stiff and soft symmetry energies within a particular model.
+However, the remedy for this discrepancy is for the authors of these models to include or better model some other physics that many of them have neglected, or it could be that the relevant physics is not constrained by other observables and needs to be constrain by further experiments and comparisons.
 
-%Similar to the experience with the symmetric matter EoS, the next steps will be to connect these mean field to to observables and begin constraining them [cite danielewics science article and Cozma private communication]
-With the \spirit TPC, we were able to measure the pion yield without resorting to extrapolations, and were able to measure the pion energy spectra of low to high energy pions with high inefficiencies.
-In the following pages we will present the total pion yield and energy spectra of this dissertation, which complements the dissertaion of Jon Barney \cite{jon}.
+For example, it is clear that many models need a  better description of the delta interactions with matter, which is critical to reproducing the total pion yields, and also the momentum dependence of the isovector mean field potentials, which have a large impact on the shapes of the energy spectra.
 
-%to provide much needed experimental data in which each code may compare too.
-The total theoretical uncertainty is too large relative to the experimental error bars to make a constraint on the density dependence of the symmetry energy.
-Any constraint would depend on the particular code one is using and would result in conflicting conclusions when using another theory.
-There is no good consensus at this time which codes should be favored over other codes, as the magnitude of the effects included in each code are still debated in order of importance.
-Here I will present the current level of agreement with the codes and remark on some of the implications it has moving forward.
+Other unconstrained physics includes aspects of the mean field potentials for nucleons, pions, and deltas.
+A primary focus is on the isovector potentials, of which only nucleon potentials for have been varied in the calculations shown below.
+Some of the models include the momentum dependent nucleonic mean field potentials, and the mean field potentials for deltas and pions.
+Calculations have already identified that the overall pion yields are influenced by the mean field potentials for the deltas \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of for delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
+
+With the \spirit TPC, we were able to measure the pion yield without resorting to extrapolations, and were able to measure the pion energy spectra of low to high energy pions accurately.
+In the following pages we will present the total pion yield and energy spectra for the most asymmetric systems, which complements the measurements of $\tin{112}{124}$ and $\tin{124}{112}$ collisions, which were at the focus of the dissertation of Jon Barney \cite{jon}.
 
@@
 We begin with the total charged pion multiplicities for the two systems of extreme isospin asymmetry: the neutron rich $\tin{132}{124}$ system and the neutron deficient  $\tin{108}{112}$ system.
-Though the previous FOPI data set provided significant constraints pertaining the the symmetry energy, they relied on extrapolations to low transverse momenta, $P_T = \sqrt{p_x^2 + p_y^2}$, below \SI{160}{\mega\electronvolt\per\clight} \cite{fopi}.
+Though the previous FOPI data set provided interesting data relevant to  the symmetry energy, they relied on extrapolations to low transverse momenta, $P_T = \sqrt{p_x^2 + p_y^2}$, below \SI{100}{\mega\electronvolt\per\clight} \cite{fopi}.
 In the \spirit TPC, we were able to measure pions to very low $P_T$ for rapidities $y > y{CM}$; where the rapidity is expressed along the beam direction (z) as,
@@
 This marks the first time that pions have been measured to such low $P_T$.
+We note that approximately 35\% of $\pi^+$ and 55\% of $\pi^-$ lie below the $P_T$ cut offs cited in the FOPI results \cite{fopi}.
+For the physics they were intending to do their extrapolations were sufficient, but for constraining the small effect of the symmetry energy, it is critical to measure all the way down to threshold for the pions to not introduce any extrapolation errors.
 
@@
 The integrated pion yield for both systems, and the $\pi^-/\pi^+$ ratio, is listed in Table~\ref{tb:pionyield}, where the systematic errors are the first error bar and the statistical error is listed next.
+This represents data with an estimated average impact parameter of $b_{avg} = \SI{2.2}{\femto\metre}$ and $b_{max} < \SI{3}{\femto\metre}$.
 Notice that the pion ratio is significantly greater than the $N/Z$ of the system.
-This is not completely unexpected; in the Delta Resonance model where $\pi^-$ come primarily from neutron-neutron collisions and $\pi^+$ from proton-proton collisions, we expect  $\pi^-/\pi^+ = (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
-For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this naive scaling assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.
+This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ ~ (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
+For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.
 
-Due to this connection between pion ratio and the isospin asymmetry at high density, the pion ratio was proposed to be a good probe of the isovector dynamics at high density.
+%Due to this connection between pion ratio and the isospin asymmetry at high density, the pion ratio was proposed to be a good probe of the isovector dynamics at high density.
 was hypothesized to be proportional to the high density N/Z ratio of the early system, where the other effects such as pion absorption and re-emission would dilute the effect, lowering the pion ratio as the system tends toward isospin equilibrium.
 It is likely however, that the dynamics plays are more significant role in pion productions via the $\Delta(1232)$ resonance than the simple $(N/Z)^2$ relation, which was also seen in \cite{fopi}.
-Some codes assume the potential of the delta is just the same as the corresponding nucleon potential, as given by the iso-spin.
-While other codes -- namly TuQMD -- introduce $\Delta$ resonance in-medium potential, which has an iso-scalar component (independent of iso-spin) and an iso-vector component.
-The nature of these two potentials is still unknown, and the role it plays in pion production has shown to be very important \cite{baoan_deltapotential, cozmaPC}.
+Some models assume the potential of the delta is just the same as the corresponding nucleon potential, as given by the iso-spin.
+While other models -- namly TuQMD -- introduce $\Delta$ resonance in-medium potential, which has an iso-scalar component (independent of iso-spin) and an iso-vector component.
+These two potentials are poorly constrained, but the role they play in pion production in collisions and in neutron stars has been shown to be very important \cite{baoan_deltapotential, cozmaPC}.
 
@@
 
-We have compared the total pion yields and ratios to the 7 common transport codes for the systems measured.
-The table of the transport codes are listed in Appendix~\ref{tb:pionyieldTheory}.
-Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the codes.
-The codes plotted here are only the soft symmetry energy since the variation in code is much larger than the variation within a code between different symmetry energies.
-While some codes make a reasonable approximation of a particular charge pion yields, no code reasonably predicts both.
+We have compared the total pion yields and ratios to the 7 common transport models for the systems measured.
+The table of the transport models are listed in Appendix~\ref{tb:pionyieldTheory}.
+Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the models.
+The models plotted here are only the soft symmetry energy since the variation in model is much larger than the variation within a model between different symmetry energies.
+While some models make a reasonable approximation of a particular charge pion yields, no model reasonably predicts both.
 
@@
 \includegraphics[width=.8\textwidth]{totalpiYield.png}
-\caption{Total pion yields as compared with 7 common transport codes.}
+\caption{Total pion yields as compared with 7 common transport models.}
 \label{fig:totalpiYield}
@@
 
-Figure~\ref{fig:totalpiRatio} shows the total single pion ratio, $R = Y(\pi^-)/Y(\pi^+)$, and the double ratio (DR), of the $\tin{132}{124}$ and $\tin{108}{112}$ system's single ratio, i.e.
-$DR = R_{\tin{132}{124}}/R_{\tin{124}{112}}$.
-Here, the variation between codes is much larger than the variation between the momentum dependence of the symmetry energy within a particular code.
-The symmetry energy variation (soft and stiff extremes) of two codes -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and in all codes in the double ratio.
-The ciruclar markers and band represents the total error bar in the data.
-Certainly it can be seen that the variation between symmetry energy extremes in the codes, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}, and the small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large disagreement between codes.
+Figure~\ref{fig:totalpiRatio} shows the total single pion ratio, $R = Y(\pi^-)/Y(\pi^+)$, and the double ratio  $DR = R_{\tin{132}{124}}/R_{\tin{124}{112}}$, of the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
+Here, the variation between models is much larger than the variation between the momentum dependence of the symmetry energy within a particular model.
+The symmetry energy variation (soft and stiff extremes) of two models -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and in all models in the double ratio.
+The circular markers and band represents the total error bar in the data.
+Certainly it can be seen that the variation between symmetry energy extremes in the models, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}, and the small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large disagreement between models.
 
@@
 \includegraphics[width=\textwidth]{totalpiRatio.png}
-\caption{Total pion ratio and double ratio compared with 7 common transport codes.}
+\caption{Total pion ratio and double ratio compared with 7 common transport models.}
 \label{fig:totalpiRatio}
@@
 The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies.
-In the $\tin{132}{124}$ data the $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
-We also expect the pion ratio is proportional to the $N/Z$ of the system.
+In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
 Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}.
@@
 It is also worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}.
-Scaling the other systems in the \spirit TPC data leads to almost the exact same value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.
+It it worth mentioning that the two experiments differ considerably in their efficiencies as was mentioned before.
+The FOPI results heavily relied on extrapolations to lower pion energies which the \spirit TPC did not rely on \cite{fopi}.
+Without a systematic error calculation or discussion with the FOPI collaboration no further detail can be given except that the \spirit TPC data seems to be reasonable expectation and in general agreement with the FOPI data set.
+Scaling the other systems in the \spirit TPC data leads to almost the exact same scaled value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.
 
@@
 This data marks the first time pion spectral data has been measured at sub-threshold energies.
-In the pion spectra, we can see the effect of the coulomb potential, which accelerates $\pi^+$ and decelerates $\pi^-$ particles, due to the positive charge of the nuclear medium.
-The low $\pi^+$ production at low energy is caused by the coulomb barrier between protons which limits $\pi^+$ production at low energies.
-The Coulomb force has the largest effect on the spectra of the pions and will play an important role when observing the general shape of the pion spectral ratio.
-Besides the initial coulomb barrier for $\pi^+$ production, the coulomb force affects only the final energy of produced pions, and has almost no role in the total pion yield.
-
-%Add figure on Glauber model 
-%Add figure on comparison to FOPI data
+In the pion spectra, we can see the effect of the Coulomb potential, which accelerates $\pi^+$ and decelerates $\pi^-$ particles, due to the positive charge of the nuclear medium.
+The low yield for the $\pi^+$ production at low energies is caused by the Coulomb barrier between protons which limits $\pi^+$ production at low energies.
+The Coulomb force has the largest effect on the spectra of the pions and will play an important role when observing the general shape of the pion spectral ratio, where it changes its shape completely.
 
 \section{Pion Spectral Ratio}
-
-\begin{comment}
-
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{singleRatio.png}
-\caption{Single ratio spectra}
-\label{fig:SRspectra}
-\end{figure}
-\end{comment}
 
@@
 In particular, it may be more sensitive to the high density regions of the early collision.
-It is generally true, that high energy pions are more likely to exit the nuclear medium  earlier, and therefore be less prone to effects such as pion absorption and re-emission; all of which dilute the sensitivity of the pion observable to the high density behavior.
-Conversly, low energy pions are more likely to stay in the nuclear medium, and be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
+It is generally true that high energy pions are more likely to exit the nuclear medium  earlier, and therefore be less prone to effects such as pion absorption and re-emission; all of which dilute the sensitivity of the pion observable to the high density behavior.
+Conversely, low energy pions are more likely to stay in the nuclear medium, and be affected by other effects such as the $\Delta$ potential in medium \cite{baoan_deltapotential}.
 Integrating these two regions to get the total pion yield dilutes the possible sensitivity the high energy pions have.
@@
 The general hyperbolic shape seen in both figures arises from the Coulomb force on the two charged pion spectra mentioned earlier in Section~\ref{sec:pionSpectra}.
-Also notice the pion ratio is smaller for less neutron poor system, as we would expect since less neutron-neutron collisions produce less $\pi^-$.
+Also notice the pion ratio is smaller for the neutron poor system, as we would expect since there are fewer neutron-neutron inelastic collisions and consequently fewer $\pi^-$.
 
@@
 In a similar way to the total pion double ratio, we can construct the spectral double ratio, the ratio between the two system's single ratios.
-In a similar way described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory --between the two systems-- to cancel out.
+As described in Section~\ref{sec:doubleRatio} we would expect systematic uncertainties in the experiment, and even in the theory --between the two systems-- to cancel out.
 For the same reasons as the pion spectral ratio, we would expect the high energy pions to be more sensitive to the high density region of the early collision.
-Here, just as in the total double ratio, the double ratio is significantly higher than any theory.
+Here, just as in the total double ratio, the spectral double ratio is significantly higher than any theory.
 Here as before, the bands in the theory represent the two extremes of the symmetry energy.

### 2020-07-17 chapters/ch_summary.tex
Added bills edits working on appendix figures

--- 2020-07-14 chapters/ch_summary.tex
@@
 \chapter{Summary}
-The need for precise experimental pion data is paramount to understanding exotic, dense matter, such as neutron stars.
-Besides direct observation of these objects, the lab is the best way to probe the density dependence of the symmetry energy, and is the only way to probe different asymmetries.
-Pions have been proposed to be sensitive to the high density regions of these collisions, making it an ideal probe to constrain the symmetry energy.
-Though simulating pion production has proved to be quite challenging, it has motivated the design of high efficiency detectors such as the \spirit TPC to measure pion emission from neutron rich heavy ion collisions.
-We preformed a campaign of experiments measuring collisions of neutron rich, radioactive beams, on stable targets in 4 different configurations -- $\tin{132}{124}, \tin{108}{112}, \tin{112}{124},$ and $\tin{124}{112}$.
-All measured with the \spirit TPC which was able to measure pions to very low energies without the need for any extrapolations and very little assumptions.
+The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars.
+Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter.
+Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta = \frac{\rho_n - \rho_p}{\rho}$.
+This information is critical to understanding the symmetry energy and associated isovector mean field potentials.
+These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars.
+Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.
+
+Pions have been proposed to be sensitive to the high density regions of these collisions, making it a compelling probe to constrain the symmetry energy.
+Moreover, questions about what are the stable phases at twice saturation densities cannot be answered without understanding the mean field potentials for pions and deltas in such matter.
+Though simulating pion production has its challenges, the prospect of answering such compelling questions has motivated the design of high efficiency detectors such as the \spirit TPC to measure pion emission from neutron rich heavy ion collisions.
+We have preformed a campaign of experiments measuring collisions of neutron rich, radioactive beams, on stable targets in 4 different configurations -- $\tin{132}{124}, \tin{108}{112}, \tin{112}{124},$ and $\tin{124}{112}$, all measured with the \spirit TPC.
 
@@
 The pion yield, ratio, and some new observables -- namely the double ratio-- were measured to within <4\% accuracy.
-We observed a deviation from the  naive $\pi^-/\pi^+ ~ (N/Z)^2$ scaling expected from the $\Delta$-resonance model discussed in Section~\ref{sec:pionObs}.
+We find that a major fraction of the pions are emitted below the thresholds of the FOPI experiment, demonstrating that the systematic errors are of the pion ratios from that previous experiments are completely dominated by the extrapolation of the FOPI data to the emission threshold.
+Based on this information, it is clear that prior conclusions based on comparison of those extrapolated data to theoretical calculations of total pion yields do not take the systematic errors of that extrapolation into account and any agreement their conclusions would contain error but to an unknown degree.
+
+  This marks the first time pions have been accurately measured in the sub-threshold region.
+We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ ~ (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
+Such a deviation is also predicted by theory, so this explanation should be retired in favor of something more quantitative and informative.
+Obtaining a better explanation from current calculations is difficult because nearly all pion production calculations are incomplete and each model has a mixture of strengths and inadequacies that are unique to each model.
+Thus, it is difficult to get the theorists to agree about the consequences of some of the model assumptions that they employ; and consequently a constraint on the density dependence of the symmetry energy.
+
+  %With this wealth of new data that we have obtained, we now begin to work with theorists to understand what new experimental pion observable is sensitive to which theoretical assumption.
+We will are exploring how the isoscalar and isovector $\Delta$ potential plays an important role in the pion production spectra and emission rates.
+  Also for the first time, the pion spectra was measured.
+The spectral ratio appears to provide information that is relevant to this question also providing much needed information about the dynamics of pions, even to very low pion energies.
+Of particular importance is the sensitivity of calculations to the high energy tails of the pion spectra.
+This is a promising observable since high energy pions  exit the nuclear matter sooner and are less affected by re-scattering and adsorption-re-emission processes, which dilute information about the EOS at the highest density.
+We expect that high energy pions will be more sensitive to the prevailing conditions at high density, while low energy pions are more sensitive to the energy available for their emission, such as the Coulomb potential, the pion optical potentials, and the $\Delta$ potential.
 
 \subsection{Outlook}
+Theoretical collaboration, which has made considerable progress improving details of the calculation, will further improve in light of this new data.
+It is clear that a serious effort must be made to include effects such as the symmetry potentials of pions and deltas; which most codes do not.
+It also appears to be possible to make first constraints on the generally unknown $\Delta$ potential by fitting the pion yield data \cite{cozmaPC}.
+Making such a constraint would allow for the possibility to start making constraints on the density dependence of the symmetry energy.
+The interactions with theoretical community that has coalesced around this effort is very exciting, and we are seeing real progress in transport theory.
+It seems very possible that the theoretical efforts will converge and provide solid interpretations of these and other data much more quickly than they did in the case of the symmetric matter EoS.

### 2020-07-17 MSU-thesis-estee.tex
Added bills edits working on appendix figures
Rough draft. Needs a once over. Bill's edits in. Appendix content in.

@@
 \usepackage{graphicx}
-\usepackage{cleveref}
 \usepackage{overpic}
@@
 
-\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/}  {images/intro/} }
+\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/}  {images/intro/} {images/appendix/} }
 

### 2020-07-17 chapters/ch_appendix.tex
Rough draft. Needs a once over. Bill's edits in. Appendix content in.

@@
 
+A similar cut variation analysis was performed on the pion spectral ratio where each bin in Figure~\ref{fig:SRsn132} and \ref{fig:SRsn108} were plotted as a function of the cut parameter.
+Learning from the total pion ratio, we would not expect much dependence on the spectral ratio, since the ratio tends to cancel any systematic effects.
+Figure~\ref{fig:clusterSpectralSn132} shows the $\tin{132}{124}$ pion spectral ratio varying the number of cluster cut, after efficiency correction.
+Very little systematic dependence is observed except for the extreme limit of $N_c$ > 26 as was seen and discussed above.
+Figure~\ref{fig:docaSpectralSn132} shows the spectral ratio, varying the DOCA cut; no systematic dependence is observed.
+While we already know there are some physics considerations causing the multiplicity dependence I have shown the cut variation of the $\tin{132}{124}$ system multiplicity in Figure~\ref{fig:multSpectralSn132}, though no systematic error was attributed.
+In all no systematic errors were significant enough to add to the pion spectral ratio for the $\tin{132}{124}$ system.
+
 \begin{figure}[!htb]
@@
     \includegraphics[width=\textwidth]{clusterSpectraSn132}
-	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut in the $\tin{132}{124}$ system.}
 	\label{fig:clusterSpectralSn132}
@@
     \includegraphics[width=\textwidth]{pocaSpectraSn132}
-	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut in the $\tin{132}{124}$ system.}
 	\label{fig:docaSpectralSn132}
@@
 
+As we will see for the $\tin{108}{112}$ system, spectral pion ratio, the story is the exact same.
+Figure~\ref{fig:clusterSpectralSn108} shows the cluster cut variation, Figure~\ref{fig:docaSpectralSn108} shows the DOCA cut variation, and Figure~\ref{fig:multSpectralSn108} shows the $\tin{108}{112}$ system multiplicity cut variation.
+
 \begin{figure}[!htb]
@@
     \includegraphics[width=\textwidth]{clusterSpectraSn108}
-	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut.}
-	\label{fig:clusterSpectralSn132}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the cluster cut in the $\tin{108}{112}$ system.}
+	\label{fig:clusterSpectralSn108}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{pocaSpectraSn108}
-	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut.}
-	\label{fig:docaSpectralSn132}
+	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the DOCA cut in the $\tin{108}{112}$ system}
+	\label{fig:docaSpectralSn108}
 \end{figure}
@@
 	\caption{The spectral pion ratio Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{108}Sn$ cut.}
-	\label{fig:multSpectralSn132}
+	\label{fig:multSpectralSn108}
+\end{figure}
+
+As with all the ratios, the double ratio also cancels systematic effects.
+Figure~\ref{fig:clusterDR} shows the cluster cut variation, Figure~\ref{fig:docaDR} shows the DOCA cut variation in the spectral double ratio observable.
+We have neglected to show the multiplicity, since we have already shown that it is derived from physics effects, and depends only on the impact parameter cut one intends to apply, not a free floating cut like the others.
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{clusterDR}
+	\caption{The spectral pion double ratio  when varying the cluster cut.}
+	\label{fig:clusterDR}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+    \includegraphics[width=\textwidth]{docaDR}
+	\caption{The spectral pion double ratio when varying the DOCA cut.}
+	\label{fig:docaDR}
 \end{figure}

### 2020-07-17 MSU-thesis-estee.tex
Almost final draft

--- 2020-06-18 MSU-thesis-estee.tex
@@
 % You must specify the title of your thesis, your name, the field of study (not department), and the year
-\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
-\author{Justin Estee}
+%\title{Constraining the High Density Nuclear Symmetry Energy with Pions}
+
+%\title{Pion Emission from Neutron Rich Heavy Ion Collisions in an Effort to constrain the symmetry energy}
+
+\title{Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy}
+
+\author{Justin Brian Estee}
 \fieldofstudy{Physics} % This should be in sentence case
@@
 % Your abstract goes here.  Master's 1 page max. PhD 2 page max.
+
+Recent astronomical measurements of neutron star mergers have reinvigorated questions about the nature of dense matter in the universe.
+The densities reached in these stars range from normal nuclear densities $\rho_o$ to upwards of $4\rho_o$ or even $9\rho_o$ in some models, where $\rho_o = \SI{1.7e14}{\gram\per\centi\metre\cubed}$; representing extremely dense, neutrons rich matter.
+The symmetry energy describes the difference in binding energy between pure neutron matter and symmetric matter with equal number of neutrons and protons.
+It is the density dependence of the symmetry energy that is related to the outward pressure, which supports the star, preventing it from collapsing under its own gravitation force.
+For years, theory and experiment  have combined to progressively constrain the Equation of State (EoS) of symmetric matter, and it was only in the last decade that the density dependence of pure neutron matter, the symmetry energy, had been constrained at normal nuclear densities and lower, $\rho < \rho_o$.
+Still large uncertainties, and discrepancies on preliminary constraints,  still remain at densities around 2$\rho_o$ which are critical importance to the understanding the properties of neutron stars.
+
+Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities and proton-neutron asymmetries.
+It has been shown that charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonances, in the early, dense parts of Heavy Ion Collisions (HICs).
+Yet, existing pion experimental data tailored to answering questions about the symmetry energy was lacking and suffers from unknown systematic errors.
+this prompted the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pions produced in neutron-rich HICs with a high efficiency.
+This Dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems measured at \SI{270}{\MeVA} taken at the RIBF facility at RIKEN.
+The total pion ratio was measured to within 4\% and the pion spectra was measured with a high degree of accuracy, marking the first time pion spectra  -- at energies relevant to the symmetry energy -- have been published.
+Also, we will compare the results with the most recent transport models which can simulate pion production.
+We will highlight some of the successes and hope to motivate a discussion in the theoretical community on how to better reproduce the pion phenomena in neutron-rich HICs, which is important for not only constraining the density dependence of the symmetry energy, but also for understanding the $\Delta$ and pion roles in neutron stars.
+
 \end{abstract}
@@
 % Your Acknowledgements are formatted like a chapter, but with no number
+\begin{comment}
 \chapter*{Acknowledgements}
@@
 Your acknowledgements here.
+
 %
 \clearpage
+\end{comment}
+
 % We need to turn single spacing back on for the contents/figures/tables lists

### 2020-07-17 chapters/ch_appendix.tex
Almost final draft

@@
     \end{tabular}
-    \caption{List of runs for the analysis. \label{tb:runList}}
+    \caption{List of all run numbers used in this analysis. \label{tb:runList}}
   \end{center}
@@
 It is usually not recommended to estimated systematic error bars using this method, but sometimes it is the only way.
-
-Here we will discuss the particular analysis using the total pion ratio as the example.
-Figure~\ref{fig:totalRatioError} shows the total pion ratio as a function of several cut variations, where one cut is varied and all the others are held constant.
-Three cuts are of particular interest, the event multiplicity, DOCA, and the number of cluster cut.
+Here we will discuss the particular analysis using the $\pi^-$ yield as an example and then discuss the $\pi^+$ the total pion ratio, and the spectral pion ratios.
+The three cuts used in our results are the event multiplicity, DOCA, and the number of cluster cut.
 
@@
          \caption{Efficiency corrected.}
-         \label{fig:pim_multCutVar}
+         \label{fig:pim_multCutVar_EC}
      \end{subfigure}
@@
          \caption{Efficiency corrected.}
-         \label{fig:pip_multCutVar}
+         \label{fig:pip_multCutVar_EC}
      \end{subfigure}
@@
          \caption{Efficiency corrected.}
-         \label{fig:ratio_multCutVar}
+         \label{fig:ratio_multCutVar_EC}
      \end{subfigure}

### 2020-07-17 chapters/ch_data_analysis.tex
Almost final draft

@@
 Due to coupling between the gating grid, the ground plane, and other effects, the impedance of the supply cables was not perfectly matched, causing an imbalance in the current where one side discharged faster than the other side.
-This caused an underdamped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum.
+This caused an under-damped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum.
 Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum after averaging over 2000 events in a given pad.
@@
  where $\phi$ is the electric potential and $\rho$ is the free charge.
-The numerical solution is provided by the Jacobi method \cite{poisson}, provided all 6 sides have defined Dirchlet boundary conditions for the potentials.
+The numerical solution is provided by the Jacobi method \cite{poisson}, provided all 6 sides have defined Dirichlet boundary conditions for the potentials.
 We neglect the wire plane region since the drift details around it are not significant for the space charge effects we are discussing here.
@@
 \includegraphics[width=\linewidth]{SC_Relation.pdf}
-\caption{Space charge relation}
+\caption{Relationship between the observable  $\Delta\mathrm{V}_\mathrm{LR}$ and the estimated charge density, and the beam intensity for each run.
+The blue and green line represents two separate runs at different beam intensities.
+The condition for  $\Delta\mathrm{V}_\mathrm{LR} = 0$ is solved for for each run.
+These values are then plotted against the beam intensity where finally a linear fit provides the interpolation for all beam intensities in between.}
 \label{fig:spacechg_relation}
@@
   \end{center}
-  \caption{Cobo timing correction.}
+  \caption{Cobo timing correction for three different layers.
+One singe pad is represented by a unique layer and row number.
+The $dy$ residual of the track fitting shows, in the upper panel, that there is a timing calibration issue.
+This offset is fixed over several runs for a given pad.
+Therefore we can correct using a time correction map for each pad; the bottom panels show the distribution after correction.
+}
   \label{fig:coboCorr}
@@
 
-The detailed software flow diagram of the embedding algorithm is shown in Fig.~\ref{fig:flow}.
+The detailed software flow diagram of the embedding algorithm is shown in Figure~\ref{fig:flow}.
 Starting from Geant4 and passing through the 3 MC digitization tasks described above, the output response of the MC track in the TPC is created.
@@
 Figure~\ref{fig:dataevent} shows this track embedded into an experimental event.
-In the following, here I will refer to experimental data with embedded MC signals as ``embedded data" for short, the MC generated response as ``MC data", and data only containing only the experimental signals as ``experimental data".
+In the following, here I will refer to experimental data with embedded MC signals as \emph{embedded data} for short, the MC generated response as \emph{MC data}, and data only containing only the experimental signals as \emph{experimental data}.
 The three sets of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
@@
 We only need to identify where the time of saturation occurs.
-The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Fig.~\ref{fig:pulseshape}.
+The characteristic signature of a saturated signal is the fast fall time of a pulse, which quickly reaches zero, as opposed to the long tail of normal pulses as seen in Figure~\ref{fig:pulseshape}.
 The long exponential tail can be effectively removed by a software technique which is similar to the electronics concept of pole-zero compensation \cite{polezero}.
@@
 This time is set to $t_{peak} - 30$ to ensure that the MC embedded signals do not overlap the real saturating data signal, which would be impossible.
-These two positions are also shown in Fig.~\ref{fig:pulseSatTag}.
+These two positions are also shown in Figure~\ref{fig:pulseSatTag}.
 
@@
 The other important cut which is made inside of the software is on the PRF distribution of each cluster discussed in Section~\ref{sec:corrTask}.
-We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Fig.~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data.
+We can also see good agreement between the MC PRF for $\pi^-$ in Figure~\ref{fig:prfpimMC}, and the data PRF, Figure~\ref{fig:prfpimData}, where the black line is the PRF fit to the experimental data.
 It is sufficient to use such a simple universal PRF function as we can describe all the crossing angle effects discussed in \ref{sec:prf}.
@@
 The efficiency then can be represented as just a function of the two remaining variables.
-Figure~\ref{fig:132_eff} shows the efficiency of $\pi^-$ and $\pi^+$ particles in the $\tin{132}{124}$ system and Fig.~\ref{fig:108_eff} for the $\tin{108}{112}$ system as a function of $\theta_{Lab}$ and momentum $p_{Lab}$.
+Figure~\ref{fig:132_eff} shows the efficiency of $\pi^-$ and $\pi^+$ particles in the $\tin{132}{124}$ system and Figure~\ref{fig:108_eff} for the $\tin{108}{112}$ system as a function of $\theta_{Lab}$ and momentum $p_{Lab}$.
 
@@
 The distribution around this mean value is not a Gaussian distribution.
-The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Guassian variable, defined as
+The variable $z$ has been used before to transform the energy loss of a particular track $dE/dx$ into a more Gaussian variable, defined as
 
@@
 
-    \caption{$\langle dE/dx\rangle$ versus p/q for $\tin{132}{124}$ system.}
+    \caption{The charged pion PID plot achieved in the $\tin{132}{124}$ system.
+PID is achieve by plotting $\langle dE/dx\rangle$ versus particle rigidity, $p/q$.
+Unique hyperbolic lines show different particle species, where negatively charged particles are plotted as negative rigidity.
+}
 \label{fig:pidpion}
@@
 The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}.
-Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Fig.~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.
+Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Figure~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.
 A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.
@@
 
-    \caption{$z_i$ projection for binned slices in p/q around the $\pi^-$.}
+    \caption{Flattened PID $z_i$ distribution for binned slices in $p/q$ around the $\pi^-$.}
 \label{fig:pimmom_flat}
@@
 
-    \caption{$z_i$ projection for binned slices in p/q around the $\pi^+$.}
+    \caption{Flattened PID $z_i$ distribution  for binned slices in p/q around the $\pi^+$.}
 \label{fig:pipmom_flat}

### 2020-07-17 chapters/ch_data_analysis_2.tex
Added bills edits working on appendix figures
Almost final draft

@@
-\chapter{Data Analysis II}
+\chapter{Data Analysis II: Cuts and final PID}
+\label{chap:data2}
+
 As Chapter~\ref{chap:data1} covered the corrections and various calibrations to the data, this Chapter will cover the various cuts taken to improve the signal in the experimental data.
@@
          \includegraphics[width=\textwidth]{beam-Sn132.png}
-         \caption{$\tin{132}{124}$ Beam particle identification.}
+         \caption{$\tin{132}{124}$ system.}
          \label{fig:beampid132}
@@
          \includegraphics[width=\textwidth]{beam-Sn108.png}
-         \caption{$\tin{132}{124}$ $\pi^-$ angular spectra.}
+         \caption{$\tin{108}{112}$ system.}
          \label{fig:beampid108}
      \end{subfigure}
-        \caption{$\tin{108}{112}$ Beam particle identification. }
+        \caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
         \label{fig:beampidTwo}
-
 \end{figure}
@@
 The resulting fragments are filtered in-flight to the desired seconary beam.
-The in-flight separation is handled by the BigRIPS fragment separator which is shown in Fig.~\ref{fig:samuraiBeamLine}.
+The in-flight separation is handled by the BigRIPS fragment separator which is shown in Figure~\ref{fig:samuraiBeamLine}.
 The dipole magnets D1 and D2 act as a velocity filter, selecting on certain magnetic rigidities $\beta\rho$.
@@
 \end{equation}
-where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Fig.~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.
+where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.
 
@@
     \end{tabular}
-    \caption{Multivariate normal distribution fit parameters for four beams.
+    \caption{2D Gaussian cuts parameters for all four systems beam PID.
     \label{beamParameters}}
@@
 
-Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Fig.~\ref{fig:beampid108} the ${}^{108}$Sn beam.
+Figure~\ref{fig:beampid132} shows a zoomed in view of the PID centered around the ${}^{132}$Sn beam and Figure~\ref{fig:beampid108} the ${}^{108}$Sn beam.
 The red lines represent the cut where particles identified inside the circle represent the beam events which are identified as the good beam events.
@@
 
-Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Fig.~\ref{fig:edge}.
+Near the edges of the detection volume, the clusters of tracks significantly deviate from the trend of the fitted track as seen in different view of the reconstructed clusters in Figure~\ref{fig:edge}.
 This happens because the last pads on the edges of the pad-plane have no neighboring pads containing charge, and therefore the last pad represents the last known position of the collected charge.
@@
     \caption{2D Gaussian fit parameters for $^{132}$Sn, $^{112}$Sn and
-      $^{108}$Sn beam angle. \label{beamAngleParameters}}
+      $^{108}$Sn beam angle cuts. \label{beamAngleParameters}}
   \end{center}
@@
  The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame.
-The z-component of all runs in each beam type are plotted around the the target region in Fig.~\ref{fig:vertexdist}.
+The z-component of all runs in each beam type are plotted around the the target region in Figure~\ref{fig:vertexdist}.
 The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position.
@@
 \end{tabular}
-\caption{Summary of measured verticies and their resolution.}
+\caption{Summary of the mean vertex location for the target position in all systems and the measured vertex resolution.}
 \label{tb:vertexresol}
@@
 \section{Impact Parameter Selection}
+\label{sec:impact}
 
@@
 \label{fig:poca}
-\caption{Distance to vertex for tracks.}
+\caption{Example diagram of the Distance of Closest Approach (DOCA) cut for tracks.}
 \end{figure}
@@
 These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.
-While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Fig.~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks.
+While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Figure~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks.
 Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly.
@@
 
-These regions can also be seen in the earlier discussion in the MC embedded efficiency in Fig.~\ref{fig:pim_eff_ex}.
+These regions can also be seen in the earlier discussion in the MC embedded efficiency in Figure~\ref{fig:pim_eff_ex}.
 We could of course include a larger region where the efficiency is small, but not zero, but it is not good practice to  correct using small efficiency values.

### 2020-07-17 chapters/ch_experiment.tex
Almost final draft

@@
 \chapter{Experiment}
+\label{chap:experiment}
+
 A Time Projection Chamber (TPC) is a detector that reconstructs the energy loss of charged particles and the location of that energy loss as in all 3 dimensions within the detector.
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=.9\textwidth]{fc_overview.png}
@@
 The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, which is the same as the other strip widths.
-The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.
+The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Figure~\ref{fig:FC_schematic}.
 Allowing for a tunable resistor allowed for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.
@@
 Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.
-Short cables transmit the signals from the surface mount connectors, through a circuit protection board called ZAP, to the inputs of the AGET chips which are mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}.
+Short cables transmit the signals from the surface mount connectors, through a circuit protection board called ZAP, to the inputs of the AGET chips which are mounted to the AsAd board as seen in Figure~\ref{fig:getzap}.
 Each AGET chip services 64 pads (63 pads are connected in our case).
@@
 In both approximations it is described by a most probable energy loss value, with a long, high-energy loss tail.
-The solid curve in Fig.~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc} \cite{bichsel}.
+The solid curve in Figure~\ref{fig:straggling} shows the energy loss distribution in Ar gas for a proton with momentum \SI{3.4}{\giga\eVperc} \cite{bichsel}.
 The dashed line is the distribution under the Landau assumptions.
@@
 The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants.
-The water never exceed  ???
-ppm  and the oxygen level never exceeded ???
-ppm.
+The water never exceed  100 ppm  and the oxygen level never exceeded 10 ppm, which are the two main contaminants in the gas which can absorb primary electrons \cite{tpcAging}.
 Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm (\SI{760}{\torr}) as a function of the reduced electric field value given in units of \si{\volt\per\centi\metre\per\torr}.
@@
 \label{sec:prf}
-Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Fig.~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
+Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Figure~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
 If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed a particular pad, $Q$, is expressed as,
@@
 For these reasons it is useful to experimentally measure the PRF and fit it with an empirical function, typically a Gaussian, to describe its behavior.
-The method for extracting the experimental PRF will be discussed latter, but by averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Fig.~\ref{fig:expprf}.
+The method for extracting the experimental PRF will be discussed latter, but by averaging over many events in the experimental data, the resulting PRF for the S$\pi$RIT TPC is shown in Figure~\ref{fig:expprf}.
 Here we see the deviations from the expected analytic Gatti distribution (black curve), whereas fitting with a two parameter Gaussian function (red curve) gives a better description of the  data, Eq.~\ref{eq:gaus}, with the two parameters being the normalization coefficient, $N_0$, width $\sigma$, and with a mean value assumed to be 0.
@@
 %scintillator bars
- The Kyoto Multiplicity Array, shown in Fig.~\ref{fig:aux}, consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
+ The Kyoto Multiplicity Array, shown in Figure~\ref{fig:aux}, consists of two arrays of plastic scintillating bars on each side of the TPC, each consisting of 30 bars.
 The entire TPC structure was designed so that light charged particles could pass through the field cage and side walls of the TPC enclosure without excessive energy loss and scattering.
@@
 %beam trigger optional
- Fig.~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure.
+ Figure~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure.
 Three of the 12 bars 1 mm in thickness and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time.
@@
 \subsection{Active Veto Array}
-The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Fig.~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
+The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Figure~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
 Because of the inherent angular dispersion of the beam there still were incoming beam events which significantly deviated from the target location.
@@
 Following the initial trigger, we had the possibility of a fast clear signal that could be initiated by a later beam particle traversing the TPC.
-If the last beam particle came within xx us after the fast trigger, electrons from this beam particle could pass through the gating grid.
+If the last beam particle came within 4 us after the fast trigger, electrons from this beam particle could pass through the gating grid.
 Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal.
@@
 
- Fig.~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
-Here the KATANA trigger and the SC SUM???
-together triggered the DAQ.
+ Figure~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
+Here the KATANA trigger and the SC SUM together triggered the DAQ.
 
@@
 
-Shown in Fig.~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
+Shown in Figure~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
 The primary goal of this data set was to calibrate the TPC in both momentum and dE/dx as will be discussed in Section~\ref{sec:cocktail}.

### 2020-07-17 chapters/ch_intro.tex
Almost final draft

@@
 
-Mankind has been interested in the nature of the matter which composes the visible universe, and what are its "constituents" or building blocks.
+Mankind has been interested in the nature of the matter which composes the visible universe, and what are its \emph{constituents} or building blocks.
 The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie.
@@
 Since the strong force makes the inter-nucleon distance approximately constant, the volume the nucleus is related to the total number of nucleons and the volume, $A$.
-Therefore the first \emph{volume} term $a_VA$ is proportional to $A$.
+Therefore the first \emph{volume} term $a_V$ is proportional to $A$.
 Since the  constituent quarks that make of the nucleons are of nearly equal mass, there is a fundamental symmetry of the strong interaction which the strong interaction between any pair of nuclei is very similar, regardless of the particular flavor of quark composition.
@@
 The term isospin was chosen because this symmetry was similar to the spin symmetry of electrons, which have spin up and down projections but are otherwise indistinguishable, unless in the presence of a symmetry breaking field.
-The isospin mathematics are is identical to the spin mathematics describing where nucleons have an isospin of 1/2 and with projections of  +1/2 for protons or -1/2 for neutrons.
+The isospin mathematics are is identical to the spin mathematics describing where nucleons have an isospin of 1/2 and with projections of  +1/2 for protons and -1/2 for neutrons.
 
@@
 The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}.
-Although the strong force does not distinguish between protons and neutrons the overall wavefunction is made up of a combination of the spacial, spin, and isospin wavefunctions.
-Since nucleons are fermions,the total wavefunction must be antisymmetric under particle exchange.
+Although the strong force does not distinguish between protons and neutrons the overall wave function is made up of a combination of the spacial, spin, and isospin wave functions.
+Since nucleons are fermions,the total wave function must be anti-symmetric under particle exchange.
 Because of this, each spacial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin asymmetry.
@@
 for a fixed number of particles $N$ and zero temperature.
-One can always extend to higher temperatures by adding the typical Boltzman dependence if needed, but here the simplification will suffice.
+One can always extend to higher temperatures by adding the typical Boltzmann dependence if needed, but here the simplification will suffice.
 The partial derivative with respects to volume can be rewritten in terms of density:
@@
 
-\section{Motivation for Thesis}
+\section{Motivation}
 In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}.
@@
 To do this a new Time Projection Chamber (TPC) was made called the SAMURAI pion Reconstruction Ion Tracker (\spirit) TPC.
+
+\section{Dissertation Outline}
+In this dissertation we will begin with a discussion of the operating principles of the \spirit TPC and the general setup of the experiment in Chapter~\ref{chap:experiment}.
+We will then talk about the corrections and calibrations applied to the data, as well as a brief introduction into the software and analysis procedure in Chapter~\ref{chap:data1}.
+This will naturally lead into a discussion in Chapter~\ref{chap:data2} about the various quality cuts taken on the data to improve the background rejection improving the signal to background ratio; this will also include the particle identification technique used to extract the pions.
+Finally in Chapter~\ref{chap:results} we will present the extracted pion yield, ratios, and spectra for the neutron-rich  $\tin{132}{124}$ system, and the neutron-poor $\tin{108}{112}$ system; which complement the systems analyzed in the dissertation by Jonathan Barney \cite{jon}.
+In Chapter~\ref{chap:summary} we will give a summary of the results and an outlook towards constraining the symmetry energy at high densities using pions.

### 2020-07-17 chapters/ch_results.tex
Almost final draft

@@
 \chapter{Results}
-%\section{Comparison to Theory}
-%Add figure of Pion spectra for different theory 
-%Reference paper for multiple theories
+\label{chap:results}
 
@@
 \end{tabular}
-\caption{Total pion yield.}
+\caption{Measured total pion yield in the $\tin{132}{124}$ and $\tin{108}{112}$ systems.}
 \label{tb:pionyield}
@@
 In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
-Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Fig.~\ref{fig:totalpiRatio}.
+Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Figure~\ref{fig:totalpiRatio}.
 Here the dependence is fitted with a 2-nd order polynomial fit.
@@
 
-\begin{comment}
-\begin{figure}[!htb]
-\centering
-\includegraphics[width=\textwidth]{pionSpectra.png}
-\caption{Pion spectra. }
-\label{fig:pionspectra}
-\end{figure}
-%need to add error analysis
-\end{comment}
-
 \begin{figure}[!htb]
@@
 \includegraphics[width=\textwidth]{pionSpectra_sn108_sum.png}
-\caption{Pion spectra $\tin{108}{112}$ system. }
+\caption{Pion spectra for the $\tin{108}{112}$ system and comparisons to 7 theoretical models. }
 \label{fig:pionspectraSn108}
@@
 \includegraphics[width=\textwidth]{pionSpectra_sn132_sum.png}
-\caption{Pion spectra $\tin{132}{124}$ system. }
+\caption{Pion spectra for the $\tin{132}{124}$ system and comparisons to 7 theoretical models. }
 \label{fig:pionspectraSn132}
@@
 \includegraphics[width=\textwidth]{singleRatio108_log.png}
-\caption{Single ratio $\tin{108}{112}$ system.}
+\caption{Pion spectral ratio $\tin{108}{112}$ system and comparisons to 7 theoretical models.}
 \label{fig:SRsn108}
@@
 \includegraphics[width=\textwidth]{singleRatio132_sum.png}
-\caption{Single ratio $\tin{132}{124}$ system.}
+\caption{Pion spectral ratio $\tin{132}{124}$ system and comparisons to 7 theoretical models.}
 \label{fig:SRsn132}

### 2020-07-17 chapters/ch_summary.tex
Almost final draft

@@
 \chapter{Summary}
+\label{chap:summary}
+
 The need for precise experimental constraints on the symmetry energy at high densities is paramount to understanding exotic, dense matter, such as neutron stars.

### 2020-07-18 chapters/ch_data_analysis.tex
Submitted to Committee

@@
 \end{figure}
+
+We can write the Gaussian fit for the pions as,
+
+\begin{equation}
+G_\pi(z) = A_\pi e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2}},
+\end{equation}
+
+and for the background spectra, in the pion cases $e^+$ and $e^-$:
+
+\begin{equation}
+G_B(z) = A_B e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2} }.
+\end{equation}
+
+The probability that a particle is a pion can be described as,
+
+\begin{equation}
+P(\pi) = \left( 1 + \frac{A_B}{A_\pi} e^{-\frac{(x-\mu_\pi)^2}{2\sigma_\pi^2}  + \frac{(x-\mu_B)^2}{2\sigma_B^2} } \right)^{-1}
+\label{eq:pionProbs}
+\end{equation}
+
+The probability value will be used later as a weighting factor when filling the histogram track-by-track as will be discussed in Section~\ref{sec:corrSpectra}.

### 2020-07-18 chapters/ch_data_analysis_2.tex
Submitted to Committee

@@
 C_a = \frac{4\pi}{\Delta\Omega},
+\label{eq:acceptCorrFactor}
 \end{equation}
@@
 \subsection{Correcting Pion Spectra (Efficiency and Acceptance)}
-
+\label{sec:corrSpectra}
 To correct for the efficiency, we apply a track-by-track correction, where the efficiency of the i-th track $\epsilon_i$ is retrieved from the efficiency database, from the parameters discussed in Section~\ref{sec:efficiency}.
@@
 C_i = \epsilon_i^{-1}.
+\label{eq:effCorrFactor}
 \end{equation}
@@
 The transformation of each  track is defined as $p^{CM} = \textbf{A}p^{Lab}$.
+
+\subsection{Pion Statistical Error}
+
+The treatment of statistical error that will be discussed here applies to the total pion yield and the binned contents of the pion spectral ratio.
+The total pion multiplicity is defined as the number of pions per event.
+Accounting for all the corrections and cuts discussed above this can be expressed as,
+
+\begin{equation}
+Y(\pi^\pm) = A\sum_{i} \frac{P(\pi)}{\epsilon_i},
+\end{equation}
+
+where $P(\pi)$ is the probability that a track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}.
+The factor $A$ is a scale factor defined as,
+
+\begin{equation}
+ A = \frac{C_a}{N_{total}},
+\end{equation}
+
+where $N_{total}$ is the total number of events, $C_a$ is the correction factor accounting for the solid angle measured as given in Eq.~\ref{eq:acceptCorrFactor}.
+The statistical uncertainty is broken into two different parts, due to the way ROOT calculates the error of each histogram bin entry with the Sumw2() function, which calculates the error of each bin as,
+
+\begin{equation}
+{\delta b}^2 = \sum_i^{N_{\pi}} {\delta w_i}^2,
+\end{equation}
+
+where the weight of the bin is $w_i = A P(\pi) \epsilon_i^{-1}$.
+This still does not account for the efficiency error which was calculated using the MC embedding method.
+The efficiency class returns the error of each efficiency bin which scales like the total number of input track $\sqrt{N_{MC}}$.
+The extra error coming from the efficiency error can be expressed as,
+
+\begin{equation}
+{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
+\end{equation}
+
+These two errors make up the total statistical error accounting for the measured number of particles, via Poisson statistics, and the efficiency error following error propagation.
+They are combined in quadrature such that the total error is $\delta \sigma^2 = \delta\epsilon^2 + \delta b^2$.
+The systematic error analysis is discussed in Appendix~\ref{sec:cutvar}.
+This discussion focuses on the systematics originating from the particular cuts we take in our analysis.
+As will be seen in the results section, the pion ratios seem to cancel out any of these effects and therefore any systematics introduced through our cuts need not be added into the total error bars.
+For the single particle yields some systematic error bars are added as guided by the cut variation analysis.

### 2020-07-18 chapters/ch_results.tex
Submitted to Committee

@@
 \includegraphics[width=\textwidth]{ypt_sn132.png}
-\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{132}{124}$ system.}
+\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{132}{124}$ system.
+The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations.
+Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
 \label{fig:ptrap_sn132}
@@
 \includegraphics[width=\textwidth]{ypt_sn108.png}
-\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system.}
+\caption{Rapidity and transverse momentum plot for $\pi^-$ and $\pi^+$ in the $\tin{108}{112}$ system.
+The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations.
+Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
 \label{fig:ptrap_sn108}
@@
 Notice that the pion ratio is significantly greater than the $N/Z$ of the system.
-This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ ~ (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
+This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ \sim (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
 For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.

### 2020-08-05 MSU-thesis_estee_MaxFormat.tex
added new format

%%This is the final version of the template I ended up using - your milage may vary.
%%In particular, the class file with this is set up to play nice with the hyperref
%%package and produce good bookmarks in the PDF file, as well as conform slightly better
%%to the grad school formatting guidelines than the original by Cliff Weil.  
%%Cliff did do the vast bulk of the work on this, however, and I would be remiss
%%if I didn't explictly point out that this is his file with some minor modifications
%%that I added.  - Dan Alt

%% Before beginning to type your dissertation, read the formatting guide, 
%% which can be found at http://grad.msu.edu/etd/docs/formattingguide.pdf
%% Also get the latest version of  msuphddissertation.cls and the template file
%% at http://www.math.msu.edu/~weil/MSU_Ph.D._Dissertation.zip
%% Send questions to weil@math.msu.edu
\documentclass{altmsuphddissertation}

%% Insert packages you wish to use except setspace and subfig. 
%% Those packages are loaded automatically.
%% IMPORTANT: Load only those packages you know you will use.
%% Some packages can cause conflicts resulting in improper formatting.

%% These are the ones I used - you probably don't need them all.  I'd keep hyperref, though. - DMA

\usepackage{caption}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage{overpic}
\usepackage{contour}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{rotating}
\usepackage{mathtools}
\usepackage{comment}\renewcommand{\arraystretch}{1.5}
\usepackage[version=3]{mhchem}
\usepackage{newtxtext,newtxmath,booktabs,siunitx}
\usepackage[export]{adjustbox}

\sisetup{input-uncertainty-signs}

%User defined commands
\newcommand{\spirit}{S$\pi$RIT }
\newcommand{\numerr}[3]{ \num[parse-numbers=false]{#1 \pm #2 (sys.) \pm #3 (stat.)}  }
\newcommand{\tin}[2]{{}^{#1}\mathrm{Sn} + {}^{#2}\mathrm{Sn}}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\renewcommand{\floatpagefraction}{.8}%
\makeatletter
\DeclareTextCommand{\textprime}{\encodingdefault}{%
  \mbox{$\m@th'\kern-\scriptspace$}% 
}
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt plus 1fil}
\makeatother
\newcolumntype{s}{>{\hsize=.5\hsize}X}

\DeclareSIUnit\torr{Torr}
\DeclareSIUnit\eVperc{\eV\per\clight}
\DeclareSIUnit\MeVA{A \mega\eV }
\DeclareSIUnit\clight{\text{\ensuremath{c}}}
\DeclareSIUnit\micron{\micro\metre}
\DeclareSIUnit\mrad{\milli\rad}
\DeclareSIUnit\gauss{G}

%% Use this if you want to put all your graphics in a specific directory - it's a relative path from
%% wherever you store this template. - DMA
\graphicspath{{images/spacecharge/} {images/} {images/mccomparison/}  {images/dataAnalysis_2/} {images/dynamicrange/} {images/dataAnalysis/} {images/experiment/} {images/results/}  {images/intro/} {images/appendix/} }

\author{Justin Brian Estee} %% Put your name in full as it is officially recognized by Michigan State University here.
\title{Charged pion emission from neutron-rich heavy ion collisions for studies on the symmetry energy} %% Put the title of your dissertation here.
\unit{Physics --- Doctor of Philosophy}
%%GRADUATE DEGREE GRANTING UNIT/PROGRAM
%% and the DEGREE  from the complete list of these at
%% http://grad.msu.edu/etd/docs/DegreeGrantingUnits.pdf

%% Put additional preamble items here.

%%%%%% LANDSCAPE PAGES  %%%%%%
%% To produce text, graphics or tables in
%% landscape mode, use the following template.
%\begin{landscape}
%\thispagestyle{empty}
%\begin{minipage}{\linewidth}
%% Insert the graphics or table that is to be
%% in landscape mode. Use the figure or table
%% environment. You may wish to center 
%% the float. 
%\end{minipage}
%\end{landacape}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%  NOTE   %%%%%%%%%%%%%%
%% PREPARING A DISSERTATION WITH THIS CLASS FILE DOES NOT %%%
%% GUARANTEE THAT THE GRADUATE SCHOOL WILL APPROVE IT %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%% WARNING %%%%%%%%%%%%%%%
%% The Graduate School requires that all text, except superscripts %%
%% and subscripts, but including text in imported %%
%% graphics files be in 12 point. For that reason it's recommended %%
%% that no text be part of any imported files. %%

%% Once your document has been filed with the Graduate School,
%% if you wish to produce a version of it whose subscripts and superscripts
%% are in traditional smaller proportion, remove the "%" sign 
%% in front of following command. 
%\DeclareMathSizes{12}{12}{10}{8}
%% To single space your document, find and remove the 
%% two commands \begin{doublespace}
%% and \end{doublespace} below.

%%\includeonly{TeX_files/chapter1,TeX_files/chapter2,TeX_files/chapter3}

\begin{document}
\bibliographystyle{altbibstyle}  %%This is my  custom bibliography style file.  Change if you like. - DMA
\frontmatter
\maketitlepage %%This command will produce the title page of your thesis.
\begin{abstract}

Recent astronomical measurements of neutron star mergers have reinvigorated questions about the nature of dense matter in the universe. The densities reached in these stars range from normal nuclear densities $\rho_o$ to upwards of $4\rho_o$ or even $9\rho_o$ in some models, where $\rho_o = \SI{1.7e14}{\gram\per\centi\metre\cubed}$; representing extremely dense, neutrons rich matter. The symmetry energy describes the difference in binding energy between pure neutron matter and symmetric matter with equal number of neutrons and protons. It is the density dependence of the symmetry energy that is related to the outward pressure, which supports the star, preventing it from collapsing under its own gravitation force. For years, theory and experiment  have combined to progressively constrain the Equation of State (EoS) of symmetric matter, and it was only in the last decade that the density dependence of pure neutron matter, the symmetry energy, had been constrained at normal nuclear densities and lower, $\rho < \rho_o$. Still large uncertainties, and discrepancies on preliminary constraints,  still remain at densities around 2$\rho_o$ which are critical importance to the understanding the properties of neutron stars.

Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities and proton-neutron asymmetries. It has been shown that charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonances, in the early, dense parts of Heavy Ion Collisions (HICs). Yet, existing pion experimental data tailored to answering questions about the symmetry energy was lacking and suffers from unknown systematic errors. this prompted the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pions produced in neutron-rich HICs with a high efficiency. This Dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems measured at \SI{270}{\MeVA} taken at the RIBF facility at RIKEN. The total pion ratio was measured to within 4\% and the pion spectra was measured with a high degree of accuracy, marking the first time pion spectra  -- at energies relevant to the symmetry energy -- have been published.  Also, we will compare the results with the most recent transport models which can simulate pion production. We will highlight some of the successes and hope to motivate a discussion in the theoretical community on how to better reproduce the pion phenomena in neutron-rich HICs, which is important for not only constraining the density dependence of the symmetry energy, but also for understanding the $\Delta$ and pion roles in neutron stars. 
\end{abstract}

%% If you wish to have a copyright page, remove the "%" in front of \begin{copyrt}
%% and remove the "%" in front of \end{copyrt}.
%% The mandatory form of the Copyright will be generated automatically. 
%% A copyright statement is optional.

%\begin{copyrt}
%\end{copyrt}

%% If you wish to have a dedication, remove the "%" in front of
%% \begin{dedication}
%% and remove the "%" in front of 
%% \end{dedication}
%% A dedication must be single-spaced and 
%% centered on the page.  Both will be done automatically.

\begin{dedication}
\centering
{\it
}
\end{dedication}

%% If you wish to have an acknowledgment, remove the "%" in front of  \begin{acknowledgment}
%% and remove the "%" in front of  \end{acknowledgment}  
\begin{acknowledgment}

\end{acknowledgment}

%% If you wish to have a preface, remove the "%" in front of \begin{preface}
%% and remove the "%" in front of \end{preface}. The formatting of
%% a preface isn't specified.
%\begin{preface}
%% Type your preface here. A preface is optional.
%\end{preface}

\TOC %% This command produces the Table of Contents. DO NOT REMOVE!

%% Remove \addvspace (extra space) between figures/tables of different chapters
%%\let\origaddvspace\addvspace
%%\renewcommand{\addvspace}[1]{}
%% If your document contains tables, remove the "%" in front of 
%%  the following line.
\LOT

%% If your document contains figures, remove the "%" in front of
%% the following line.
\LOF
%%\renewcommand{\addvspace}[1]{\origaddvspace{#1}}
%% If any of your figures contain color, you must
%% include the following disclaimer in the caption of your first figure.
%% "For interpretation of the references to color in this and all other figures, 
%% the reader is referred to the electronic version of this thesis."

%%%% LIST OF SYMBOLS OR LIST OF ABBREVIATIONS %%%%
%% If you wish to have a list of symbols or a list of abbreviations, 
%% it should be here. For a list of symbols remove the "%" in front of 
%% \begin{symbols} and remove the "%" in front of \end{symbols}.
%\begin{symbols}
%% Type your list using a list environment here.
%\end{symbols}
%% Similarly for a list of abbreviations remove the "%" in front of 
%% \begin{abbrev} and remove the "%" in front of \end{abbrev}.

%% Replace ANL and ARTEMIS with your own abbreviations.  Or leave them in.  It's up to you! - DMA
%\begin{abbrev}
%\begin{itemize}
%\end{itemize}
%\end{abbrev}

%% The list will be included in the TOC as
%% KEY TO SYMBOLS or KEY TO ABBREVIATIONS
%%%%%%%%%%

\newpage
\pagenumbering{arabic}
\begin{doublespace}

%%If the introduction is to be an unnumbered chapter, put it here.

%% Put the body of your dissertation here. 
%% DO NOT include  the bibliography

\mainmatter
%%I chose to have each of my chapters in a separate file in a subdirectory called ``chapters''.  If you want to do that, 
%%here's how: - DMA
\include{chapters_altFormat/ch_intro}
\include{chapters_altFormat/ch_experiment}
\include{chapters_altFormat/ch_data_analysis}
\include{chapters_altFormat/ch_data_analysis_2}
\include{chapters_altFormat/ch_results}
\include{chapters_altFormat/ch_summary}
%%%%%%%    APPENDICES    %%%%%%%%%%
%% Uncomment ``\appendix'' for one appendix, ``\appendices'' for more than one, and neither for none.
%%\appendix
\appendices
\include{chapters_altFormat/ch_appendix}
%%If you're not using the \include method, 
%% just type \chapter{<appendix 1 name>}
%% and enter the text of the appendix as you would a chapter.

\end{doublespace}

%%%%%%  Bibliography %%%%%
%% A bibliography is required. By default it is called, "Bibliography"
%% You may use \UTF{00D2}Literature Cited\UTF{00D3}, \UTF{00D2}Works Cited\UTF{00D3} or \UTF{00D2}References\UTF{00D3} 
%% instead of \UTF{00D2}Bibliography\UTF{00D3} if that is the convention in your discipline. 
%% To do so, place your choice in the  empty argument 
%% of the following command and remove the "%".
%\renewcommand{\bibname}{}

%% The bibliography may be made using BibTeX.
%% To do so the necessary commands must be entered in the 
%% preamble and here. 
%% Note - do NOT include an extension to the filename below - If your bibtex file is ``BibliographyFilename.bib'', do this: - DMA
\bibliography{mybibfile}

%% If the Bibliography is made from scratch,
%% remove the "%" in front of "\begin{thebibliography}{???}"
%% replacing the ??? with the appropriate entry and 
%% remove the "%" in front of "\end{thebibliography}"
%\begin{thebibliography}{}
%%  Enter the bibliography here.
%\end{thebibliography}
%% In either case, the bibliography is automatically entered
%% in the Table of Contents.
\end{document}

### 2020-08-05 chapters_altFormat/ch_appendix.tex
added new format

--- 2020-07-17 chapters/ch_appendix.tex
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pimTotal}
-         \caption{No efficiency correction.}
-         \label{fig:pim_multCutVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pimTotal}
-         \caption{Efficiency corrected.}
-         \label{fig:pim_multCutVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_NoEC_pimTotal} \label{fig:pim_multCutVar_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_pimTotal} \label{fig:pim_multCutVar_EC}}}%
 	\caption{Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_NoEC_pimTotal}
-         \caption{Before efficiency correction.}
-         \label{fig:pim_clustVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_pimTotal}
-         \caption{After efficiency correction.}
-         \label{fig:pim_clustVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[Before efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_NoEC_pimTotal} \label{fig:pim_clustVar_NoEC}}}%
+    \qquad
+    \subfloat[After efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_pimTotal} \label{fig:pim_clustVar_EC}}}%
 	\caption{Y($\pi^-$) when varying the number of cluster cut of the tracks.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_NoEC_pimTotal}
-         \caption{No efficiency correction.}
-         \label{fig:pim_cutVarPOCA_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_pimTotal}
-         \caption{Efficiency corrected.}
-         \label{fig:pim_cutVarPOCA_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_NoEC_pimTotal} \label{fig:pim_cutVarPOCA_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_pimTotal} \label{fig:pim_cutVarPOCA_EC}}}%
 	\caption{Y($\pi^-$) when varying the dOCA cut.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_pipTotal}
-         \caption{No efficiency correction.}
-         \label{fig:pip_multCutVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_pipTotal}
-         \caption{Efficiency corrected.}
-         \label{fig:pip_multCutVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_NoEC_pipTotal} \label{fig:pip_multCutVar_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_pipTotal} \label{fig:pip_multCutVar_EC}}}%
  	\caption{Y($\pi^+$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_NoEC_pipTotal}
-         \caption{Before efficiency correction.}
-         \label{fig:pip_clustVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_pipTotal}
-         \caption{After efficiency correction.}
-         \label{fig:pip_clustVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[Before efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_NoEC_pipTotal} \label{fig:pip_clustVar_NoEC}}}%
+    \qquad
+    \subfloat[After efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_pipTotal} \label{fig:pip_clustVar_EC}}}%
 	\caption{Y($\pi^+$) when varying the number of cluster cut of the tracks.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_NoEC_pipTotal}
-         \caption{No efficiency correction.}
-         \label{fig:pip_cutVarPOCA_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_pipTotal}
-         \caption{Efficiency corrected.}
-         \label{fig:pip_cutVarPOCA_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_NoEC_pipTotal} \label{fig:pip_cutVarPOCA_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_pipTotal} \label{fig:pip_cutVarPOCA_EC}}}%
 	\caption{Y($\pi^+$) when varying the dOCA cut.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_NoEC_TotalRatio}
-         \caption{No efficiency correction.}
-         \label{fig:ratio_multCutVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{132Sn_Multiplicity_TotalRatio}
-         \caption{Efficiency corrected.}
-         \label{fig:ratio_multCutVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_NoEC_TotalRatio} \label{fig:ratio_multCutVar_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{132Sn_Multiplicity_TotalRatio} \label{fig:ratio_multCutVar_EC}}}%
 	\caption{Y($\pi^+$)/Y($\pi^-$) when varying the ${}^{132}$Sn charged particle multiplicity cut. }
@@
 
-\begin{figure}[!htb]
+ \begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_NoEC_TotalRatio}
-         \caption{Before efficiency correction.}
-         \label{fig:ratio_clustVar_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{Clusters_TotalRatio}
-         \caption{After efficiency correction.}
-         \label{fig:ratio_clustVar_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[Before efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_NoEC_TotalRatio} \label{fig:ratio_clustVar_NoEC}}}%
+    \qquad
+    \subfloat[After efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{Clusters_TotalRatio} \label{fig:ratio_clustVar_EC}}}%
 	\caption{Y($\pi^+$)/Y($\pi^-$) when varying the number of cluster cut of the tracks.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_NoEC_TotalRatio}
-         \caption{No efficiency correction.}
-         \label{fig:ratio_cutVarPOCA_NoEC}
-     \end{subfigure}
-     \hfill
-    \begin{subfigure}[b]{\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{POCA_TotalRatio}
-         \caption{Efficiency corrected.}
-         \label{fig:ratio_cutVarPOCA_EC}
-     \end{subfigure}
-     \hfill
+    \subfloat[No efficiency correction.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_NoEC_TotalRatio} \label{fig:ratio_cutVarPOCA_NoEC}}}%
+    \qquad
+    \subfloat[Efficiency corrected.]{{\includegraphics[width=.46\textwidth,valign=c]{POCA_TotalRatio} \label{fig:ratio_cutVarPOCA_EC}}}%
 	\caption{Y($\pi^+$)/Y($\pi^-$) when varying the dOCA cut.}

### 2020-08-05 chapters_altFormat/ch_data_analysis.tex
added new format

--- 2020-07-18 chapters/ch_data_analysis.tex
@@
 
-\begin{figure}[!htb]
-     \centering
-     \begin{subfigure}[b]{0.24\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{psa0.png}
-         \caption{$\mathrm{1^{st}}$ step of the PSA}
-         \label{fig:psa0}
-     \end{subfigure}
-     \hfill
-     \centering
-     \begin{subfigure}[b]{0.24\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{psa1.png}
-         \caption{$\mathrm{2^{nd}}$ iteration}
-         \label{fig:psa1}
-     \end{subfigure}
-     \hfill
-      \centering
-     \begin{subfigure}[b]{0.24\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{psa2.png}
-         \caption{$\mathrm{3^{rd}}$ iteration}
-         \label{fig:psa2}
-     \end{subfigure}
-     \hfill
-      \centering
-     \begin{subfigure}[b]{0.24\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{psa3.png}
-         \caption{$\mathrm{4^{th}}$ iteration}
-         \label{fig:psa3}
-     \end{subfigure}
-     \hfill
-     \centering
-     \begin{subfigure}[b]{.5\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{psa4.png}
-         \caption{Final output of the PSA}
-         \label{fig:psa4}
-     \end{subfigure}
-     \hfill
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[$\mathrm{1^{st}}$ step of the PSA]{{\includegraphics[width=.4\textwidth,valign=t]{psa0.png} \label{fig:psa0}}}%
+    \qquad
+    \subfloat[$\mathrm{2^{nd}}$ iteration]{{\includegraphics[width=.24\textwidth,valign=c]{psa1.png} \label{fig:psa1}}}%
+    \subfloat[$\mathrm{3^{nd}}$ iteration]{{\includegraphics[width=.24\textwidth,valign=c]{psa2.png} \label{fig:psa2}}}%
+    \subfloat[$\mathrm{4^{nd}}$ iteration]{{\includegraphics[width=.24\textwidth,valign=c]{psa3.png} \label{fig:psa3}}}%
+    \subfloat[Final output of the PSA]{{\includegraphics[width=.24\textwidth,valign=c]{psa4.png} \label{fig:psa}}}%
 
@@
 
-\begin{comment}
 \begin{figure}[ht!]s
@@
 \end{figure}
-\end{comment}
 
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_small_angle.pdf} 
-        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, before the space charge correciton.} \label{fig:mom_S_before}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_large_angle.pdf} 
-        \caption{Comparison of the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} > \ang{40}$, before the space charge correction.} \label{fig:mom_L_before}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_aftercor_small_angle.pdf} 
-        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction. } \label{fig:mom_S_after}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_aftercor_large_angle.pdf} 
-        \caption{Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction.} \label{fig:mom_L_after}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, before the space charge correciton.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_small_angle.pdf} \label{fig:psa0}}}%
+    \qquad
+    \subfloat[Comparison of the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} > \ang{40}$, before the space charge correction.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_large_angle.pdf} \label{fig:mom_L_before}}}\\
+    \subfloat[Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_aftercor_small_angle.pdf} \label{fig:mom_S_after}}}%
+    \qquad
+    \subfloat[Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, after the space charge correction.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_aftercor_large_angle.pdf}  \label{fig:mom_L_after}}}%
+
  \label{fig:mom_sc}
@@
 
-\begin{comment}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_frac_small_angle.pdf} 
-        \caption{Generic} \label{fig:mom_S_1D}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{BDC_P_frac_large_angle.pdf} 
-        \caption{Competitors} \label{fig:mom_L_1D}
-    \end{subfigure}
-
-\label{fig:mom_1D}
-\end{figure}
-\end{comment}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{sn132_SC} 
-        \caption{$\tin{132}{123}$ system.} \label{fig:132momdist_sc}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{sn108_SC} 
-        \caption{$\tin{108}{112}$ system.} \label{fig:108momdist_sc}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[$\tin{132}{123}$ system.]{{\includegraphics[width=\textwidth,valign=c]{sn132_SC} \label{fig:132momdist_sc}}}%
+    \qquad
+    \subfloat[$\tin{108}{112}$ system.]{{\includegraphics[width=\textwidth,valign=c]{sn108_SC} \label{fig:108momdist_sc}}}%
 	\caption{Momentum distributions of the Left and Right sides of the TPC for $\pi^+$ and $\pi^-$ particles.}
@@
 
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset6before_row_at_layer10.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset5before_row_at_layer30.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset4before_row_at_layer50.png}
-
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset6after_row_at_layer10.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset5after_row_at_layer30.png}
-    \includegraphics[width=0.33\textwidth]{run2907_yoffset4after_row_at_layer50.png}
-
-  \end{center}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset6before_row_at_layer10.png} \label{}}}%
+    \qquad
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset5before_row_at_layer30.png} \label{}}}%
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset4before_row_at_layer50.png} \label{}}}%\\
+
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset6after_row_at_layer10.png} \label{}}}%
+	\qquad
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset5after_row_at_layer30.png} \label{}}}%
+    \subfloat[]{{\includegraphics[width=.3\textwidth,valign=c]{run2907_yoffset4after_row_at_layer50.png} \label{}}}%
+
 	\caption{Cobo timing correction for three different layers.
@@
 
-\begin{figure}[!htb]
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{run2905_yoffset1before_all.jpg} 
-        \caption{Before applying the timing correction.} \label{fig:yoff_allBefore}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{run2905_yoffset1after_all.png}
-        \caption{After applying the timing correction.} \label{fig:yoff_allAfter}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Before applying the timing correction.]{{\includegraphics[width=.46\textwidth,valign=c]{run2905_yoffset1before_all.jpg} \label{fig:yoff_allBefore}}}%
+    \qquad
+    \subfloat[After applying the timing correction.]{{\includegraphics[width=.46\textwidth,valign=c]{run2905_yoffset1after_all.png} \label{fig:yoff_allAfter}}}%
+
     \caption{Y-residual distributions for all pads.}
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{mcresponse.png}
-        \caption{200 MeV/c $\pi^-$ MC digitized track} \label{fig:mcevent}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{event5.png} 
-        \caption{Experimental data event.} \label{fig:dataevent}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[200 MeV/c $\pi^-$ MC digitized track]{{\includegraphics[width=.46\textwidth,valign=c]{mcresponse.png} \label{fig:mcevent}}}%
+    \qquad
+    \subfloat[Experimental data event.]{{\includegraphics[width=.46\textwidth,valign=c]{event5.png} \label{fig:dataevent}}}%
     \caption{Charge readout of the pad plane, showing the max ADC in a given pad, for both the MC simulation and experimental data.}
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{perspective_embed}
-        \caption{Perspective view.} \label{fig:persEmbed}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.3\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{top_embed} 
-        \caption{Top down view.} \label{fig:topEmbed}
-    \end{subfigure}
-     \hfill
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{side_embed} 
-        \caption{Side view.} \label{fig:sideEmbed}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Perspective view.]{{\includegraphics[width=.5\textwidth,valign=c]{perspective_embed} \label{fig:persEmbed}}}%
+    \qquad
+    \subfloat[Top down view.]{{\includegraphics[width=.3\textwidth,valign=c]{top_embed} \label{fig:topEmbed}}}%
+    \subfloat[Side view.]{{\includegraphics[width=.55\textwidth,valign=c]{side_embed} \label{fig:sideEmbed}}}%
+
     \caption{A 200 MeV/c $\pi^-$ embedded into a nuclear collision type event. The embedded track identified by the software is highlighted by the solid green line. }
-
 	\label{fig:embedtrack}
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{deltaE_topview_neg} 
-        \caption{Top view} \label{fig:deltaE_topview}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{deltaE_sideview_neg} 
-        \caption{Side view} \label{fig:deltaE_sideview}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Top view]{{\includegraphics[width=.46\textwidth,valign=c]{deltaE_topview_neg} \label{fig:deltaE_topview}}}%
+    \qquad
+    \subfloat[Side view]{{\includegraphics[width=.46\textwidth,valign=c]{deltaE_sideview_neg} \label{fig:deltaE_sideview}}}%
+
   	 \caption{Geant4 simulation of ${}^{108}$Sn beam at 270 MeV/A in P10 gas. Notice the extent of the delta electrons in the vertical direction as compared to the horizontal extent. }
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_200_pol.png}
-        \caption{Polar efficiency plot where $\theta$ is represented by the radius of the circle and $\phi$ is measured as the polar angle made between the y and x axis; this view is  best understood as looking at the TPC from the downstream point of view.} \label{fig:pim_pol_eff_ex}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_200_efficiency.png} 
-        \caption{Cartesian representation of the same efficiency plot  as in the polar plot.} \label{fig:pim_flat_eff_ex}
-    \end{subfigure}
-
+ \begin{figure}[!htb]%
+    \centering
+    \subfloat[Polar efficiency plot where $\theta$ is represented by the radius of the circle and $\phi$ is measured as the polar angle made between the y and x axis; this view is  best understood as looking at the TPC from the downstream point of view.]{{\includegraphics[width=.46\textwidth,valign=c]{pim_200_pol.png} \label{fig:pim_pol_eff_ex}}}%
+    \qquad
+    \subfloat[Cartesian representation of the same efficiency plot  as in the polar plot.]{{\includegraphics[width=.46\textwidth,valign=c]{pim_200_efficiency.png} \label{fig:pim_flat_eff_ex}}}%
 	\caption{Efficiency calculations for $\pi^-$ particle traveling at \SI{200}{\mega\eVperc}.}
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_132_efficiency.png}
-        \caption{$\pi^-$ efficiency for $\tin{132}{124}$} \label{fig:pim_132_eff}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pip_132_efficiency.png} 
-        \caption{$\pi^+$ efficiency for $\tin{132}{124}$} \label{fig:pip_132_eff}
-    \end{subfigure}
-
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[$\pi^-$ efficiency for $\tin{132}{124}$]{{\includegraphics[width=.46\textwidth,valign=c]{pim_132_efficiency.png} \label{fig:pim_132_eff}}}%
+    \qquad
+    \subfloat[$\pi^+$ efficiency for $\tin{132}{124}$]{{\includegraphics[width=.46\textwidth,valign=c]{pip_132_efficiency.png} \label{fig:pip_132_eff}}}%
     \caption{Pion efficiencies in the $\tin{132}{124}$ system. }
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_132_efficiency.png}
-        \caption{$\pi^-$ efficiency for $\tin{108}{112}$} \label{fig:pim_108_eff}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pip_108_efficiency.png} 
-        \caption{$\pi^+$ efficiency for $\tin{108}{112}$} \label{fig:pip_108_eff}
-    \end{subfigure}
-
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[$\pi^-$ efficiency for $\tin{108}{112}$]{{\includegraphics[width=.46\textwidth,valign=c]{pim_132_efficiency.png} \label{fig:pim_108_eff}}}%
+    \qquad
+    \subfloat[$\pi^+$ efficiency for $\tin{108}{112}$]{{\includegraphics[width=.46\textwidth,valign=c]{pip_108_efficiency.png} \label{fig:pip_108_eff}}}%
  	\caption{ Pion efficiencies in the $\tin{108}{112}$ system. }
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pidPion.png}
-        \caption{Output PID pions.} \label{fig:pidpion_raw}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pidPion_zi.png} 
-        \caption{PID plot for $\pi^-$ and $\pi^+$ species in $z_i$. } \label{fig:pipion_zi}
-    \end{subfigure}
-
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Output PID pions.]{{\includegraphics[width=.46\textwidth,valign=c]{pidPion.png} \label{fig:pidpion_raw}}}%
+    \qquad
+    \subfloat[PID plot for $\pi^-$ and $\pi^+$ species in $z_i$. ]{{\includegraphics[width=.46\textwidth,valign=c]{pidPion_zi.png} \label{fig:pipion_zi}}}%
  	\caption{The charged pion PID plot achieved in the $\tin{132}{124}$ system.
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_mom2.png}
-        \caption{Bin -54 MeV/c < p/q < -52. } \label{fig:pimmom2}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pim_mom1.png} 
-        \caption{Bin -68 MeV/c < p/q < -66.} \label{fig:pimmom1}
-    \end{subfigure}
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Bin -54 MeV/c < p/q < -52. ]{{\includegraphics[width=.46\textwidth,valign=c]{pim_mom2.png} \label{fig:pimmom2}}}%
+    \qquad
+    \subfloat[Bin -68 MeV/c < p/q < -66.]{{\includegraphics[width=.46\textwidth,valign=c]{pim_mom1.png} \label{fig:pimmom1}}}%
 
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pip_mom1.png}
-        \caption{Bin 74 MeV/c < p/q < 76. } \label{fig:pipmom1}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{pip_mom2.png} 
-        \caption{Bin 94 MeV/c < p/q < 86.} \label{fig:pipmom2}
-    \end{subfigure}
-
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Bin 74 MeV/c < p/q < 76.]{{\includegraphics[width=.46\textwidth,valign=c]{pip_mom1.png} \label{fig:pipmom1}}}%
+    \qquad
+    \subfloat[Bin 94 MeV/c < p/q < 86.]{{\includegraphics[width=.46\textwidth,valign=c]{pip_mom2.png} \label{fig:pipmom2}}}%
  	\caption{Flattened PID $z_i$ distribution  for binned slices in p/q around the $\pi^+$.}

### 2020-08-05 chapters_altFormat/ch_data_analysis_2.tex
added new format

--- 2020-07-18 chapters/ch_data_analysis_2.tex
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{beam-Sn132.png} \label{fig:beampid132}}}%
+    \qquad
+    \subfloat[$\tin{108}{112}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{beam-Sn108.png} \label{fig:beampid108}}}%
 
-     \centering
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{beam-Sn132.png}
-         \caption{$\tin{132}{124}$ system.}
-         \label{fig:beampid132}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{beam-Sn108.png}
-         \caption{$\tin{108}{112}$ system.}
-         \label{fig:beampid108}
-     \end{subfigure}
 	\caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
+    \centering
+    \subfloat[Top view of the tracks showing the edge effect on the left and right sides of the TPC.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterLR.png} \label{fig:clusterLR}}}%
+    \qquad
+    \subfloat[Side view of the TPC showing the edge effect near the top.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterTop.png} \label{fig:clusterTop}}}%
+    \subfloat[Side view of the TPC showing the edge effect near the bottom.]{{\includegraphics[width=.45\textwidth,valign=c]{clusterBottom.png} \label{fig:clusterBottom}}}%
 
-    \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{clusterLR.png} 
-        \caption{Top view of the tracks showing the edge effect on the left and right sides of the TPC.} 	   \label{fig:clusterLR}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{clusterTop.png} 
-        \caption{Side view of the TPC showing the edge effect near the top.} \label{fig:clusterTop}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{clusterBottom.png} 
-        \caption{Side view of the TPC showing the edge effect near the bottom.} \label{fig:clusterBottom}
-    \end{subfigure}
   	\caption{Example showing the edge effect has on the cluster positions near the extreme edges of the TPC volume. }    
@@
 
-\begin{figure}[!htb]
-
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{sideviewHighDenCut.jpg}
-         \caption{Side view.}
-         \label{fig:sideHigh}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{topviewHighDenCut.jpg}
-         \caption{Top view.}
-         \label{fig:topHigh}
-     \end{subfigure}
+    \subfloat[Side view.]{{\includegraphics[width=.46\textwidth,valign=c]{sideviewHighDenCut.jpg} \label{fig:sideHigh}}}%
+    \qquad
+    \subfloat[Top view.]{{\includegraphics[width=.46\textwidth,valign=c]{topviewHighDenCut.jpg} \label{fig:topHigh}}}%
  	  \label{fig:highcut}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{beamAngle-Sn108.png} 
-        \caption{Beam angle of the $\tin{108}{124}$ system} \label{fig:beamangle108}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{beamAngle-Sn112.png} 
-        \caption{Beam angle of the $\tin{112}{124}$ system} \label{fig:beamangle112}
-    \end{subfigure}
+    \subfloat[Beam angle of the $\tin{108}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn108.png} \label{fig:beamangle108}}}%
+    \qquad
+    \subfloat[Beam angle of the $\tin{112}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn112.png} \label{fig:beamangle112}}}%
 
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{beamAngle-Sn124.png} 
-        \caption{Beam angle of the $\tin{124}{112}$ system} \label{fig:beamangle124}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{beamAngle-Sn132.png} 
-        \caption{Beam angle of the $\tin{132}{124}$ system} \label{fig:beamangle132}
-    \end{subfigure}
+    \subfloat[Beam angle of the $\tin{124}{112}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn124.png} \label{fig:beamangle124}}}%
+    \qquad
+    \subfloat[Beam angle of the $\tin{132}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn132.png} \label{fig:beamangle132}}}%
+
 	\caption{Beam angle at the face of the target in the TPC for all 4 systems.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{vertex-Sn108.png} 
-        \caption{$\tin{108}{124}$ system.} \label{fig:vertex108}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{vertex-Sn112.png} 
-        \caption{$\tin{112}{124}$ system.} \label{fig:vertex112}
-    \end{subfigure}
+    \subfloat[$\tin{108}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn108.png} \label{fig:vertex108}}}% 
+    \qquad
+    \subfloat[$\tin{112}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn112.png} \label{fig:vertex112}}}% 
+    \\
+    \subfloat[$\tin{124}{112}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn124.png} \label{fig:vertex124}}}% 
+    \\
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{vertex-Sn132.png} \label{fig:vertex132}}}%
 
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{vertex-Sn124.png} 
-        \caption{$\tin{124}{112}$ system.} \label{fig:vertex124}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{vertex-Sn132.png} 
-        \caption{$\tin{132}{124}$ system.} \label{fig:vertex132}
-    \end{subfigure}
   	\caption{Vertex distributions, around the target region, of all experimental runs for each secondary beam system. }
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{angularCut-pim-Sn108.png}
-         \caption{$\pi^-$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
-         \label{fig:pim108angle}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{angularCut-pim-Sn132.png}
-         \caption{$\pi^-$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
-         \label{fig:pim132angle}
-     \end{subfigure}
+    \subfloat[$\pi^-$ angular cuts in the $\tin{108}{112}$ system.
+$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
+    \qquad
+    \subfloat[$\pi^-$ angular cuts in the $\tin{132}{124}$ system.
+$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
+
 	\caption{Measured $\pi^-$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{angularCut-pip-Sn108.png}
-         \caption{$\pi^+$ angular cuts in the $\tin{108}{112}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
-         \label{fig:pip108angle}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.49\textwidth}
-         \centering
-         \includegraphics[width=\textwidth]{angularCut-pip-Sn132.png}
-         \caption{$\pi^+$ angular cuts in the $\tin{132}{124}$ system. $\theta_{CM} < 90$ and < $\phi$ <}
-         \label{fig:pip132angle}
-     \end{subfigure}
+    \subfloat[$\pi^+$ angular cuts in the $\tin{108}{112}$ system.
+$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
+    \qquad
+    \subfloat[$\pi^+$ angular cuts in the $\tin{132}{124}$ system.
+$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
      \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}

### 2020-08-05 chapters_altFormat/ch_experiment.tex
added new format

--- 2020-07-17 chapters/ch_experiment.tex
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{wires_open.pdf} 
-        \caption{Wires open} \label{fig:wires_open}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.42\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{wires_closed.pdf} 
-        \caption{Wires closed.} \label{fig:wires_closed}
-    \end{subfigure}
-\caption{Cartoon depiction of electron drift lines going through the various wire planes.}
-\label{fig:wires}
+    \subfloat[Wires open]{{\includegraphics[width=.45\textwidth,valign=t]{wires_open.pdf} \label{fig:wires_open}}}%
+    \qquad
+    \subfloat[Wires closed]{{\includegraphics[width=.39\textwidth,valign=t]{wires_closed.pdf} \label{fig:wires_closed}}}%
+    \caption{Cartoon depiction of electron drift lines going through the various wire planes.}%
+    \label{fig:wires}%
 \end{figure}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{averageTransparency.png} 
-        \caption{On configurations where all wires are the same voltage.} \label{fig:ggAvgTrans}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{transparencyDeltaV.png} 
-        \caption{Off configuration where adjacent wires have a voltage difference of $2 \Delta V$.} \label{fig:ggDeltaVTrans}
-    \end{subfigure}
-
+    \subfloat[On configurations where all wires are the same voltage.]{{\includegraphics[width=.45\textwidth,valign=t]{averageTransparency.png} \label{fig:ggAvgTrans}}}%
+    \qquad
+    \subfloat[Off configuration where adjacent wires have a voltage difference of $2 \Delta V$.]{{\includegraphics[width=.45\textwidth,valign=t]{transparencyDeltaV.png} \label{fig:ggDeltaVTrans}}}%
 \caption{Electron transparency as measured and simulated in the on and off configurations of the gating grid.}    
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!htb]%
     \centering
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{DataTrigger1.png} 
-        \caption{${}^{124}$Xe primary beam trigger.} \label{fig:dataTrigger1}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.45\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{DataTrigger2.png} 
-        \caption{${}^{238}$U primary beam trigger.} \label{fig:dataTrigger2}
-    \end{subfigure}
+    \subfloat[${}^{124}$Xe primary beam trigger.]{{\includegraphics[width=.45\textwidth,valign=t]{DataTrigger1.png} \label{fig:dataTrigger1}}}%
+    \qquad
+    \subfloat[${}^{238}$U primary beam trigger.]{{\includegraphics[width=.45\textwidth,valign=t]{DataTrigger2.png} \label{fig:dataTrigger2}}}%
 	\caption{Master triggers in each primary beam setting.}    

### 2020-08-05 chapters_altFormat/ch_intro.tex
added new format

--- 2020-07-17 chapters/ch_intro.tex
@@
 \chapter{Introduction}
-On August 17, 2017, a collision of two  neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with many other astronomical observatories, marking a new era of advanced multi-messenger observations \cite{ns_2017}, dramatically increasing already very strong interesting in  questions concerning the nature of dense matter in the universe.
+On August 17, 2017, a collision of two neutron stars produced gravitational waves which were measured by the LIGO and Advanced Virgo detectors, in combination with many other astronomical observatories, marking a new era of advanced multi-messenger observations \cite{ns_2017}; reinvigorating questions concerning the nature of dense matter in the universe.
 
-Mankind has been interested in the nature of the matter which composes the visible universe, and what are its \emph{constituents} or building blocks.
+Mankind has been continuously interested in the nature of the matter which composes the visible universe, and what are its constituents or building blocks.
 The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie.
-His studies led him to conclude that atoms consist of a dense positively charged atomic nucleus surrounded by a dilute electron cloud.
-Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to positively charged protons as the basic constituents of nuclear matter.
-Remarkably, the nucleus itself accounts for 99.9\% of the mass of the atom and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.
+His studies led him to conclude that atoms consist of a dense, positively charged atomic nucleus, surrounded by a dilute electron cloud.
+Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to the known positively charged particle -- protons -- as the basic constituents of nuclear matter.
+Remarkably, the nucleus accounts for 99.9\% of the mass of the atom, and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.
 
@@
 This strong force is a manifestation of the fundamental interactions between the quarks which constitute each nucleon.
-Three quarks of two types or \emph{flavors} labeled \emph{up} and \emph{down} with nearly equal masses are needed to form each nucleon.
+Three quarks, of two types or \emph{flavors} labeled \emph{up} and \emph{down}, with nearly equal masses form each nucleon.
 At low energies the internal quark structure of nucleons is less important and nucleons can be thought of as fundamental particles.
@@
 Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich.
-Thus the asymmetry term above is proportional to the factor $\frac{(N - Z)^2}{A}$ which comes directly from Pauli blocking.
+Thus the asymmetry term above is proportional to the factor $(N - Z)^2/A$ which comes directly from Pauli blocking.
 While Eq.~\ref{eq:semiEmp} describes normal nuclear matter, at saturation density, quite well, a more general description must be made when going to higher or lower densities.
@@
 
-Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately \SI{11(2)}{\kilo\metre}, though they weigh about 1.5 times the mass of our sun.
+Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately $11^{+2}_{-2}$ \si{\kilo\metre}, though they weigh about 1.5 times the mass of our sun.
 Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the balancing outward force arising from the strong force which exerts a large outward pressure arising from the symmetry energy at high densities.

### 2020-08-05 chapters_altFormat/ch_summary.tex
added new format

--- 2020-07-17 chapters/ch_summary.tex
@@
 Direct observation of these objects provides information about the total equation of state, which is primarily sensitive to the pressure in neutron matter.
-Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta = \frac{\rho_n - \rho_p}{\rho}$.
+Only by laboratory measurements can one control the asymmetry of matter and isolate its dependence on the isospin asymmetry $\delta =(\rho_n - \rho_p)/\rho$.
 This information is critical to understanding the symmetry energy and associated isovector mean field potentials.
@@
   This marks the first time pions have been accurately measured in the sub-threshold region.
-We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ ~ (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
+We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ \sim (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
 Such a deviation is also predicted by theory, so this explanation should be retired in favor of something more quantitative and informative.

### 2020-08-05 macrosForThesis/PionSpectra/PlotDataTheory/publishPionsSpectra108Sn_matchHighEnergy.C

@@ -1,0 +1,407 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+
+
+void publishPionsSpectra108Sn_matchHighEnergy()
+{
+
+ TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft.root");
+  TGraphErrors *pBUU_soft_no_pim = (TGraphErrors *) pBUU_soft_no_f->Get("pim108");
+  TGraphErrors *pBUU_soft_no_pip = (TGraphErrors *) pBUU_soft_no_f->Get("pip108");
+  
+  TFile *pBUU_stiff_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUStiff.root");
+  TGraphErrors *pBUU_stiff_no_pim = (TGraphErrors *) pBUU_stiff_no_f->Get("pim108");
+  TGraphErrors *pBUU_stiff_no_pip = (TGraphErrors *) pBUU_stiff_no_f->Get("pip108");
+  
+  TFile *amdJAM55_f = TFile::Open("/home/justin/Homework/AMDJAM/highStat/input_for_esym.info/AMDJAML55new.root");
[…]

### 2020-08-05 macrosForThesis/PionSpectra/SpaceChargeVariation/plotSpaceCharge.C

@@ -26,9 +26,9 @@ void plotSpaceCharge()
   TH1D *pim_132_plus = (TH1D *)sn132_plus->Get("pimkecm_ec");
   TH1D *pip_132_plus = (TH1D *)sn132_plus->Get("pipkecm_ec");
 
-  //  pim_132_nom->Divide(pip_132_nom);
-  //  pim_132_minus->Divide(pip_132_minus);
-  //  pim_132_plus->Divide(pip_132_plus);
+  pim_132_nom->Divide(pip_132_nom);
+  pim_132_minus->Divide(pip_132_minus);
+  pim_132_plus->Divide(pip_132_plus);
 
   pim_108_nom->Divide(pip_108_nom);
   pim_108_minus->Divide(pip_108_minus);

### 2020-08-07 MSU-thesis_estee_MaxFormat.tex
Put in Pawel and Vladimir edits

@@
 For years, theory and experiment  have combined to progressively constrain the Equation of State (EoS) of symmetric matter, and it was only in the last decade that the density dependence of pure neutron matter, the symmetry energy, had been constrained at normal nuclear densities and lower, $\rho < \rho_o$.
-Still large uncertainties, and discrepancies on preliminary constraints,  still remain at densities around 2$\rho_o$ which are critical importance to the understanding the properties of neutron stars.
+Still large uncertainties, and discrepancies on preliminary constraints, remain at densities around 2$\rho_o$ which are critical importance to the understanding the properties of neutron stars.
 
@@
 Yet, existing pion experimental data tailored to answering questions about the symmetry energy was lacking and suffers from unknown systematic errors.
-this prompted the construction of the \spirit Time Projection Chamber (TPC) which was constructed to measure charged pions produced in neutron-rich HICs with a high efficiency.
+This prompted the construction of the Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) which was constructed to measure charged pions produced in neutron-rich HICs with a high efficiency.
 This Dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems measured at \SI{270}{\MeVA} taken at the RIBF facility at RIKEN.
-The total pion ratio was measured to within 4\% and the pion spectra was measured with a high degree of accuracy, marking the first time pion spectra  -- at energies relevant to the symmetry energy -- have been published.
+The total pion ratio was measured to within 4\% and the pion spectra are measured with a high degree of accuracy, marking the first time pion spectra  -- at energies relevant to the symmetry energy -- have been published.
 Also, we will compare the results with the most recent transport models which can simulate pion production.

### 2020-08-07 chapters_altFormat/ch_data_analysis.tex
Put in Pawel and Vladimir edits

@@
 Three basic calibrations are performed on the raw data level within this task.
-The gain calibration calibrates the electronics will be described in Section\ref{sec:elecCalib}.
+The calibration of the electronics will be described in Section\ref{sec:elecCalib}.
 The anode gain calibration which will be discussed in Section~\ref{sec:anodeCalib}, and the gating grid noise subtraction and time bucket cut which is described in Section~\ref{sec:ggnoisesub}.
@@
 %cobo gain matching
-An algorithm loops over the raw time bucket spectrum to find all the unique pules in each pad.
+An algorithm loops over the raw time bucket spectrum to find all the unique pulses in each pad.
 It is very likely that there are several pulses in a pad coming from multiple tracks passing under the same pad which are separated only in arrival time.
@@
 Here, the second pulse is found and fitted again with the same procedure; the red line will be subtracted from the raw ADC spectrum.
-The black line represents the sum of all the fitted pules in each step.
-This continues until all pules are found or until the PSA reaches the last bin in the spectrum.
+The black line represents the sum of all the fitted pulses in each step.
+This continues until all pulses are found or until the PSA reaches the last bin in the spectrum.
 Figure~\ref{fig:psa4} shows the final result where 4 fitted pulses were found.
@@
 The y-coordinate is calculated using the known drift velocity $v$ and time of the signal $t_o$ as $y = v\cdot t_0$, $t_0$.
-The hit class forms the fundamental structure from which all tracking is performed since it contains the x,y,z spacial coordinates, and charge information $Q$.
+The hit class forms the fundamental structure from which all tracking is performed since it contains the x,y,z spatial coordinates, and charge information $Q$.
 
@@
 
-  	\caption{Example of the pulse shape algorithm through each step of in processing a given pad. }
+  	\caption{Example of the pulse shape algorithm through each step in processing a given pad. }
 	\label{fig:psaTask}
@@
 This is a particular transformation in which unique helices in the Cartesian space form great circles on the Riemann sphere.
-This is particularly useful since hits which correspond to a track form a helices the Cartesian space of the TPC coordinates.
+This is particularly useful since hits which correspond to a track form a helix in the Cartesian space of the TPC coordinates.
 It is difficult to search for tracks in the hit cloud by fitting a general helix to different combinations of hits.
@@
 In this example, three clusters at the bottom are clustered in the x-direction and for the upper three clustered in the z-direction.
-The cluster representing each group of hits represent the average position of the track.
+The cluster for each group of hits represents the average position of the track.
 The track fit represents the position of the avalanche, $\bar{x}$,  where the position from the center to each pad to the $\bar{x}$ position is defined as $\lambda_i$.}
@@
 You can imagine if we calculated the clusters only along the x-axis, tracks with $\theta \approx 0^{\circ}$ the x-position is not well defined.
-The cluster position and charge is stored into a new container representing the cluster called the STHitCluster class.
+The cluster position and charge are stored into a new container representing the cluster called the STHitCluster class.
 All the clusters which belong to a particular helix track is also stored in the corresponding STHelixTrack class.
@@
 GENFIT utilizes a Kalman fitting algorithm which returns the estimated momentum and charge values for a given track \cite{genfit}.
-These values are then stored in into a new track container called the STRecoTrack class.
+These values are then stored into a new track container called the STRecoTrack class.
 
@@
 Due to coupling between the gating grid, the ground plane, and other effects, the impedance of the supply cables was not perfectly matched, causing an imbalance in the current where one side discharged faster than the other side.
-This caused an under-damped oscillating current in the grid, which created and  oscillating residual noise early in the time bucket spectrum.
+This caused an under-damped oscillating current in the grid, which created an oscillating residual noise early in the time bucket spectrum.
 Figure~\ref{fig:ggNoiseSub} shows in the upper panel shows the ADC time bucket spectrum after averaging over 2000 events in a given pad.
@@
 \end{tabular}
-\caption{Summary of expected cocktail for the calibration run taken with the Al target. }
+\caption{Summary of expected cocktail for the calibration run taken with the Aluminum target. }
 \label{tb:cocktail100tar}
@@
 
-A cocktail of several light charged particles (p,d,t,${}^{3}$He,${}^{4}$He) of well defined magnetic rigidity was produced was produced and measured in the TPC to provide a momentum calibration of the TPC.
+A cocktail of several light charged particles (p,d,t,${}^{3}$He,${}^{4}$He) of well defined magnetic rigidity was produced and measured in the TPC to provide a momentum calibration of the TPC.
 The magnetic and slit settings of the dipoles in the BigRIPS spectrometer were set so that the momentum resolution of the beam was defined to better than $\delta p/p$ < 1\%.
@@
 This is shown in Tables~cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
-We could address this discrepancy by incorporating the smaller horizontal magnetic field components at the edge of the dipole and taking into account the electron drift velocities in the direction of $\vec{E}\times\vec{B}$ direction as seen from Eq.\ref{eq:elecdrift}.
+We could address this discrepancy by incorporating the smaller horizontal magnetic field components at the edge of the dipole and taking into account the electron drift velocities in the direction of $\vec{E}\times\vec{B}$ as seen from Eq.\ref{eq:elecdrift}.
 The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
@@
 
-(Tables~\cref{tb:cocktail100tar,tb:cocktail100}), protons see a slight improvement of about ~1\% where as the deuterons are over corrected in both settings.
+(Tables \ref{tb:cocktail100tar} and \ref{tb:cocktail100}), protons see a slight improvement of about ~1\% whereas the deuterons are over corrected in both settings.
 In general, all these light particles corrected momentum are of the order of 1\% smaller than value given by the rigidity selection in the BigRIPS separator.
@@
 \label{sec:anodeCalib}
-As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment due to high currents being observed on the wires due to a leak in the gating grid that has since been corrected.
+As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment, high currents being observed on the wires due to a leak in the gating grid that has since been corrected \cite{jon}.
 Out of all the runs used in the analysis in this thesis -- see \ref{tb:runList} -- the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all other runs.
@@
 
-Figure~\ref{fig:highlowcal} shows the correlation plot of $\langle dE/dx\angle$ of the high versus the low gain sections.
-The effects of the high gain channel saturation can around a value of 400 ADC\si{\per\milli\metre} and plateaus, where as the low gain does not; this region is left out of the fit.
+Figure~\ref{fig:highlowcal} shows the correlation plot of $\langle dE/dx\rangle$ of the high versus the low gain sections.
+The effects of the high gain channel saturation around a value of 400 ADC\si{\per\milli\metre} and plateaus, whereas the low gain does not; this region is left out of the fit.
 A linear fit was performed for the calibration between the high gain channel $h_c = G\cdot l_c$, where $l_c$ is the value of the low gain channel and $G$ is the gain factor.
-In the case where the low anode voltage was \SI{1214}{\volt}, the $G=9.8$, where in the case of \SI{1085}{\volt}, the $G = $.
+In the case where the low anode voltage was \SI{1214}{\volt}, the gain factor was $G=9.8$.
+
+%where in the case of \SI{1085}{\volt}, the $G = $.
 
  A map defining the gain calibration of each pad is pre-loaded into the software.
-In the decoder task the raw ADC spectra was multiplied by this gain factor for each channel.
+In the decoder task the raw ADC spectra were multiplied by this gain factor for each channel.
 The PSA thresholds of those channels also were multiplied by the same gain factor since the noise levels were also artificially magnified as well.
@@
 Typically in higher energy experiments fundamental particles are produced with charge $\pm e$.
-Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum ionization in the energy loss.
+Also, the particles are traveling at higher energies in which the energy losses are near or close to the minimum ionization.
 In these cases the dynamic range of most TPC electronics can cover a wide range of particles.
@@
 There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics \cite{feanics,feanics2}.
-Never the less, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware.
+Nevertheless, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware.
 Which can be particularly useful for experiments which have already taken data with older electronics technologies.
@@
 In this case the effective dynamic range is not the single channel dynamic range, but the relative dynamic range between central pad --holding the largest charge-- and the adjacent pads --holding the smaller charges in the PRF distribution.
-For example, to measure minimum ionizing particles, the Signal-to-Noise-Ratio (SNR) of the pad with the smallest charge in the distribution should be some reasonable value above noise, so that the full charge distribution can be measured.
-From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds about 80\% of the total charge, where as the two adjacent pads each hold the remaining 10\%.
+For example, to measure minimum ionizing particles, the Signal-to-Noise-Ratio (SNR) of the pad with the smallest charge in the distribution should be some reasonable value, so that the full charge distribution can be measured.
+From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds about 80\% of the total charge, whereas the two adjacent pads each hold the remaining 10\%.
 In the \spirit TPC the electronics gain was set so that the the pads have a SNR of 6:1 for minimum ionizing tracks, and therefore the central pad had a SNR of 50:1.
@@
 The Gaussian fit was performed on each PRF for a crossing angle steps of  $\ang{5}$ ranging from $\ang{0} < \theta \leq \ang{90}$.
-Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function -- Eq.~\ref{eq:prfgaus} -- which are are plotted versus $\theta$.
+Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function -- Eq.~\ref{eq:prfgaus} -- which are plotted versus $\theta$.
 A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value, which is shown as the black line.
@@
 The minimum $\chi^2$ value returns the best estimate for the unknown saturated charge values.
-The saturated hit charges are updated to reflect these estimates and the cluster position and charge is updated accordingly.
+The saturated hit charges are updated to reflect these estimates and the cluster position and charge are updated accordingly.
 
@@
 
-To provide evidence of the success of this technique, we observe tracks which saturate pads in the high anode wire voltage region are not saturated in the low anode voltage region.
+To provide evidence of the success of this technique, we observe tracks which saturate pads in the high anode wire voltage region but are not saturated in the low anode voltage region.
 While this desaturation technique avoids the need to lower the gain of any region, the low anode voltage region proved to be a direct measurement of the success of this technique.
@@
 Figure~\ref{fig:lowvshigh} shows the the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
-The effect of saturation can be seen in the high gain region for the uncorrected data above values of \SI{400}{ADC \per \milli\metre} where the values plateau, where as the low gain region still returns accurate values.
+The effect of saturation can be seen in the high gain region for the uncorrected data above values of \SI{400}{ADC \per \milli\metre} where the values plateau, whereas the low gain region still returns accurate values.
 Below this value the electronics are not saturated, and therefore the high and low gain sections agree.
@@
 After applying the desaturation technique -- in the right panel-- we see a large improvement, most notably for the He and Li particles, which suffer the most from saturation.
-Even the ${}^{6}Li$ and ${}^{7}Li$ particles can be separated and a more subtle improvement of the lighter particles (p, d, t), can also be seen at lower momenta.
+Even the ${}^{6}$Li and ${}^{7}$Li particles can be separated and a more subtle improvement of the lighter particles (p, d, t), can also be seen at lower momenta.
 In these regions, there was little to no PID resolution before desaturation technique was applied.
@@
 
-As the particles passes through the gas inside of the field cage, they ionize the gas creating electron-ion pairs.
+As the particles pas through the gas inside of the field cage, they ionize the gas creating electron-ion pairs.
 The drift velocities of the ions are typically \SI{e4} times slower than electron drift velocities \cite{blumrol}.
@@
 But as discussed in Section~\ref{sec:wireplanes}, the gating grid captures all of the ions from this region.
-The other source of ions come from the primary ionization produced by the beam and reaction products in the detector gas.
+The other source of ions comes from the primary ionization produced by the beam and reaction products in the detector gas.
 The energy loss  $\langle dE/dx\rangle \propto Z^2$, where Z is the charge of the particle type.
-Because the charge of the un-reacted beam is around Z~50, the ionization due to the beam is a factor of \num{2e3} times that of the light charged particles which mostly are of charge Z~1.
+Because the charge of the un-reacted beam is around Z=50, the ionization due to the beam is a factor of \num{2e3} times that of the light charged particles which mostly are of charge Z=1.
 Therefore the ions resulting from the un-reacted beam is the largest source of positive ions in the TPC.
@@
  The beam is positioned about \SI{25}{\centi\metre} below the anode plane and \SI{29.6}{\centi\metre} above the cathode in the TPC.
-It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane where as it takes the ions \num{5e4}\si{\micro\sec} to drift to the cathode.
+It takes electrons approximately 5\si{\micro\sec} to drift to the anode plane whereas it takes the ions \num{5e4}\si{\micro\sec} to drift to the cathode.
 The beam rate in the experiment was approximately \SI{10}{\kilo\hertz}, which has an average occurrence of 1 beam every \SI{100}{\micro\sec}.
@@
 The linear fits relating the space charge density to the beam rate for each system is shown in Fig.~\ref{fig:spacechg_br_all} where the space charge value is inferred from the measured beam rate.
-Notice that in the $\tin{112}{124}$ system we assume a constant the space charge correction, since the variation in $\Delta\mathrm{V}_\mathrm{LR}$ rate did not span a large enough range to warrant a more detailed analysis.
-
-To reduce the need to compute a new electric field for each space charge value, notice that $\vec{E}\propto \rho$.
+Notice that in the $\tin{112}{124}$ system we assume a constant for the space charge density, since the variation in $\Delta\mathrm{V}_\mathrm{LR}$ rate did not span a large enough range to warrant a more detailed analysis.
+
+To reduce the need to compute a new electric field for each space charge value, notice that $E \propto \rho$.
 We can therefore solve the electric field for a certain reference charge value $\rho_o$, and scale the solution for any other free charge $\rho$ linearly by the ratio $\rho/\rho_o$.
@@
 Adding the BDC vertex greatly improves the momentum resolution of the track fitting.
-Since we know that the space charge affects right and left-going tracks differently, both tracks appear to diverge from the original vertex location appearing to no longer originate from the BDC vertex which is independent of the space charge effects.
+Since we know that the space charge affects right and left-going tracks differently, both tracks diverge from the original vertex location appearing to no longer originate from the BDC vertex which is independent of the space charge effects.
 Without correcting for the space charge effects adding the BDC introduces a systematic shift in the momentum when comparing right and left-going tracks as compared with the momentum value before adding the BDC.
@@
 This is because for small deviations in the track -- at smaller polar angles -- leads to small deviations at the target location.
-Therefore the momentum without does not disagree as much as with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
-Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > 40 \deg$ which are more sensitive to small changes in the track -- corresponding to larger deviations at the target location.
+Therefore the momentum without the BDC as an extra constraint does not disagree as much with the BDC point as shown in Fig.~\ref{fig:mom_S_before}.
+Figure~\ref{fig:mom_L_before} shows the momentum value of tracks going at polar angles of $\theta > \ang{40}$ which are more sensitive to small changes in the track -- corresponding to larger deviations at the target location.
 Therefore we get large systematic shifts in the momentum after including the BDC.
@@
 Recall that only the relative distance between left and right-going tracks was minimized, and there was no guarantee that the corrected tracks coincide with the absolute BDC position at the target.
-The others being the agreement with the expected cocktail calibration beam as described in Section~\ref{sec:cocktail}.
+The others being in agreement with the expected cocktail calibration beam as described in Section~\ref{sec:cocktail}.
 
 Here we will discuss some of the final results of the pion kinetic energy spectrum as it pertains to the verification of the space charge analysis.
-The details of the pion PID analysis will be discussed in detail in Section~\ref{sec:pid}.
+The details of the pion PID analysis will be discussed in Section~\ref{sec:pid}.
 Figure~\ref{fig:sc_momdist} shows the $\pi^+$ and $\pi^-$ kinetic energy distributions in the center of mass system for the $\tin{132}{124}$ and $\tin{108}{112}$ systems.
@@
 For central collisions, one would expect no difference between the two momentum distributions due to the symmetry of the emission.
-As seen in the left panels of both figures, the data before the space charge correction is shown, where as in the right panels the data after the space charge correction is shown.
+As seen in the left panels of both figures, the data before the space charge correction is shown, whereas in the right panels the data after the space charge correction is shown.
 There is a significant improvement in the matching of the left and right side momentum distributions for both charged pion species, in both systems, especially for the high energy tail of the pions.
@@
 The resulting corrected distribution is shown in Fig.~\ref{fig:coboCorr} for the same layer set.
-Figure~\ref{fig:yoff} also show the summary of all the pads before the correction and after the correction.
+Figure~\ref{fig:yoff} also shows the summary of all the pads before the correction and after the correction.
 There is a significant improvement in the width of the distribution going from \SI{1.5}{\milli\metre} to \SI{0.6}{\milli\metre}.
@@
 The first simulation utilizes Geant4 to simulate the interactions of a particle passing through the various materials in the TPC.
-A scale model of the field cage was made, and the correct materials types were put in, with the correct gas mixture and density of P-10 gas at a density of 1 atm.
+A scale model of the field cage was made, and the correct materials types were put in, with the correct gas mixture of P-10 gas at a pressure of 1 atm.
 The magnetic field map of the SAMURAI dipole magnet was also imported into Geant4 as well after assuming rotational symmetry along the axis perpendicular to the pole face -- since only 1/4 of the magnet was simulated.
@@
 where $\sigma_x = 3.4$ and $\sigma_z = 3.5$, and $x_0$ and $z_0$ are the final position of the drifted electron.
-The Gaussian width parameters were determined through an iterative comparison matching the PRFs of the MC and experimental data sets.
+These Gaussian parameters were determined as a best fit of the MC PRF and the experimental data PRF.
 
@@
 Figure~\ref{fig:pulseshape} shows the pulse shape which was extracted from the experimental signals in the data.
-These were signals that did not saturate the electronics and was averaged over a wide range of ADC values.
+These were signals that did not saturate the electronics and were averaged over a wide range of ADC values.
 Here it is normalized so the maximum height is 1.
@@
 
-There are several effects in a complex experimental set up that influence the data in either unknown or un-quantifiable ways.
+There are several effects in a complex experimental setup that influence the data in either unknown or un-quantifiable ways.
 One such effect is the bias of the triggering system, here the Kyoto and Katana multiplicity arrays preferentially select data that are emitted in a particular reaction plane.
@@
 Figure~\ref{fig:dataevent} shows this track embedded into an experimental event.
-In the following, here I will refer to experimental data with embedded MC signals as \emph{embedded data} for short, the MC generated response as \emph{MC data}, and data only containing only the experimental signals as \emph{experimental data}.
-The three sets of data are each independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
+In the following, here I will refer to experimental data with embedded MC signals as \emph{embedded data} for short, the MC generated response as \emph{MC data}, and data only containing the experimental signals as \emph{experimental data}.
+The three sets of data are independently analyzed by the PSA algorithm described in Section~\ref{sec:psa},  which finds all the hits associated with each data set.
 These three independent hit data sets are temporarily stored.
@@
 This is critical for minimum ionizing particles such as pions; this can only be accomplished by first removing almost all of the experimental hits as was done in the first step.
-Each embedded hit that passes this step of criteria is tagged as a originating from a MC hit.
+Each embedded hit that passes this step of criteria is tagged as originating from a MC hit.
 From here, the embedded data is treated as if it were real data passing through the same software analysis.
@@
 For example, several things may happen along the process of embedding that may disqualify a track as a candidate in this na\"ive tagging approach.
-A track could break up, lose or share its charge with an adjacent track, or it may be not be identified at all for a variety of other reasons.
+A track could break up, lose or share its charge with an adjacent track, or it may not be identified at all for a variety of other reasons.
 For the final reconstructed track to be a candidate for the input MC track, it must satisfy two conditions, $N_{sat} > 5$ and $N_{\mathrm{sat}}/N_{\mathrm{Total}} \geq .5$, where $N_{\mathrm{sat}}$ is the number of saturated clusters in a track, and $N_{\mathrm{Total}}$ is the number of total clusters.
@@
 Typically only one of the tracks in the set will satisfy one of the reasonable quality cut conditions in an analysis which will be discussed further in Section~\ref{sec:qualitycut}.
-For example, the most reasonable assumption is the track originating from the vertex is the most correct in this case, where as the split fraction of the track wont originate from the vertex but instead have a large distance to vertex value.
+For example, the most reasonable assumption is the track originating from the vertex is the most correct in this case, whereas the split fraction of the track wont originate from the vertex but instead have a large distance to vertex value.
 In this thesis, the track with the minimum distance to vertex is identified as the track in which the efficiency analysis will be performed.
@@
 This occurs for all tracks, even for minimum ionizing tracks.
-In this case, the saturation is infrequent and typically not an issue as there are many other clusters are not saturated.
+In this case, the saturation is infrequent and typically not an issue as there are many other clusters that are not saturated.
 As the charge of the particle gets higher, or the momentum gets lower, the  energy loss value also becomes higher, and a significant fraction of pads in a track may be saturated.
@@
 There is a large amount of high energy electrons, commonly referred to as \emph{delta rays},  which are produced from the beam passing through the gas.
-These electrons scale with the charge of a particle, in which the Sn beam produces several high energy electrons \cite{pdg}.
+The number of electrons produced scale with the charge of a particle, in which the Sn beam produces several high energy electrons \cite{pdg}.
 In the presence of the magnetic field, almost all of the electrons traveling perpendicular to the field curl up and cannot travel very far, even for very high energy electrons.
@@
 
-Since the \spirit TPC is a fixed target experiment it's angular coverage is certainly not 4$\pi$.
-Because the target is several cm away from the widow of the field cage the geometric acceptance is not even 2$\pi$.
+Since the \spirit TPC is a fixed target experiment its angular coverage is certainly not 4$\pi$.
+Because the target is several cm away from the window of the field cage the geometric acceptance is not even 2$\pi$.
 The rectangular design complicates the calculation of the geometric acceptance of the TPC.
@@
 The bin-by-bin correction method is only valid if this effect is small as compared with the size of the bin.
-If not a more complicated un-folding procedure is needed.
+If not a more complicated unfolding procedure is needed.
 A track is defined as successfully reconstructed even it it has migrated out of the bin, and has successfully passed all cuts.
@@
 It's reasonable to assume the track with the minimum distance to vertex is the track that best represents the MC input track.
-The embedded MC tracks are generated as a uniform distribution in $\theta$,$\phi$, and $p$, to ensure each bin has as similar number of input tracks $N$, giving approximately the same statistical error.
+The embedded MC tracks are generated as a uniform distribution in $\theta$,$\phi$, and $p$, to ensure each bin has a similar number of input tracks $N$, giving approximately the same statistical error.
 Here the efficiency distribution is independent of the initial input distribution since we were careful to define reconstructed tracks, as those which may have migrated out of the bin.
@@
 Figure~\ref{fig:pim_eff_ex} shows the efficiency calculated for $\pi^-$ tracks traveling with momentum \SI{200}{\mega\eVperc} as a function of the two laboratory angles.
-It is easiest to visualize the efficiency distribution first in a polar representation which is most like viewing the actual TPC and progressing to a visualization that is easier to see the values.
+It is easiest to visualize the efficiency distribution first in a polar representation which is most like viewing the actual TPC and progressing to a visualization that is better suited to see the values.
 Figure~\ref{fig:pim_pol_eff_ex} shows the efficiency values plotted in a polar representation where $\theta_{Lab}$ is represented as the radial dimension and $\phi$ is represented as the polar angle.
@@
 Figure~\ref{fig:pim_flat_eff_ex} shows the same efficiency plot but in a Cartesian plot.
-Here we can see even better the acceptance of downward going tracks , $\phi$=\ang{270}, is greater than the dip in upward going tracks, $\phi$=\ang{90}, resulting from the field cage being shorter in the top than in the bottom, this was due to the space the electronics took up on the top of the TPC, making the field cage asymmetric up and down.
+Here we can see even better the acceptance of downward going tracks, $\phi$=\ang{270}, is greater than the dip in upward going tracks, $\phi$=\ang{90}, resulting from the field cage being shorter in the top than in the bottom, this was due to the space the electronics took up on the top of the TPC, making the field cage asymmetric up and down.
 
@@
 
-where the parameters $p_0-p_4$ are free parameters, $\beta= p/\sqrt{p^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum.
+where the parameters $p_0-p_4$ are free parameters, $\beta= pc/\sqrt{{pc}^2 + m^2c^4}$, where $m$ is the mass of the particle and $p$ is the momentum.
 The PID lines vary slightly as a function of the emission angle of each track.
@@
 The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}.
-Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, where as Figure~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.
+Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, whereas Figure~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.
 A Gaussian fit is performed to the pions and to the $e$ spectrum to estimate the background contribution.

### 2020-08-07 chapters_altFormat/ch_data_analysis_2.tex
Put in Pawel and Vladimir edits

@@
 \end{equation}
-where x=A/Q, y=Z, $\mu$ is the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.
+where x=A/Q, y=Z, $\mu$ are the mean values, and $\sigma$ are the Gaussian widths of the two variables. The gates drawn in Figure~\ref{fig:beampidTwo} are summarized in Table~\ref{beamParameters}.
 
@@
 This also occurs for the first and last time bucket in the vertical direction.
-While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough cause issues in the momentum reconstruction.
+While the number of affected clusters is small, compared with the total number in the track, the deviation at the end of the track is enough to cause issues in the momentum reconstruction.
 Simple cuts were taken to graphically remove these clusters around the left,right, top, and bottom of the TPC.
@@
  The target position was physically expected to be \SI{-13.2}{\milli\metre} in the TPC coordinate frame.
-The z-component of all runs in each beam type are plotted around the the target region in Figure~\ref{fig:vertexdist}.
+The z-component of all runs in each beam type are plotted around the target region in Figure~\ref{fig:vertexdist}.
 The mean position of the vertex from the reconstructed data is around \SI{-14.76}{\milli\metre} which is about \SI{1.6}{\milli\metre} off from the expected target position.
@@
 
-The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei.
+The impact parameter is a theoretical quantity defined as the distance between two centers of colliding nuclei if they were to follow a straight line path.
 Though there is no way to directly measure the impact parameter in an experiment, the track multiplicity is indirectly related to the estimated impact parameter of a collision \cite{impactpar}.
@@
 
-where $b_{\mathrm{max}}$ represents the maximum impact parameter detected by the TPC, $b$ is the impact parameter of the vent,  and $dP(N_C/dN_C$ is the normalized multiplicity distribution \cite{reducedimpact}.
+where $b_{\mathrm{max}}$ represents the maximum impact parameter detected by the TPC, $b$ is the impact parameter of the event,  and $dP(N_C/dN_C$ is the normalized multiplicity distribution \cite{reducedimpact}.
 The reduced impact parameter ranges from $\hat{b}=0$ for the most central collisions and to $\hat{b}=1$ for the most peripheral.
-A  detailed analysis was performed, determining the maximum cross section $\sigma_{max}$, for each system, in which $b_{\mathrm{max} = \sqrt{\sigma_{max}/\pi}}$.
+A  detailed analysis was performed, determining the maximum cross section $\sigma_{max}$, for each system, in which $b_{\mathrm{max}} = \sqrt{\sigma_{max}/\pi}$.
 The detailed analysis is given in \cite{jon}.
@@
 The solid angle covered by the $\pi^-$ cuts is \SI{2.077}{\steradian} and for $\pi^+$ cuts \SI{2.042}{\steradian}.
-Assuming the pion emission is isotropic, we multiply each observed pion by a correction factor, $C_a$, to correct for the full 4$\pi$ angular coverage,
+Assuming the pion emission is isotropic, we weight each observed pion by a correction factor, $C_a$, to correct for the full 4$\pi$ angular coverage,
 
@@
 The 4-momentum vector in the lab frame is defined as $\textbf{P} = (E/c,p_x,p_y,p_z)$.
-Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined as the Lorentz transformation matrix $A$,
+Where the corresponding Lorentz transform into the CM frame along the beam (z-axis) is defined by the matrix $A$,
 
@@
 
-where $\beta$, describes the velocity of the CM system, and $\gamma=\sqrt{1-\beta^2}^{-1}$.
+where $\beta$, describes the velocity of the CM system, and $\gamma=1/ \sqrt{1-\beta^2}$.
 The parameter $\beta$ can be determined from the total momentum of the system in the Laboratory frame $P = \sqrt{ T_{P}^2 + 2M_{T}T_{T}}$ and the total energy of the system $E = T_{P} + M_{P} + M_{T}$ -- where $T_{P}$ is the projectile kinetic energy in the Lab frame, $M_{T}$, and $M_{P}$ are the total mass of the target and projectile, where $\beta = -P/E$; the (-) sign denotes the correct direction for transforming from the Lab to the CM frame.
@@
 
-where $P(\pi)$ is the probability that a track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}.
+where $P(\pi)$ is the probability that a given track is a pion described in Eq.~\ref{eq:pionProbs}, $\epsilon_i(p_{Lab}, \theta_{Lab})$ is the efficiency of particle $i$ given by its momentum $p_{Lab}$ and polar angle $\theta_{Lab}$ as discussed in Section~\ref{sec:efficiency}.
 The factor $A$ is a scale factor defined as,
@@
 \begin{equation}
-{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
+{\delta \epsilon }^2 = A^2 \sum_i^{N_{\pi}} \left[ \left( \frac{P(\pi)^{\pi} }{\epsilon_i} \right)^2  \left(  \frac{P(\pi)^{\pi} }{\epsilon_i} + \frac{\delta \epsilon_i}{\epsilon_i} \right)^2   \right]
 \end{equation}

### 2020-08-07 chapters_altFormat/ch_experiment.tex
Put in Pawel and Vladimir edits

@@
 
-Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem.
+Figure~\ref{fig:TPC_schematic} shows the schematic of the effective resistances of the TPC subsystem.
 The cathode is connected to the high voltage (HV) supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$.
@@
 \includegraphics[width=\textwidth]{gg_onoff.pdf}
-\caption{Garfield simulations of the on and off configuration of the gating grid.}
+\caption{Garfield simulations of the on and off configuration of the gating grid with the drift lines of electrons through all of the grids.}
 \label{fig:gg_onoff}
@@
 Theoretically the most negative value represents 100\% electron transparency and was used as the reference run in which we scaled all measurements to.
-The measured electron transparency, $T$, was estimated as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run.
+The measured electron transparency, $T$, was estimated as $T_i = \langle dE/dx\rangle_i /{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run, and $\langle dE/dx\rangle_i$ is the average energy loss for the $i^{th}$ run.
 Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation \cite{garfield++}.
@@
 The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first.
-In later runs, most of the trigger functions were performed the in Katana trigger box.
-The master trigger for the DAQ was different for each primary beam as the experiment got progressively better.
-During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, where as in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
+In later runs, most of the trigger functions were performed in the Katana trigger box.
+The master trigger for the DAQ was different for each primary beam.
+During the ${}^{124}$Xe primary beam, the KATANA trigger box was an input into the trigger logic, whereas in the ${}^{238}$U primary beam, the KATANA trigger box functioned as the trigger logic utilizing the internal trigger electronics.
 In either case the differences in the trigger logic were very minor and they both behaved practically the same except for minor details on the gating-grid trigger \cite{jon}.

### 2020-08-07 chapters_altFormat/ch_intro.tex
Put in Pawel and Vladimir edits

@@
 
-Due to the nature of the strong force, nuclear matter can be thought of as an in-compressible liquid, in much the same way water exhibits incompressibility.
+Due to the nature of the strong force, nuclear matter can be thought of as an incompressible liquid, in much the same way water exhibits incompressibility.
 This analogous picture treating nuclei as liquid droplets is remarkably successful at describing the nuclear binding energy $E_{B}$, which is the energy it would take to disassemble a nucleus into its constituents.
@@
 The term isospin was chosen because this symmetry was similar to the spin symmetry of electrons, which have spin up and down projections but are otherwise indistinguishable, unless in the presence of a symmetry breaking field.
-The isospin mathematics are is identical to the spin mathematics describing where nucleons have an isospin of 1/2 and with projections of  +1/2 for protons and -1/2 for neutrons.
+The isospin mathematics are identical to the spin mathematics describing whether nucleons have an isospin of 1/2 and with projections of  +1/2 for protons and -1/2 for neutrons.
 
@@
 The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}.
-Although the strong force does not distinguish between protons and neutrons the overall wave function is made up of a combination of the spacial, spin, and isospin wave functions.
+Although the strong force does not distinguish between protons and neutrons the overall wave function is made up of a combination of the spatial, spin, and isospin wave functions.
 Since nucleons are fermions,the total wave function must be anti-symmetric under particle exchange.
-Because of this, each spacial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin asymmetry.
+Because of this, each spatial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin.
 Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich.
@@
 The remaining 60 percent primarily occurs because the nuclear force is most attractive between nucleons in the same state, a configuration that is only allowed if one of the nueleons is a proton and the other is a neutron.
-In other words, the di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, where as the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more strongly bound.
+In other words, the di-neutron and di-proton form part of the isospin triplet only allowing for the total isospin $T=1$, whereas the deuteron (neutron-proton) system may form $T={0,1}$ in the singlet or the triplet, with the singlet being more strongly bound.
 
 Large macroscopic objects such as neutron stars are composed of mostly pure neutron matter \cite{neutronstar} with a typical size of approximately $11^{+2}_{-2}$ \si{\kilo\metre}, though they weigh about 1.5 times the mass of our sun.
-Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the balancing outward force arising from the strong force which exerts a large outward pressure arising from the symmetry energy at high densities.
+Such a compact star would normally collapse inward, under its own gravitaitonal force, except for the large outward pressure -- arising from the repulsive nature of the strong nuclear force at high densities.
 The densities reached in a neutron star's interior is a matter of debate.
@@
 \begin{equation}
-P = \Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N}
+P = -\Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N}
 \label{eq:pressEos}
@@
 \begin{equation}
-P = -\rho^2 \frac{dE}{d\rho}\vert_{T=0,N}.
+P = \rho^2 \frac{\delta E}{\delta\rho}\vert_{T=0,N}.
 \label{eq:densEos}
@@
 In contrast, the symmetry energy has only been accurately constrained for densities at or below $\rho_0$ and only preliminary constraints have been obtained at higher densities.
-Figure~\ref{fig:symDen} shows the current status of laboratory constrains on on the density dependence of the symmetry energy.
+Figure~\ref{fig:symDen} shows the current status of laboratory constraints on the density dependence of the symmetry energy.
 The red region labeled 4.) for $\rho < \rho_o$, corresponds to constraints from isobaric analog states \cite{dan2014}.
@@
 Besides observing neutron stars directly, heavy-ion collisions (HIC) provide the only way we can probe the density dependence of the symmetry energy in the laboratory setting.
-When two nuclei collide in a collision, in the very early stages they compress to form a high density region where the nuclei overlap.
+When two nuclei collide, in the very early stages they compress to form a high density region where the nuclei overlap.
 This momentary density can reach up to $3\rho_0$ depending on the incident beam energy, but only a small volume attains such densities.
@@
 The pressure arising from the symmetry energy depends on the curvature of the symmetry energy at a given density as described in Eq.~\ref{eq:densEos}.
-If the density dependence of the symmetry energy is positive at high densities the symmetry energy would in general force neutrons out of the high density system.
+If the density dependence of the symmetry energy grows with a positive derivative at high densities, the symmetry energy would in general force neutrons out of the high density system.
 If the symmetry energy decreases at high densities, making the derivative  negative, then the symmetry energy term would tend to attract neutrons.
 The symmetry energy is referred to as \emph{soft} if the second order term in the Taylor expansion -- at saturation density-- is negative, and \emph{stiff} in the case of a curvature that is close to zero or positive.
-It is the pressure coming from the symmetry energy that is driving the relative dynamics of neutrons and protons, where a stiff symmetry energy expels neutrons from dense matter and vice versa for as soft symmetry energy.
+It is the pressure coming from the symmetry energy that is driving the relative dynamics of neutrons and protons, where a stiff symmetry energy expels neutrons from dense matter and vice versa for the soft symmetry energy.
 
@@
 \label{sec:pionObs}
-Another approach is to find an observable that can be detected with high efficiencies, unlike the neutron, and is more sensitive to the just the high density region.
-For these reasons pions are a very interesting obserabvle, as they are produced though an intermediate step via production of $\Delta$ resonances.
-Here the reaction $ NN \leftrightarrow \Delta$ takes place, where nucleon-nucleon collisions form an excited $\Delta$(1232) baryon resonance from one of the nucleons, which then decays shortly thereafter into a pion $\Delta \leftrightarrow \pi N$.
-The threshold for $\Delta$ resonance production in nucleon-nucleon collisions, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
+Another approach is to find an observable that can be detected with high efficiencies, unlike the neutron, and is more sensitive to just the high density region.
+For these reasons pions are a very interesting observable.
+Pions are produced though an intermediate step via production of $\Delta$ resonances.
+Here the reaction $ NN \leftrightarrow \Delta N$ takes place, where nucleon-nucleon collisions form $\Delta$(1232) baryon resonance from one of the nucleons, which then decays shortly thereafter into a pion $\Delta \leftrightarrow \pi N$.
+The threshold for  an on-shell $\Delta$ resonance production in nucleon-nucleon collisions, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
 In large nuclei, the internal motion of the nucleons is substantial and this modifies the production rates.
@@
 
-Nuclear collisions are typically calculated via semi-classical theories based on a semi-classical extension of the Boltzmann equation or by semi-classical modifications of molecular dynamics \cite{bertsch}.
+Nuclear collisions are typically calculated via semi-classical theories based on a extension of the Boltzmann equation or by modifications of molecular dynamics \cite{bertsch}.
 These theories predict that most of the $\Delta$'s are produced in the early, dense regions of the collision \cite{mingzhang}.
-Figure~\ref{fig:deltaProduction} shows the average local density in panel (c)  in which $\Delta$'s are produced and the total  number of $\Delta$'s produced in the system (b), as a function of time in the simulation, for Au + Au collisions at \SI{400}{\MeVA}.
+Panel (c) of Figure~\ref{fig:deltaProduction} shows the average local density where $\Delta$'s are produced, and the total  number of $\Delta$'s produced in the system (b), as a function of time in the simulation, for Au + Au collisions at \SI{400}{\MeVA}.
 Panel (a) shows the distribution of density at moment of creation for each $\Delta$.
@@
 
-The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordon coefficients as shown in Appendix~\ref{appen:deltadecay}.
+The branching ratio of the various flavors of $\Delta$'s are given by the Clebsh-Gordan coefficients as shown in Appendix~\ref{appen:deltadecay}.
 Here we see that in general proton-proton collisions give rise primarily to  $\pi^+$ and neutron-neutron collisions give rise to primarily $\pi^-$.
@@
 Since this is a three body process, the total pion absorption (removing them from the system) is a smaller effect than the absorption and re-emission process.
-Aside from these two effects which degrade the observable, in general the $\pi^-$ are connected with neutrons and the $\pi^+$ production with proton behavior in the high density, early collision; effectively turning the problematic neutron measurement into a charged $\pi^-$ measurement, which is much easier to measure experimentally.
-It is also directly related to mostly the high density region where the $\Delta$'s are produced.
+Aside from these two effects which degrade the observable, in general the $\pi^-$ are connected with neutrons and the $\pi^+$ production with proton behavior in the high density, early collision; effectively turning the problematic neutron into a measuring charged $\pi^-$ , which is much easier to measure experimentally.
+It is also directly related the high density region where the $\Delta$'s are produced.
 
@@
 Yet in these previous experiments, only the total pion yields were published and no pion spectra were published.
-It was also known that the efficiency analysis of the FOPI detector was particularly difficult especially for low energy pions.
+It was also known that the efficiency analysis of the FOPI collaboration which was particularly difficult especially for low energy pions.
 In fact the published pion multiplicities are not the measured values but rather extrapolated to low pion energies.
-We know now that below the cut off cited in \cite{fopi}, approximately 35\% percent of the $\pi^+$ multiplicity and 55\% percent of the $\pi^-$ multiplicity, showing the importance of measuring down to low pion energies.
+We know now that below the cutoff cited in \cite{fopi}, approximately 35\% of the $\pi^+$ multiplicity and 55\% of the $\pi^-$ multiplicity, showing the importance of measuring down to low pion energies.
 While the pion production increases with the beam energy, this increase is countered by the fact that the effects of the symmetry energy decrease with energy \cite{fopi}.

### 2020-08-07 chapters_altFormat/ch_results.tex
Put in Pawel and Vladimir edits

--- 2020-07-18 chapters/ch_results.tex
@@
 Though these box calculations model an equilibrated system at finite density, which is  not replicated by experiment, the solutions  of  pion and $\Delta$ yields can be analytically calculated from the ansatz of statistical equilibrium, providing a good benchmark.
-This allowed for each model to systematically go through the numerical treatment and details in each model such as initialization of nuclei, stability of the model, numerical handing of pauli-blocking, etc.
+This allowed for each model to systematically go through the numerical treatment and details in each model such as initialization of nuclei, stability of the model, numerical handing of Pauli-blocking, etc.
 After satisfying these benchmark calculations, each model was used with what each author considered the most reliable input and model assumption.
@@
 \emph{in-medium effects}, typically introduced by including a pion optical potential, which describes the pion scattering and absorption.
-Some models include the iso-scalar and the iso-vector delta potential, which are not very well constrained, but can be important in the production of pions \cite{baoan_deltapotential,inmedPionKo,inmedPionFeng}.
+Some models include the isoscalar and the isovector delta potential, which are not very well constrained, but can be important in the production of pions \cite{baoan_deltapotential,inmedPionKo,inmedPionFeng}.
 
@@
 As will be the re-occurring theme, there is a large variation in the predicted pion observables between theoretical models; this is greater than the variation between the stiff and soft symmetry energies within a particular model.
-However, the remedy for this discrepancy is for the authors of these models to include or better model some other physics that many of them have neglected, or it could be that the relevant physics is not constrained by other observables and needs to be constrain by further experiments and comparisons.
+However, the remedy for this discrepancy is for the authors of these models to include or better model some other physics that many of them have neglected, or it could be that the relevant physics is not constrained by other observables and needs to be constrained by further experiments and comparisons.
 
@@
 Other unconstrained physics includes aspects of the mean field potentials for nucleons, pions, and deltas.
-A primary focus is on the isovector potentials, of which only nucleon potentials for have been varied in the calculations shown below.
+A primary focus is on the isovector potentials, of which only nucleon potentials has been varied in the calculations shown below.
 Some of the models include the momentum dependent nucleonic mean field potentials, and the mean field potentials for deltas and pions.
-Calculations have already identified that the overall pion yields are influenced by the mean field potentials for the deltas \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of for delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
+Calculations have already identified that the overall pion yields are influenced by the mean field potentials for the deltas \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
 
@@
 The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations.
-Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
+whereas the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
 \label{fig:ptrap_sn132}
@@
 The measured half of the distribution describes the forward rapidities $y_o > 0$ and has no other extrapolations.
-Where as the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
+whereas the reflected side shows the measured distribution assuming symmetry about $y_o = 0$.}
 \label{fig:ptrap_sn108}
@@
 The table of the transport models are listed in Appendix~\ref{tb:pionyieldTheory}.
-Figure~\ref{fig:totalpiYield} shows the total pion yield for the four systems measured as compared with the models.
+Figure~\ref{fig:totalpiYield} show the total pion yield for the four systems measured as compared with the models.
 The models plotted here are only the soft symmetry energy since the variation in model is much larger than the variation within a model between different symmetry energies.
@@
 The symmetry energy variation (soft and stiff extremes) of two models -- $\chi$BUU and TuQMD -- is plotted as a wide band in the single ratio and in all models in the double ratio.
-The circular markers and band represents the total error bar in the data.
+The circular markers and band represent the total error bar in the data.
 Certainly it can be seen that the variation between symmetry energy extremes in the models, though small, still exists; as initially predicted \cite{baoan_piprod1,baoan_piprod2}, and the small error bars in the data would facilitate a detailed analysis to extract the high density behavior of the symmetry energy, if not for large disagreement between models.
@@
 The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies.
-In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ where as in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
+In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ whereas in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
 Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Figure~\ref{fig:totalpiRatio}.
@@
 \centering
-\includegraphics[width=\textwidth]{pionSpectra_sn108_sum.png}
+\includegraphics[width=.8\textwidth]{pionSpectra_sn108_sum.png}
 \caption{Pion spectra for the $\tin{108}{112}$ system and comparisons to 7 theoretical models. }
@@
 \centering
-\includegraphics[width=\textwidth]{pionSpectra_sn132_sum.png}
+\includegraphics[width=.8\textwidth]{pionSpectra_sn132_sum.png}
 \caption{Pion spectra for the $\tin{132}{124}$ system and comparisons to 7 theoretical models. }
@@
 
-Figures~\ref{fig:pionspectraSn132} and \ref{fig:pionspectraSn108} shows pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems respectively; corrected for efficiency and accounting for the solid angle of 4$\pi$.
+Figures~\ref{fig:pionspectraSn132} and \ref{fig:pionspectraSn108} show pion CM kinetic energy spectra for both the $\tin{132}{124}$ and $\tin{108}{112}$ systems respectively; corrected for efficiency and accounting for the solid angle of 4$\pi$.
 This data marks the first time pion spectral data has been measured at sub-threshold energies.
@@
 \centering
-\includegraphics[width=\textwidth]{singleRatio108_log.png}
+\includegraphics[width=.8\textwidth]{singleRatio108_log.png}
 \caption{Pion spectral ratio $\tin{108}{112}$ system and comparisons to 7 theoretical models.}
@@
 \centering
-\includegraphics[width=\textwidth]{singleRatio132_sum.png}
+\includegraphics[width=.8\textwidth]{singleRatio132_sum.png}
 \caption{Pion spectral ratio $\tin{132}{124}$ system and comparisons to 7 theoretical models.}
@@
 \centering
-\includegraphics[width=\textwidth]{doubleRatio_sum.png}
+\includegraphics[width=.8\textwidth]{doubleRatio_sum.png}
 \caption{Comparing the double spectral ratio, $R_{\tin{132}{124}}/R_{\tin{108}{112}}$, to the 7 theoretical transport simulations.}

### 2020-08-07 chapters_altFormat/ch_summary.tex
Put in Pawel and Vladimir edits

@@
 This information is critical to understanding the symmetry energy and associated isovector mean field potentials.
-These mean field potential contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars.
+These mean field potentials contribute strongly to the chemical potentials, which control the phase transitions and the internal structure within stars.
 Heavy ion collisions provide the only laboratory environment to study these issues at the suprasaturation densities relevant to the neutron star interior.
@@
 We will are exploring how the isoscalar and isovector $\Delta$ potential plays an important role in the pion production spectra and emission rates.
-  Also for the first time, the pion spectra was measured.
-The spectral ratio appears to provide information that is relevant to this question also providing much needed information about the dynamics of pions, even to very low pion energies.
+  Also for the first time, the pion spectra were measured.
+The spectral ratio appears to provide information that is relevant to this question  about the dynamics of pions, even to very low pion energies.
 Of particular importance is the sensitivity of calculations to the high energy tails of the pion spectra.
@@
 Making such a constraint would allow for the possibility to start making constraints on the density dependence of the symmetry energy.
-The interactions with theoretical community that has coalesced around this effort is very exciting, and we are seeing real progress in transport theory.
+The interactions with theoretical community that has coalesced around this effort are very exciting, and we are seeing real progress in transport theory.
 It seems very possible that the theoretical efforts will converge and provide solid interpretations of these and other data much more quickly than they did in the case of the symmetric matter EoS.

### 2020-08-07 macrosForThesis/DigiTasks/PulseShapePlot/plotPulse.C

@@ -78,14 +78,18 @@ void plotPulse()
   leg->SetBorderSize(0);
   leg->SetFillColor(0);
   
+  /*
   TLegend *leg2 = new TLegend(.5,.5,.92,.9);
   leg2->AddEntry(sat_b,"Saturated Signal Pole Zero","L");
   leg2->AddEntry(not_sat_b,"Normal Signal Pole Zero","L");
+  leg2->AddEntry(sat,"Saturated Signal","L");
+  leg2->AddEntry(not_sat,"Normal Signal","L");
   leg2->AddEntry(sat_tb_point,"Saturated Tb","P");
   leg2->AddEntry(sat_mc_tb_point,"Saturated Embedding Threshold","P");
   leg2->SetBorderSize(0);
   leg2->SetFillColor(0);
-
+  */
+  
   sat->SetNpx(10000);
   not_sat->SetNpx(10000);  
 
@@ -141,6 +145,17 @@ void plotPulse()
   style::make(not_sat_clone);
   c2->SaveAs("satpulse.png");
   
+  TLegend *leg2 = new TLegend(.5,.5,.92,.9);
+  leg2->AddEntry(sat_b,"Saturated Signal Pole Zero","L");
+  leg2->AddEntry(not_sat_b,"Normal Signal Pole Zero","L");
+  leg2->AddEntry(sat_clone,"Saturated Signal","L");
+  leg2->AddEntry(not_sat_clone,"Normal Signal","L");
+  leg2->AddEntry(sat_tb_point,"Saturated Tb","P");
+  leg2->AddEntry(sat_mc_tb_point,"Saturated Embedding Threshold","P");
+  leg2->SetBorderSize(0);
+  leg2->SetFillColor(0);
+
+
   c3->cd();
   sat_clone->GetXaxis()->SetRangeUser(0,100);
   sat_clone->GetYaxis()->SetRangeUser(-.05,.2); 

### 2020-08-09 chapters_altFormat/ch_data_analysis.tex
Added Betty's edits

@@
 Here we will briefly describe the outline of the main tasks in the S$\pi$RITROOT software and then will discuss of each task later.
-The level of discussion concerning the software is really a practical, base understanding, appropriate for the discussion of this dissertation.
+The level of discussion concerning the software is really a practical, base understanding, appropriate for the discussion of this Dissertation.
 For more details about the tracking algorithm the reader is refered to \cite{spiritroot_paper}, and to actual software which is published on GitHub \cite{spiritroot_git}.
@@
 Since the STHit class inherits from the STPad class this information is also carried over.
-The y-coordinate is calculated using the known drift velocity $v$ and time of the signal $t_o$ as $y = v\cdot t_0$, $t_0$.
+The y-coordinate is calculated using the known drift velocity $v$ and time of the signal $t_o$ as $y = v\cdot t_0$.
 The hit class forms the fundamental structure from which all tracking is performed since it contains the x,y,z spatial coordinates, and charge information $Q$.
@@
 The hits within a helix track are dense and the position resolution of an individual hit is not very precise.
-As discussed in Section~\ref{sec:prf}, the TPC achieves sub-millimeter position resolution by combining neighboring pads into clusters using the the Pad Response Function (PRF).
+As discussed in Section~\ref{sec:prf}, the TPC achieves sub-millimeter position resolution by combining neighboring pads into clusters using the Pad Response Function (PRF).
 In the helix track finding task, the hits are further reduced into more precise clusters.
@@
 Clusters which do not follow this PRF gate usually have been corrupted in some way.
-Typically this occurs when charges from other nearby tracks mix distorting the hit distribution which corrupts the energy loss and position of the cluster.
+Typically this occurs when charges from other nearby tracks mix, distorting the hit distribution which corrupts the energy loss and position of the cluster.
 This tends to occur most frequently in the dense region surrounding the target, where the track density is high.
@@
 The energy loss of the track is calculated by the truncated mean method described in Eq.~\ref{eq:truncmean}.
-The GENFIT task represents the final part of the analysis which pertaining to track reconstruction.
+The GENFIT task represents the final part of the analysis pertaining to track reconstruction.
 From here the PID can be constructed using the energy loss, momentum, and charge information.
@@
 The RAVE vertex is typically referred to as the TPC vertex in this dissertation.
-It is used primarily to find if an event is on target by observing the z-distribution as will be discussed in Section~\ref{sec:vertexcut}.
+It is used primarily to find if an event is on target as will be discussed in Section~\ref{sec:vertexcut}.
 Off target events stemming from reactions in the active collimator or with the counter gas are rejected this way.
@@
 The gating grid noise profile was measured by recording experimental data, only turning off the anode wires.
-By doing so there are no signals originating from tracks in the TPC and there is only the gating grid noise, since the grid is still being triggered.
+In doing so there are no signals originating from tracks in the TPC and there is only the gating grid noise, since the grid is still being triggered.
 The gating grid noise was stable for a given pad and the average value --shown as the red line-- can be calculated after averaging over several thousands of events and then removed by subtracting the profile from the real data for each pad.
@@
 The pedestal -- or base line ADC-- is around 3800 ADC in the upper panel, is automatically subtracted when subtracting the gating grid noise profile since the subtracted ADC spectra is centered around 0.
-Still some residual gating grid noise remains even after the subtraction, but it is significantly reduced and its extent is over a much smaller time bucket region.
+Still, some residual gating grid noise remains even after the subtraction, but it is significantly reduced and its extent is over a much smaller time bucket region.
 This failure happens in regions where the derivative of the gating grid noise is large.
@@
 A thick Aluminum target was also inserted to provide a slightly lower energy calibration point for part of the lower rigidity setting, effectively creating three energy calibration points over several particle species.
-In some cases the certain particles were not able to propagate or were strongly attenuated and resulted in measurements of poor statistical precision.
-
-Since the cocktail beam was a mono-energetic source of particles and rigidity selection in the BigRIPS corresponding to an actual momentum resolution of better than 1\%, we can interpret the  observed momentum resolution as a measurement of the intrinsic TPC momentum resolution of the TPC for these particles.
+In some cases certain particles were not able to propagate or were strongly attenuated and resulted in measurements of poor statistical precision.
+
+Since the cocktail beam was a mono-energetic source of particles and rigidity selection in the BigRIPS corresponding to an actual momentum resolution of better than 1\%, we can interpret the  observed momentum resolution as a measurement of the intrinsic TPC momentum for these particles.
 The momentum resolution in general depends on several factors such as the particle's angle, momentum, charge, track multiplicity, etc.
@@
 We found that the momenta of light particles in the  cocktail calibration beam differed significantly (up to 5\% ) from the expected values.
-This is shown in Tables~cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
+This is shown in Tables~\cref{tb:cocktail100tar,tb:cocktail100,tb:cocktail300}.
 We could address this discrepancy by incorporating the smaller horizontal magnetic field components at the edge of the dipole and taking into account the electron drift velocities in the direction of $\vec{E}\times\vec{B}$ as seen from Eq.\ref{eq:elecdrift}.
 The $\vec{E}\times\vec{B}$ drift velocity causes the electron trajectories to shift toward the +x-axis in the TPC coordinates causing particles of positive charge (going in the -x-axis) to have a higher measured momenta than in reality.
-The details of the correcting for the $\vec{E}\times\vec{B}$ effect will be discussed later in Section~\ref{sec:spacecharge} in a more general way which also includes the correction for the presence of positive space charge in the field cage volume.
+The details of the correction for the $\vec{E}\times\vec{B}$ effect will be discussed later in Section~\ref{sec:spacecharge} in a more general way which also includes the correction for the presence of positive space charge in the field cage volume.
 The same correction technique was applied here, where the cocktail beam is the special case of zero space charge.
@@
 \label{sec:anodeCalib}
-As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment, high currents being observed on the wires due to a leak in the gating grid that has since been corrected \cite{jon}.
-Out of all the runs used in the analysis in this thesis -- see \ref{tb:runList} -- the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all other runs.
+As discussed in Section~\ref{sec:wireplanes} the voltage of the anode sections 12 and 14 were reduced during the experiment as high currents were observed on the wires due to a leak in the gating grid that has since been corrected \cite{jon}.
+Out of all the runs used in the analysis in this thesis -- see Appendix~\ref{tb:runList} -- the anode sections 12 and 14 were lowered to \SI{1085}{\volt} for runs 2272-2371 and set to \SI{1214}{\volt} for all other runs.
 By lowering the voltage on these anode wires, the gas gain is lowered as compared with all the other anode wire plane sections which operate at \SI{1460}{\volt}.
@@
 Figure~\ref{fig:highlowcal} shows the correlation plot of $\langle dE/dx\rangle$ of the high versus the low gain sections.
-The effects of the high gain channel saturation around a value of 400 ADC\si{\per\milli\metre} and plateaus, whereas the low gain does not; this region is left out of the fit.
+The high gain channel saturates around a value of 400 ADC\si{\per\milli\metre} and plateaus, whereas the low gain does not; this region is left out of the fit.
 A linear fit was performed for the calibration between the high gain channel $h_c = G\cdot l_c$, where $l_c$ is the value of the low gain channel and $G$ is the gain factor.
@@
 Since the energy loss in the Bethe-Bloch equation, Eq.~\ref{eq:bb}, is proportional to $Z^2$, the range of energy losses reflects the possible range of z values.
-Even more, in low energy nuclear HIC of intermediate beam energies -- around \SI{300}{\MeVA} -- produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
-In this case, the limited dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases, leading to very large and even saturated pulses in the electronics.
+HIC of intermediate beam energies -- around \SI{300}{\MeVA} -- produce low velocity particles which exist in the $1/\beta^2$ region of Eq.~\ref{eq:bb}, where the energy loss grows dramatically.
+In this case, the dynamic range of electronics significantly limit the PID as the charge of a particle increases and the velocity decreases, leading to very large and even saturated pulses in the electronics.
 
@@
 The reconstruction of such tracks will suffer.
-There are ongoing efforts in the nuclear community to develop new electronics which hope to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics \cite{feanics,feanics2}.
-Nevertheless, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware.
-Which can be particularly useful for experiments which have already taken data with older electronics technologies.
+There are ongoing efforts in the nuclear community to develop new electronics to mitigate these issues by developing more sophisticated  pre-amplifiers and electronics \cite{feanics,feanics2}.
+Nevertheless, it is useful to develop a software technique which extends the dynamic range of TPC electronics without the use of new external hardware, which can be particularly useful for experiments which have already taken data with older electronics technologies.
 In this section we will outline a novel software technique that takes advantage of the PRF described in Section~\ref{sec:prf} and can effectively extend the dynamic range of the TPC electronics.
 
-In TPCs, the effective dynamic range is very different from the single channel dynamic range depending on how the TPC measurement is used.
+The effective dynamic range is very different from the single channel dynamic range depending on how the TPC measurement is used.
 Typically TPCs are operated inside of a magnetic field for the purpose of reconstructing the momentum of a track, which requires sub-millimeter precision in the position determination along the track path.
@@
 From the PRF in Section~\ref{sec:prf}, we know the central pad in the cluster holds about 80\% of the total charge, whereas the two adjacent pads each hold the remaining 10\%.
-In the \spirit TPC the electronics gain was set so that the the pads have a SNR of 6:1 for minimum ionizing tracks, and therefore the central pad had a SNR of 50:1.
+In the \spirit TPC the electronics gain was set so that the pads have a SNR of 6:1 for minimum ionizing tracks, and therefore the central pad had a SNR of 50:1.
 The maximum SNR in the central pad -- before the electronics saturate -- the SNR is 800:1.
@@
 The Gaussian fit was performed on each PRF for a crossing angle steps of  $\ang{5}$ ranging from $\ang{0} < \theta \leq \ang{90}$.
-Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the the Gaussian function -- Eq.~\ref{eq:prfgaus} -- which are plotted versus $\theta$.
+Figure~\ref{fig:normsigma} shows the two parameters resulting from fitting the Gaussian function -- Eq.~\ref{eq:prfgaus} -- which are plotted versus $\theta$.
 A $4^{th}$ order polynomial fit of these parameters allowed for interpolating for any given $\theta$ value, which is shown as the black line.
@@
 
-Figure~\ref{fig:lowvshigh} shows the the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
+Figure~\ref{fig:lowvshigh} shows the $\langle dE/dx\rangle$ values of the high gain region compared with the calibrated low gain region.
 The effect of saturation can be seen in the high gain region for the uncorrected data above values of \SI{400}{ADC \per \milli\metre} where the values plateau, whereas the low gain region still returns accurate values.
@@
 \includegraphics[width=\linewidth]{spacechg_cartoon.png}
-\caption{Location of space charge in 132 Sn}
+\caption{Cartoon diagram of the location of space charge for the ${}^{132}$Sn beam.}
 \label{fig:spacechg_cartoon}
@@
 Figure~\ref{fig:spacechg_cartoon} gives an idea of the shape of the sheet of space charge carved out by the beam path.
+Figure~\ref{fig:beampaths} show the beam paths of all beams as extracted from the data.
 The ions from each beam create a line charge which drifts towards the cathode with a constant velocity.
@@
     \centering
-    \subfloat[$\tin{132}{123}$ system.]{{\includegraphics[width=\textwidth,valign=c]{sn132_SC} \label{fig:132momdist_sc}}}%
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=\textwidth,valign=c]{sn132_SC} \label{fig:132momdist_sc}}}%
     \qquad

### 2020-08-09 chapters_altFormat/ch_data_analysis_2.tex
Added Betty's edits

@@
 As Chapter~\ref{chap:data1} covered the corrections and various calibrations to the data, this Chapter will cover the various cuts taken to improve the signal in the experimental data.
-Some cuts which identify the incoming beam and the event on target are the minimum cut in order to perform an analysis.
+Some cuts which identify the incoming beam and the event on target are the minimum cut required to perform an analysis.
 The cuts pertaining to the track reconstruction quality are also required to suppress as much of the background noise as possible, most of which comes from reconstruction inefficiencies which the cuts presented here address.
@@
 
-	\caption{Beam PID achieved of the the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
+	\caption{Beam PID achieved of the neutron rich and poor systems, from analysis of the BigRIPs spectrometer \cite{jon} }
 	\label{fig:beampidTwo}
@@
     \centering
-    \subfloat[Beam angle of the $\tin{108}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn108.png} \label{fig:beamangle108}}}%
+    \subfloat[$\tin{108}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn108.png} \label{fig:beamangle108}}}%
     \qquad
-    \subfloat[Beam angle of the $\tin{112}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn112.png} \label{fig:beamangle112}}}%
+    \subfloat[ $\tin{112}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn112.png} \label{fig:beamangle112}}}%
 
-    \subfloat[Beam angle of the $\tin{124}{112}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn124.png} \label{fig:beamangle124}}}%
+    \subfloat[ $\tin{124}{112}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn124.png} \label{fig:beamangle124}}}%
     \qquad
-    \subfloat[Beam angle of the $\tin{132}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn132.png} \label{fig:beamangle132}}}%
+    \subfloat[$\tin{132}{124}$ system]{{\includegraphics[width=.45\textwidth,valign=c]{beamAngle-Sn132.png} \label{fig:beamangle132}}}%
 
@@
 
-Figure~\ref{fig:beamangle} shows the distribution of beam angles for ${}^{132}$Sn and ${}^{108}$Sn systems.
+Figure~\ref{fig:beamangle} shows the distribution of beam angles for all systems.
 Graphical cuts represented by the red lines where events lying in the cuts represent beams that were reconstructed correctly by the beam tracking software outlined in \cite{jon}.
@@
 
-Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that effect the efficiency of track reconstruction.
+Since the TPC does not have spherical symmetry, there are regions of bad geometric acceptance that affect the efficiency of track reconstruction.
 Most notably upward and downward-going tracks will have the shortest track length and will correspond to low regions of efficiency as seen in \ref{sec:efficiency}.
@@
 These angles are actually the same as the spherical $\phi$ angle, where the left and right, are most efficient regions, correspond to  $\ang{0}\substack{+24 \\ -34}$ and $\ang{180}\substack{+34 \\ -24}$.
-While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut off around these angles can be seen in Figure~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks.
+While the exact area of zero efficiency depends on polar angle $\theta_{Lab}$, a clear cut around these angles can be seen in Figure~\ref{fig:angleEffExplanation} in the angular distribution of the measured tracks.
 Here, the experimental data was fitted with a Fermi distribution to find the inflection point of where the efficiency drops off significantly.
@@
     \centering
-    \subfloat[$\pi^-$ angular cuts in the $\tin{108}{112}$ system.
-$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system. ]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
     \qquad
-    \subfloat[$\pi^-$ angular cuts in the $\tin{132}{124}$ system.
-$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
+    \subfloat[$\tin{132}{124}$ system. ]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
 
@@
     \centering
-    \subfloat[$\pi^+$ angular cuts in the $\tin{108}{112}$ system.
-$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
     \qquad
-    \subfloat[$\pi^+$ angular cuts in the $\tin{132}{124}$ system.
-$\theta_{CM} < 90$ and < $\phi$ <]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
      \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
@@
 We also measured two systems $\tin{112}{124}$ and its inverse $\tin{124}{112}$.
-The forward emission in the $\tin{124}{112}$ system is really the same as the backward emission of the $\tin{112}{124}$ system and vice versa.
+The forward emission in the $\tin{124}{112}$ system is the same as the backward emission of the $\tin{112}{124}$ system and vice versa.
 It was shown for pions emitted in central collisions the forward and backward emission was the same \cite{jon}, proving that at least for central collisions pions are emitted isotropically to a good approximation.

### 2020-08-09 chapters_altFormat/ch_experiment.tex
Added Betty's edits

@@
 TPCs have been employed in gases, liquid, and even solid detectors \cite{starTPC,arTPC,eosTPC}.
-In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Thesis, along with the rest of the experimental setup.
+In the following, we outline the physical principles involved in measurements within a gaseous TPC, and discuss the particular details of the TPC used in this Dissertation, along with the rest of the experimental setup.
 
@@
 This correlation allows the particle type to be identified as discussed later in Section~\ref{sec:pid}.
-In this chapter we will provide more details about the particle detection within the \spirit TPC  that is relevant to the reactions studied in this thesis.
+In this chapter we will provide more details about the particle detection within the \spirit TPC  that is relevant to the reactions studied in this Dissertation.
 
@@
 
-The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) was developed to measure pions and other light charge particles produced in collisions of radioactive rare isotope beams with stable targets in fixed target experiments.
-The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
+The SAMURAI Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) was developed to measure pions and other light charged particles produced in collisions of radioactive rare isotope beams with stable targets in fixed target experiments.
+The TPC is built on an Aluminum angle iron skeleton with thin Aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC.
 The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}.
@@
 The field cage contains the detector gas and defines the constant electric field.
-It is rigidly mounted to the large aluminum top plate of the TPC gas containent vessel.
+It is rigidly mounted to the large Aluminum top plate of the TPC enclosure.
 The field cage is electrically isolated from the top plate by a lexan top perimeter ring with o-rings to provide as gas seal.
 The pad plane and wire plane structures are  mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate.
-Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurement \cite{jon}.
+Several Aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurement \cite{jon}.
 Holes in the top plate allowed signals from the individual charge sensitive pads on the pad plane to be transmitted via cables to the electronics.
@@
 \subsection{Enclosure}
-The skeleton of the enclosure is composed of a rigid rectangular aluminum angle-iron frame.
+The skeleton of the enclosure is composed of a rigid rectangular Aluminum angle-iron frame.
 To this frame, six rectangular sides are attached.
-The downstream window and the two side windows are constructed of a aluminum window frames with a 1 mm  aluminum sheet metal window.
+The downstream window and the two side windows are constructed of a Aluminum window frames with a 1 mm  Aluminum sheet metal window.
 These thin sheet metal windows allow charged particles and neutrons to exit the TPC without significant energy loss or scattering.
@@
 The array of circuit boards were glued and supported by Lexan pieces.
-We did not use the common PCB substrate material FR4, which contains a Bromine impregnated epoxy which can outgas Bromine, which can absorb the secondary electrons from the track, thereby degrading the signal \cite{tpcAging}.
+We did not use the common PCB substrate material FR4, which contains a Bromine impregnated epoxy which can outgas Bromin and absorb the secondary electrons from the track, thereby degrading the signal \cite{tpcAging}.
 Instead, a halogen free material, Cryogenic-G10, was used for the board material.
 The downstream wall mainly consists of a large, thin exit window.
-The window  is constructed of   \SI{10}{\micro\metre} Kapton upon which are  evaporated aluminum strips that define the electric field.
-At the bottom of field cage is the cathode that is  constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure.
-On the other end the boards were epoxied into an aluminum top perimeter which also served as the last electrode ring in the TPC.
+The window  is constructed of   \SI{10}{\micro\metre} Kapton upon which are  evaporated Aluminum strips that define the electric field.
+At the bottom of field cage is the cathode that is  constructed from an Aluminum honeycomb laminate, composed of two Aluminum sheets bonded to an Aluminum honeycomb core, providing a lightweight yet rigid structure.
+On the other end the boards were epoxied into an Aluminum top perimeter which also served as the last electrode ring in the TPC.
 Together with the cathode bottom, the field cage proved to be a rigid lightweight structure.
@@
 The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart.
-Since the field cage was build with an array of PCB boards, connections from adjacent strips were connected all the way around until all the strips formed a continuous ring.
-The interfaces were the side pieces met were connected by G-10 corner pieces where conductive paint strips were painted on.
+Since the field cage was built with an array of PCB boards, connections from adjacent strips were connected all the way around until all the strips formed a continuous ring.
+The interfaces where the side pieces met were connected by G10 corner pieces where conductive paint strips were painted on.
 The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip.
@@
 The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Figure~\ref{fig:FC_schematic}.
-Allowing for a tunable resistor allowed for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.
+Allowing for a tunable resistor for fine tuning of the electric field in the region between the top perimeter and the wire planes as will be discussed in Section~\ref{sec:wireplanes}.
 
@@
 Typically the gating grid is always in the closed configuration, only opening it momentarily when the data acquisition trigger criteria is met to take data.
-By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, which would flood the field cage with space charge, distorting the electrons represnting the track paths.
-We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume.
+By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, flooding the field cage with space charge, distorting the electrons represnting the track paths.
+We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift 1 TPC volume.
 After this, we close the gating grid to prevent the back-flow of ions from the avalanche region from that event.
@@
 They are biased to high voltages of about  \SI{e3}{\volt} which creates a very high electric field very close to the anode wire.
-As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gained between collisions exceeds the ionization potential of the gas nocking out more electron-ion pairs in the gas.
-This happens over several collisions creating an avalanche effect amplifying the electron current.
+As the electron approaches the anode wire, the field gets strong enough that the kinetic energy gained between collisions exceeds the ionization potential of the gas knocking out more electron-ion pairs in the gas.
+This happens over succesive  collisions creating an avalanche effect amplifying the electron current.
 The final number of electrons produced depends on the anode wire voltage and the gas properties.
@@
 We will refer to the voltage \SI{1460}{\volt} as the \emph{high voltage} and \SI{1214}{\volt} as the \emph{low voltage}.
-Two sections were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.
+Two sections of the anode wire plane, were biased with the lower voltage setting to minimize the effects of a leakage of secondary electrons around the end of the gating grid \cite{jon}.
 
@@
 Care must be taken to choose the resister ($R_{TP2}$) that connects the field cage divider chain to the top plate.
-The end of the field cage nearest to the top plate is at the same potential as the aluminum top perimeter piece to which the field cage is glued.
+The end of the field cage nearest to the top plate is at the same potential as the Aluminum top perimeter piece to which the field cage is glued.
 This top perimeter piece forms a horizontal surface that is an equipotential representing the last ring.
@@
 This requires the value for $R_{TP2}$ to be chosen correctly.
-Since this simple point can be easily forgotten, we now discuss this choice in greater detail.
 
@@
 \end{equation}
-where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle \cite{blumrol}.
+where $N$ is the number density of electrons in the medium, $e$, the elementary charge, $mc^2$, is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle \cite{blumrol}.
 Yet there is a large variation in energy loss around this mean value.
@@
 The mean energy loss $\langle\Delta\rangle$ is significantly shifted from the most probable value $\Delta_p$, due to the long high energy tail.
-Because of this long tail, for a finite set of energy loss measurements along a given track, the mean value is a very unreliable.
+Because of this long tail, for a finite set of energy loss measurements along a given track, the mean value is very unreliable.
 The most probable energy loss is the better observable which can be obtained either through fitting of the observed distribution or through the truncated mean method.
@@
 
-where $f_r$ is the cut off fraction. In this thesis, a value of 0.7 was used and it was also used in the STAR TPC collaboration \cite{starsyst}.
+where $f_r$ is the cut off fraction. In this Dissertation, a value of 0.7 was used. This was the same value used in the STAR TPC collaboration \cite{starsyst}.
 
@@
 \label{sec:prf}
-Each electron avalanche produces an two-dimensional image charge on the pad plane, as shown in the cartoon in Figure~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
+Each electron avalanche produces a two-dimensional image charge on the pad plane, as shown in the cartoon in Figure~\ref{fig:2DPRF}, where the projection of the charge distribution onto the $x$ and $z$ axis of this distribution are labeled as $\rho(x)$ and $\rho(z)$ respectively.
 If $\rho(x,z)$ represents the charge distribution on the pad-plane, the total charge observed a particular pad, $Q$, is expressed as,
@@
 
-The \spirit TPC was designed to fit exactly into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line.
+The \spirit TPC was designed to fit into the dipole gap of the  dipole magnet at the end of the BigRIPS beam line.
 Figure~\ref{fig:experiment} shows a drawing of the \spirit TPC inside of the SAMURAI magnet chamber which was rotated to the $\ang{0}$ configuration.
@@
 Since the magnet was also measured with the same system after installation, we can match the two systems to get the absolute position of the TPC, and its internal components, relative to the magnet frame.
-The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate, which is much more precise than needed or Show data.
+The position resolution of this type of system was estimated to be around \SI{200}{\micro\metre} for each coordinate.
 
-The origin of the \spirit TPC was defined to be at the center of the TPC pad plane in the x-direction and at the edge of the last upstream pad.
+The origin of the \spirit TPC was defined to be at the center of the TPC pad plane in the x-direction and at the edge of the first upstream pad.
 The center of the magnet is defined as the center of the magnet in x and z-directions and the middle of the dipole gap in the y-direction.
-The origin of the \spirit TPC in the magnet frame is (1.794,205.502,-580.526) for x,y,z-coordinates in units of \si{\milli\metre}.
-The error in each coordinate estimated to be (.2,.1,.4) \si{\milli\metre} respectively.
+The origin of the \spirit TPC in the magnet frame is (1.8,205.5,-580.5) for x,y,z-coordinates in units of \si{\milli\metre}.
+The error in each coordinate estimated to be (0.2,0.1,0.4) \si{\milli\metre} respectively.
 
@@
 \caption{Exploded views of Kyoto and KATANA arrays.The Kyoto array elements are in tan and are located on the left an right side of the TPC.
-When they are in place, they are nested against the thin aluminum side windows of the TPC.
+When they are in place, they are nested against the thin Aluminum side windows of the TPC.
 The KATANA array is in green and is located adjacent to the downstream window of the TPC.}
@@
 
-\subsection{Krakow KATANA Veto and Multiplicity Array}
+\subsection{KATANA Veto and Multiplicity Array}
 %beam veto
@@
  Figure~\ref{fig:aux} also shows the Krakow KATANA array, which consists of 12 plastic scintillating bars mounted adjacent to the downstream wall of the TPC enclosure.
-Three of the 12 bars 1 mm in thickness and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time.
+Three of the 12 bars, which are 1 mm in thickness, and operated as a beam veto in the event the beam did not make a nuclear collision with the target; this occurred approximately 99\% of the time.
 The 9 other bars were 1 cm in thickness and provided additional information about the light particle multiplicity array similar to the Kyoto array.
@@
 The beam was tuned by two sets of quadrupole  magnets, STQ 1 and STQ2 (as seen in Figure~\ref{fig:samuraiBeamLine}), so that the beam spot was focused on the TPC target location.
-Because of the inherent angular dispersion of the beam there still were incoming beam events which significantly deviated from the target location.
+Because of the inherent angular dispersion of the beam, some events significantly deviated from the target location.
 These could be suppressed in the analyses by a gate on the reaction vertex; however, there remained a concern about the dead time such target frame events would contribute and how this would limit the amount of good data we could obtain.
 To veto these type of events an active veto array was set at the entrance of the TPC consisting of four small scintillating bars arranged to be slightly larger than the target size.
-The threshold was set so that any beam particle which passed through any of the bars it would send a trigger signal to not trigger the system since the beam path would not be on target but on some other material inside the TPC.
-Minimizing these active veto rates also proved to be useful when tuning the beam so as to be centered on the target.
+The threshold was set so that any beam particle which passed through any of the bars it would send a signal not to trigger the system since the beam path would not be on target but on some other material inside the TPC.
+Minimizing these active veto rates also proved to be useful when tuning the beam to be more centered on the target.
 
@@
 Such data would be useless, so we closed the gating grid and stopped the data acquisition with a fast clear signal.
-Additional fast clears could also initiated if the Katana trigger box did not generate a trigger.
+Additional fast clears  also was initiated if the Katana trigger box did not generate a trigger.
 Figure~\ref{fig:trigLogic} shows the logic of both of these triggers.
 
-The Katana trigger box took some time to completely master, it did not take on the major trigger responsiblities at first.
+The Katana trigger box initially did not take on the major trigger responsiblities.
 In later runs, most of the trigger functions were performed in the Katana trigger box.
@@
 
-Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal.
+Figure~\ref{fig:dataTrigger1} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output required a Start Counter, KATANA multiplicity, no Veto, and no DAQ busy signal.
 The KATANA trigger, Kyoto Multiplicity, and Start Counter together triggered the DAQ.
 
- Figure~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output was there must be a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
+ Figure~\ref{fig:dataTrigger2} summarizes the ${}^{132}$Xe primary beam, where the condition to produce a true KATANA trigger output required a Start Counter, Kyoto or KATANA multiplicity, no Veto, and no DAQ busy signal.
 Here the KATANA trigger and the SC SUM together triggered the DAQ.
@@
 When opening the gating grid, it is assumed the full volume of the TPC will be read out, therefore  a \SI{11}{\micro\second} gate busy signal is produced; which is slightly more than the time it takes for all the electrons to drift in the field cage.
-In the case were the gating grid should be closed, either due to the fast clear circuit or the end of the TPC measurement, a \SI{5}{\micro\second} busy signal is produced to allow for the gating grid to settle to a steady state closed configuration, and to clear the drift volume of any residual electrons from the beam.
+In the case where the gating grid should be closed, either due to the fast clear circuit or the end of the TPC measurement, a \SI{5}{\micro\second} busy signal is produced to allow for the gating grid to settle to a steady state, closed configuration, and to clear the drift volume of any residual electrons from the beam.
 Both of these gates are included with the DAQ in an OR configuration which makes the overall busy signal.
 
-\section{Collision Data Taken}
+\section{Collision Data}
 
@@
 
-Shown in Figure~\ref{fig:cocktail} is a typical cocktail event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
+Shown in Figure~\ref{fig:cocktail} is a typical calibration event, where one particle enters the TPC volume at a time and parallel to the pad plane, representing an ideal case for momentum and $dE/dx$ determination; as it does not suffer from inefficiencies of high multiplicity events seen in the collision experimental data.
 The primary goal of this data set was to calibrate the TPC in both momentum and dE/dx as will be discussed in Section~\ref{sec:cocktail}.
@@
 \includegraphics[width=\linewidth]{cocktail.png}
-\caption{Pad plane projection for a cocktail event in the TPC.}
+\caption{Pad plane projection for a calibration event in the TPC.}
 \label{fig:cocktail}

### 2020-08-09 chapters_altFormat/ch_intro.tex
Added Betty's edits

@@
 Mankind has been continuously interested in the nature of the matter which composes the visible universe, and what are its constituents or building blocks.
-The seminal discovery was made in 1911 by Ernest Rutherford who scattered alpha particles from a gold foil using particles emitted by an alpha source from Marie Curie.
+The seminal discovery was made in 1911 by Ernest Rutherford who scattered Alpha particles from a gold foil using particles emitted by an Alpha source from Marie Curie.
 His studies led him to conclude that atoms consist of a dense, positively charged atomic nucleus, surrounded by a dilute electron cloud.
 Later,  James Chadwick  discovered that the nucleus contained neutral particles -- neutrons  -- in addition to the known positively charged particle -- protons -- as the basic constituents of nuclear matter.
-Remarkably, the nucleus accounts for 99.9\% of the mass of the atom, and only \num{e-12} of the total volume, making it the densest stable form of matter on Earth.
+Remarkably, the nucleus accounts for 99.9\% of the mass of the atom, and only \num{e-12} of the total volume, making it the most dense stable form of matter on Earth.
 
@@
 The strong force between nucleons is attractive only for a small distance of approximately \SIrange{1}{2}{\femto\metre} and becomes very repulsive at even shorter distances, making the nucleus very difficult to compress.
-It is for this reason that the distance between any pair of nucleons is a near constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the lightest to the heaviest nuclei \cite{krane}.
+It is for this reason that the distance between any pair of nucleons is nearly a constant value, and therefore the central density value is relatively constant over a wide range of nuclei, from the lightest to the heaviest nuclei \cite{krane}.
 This density is referred to as the  \emph{saturation density}, since the central density seems to saturate at a common value, namely, $\rho_0 = \SI{1.7e14}{\gram\per\centi\metre\cubed}$ or \num{0.16} nucleons \si{\per\femto\metre\cubed}.
@@
 
-Since the strong force makes the inter-nucleon distance approximately constant, the volume the nucleus is related to the total number of nucleons and the volume, $A$.
-Therefore the first \emph{volume} term $a_V$ is proportional to $A$.
+Since the strong force makes the inter-nucleon distance approximately constant, the volume term ,$a_V$, is proportional to the volume of the nucleus which is related to the total number of nucleons $A$.
 Since the  constituent quarks that make of the nucleons are of nearly equal mass, there is a fundamental symmetry of the strong interaction which the strong interaction between any pair of nuclei is very similar, regardless of the particular flavor of quark composition.
 This symmetry is called \emph{isospin symmetry}.
-The term isospin was chosen because this symmetry was similar to the spin symmetry of electrons, which have spin up and down projections but are otherwise indistinguishable, unless in the presence of a symmetry breaking field.
+The term isospin was chosen because this symmetry is similar to the spin symmetry of electrons, which have spin up and down projections but are otherwise indistinguishable, unless in the presence of a symmetry breaking field.
 The isospin mathematics are identical to the spin mathematics describing whether nucleons have an isospin of 1/2 and with projections of  +1/2 for protons and -1/2 for neutrons.
@@
 Continuing on with the Bethe-Weizacher formula, we see that there are several correction terms accounting for such effects like the surface, $a_s$, where nucleons near the surface have fewer surrounding neighbors as nucleons inside.
-Since the volume of the nuclei scales as $A$, radius of the nuclei would scale as $A^{1/3}$.
-Also there is a correction accounting for the Coulomb force, $a_c$, which is proportional to the radial extent of the nuclei $R^{-1}$; which scales as $A^{1/3}$.
+Since the volume of the nuclei scales as $A$, radius of the nuclei would scale as $A^{\frac{1}{3}}$.
+Also there is a correction accounting for the Coulomb force, $a_c$, which is proportional to the radial extent of the nuclei $R^{-1}$; which scales as $A^{\frac{1}{3}}$.
 The asymmetry term, $a_A$, is related to the difference in energy between adding a neutron or a proton; it is typically referred to as the \emph{symmetry energy}.
@@
 Because of this, each spatial and spin state of a nucleon can only be occupied by one neutron and one proton, because there still is isospin.
-Because of this it is more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich.
+It is therefore more energetically favorable to form neutron-proton pairs than to become more and more neutron or proton rich.
 Thus the asymmetry term above is proportional to the factor $(N - Z)^2/A$ which comes directly from Pauli blocking.
@@
 \begin{equation}
-P = -\Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N}
+P = -\Big(\frac{\delta E}{\delta V}\Big)\vert_{T=0,N} ,
 \label{eq:pressEos}
@@
 When two nuclei collide, in the very early stages they compress to form a high density region where the nuclei overlap.
-This momentary density can reach up to $3\rho_0$ depending on the incident beam energy, but only a small volume attains such densities.
+This momentary density can reach up to $3\rho_0$ ,depending on the incident beam energy, over a small volume where the nuclei overlap.
 HICs also provide the only way we can probe the isospin asymmetry dependence of the nuclear EoS.
@@
 For these reasons pions are a very interesting observable.
-Pions are produced though an intermediate step via production of $\Delta$ resonances.
-Here the reaction $ NN \leftrightarrow \Delta N$ takes place, where nucleon-nucleon collisions form $\Delta$(1232) baryon resonance from one of the nucleons, which then decays shortly thereafter into a pion $\Delta \leftrightarrow \pi N$.
+Pions are produced through an intermediate step via production of $\Delta$ resonances.
+Here the reaction $ NN \leftrightarrow \Delta N$ takes place, where nucleon-nucleon collisions form $\Delta$(1232) baryon resonance from one of the nucleons, which then decays shortly thereafter into a pion and nucelon, $\Delta \leftrightarrow \pi N$.
 The threshold for  an on-shell $\Delta$ resonance production in nucleon-nucleon collisions, with a mass of \SI{1232}{\mega\electronvolt\per\clight\squared}, corresponds to a laboratory beam of \SI{290}{\MeVA} in a fixed target experiment.
@@
 In an effort to answer the high density behavior of the symmetry energy, we designed a new detector and a set of experiments of Sn + Sn collisions at \SI{270}{\MeVA}.
-Here we utilized inverse kinematics where the beam is made of a radio-active neutron rich beam, impinging on a stable Sn target.
+Here we utilized inverse kinematics where the beam is made of a radioactive neutron rich beam, impinging on a stable Sn target.
 We can probe the isospin asymmetry of the symmetry energy by changing the neutron-proton asymmetry of the incoming beam.
 Pion production has been studied before in stable beams for beam energies of \SI{400}{\MeVA} and above \cite{fopi}.
-Yet in these previous experiments, only the total pion yields were published and no pion spectra were published.
-It was also known that the efficiency analysis of the FOPI collaboration which was particularly difficult especially for low energy pions.
+Yet in these previous experiments, only the total pion yields were published and not the pion spectra.
+It was also known that the efficiency analysis of the FOPI collaboration \cite{fopi} which was particularly difficult especially for low energy pions.
 In fact the published pion multiplicities are not the measured values but rather extrapolated to low pion energies.
-We know now that below the cutoff cited in \cite{fopi}, approximately 35\% of the $\pi^+$ multiplicity and 55\% of the $\pi^-$ multiplicity, showing the importance of measuring down to low pion energies.
+We know now that below the FOPI cutoff cited in \cite{fopi}, approximately 35\% of the $\pi^+$ multiplicity and 55\% of the $\pi^-$ multiplicity, showing the importance of measuring down to low pion energies.
 While the pion production increases with the beam energy, this increase is countered by the fact that the effects of the symmetry energy decrease with energy \cite{fopi}.

### 2020-08-09 chapters_altFormat/ch_results.tex
Added Betty's edits

@@
 Some of the models include the momentum dependent nucleonic mean field potentials, and the mean field potentials for deltas and pions.
-Calculations have already identified that the overall pion yields are influenced by the mean field potentials for the deltas \cite{cozmaPC}, which may not be surprising since a similar sensitivity has govern the occurrence of delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
+Calculations have already identified that the overall pion yields are influenced by the mean field potentials for the deltas \cite{cozmaPC}, which may not be surprising since a similar sensitivity has governed the occurrence of delta and pionic matter within neutron stars \cite{deltaNS,pionNS}.
 
@@
 Though the previous FOPI data set provided interesting data relevant to  the symmetry energy, they relied on extrapolations to low transverse momenta, $P_T = \sqrt{p_x^2 + p_y^2}$, below \SI{100}{\mega\electronvolt\per\clight} \cite{fopi}.
-In the \spirit TPC, we were able to measure pions to very low $P_T$ for rapidities $y > y{CM}$; where the rapidity is expressed along the beam direction (z) as,
+In the \spirit TPC, we were able to measure pions to very low $P_T$ for rapidities $y > y_{CM}$; where the rapidity is expressed along the beam direction (z) as,
 
@@
 \begin{equation}
-y_o = \frac{y_{\pi CM}}{y_{beam} - y_{CM}},
+y_o = \frac{y_{\pi_{CM}}}{y_{beam} - y_{CM}},
 \end{equation}
@@
 This is not completely unexpected; in the delta resonance model mentioned in Section~\ref{sec:pionObs}, we expect  $\pi^-/\pi^+ \sim (N/Z)^2$ \cite{baoan_piprod1,baoan_piprod2}, where N/Z is the ratio of neutron to protons in the dense region where pions are produced.
-For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may a impact that approximation significantly.
+For the $\tin{132}{124}$ system $(N/Z)^2$ = 2.56 and for $\tin{108}{112}$ $(N/Z)^2$ = 1.44, but this n\"aive model assumes no migration of nucleons in or out of the high density region where the isovector mean field and other transport properties may impact that approximation significantly.
 
@@
 The FOPI collaboration has measured the total pion multiplicity resulting from ${}^{197}Au + {}^{197}Au$ collisions at several, higher, beam energies.
-In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ whereas in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.493$.
+In the $\tin{132}{124}$ data the asymmetry is $N/Z=1.56$ whereas in the ${}^{197}Au + {}^{197}Au$ the $N/Z=1.49$.
 Since 4 beams were measured in this experiment, the $N/Z$ dependence was measured as seen in Figure~\ref{fig:totalpiRatio}.
@@
 It is also worth mentioning that the pion ratio observed in the FOPI \SI{400}{\MeVA} setting was already considerably higher than what is expected from the $(N/Z)^2$ delta resonance model \cite{baoan_piprod1,baoan_piprod2}.
-It it worth mentioning that the two experiments differ considerably in their efficiencies as was mentioned before.
+It is worth mentioning that the two experiments differ considerably in their efficiencies as was mentioned before.
 The FOPI results heavily relied on extrapolations to lower pion energies which the \spirit TPC did not rely on \cite{fopi}.
-Without a systematic error calculation or discussion with the FOPI collaboration no further detail can be given except that the \spirit TPC data seems to be reasonable expectation and in general agreement with the FOPI data set.
+Without a systematic error calculation or discussion with the FOPI collaboration no further detail can be given except that the \spirit TPC data seems to be within the expectation and in general agreement with the FOPI data set.
 Scaling the other systems in the \spirit TPC data leads to almost the exact same scaled value of $\mathrm{Y}(\pi^-)/\mathrm{Y}(\pi^+) = \num{4.82}$ and would be redundant.

### 2020-08-09 chapters_altFormat/ch_summary.tex
Added Betty's edits

@@
   This marks the first time pions have been accurately measured in the sub-threshold region.
-We observed significant a deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ \sim (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
+We observed a significant deviation from the  n\"aive expectation of the simple $\Delta$ resonance model, where $\pi^-/\pi^+ \sim (N/Z)^2$, as discussed in Section~\ref{sec:pionObs}.
 Such a deviation is also predicted by theory, so this explanation should be retired in favor of something more quantitative and informative.

### 2020-08-14 MSU-thesis_estee_MaxFormat.tex
Almost finished. Acknowlegements added

@@
 {\it
+To my wife for her unchanging support and patience\\ through this journey,
+and to my flowering bud, Hanna. 
 }
@@
 \begin{acknowledgment}
+This thesis encompassed many stages all the way from detector development and design, to setting up the experiments and the analysis of the data.
+This was a challenging thesis and its success can only be attributed to the collective effort of a relatively small group of very talented individuals.
+Without these people, this thesis would have not been successful, and the data would not have been of such high quality.
+
+First, I would like to thank my advisor Professor Bill Lynch.
+It is through his extensive knowledge and experience that has guided me through this project and for giving me the opportunities to present my research at a wide range of conferences.
+I would like to thank him also for his though provoking dialogue, and for his dedication to making this project a success.
+I would like to also thank Professor Betty Tsang for her leadership, advice, and for keeping this project always on track to meet major milestones.
+I'm very thankful for the many thoughtful conversations and opportunities she has given me to really travel the world and network with many other people.
+
+To my guidance committee, Profs.
+Pawe\l~ Danielewicz, Vladimir Zelevinsky, and Wade Fisher; I chose you all to be a part of my committee because of the impact you had made in my courses.
+Thank you for your time and effort you put into your lectures, and the effort put into advising and reviewing this thesis.
+I will cherish the memories.
+
+I would like to also thank my greater collaboration, starting with the Japanese.
+Thank you Tadaaki Isobe and Mizuki Kurata-Nishimura, Professor Tetsuya Murakami, and Masanori Kaneko for leading the effort in Japan and seeing this experiment through.
+Isobe-san, I wanted to thank you especially for the hard work of setting up the DAQ and electronics of the TPC which really preformed flawlessly, considering how many subsystems were merged and the newness of the GET electronics.
+Mizuki-san, I always enjoyed working with you.
+You are a hard worker and without you the experimental setup, testing, and analysis could not have finished.
+Murakami-san, thank you for your support, advice, and wealth of knowledge.
+I considered you my advisor when I was away from home.
+Kaneko-san, thank you for working so diligently on the Kyoto array and the analysis.
+To Hidetada Baba, thank you for working on the electronics of our experiment, DAQ, and for introducing me to the wonders of Japanese Baseball.
+To all my friends at the University of Tokyo CNS for your friendship, softball, and hospitality.
+I also wanted to thank our Polish colleagues Jerzy \L ukasik, Piotr Paw\l owski, Krzysztof Pelczar, and Pawe\l \L asko for their hard work in developing the KATANA array and for setting up the trigger in the experiment.
+
+I wanted to thank Jon Barney, my good friend; without you this project really would not be possible.
+I could always trust the work you did since you always dedicated yourself to this project.
+Thank you for getting me through the humps in graduate school and for always lending an ear.
+Thank you to Genie Jhang and JungWoo Lee, you wrote beautiful software for this project, and for the hard work done on the analysis, you both are really in a league of your own.
+It was no small task to write the majority of the code by yourselves.
+I've had the pleasure of working with Tommy Tsang who contributed a lot to the analysis of this project and worked double of the rest of us.
+Thank you also to Rensheng Wang.
+I've enjoyed our friendship and you are truly a great researcher and I could always count on you to help me think through a problem I could not solve.
+
+I wanted to thank all of my postdocs whom I have not yet mentioned.
+To Bec Shane and Alan McIntosh, thank you for being the foundation of this project and for the long hours you poured into this.
+To Giordano Cerizza for the work that you did on the software, advice, and for our friendship.
+To Clemantine Santamaria, thank you for your friendship, and for your hard work setting up the experiment and for ensuring that our trigger ran well.
+
+I am forever grateful to my mother and father for their continued support through my education.
+I am very blessed to have them as my parents.
+I wanted to thank my Aunts, Uncles, and Grandparents who have helped me financially and supported me throughout my schooling.
+To my Uncle Brian York, for his advice through the years and stimulating discussions.
+
+Thank you to all the HiRA students through the years.
+I cherished my time with Juan Manfredi a true friend and great researcher.
+Thank you to the NSCL staff and RIKEN staff for making the construction and operation of this detector a success.
+Especially the BigRIPS team, RIBF cyclotron operators, and from the NSCL, John Puro, Jay Pline, Brian Drewyor, Ben Shuart, and John Yurkon for all of their advice, consulting, and ordering of parts.
+
+I would also like to thank the funding sources which have made this all possible: the U.S.
+DOE under Grant Nos.DE-SC0014530, DE-NA0002923, US NSF GrantNo.PHY-1565546, and the Japanese MEXT KAKENHI grant No.24105004.
+The NSF EAPSI fellowship supporting my research at RIKEN for the summer of 2015(NSF award 1515450).
 

### 2020-08-14 chapters_altFormat/ch_data_analysis.tex
Almost finished. Acknowlegements added

@@
     \subfloat[$\mathrm{4^{nd}}$ iteration]{{\includegraphics[width=.24\textwidth,valign=c]{psa3.png} \label{fig:psa3}}}%
-    \subfloat[Final output of the PSA]{{\includegraphics[width=.24\textwidth,valign=c]{psa4.png} \label{fig:psa}}}%
+    \subfloat[Final output of the PSA]{{\includegraphics[width=.24\textwidth,valign=c]{psa4.png} \label{fig:psa4}}}%
 
@@
 \ra{1.3}
-\begin{tabular}{@{}rrrrrrr@{}}\toprule
+\begin{tabular}{@{}ccccccc@{}}\toprule
 & \multicolumn{3}{c}{$100 MeV Target$}\\
@@
 \ra{1.3}
-\begin{tabular}{@{}rrrrrrr@{}}\toprule
+\begin{tabular}{@{}ccccccc@{}}\toprule
 & \multicolumn{3}{c}{$100 MeV$}\\
@@
 \ra{1.3}
-\begin{tabular}{@{}rrrrrrr@{}}\toprule
+\begin{tabular}{@{}ccccccc@{}}\toprule
 & \multicolumn{3}{c}{$300 MeV$}\\
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{gaincalib.png}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{highlowcal.png}
@@
 \begin{figure}[ht!]
+\centering
 \includegraphics[width=\linewidth]{intrographic}
@@
 
-\begin{figure}[ht!]s
-\begin{overpic}[width=\linewidth]{fig5.pdf}
-\put(61,55){\contour{white}{ PRF${}_{\mathrm{Gaus}}(\lambda)$ eq. \ref{eq:gaus}  }}
-\put(61,49){\contour{white}{ PRF${}_{\mathrm{Gatti}}(\lambda)$ eq. \ref{eq:gatti} }}
-\end{overpic}
-\caption{Experimental pad response function of many events for a crossing angle of $85^{\circ} < \theta \leq 90^{\circ}$.  }
-\label{fig:expprf}
-\end{figure}
-
 The shape of the PRF depends on the crossing angle of the track, which determines how wide the charge is distributed along the wire \cite{gatti}.
@@
 \begin{figure}[ht!]
+\centering
 \vspace{5mm}
@@
 \begin{figure}[ht!]
+\centering
 \includegraphics[width=\linewidth]{saturated_pads}
@@
 \centering
-\includegraphics[width=\linewidth]{spacechg_cartoon.png}
+\includegraphics[width=.8\linewidth]{spacechg_cartoon.png}
 \caption{Cartoon diagram of the location of space charge for the ${}^{132}$Sn beam.}
@@
 \centering
-\includegraphics[width=\linewidth]{beampath.png}
+\includegraphics[width=.8\linewidth]{beampath.png}
 \caption{Beam path of the experiments}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{SC_Relation.pdf}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{spaceChgDensity.png}
@@
 \begin{figure}[!htb]
+\centering
 \includegraphics[width=\linewidth]{VLRresidual.png}
@@
     \centering
-    \subfloat[Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, before the space charge correciton.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_small_angle.pdf} \label{fig:psa0}}}%
+    \subfloat[Compare the momentum before and after including the BDC vertex for tracks with $\theta_{Lab} < \ang{40}$, before the space charge correciton.]{{\includegraphics[width=.45\textwidth,valign=c]{BDC_P_small_angle.pdf} \label{fig:mom_S_before}}}%
     \qquad

### 2020-08-14 chapters_altFormat/ch_data_analysis_2.tex
Almost finished. Acknowlegements added

@@
 \includegraphics[scale=.5]{error_b_full.pdf}
-\caption{Estimate of the impact parameter from the track multiplicity.}
+\caption{Estimate of the impact parameter from the track multiplicity from \cite{jon}.}
 \label{fig:impactPar}
@@
     \centering
-    \subfloat[ $\tin{108}{112}$ system. ]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system. ]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
     \qquad
-    \subfloat[$\tin{132}{124}$ system. ]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
+    \subfloat[$\tin{132}{124}$ system. ]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
 
@@
     \centering
-    \subfloat[ $\tin{108}{112}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
     \qquad
-    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.46\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
      \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
@@
 
-\begin{equation}
-A = \begin{pmatrix}
+\[ \left( \begin{array}{cccc}
 1 & 0 & 0 & 0\\
 0 & 1 & 0 & 0\\
-0 & 0 & \gamma & -\beta \gamma\\
+0 & 0 & \gamma & -\beta \gamma \\
 0 & 0 & -\beta \gamma & \gamma
-\end{pmatrix},
-\end{equation}
+\end{array} \right) \]
 

### 2020-08-14 chapters_altFormat/ch_experiment.tex
Almost finished. Acknowlegements added

@@
  \centering
-\ra{1.3}
-\begin{tabular}{@{}rrrrrrr@{}}\toprule 
-\multicolumn{2}{c}{GET electronics settings}\\
+\ra{1.1}
+\begin{tabular}{@{}ccccccc@{}}\toprule 
+\multicolumn{7}{c}{GET electronics settings}\\
 \midrule
- Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane & Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
+ Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane \si{\milli\metre}& Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
  Anode  & Au-plated W   &  20  &  4  &  4   &  0.5  &  1460  \\
@@
 By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, flooding the field cage with space charge, distorting the electrons represnting the track paths.
-We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift 1 TPC volume.
+We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume.
 After this, we close the gating grid to prevent the back-flow of ions from the avalanche region from that event.
@@
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{gain.png}
+\centering
+\includegraphics[width=.8\textwidth]{gain.png}
 \caption{Number of electrons produced in a single avalanche on an anode wire.
@@
 \centering
-\includegraphics[width=\textwidth]{3boards_ps.jpg}
+\includegraphics[width=.8\textwidth]{3boards_ps.jpg}
 \caption{Figure of the pad plane boards being glued to the top plate. }
@@
 \begin{figure}[!htb]
-\includegraphics[width=\linewidth]{data.png}
+\centering
+\includegraphics[width=.8\linewidth]{data.png}
 \caption{Pad plane projection for a collision event in the TPC.
@@
 \begin{figure}[!htb]
-\includegraphics[width=\linewidth]{cocktail.png}
+\centering
+\includegraphics[width=.8\linewidth]{cocktail.png}
 \caption{Pad plane projection for a calibration event in the TPC.}

### 2020-08-14 nicematrix-french.tex
Almost finished. Acknowlegements added

% -*- coding: utf-8 ; -*-
\documentclass[dvipsnames]{article}% dvipsnames is for xcolor (loaded by Tikz, loaded by nicematrix)
\usepackage{xltxtra}
\usepackage[french]{babel}
\frenchsetup{og = « , fg = »}

\usepackage{geometry}
\geometry{left=2.8cm,right=2.8cm,top=2.5cm,bottom=2.5cm,papersize={21cm,29.7cm}}

\usepackage{nicematrix}

\usepackage{tikz}
\usetikzlibrary{fit}

\usepackage{enumitem}
\usepackage{siunitx}
\usepackage{verbatim}

% We use \MakeShortVerb of shortvrb and not \DefineShortVerb of fancyvrb
% because we don't want the contents of short verbatim colored in gray
\usepackage{shortvrb}
\MakeShortVerb{\|}

\usepackage{fancyvrb}
\fvset{commandchars=\~\#\@,formatcom=\color{gray}}

\usepackage{titlesec}
\titlespacing*{\section}{0pt}{6.5ex plus 1ex minus .2ex}{4.3ex plus .2ex}
\titlespacing*{\subsection}{0pt}{4.5ex plus 1ex minus .2ex}{2ex plus .2ex}

\usepackage{multicol}
\setlength{\columnseprule}{0.4pt}

\def\interitem{\vspace{7mm plus 2 mm minus 3mm}}          
\def\emphase{\bgroup\color{RoyalPurple}\let\next=}

\usepackage{footnote}
\usepackage{booktabs}
\usepackage{varwidth}
\usepackage[hyperfootnotes = false]{hyperref}

\hypersetup
  {
    pdfinfo = 
      {
        Title = L’extension nicematrix ,
        Subject = Une extension LaTeX ,
        Author = F. Pantigny 
      }
  }

\NewDocumentEnvironment {scope} {} {} {}

\NewDocumentCommand {\pkg} {m} {\textsf{#1}}
\NewDocumentCommand {\cls} {m} {\textsf{#1}}

\setlength{\parindent}{0pt}

\begin{document}

\VerbatimFootnotes

\title{L'extension \pkg{nicematrix}
       \thanks{Ce document correspond à la version~\myfileversion\space de \pkg{nicematrix},
               en date du~\myfiledate.}} 
\author{F. Pantigny \\ \texttt{fpantigny@wanadoo.fr}}

\maketitle

\begin{abstract}
L'extension LaTeX \pkg{nicematrix} fournit de nouveaux environnements similaires aux environnements classiques
|{tabular}|, |{array}| et |{matrix}| de \pkg{array} et \pkg{amsmath} mais avec des fonctionnalités plus étendues.
\end{abstract}

\vspace{1cm}

\hspace{1cm}
$\begin{bNiceArray}{cccc}[first-row,first-col,
                          code-for-first-col=\color{blue}\scriptstyle,
                          code-for-first-row=\color{blue}\scriptstyle,
                          columns-width = auto]
       & C_1     & C_2    & \Cdots  & C_n    \\
L_1    & a_{11}  & a_{12}  & \Cdots  & a_{1n} \\
L_2    & a_{21}  & a_{22}  & \Cdots  & a_{2n} \\
\Vdots & \Vdots & \Vdots  & \Ddots  & \Vdots \\
L_n    & a_{n1}  & a_{n2}  & \Cdots  & a_{nn} 
\end{bNiceArray}$
\hspace{2cm}
\begin{NiceTabular}[c]{lSSSS}%
[code-before = \rowcolor{red!15}{1-2} \rowcolors{3}{blue!15}{}]
\toprule
\Block{2-1}{Produit} & \multicolumn{3}{c}{dimensions (cm)} & \Block{2-1}{\rotate Prix} \\
\cmidrule(rl){2-4} 
 & L & l & h \\
\midrule
petit    & 3   & 5.5  & 1   & 30    \\
standard & 5.5 & 8    & 1.5 & 50.5  \\
premium  & 8.5 & 10.5 & 2   & 80    \\
extra    & 8.5 & 10   & 1.5 & 85.5  \\
spécial  & 12  & 12   & 0.5 & 70    \\
\bottomrule
\end{NiceTabular}

\vspace{1cm}

L'extension \pkg{nicematrix} est entièrement contenue dans le fichier |nicematrix.sty|. Ce fichier peut être placé
dans le répertoire courant ou dans une arborescence |texmf|. Le mieux reste néanmoins d'installer \pkg{nicematrix}
avec une distribution TeX comme MiKTeX ou TeXlive.

\bigskip
Cette extension peut être utilisée avec |xelatex|, |lualatex| et |pdflatex| mais aussi avec le cheminement
classique |latex|-|dvips|-|ps2pdf| (ou Adobe Distiller).

\bigskip
Cette extension nécessite et charge les extensions \pkg{l3keys2e}, \pkg{xparse}, \pkg{array}, \pkg{amsmath} et
\pkg{pgfcore} ainsi que le module \pkg{shapes} de \textsc{pgf} (l'extension \pkg{tikz}, qui est une surcouche de
\textsc{pgf} n'est \emph{pas} chargée). L'utilisateur final n'a qu'à charger l'extension \pkg{nicematrix} avec
l'instruction habituelle : |\usepackage{nicematrix}|.

\bigskip
L'idée de \pkg{nicematrix} est de créer des nœuds \textsc{pgf} derrière les cases et les positions des filets des
tableaux créés par \pkg{array} et de les utiliser pour développer de nouvelles fonctionnalités. Comme toujours avec
\textsc{pgf}, les coordonnées de ces nœuds sont écrites dans le fichier |.aux| pour être utilisées à la
compilation suivante. C'est pourquoi l'utilisation de \pkg{nicematrix} nécessite \textbf{plusieurs compilations
  successives}.

\bigskip
La plupart des fonctionnalités de \pkg{nicematrix} sont accessibles sans avoir à utiliser explicitement
\textsc{pgf} ou Tikz (ce dernier n'est d'ailleurs pas chargé par défaut).

\bigskip
Une commande|\NiceMatrixOptions| est fournie pour régler les options (la portée des options fixées par cette
commande est le groupe TeX courant : elles sont semi-globales).

\bigskip
{\bfseries \color{red} Important

Depuis la version 5.0 de \pkg{nicematrix}, on doit utiliser les lettres |l|, |c| et
|r| dans les préambules des environnements et non plus les lettres |L|, |C| et
|R|.

Pour assurer la compatibilité avec les versions précédentes, \pkg{nicematrix}
doit être chargée avec l'option |define-L-C-R|.
\begin{Verbatim}
\usepackage[define-L-C-R]{nicematrix}
\end{Verbatim}
}

\newpage

\section{Les environnements de cette extension}

L'extension \pkg{nicematrix} définit les nouveaux environnements suivants :

\medskip
\begin{ttfamily}
\setlength{\tabcolsep}{3mm}
\begin{tabular}{llll}
\{NiceTabular\}  & \{NiceArray\}  & \{NiceMatrix\}   \\
\{NiceTabular*\} & \{pNiceArray\} & \{pNiceMatrix\}  \\
               & \{bNiceArray\} & \{bNiceMatrix\}  \\
               & \{BNiceArray\} & \{BNiceMatrix\}  \\
               & \{vNiceArray\} & \{vNiceMatrix\}  \\
               & \{VNiceArray\} & \{VNiceMatrix\}  
\end{tabular}
\end{ttfamily}

\medskip
Les environnements |{NiceArray}|, |{NiceTabular}| et |{NiceTabular*}| sont
similaires aux environnements |{array}|, |{tabular}| et |{tabular*}| de
l'extension \pkg{array} (qui est chargée par \pkg{nicematrix}).

\medskip
Les environnements |{pNiceArray}|, |{bNiceArray}|, etc. n'ont pas d'équivalents
dans \pkg{array}.

\medskip
Les environnements |{NiceMatrix}|, |{pNiceMatrix}|, etc. sont similaires aux
environnements correspondants de l'\pkg{amsmath} (qui est chargée par
\pkg{nicematrix}) : |{matrix}|, |{pmatrix}|, etc.

\medskip
Tous les environnements de l'extension \pkg{nicematrix} acceptent, entre
crochets, une liste optionnelle de paires de la forme \textsl{clé=valeur}.
{\bfseries Il doit n'y avoir aucun espace devant le crochet ouvrant (|[|) de
  cette liste d'options.}

\bigskip
\textbf{Important}

Avant la version 5.0, il fallait, pour des raisons techniques, utiliser les
lettres |L|, |C| et |R| au lieu de |l|, |c| et |r| dans les préambules des
environnements de \pkg{nicematrix}.

Pour pouvoir continuer à utiliser ces lettres |L|, |C| et |R|, \pkg{nicematrix}
doit être chargé avec l'option |define-L-C-R|.
\begin{Verbatim}
\usepackage[define-L-C-R]{nicematrix}
\end{Verbatim}

\section{L'espace vertical entre les rangées}

Il est bien connu que certaines rangées des tableaux créés par défault avec
LaTeX sont trop proches l'une de l'autre. On en donne ci-dessous un exemple
classique.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
$\begin{pmatrix}
\frac12 & -\frac12 \\
\frac13 & \frac14 \\
\end{pmatrix}$
\end{BVerbatim}
$\begin{pmatrix}
\frac12 & -\frac12 \\
\frac13 & \frac14 \\
\end{pmatrix}$

\bigskip
En s'inspirant de l'extension \pkg{cellspace} qui traite de ce problème,
l'extension \pkg{nicematrix} propose deux clés |cell-space-top-limit| et
|cell-space-bottom-limit| qui sont similaires aux deux paramètres
|\cellspacetoplimit| et |\cellspacebottomlimit| proposés par \pkg{cellspace}. La
valeur initiale de ces paramètres est $0$~pt pour que les environnements de
\pkg{nicematrix} aient par défaut le même comportement que ceux de \pkg{array}
et de l'\pkg{amsmath} mais une valeur de $1$~pt serait un bon choix. On
conseille de régler leurs valeurs avec la commande
|\NiceMatrixOptions|.\footnote{On remarquera que ces paramètres s'appliquent
  aussi aux colonnes de type |S| de \pkg{siunitx} alors que \pkg{cellspace}
  n'est pas utilisable avec ces colonnes.}

\medskip
\begin{Verbatim}
\NiceMatrixOptions{~emphase#cell-space-top-limit = 1pt,cell-space-bottom-limit = 1pt@}
\end{Verbatim}

\begin{BVerbatim}[baseline=c,boxwidth=9cm]
$\begin{pNiceMatrix}
\frac12 & -\frac12 \\
\frac13 & \frac14 \\
\end{pNiceMatrix}$
\end{BVerbatim}
\begin{scope}
\NiceMatrixOptions{
  cell-space-top-limit = 1pt ,
  cell-space-bottom-limit = 1pt ,
}
$\begin{pNiceMatrix}
\frac12 & -\frac12 \\
\frac13 & \frac14 \\
\end{pNiceMatrix}$
\end{scope}

\bigskip
\section{La clé baseline}

L'extension \pkg{nicematrix} propose une option |baseline| pour la position
verticale des tableaux. Cette option |baseline| prend comme valeur un entier qui
indique le numéro de rangée dont la ligne de base servira de ligne de base pour
le tableau.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
$A = \begin{pNiceMatrix}[~emphase#baseline=2@]
\frac{1}{\sqrt{1+p^2}} & p & 1-p \\
1 & 1 & 1 \\
1 & p & 1+p
\end{pNiceMatrix}$
\end{BVerbatim}
$A = \begin{pNiceMatrix}[baseline=2]
\frac{1}{\sqrt{1+p^2}} & p & 1-p \\
1 & 1 & 1 \\
1 & p & 1+p
\end{pNiceMatrix}$

\interitem L'option |baseline| peut aussi prendre les trois valeurs spéciales
|t|, |c| et |b|. Ces trois lettres peuvent aussi être utilisées de manière
absolue comme pour l'option des environnements |{tabular}| et |{array}| de
\pkg{array}. La valeur initiale de |baseline| est~|c|.

\bigskip
Dans l'exemple suivant, on utilise l'option |t| (synonyme de |baseline=t|)
immédiatement après un |\item| de liste. On remarquera que la présence d'un
|\hline| initial n'empêche pas l'alignement sur la ligne de base de la première
rangée (avec |{tabular}| ou |{array}| de \pkg{array}, il faut utiliser
|\firsthline|.

\smallskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\begin{enumerate}
\item un item
\smallskip
\item \renewcommand{\arraystretch}{1.2}
$\begin{NiceArray}[t]{lcccccc}
\hline
n   & 0 & 1 & 2 & 3 & 4  & 5 \\
u_n & 1 & 2 & 4 & 8 & 16 & 32 
\hline
\end{NiceArray}$
\end{enumerate}
\end{BVerbatim}
%
\begin{minipage}{5cm}
\begin{enumerate}
\item un item
\smallskip
\item \renewcommand{\arraystretch}{1.2}
$\begin{NiceArray}[t]{lcccccc}
\hline
n   & 0 & 1 & 2 & 3 & 4  & 5  \\
u_n & 1 & 2 & 4 & 8 & 16 & 32 \\
\hline
\end{NiceArray}$
\end{enumerate}
\end{minipage}

\bigskip
Il est également possible d'utiliser les outils de \pkg{booktabs}: |\toprule|,
|\bottomrule|, |\midrule|, etc.\par\nobreak

\smallskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\begin{enumerate}
\item an item
\smallskip
\item 
$\begin{NiceArray}[t]{lcccccc}
~emphase#\toprule@
n   & 0 & 1 & 2 & 3 & 4  & 5 \\
~emphase#\midrule@
u_n & 1 & 2 & 4 & 8 & 16 & 32 
~emphase#\bottomrule@
\end{NiceArray}$
\end{enumerate}
\end{BVerbatim}
\begin{minipage}{5cm}
\begin{enumerate}
\item an item
\smallskip
\item 
$\begin{NiceArray}[t]{lcccccc}
\toprule
n   & 0 & 1 & 2 & 3 & 4  & 5  \\
\midrule
u_n & 1 & 2 & 4 & 8 & 16 & 32 \\
\bottomrule
\end{NiceArray}$
\end{enumerate}
\end{minipage}

\section{Les blocs}

\label{Block}
Dans les environnements de \pkg{nicematrix}, on peut utiliser la commande
|\Block| pour placer un élément au centre d'un rectangle de cases fusionnées. La
commande |\Block| ne crée pas d'espace par elle-même.

La commande |\Block| doit être utilisée dans la case supérieure gauche du bloc
avec deux arguments. Le premier argument est la taille de ce bloc avec la
syntaxe $i$-$j$ où $i$ est le nombre de rangées et $j$ le nombre de colonnes du
bloc. Le deuxième argument, est, sans surprise, le contenu du bloc.

Dans |{NiceTabular}|, le contenu est composé en mode texte. Dans les autres
environnements, il est composé en mode mathématique.

\medskip
\begin{BVerbatim}
\begin{NiceTabular}{cccc}
rose      & tulipe & marguerite & dahlia \\
violette  & ~emphase#\Block{2-2}{\LARGE\color{blue} fleurs}@  & & souci \\
pervenche & & & lys \\
arum      & iris & jacinthe & muguet 
\end{NiceTabular}
\end{BVerbatim}

\medskip
\begin{center}
\begin{NiceTabular}{cccc}
rose      & tulipe & marguerite & dahlia \\
violette  & \Block{2-2}{\LARGE\color{blue} fleurs}  & & souci \\
pervenche & & & lys \\
arum      & iris & jacinthe & muguet 
\end{NiceTabular}
\end{center}

\interitem On remarquera que le centrage horizontal du contenu des blocs est
correct même si un espacement des colonnes a été demandé par une instruction
comme |!{\qquad}| dans le préambule (ce n'est pas le cas avec |\multicolumn|).
Dans l'exemple suivant, l'en-tête «Premier groupe» est correctement centré.

\medskip
\begin{center}
\fvset{commandchars=\~\#\+}
\begin{BVerbatim}
\begin{NiceTabular}{@{}c!{\qquad}ccc~emphase#!{\qquad}+ccc~@{}}
\toprule
& ~emphase#\Block{1-3}{Premier groupe}+ & & & \Block{1-3}{Deuxième groupe} \\
Rang & 1A & 1B & 1C & 2A & 2B & 2C \\
\midrule
 1 & 0.657 & 0.913 & 0.733 & 0.830 & 0.387 & 0.893\\
 2 & 0.343 & 0.537 & 0.655 & 0.690 & 0.471 & 0.333\\
 3 & 0.783 & 0.885 & 0.015 & 0.306 & 0.643 & 0.263\\
 4 & 0.161 & 0.708 & 0.386 & 0.257 & 0.074 & 0.336\\
\bottomrule
\end{NiceTabular}
\end{BVerbatim}
\end{center}

\bigskip
\begin{center}
\begin{NiceTabular}{@{}c!{\qquad}ccc!{\qquad}ccc@{}}
\toprule
& \Block{1-3}{Premier groupe} & & & \Block{1-3}{Deuxième groupe} \\
Rang & 1A & 1B & 1C & 2A & 2B & 2C \\
\midrule
 1 & 0.657 & 0.913 & 0.733 & 0.830 & 0.387 & 0.893\\
 2 & 0.343 & 0.537 & 0.655 & 0.690 & 0.471 & 0.333\\
 3 & 0.783 & 0.885 & 0.015 & 0.306 & 0.643 & 0.263\\
 4 & 0.161 & 0.708 & 0.386 & 0.257 & 0.074 & 0.336\\
\bottomrule
\end{NiceTabular}
\end{center}

\interitem
On peut aussi utiliser la commande |\Block| dans des matrices mathématiques.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.6cm]
$\begin{bNiceArray}{ccc|c}[margin]
~emphase#\Block{3-3}{A}@ & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
\hline
0 & \Cdots& 0 & 0
\end{bNiceArray}$
\end{BVerbatim}
$\begin{bNiceArray}{ccc|c}[margin]
\Block{3-3}{A} & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
\hline
0 & \Cdots& 0 & 0
\end{bNiceArray}$

\bigskip
On peut souhaiter agrandir la taille du «$A$» placé dans le bloc de l'exemple
précédent. Comme il est composé en mode mathématique, on ne peut pas directement
utiliser une commande comme |\large|, |\Large| ou |\LARGE|. C'est pourquoi une
option à mettre entre chevrons est proposée par |\Block| pour spécifier du code
LaTeX qui sera inséré \emph{avant} le début du mode mathématique.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.6cm]
$\begin{bNiceArray}{ccc|c}[margin]
\Block{3-3}~emphase#<\LARGE>@{A} & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
\hline
0 & \Cdots& 0 & 0
\end{bNiceArray}$
\end{BVerbatim}
$\begin{bNiceArray}{ccc|c}[margin]
\Block{3-3}<\LARGE>{A} & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
\hline
0 & \Cdots& 0 & 0
\end{bNiceArray}$

\section{Les filets horizontaux et verticaux}

Les techniques habituelles pour tracer des filets peuvent être utilisées dans
les environnements de \pkg{nicematrix}, à l'exception de |\vline|. Il y a
néanmoins quelques petites différences de comportement avec les environnements
classiques.

\bigskip
\subsection{Quelques différences avec les environnements classiques}

\subsubsection{Les filets verticaux}

Dans les environnements de \pkg{nicematrix}, les filets verticaux spécifiés par
\verb+|+ dans le préambule des environnements ne sont jamais coupés, même en cas
de ligne incomplète ou de double filet horizontal spécifié par |\hline\hline|
(il n'y a pas besoin d'utiliser \pkg{hhline}).\footnote{Il s'agit là du comportement depuis la version 5.1 de
  \pkg{nicematrix}. Dans les versions précédentes, le comportement était
  le comportement par défaut de \pkg{array}.}

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{NiceTabular}{|c|c|} \hline
Premier & Deuxième \\ ~emphase#\hline\hline@
Paul \\ \hline
Marie & Pauline \\ \hline
\end{NiceTabular}
\end{BVerbatim}
%
\begin{NiceTabular}{|c|c|}[c] \hline
Premier & Deuxième \\ \hline\hline
Paul \\ \hline
Marie & Pauline \\ \hline
\end{NiceTabular}

\bigskip
Si vous utilisez \pkg{booktabs} (qui fournit |\toprule|, |\midrule|,
|\bottomrule|, etc.) et que vous tenez absolument à mettre des filets verticaux
(ce qui est contraire à l'esprit à \pkg{booktabs}), vous constaterez que les
filets tracés par \pkg{nicematrix} sont compatibles avec \pkg{booktabs}.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.5cm]
$\begin{NiceArray}{~emphase#|cccc|@} \toprule
a & b & c & d \\ \midrule
1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\ \bottomrule
\end{NiceArray}$
\end{BVerbatim}
%
$\begin{NiceArray}{|cccc|}
\toprule
a & b & c & d \\
\midrule
1 & 2 & 3 & 4 \\
1 & 2 & 3 & 4 \\
\bottomrule
\end{NiceArray}$

\bigskip
Il reste néanmoins possible de définir un spécificateur, nommé par exemple
\verb|I|, pour tracer des filets verticaux avec le comportement standard de
\pkg{array} :
%
\begin{Verbatim}
\newcolumntype{I}{!{\vrule}}
\end{Verbatim}

En fait, en l'occurrence, il serait peut-être plus habile d'insérer
|\OnlyMainNiceMatrix| (cf. p.~\pageref{OnlyMainNiceMatrix}) :
\begin{Verbatim}
\newcolumntype{I}{!{\OnlyMainNiceMatrix{\vrule}}}
\end{Verbatim}

\subsubsection{La commande  \textbackslash cline}

\label{remark-cline}

Les traits verticaux et horizontaux que l'on insère avec |\hline| et le
spécificateur de colonne «\verb+|+» de \pkg{array} rendent le tableau plus large
ou plus long d'une quantité égale à la largeur du trait (avec \pkg{array} et
aussi avec \pkg{nicematrix}).

\smallskip
Pour des raisons historiques, il n'en est pas de même pour la commande |\cline|,
comme on peut le voir avec l'exemple suivant.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\setlength{\arrayrulewidth}{2pt}
\begin{tabular}{cccc} \hline
A&B&C&D \\ ~emphase#\cline{2-2}@
A&B&C&D \\ \hline
\end{tabular}
\end{BVerbatim}
%
\begin{scope}
\setlength{\arrayrulewidth}{2pt}
\begin{tabular}[c]{cccc}
\hline
A&B&C&D \\
\cline{2-2}
A&B&C&D \\
\hline
\end{tabular}
\end{scope}

\medskip
Dans les environnements de \pkg{nicematrix}, cette situation est corrigée (il
est néanmoins possible de revenir au comportement par défaut de |\cline| avec la
clé |standard-cline|).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\setlength{\arrayrulewidth}{2pt}
\begin{NiceTabular}{cccc} \hline
A&B&C&D \\ ~emphase#\cline{2-2}@
A&B&C&D \\ \hline
\end{NiceTabular}
\end{BVerbatim}
%
\begin{scope}
\setlength{\arrayrulewidth}{2pt}
\begin{NiceTabular}[c]{cccc}
\hline
A&B&C&D \\
\cline{2-2}
A&B&C&D \\
\hline
\end{NiceTabular}
\end{scope}

\subsection{L'épaisseur et la couleur des filets}

Les environnements de \pkg{nicematrix} proposent une clé |rules/width| pour
fixer la largeur (en fait l'épaisseur) des filets dans l'environnement. En fait,
cette clé ne fait que fixer la valeur de |\arrayrulewidth|.

\smallskip
On sait que \pkg{colortbl} propose la commande |\arrayrulecolor| pour spécifier
la couleur de ces filets.

\smallskip
Avec \pkg{nicematrix}, il est possible de spécifier une couleur même si
\pkg{colortbl} n'est pas chargé. Par souci de compatibilité, la commande est
nommée également |\arrayrulecolor|. Les environnements de \pkg{nicematrix}
proposent également une clé |rules/color| qui permet de fixer cette couleur pour
l'environnement en question.

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[baseline=c,boxwidth=9.5cm]
\begin{NiceTabular}{|ccc|}[~emphase#rules/color=[gray]{0.9},rules/width=1pt@]
\hline
rose & tulipe & lys \\
arum & iris & violette \\
muguet & dahlia & souci \\
\hline
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{|ccc|}[rules/color=[gray]{0.9},rules/width=1pt]
\hline
rose & tulipe & lys \\
arum & iris & violette \\
muguet & dahlia & souci \\
\hline
\end{NiceTabular}
\end{scope}

\medskip
Si on veut définir de nouveaux spécificateurs de colonnes pour des filets (par
exemple plus épais ou bien d'une couleur spécifique), on aura peut-être intérêt
à utiliser la commande |\OnlyMainNiceMatrix| décrite
p.~\pageref{OnlyMainNiceMatrix}.

\bigskip
\subsection{Les clés hlines et vlines}

La clé |hlines| demande un tracé de tous les filets horizontaux et la clé
|vlines| demande un tracé de tous les filets verticaux. En fait, dans les
environnements avec délimiteurs (comme |{pNiceMatrix}| ou |{bNiceArray}|), les
filets extérieurs ne sont pas tracés (ce qui est le comportement certainement
attendu).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.6cm]
$\begin{pNiceMatrix}[~emphase#vlines@,rules/width=0.2pt]
1 & 2 & 3 & 4 & 5 & 6 \\
1 & 2 & 3 & 4 & 5 & 6 \\
1 & 2 & 3 & 4 & 5 & 6 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}[vlines,rules/width=0.2pt]
1 & 2 & 3 & 4 & 5 & 6 \\
1 & 2 & 3 & 4 & 5 & 6 \\
1 & 2 & 3 & 4 & 5 & 6 
\end{pNiceMatrix}$

\subsection{La clé hvlines}
\label{hvlines}

La clé |hvlines| demande le tracé de tous les filets horizontaux et verticaux
\emph{sauf dans les blocs}.\footnote{En fait, quand la clé |hvlines| (ou la clé
  |hvlines-except-corners| décrite juste après) est utilisée, les filets ne sont
  pas non plus tracés dans les blocs virtuels délimités par des cases reliées
  par des lignes pointillées (cf. p. \pageref{dotted-and-hvlines}).}

\medskip
\begin{Verbatim}
\begin{NiceTabular}{cccc}[~emphase#hvlines@,rules/color=blue,rules/width=1pt]
rose      & tulipe & marguerite & dahlia \\
violette  & \Block{2-2}{\LARGE\color{blue} fleurs}  & & souci \\
pervenche & & & lys \\
arum      & iris & jacinthe & muguet 
\end{NiceTabular}
\end{Verbatim}
%
\begin{center}
\begin{NiceTabular}{cccc}[hvlines,rules/color=blue,rules/width=1pt]
rose      & tulipe & marguerite & dahlia \\
violette  & \Block{2-2}{\LARGE\color{blue} fleurs}  & & souci \\
pervenche & & & lys \\
arum      & iris & jacinthe & muguet 
\end{NiceTabular}
\end{center}

\subsection{La clé hvlines-except-corners}

La clé |hvlines-except-corners| demande le tracé de tous les filets horizontaux
et verticaux \emph{sauf dans les blocs} et sauf dans les coins vides.

\medskip
\begin{Verbatim}
\begin{NiceTabular}{*{6}{c}}[~emphase#hvlines-except-corners@,cell-space-top-limit=3pt]
  &   &   &   & A \\
  &   & A & A & A \\
  &   &   & A \\
  &   & A & A & A & A \\
A & A & A & A & A & A \\
A & A & A & A & A & A \\
  & \Block{2-2}{B} & & A \\
  &   &   & A \\
  & A & A & A \\
\end{NiceTabular}
\end{Verbatim}
%
\begin{center}
\begin{NiceTabular}{*{6}{c}}[hvlines-except-corners,cell-space-top-limit=3pt]
  &   &   &   & A \\
  &   & A & A & A \\
  &   &   & A \\
  &   & A & A & A & A \\
A & A & A & A & A & A \\
A & A & A & A & A & A \\
  & \Block{2-2}{B} & & A \\
  &   &   & A \\
  & A & A & A \\
\end{NiceTabular}
\end{center}

Comme on le voit, un «coin vide» est constitué de la réunion de tous les
rectangles vides partant de la case située exactement dans le coin.

\bigskip
\begin{BVerbatim}[boxwidth=11cm,baseline=c]
\begin{NiceTabular}{*{6}{c}}%
  [~emphase#hvlines-except-corners@,cell-space-top-limit=3pt]
1\\
1&1\\
1&2&1\\
1&3&3&1\\
1&4&6&4&1\\
1&5&10&10&5&1
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{*{6}{c}}%
  [c,hvlines-except-corners,cell-space-top-limit=3pt]
1\\
1&1\\
1&2&1\\
1&3&3&1\\
1&4&6&4&1\\
1&5&10&10&5&1
\end{NiceTabular}

\subsection{La commande \textbackslash diagbox}

La commande |\diagbox| (inspirée par l'extension \pkg{diagbox}) permet, quand
elle est utilisée dans une case, de couper cette case selon une diagonale
descendante.\footnote{L'auteur de ce document considère que ce type de
  construction est un piètre choix graphique.}

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{NiceArray}{*{5}{c}}[hvlines]
~emphase#\diagbox{x}{y}@  & e & a & b & c \\
e & e & a & b & c \\
a & a & e & c & b \\
b & b & c & e & a \\
c & c & b & a & e 
\end{NiceArray}$
\end{BVerbatim}
$\begin{NiceArray}{*{5}{c}}[hvlines]
\diagbox{x}{y}  & e & a & b & c \\
e & e & a & b & c \\
a & a & e & c & b \\
b & b & c & e & a \\
c & c & b & a & e 
\end{NiceArray}$

\medskip
Cette commande |\diagbox| peut aussi être utilisée dans un |\Block|.

\subsection{Filets en pointillés}

Dans les environnements de \pkg{nicematrix}, il est possible d'utiliser la
commande |\hdottedline| (fournie par \pkg{nicematrix}) qui est l'équivalent pour
les pointillés des commandes |\hline| et |\hdashline| (cette dernière étant une
commande de \pkg{arydshln}).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9.5cm]
\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
~emphase#\hdottedline@
6 & 7 & 8 & 9 & 10 \\
11 & 12 & 13 & 14 & 15 
\end{pNiceMatrix}
\end{BVerbatim}
$\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
\hdottedline
6 & 7 & 8 & 9 & 10 \\
11 & 12 & 13 & 14 & 15 
\end{pNiceMatrix}$

\bigskip
Dans les environnements avec un préambule explicite (comme |{NiceTabular}|,
|{NiceArray}|, etc.), il est possible de dessiner un trait vertical en
pointillés avec le spécificateur «|:|».

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9.5cm]
\begin{pNiceArray}{cccc~emphase#:@c}
1 & 2 & 3 & 4 & 5 \\
6 & 7 & 8 & 9 & 10 \\
11 & 12 & 13 & 14 & 15 
\end{pNiceArray}
\end{BVerbatim}
$\begin{pNiceArray}{cccc:c}
1 & 2 & 3 & 4 & 5 \\
6 & 7 & 8 & 9 & 10 \\
11 & 12 & 13 & 14 & 15 
\end{pNiceArray}$

\bigskip
Il est possible de changer dans \pkg{nicematrix} la lettre utilisée pour
indiquer dans le préambule un trait vertical en pointillés avec l'option
|letter-for-dotted-lines| disponible dans |\NiceMatrixOptions|.

\bigskip
\emph{Remarque} : Quand l'extension \pkg{array} (sur laquelle s'appuie
\pkg{nicematrix}) est chargée, les traits verticaux et horizontaux que l'on
insère rendent le tableau plus large ou plus long d'une quantité égale à la
largeur du trait\footnote{En fait, cela est vrai pour |\hline| et «\verb+|+»
  mais pas pour |\cline| : cf. p.~\pageref{remark-cline}}. Avec
\pkg{nicematrix}, les lignes en pointillés tracées par |\hdottedline| et «|:|»
ont le même effet.

\section{Les couleurs des rangées et des colonnes}

\subsection{Utilisation de colortbl}

Rappelons que l'extension \pkg{colortbl} peut être chargée directement par
|\usepackage{colortbl}| ou en chargeant l'extension \pkg{xcolor} avec l'option
|table| : |\usepackage[table]{xcolor}|.

\medskip
Comme l'extension \pkg{nicematrix} est fondée sur \pkg{array}, il est possible
d'utiliser \pkg{colortbl} avec \pkg{nicematrix}.

\medskip
Il y a néanmoins deux inconvénients :
\begin{itemize}
\item L'extension \pkg{colortbl} patche \pkg{array}, ce qui entraîne des
incompatibilités (par exemple avec la commande |\hdotsfor|).

\item L'extension \pkg{colortbl} construit le tableau ligne par ligne, en
alternant rectangles colorés, filets et contenu des cases. Le \textsc{pdf}
résultant déroute certains lecteurs de \textsc{pdf} et on a des artefacts
d'affichage : des filets verticaux semblent disparaître ; au contraire, de fines
lignes blanches peuvent apparaître. L'extension \pkg{nicematrix} proposent des
outils qui évitent ces inconvénients.
\end{itemize}

\subsection{Les outils de nicematrix dans le code-before}

L'extension \pkg{nicematrix} proposent des outils (indépendants de
\pkg{colortbl}) pour tracer d'abord les rectangles colorés, puis le contenu des
cases et les filets. Cette manière de faire est plus dans l'esprit du «modèle du
peintre» des formats PostScript et \textsc{pdf} et convient donc mieux aux
lecteurs de \textsc{pdf}. L'inconvénient est qu'elle nécessite plusieurs
compilations successives.

\medskip
L'extension \pkg{nicematrix} fournit une clé |code-before| pour du code qui sera
exécuté avant le tracé du tableau. De nouvelles commandes y sont disponibles :
|\cellcolor|, |\rectanglecolor|, |\rowcolor|, |\columncolor|, |\rowcolors| et
|\chessboardcolors|.
\label{code-before}

\medskip
Toutes ces commandes acceptent un argument optionnel (entre crochets et en
première position) qui est le modèle colorimétrique pour la spécification des
couleurs.

\medskip
\begin{itemize}
\item La commande |\cellcolor| tient son nom de la commande |\cellcolor| de
\pkg{colortbl}.

Elle prend en arguments obligatoires une couleur et une liste de cases sous le
format $i$-$j$ où $i$ est le numéro de ligne et $j$ le numéro de colonne.

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[boxwidth=10cm,baseline=c]
\begin{NiceTabular}{|c|c|c|}[code-before = ~emphase#\cellcolor{red!15}{3-1,2-2,1-3}@]
\hline
a & b & c \\ \hline
e & f & g \\ \hline
h & i & j \\ \hline
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{|c|c|c|}[code-before = \cellcolor{red!15}{3-1,2-2,1-3}]
\hline
a & b & c \\ \hline
e & f & g \\ \hline
h & i & j \\ \hline
\end{NiceTabular}
\end{scope}

\bigskip
\item La commande |\rectanglecolor| prend trois arguments obligatoires. Le
premier est la couleur, les deux suivants fournissent la case en haut à gauche
et la case en bas à droite du rectangle.

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[boxwidth=10cm,baseline=c]
\begin{NiceTabular}{|c|c|c|}[code-before = ~emphase#\rectanglecolor{blue!15}{2-2}{3-3}@]
\hline
a & b & c \\ \hline
e & f & g \\ \hline
h & i & j \\ \hline
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{|c|c|c|}[code-before = \rectanglecolor{blue!15}{2-2}{3-3}]
\hline
a & b & c \\ \hline
e & f & g \\ \hline
h & i & j \\ \hline
\end{NiceTabular}
\end{scope}

\bigskip
\item La commande |\rowcolor| doit son nom à la commande |\rowcolor| de
\pkg{colortbl}. Son premier argument obligatoire est la couleur et le deuxième
est une liste de numéros de rangées ou bien d'intervalles de rangées sous la
forme $a$-$b$ (un intervalle de la forme $a$- représente toutes les rangées à
partir de la rangée~$a$).

\medskip

\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[boxwidth=9cm,baseline=c]
$\begin{NiceArray}{lll}[hvlines, ~emphase#code-before = \rowcolor{red!15}{1,3-5,8-}@]
a_1 & b_1 & c_1 \\
a_2 & b_2 & c_2 \\
a_3 & b_3 & c_3 \\
a_4 & b_4 & c_4 \\
a_5 & b_5 & c_5 \\
a_6 & b_6 & c_6 \\
a_7 & b_7 & c_7 \\
a_8 & b_8 & c_8 \\
a_9 & b_9 & c_9 \\
a_{10} & b_{10} & c_{10} \\
\end{NiceArray}$
\end{BVerbatim}
%
$\begin{NiceArray}{lll}%
[baseline=4,hvlines, code-before = \rowcolor{red!15}{1,3-5,8-}]
a_1 & b_1 & c_1 \\
a_2 & b_2 & c_2 \\
a_3 & b_3 & c_3 \\
a_4 & b_4 & c_4 \\
a_5 & b_5 & c_5 \\
a_6 & b_6 & c_6 \\
a_7 & b_7 & c_7 \\
a_8 & b_8 & c_8 \\
a_9 & b_9 & c_9 \\
a_{10} & b_{10} & c_{10} \\
\end{NiceArray}$
\end{scope}

\bigskip
\item La commande |\columncolor| doit son nom à la commande |\columncolor| de
\pkg{colortbl}. Sa syntaxe est similaire à celle de |\rowcolor|.

\bigskip
\item La commande |\rowcolors| (avec un \emph{s}) doit son nom à la commande
|\rowcolors| de \pkg{xcolor}\footnote{La commande |\rowcolors| de \pkg{xcolor}
  est disponible quand \pkg{xcolor} est chargé avec l'option |table|.}. Le
\emph{s} rappelle qu'il y a deux couleurs. Elle colorie alternativement les
rangées avec les deux couleurs à partir de la rangée dont le numéro est donné en
premier argument (obligatoire).

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\begin{NiceTabular}{lr}[hlines,code-before = ~emphase#\rowcolors{1}{blue!10}{}@]
Pierre & 12 \\
Jacques & 8 \\
Stéphanie & 18 \\
Amélie & 20 \\
Henri & 14 \\
Estelle & 15 
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{lr}[hlines,code-before = \rowcolors{1}{blue!10}{},baseline=3]
Pierre & 12 \\
Jacques & 8 \\
Stéphanie & 18 \\
Amélie & 20 \\
Henri & 14 \\
Estelle & 15 
\end{NiceTabular}
\end{scope}

\bigskip
\item La commande |\chessboardcolors| prend en arguments obligatoires deux
couleurs et colorie les cases en quinconces avec les deux couleurs.

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
$\begin{pNiceMatrix}[r,margin, ~emphase#code-before=\chessboardcolors{red!15}{blue!15}@]
1 & -1 & 1 \\
-1 & 1 & -1 \\
1 & -1 & 1 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}%
[baseline=1, r, margin, code-before = \chessboardcolors{red!15}{blue!15}]
1 & -1 & 1 \\
-1 & 1 & -1 \\
1 & -1 & 1 
\end{pNiceMatrix}$
\end{scope}

\medskip
On a utilisé la clé |r| qui impose que toutes les colonnes soient alignées à
droite (cf. p.~\pageref{key-R}).
\end{itemize}

\bigskip
On remarquera que ces commandes sont compatibles avec les commandes de
\pkg{booktabs} (|\toprule|, |\midrule|, |\bottomrule|, etc).

\medskip
\begin{scope}
\hfuzz=10cm
\begin{BVerbatim}[baseline=c,boxwidth=8.5cm]
\begin{NiceTabular}{lSSSS}%
[code-before = \rowcolor{red!15}{1-2} \rowcolors{3}{blue!15}{}]
~emphase#\toprule@
\Block{2-1}{Produit} \\
\Block{1-3}{dimensions (cm)} & & & 
\Block{2-1}{\rotate Prix} \\
~emphase#\cmidrule(rl){2-4}@
 & L & l & h \\
~emphase#\midrule@
petit    & 3   & 5.5  & 1   & 30    \\
moyen    & 5.5 & 8    & 1.5 & 50.5  \\
premium  & 8.5 & 10.5 & 2   & 80    \\
extra    & 8.5 & 10   & 1.5 & 85.5  \\
spécial  & 12  & 12   & 0.5 & 70    \\
~emphase#\bottomrule@
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}[c]{lSSSS}%
[code-before = \rowcolor{red!15}{1-2} \rowcolors{3}{blue!15}{}]
\toprule
\Block{2-1}{Produit} & 
\Block{1-3}{dimensions (cm)} & & & 
\Block{2-1}{\rotate Prix} \\
\cmidrule(rl){2-4} 
 & L & l & h \\
\midrule
petit    & 3   & 5.5  & 1   & 30    \\
moyen    & 5.5 & 8    & 1.5 & 50.5  \\
premium  & 8.5 & 10.5 & 2   & 80    \\
extra    & 8.5 & 10   & 1.5 & 85.5  \\
spécial  & 12  & 12   & 0.5 & 70    \\
\bottomrule
\end{NiceTabular}
\end{scope}

\medskip
On a utilisé le type de colonne |S| de \pkg{siunitx}.

\subsection{Outils de coloriage avec la syntaxe de colortbl}

On peut accéder aux outils de coloriage précédents avec une syntaxe proche de
celle proposée par \pkg{colortbl}. On doit pour cela utiliser la clé
|colortbl-like| dans l'environnement concerné.\footnote{Pour le moment, cette
  clé n'est pas disponible disponible dans |\NiceMatrixOptions|.}

On a alors accès aux trois commandes suivantes :
\begin{itemize}
\item |\cellcolor| qui colorie la case courante ;
\item |\rowcolor| à utiliser dans une case et qui colorie le reste de la rangée
;
\item |\columncolor| à utiliser dans le préambule de la même manière que la
commande éponyme de \pkg{colortbl}\footnote{Contrairement à la commande
|\columncolor| de \pkg{colortbl}, celle-ci peut apparaître à l'intérieur d'une
autre commande, elle-même utilisée dans le préambule.}.
\end{itemize}

\medskip
\begin{Verbatim}
\begin{NiceTabular}[colortbl-like]{>{~emphase#\columncolor{blue!15}}@ccc}
\toprule
~emphase#\rowcolor{red!15}@
Nom & Prénom & Année de naissance \\
\midrule
Achard & Jacques & 5 juin 1962 \\
Lefebvre & Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{Verbatim}

\begin{center}
\begin{NiceTabular}[colortbl-like]{>{\columncolor{blue!15}}ccc}
\toprule
\rowcolor{red!15}
Nom & Prénom & Année de naissance \\
\midrule
Achard & Jacques & 5 juin 1962 \\
Lefebvre & Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{center}

\section{La largeur des colonnes}
\label{width}

Dans les environnements avec un préambule explicite (comme |{NiceTabular}|,
|{NiceArray}|, etc.), il est possible de fixer la largeur d'une colonne avec les
lettres classiques |w| et |W| de l'extension \pkg{array}.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\begin{NiceTabular}{~emphase#W{c}{2cm}@cc}[hvlines]
Paris  & New York & Madrid \\
Berlin & London   & Roma    \\
Rio    & Tokyo    & Oslo 
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{W{c}{2cm}cc}[hvlines]
Paris  & New York & Madrid \\
Berlin & London   & Roma    \\
Rio    & Tokyo    & Oslo 
\end{NiceTabular}

\bigskip
Dans les environnements de \pkg{nicematrix}, il est aussi possible de fixer la
largeur \emph{minimale} de toutes les colonnes de la matrice directement avec
l'option |columns-width|.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{pNiceMatrix}[~emphase#columns-width = 1cm@]
1  & 12 & -123 \\
12 & 0  & 0    \\
4  & 1  & 2 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}[columns-width = 1cm]
1  & 12 & -123 \\
12 & 0  & 0    \\
4  & 1  & 2 
\end{pNiceMatrix}$

\medskip
Notez que l'espace inséré entre deux colonnes (égal à 2 |\tabcolsep| dans
|{NiceTabular}| et à 2 |\arraycolsep| dans les autres environnements) n'est pas
supprimé (il est évidemment possible de le supprimer en mettant |\tabcolsep| ou
|\arraycolsep| à~$0$ avant).

\bigskip
Il est possible de donner la valeur spéciale |auto| à l'option |columns-width|:
toutes les colonnes du tableau auront alors une largeur égale à la largeur de la
case la plus large du tableau.\footnote{Le résultat est atteint dès la première
  compilation (mais PGF/Tikz écrivant des informations dans le fichier |.aux|,
  un message demandant une deuxième compilation apparaîtra).}\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{pNiceMatrix}[~emphase#columns-width = auto@]
1  & 12 & -123 \\
12 & 0  & 0    \\
4  & 1  & 2 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}[columns-width = auto]
1  & 12 & -123 \\
12 & 0  & 0    \\
4  & 1  & 2 
\end{pNiceMatrix}$

\bigskip
Sans surprise, il est possible de fixer la largeur minimale de toutes les
colonnes de toutes les matrices dans une certaine portion de document avec la
commande |\NiceMatrixOptions|.\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=8.5cm]
~emphase#\NiceMatrixOptions{columns-width=10mm}@
$\begin{pNiceMatrix}
a & b \\ c & d 
\end{pNiceMatrix}
= 
\begin{pNiceMatrix}
1   & 1245 \\ 345 & 2 
\end{pNiceMatrix}$
\end{BVerbatim}
\begin{scope}
\NiceMatrixOptions{columns-width=10mm}
$\begin{pNiceMatrix}
a & b \\
c & d 
\end{pNiceMatrix}
= 
\begin{pNiceMatrix}
1   & 1245 \\
345 & 2 
\end{pNiceMatrix}$
\end{scope}

\bigskip
Mais il est aussi possible de fixer une zone dans laquelle toutes les matrices
auront leurs colonnes de la même largeur, égale à la largeur de la case la plus
large de toutes les matrices de la zone. Cette construction utilise
l'environnement |{NiceMatrixBlock}| avec l'option
|auto-columns-width|\footnote{Pour le moment, c'est le seul usage de
  l'environnement |{NiceMatrixBlock}| mais il pourrait y en avoir davantage dans
  le futur.}. L'environnement |{NiceMatrixBlock}| n'a pas de rapport direct avec
la commande |\Block| présentée précédemment dans ce document
(cf.~p.~\pageref{Block}).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=8.5cm]
~emphase#\begin{NiceMatrixBlock}[auto-columns-width]@
$\begin{array}{c}
\begin{bNiceMatrix}
 9 & 17 \\ -2 & 5 
 \end{bNiceMatrix} \\ \\
\begin{bNiceMatrix}
 1   & 1245345 \\  345 & 2 
\end{bNiceMatrix}
\end{array}$
~emphase#\end{NiceMatrixBlock}@
\end{BVerbatim}
\begin{NiceMatrixBlock}[auto-columns-width]
$\begin{array}{c}
\begin{bNiceMatrix}
 9 & 17 \\ -2 & 5 
 \end{bNiceMatrix} \\ \\
\begin{bNiceMatrix}
 1   & 1245345 \\  345 & 2 
\end{bNiceMatrix}
\end{array}$
\end{NiceMatrixBlock}

\medskip
\textbf{Plusieurs compilations peuvent être nécessaires pour obtenir le résultat désiré.}

\bigskip
\section{Les rangées et colonnes extérieures}
Les environnements de \pkg{nicematrix} permettent de composer des rangées et des
colonnes «extérieures» grâce aux options |first-row|, |last-row|, |first-col| et
|last-col|.
\label{exterior}

Si elle est présente, la «première rangée» (extérieure) est numérotée par $0$
(et non $1$). Il en est de même pour la «première colonne».

\begin{Verbatim}
$\begin{pNiceMatrix}[~emphase#first-row,last-row,first-col,last-col@,nullify-dots]
       & C_1    & \Cdots &        & C_4    &        \\
L_1    & a_{11} & a_{12} & a_{13} & a_{14} & L_1    \\
\Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
       & a_{31} & a_{32} & a_{33} & a_{34} &        \\
L_4    & a_{41} & a_{42} & a_{43} & a_{44} & L_4    \\
       & C_1    & \Cdots &        & C_4    &     
\end{pNiceMatrix}$
\end{Verbatim}
%
\[\begin{pNiceMatrix}[first-row,last-row,first-col,last-col,nullify-dots]
       & C_1    & \Cdots &        & C_4    &        \\
L_1    & a_{11} & a_{12} & a_{13} & a_{14} & L_1    \\
\Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
       & a_{31} & a_{32} & a_{33} & a_{34} &        \\
L_4    & a_{41} & a_{42} & a_{43} & a_{44} & L_4    \\
       & C_1    & \Cdots &        & C_4    &     
\end{pNiceMatrix}\]

\medskip
Les lignes pointillées ont été tracées avec les outils présentés p.~\pageref{Cdots}.

\bigskip
Il y a plusieurs remarques à formuler.
%
\begin{itemize}[beginpenalty=10000]
\item Si on utilise un environnement avec préambule explicite (c'est-à-dire
|{NiceArray}| ou l'une de ses variantes), on ne doit pas mettre dans ce
préambule de spécification de colonne pour les éventuelles première et dernière
colonne : ce sera automatiquement (et nécessairement) une colonne |R| pour la
première colonne et une colonne |L| pour la dernière.

\item On peut se demander comment \pkg{nicematrix} détermine le nombre de rangées et de colonnes nécessaires à la
composition de la «dernière rangée» et de la «dernière colonne».

\begin{itemize}
\item Dans le cas d'un environnement avec préambule, comme |{NiceTabular}| ou
|{pNiceArray}|, le nombre de colonnes se déduit évidemment du préambule.

\item Dans le cas où l'option |light-syntax| (cf. p. \pageref{light-syntax}) est
utilisée, \pkg{nicematrix} profite du fait que cette option nécessite de toutes
manières le chargement complet du contenu de l'environnement (d'où
l'impossibilité de mettre du verbatim dans ce cas-là) avant composition du
tableau. L'analyse du contenu de l'environnement donne le nombre de rangées
(mais pas le nombre de colonnes).

\item Dans les autres cas, \pkg{nicematrix} détermine le nombre de rangées et de
colonnes à la première compilation et l'écrit dans le fichier |.aux| pour
pouvoir l'utiliser à la compilation suivante.

\textsl{Néanmoins, il est possible de donner le numéro de la dernière rangée et
  le numéro de la dernière colonne en arguments des options |last-row| et
  |last-col|, ce qui permettra d'accélérer le processus complet de compilation.}
C'est ce que nous ferons dans la suite.
\end{itemize}

\end{itemize}

\bigskip
On peut contrôler l'apparence de ces rangées et colonnes avec les options
|code-for-first-row|, |code-for-last-row|, |code-for-first-col| et
|code-for-last-col|. Ces options sont des listes de tokens qui seront insérées
au début de chaque case de la rangée ou de la colonne considérée.

\begin{Verbatim}
\NiceMatrixOptions{~emphase#code-for-first-row@ = \color{red},
                   ~emphase#code-for-first-col@ = \color{blue},
                   ~emphase#code-for-last-row@ = \color{green},
                   ~emphase#code-for-last-col@ = \color{magenta}}
$\begin{pNiceArray}{cc|cc}[first-row,last-row=6,first-col,last-col,nullify-dots]
       & C_1    & \Cdots &        & C_4    &        \\
L_1    & a_{11} & a_{12} & a_{13} & a_{14} & L_1    \\
\Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
\hline
       & a_{31} & a_{32} & a_{33} & a_{34} &        \\
L_4    & a_{41} & a_{42} & a_{43} & a_{44} & L_4    \\
       & C_1    & \Cdots &        & C_4    &     
\end{pNiceArray}$
\end{Verbatim}
%
\begin{scope}
\NiceMatrixOptions{code-for-first-row = \color{red},
                   code-for-first-col = \color{blue},
                   code-for-last-row = \color{green},
                   code-for-last-col = \color{magenta}}
\begin{displaymath}
\begin{pNiceArray}{cc|cc}[first-row,last-row=5,first-col,last-col,nullify-dots]
       & C_1    & \multicolumn1c{\Cdots} &        & C_4    &        \\
L_1    & a_{11} & a_{12} & a_{13} & a_{14} & L_1    \\
\Vdots & a_{21} & a_{22} & a_{23} & a_{24} & \Vdots \\
\hline
       & a_{31} & a_{32} & a_{33} & a_{34} &        \\
L_4    & a_{41} & a_{42} & a_{43} & a_{44} & L_4    \\
       & C_1    & \multicolumn1c{\Cdots} &        & C_4    &     
\end{pNiceArray}
\end{displaymath}
\end{scope}

\emph{Remarques}
\begin{itemize}[beginpenalty=10000]
\item Comme on peut le voir dans l'exemple précédent, les filets horizontaux et
verticaux ne s'étendent pas dans les rangées et colonnes extérieures.

Néanmoins, si on veut définir de nouveaux spécificateurs de colonnes pour des
filets (par exemple plus épais), on aura sans doute intérêt à utiliser la
commande |\OnlyMainNiceMatrix| décrite p.~\pageref{OnlyMainNiceMatrix}.
\item Une spécification de couleur présente dans |code-for-first-row| s'applique
à une ligne pointillée tracée dans cette «première rangée» (sauf si une valeur a
été donnée à |xdots/color|). Idem pour les autres.
\item Sans surprise, une éventuelle option |columns-width| (décrite
p.~\pageref{width}) ne s'applique pas à la «première colonne» ni à la «dernière
colonne».
\item Pour des raisons techniques, il n'est pas possible d'utiliser l'option de
la commande |\\| après la «première rangée» ou avant la «dernière rangée» (le
placement des délimiteurs serait erroné).
\end{itemize}

\section{Les lignes en pointillés continues}

\label{Cdots}

À l'intérieur des environnements de l'extension \pkg{nicematrix}, de nouvelles
commandes sont définies : |\Ldots|, |\Cdots|, |\Vdots|, |\Ddots|, et |\Iddots|.
Ces commandes sont conçues pour être utilisées à la place de |\dots|, |\cdots|,
|\vdots|, |\ddots| et |\iddots|.\footnote{La commande |\iddots|, définie dans
  \pkg{nicematrix}, est une variante de |\ddots| avec les points allant vers le
  haut. Si \pkg{mathdots} est chargée, la version de \pkg{mathdots} est
  utilisée. Elle correspond à la commande |\adots| de \pkg{unicode-math}.}
\newcounter{fniddots} \setcounter{fniddots}{\thefootnote}

\smallskip
Chacune de ces commandes doit être utilisée seule dans la case du tableau et
elle trace une ligne en pointillés entre les premières cases non
vides\footnote{La définition précise de ce qui est considéré comme une «case
  vide» est donnée plus loin (cf. p.~\pageref{empty-cells}).} situées de part et
d'autre de la case courante. Bien entendu, pour |\Ldots| et |\Cdots|, c'est une
ligne horizontale ; pour |\Vdots|, c'est une ligne verticale et pour |\Ddots| et
|\Iddots|, ce sont des lignes diagonales. On peut changer la couleur d'une ligne
avec l'option~|color|.\footnote{Il est aussi possible de changer la couleur de
  toutes ces lignes pointillées avec l'option |xdots/color| (\textsl{xdots} pour
  rappeler que cela s'applique à |\Cdots|, |\Ldots|, |Vdots|, etc.) : cf.
  p.~\pageref{customization}). }\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{bNiceMatrix}
a_1      & \Cdots &        & & a_1     \\
\Vdots   & a_2    & \Cdots & & a_2     \\
         & \Vdots & \Ddots[color=red] \\
\\
a_1      & a_2    &        & & a_n 
\end{bNiceMatrix}
\end{BVerbatim}
$\begin{bNiceMatrix}
a_1      & \Cdots &        & & a_1   \\
\Vdots   & a_2    & \Cdots & & a_2   \\
         & \Vdots & \Ddots[color=red] \\
\\
a_1      & a_2    &        & & a_n 
\end{bNiceMatrix}$

\interitem Pour représenter la matrice nulle, on peut choisir d'utiliser le
codage suivant :\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{bNiceMatrix}
0      & \Cdots & 0      \\
\Vdots &        & \Vdots \\
0      & \Cdots & 0 
\end{bNiceMatrix}
\end{BVerbatim}
$\begin{bNiceMatrix}
0      & \Cdots & 0      \\
\Vdots &        & \Vdots \\
0      & \Cdots & 0 
\end{bNiceMatrix}$

\bigskip
On peut néanmoins souhaiter une matrice plus grande. Habituellement, dans un tel
cas, les utilisateurs de LaTeX ajoutent une nouvelle ligne et une nouvelle
colonne. Il est possible d'utiliser la même méthode avec \pkg{nicematrix}
:\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{bNiceMatrix}
0      & \Cdots & \Cdots & 0      \\
\Vdots &        &        & \Vdots \\
\Vdots &        &        & \Vdots \\
0      & \Cdots & \Cdots & 0 
\end{bNiceMatrix}
\end{BVerbatim}
$\begin{bNiceMatrix}
0      & \Cdots & \Cdots & 0      \\
\Vdots &        &        & \Vdots \\
\Vdots &        &        & \Vdots \\
0      & \Cdots & \Cdots & 0 
\end{bNiceMatrix}$

\bigskip
Dans la première colonne de cet exemple, il y a deux instructions |\Vdots| mais,
bien entendu, une seule ligne en pointillés sera.

\bigskip
En fait, dans cet exemple, il aurait été possible de tracer la même matrice plus
rapidement avec le codage suivant :\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{bNiceMatrix}
0       & \Cdots &        & 0      \\
\Vdots  &        &        &        \\
        &        &        & \Vdots \\
0       &        & \Cdots & 0 
\end{bNiceMatrix}
\end{BVerbatim}
$\begin{bNiceMatrix}
0       & \Cdots &        & 0      \\
\Vdots  &        &        &        \\
        &        &        & \Vdots \\
0       &        & \Cdots & 0 
\end{bNiceMatrix}$

\bigskip
Il y a aussi d'autres moyens de changer la taille d'une matrice. On pourrait
vouloir utiliser l'argument optionnel de la commande~|\\| pour l'espacement
vertical et la commande~|\hspace*| dans une case pour l'espacement
horizontal.\footnote{Dans \pkg{nicematrix}, il faut utiliser |\hspace*| et non
  |\hspace| car \pkg{nicematrix} utilise \pkg{array}. Remarquons aussi que l'on
  peut également régler la largeur des colonnes en utilisant l'environnement
  |{NiceArray}| (ou une de ses variantes) avec une colonne de type~|w| ou |W|:
  cf. p.~\pageref{width}}

Toutefois, une commande~|\hspace*| pourrait interférer dans la construction des
lignes en pointillés. C'est pourquoi l'extension \pkg{nicematrix} fournit une
commande~|\Hspace| qui est une variante de |\hspace| transparente pour la
construction des lignes en pointillés de \pkg{nicematrix}.\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
\begin{bNiceMatrix}
0      & \Cdots & ~emphase#\Hspace*{1cm}@ & 0      \\
\Vdots &        &               & \Vdots \\~emphase#[1cm]@
0      & \Cdots &               & 0    
\end{bNiceMatrix}
\end{BVerbatim}
$\begin{bNiceMatrix}
0      & \Cdots & \Hspace*{1cm} & 0      \\
\Vdots &        &               & \Vdots \\[1cm]
0      & \Cdots &               & 0    
\end{bNiceMatrix}$

\subsection{L'option nullify-dots}

Considérons la matrice suivante qui a été composée classiquement avec
l'environnement |{pmatrix}| de \pkg{amsmath}.\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$A = \begin{pmatrix}
h & i & j & k & l & m \\
x &   &   &   &   & x 
\end{pmatrix}$
\end{BVerbatim}
$A = \begin{pmatrix}
h & i & j & k & l & m \\
x   &     &     &    &     & x 
\end{pmatrix}$

\bigskip
Si nous ajoutons des instructions |\ldots| dans la seconde rangée, la géométrie
de la matrice est modifiée.\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$B = \begin{pmatrix}
h & i & j & k & l & m \\
x & \ldots & \ldots  & \ldots & \ldots & x 
\end{pmatrix}$
\end{BVerbatim}
$B = \begin{pmatrix}
h & i & j & k & l & m \\
x   & \ldots   & \ldots  & \ldots & \ldots & x 
\end{pmatrix}$

\bigskip
Par défaut, avec \pkg{nicematrix}, si nous remplaçons |{pmatrix}| par
|{pNiceMatrix}| et |\ldots| par |\Ldots|, la géométrie de la matrice n'est pas
changée.\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$C = \begin{pNiceMatrix}
h & i & j & k & l & m \\
x & \Ldots & \Ldots & \Ldots & \Ldots & x 
\end{pNiceMatrix}$
\end{BVerbatim}
$C = \begin{pNiceMatrix}
h & i & j & k & l & m \\
x & \Ldots & \Ldots  & \Ldots & \Ldots & x 
\end{pNiceMatrix}$

\bigskip
On pourrait toutefois préférer la géométrie de la première matrice $A$ et
vouloir avoir la même géométrie avec une ligne en pointillés continue dans la
seconde rangée. C'est possible en utilisant l'option |nullify-dots| (et une
seule instruction |\Ldots| suffit).\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$D = \begin{pNiceMatrix}[~emphase#nullify-dots@]
h & i & j & k & l & m \\
x & \Ldots & & & & x 
\end{pNiceMatrix}$
\end{BVerbatim}
$D = \begin{pNiceMatrix}[nullify-dots]
h & i & j & k & l & m \\
x & \Ldots & &  &  & x 
\end{pNiceMatrix}$

\medskip
L'option |nullify-dots| «smashe» les instructions |\Ldots| (et ses variantes)
horizontalement mais aussi verticalement.

\subsection{Les commandes \textbackslash Hdotsfor et \textbackslash Vdotsfor}

Certaines personnes utilisent habituellement la commande |\hdotsfor| de
l'extension \pkg{amsmath} pour tracer des lignes en pointillés horizontales dans
une matrice. Dans les environnements de \pkg{nicematrix}, il convient d'utiliser
|\Hdotsfor| à la place pour avoir les lignes en pointillés similaires à toutes
celles tracées par l'extension \pkg{nicematrix}.

Comme avec les autres commandes de \pkg{nicematrix} (comme |\Cdots|, |\Ldots|,
|\Vdots|, etc.), la ligne en pointillés tracée par |\Hdotsfor| s'étend jusqu'au
contenu des cases de part et d'autre.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=7cm]
$\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
1 & ~emphase#\Hdotsfor{3}@ & 5 \\
1 & 2 & 3 & 4 & 5 \\
1 & 2 & 3 & 4 & 5 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
1 & \Hdotsfor{3} & 5 \\
1 & 2 & 3 & 4 & 5 \\
1 & 2 & 3 & 4 & 5 
\end{pNiceMatrix}$

\bigskip
Néanmoins, si ces cases sont vides, la ligne en pointillés s'étend seulement
dans les cases spécifiées par l'argument de |\Hdotsfor| (par conception).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=7cm]
$\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
  & ~emphase#\Hdotsfor{3}@ \\
1 & 2 & 3 & 4 & 5 \\
1 & 2 & 3 & 4 & 5 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}
1 & 2 & 3 & 4 & 5 \\
  & \Hdotsfor{3} \\
1 & 2 & 3 & 4 & 5 \\
1 & 2 & 3 & 4 & 5 
\end{pNiceMatrix}$

\medskip
Remarque : Contrairement à la commande |\hdotsfor| de \pkg{amsmath}, la commande
|\Hdotsfor| est utilisable lorsque l'extension \pkg{colortbl} est chargée (mais
vous risquez d'avoir des problèmes si vous utilisez |\rowcolor| sur la même
rangée que |\Hdotsfor|).

\bigskip
L'extension \pkg{nicematrix} propose aussi une commande |\Vdotsfor| similaire à
|\Hdotsfor| mais traçant des lignes verticales.

\bigskip
L'exemple suivant utilise à la fois |\Hdotsfor| et |\Vdotsfor|:

\begin{Verbatim}[formatcom=\small\color{gray}]
\begin{bNiceMatrix}
C[a_1,a_1] & \Cdots & C[a_1,a_n] 
  & \hspace*{20mm} & C[a_1,a_1^{(p)}] & \Cdots & C[a_1,a_n^{(p)}] \\
\Vdots & \Ddots & \Vdots 
  & ~emphase#\Hdotsfor{1}@ & \Vdots & \Ddots & \Vdots \\ 
C[a_n,a_1] & \Cdots & C[a_n,a_n] 
  & & C[a_n,a_1^{(p)}] & \Cdots & C[a_n,a_n^{(p)}] \\
\rule{0pt}{15mm} & ~emphase#\Vdotsfor{1}@ & & \Ddots & & ~emphase#\Vdotsfor{1}@ \\
C[a_1^{(p)},a_1] & \Cdots & C[a_1^{(p)},a_n] 
  & & C[a_1^{(p)},a_1^{(p)}] & \Cdots & C[a_1^{(p)},a_n^{(p)}] \\
\Vdots & \Ddots & \Vdots 
  & ~emphase#\Hdotsfor{1}@ & \Vdots & \Ddots & \Vdots \\ 
C[a_n^{(p)},a_1] & \Cdots & C[a_n^{(p)},a_n] 
  & & C[a_n^{(p)},a_1^{(p)}] & \Cdots & C[a_n^{(p)},a_n^{(p)}]
\end{bNiceMatrix}
\end{Verbatim}

\[\begin{bNiceMatrix}
C[a_1,a_1] & \Cdots & C[a_1,a_n] & \hspace*{20mm} & C[a_1,a_1^{(p)}] & \Cdots & C[a_1,a_n^{(p)}] \\
\Vdots & \Ddots & \Vdots & \Hdotsfor{1} &  \Vdots & \Ddots & \Vdots \\ 
C[a_n,a_1] & \Cdots & C[a_n,a_n] & & C[a_n,a_1^{(p)}] & \Cdots & C[a_n,a_n^{(p)}] \\
\rule{0pt}{15mm} & \Vdotsfor{1} & & \Ddots & & \Vdotsfor{1} \\
C[a_1^{(p)},a_1] & \Cdots & C[a_1^{(p)},a_n] & & C[a_1^{(p)},a_1^{(p)}] & \Cdots & C[a_1^{(p)},a_n^{(p)}] \\
\Vdots & \Ddots & \Vdots & \Hdotsfor{1} & \Vdots & \Ddots & \Vdots \\ 
C[a_n^{(p)},a_1] & \Cdots & C[a_n^{(p)},a_n] & & C[a_n^{(p)},a_1^{(p)}] & \Cdots & C[a_n^{(p)},a_n^{(p)}]
\end{bNiceMatrix}\]

\subsection{Comment créer les lignes en pointillés de manière transparente}

L'extension \pkg{nicematrix} fournit une option appelée |transparent| qui permet
d'utiliser du code existant de manière transparente dans les environnements de
l'\pkg{amsmath} : |{matrix}|, |{pmatrix}|, etc. En fait, cette option est un
alias pour la conjonction de deux options : |renew-dots| et
|renew-matrix|.\footnote{Comme toutes les autres options, les options
  |renew-dots|, |renew-matrix| et |transparent| peuvent être fixées avec la
  commande |\NiceMatrixOptions|, mais elles peuvent aussi être passées en option
  du |\usepackage| (ce sont les trois seules).}

\smallskip

\begin{itemize}
\item L'option |renew-dots|\par\nobreak

Avec cette option, les commandes |\ldots|, |\cdots|, |\vdots|, |\ddots|,
|\iddots|\footnotemark[\thefniddots] et |\hdotsfor| sont redéfinies dans les
environnements de \pkg{nicematrix} et agissent alors comme |\Ldots|, |\Cdots|,
|\Vdots|, |\Ddots|, |\Iddots| et |\Hdotsfor| ; la commande |\dots| (points de
suspension «automatiques» de |amsmath|) est aussi redéfinie et se comporte comme
|\Ldots|.

\item L'option |renew-matrix|\par\nobreak

Avec cette option, l'environnement |{matrix}| est redéfini et se comporte comme
|{NiceMatrix}| et il en est de même pour les cinq variantes.
\end{itemize}

\bigskip 
Par conséquent, avec l'option |transparent|, un code classique donne directement
le résultat fourni par \pkg{nicematrix}.\par\nobreak

\bigskip
\begin{scope}
\NiceMatrixOptions{transparent}
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
~emphase#\NiceMatrixOptions{transparent}@
\begin{pmatrix}
1      & \cdots & \cdots & 1      \\
0      & \ddots &        & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
0      & \cdots & 0      & 1
\end{pmatrix}
\end{BVerbatim}
$\begin{pmatrix}
1      & \cdots & \cdots & 1      \\
0      & \ddots &        & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
0      & \cdots & 0      & 1
\end{pmatrix}$
\end{scope}

\subsection{Les labels des lignes en pointillés}

Les commandes |\Ldots|, |\Cdots|, |\Vdots|, |\Ddots|, |\Iddots| et |\Hdotsfor|
(ainsi que la commande |\line| dans le |code-after| décrite
p.~\pageref{line-in-code-after}) peuvent en fait prendre deux arguments
optionnels spécifiés par les caractères |_| et |^| pour des labels situés
au-dessous et au-dessus de la ligne. Les arguments sont composés en mode
mathématique avec |\scriptstyle|.

\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{bNiceMatrix}
1 & \hspace*{1cm}            & 0 \\[8mm]
  & ~emphase#\Ddots^{n \text{ times}}@ &   \\
0 &                          & 1 
\end{bNiceMatrix}$
\end{BVerbatim}
$\begin{bNiceMatrix}
1 & \hspace*{1cm}            & 0 \\[8mm]
  & \Ddots^{n \text{ times}} &   \\
0 &                          & 1 
\end{bNiceMatrix}$

\subsection{Personnalisation des lignes en pointillés}

\label{customization}
Les lignes pointillées tracées par |\Ldots|, |\Cdots|, |\Vdots|, |\Ddots|,
|\Iddots| et |\Hdotsfor| (ainsi que par la commande |\line| dans le |code-after|
décrite p.~\pageref{line-in-code-after}) peuvent être paramétrées par trois
options (que l'on met entre crochets après la commande) :
%
\begin{itemize}
\item |color| ;
\item |shorten| ; 
\item |line-style|.
\end{itemize}

Ces options peuvent aussi être fixées avec |\NiceMatrixOptions| ou bien au
niveau d'un environnement mais elles doivent alors être préfixées par |xdots|,
ce qui fait que leurs noms deviennent :
%
\begin{itemize}
\item |xdots/color| ;
\item |xdots/shorten| ; 
\item |xdots/line-style|.
\end{itemize}
%
Pour la clarté, dans la suite, on utilisera ces noms-là.

\bigskip
\textbf{L'option xdots/color}\par\nobreak

\smallskip
L'option |xdots/color| indique bien entendu la couleur de la ligne tracée. On
remarquera néanmoins que les lignes tracées dans les rangées et colonnes
extérieures (décrites plus loin) bénificient d'un régime spécial : cf.
p.~\pageref{exterior}.

\bigskip
\textbf{L'option xdots/shorten}\par\nobreak

\smallskip
L'option |xdots/shorten| indique la marge qui est laissée aux deux extrémités de
la ligne. Le nom s'inspire des options «|shorten >|» et «|shorten <|» de Tikz,
mais il faut remarquer que \pkg{nicematrix} ne propose que |xdots/shorten|. La
valeur initiale de ce paramètre est de $0.3$~em (il est conseillé d'utiliser une
unité de mesure dépendante de la fonte courante).

\bigskip
\textbf{L'option xdots/line-style}\par\nobreak

\smallskip
Il faut savoir que, par défaut, les lignes de Tikz tracées avec le paramètre
|dotted| sont composées de points carrés et non pas ronds.\footnote{La raison de
  départ est que le format \textsc{pdf} comporte un système de description de
  lignes en tiretés, qui, puisqu'il est incorporé dans le \textsc{pdf}, est
  affiché très rapidement par les lecteurs de \textsc{pdf}. Il est facile à
  partir de ce type de ligne de créer des lignes de points carrés alors qu'une
  ligne de points ronds doit être construite explicitement point par point.}

\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\tikz \draw [dotted] (0,0) -- (5,0) ;
\end{BVerbatim}
\tikz \draw [dotted] (0,0) -- (5,0) ;

\bigskip
Voulant proposer des lignes avec des points ronds dans le style de celui de
|\ldots| (au moins celui des fontes \emph{Computer Modern}), l'extension
\pkg{nicematrix} contient en interne son propre système de ligne en pointillés
(qui, au passage, n'utilise que \pkg{pgf} et non \pkg{tikz}). Ce style est
appelé le style |standard|. Cette valeur est la valeur initiale du paramètre
|xdots/line-style|.

\bigskip
Néanmoins (quand Tikz est chargé), on peut utiliser pour |xdots/line-style|
n'importe quel style proposé par Tikz, c'est-à-dire n'importe quelle suite
d'options Tikz applicables à un chemin (à l'exception de «|color|», 
«|shorten >|» et «|shorten <|»).

\medskip
Voici par exemple une matrice tridiagonale avec le style |loosely dotted| :\par\nobreak

\medskip
\begin{BVerbatim}[baseline=c]
$\begin{pNiceMatrix}[nullify-dots,~emphase#xdots/line-style=loosely dotted@]
a      & b      & 0      &        & \Cdots & 0      \\ 
b      & a      & b      & \Ddots &        & \Vdots \\
0      & b      & a      & \Ddots &        &        \\
       & \Ddots & \Ddots & \Ddots &        & 0      \\
\Vdots &        &        &        &        & b      \\
0      & \Cdots &        & 0      & b      & a
\end{pNiceMatrix}$
\end{BVerbatim}

\[\begin{pNiceMatrix}[nullify-dots,xdots/line-style=loosely dotted]
a      & b      & 0      &        & \Cdots & 0      \\ 
b      & a      & b      & \Ddots &        & \Vdots \\
0      & b      & a      & \Ddots &        &        \\
       & \Ddots & \Ddots & \Ddots &        & 0      \\
\Vdots &        &        &        &        & b      \\
0      & \Cdots &        & 0      & b      & a
\end{pNiceMatrix}\]

\subsection{Les lignes pointillées et la clé hvlines}

\label{dotted-and-hvlines}

On a dit (cf. p. \pageref{hvlines}) que la clé |hvlines| trace tous les filets
horizontaux et verticaux, exceptés dans les blocs. En fait, avec cette clé, les
filets ne sont pas non plus tracés dans les blocs virtuels délimités par des
cases reliées par des lignes pointillées.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.6cm]
$\begin{bNiceMatrix}[margin,~emphase#hvlines@]
\Block{3-3}<\LARGE>{A} & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
0 & \Cdots& 0 & 0
\end{bNiceMatrix}$
\end{BVerbatim}
$\begin{bNiceMatrix}[margin,hvlines]
\Block{3-3}<\LARGE>{A} & & & 0 \\
& \hspace*{1cm} & & \Vdots \\
& & & 0 \\
0 & \Cdots& 0 & 0
\end{bNiceMatrix}$

\section{Le code-after}

\label{code-after}

L'option |code-after| peut être utilisée pour indiquer du code qui sera exécuté
après la construction de la matrice.\footnote{Il existe aussi une clé
  |code-before| décrite p.~\pageref{code-before}.}

Une commande spéciale, nommée |\line| est disponible pour tracer directement des
lignes en pointillés entre les cases. Elle prend deux arguments correspondant
aux deux cases à relier, chacun de la forme $i$-$j$ où $i$ est le numéro de
ligne et $j$ est le numéro de colonne. Elle peut par exemple être utilisée pour
tracer une ligne entre deux cases adjacentes comme dans l'exemple suivant.
\label{line-in-code-after}

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=11cm]
\NiceMatrixOptions{xdots/shorten = 0.6 em}
\begin{pNiceMatrix}[~emphase#code-after=\line{2-2}{3-3}@]
I       & 0      & \Cdots  &0     \\
0       & I      & \Ddots  &\Vdots\\
\Vdots  &\Ddots  & I       &0     \\
0       &\Cdots  & 0       &I
\end{pNiceMatrix}
\end{BVerbatim}
\begin{scope}
\NiceMatrixOptions{xdots/shorten = 0.6 em}
$\begin{pNiceMatrix}[code-after=\line{2-2}{3-3}]
I       & 0      & \Cdots  &0     \\
0       & I      & \Ddots  &\Vdots\\
\Vdots  &\Ddots  & I       &0     \\
0       &\Cdots  & 0       &I
\end{pNiceMatrix}$
\end{scope}

\medskip
Pour améliorer la lisibilité du code, une syntaxe alternative est proposée : on
peut spécifier les instructions du |code-after| à la fin de l'environnement,
après le mot-clé |\CodeAfter|\footnote{Dans certaines circonstances, on doit
  mettre |\omit \CodeAfter|, |\omit| étant un mot-clé de TeX qui annule le motif
  de la cellule courante.}. Pour un exemple, voir page
\pageref{exemple-CodeAfter}.

\section{Les notes dans les tableaux}

\label{s:notes}

\subsection{Les notes de pied de page}

\smallskip
L'extension \pkg{nicematrix} permet, en utilisant \pkg{footnote} ou bien
\pkg{footnotehyper}, d'extraire les notes insérées avec |\footnote| dans un
environnement de \pkg{nicematrix} pour les reporter en pied de page avec les
autres notes du document.

\smallskip
Si \pkg{nicematrix} est chargée avec l'option |footnote| (avec
|\usepackage[footnote]{nicematrix}| ou avec |\PassOptionsToPackage|),
l'extension \pkg{footnote} est chargée (si elle ne l'est pas déjà) et elle est
utilisée pour extraire les notes de pied de page.

\smallskip
Si \pkg{nicematrix} est chargée avec l'option |footnotehyper|, l'extension
\pkg{footnotehyper} est chargée (si elle ne l'est pas déjà) et elle est utilisée
pour extraire les notes de pied de page.

\smallskip
Attention : Les extensions \pkg{footnote} et \pkg{footnotehyper} sont
incompatibles. L'extension \pkg{footnotehyper} est le successeur de l'extension
\pkg{footnote} et devrait être utilisée préférentiellement. L'extension
\pkg{footnote} a quelques défauts ; en particulier, elle doit être chargée après
l'extension \pkg{xcolor} et elle n'est pas parfaitement compatible avec
\pkg{hyperref}.

\subsection{Les notes de tableaux}

\label{tabularnote}

L'extension \pkg{nicematrix} propose aussi une commande |\tabularnote| qui
permet de spécifier des notes qui seront composées à la fin du tableau avec une
longueur de ligne égale à la largeur du tableau (hors éventuelles colonnes
extérieures). Sans surprise, cette commande n'est disponible que dans les
environnements sans délimiteurs, c'est-à-dire |{NiceTabular}|, |{NiceArray}| et
|{NiceMatrix}|.

En fait, cette commande n'est disponible que si l'extension \pkg{enumitem} a été
chargée (avant ou après \pkg{nicematrix}). Les notes sont en effet composées en
fin de tableau selon un type de liste défini par l'extension \pkg{enumitem}.

\begin{scope}
\fvset{commandchars=\~\#\!}
\begin{Verbatim}
\begin{NiceTabular}{@{}llr@{}}[first-row,code-for-first-row = \bfseries]
\toprule
Nom & Prénom & Date de naissance \\
\midrule
Achard~emphase#\tabularnote{La famille Achard est une très ancienne famille du Poitou.}!
& Jacques & 5 juin 1962 \\
Lefèbvre~emphase#\tabularnote{Le patronyme Lefèbvre est une altération de Lefébure.}!
& Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{Verbatim}
\end{scope}

\begin{center}
\begin{NiceTabular}{@{}llr@{}}[first-row,code-for-first-row = \bfseries]
\toprule
Nom & Prénom & Date de naissance \\
\midrule
Achard\tabularnote{La famille Achard est une très ancienne famille du Poitou.} 
& Jacques & 5 juin 1962 \\
Lefebvre\tabularnote{Le patronyme Lefebvre est une altération de Lefébure.} 
& Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{center}

\begin{table}
\setlength{\belowcaptionskip}{1ex}
\centering
\caption{Utilisation de \texttt{\textbackslash tabularnote}\tabularnote{On peut
    mettre une note dans le titre.}}
\label{t:tabularnote}
\begin{NiceTabular}{@{}llc@{}}[notes/bottomrule]
\toprule
Nom & Prénom & Durée de vie \\
\midrule
Barrère & Bertrand &  86\\
Nightingale\tabularnote{Souvent considérée comme la première
  infirmière.}\tabularnote{Surnommée «la Dame à la Lampe».}
& Florence & 90 \\
Schœlcher & Victor & 89\tabularnote{L'appel de note déborde à droite.}\\
Touchet & Marie & 89 \\
Wallis & John & 87 \\
\bottomrule
\end{NiceTabular}
\end{table}

\bigskip
\begin{itemize}
\item La commande |\tabularnote| est en fait utilisable avant
l'environnement de \pkg{nicematrix}, le but étant de pouvoir l'utiliser sur le
titre inséré par |\caption| dans un environnement |{table}| de LaTeX.

\item Si plusieurs commandes |\tabularnote{...}| se suivent \emph{sans aucun espace
  entre elles}, les appels de notes correspondants sont composés ensemble,
séparés par une virgule (comme avec l'option |multiple| de \pkg{footmisc} pour
les notes de pied de page).

\item Si une commande |\tabularnote{...}| se trouve exactement à la fin d'une cellule
(sans aucun espace après), l'appel de note est composé en débordement vers la
droite (cela peut permettre de mieux conserver l'alignement des contenus d'une
colonne).

\item Si la clé |notes/para| est utilisée, les notes sont composées à la fin du
tableau en un seul paragraphe.

\item Si l'extension \pkg{booktabs} a été chargée (avant ou après
\pkg{nicematrix}), la clé |notes/bottomrule| permet de faire tracer un
|\bottomrule| de \pkg{booktabs} \emph{après} les notes.

\item Il est possible de référencer une note de tableau (avec la commande
|\label| placée après le |\tabularnote|).
\end{itemize}
%
Voir sur la table \ref{t:tabularnote}, p.~\pageref{t:tabularnote}, certaines de ces
remarques illustrées. Cette table a été composée avec le code suivant.

\begin{center}
\fvset{commandchars=\~\#\!}
\begin{BVerbatim}[formatcom=\small\color{gray}]
\begin{table}
\setlength{\belowcaptionskip}{1ex}
\centering
\caption{Utilisation de \texttt{\textbackslash
    tabularnote}~emphase#\tabularnote{On peut!
     ~emphase# mettre une note dans le titre.}!}
\label{t:tabularnote}
\begin{NiceTabular}{@{}llc@{}}[~emphase#notes/bottomrule!]
\toprule
Nom & Prénom & Durée de vie \\
\midrule
Barrère & Bertrand &  86\\
Nightingale~emphase#\tabularnote{Souvent considérée comme la première!
  ~emphase#infirmière.}\tabularnote{Surnommée «la Dame à la Lampe».}!
& Florence & 90 \\
Schœlcher & Victor & 89~emphase#\tabularnote{L'appel de note déborde à droite.}!\\
Touchet & Marie & 89 \\
Wallis & John & 87 \\
\bottomrule
\end{NiceTabular}
\end{table}
\end{BVerbatim}
\end{center}

\subsection{Personnalisation des notes de tableau}

Les notes de tableau peuvent être personnalisées grâce à un ensemble de clés
disponibles dans |\NiceMatrixOptions|. Ces clés ont un nom préfixé par |notes| :
\begin{itemize}
\item |notes/para| 
\item |notes/bottomrule| 
\item |notes/style| 
\item |notes/label-in-tabular| 
\item |notes/label-in-list| 
\item |notes/enumitem-keys| 
\item |notes/enumitem-keys-para| 
\item |notes/code-before|
\end{itemize}
Pour la commmodité, il est aussi possible de fixer ces clés dans
|\NiceMatrixOptions| via une clé |notes| qui prend en argument une liste de
paires \textsl{clé=valeur} où le nom des clés n'a plus à être préfixé par
|notes|:
\begin{center}
\begin{BVerbatim}[formatcom = \small \color{gray}]
\NiceMatrixOptions
  {
    notes = 
     {
       bottomrule ,
       style = ... ,
       label-in-tabular = ... ,
       enumitem-keys = 
        {
          labelsep = ... ,
          align = ... ,
          ...
        }
     }
  }
\end{BVerbatim}
\end{center}

\bigskip
On détaille maintenant ces clés.

\begin{itemize}[itemsep=\medskipamount]
\item La clé |notes/para| demande la composition des notes en fin de tableau en un seul
paragraphe.

Valeur initiale : |false|

Cette clé est également accessible dans un environnement individuel.

\item La clé |notes/bottomrule| permet de faire tracer un |\bottomrule| de \pkg{booktabs}
\emph{après} les notes. Ce trait n'est tracé que s'il y a effectivement des
notes dans le tableau. L'extension \pkg{booktabs} doit avoir été chargée (avant
ou après l'extension \pkg{nicematrix}). Dans le cas contraire, une erreur est générée.

Valeur initiale : |false|

Cette clé est également accessible dans un environnement individuel.

\item La clé |notes/style| est une commande dont l'argument est spécifié par
|#1| et qui indique le style de numérotation des notes. C'est ce style qui est
utilisé par |\ref| pour faire référence à une note de tableau pour laquelle on a
utilisé un |\label|. Ce sont les labels mis en forme avec ce style qui sont
séparés par des virgules quand on utilise plusieurs commandes |\tabularnote|
successivement. Le marqueur |#1| est censé correspondre à un nom de compteur
LaTeX.

Valeur initiale : |\emph{\alph{#1}}|

Une autre valeur possible pourrait être tout simplement |\arabic{#1}|

\item La clé |notes/label-in-tabular| est une commande dont l'argument est
spécifié par |#1| et qui sert au formatage de l'appel de note dans le tableau.
En interne, le numéro de note a déjà été formaté par |notes/style| avant d'être
passé en argument à cette commande.

Valeur initiale : |\textsuperscript{#1}|

Pour la composition du français, il est de tradition de mettre un petit espace
avant l'appel de note. On peut faire ce réglage de la manière suivante :

\begin{Verbatim}
\NiceMatrixOptions{notes/label-in-tabular = \,\textsuperscript{~#1}}
\end{Verbatim}

\item La clé |notes/label-in-list| est une commande dont l'argument est
spécifié par |#1| et qui sert au formatage du numéro de note dans la liste des
notes en fin de tableau. En interne, le numéro de note a déjà été formaté par
|notes/style| avant d'être passé en argument à cette commande.

Valeur initiale : |\textsuperscript{#1}|

Pour la composition du français, on ne compose pas les labels des notes en
lettres supérieures dans la liste des notes. On pourra donc prendre le réglage
suivant :
\begin{Verbatim}
\NiceMatrixOptions{notes/label-in-list = ~#1.\nobreak\hspace{0.25em}}
\end{Verbatim}
La commande |\nobreak| est pour le cas où l'option |para| est utilisée.

\item Les notes sont composées en fin de tableau en utilisant en interne un
style de liste de \pkg{enumitem}.

La clé |notes/enumitem-keys| permet de spécifier une liste de paires
\textsl{clé=valeur} (au format de \pkg{enumitem}) permettant de paramétrer ce
type de liste.

Valeur initiale :\quad |noitemsep , leftmargin = * , align = left , labelsep = 0pt|

Cette valeur initiale contient la spécification |align = left| qui demande que
le label de la note soit composé à gauche dans la boîte qui lui est dévolue.

Ce réglage a l'avantage d'avoir les notes calées à gauche, ce qui est plaisant
si on compose des tableaux dans l'esprit de \pkg{booktabs} (voir par exemple la
table \ref{t:tabularnote}, p.~\pageref{t:tabularnote}).

\item La clé |notes/enumitem-keys-para| est similaire à la précédente mais elle
est utilisée pour le type de liste qui sera utilisé quand l'option |para| est
choisie. Bien entendu, quand cette option |para| est active, c'est une liste de
type |inline| (suivant le vocabulaire de \pkg{enumitem}) qui est utilisée et les
paires \textsl{clé=valeur} doivent donc correspondre à une telle liste de type
|inline|.

Valeur initiale :\quad |afterlabel = \nobreak, itemjoin = \quad|

\item La clé |notes/code-before| est une liste de tokens qui seront insérés
avant la composition de la liste de notes.

Valeur initiale : \textsl{vide}

Si on souhaite, par exemple, que les notes soient composées en gris et en
|\footnotesize|, c'est cette clé qu'il faut utiliser.
\begin{Verbatim}
\NiceMatrixOptions{notes/code-before = \footnotesize \color{gray}}
\end{Verbatim}
On peut aussi mettre dans cette clé |\raggedright| ou |\RaggedRight| (cette
dernière est une commande de \pkg{ragged2e}).
\end{itemize}

\bigskip
Pour un exemple de personnalisation des notes de tableau, voir p.~\pageref{ex:notes}.

\subsection{Utilisation de \{NiceTabular\} avec threeparttable}

Si vous souhaitez utiliser l'environnement |{NiceTabular}| dans un environnement
|{threeparttable}| de l'extension éponyme, vous devez patcher l'environnement
|{threeparttable}| avec le code suivant.
\begin{Verbatim}[commandchars=\~\#\!]
\makeatletter
\AtBeginEnvironment{threeparttable}{\TPT@hookin{NiceTabular}}
\makeatother
\end{Verbatim}

La commande |\AtBeginEnvironment| est une commande de l'extension \pkg{etoolbox}
qui doit donc avoir été chargée au préalable.

\section{Autres fonctionnalités}

\subsection{Utilisation du type de colonne S de siunitx}

Si l'extension \pkg{siunitx} est chargée (avant ou après \pkg{nicematrix}), il est possible d'utiliser les colonnes
de type |S| de \pkg{siunitx} dans les environnements de \pkg{nicematrix}. L'implémentation n'utilise explicitement
aucune macro privée de \pkg{siunitx}.

\medskip
\begin{BVerbatim}[baseline = c, boxwidth = 10.6cm]
$\begin{pNiceArray}{~emphase#S@cW{c}{1cm}c}[nullify-dots,first-row]
{C_1} & \Cdots &  & C_n \\
2.3  & 0 & \Cdots & 0 \\
12.4 & \Vdots & & \Vdots \\
1.45 \\
7.2  & 0 & \Cdots & 0 
\end{pNiceArray}$
\end{BVerbatim}
$\begin{pNiceArray}{ScW{c}{1cm}c}[nullify-dots,first-row]
{C_1} & \Cdots &  & C_n \\
2.3  & 0 & \Cdots & 0 \\
12.4 & \Vdots & & \Vdots \\
1.45 \\
7.2  & 0 & \Cdots & 0 
\end{pNiceArray}$

\medskip
En revanche, les colonnes |d| de l'extension \pkg{dcolumn} ne sont pas prises en
charge par \pkg{nicematrix}.

\subsection{Option d'alignement dans \{NiceMatrix\}}

\label{key-R}

Les environnements sans préambule (|{NiceMatrix}|, |{pNiceMatrix}|,
|{bNiceMatrix}|, etc.) proposent les options |l| et |r| qui imposent des
colonnes alignées à gauche ou à droite.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{bNiceMatrix}[r]
\cos x & - \sin x \\
\sin x & \cos x 
\end{bNiceMatrix}$
\end{BVerbatim}
$\begin{bNiceMatrix}[r]
\cos x & - \sin x \\
\sin x & \cos x 
\end{bNiceMatrix}$

\medskip
Il existe aussi l'option |S| qui impose que toutes les colonnes soient des
colonnes |S| de \pkg{siunitx} (si cette extension est chargée).\footnote{Cela
  reprend une partie des fonctionnalités proposées par les environnements
  |{pmatrix*}|, |{bmatrix*}|, etc. de \pkg{mathtools}.}

\subsection{La commande \textbackslash rotate}

Utilisée au début d'une case, la commande |\rotate| (fournie par \pkg{nicematrix}) compose le contenu après une 
rotation de 90° dans le sens direct.

Dans l'exemple suivant, on l'utilise dans le |code-for-first-row|.

\bigskip

\begin{BVerbatim}[baseline=c,boxwidth=12cm]
\NiceMatrixOptions%
 {code-for-first-row = \scriptstyle ~emphase#\rotate@ \text{image de },
  code-for-last-col = \scriptstyle }
$A = \begin{pNiceMatrix}[first-row,last-col=4]
e_1 & e_2 & e_3       \\
1   & 2   & 3   & e_1 \\
4   & 5   & 6   & e_2 \\
7   & 8   & 9   & e_3 
\end{pNiceMatrix}$
\end{BVerbatim}
\begin{varwidth}{10cm}
\NiceMatrixOptions%
 {code-for-first-row = \scriptstyle\rotate \text{image de },
  code-for-last-col = \scriptstyle }
$ A = \begin{pNiceMatrix}[first-row,last-col=4]
e_1 & e_2 & e_3 \\
1   & 2   & 3  & e_1 \\
4   & 5   & 6  & e_2 \\
7   & 8   & 9  & e_3 \\
\end{pNiceMatrix}$
\end{varwidth}

\bigskip
Si la commande |\rotate| est utilisée dans la «dernière rangée» (extérieure à la
matrice), les éléments qui subissent cette rotation sont alignés vers le haut.

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=12cm]
\NiceMatrixOptions%
 {code-for-last-row = \scriptstyle ~emphase#\rotate@ ,
  code-for-last-col = \scriptstyle }
$A = \begin{pNiceMatrix}[last-row,last-col=4]
1   & 2   & 3   & e_1 \\
4   & 5   & 6   & e_2 \\
7   & 8   & 9   & e_3 \\
\text{image de } e_1 & e_2 & e_3 
\end{pNiceMatrix}$
\end{BVerbatim}
\begin{varwidth}{10cm}
\NiceMatrixOptions%
 {code-for-last-row = \scriptstyle\rotate ,
  code-for-last-col = \scriptstyle }%
$A = \begin{pNiceMatrix}[last-row,last-col=4]
1   & 2   & 3  & e_1 \\
4   & 5   & 6  & e_2 \\
7   & 8   & 9  & e_3 \\
\text{image de } e_1 & e_2 & e_3 
\end{pNiceMatrix}$
\end{varwidth}

\subsection{L'option small}

\label{small}

Avec l'option |small|, les environnements de l'extension \pkg{nicematrix} sont
composés d'une manière proche de ce que propose l'environnement |{smallmatrix}|
de l'\pkg{amsmath} (et les environnements |{psmallmatrix}|, |{bsmallmatrix}|,
etc. de \pkg{mathtools}).

\bigskip
\begin{Verbatim}
$\begin{bNiceArray}{cccc|c}[~emphase#small@, 
                            last-col, 
                            code-for-last-col = \scriptscriptstyle, 
                            columns-width = 3mm ] 
1 & -2 & 3 & 4 & 5 \\
0 & 3  & 2 & 1 & 2 & L_2 \gets 2 L_1 - L_2 \\
0 & 1  & 1 & 2 & 3 & L_3 \gets L_1 + L_3 
\end{bNiceArray}$
\end{Verbatim}
%
\[\begin{bNiceArray}{cccc|c}[small, last-col, code-for-last-col = \scriptscriptstyle, columns-width=3mm]
1 & -2 & 3 & 4 & 5 \\
0 & 3  & 2 & 1 & 2 & L_2 \gets 2 L_1 - L_2 \\
0 & 1  & 1 & 2 & 3 & L_3 \gets L_1 + L_3 
\end{bNiceArray}\]

\bigskip
On remarquera néanmoins que l'environnement |{NiceMatrix}| avec l'option |small|
ne prétend pas être composé exactement comme l'environnement |{smallmatrix}|.
C'est que les environnements de \pkg{nicematrix} sont tous fondés sur |{array}|
(de \pkg{array}) alors que ce n'est pas le cas de |{smallmatrix}| (fondé
directement sur un |\halign| de TeX).

\medskip
En fait, l'option |small| correspond aux réglages suivants :
\begin{itemize}
\item les composantes du tableau sont composées en |\scriptstyle| ; 
\item |\arraystretch| est fixé à $0.47$ ; 
\item |\arraycolsep| est fixé à $1.45$~pt ; 
\item les caractéristiques des lignes en pointillés sont également modifiées.
\end{itemize}

\subsection{Les compteurs iRow et jCol}

Dans les cases du tableau, il est possible d'utiliser les compteurs LaTeX |iRow|
et |jCol| qui représentent le numéro de la rangée courante et le numéro de la
colonne courante\footnote{On rappelle que le numéro de la «première rangée» (si
  elle existe) est $0$ et que le numéro de la «première colonne» (si elle
  existe) est $0$ également.}. Bien entendu, l'utilisateur ne doit pas modifier
les valeurs de ces compteurs qui sont utilisés en interne par \pkg{nicematrix}.

Dans le |code-before| (cf. p. \pageref{code-before}) et dans le |code-after|
(cf. p. \pageref{code-after}), |iRow| représente le nombre total de rangées
(hors éventuelles rangées extérieures) et |jCol| le nombre total de colonnes
(hors potentielles colonnes extérieures).

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=10.6cm]
$\begin{pNiceMatrix}%
    [first-row,
     first-col,
     code-for-first-row = \mathbf{~emphase#\alph{jCol}@} ,
     code-for-first-col = \mathbf{~emphase#\arabic{iRow}@} ]
&   &    &    &   \\
& 1 & 2  & 3  & 4 \\
& 5 & 6  & 7  & 8 \\
& 9 & 10 & 11 & 12
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}[first-row,
                   first-col,
                   code-for-first-row = \mathbf{\alph{jCol}} ,
                   code-for-first-col = \mathbf{\arabic{iRow}} ]
&   &    &    &   \\
& 1 & 2  & 3  & 4 \\
& 5 & 6  & 7  & 8 \\
& 9 & 10 & 11 & 12
\end{pNiceMatrix}$

\medskip
Si des compteurs LaTeX nommés |iRow| ou |jCol| sont créés dans le document par
d'autres extensions que \pkg{nicematrix} (ou tout simplement par l'utilisateur
final), ces compteurs sont masqués dans les environnements de \pkg{nicematrix}.

\bigskip
L'extension \pkg{nicematrix} propose aussi des commandes pour composer
automatiquement des matrices à partir d'un motif général. Ces commandes sont
nommées |\AutoNiceMatrix|, |\pAutoNiceMatrix|, |\bAutoNiceMatrix|,
|\vAutoNiceMatrix|, |\VAutoNiceMatrix| et |\BAutoNiceMatrix|.

Chacune de ces commandes prend deux arguments obligatoires : le premier est la
taille de la matrice, sous la forme $n$-$p$, où $n$ est le nombre de rangées et
$p$ est le nombre de colonnes et le deuxième est le motif (c'est-à-dire
simplement des tokens qui seront insérés dans chaque case de la matrice,
exceptées celles des éventuelles rangées et colonnes extérieures).

\medskip
\begin{Verbatim}
$C = ~emphase#\pAutoNiceMatrix@{3-3}{C_{\arabic{iRow},\arabic{jCol}}}$
\end{Verbatim}

\[C = \pAutoNiceMatrix{3-3}{C_{\arabic{iRow},\arabic{jCol}}}\]

\subsection{L'option light-syntax}

\label{light-syntax}

L'option |light-syntax| (inspirée de l'extension \pkg{spalign}) permet d'alléger
la saisie des matrices, ainsi que leur lisibilité dans le source TeX. Lorsque
cette option est activée, on doit utiliser le point-virgule comme marqueur de
fin de rangée et séparer les colonnes par des espaces ou des tabulations. On
remarquera toutefois que, comme souvent dans le monde TeX, les espaces après les
séquences de contrôle ne sont pas comptées et que les éléments entre accolades
sont considérés comme un tout.

\medskip
\begin{scope}
\begin{BVerbatim}[baseline=c,boxwidth=10cm]
$\begin{bNiceMatrix}[~emphase#light-syntax@,first-row,first-col]
{} a             b                 ;
a  2\cos a       {\cos a + \cos b} ;
b \cos a+\cos b  { 2 \cos b }
\end{bNiceMatrix}$
\end{BVerbatim}
\end{scope}
% 
$\begin{bNiceMatrix}[light-syntax,first-row,first-col]
{} a             b                 ;
a  2\cos a       {\cos a + \cos b} ;
b \cos a+\cos b  { 2 \cos b }
\end{bNiceMatrix}$

\medskip
On peut changer le caractère utilisé pour indiquer les fins de rangées avec
l'option |end-of-row|. Comme dit précédemment, la valeur initiale de ce
paramètre est un point-virgule.

\medskip
Lorsque l'option |light-syntax| est utilisée, il n'est pas possible de mettre
d'éléments en verbatim (avec par exemple la commande
|\verb|) dans les cases du tableau.\footnote{La raison en est que lorsque l'option |light-syntax|
  est utilisée, le contenu complet de l'environnement est chargé comme un
  argument de commande TeX. L'environnement ne se comporte plus comme un «vrai»
  environnement de LaTeX qui se contente d'insérer des commandes avant et
  après.}

\subsection{L'environnement \{NiceArrayWithDelims\}}

En fait, l'environnement |{pNiceArray}| et ses variantes sont fondés sur un
environnement plus général, appelé |{NiceArrayWithDelims}|. Les deux premiers
arguments obligatoires de cet environnement sont les délimiteurs gauche et droit
qui seront utilisés dans la construction de la matrice. Il est possible
d'utiliser |{NiceArrayWithDelims}| si on a besoin de délimiteurs atypiques ou
asymétriques.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=11cm]
$\begin{~emphase#NiceArrayWithDelims@}
   {\downarrow}{\uparrow}{ccc}[margin]
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\end{~emphase#NiceArrayWithDelims@}$
\end{BVerbatim}
%
$\begin{NiceArrayWithDelims}
   {\downarrow}{\uparrow}{ccc}[margin]
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\end{NiceArrayWithDelims}$

\section{Utilisation de Tikz avec nicematrix}

\label{name}

\subsection{Les nœuds correspondant aux contenus des cases}

L'extension \pkg{nicematrix} crée un nœud PGF-Tikz pour chaque case (non vide)
du tableau considéré. Ces nœuds sont utilisés, entre autres, pour tracer les
lignes en pointillés entre les cases du tableau.

\smallskip
Tous les nœuds du document doivent avoir des noms deux à deux distincts et le
nom de ces nœuds doit donc faire intervenir le numéro de l'environnement
courant. Les environnements créés par \pkg{nicematrix} sont en effet numérotés
par un compteur global interne.

\smallskip
Si l'environnement concerné a le numéro $n$, alors le nœud de la rangée~$i$ et
de la colonne~$j$ a pour nom |nm-|$n$|-|$i$|-|$j$.

\smallskip
La commande |\NiceMatrixLastEnv| donne le numéro du dernier de ces
environnements (pour LaTeX, il s'agit d'une commande — complètement développable
— et non d'un compteur).

\smallskip
Il est néanmoins recommandé de passer plutôt par la clé |name|. Celle-ci permet
de donner un nom à l'environnement. Une fois l'environnement nommé, les nœuds
sont accessibles à travers les noms «\textsl{nom}-$i$-$j$» où \textsl{nom} est
le nom donné au tableau et $i$ et $j$ les numéros de rangée et de colonne de la
case considérée. On peut les utiliser avec \textsc{pgf} mais l'utilisateur final
préférera sans doute utiliser Tikz (qui est une sur-couche de \textsc{pgf}). Il
faut néanmoins se souvenir que \pkg{nicematrix} ne charge pas Tikz par défault.

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=11cm]
$\begin{pNiceMatrix}[name=~emphase#ma-matrice@]
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\end{pNiceMatrix}$
\tikz[remember picture,overlay] 
     \draw ~emphase#(ma-matrice-2-2)@ circle (2mm) ; 
\end{BVerbatim}
$\begin{pNiceMatrix}[name=ma-matrice]
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\end{pNiceMatrix}$
\tikz[remember picture,overlay] 
     \draw (ma-matrice-2-2) circle (2mm) ;

\medskip
Ne pas oublier les options |remember picture| et |overlay|.

\bigskip
Dans le |code-after|, et si Tikz est chargé, les choses sont plus simples. On
peut (et on doit) désigner les nœuds sous la forme $i$-$j$ : il n'y a pas besoin
de préciser l'environnement qui est évidemment l'environnement courant.

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=11cm]
$\begin{pNiceMatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\CodeAfter
\tikz \draw ~emphase#(2-2)@ circle (2mm) ; 
\end{pNiceMatrix}$
\end{BVerbatim}
$\begin{pNiceMatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9 
\CodeAfter
\tikz \draw (2-2) circle (2mm) ; 
\end{pNiceMatrix}$

\bigskip
Dans l'exemple suivant, nous avons surligné tous les nœuds de la matrice (on
explique plus loin la technique utilisée : cf. p. \pageref{highlight}).

\[\begin{pNiceMatrix}
a & a + b & a + b + c \\
a & a     & a + b  \\
a & a     & a 
\CodeAfter
\begin{tikzpicture}[ 
                     every node/.style = 
                       { 
                         blend mode = multiply,
                         inner sep = 0 pt ,
                         fill = red!15
                       }
                   ]
\node [fit = (1-1)] {} ;
\node [fit = (1-3)] {} ;
\node [fit = (2-2)] {} ;
\node [fit = (3-1)] {} ;
\node [fit = (3-3)] {} ;
\node [fit = (1-2)] {} ;
\node [fit = (2-1)] {} ;
\node [fit = (2-3)] {} ;
\node [fit = (3-2)] {} ;
\end{tikzpicture}
\end{pNiceMatrix}\]

\subsection{Les «nœuds moyens» et les «nœuds larges»}

En fait, l'extension \pkg{nicematrix} peut créer deux séries de nœuds
supplémentaires (\emph{extra nodes} en anglais) : les «nœuds moyens»
(\emph{medium nodes} en anglais) et les «nœuds larges» (\emph{large nodes} en
anglais). Les premiers sont créés avec l'option |create-medium-nodes| et les
seconds avec l'option |create-large-nodes|.\footnote{Il existe aussi l'option
  |create-extra-nodes| qui est un alias pour la conjonction de
  |create-medium-nodes| et |create-large-nodes|.}

\medskip
Ces nœuds ne sont pas utilisés par défault par \pkg{nicematrix}.

\medskip
Les noms des «nœuds moyens» s'obtiennent en ajoutant le suffixe «|-medium|» au
nom des nœuds normaux. Dans l'exemple suivant, on a surligné tous les «nœuds
moyens». Nous considérons que cet exemple se suffit à lui-même comme définition
de ces nœuds.
\[\begin{pNiceMatrix}[
 create-medium-nodes,
 code-after = {\begin{tikzpicture}
                  [every node/.style = {fill = red!15,
                                        blend mode = multiply,
                                        inner sep = 0 pt},
                   name suffix = -medium]
               \node [fit = (1-1)] {} ;
               \node [fit = (1-2)] {} ;
               \node [fit = (1-3)] {} ;
               \node [fit = (2-1)] {} ;
               \node [fit = (2-2)] {} ;
               \node [fit = (2-3)] {} ;
               \node [fit = (3-1)] {} ;
               \node [fit = (3-2)] {} ;
               \node [fit = (3-3)] {} ;
               \end{tikzpicture}}]
a & a + b & a + b + c \\
a & a     & a + b  \\
a & a     & a 
\end{pNiceMatrix}\]

\medskip
Les noms des «nœuds larges» s'obtiennent en ajoutant le suffixe «|-large|» au
nom des nœuds normaux. Dans l'exemple suivant, on a surligné tous les «nœuds
larges». Nous considérons que cet exemple se suffit à lui-même comme définition
de ces nœuds.\footnote{Il n'y a pas de «nœuds larges» créés dans les rangées et
  colonnes extérieures (pour ces rangées et colonnes, voir
  p.~\pageref{exterior}).}

\[\begin{pNiceMatrix}[
 create-large-nodes,
 code-after = {\begin{tikzpicture}
                  [every node/.style = {blend mode = multiply,
                                        inner sep = 0 pt},
                   name suffix = -large]
               \node [fit = (1-1),fill = red!15] {} ;
               \node [fit = (1-3),fill = red!15] {} ;
               \node [fit = (2-2),fill = red!15] {} ;
               \node [fit = (3-1),fill = red!15] {} ;
               \node [fit = (3-3),fill = red!15] {} ;
               \node [fit = (1-2),fill = blue!15] {} ;
               \node [fit = (2-1),fill = blue!15] {} ;
               \node [fit = (2-3),fill = blue!15] {} ;
               \node [fit = (3-2),fill = blue!15] {} ;
               \end{tikzpicture}}]
a & a + b & a + b + c \\
a & a     & a + b  \\
a & a     & a 
\end{pNiceMatrix}\]

\medskip
Les «nœuds larges» de la première colonne et de la dernière colonne peuvent
apparaître trop petits pour certains usages. C'est pourquoi il est possible
d'utiliser les options |left-margin| et |right-margin| pour ajouter de l'espace
des deux côtés du tableau et aussi de l'espace dans les «nœuds larges» de la
première colonne et de la dernière colonne. Dans l'exemple suivant, nous avons
utilisé les options |left-margin| et |right-margin|.\footnote{Les options
  |left-margin| et |right-margin| prennent des dimensions comme valeurs mais, si
  aucune valeur n'est donnée, c'est la valeur par défaut qui est utilisée et
  elle est égale à |\arraycolsep| (valeur initial : 5~pt). Il existe aussi une
  option |margin| pour fixer à la fois |left-margin| et |right-margin|.}
\[\begin{pNiceMatrix}[
 create-large-nodes,left-margin,right-margin,
 code-after = {\begin{tikzpicture}
                  [every node/.style = {blend mode = multiply,
                                        inner sep = 0 pt},
                   name suffix = -large]
               \node [fit = (1-1),fill = red!15] {} ;
               \node [fit = (1-3),fill = red!15] {} ;
               \node [fit = (2-2),fill = red!15] {} ;
               \node [fit = (3-1),fill = red!15] {} ;
               \node [fit = (3-3),fill = red!15] {} ;
               \node [fit = (1-2),fill = blue!15] {} ;
               \node [fit = (2-1),fill = blue!15] {} ;
               \node [fit = (2-3),fill = blue!15] {} ;
               \node [fit = (3-2),fill = blue!15] {} ;
               \end{tikzpicture}}]
a & a + b & a + b + c \\
a & a     & a + b  \\
a & a     & a 
\end{pNiceMatrix}\]

\medskip
Il est aussi possible d'ajouter de l'espace sur les côtés du tableau avec les
options |extra-left-margin| et |extra-right-margin|. Ces marges ne sont pas
incorporées dans les «nœuds larges». Dans l'exemple suivant, nous avons utilisé
|extra-left-margin| et |extra-right-margin| avec la valeur $3$~pt.
\[\begin{pNiceMatrix}[
 create-large-nodes,left-margin,right-margin,extra-right-margin=3pt,extra-left-margin=3pt,
 code-after = {\begin{tikzpicture}
                  [every node/.style = {blend mode = multiply,
                                        inner sep = 0 pt},
                   name suffix = -large]
               \node [fit = (1-1),fill = red!15] {} ;
               \node [fit = (1-3),fill = red!15] {} ;
               \node [fit = (2-2),fill = red!15] {} ;
               \node [fit = (3-1),fill = red!15] {} ;
               \node [fit = (3-3),fill = red!15] {} ;
               \node [fit = (1-2),fill = blue!15] {} ;
               \node [fit = (2-1),fill = blue!15] {} ;
               \node [fit = (2-3),fill = blue!15] {} ;
               \node [fit = (3-2),fill = blue!15] {} ;
               \end{tikzpicture}}]
a & a + b & a + b + c \\
a & a     & a + b  \\
a & a     & a 
\end{pNiceMatrix}\]

\bigskip
\textbf{Attention} : Ces nœuds sont reconstruits à partir des contenus des cases
et ne correspondent donc pas nécessairement aux cases délimitées par des filets.

\bigskip
\begin{minipage}[c]{7.5cm}
Voici un tableau qui a été composé de la manière suivante :

\medskip
\begin{BVerbatim}
\large
\begin{NiceTabular}{wl{2cm}ll}[hvlines]
fraise & amande & abricot \\
prune & pêche & poire  \\[1ex]
noix & noisette & brugnon
\end{NiceTabular}
\end{BVerbatim}
\end{minipage}
\hspace{1cm}
\begin{scope}
\large
\begin{NiceTabular}[c]{wl{2cm}ll}[hvlines]
fraise & amande & abricot \\
prune & pêche & poire  \\[1ex]
noix & noisette & brugnon
\end{NiceTabular}
\end{scope}

\vspace{1cm}
\begin{minipage}[c]{7cm}
Ci-contre, on a colorié toutes les cases de ce tableau avec |\chessboardcolors|.
\end{minipage}
\hspace{1.5cm}
\begin{scope}
\large
\begin{NiceTabular}[c]{wl{2cm}ll}[hvlines,code-before = \chessboardcolors{red!15}{blue!15}]
fraise & amande & abricot \\
prune & pêche & poire  \\[1ex]
noix & noisette & brugnon
\end{NiceTabular}
\end{scope}

\vspace{1cm}
\begin{minipage}[c]{7cm}
Voici maintenant tous les «nœuds larges» de ce tableau (sans utilisation de
|margin| ni de |extra-margin|).
\end{minipage}
\hspace{1.5cm}
\begin{scope}
\large
\begin{NiceTabular}[c]{w{l}{2cm}ll}[hvlines,
 create-large-nodes,
 code-after = {\begin{tikzpicture}
                  [every node/.style = {blend mode = multiply,
                                        inner sep = 0 pt},
                   name suffix = -large]
               \node [fit = (1-1),fill = red!15] {} ;
               \node [fit = (1-3),fill = red!15] {} ;
               \node [fit = (2-2),fill = red!15] {} ;
               \node [fit = (3-1),fill = red!15] {} ;
               \node [fit = (3-3),fill = red!15] {} ;
               \node [fit = (1-2),fill = blue!15] {} ;
               \node [fit = (2-1),fill = blue!15] {} ;
               \node [fit = (2-3),fill = blue!15] {} ;
               \node [fit = (3-2),fill = blue!15] {} ;
               \end{tikzpicture}}]
fraise & amande & abricot \\
prune & pêche & poire  \\[1ex]
noix & noisette & brugnon
\end{NiceTabular}
\end{scope}

\subsection{Les nœuds «row» et «col»}

L'extension \pkg{nicematrix} crée un nœud PGF-Tikz indiquant la position
potentielle de chaque filet horizontal (avec les noms |row-|$i$) et de chaque
filet vertical (avec les noms |col-|$j$), comme décrit sur la figure ci-dessous.
Ces nœuds sont accessibles dans le |code-before| et dans le |code-after|.
\begin{center}
\begin{NiceTabular}{ccc}[hvlines,rules/width=1pt,rules/color=gray]
rose & tulipe & lys \\
arum & iris & violette \\
muguet & dahlia & souci 
\CodeAfter
\tiny
\begin{tikzpicture}
\foreach \i in {1,2,3,4}
  { 
    \fill [red] (row-\i) circle (0.5mm) ; 
    \node [red,anchor=east] at (row-\i) {row-\i} ; 
  } 
\foreach \j in {1,2,3,4}
  { 
    \fill [blue] (col-\j) circle (0.5mm) ; 
    \node [blue,anchor=north] at (col-\j) {col-\j} ; 
  } 
\end{tikzpicture}
\end{NiceTabular}
\end{center}

\bigskip
Si on utilise Tikz (on rappelle que \pkg{nicematrix} ne charge pas Tikz par
défaut), on peut donc accéder (dans le |code-before| et le |code-after|) à
l'intersection du filet horizontal~$i$ et du filet vertical~$j$ avec la syntaxe
|(row-|$i$\verb+-|col-+$j$|)|.

\medskip
\begin{Verbatim}
\[\begin{NiceMatrix}[
  code-before = 
    { 
~emphase#      \tikz \draw [fill = red!15] @
~emphase#         (row-7-|col-4) -- (row-8-|col-4) -- (row-8-|col-5) -- @
~emphase#         (row-9-|col-5) -- (row-9-|col-6) |- cycle ; @
    }
]
1 \\
1 & 1 \\
1 & 2 &  1 \\
1 & 3 &  3 &  1 \\
1 & 4 &  6 &  4 &  1 \\
1 & 5 & 10 & 10 &  5 &  1 \\
1 & 6 & 15 & 20 & 15 &  6 &  1 \\
1 & 7 & 21 & 35 & 35 & 21 &  7 & 1 \\
1 & 8 & 28 & 56 & 70 & 56 & 28 & 8 & 1
\end{NiceMatrix}\]
\end{Verbatim}
% 
\[\begin{NiceMatrix}[
  code-before = 
    { 
      \tikz \draw [fill = red!15] 
         (row-7-|col-4) -- (row-8-|col-4) -- (row-8-|col-5) -- 
         (row-9-|col-5) -- (row-9-|col-6) |- cycle ; 
    }
]
1 \\
1 & 1 \\
1 & 2 &  1 \\
1 & 3 &  3 &  1 \\
1 & 4 &  6 &  4 &  1 \\
1 & 5 & 10 & 10 &  5 &  1 \\
1 & 6 & 15 & 20 & 15 &  6 &  1 \\
1 & 7 & 21 & 35 & 35 & 21 &  7 & 1 \\
1 & 8 & 28 & 56 & 70 & 56 & 28 & 8 & 1
\end{NiceMatrix}\]

\section{Outils pour les développeurs}

Comme son nom l'indique, la liste de tokens |\g_nicematrix_code_after_tl| est
publique (conformément aux conventions de LaTeX3, toute variable dont le nom
débute par |\g_nicematrix| ou |\l_nicematrix| est publique alors que toute variable
dont le nom débute par |\g__nicematrix| ou par |\l__nicematrix| est privée).

\medskip
Cette variable contient le code qui constitue le «|code-after|». Le développeur
peut donc l'utiliser pour y ajouter du code à partir d'une cellule du tableau
(l'affectation devra être globale, ce qui permettra de sortir de la cellule, qui
est un groupe au sens de TeX).

\medskip
\emph{Exemple} : On souhaite écrire une commande |\crossbox| qui barre en croix
la cellule courante. Cette commande prendra en argument optionnel une liste de
couples \textsl{clé}-\textsl{valeur} qui sera passée à Tikz avant que la croix
ne soit tracée.

On peut alors programmer cette commande |\crossbox| de la manière suivante, qui
utilise explicitement la variable publique |\g_nicematrix_code_after_tl|.

\begin{scope}
\fvset{commandchars=\§\¤\μ}
\begin{Verbatim}
\ExplSyntaxOn
\cs_new_protected:Nn \__pantigny_crossbox:nnn
  {
    \begin { tikzpicture }
    \draw [ #3 ] 
          ( row-#1 -| col-\int_eval:n { #2 + 1 } ) 
       -- ( row-\int_eval:n { #1 + 1 } -| col-#2 ) 
          ( row-#1 -| col-\int_eval:n #2 ) 
       -- ( row-\int_eval:n { #1 + 1 } -| col-\int_eval:n { #2 + 1 } ) ; 
    \end { tikzpicture }
  }

\NewDocumentCommand \crossbox { ! O { } }
  {
    \tl_gput_right:Nx §emphase¤\g_nicematrix_code_after_tlμ
      { 
        \__pantigny_crossbox:nnn 
          { \int_use:c { c@iRow } } 
          { \int_use:c { c@jCol } }
          { \exp_not:n { #1 } }
      }
  }
\ExplSyntaxOff
\end{Verbatim}
\end{scope}

\ExplSyntaxOn
\cs_new_protected:Nn \__pantigny_crossbox:nnn
  {
    \begin { tikzpicture }
    \draw [ #3 ] 
          ( row-#1 -| col-\int_eval:n { #2 + 1 } ) 
       -- ( row-\int_eval:n { #1 + 1 } -| col-#2 ) 
          ( row-#1 -| col-\int_eval:n #2 ) 
       -- ( row-\int_eval:n { #1 + 1 } -| col-\int_eval:n { #2 + 1 } ) ; 
    \end { tikzpicture }
  }

\NewDocumentCommand \crossbox { ! O { } }
  {
    \tl_gput_right:Nx \g_nicematrix_code_after_tl
      { 
        \__pantigny_crossbox:nnn 
          { \int_use:c { c@iRow } } 
          { \int_use:c { c@jCol } }
          { \exp_not:n { #1 } }
      }
  }
\ExplSyntaxOff

\bigskip
Voici un exemple d'utilisation :

\medskip
\begin{BVerbatim}[baseline=c,boxwidth=9cm]
\begin{NiceTabular}{ccc}[hvlines]
merlan & requin & cabillaud \\
baleine & ~emphase#\crossbox[red]@ & morue \\
mante & raie & poule 
\end{NiceTabular}
\end{BVerbatim}
\begin{NiceTabular}{ccc}[hvlines]
merlan & requin & cabillaud \\
baleine & \crossbox[red] & morue \\
mante & raie & poule 
\end{NiceTabular}

\section{Remarques techniques}

Première remarque : l'extension \pkg{nicematrix} doit être chargée après
l'extension \pkg{underscore}. Si ce n'est pas le cas, une erreur est levée.

\subsection{Pour définir de nouveaux types de colonnes}

\label{OnlyMainNiceMatrix}

L'extension \pkg{nicematrix} fournit la commande |\OnlyMainNiceMatrix| qui est
destinée à être utilisée dans des définitions de nouveaux types de colonnes. Son
argument n'est exécuté que si on se place dans la partie principale du tableau,
c'est-à-dire que l'on n'est pas dans l'une des éventuelles rangées extérieures.

Par exemple, si on souhaite définir un type de colonne |?| pour tracer un trait fort (noir) d'épaisseur 1~pt, on
pourra écrire\footnote{La commande |\vrule| est une commande de TeX (et non de LaTeX).} :
\begin{Verbatim}
\newcolumntype{?}{!{\OnlyMainNiceMatrix{\vrule width 1 pt}}}
\end{Verbatim}

Le trait fort correspondant ne s'étendra pas dans les rangées
extérieures.\footnote{Un tel filet utilisant les techniques classiques de
  \pkg{array}, il ne traversera pas les doubles filets horizontaux de |\hline\hline|.}

\medskip
\begin{scope}
\newcolumntype{?}{!{\OnlyMainNiceMatrix{\vrule width 1 pt}}}

\begin{BVerbatim}[baseline = c, boxwidth = 10.5cm]
$\begin{pNiceArray}{cc?cc}[first-row,last-row=3]
C_1 & C_2 & C_3 & C_4 \\
a & b & c & d \\
e & f & g & h \\
C_1 & C_2 & C_3 & C_4
\end{pNiceArray}$
\end{BVerbatim}
$\begin{pNiceArray}{cc?cc}[first-row,last-row=3]
C_1 & C_2 & C_3 & C_4 \\
a & b & c & d \\
e & f & g & h \\
C_1 & C_2 & C_3 & C_4
\end{pNiceArray}$
\end{scope}

\medskip
Le spécificateur |?| ainsi créé est aussi utilisable dans les environnements
|{tabular}| et |{array}| (de \pkg{array}) et, dans ce cas, |\OnlyMainNiceMatrix|
est sans effet.

\subsection{Lignes diagonales}

Par défaut, toutes les lignes diagonales\footnote{On parle des lignes créées par
  |\Ddots| et non des lignes créées par une commande |\line| dans le
  |code-after|.} d'un même tableau sont «parallélisées». Cela signifie que la
première diagonale est tracée et que, ensuite, les autres lignes sont tracées
parallèlement à la première (par rotation autour de l'extrémité la plus à gauche
de la ligne). C'est pourquoi la position des instructions |\Ddots| dans un
tableau peut avoir un effet marqué sur le résultat final.

\medskip
Dans les exemples suivants, la première instruction |\Ddots| est marquée en couleur :

\medskip
\begin{scope}
\begin{minipage}{9.5cm}
Exemple avec parallélisation (comportement par défaut):
\begin{Verbatim}
$A = \begin{pNiceMatrix}
1      & \Cdots &        & 1      \\
a+b    & ~emphase#\Ddots@~ &        & \Vdots \\
\Vdots & \Ddots &        &        \\
a+b    & \Cdots & a+b    & 1
\end{pNiceMatrix}$
\end{Verbatim}
\end{minipage}
$A = \begin{pNiceMatrix}
1      & \Cdots &     & 1      \\
a+b    & \Ddots &     & \Vdots \\
\Vdots & \Ddots &     &        \\
a+b    & \Cdots & a+b & 1
\end{pNiceMatrix}$

\bigskip
\NiceMatrixOptions{parallelize-diags=true}%
\begin{minipage}{9.5cm}
\begin{Verbatim}
$A = \begin{pNiceMatrix}
1      & \Cdots &        & 1      \\
a+b    &        &        & \Vdots \\
\Vdots & ~emphase#\Ddots@~ & \Ddots &        \\
a+b    & \Cdots & a+b    & 1
\end{pNiceMatrix}$
\end{Verbatim}
\end{minipage}
$A = \begin{pNiceMatrix}
1      & \Cdots &        & 1      \\
a+b    &        &        & \Vdots \\
\Vdots & \Ddots & \Ddots &        \\
a+b    & \Cdots & a+b    & 1
\end{pNiceMatrix}$

\bigskip
Il est possible de désactiver la parallélisation avec l'option
|parallelize-diags| mise à |false|: \par\nobreak

\medskip
\NiceMatrixOptions{parallelize-diags=false}%
\begin{minipage}{9.5cm}
Le même exemple sans parallélisation :
\end{minipage}
$A = \begin{pNiceMatrix}
1      & \Cdots  &     & 1      \\
a+b    & \Ddots  &     & \Vdots \\
\Vdots & \Ddots  &     &        \\
a+b    & \Cdots  & a+b & 1
\end{pNiceMatrix}$

\end{scope}

\subsection{Les cases «vides»}

\label{empty-cells}
Une instruction comme |\Ldots|, |\Cdots|, etc. essaye de déterminer la première
case vide de part et d'autre de la case considérée. Néanmoins, une case vide
n'est pas nécessairement sans contenu dans le codage TeX (c'est-à-dire sans
aucun token entre les deux esperluettes~|&|). En effet, une case dont le contenu
est |\hspace*{1cm}| peut être considérée comme vide.

\interitem
Pour \pkg{nicematrix}, les règles précises sont les suivantes :

\begin{itemize}
\item Une case implicite est vide. Par exemple, dans la matrice suivante

\begin{Verbatim}
\begin{pmatrix}
a & b \\
c 
\end{pmatrix}
\end{Verbatim}

la dernière case (deuxième rangée et deuxième colonne) est vide.

\medskip
\item Chaque case avec un rendu par TeX de largeur nulle est vide.

\medskip
\item Une case avec une commande |\Hspace| (ou |\Hspace*|) est vide. Cette
commande |\Hspace| est une commande définie par l'extension \pkg{nicematrix}
avec la même signification que |\hspace| excepté que la case où cette commande
est utilisée est considérée comme vide. Cette commande peut être utilisée pour
fixer la largeur des colonnes sans interférer avec le tracé des lignes en
pointillés par \pkg{nicematrix}.

\end{itemize}

\subsection{L'option exterior-arraycolsep}

L'environnement |{array}| insère un espace horizontal égal à |\arraycolsep|
avant et après chaque colonne. En particulier, il y a un espace égal à
|\arraycolsep| avant et après le tableau. Cette caractéristique de
l'environnement |{array}| n'était probablement pas une bonne idée\footnote{Dans
  la documentation de |{amsmath}|, on peut lire : {\itshape The extra space of
    |\arraycolsep| that \pkg{array} adds on each side is a waste so we remove it
    [in |{matrix}|] (perhaps we should instead remove it from array in general,
    but that's a harder task).}}. L'environnement |{matrix}| et ses variantes
(|{pmatrix}|, |{vmatrix}|, etc.) de \pkg{amsmath} préfèrent supprimer ces
espaces avec des instructions explicites |\hskip -\arraycolsep|\footnote{Et non
  en insérant |@{}| de part et d'autre du préambule, ce qui fait que la longueur
  des |\hline| n'est pas modifiée et elle peut paraître trop longue, surtout
  avec des crochets.}. L'extension \pkg{nicematrix} fait de même dans
\emph{tous} ses environnements y compris l'environnement |{NiceArray}|.
Néanmoins, si l'utilisateur souhaite que l'environnement |{NiceArray}| se
comporte par défaut comme l'environnement |{array}| de \pkg{array} (par exemple
pour faciliter l'adaptation d'un document existant), il peut contrôler ce
comportement avec l'option |exterior-arraycolsep| accessible via la commande
|\NiceMatrixOptions|. Avec cette option, des espaces extérieurs de longueur
|\arraycolsep| seront insérés dans les environnements |{NiceArray}| (les autres
environnements de l'extension \pkg{nicematrix} ne sont pas affectés).

\subsection{Incompatibilités}

L'extension \pkg{nicematrix} n'est pas parfaitement compatible avec l'extension \pkg{arydshln} (parce que cette
extension redéfinit de nombreuses commandes internes de \pkg{array}).

\section{Exemples}

\subsection{Notes dans les tableaux}

\label{ex:notes}

Les outils de \pkg{nicematrix} pour les notes dans les tableaux ont été
présentés à la partie \ref{s:notes} p.~\pageref{s:notes}.

\bigskip
Imaginons que l'on souhaite numéroter les notes de tableau (celles construites
avec |\tabularnote|) avec des astérisques.\footnote{Bien entendu, il faut qu'il
  y en ait très peu : trois paraît un maximum.}

On commence par écrire une commande |\stars| similaire aux commandes classiques
|\arabic|, |\alph|, |\Alph|, etc. mais qui produit un nombre d'astérisques égal
à son argument.\footnote{Ou plutôt : à la valeur de son argument.}
\begin{Verbatim}
\ExplSyntaxOn
\NewDocumentCommand ~emphase#\stars@ { m }
  { \prg_replicate:nn { \value { ~#1 } } { $ \star $ } } 
\ExplSyntaxOff
\end{Verbatim}
%
Bien entendu, on change le style des notes avec la clé |notes/style|. Mais, il
serait bon aussi de changer certains paramètres du type de liste (au sens de
\pkg{enumitem}) utilisé pour composer les notes après le tableau. On demande de
composer les labels avec une largeur égale à celle du plus grand des labels. Or,
le label le plus large est bien entendu celui avec le maximum d'astéristiques.
On connaît ce nombre : il est égal à |\value{tabularnote}| (car |tabularnote|
est le compteur LaTeX utilisé par |\tabularnote| et il est donc égal à la fin au
nombre total de notes dans le tableau). On utilise alors la clé |widest*| de 
\pkg{enumitem} pour demander une largeur de label correspondante :
|widest*=\value{tabularnote}|. 
\begin{Verbatim}
\NiceMatrixOptions 
  {
    notes = 
     {
       ~emphase#style = \stars{~#1} , @
       ~emphase#enumitem-keys = @
       ~emphase# { @
       ~emphase#   widest* = \value{tabularnote} ,@
       ~emphase#   align = right @
       ~emphase# } @
     }
  }
\end{Verbatim}

\begin{scope}
\ExplSyntaxOn
\NewDocumentCommand \stars { m }
  { \prg_replicate:nn { \value { #1 } } { $ \star $ } } 
\NiceMatrixOptions 
  {
    notes = 
     {
       style = \stars{#1} ,
       enumitem-keys = 
        {
          widest* = \value{tabularnote} ,
          align = right 
        }
     }
  }
\ExplSyntaxOff
\fvset{commandchars=\~\#\!}
\begin{Verbatim}
\begin{NiceTabular}{@{}llr@{}}[first-row,code-for-first-row = \bfseries]
\toprule
Nom & Prénom & Date de naissance \\
\midrule
Achard~emphase#\tabularnote{La famille Achard est une très ancienne famille du Poitou.}!
& Jacques & 5 juin 1962 \\
Lefèbvre~emphase#\tabularnote{Le patronyme Lefèbvre est une altération de Lefébure.}!
& Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{Verbatim}

\begin{center}
\begin{NiceTabular}{@{}llr@{}}[first-row,code-for-first-row = \bfseries]
\toprule
Nom & Prénom & Date de naissance \\
\midrule
Achard\tabularnote{La famille Achard est une très ancienne famille du Poitou.} 
& Jacques & 5 juin 1962 \\
Lefebvre\tabularnote{Le patronyme Lefebvre est une altération de Lefébure.} 
& Mathilde & 23 mai 1988 \\
Vanesse & Stéphanie & 30 octobre 1994 \\
Dupont & Chantal & 15 janvier 1998 \\
\bottomrule
\end{NiceTabular}
\end{center}
\end{scope}

\subsection{Lignes en pointillés}

Une matrice de permutation.\par\nobreak

À titre d'exemple, on a augmenté la valeur du paramètre
|xdots/shorten|.\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c]
$\begin{pNiceMatrix}[~emphase#xdots/shorten=6em@]
0       & 1 & 0 &        & \Cdots &   0    \\
\Vdots  &   &   & \Ddots &        & \Vdots \\
        &   &   & \Ddots &        &        \\
        &   &   & \Ddots &        &   0    \\
0       & 0 &   &        &        &   1    \\
1       & 0 &   & \Cdots &        &   0    
\end{pNiceMatrix}$
\end{BVerbatim}
\hspace{2.5cm}
$\begin{pNiceMatrix}[xdots/shorten=0.6em]
0       & 1 & 0 &        & \Cdots &   0    \\
\Vdots  &   &   & \Ddots &        & \Vdots \\
        &   &   & \Ddots &        &        \\
        &   &   & \Ddots &        &   0    \\
0       & 0 &   &        &        &   1    \\
1       & 0 &   & \Cdots &        &   0    
\end{pNiceMatrix}$

\vspace{2cm}

Un exemple avec |\Iddots|. On a augmenté encore davantage la valeur de |xdots/shorten|.\par\nobreak
\bigskip
\begin{BVerbatim}[baseline=c]
$\begin{pNiceMatrix}[~emphase#xdots/shorten = 0.9em@]
1       & \Cdots  &         & 1      \\
\Vdots  &         &         & 0      \\
        & ~emphase#\Iddots@ & ~emphase#\Iddots@ & \Vdots \\
1       & 0       & \Cdots  & 0 
\end{pNiceMatrix}$
\end{BVerbatim}
\hspace{4cm}
$\begin{pNiceMatrix}[xdots/shorten = 0.9em]
1       & \Cdots  &         & 1      \\
\Vdots  &         &         & 0      \\
        & \Iddots & \Iddots & \Vdots \\
1       & 0       & \Cdots  & 0 
\end{pNiceMatrix}$

\vspace{2cm}
Un exemple avec |\multicolumn|:\par\nobreak
\bigskip
\begin{BVerbatim}
\begin{BNiceMatrix}[nullify-dots]
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\
\Cdots &  & ~emphase#\multicolumn{6}{c}{10 \text{ autres lignes}}@ & \Cdots \\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10
\end{BNiceMatrix}
\end{BVerbatim}

\bigskip
\[\begin{BNiceMatrix}[nullify-dots]
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\
\Cdots &  & \multicolumn{6}{c}{10 \text{ autres lignes}} & \Cdots \\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10
\end{BNiceMatrix}\]

\vspace{2cm}
Un exemple avec |\Hdotsfor|:\par\nobreak

\bigskip
\begin{BVerbatim}[baseline=c,boxwidth=11cm]
\begin{pNiceMatrix}[nullify-dots]
0 & 1 & 1 & 1 & 1 & 0 \\
0 & 1 & 1 & 1 & 1 & 0 \\
\Vdots  & ~emphase#\Hdotsfor{4}@ & \Vdots \\
 & ~emphase#\Hdotsfor{4}@ & \\
 & ~emphase#\Hdotsfor{4}@ & \\
 & ~emphase#\Hdotsfor{4}@ & \\
0 & 1 & 1 & 1 & 1 & 0 
\end{pNiceMatrix}
\end{BVerbatim}
$\begin{pNiceMatrix}[nullify-dots]
0 & 1 & 1 & 1 & 1 & 0 \\
0 & 1 & 1 & 1 & 1 & 0 \\
\Vdots  & \Hdotsfor{4} & \Vdots \\
 & \Hdotsfor{4} & \\
 & \Hdotsfor{4} & \\
 & \Hdotsfor{4} & \\
0 & 1 & 1 & 1 & 1 & 0 
\end{pNiceMatrix}$

\vspace{2cm}
Un exemple pour le résultant de deux polynômes :
\par\nobreak
\bigskip
\begin{BVerbatim}
\setlength{\extrarowheight}{1mm}
\begin{vNiceArray}{cccc:ccc}[columns-width=6mm]
a_0   &      &&       &b_0    &      &     \\
a_1   &\Ddots&&       &b_1    &\Ddots&     \\
\Vdots&\Ddots&&       &\Vdots &\Ddots&b_0  \\
a_p   &      &&a_0    &       &      &b_1   \\
      &\Ddots&&a_1    &b_q    &      &\Vdots\\
      &      &&\Vdots &       &\Ddots&      \\
      &      &&a_p     &       &      &b_q    
\end{vNiceArray}
\end{BVerbatim}

\bigskip

\begin{scope}
\setlength{\extrarowheight}{1mm}
\[\begin{vNiceArray}{cccc:ccc}[columns-width=6mm]
a_0   &      &&       &b_0    &      &     \\
a_1   &\Ddots&&       &b_1    &\Ddots&     \\
\Vdots&\Ddots&&       &\Vdots &\Ddots&b_0  \\
a_p   &      &&a_0    &       &      &b_1   \\
      &\Ddots&&a_1    &b_q    &      &\Vdots\\
      &      &&\Vdots &       &\Ddots&      \\
      &      &&a_p     &       &      &b_q    
\end{vNiceArray}\]
\end{scope}

\vspace{2cm}
Un exemple avec un système linéaire:\par\nobreak

\begin{Verbatim}
$\begin{pNiceArray}{*6c|c}[nullify-dots,last-col,code-for-last-col=\scriptstyle]
1      & 1 & 1 &\Cdots &   & 1      & 0      & \\
0      & 1 & 0 &\Cdots &   & 0      &        & L_2 \gets L_2-L_1 \\
0      & 0 & 1 &\Ddots &   & \Vdots &        & L_3 \gets L_3-L_1 \\
       &   &   &\Ddots &   &        & \Vdots & \Vdots \\
\Vdots &   &   &\Ddots &   & 0      & \\
0      &   &   &\Cdots & 0 & 1      & 0      & L_n \gets L_n-L_1 
\end{pNiceArray}$
\end{Verbatim}

\[\begin{pNiceArray}{*6c|c}[nullify-dots,last-col,code-for-last-col=\scriptstyle]
1      & 1 & 1 &\Cdots &   & 1      & 0      & \\
0      & 1 & 0 &\Cdots &   & 0      &        & L_2 \gets L_2-L_1 \\
0      & 0 & 1 &\Ddots &   & \Vdots &        & L_3 \gets L_3-L_1 \\
       &   &   &\Ddots &   &        & \Vdots & \Vdots \\
\Vdots &   &   &\Ddots &   & 0      & \\
0      &   &   &\Cdots & 0 & 1      & 0      & L_n \gets L_n-L_1 
\end{pNiceArray}\]

\subsection{Des lignes pointillées qui ne sont plus pointillées}

L'option |line-style| permet de changer le style des lignes tracées par
|\Ldots|, |\Cdots|, etc. On peut de ce fait tracer des lignes qui ne sont plus
pointillées.

\begin{Verbatim}
\NiceMatrixOptions
  {nullify-dots,code-for-first-col = \color{blue},code-for-first-col=\color{blue}}
$\begin{pNiceMatrix}[first-row,first-col]
       &   & \Ldots[line-style={solid,<->},shorten=0pt]^{n \text{ columns}} \\
       & 1 & 1 & 1 & \Ldots & 1 \\
       & 1 & 1 & 1 &  & 1 \\
\Vdots[line-style={solid,<->}]_{n \text{ rows}} 
       & 1 & 1 & 1 &  & 1 \\
       & 1 & 1 & 1 &  & 1 \\
       & 1 & 1 & 1 & \Ldots & 1 
\end{pNiceMatrix}$
\end{Verbatim}

\begin{scope}
\NiceMatrixOptions
  {nullify-dots,code-for-first-col = \color{blue},code-for-first-row=\color{blue}}
\[\begin{pNiceMatrix}[first-row,first-col]
       &   & \Ldots[line-style={solid,<->},shorten=0pt]^{n \text{ columns}} \\
       & 1 & 1 & 1 & \Ldots & 1 \\
       & 1 & 1 & 1 &  & 1 \\
\Vdots[line-style={solid,<->}]_{n \text{ rows}} & 1 & 1 & 1 &  & 1 \\
       & 1 & 1 & 1 &  & 1 \\
       & 1 & 1 & 1 & \Ldots & 1 
\end{pNiceMatrix}\]
\end{scope}

\subsection{Largeur des colonnes}

\medskip
Dans l'exemple suivant, nous utilisons |{NiceMatrixBlock}| avec l'option
|auto-columns-width| parce que nous voulons la même largeur (automatique) pour
toutes les colonnes.

\bigskip
\begin{BVerbatim}
~emphase#\begin{NiceMatrixBlock}[auto-columns-width]@
\NiceMatrixOptions
 { last-col,code-for-last-col = \color{blue}\scriptstyle,light-syntax}
\setlength{\extrarowheight}{1mm}
$\begin{pNiceArray}{cccc:c}
  1  1  1   1   1 { } ;
  2  4  8  16   9 ;
  3  9 27  81  36 ;
  4 16 64 256 100 
\end{pNiceArray}$
\medskip
$\begin{pNiceArray}{cccc:c}
  1  1  1   1  1 ;
  0  2  6  14  7        { L_2 \gets -2 L_1 + L_2 } ;
  0  6 24  78 33        { L_3 \gets -3 L_1 + L_3 } ;
  0 12 60 252 96        { L_4 \gets -4 L_1 + L_4 } 
\end{pNiceArray}$
...
~emphase#\end{NiceMatrixBlock}@
\end{BVerbatim}

\bigskip

\begin{multicols}{2}
\begin{NiceMatrixBlock}[auto-columns-width]
\NiceMatrixOptions{last-col,code-for-last-col = \color{blue}\scriptstyle,light-syntax}
\setlength{\extrarowheight}{1mm}

\enskip $\begin{pNiceArray}{cccc:c}
1  1  1   1   1 { } ;
2  4  8  16   9 ;
3  9 27  81  36 ;
4 16 64 256 100 
\end{pNiceArray}$

\medskip

\enskip $\begin{pNiceArray}{cccc:c}
1  1  1   1  1 ;
0  2  6  14  7 { L_2 \gets -2 L_1 + L_2 } ;
0  6 24  78 33 { L_3 \gets -3 L_1 + L_3 } ;
0 12 60 252 96 { L_4 \gets -4 L_1 + L_4 } 
\end{pNiceArray}$

\medskip

\enskip $\begin{pNiceArray}{cccc:c}
1 1  1  1 1 ;
0 1  3  7 \frac72    { L_2 \gets \frac12 L_2 } ;
0 3 12 39 \frac{33}2 { L_3 \gets \frac12 L_3 } ;
0 1  5 21 8          { L_4 \gets \frac1{12} L_4 }
\end{pNiceArray}$

\medskip

\enskip $\begin{pNiceArray}{cccc:c}
1 1  1   1 1       ;
0 1  3   7 \frac72 ;
0 0  3  18 6        { L_3 \gets -3 L_2 + L_3 } ;
0 0 -2 -14 -\frac92 { L_4 \gets L_2 - L_4 }
\end{pNiceArray}$

\medskip

\enskip $\begin{pNiceArray}{cccc:c}
1 1  1   1 1       ;
0 1  3   7 \frac72 ;
0 0  1   6 2        { L_3 \gets \frac13 L_3 } ;
0 0 -2 -14 -\frac92 
\end{pNiceArray}$

\medskip

\enskip $\begin{pNiceArray}{cccc:c}
1 1 1  1 1       ;
0 1 3  7 \frac72 ;
0 0 1  6 2       ;
0 0 0 -2 -\frac12   { L_4 \gets L_3 + L_4 }
\end{pNiceArray}$
\end{NiceMatrixBlock}
\end{multicols}

\subsection{Comment surligner les cases}

\label{highlight}

\medskip
Les exemples suivants nécessitent d'avoir chargé Tikz (\pkg{nicematrix} ne
charge que \textsc{pgf}) ainsi que la bibliothèque Tikz |fit|, ce qui peut se
faire avec les deux instructions suivantes dans le préambule du document :

\begin{verbatim}
\usepackage{tikz}
\usetikzlibrary{fit}
\end{verbatim}

\medskip
Pour mettre en évidence une case, il est possible de «dessiner» l'un des nœuds
(le «nœud normal», le «nœud moyen» ou le «nœud large»). Dans l'exemple suivant,
on utilise les «nœuds larges» de la diagonale de la matrice (avec la clé de Tikz
«|name suffix|», il est facile d'utiliser les «nœuds larges»).

Nous redessinons les nœuds avec de nouveaux nœuds en utilisant la bibliothèque
\pkg{fit} de Tikz. Comme nous voulons recréer des nœuds identiques aux premiers,
nous devons fixer |inner sep = 0pt| (si on ne fait pas cela, les nouveaux nœuds
seront plus grands que les nœuds d'origine créés par \pkg{nicematrix}).

\label{exemple-CodeAfter}

\begin{Verbatim}
$\begin{pNiceArray}{>{\strut}cccc}[create-large-nodes,margin,extra-margin = 2pt]
  a_{11} & a_{12} & a_{13} & a_{14} \\
  a_{21} & a_{22} & a_{23} & a_{24} \\
  a_{31} & a_{32} & a_{33} & a_{34} \\
  a_{41} & a_{42} & a_{43} & a_{44} 
\CodeAfter
  \begin{tikzpicture}[~emphase#name suffix = -large@,
                      every node/.style = {draw,~emphase#inner sep = 0 pt@}]
     \node [fit = (1-1)] {} ; 
     \node [fit = (2-2)] {} ; 
     \node [fit = (3-3)] {} ; 
     \node [fit = (4-4)] {} ; 
  \end{tikzpicture}
\end{pNiceArray}$
\end{Verbatim}
%
\[\begin{pNiceArray}{>{\strut}cccc}[create-large-nodes,margin,extra-margin = 2pt]
a_{11} & a_{12} & a_{13} & a_{14} \\
a_{21} & a_{22} & a_{23} & a_{24} \\
a_{31} & a_{32} & a_{33} & a_{34} \\
a_{41} & a_{42} & a_{43} & a_{44} 
\CodeAfter
\begin{tikzpicture}[name suffix = -large,
                    every node/.style = {draw,inner sep = 0 pt}]
   \node [fit = (1-1)] {} ; 
   \node [fit = (2-2)] {} ; 
   \node [fit = (3-3)] {} ; 
   \node [fit = (4-4)] {} ; 
\end{tikzpicture}
\end{pNiceArray}\]

On remarquera que les traits que l'on vient de tracer sont dessinés \emph{après}
la matrice sans modifier la position des composantes de celle-ci. En revanche,
les traits tracés par |\hline|, le spécificateur «\verb+|+» ou les options
|hlines|, |vlines| et |hvlines| «écartent» les composantes de la
matrice.\footnote{Pour la commande |\cline|, voir la remarque
  p.~\pageref{remark-cline}.}

\vspace{1cm} Il est possible de colorier une rangée avec |\rowcolor| dans le
|code-before| (ou avec |\rowcolor| de \pkg{colortbl} dans la première case de la
rangée ). Les possibilités de réglages sont néanmoins limitées. C'est pourquoi
nous présentons ici une autre méthode pour surligner une rangée de la matrice.
Nous créons un nœud Tikz rectangulaire qui englobe les nœuds de la deuxième
rangée en utilisant les outils de la bibliothèque Tikz \pkg{fit}. Ce nœud est
rempli après la construction de la matrice. Pour que l'on puisse voir le texte
\emph{sous} le nœud, nous devons utiliser la transparence avec le |blend mode|
égal à |multiply|.

\tikzset{highlight/.style={rectangle,
                           fill=red!15,
                           blend mode = multiply, 
                           rounded corners = 0.5 mm, 
                           inner sep=1pt,
                           fit=#1}}

\medskip
\begin{Verbatim}
\tikzset{highlight/.style={rectangle,
                           fill=red!15,
                           ~emphase#blend mode = multiply@,
                           rounded corners = 0.5 mm, 
                           inner sep=1pt,
                           fit=~#1}}

$\begin{bNiceMatrix}[~emphase#code-after = {\tikz \node [highlight = (2-1) (2-3)] {} ;}@]
0 & \Cdots & 0 \\
1 & \Cdots & 1 \\
0 & \Cdots & 0 
\end{bNiceMatrix}$
\end{Verbatim}

\[\begin{bNiceMatrix}[code-after = {\tikz \node [highlight = (2-1) (2-3)] {} ;}]
0 & \Cdots & 0 \\
1 & \Cdots & 1 \\
0 & \Cdots & 0 
\end{bNiceMatrix}\]

\bigskip
Ce code échoue avec |latex|-|dvips|-|ps2pdf| parce que Tikz pour |dvips|, pour
le moment, ne prend pas en charge les \emph{blend modes}. Néanmoins, le code
suivant, dans le préambule du document LaTeX, devrait activer les \emph{blend
  modes} pour ce mode de compilation.

\begin{scope} \small
|\ExplSyntaxOn|

|\makeatletter|

|\tl_set:Nn \l_tmpa_tl {pgfsys-dvips.def}|

|\tl_if_eq:NNT \l_tmpa_tl \pgfsysdriver|

|  {\cs_set:Npn\pgfsys@blend@mode#1{\special{ps:~/\tl_upper_case:n #1~.setblendmode}}}|

|\makeatother|

|\ExplSyntaxOff|
\end{scope}

\vspace{1cm} On rappelle que dans le cas d'un ensemble de cases fusionnées (avec
la commande |\Block|), un nœud Tikz est créé pour l'ensemble des cases avec pour
nom $i$|-|$j$|-block| où $i$ et $j$ sont les numéros de ligne et de colonne de
la case en haut à gauche (où a été utilisée la commande |\Block|). Si on a
demandé la création des nœuds |medium|, alors un nœud de ce type est aussi créé
pour ce bloc avec un nom suffixé par |-medium|.

\vspace{1cm}
On considère maintenant la matrice suivante que l'on a appelée |exemple|.

\medskip
\begin{Verbatim}
$\begin{pNiceArray}{ccc}[~emphase#name=exemple@,last-col,create-medium-nodes]
a & a + b & a + b + c & L_1 \\
a & a     & a + b     & L_2 \\
a & a     & a         & L_3
\end{pNiceArray}$
\end{Verbatim}

\[\begin{pNiceArray}{ccc}[last-col]
a & a + b & a + b + c & L_1 \\
a & a     & a + b     & L_2 \\
a & a     & a         & L_3
\end{pNiceArray}\]

\bigskip
Si on veut surligner chaque rangée de la matrice, on peut utiliser la technique précédente trois fois.

\begin{Verbatim}
\tikzset{mes-options/.style={remember picture, 
                             overlay,
                             name prefix = exemple-,
                             highlight/.style = {fill = red!15,
                                                 blend mode = multiply,
                                                 inner sep = 0pt,
                                                 fit = ~#1}}}
\end{Verbatim}

\tikzset{mes-options/.style={remember picture, 
                             overlay,
                             name prefix = exemple-,
                             highlight/.style = {fill = red!15,
                                                 blend mode = multiply,
                                                 inner sep = 0pt,
                                                 fit = #1}}}

\begin{Verbatim}
\begin{tikzpicture}[mes-options]
\node [highlight = (1-1) (1-3)] {} ;
\node [highlight = (2-1) (2-3)] {} ;
\node [highlight = (3-1) (3-3)] {} ;
\end{tikzpicture}
\end{Verbatim}

\medskip
On obtient la matrice suivante.

\[\begin{pNiceArray}{ccc}[
     last-col,
     code-after = {\begin{tikzpicture}[highlight/.style = {fill = red!15,
                                                           blend mode = multiply,
                                                           inner sep = 0pt,
                                                           fit = #1}]
                   \node [highlight = (1-1) (1-3)] {} ;
                   \node [highlight = (2-1) (2-3)] {} ;
                   \node [highlight = (3-1) (3-3)] {} ;
                   \end{tikzpicture}}]
a & a + b & a + b + c & L_1 \\
a & a     & a + b     & L_2 \\
a & a     & a         & L_3
\end{pNiceArray}\]

\medskip
Le résultat peut paraître décevant. On peut l'améliorer en utilisant les «nœuds
moyens» au lieu des «nœuds normaux».

\begin{Verbatim}
\begin{tikzpicture}[mes-options, ~emphase#name suffix = -medium@]
\node [highlight = (1-1) (1-3)] {} ;
\node [highlight = (2-1) (2-3)] {} ;
\node [highlight = (3-1) (3-3)] {} ;
\end{tikzpicture}
\end{Verbatim}

\medskip
On obtient la matrice suivante.

\[\begin{pNiceArray}{ccc}[
     last-col,
     create-medium-nodes,
     code-after = {\begin{tikzpicture}[highlight/.style = {fill = red!15,
                                                           blend mode = multiply,
                                                           inner sep = 0pt,
                                                           fit = #1},
                                       name suffix = -medium]
                   \node [highlight = (1-1) (1-3)] {} ;
                   \node [highlight = (2-1) (2-3)] {} ;
                   \node [highlight = (3-1) (3-3)] {} ;
                   \end{tikzpicture}}]
a & a + b & a + b + c & L_1 \\
a & a     & a + b     & L_2 \\
a & a     & a         & L_3
\end{pNiceArray}\]

\vspace{1cm}

Dans l'exemple suivant, on utilise les «nœuds larges» pour surligner une zone de
la matrice.\par\nobreak
\begin{Verbatim}
\begin{pNiceArray}{>{\strut}cccc}[create-large-nodes,margin,extra-margin=2pt]
  A_{11} & A_{12} & A_{13} & A_{14} \\
  A_{21} & A_{22} & A_{23} & A_{24} \\
  A_{31} & A_{32} & A_{33} & A_{34} \\
  A_{41} & A_{42} & A_{43} & A_{44}  
\CodeAfter
  \tikz \path [~emphase#name suffix = -large@,fill = red!15, blend mode = multiply]
      (1-1.north west)
   |- (2-2.north west)
   |- (3-3.north west)
   |- (4-4.north west)
   |- (4-4.south east)
   |- (1-1.north west) ; 
\end{pNiceArray}
\end{Verbatim}

\[\begin{pNiceArray}{>{\strut}cccc}[create-large-nodes,margin,extra-margin=2pt]
  A_{11} & A_{12} & A_{13} & A_{14} \\
  A_{21} & A_{22} & A_{23} & A_{24} \\
  A_{31} & A_{32} & A_{33} & A_{34} \\
  A_{41} & A_{42} & A_{43} & A_{44}  
\CodeAfter
  \tikz \path [name suffix = -large,fill = red!15, blend mode = multiply]
      (1-1.north west)
   |- (2-2.north west)
   |- (3-3.north west)
   |- (4-4.north west)
   |- (4-4.south east)
   |- (1-1.north west) ; 
\end{pNiceArray}\]

\subsection{Utilisation directe des nœuds Tikz}

Dans l'exemple suivant, on souhaite illustrer le produit mathématique de deux matrices.

\medskip
L'utilisation de |{NiceMatrixBlock}| avec l'option |auto-columns-width| va
permettre que toutes les colonnes aient la même largeur ce qui permettra un
alignement des deux matrices superposées.
\begin{Verbatim}
\begin{NiceMatrixBlock}[auto-columns-width]
\end{Verbatim}

\begin{Verbatim}
\NiceMatrixOptions{nullify-dots}
\end{Verbatim}

Les trois matrices vont être disposées les unes par rapport aux autres grâce à un tableau de LaTeX.
\begin{Verbatim}
$\begin{array}{cc}
& 
\end{Verbatim}

La matrice $B$ a une «première rangée» (pour $C_j$) d'où l'option |first-row|.
\begin{Verbatim}
\begin{bNiceArray}{c>{\strut}cccc}[name=B,first-row]
       &          & ~emphase#C_j@                      \\
b_{11} & \Cdots   & b_{1j} & \Cdots & b_{1n} \\
\Vdots &          & \Vdots &        & \Vdots \\
       &          & b_{kj}                   \\
       &          & \Vdots                   \\
 b_{n1}  & \Cdots & b_{nj} & \Cdots & b_{nn} 
\end{bNiceArray} \\ \\
\end{Verbatim}

La matrice $A$ a une «première colonne» (pour $L_i$) d'où l'option |first-col|.
\begin{Verbatim}
\begin{bNiceArray}{cc>{\strut}ccc}[name=A,first-col]
    & a_{11} & \Cdots &        &        & a_{1n} \\
    & \Vdots &        &        &        & \Vdots \\
~emphase#L_i@ & a_{i1} & \Cdots & a_{ik} & \Cdots & a_{in} \\
    & \Vdots &        &        &        & \Vdots \\
    & a_{n1} & \Cdots &        &        & a_{nn} 
\end{bNiceArray}
& 
\end{Verbatim}

Dans la matrice produit, on remarquera que les lignes en pointillés sont «semi-ouvertes».
\begin{Verbatim}
\begin{bNiceArray}{cc>{\strut}ccc}
       & &        & & \\
       & & \Vdots     \\
\Cdots & & c_{ij}     \\
\\
\\
\end{bNiceArray} 
\end{array}$

\end{NiceMatrixBlock}
\end{Verbatim}

\begin{Verbatim}                             
\begin{tikzpicture}[remember picture, overlay]
 \node [highlight = (A-3-1) (A-3-5) ] {} ; 
 \node [highlight = (B-1-3) (B-5-3) ] {} ; 
 \draw [color = gray] (A-3-3) to [bend left] (B-3-3) ; 
\end{tikzpicture}
\end{Verbatim}

\begin{NiceMatrixBlock}[auto-columns-width]
\NiceMatrixOptions{nullify-dots}
$\begin{array}{cc}
& 
\begin{bNiceArray}{c>{\strut}cccc}[name=B,first-row]
      &        & C_j \\
b_{11} & \Cdots & b_{1j} & \Cdots & b_{1n} \\
\Vdots &       & \Vdots &       & \Vdots \\
       &       & b_{kj}  \\
       &       & \Vdots \\
 b_{n1}  & \Cdots & b_{nj} & \Cdots & b_{nn} 
\end{bNiceArray} \\ \\
\begin{bNiceArray}{cc>{\strut}ccc}[name=A,first-col]
    & a_{11} & \Cdots &  &  & a_{1n} \\
    & \Vdots &       &  &  & \Vdots \\
L_i & a_{i1} & \Cdots & a_{ik} & \Cdots & a_{in} \\
    & \Vdots &       &  &  & \Vdots \\
    & a_{n1} & \Cdots &  &  & a_{nn} \\
\end{bNiceArray}
& 
\begin{bNiceArray}{cc>{\strut}ccc}
       &  &        & & \\
       &  & \Vdots     \\
\Cdots &  & c_{ij}      \\
\\
\\
\end{bNiceArray} 
\end{array}$

\end{NiceMatrixBlock}

\begin{tikzpicture}[remember picture, overlay]
 \node [highlight = (A-3-1) (A-3-5) ] {} ; 
 \node [highlight = (B-1-3) (B-5-3) ] {} ; 
 \draw [color = gray] (A-3-3) to [bend left] (B-3-3) ; 
\end{tikzpicture}

\section*{Autre documentation}

Le document |nicematrix.pdf| (fourni avec l'extension \pkg{nicematrix}) contient
une traduction anglaise de la documentation ici présente, ainsi que le code
source commenté et un historique des versions.

\end{document}

% Local Variables:
% TeX-fold-mode: t
% TeX-fold-preserve-comments: nil
% flyspell-mode: true
% fill-column: 80
% End:

### 2020-08-14 MSU-thesis_estee_MaxFormat.tex
Submit to University

@@
 
+\hypersetup{draft}
 \sisetup{input-uncertainty-signs}
@@
 
-Besides measuring neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities and proton-neutron asymmetries.
+Besides observing neutrons stars directly, the laboratory provides the only way to study dense nuclear matter at different densities and proton-neutron asymmetries.
 It has been shown that charged pions are produced from nucleon-nucleon collisions through the short-lived intermediate $\Delta$ resonances, in the early, dense parts of Heavy Ion Collisions (HICs).
@@
 {\it
-To my wife for her unchanging support and patience\\ through this journey,
+To my wife for her unwavering support and patience,\\
 and to my flowering bud, Hanna. 
@@
 I wanted to thank my Aunts, Uncles, and Grandparents who have helped me financially and supported me throughout my schooling.
-To my Uncle Brian York, for his advice through the years and stimulating discussions.
+To my Uncle Brian York, for his advice through the years and his support throughout this journey.
 

### 2020-08-14 chapters_altFormat/ch_intro.tex
Submit to University

@@
 
-Guided by Eq.~\ref{eq:semiEmp}, we can separate the energy density $E$ of a system into two components,
+Guided by Eq.~\ref{eq:semiEmp}, we can separate the energy per nucleon, $E$, of a system into two components,
 
@@
 The symmetry energy is referred to as \emph{soft} if the second order term in the Taylor expansion -- at saturation density-- is negative, and \emph{stiff} in the case of a curvature that is close to zero or positive.
-It is the pressure coming from the symmetry energy that is driving the relative dynamics of neutrons and protons, where a stiff symmetry energy expels neutrons from dense matter and vice versa for the soft symmetry energy.
+It is the pressure coming from the symmetry energy that is driving the relative dynamics of neutrons and protons, where a stiff symmetry energy expels neutrons from dense matter and expels them less for a softer symmetry energy.
 

### 2020-08-14 chapters_altFormat/ch_results.tex
Submit to University

@@
 \includegraphics[width=.8\textwidth]{singleRatio108_log.png}
-\caption{Pion spectral ratio $\tin{108}{112}$ system and comparisons to 7 theoretical models.}
+\caption{Pion spectral ratio $\tin{108}{112}$ system and comparisons to 6 theoretical models.}
 \label{fig:SRsn108}
@@
 \includegraphics[width=.8\textwidth]{singleRatio132_sum.png}
-\caption{Pion spectral ratio $\tin{132}{124}$ system and comparisons to 7 theoretical models.}
+\caption{Pion spectral ratio $\tin{132}{124}$ system and comparisons to 6 theoretical models.}
 \label{fig:SRsn132}

### 2020-08-14 macrosForThesis/PionSpectra/PlotDataTheory/publish108SingleSum.C

@@ -31,6 +31,9 @@ TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
 
 void publish108SingleSum()
 {
+  gStyle->SetLineWidth(8);
+
+
   TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
   TGraphErrors *pBUU_soft_no_DR = (TGraphErrors *) pBUU_soft_no_f->Get("piRatio108");
 
@@ -270,10 +273,10 @@ void publish108SingleSum()
   TLegend *leg_soft = new TLegend(.67,.56,.97,.85);
   leg_soft->AddEntry(tuQMD_sum,"TuQMD","FL");
   leg_soft->AddEntry(urQMD_sum,"urQMD","FL");
-  //leg_soft->AddEntry(imQMD_sum,"ImQMD","FL");
+  leg_soft->AddEntry(imQMD_sum,"IQMD-BNU","FL");
   leg_soft->AddEntry(xbuu_sum,"#chiBUU","FL");
-  //  leg_soft->AddEntry(amdJAM_sum,"AMD + JAM","FL");
-  //  leg_soft->AddEntry(pBUU_sum,"pBUU","FL");
+  leg_soft->AddEntry(amdJAM_sum,"AMD + JAM","FL");
+  leg_soft->AddEntry(pBUU_sum,"pBUU","FL");
 
   leg_soft->SetBorderSize(0);
   leg_soft->SetFillStyle(0);
@@ -289,11 +292,11 @@ void publish108SingleSum()
 
   singleRatio_sn108->GetXaxis()->SetNdivisions(505);
   singleRatio_sn108->Draw("L E1");
-  //  pBUU_sum->Draw("same  LE3 ");
-  //  imQMD_sum->Draw("same  LE3");
+  pBUU_sum->Draw("same  LE3 ");
+  imQMD_sum->Draw("same  LE3");
   xbuu_sum->Draw("same  LE3 ");
   tuQMD_sum->Draw("same  LE3");
-  //  amdJAM_sum->Draw("same  LE3 ");
+  amdJAM_sum->Draw("same  LE3 ");
   urQMD_sum->Draw("same  LE3");
   singleRatio_sn108->DrawCopy("same L E1");  
   leg_soft->Draw();

### 2020-08-14 macrosForThesis/PionSpectra/PlotDataTheory/publish132SingleSum.C

@@ -31,6 +31,8 @@ TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
 
 void publish132SingleSum()
 {
+  gStyle->SetLineWidth(8);
+
   TFile *pBUU_soft_no_f = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
   TGraphErrors *pBUU_soft_no_DR = (TGraphErrors *) pBUU_soft_no_f->Get("piRatio132");
 
@@ -285,11 +287,11 @@ void publish132SingleSum()
 
   TLegend *leg_soft = new TLegend(.67,.56,.97,.85);
   leg_soft->AddEntry(tuQMD_sum,"TuQMD","FL");
-  //  leg_soft->AddEntry(urQMD_sum,"urQMD","FL");
-  leg_soft->AddEntry(imQMD_sum,"ImQMD","FL");
+  leg_soft->AddEntry(urQMD_sum,"urQMD","FL");
+  leg_soft->AddEntry(imQMD_sum,"IQMD-BNU","FL");
   leg_soft->AddEntry(xbuu_sum,"#chiBUU","FL");
-  //  leg_soft->AddEntry(amdJAM_sum,"AMD + JAM","FL");
-  //  leg_soft->AddEntry(pBUU_sum,"pBUU","FL");
+  leg_soft->AddEntry(amdJAM_sum,"AMD + JAM","FL");
+  leg_soft->AddEntry(pBUU_sum,"pBUU","FL");
 
   leg_soft->SetBorderSize(0);
   leg_soft->SetFillStyle(0);
@@ -305,12 +307,12 @@ void publish132SingleSum()
 
   singleRatio_sn132->GetXaxis()->SetNdivisions(505);
   singleRatio_sn132->Draw("L E1");
-  //  pBUU_sum->Draw("same  LE3 ");
+  pBUU_sum->Draw("same  LE3 ");
   imQMD_sum->Draw("same  LE3");
   xbuu_sum->Draw("same  LE3 ");
   tuQMD_sum->Draw("same  LE3");
-  //  amdJAM_sum->Draw("same  LE3 ");
-  //  urQMD_sum->Draw("same  LE3");
+  amdJAM_sum->Draw("same  LE3 ");
+  urQMD_sum->Draw("same  LE3");
   singleRatio_sn132->DrawCopy("same L E1");  
   leg_soft->Draw();
   leg_stiff->Draw();

### 2020-08-14 macrosForThesis/PionSpectra/PlotDataTheory/publishDoubleRatioSum.C

@@ -282,8 +282,8 @@ void publishDoubleRatioSum()
 
   TLegend *leg_soft = new TLegend(.2,.6,.6,.9);
   leg_soft->AddEntry(tuQMD_sum,"TuQMD","FL");
-  //  leg_soft->AddEntry(urQMD_sum,"urQMD","FL");
-  leg_soft->AddEntry(imQMD_sum,"ImQMD","FL");
+  leg_soft->AddEntry(urQMD_sum,"urQMD","FL");
+  leg_soft->AddEntry(imQMD_sum,"IQMD","FL");
   leg_soft->AddEntry(xbuu_sum,"#chiBUU","FL");
   leg_soft->AddEntry(amdJAM_sum,"AMD + JAM","FL");
   leg_soft->AddEntry(pBUU_sum,"pBUU","FL");
@@ -306,7 +306,7 @@ void publishDoubleRatioSum()
   xbuu_sum->Draw("same  LE3 ");
   tuQMD_sum->Draw("same  LE3");
   amdJAM_sum->Draw("same  LE3 ");
-  //  urQMD_sum->Draw("same  LE3");
+  urQMD_sum->Draw("same  LE3");
   doubleRatio->DrawCopy("same L E1");  
   leg_soft->Draw();
   leg_stiff->Draw();

### 2020-08-14 macrosForThesis/PionSpectra/PlotDataTheory/publishPionsSpectra108Sn.C

@@ -137,6 +137,34 @@ void publishPionsSpectra108Sn()
   int pBUUSoft_msz = 3;
   int pBUUSoft_lw = 10;
   int pBUUSoft_ls = 10;
+
+  
+  int imQMD_Soft_mc = kCyan;
+  int imQMD_Soft_fc = kCyan;
+  int imQMD_Soft_fs = 1;
+  int imQMD_Soft_ms = 26;
+  int imQMD_Soft_msz = 3;
+  int imQMD_Soft_lw = 10;
+  int imQMD_Soft_ls = 7;
+
+  imQMD_pim_sum->SetFillColorAlpha(imQMD_Soft_fc,.5);
+  imQMD_pim_sum->SetFillStyle(imQMD_Soft_fs);
+  imQMD_pim_sum->SetMarkerStyle(imQMD_Soft_ms);
+  imQMD_pim_sum->SetMarkerColor(imQMD_Soft_mc);
+  imQMD_pim_sum->SetMarkerSize(imQMD_Soft_msz);
+  imQMD_pim_sum->SetLineStyle(imQMD_Soft_ls);
+  imQMD_pim_sum->SetLineWidth(imQMD_Soft_lw);
+  imQMD_pim_sum->SetLineColor(imQMD_Soft_mc);
+
+  imQMD_pip_sum->SetFillColorAlpha(imQMD_Soft_fc,.5);
+  imQMD_pip_sum->SetFillStyle(imQMD_Soft_fs);
+  imQMD_pip_sum->SetMarkerStyle(imQMD_Soft_ms);
+  imQMD_pip_sum->SetMarkerColor(imQMD_Soft_mc);
+  imQMD_pip_sum->SetMarkerSize(imQMD_Soft_msz);
+  imQMD_pip_sum->SetLineStyle(imQMD_Soft_ls);
+  imQMD_pip_sum->SetLineWidth(imQMD_Soft_lw);
+  imQMD_pip_sum->SetLineColor(imQMD_Soft_mc);
+
   
   tuQMD_pim_sum->SetFillColorAlpha(tuQMD_Soft_fc,.5);
   tuQMD_pim_sum->SetFillStyle(tuQMD_Soft_fs);
@@ -270,6 +298,7 @@ void publishPionsSpectra108Sn()
   leg->AddEntry(urQMD_pim_sum,"urQMD","FL");
   leg->AddEntry(xbuu_pim_sum,"#chiBUU","FL");
[…]

### 2020-08-14 macrosForThesis/PionSpectra/PlotDataTheory/publishPionsSpectra132Sn.C

@@ -137,6 +137,33 @@ void publishPionsSpectra132Sn()
   int pBUUSoft_msz = 3;
   int pBUUSoft_lw = 10;
   int pBUUSoft_ls = 10;
+
+  int imQMD_Soft_mc = kCyan;
+  int imQMD_Soft_fc = kCyan;
+  int imQMD_Soft_fs = 1;
+  int imQMD_Soft_ms = 26;
+  int imQMD_Soft_msz = 3;
+  int imQMD_Soft_lw = 10;
+  int imQMD_Soft_ls = 7;
+
+  imQMD_pim_sum->SetFillColorAlpha(imQMD_Soft_fc,.5);
+  imQMD_pim_sum->SetFillStyle(imQMD_Soft_fs);
+  imQMD_pim_sum->SetMarkerStyle(imQMD_Soft_ms);
+  imQMD_pim_sum->SetMarkerColor(imQMD_Soft_mc);
+  imQMD_pim_sum->SetMarkerSize(imQMD_Soft_msz);
+  imQMD_pim_sum->SetLineStyle(imQMD_Soft_ls);
+  imQMD_pim_sum->SetLineWidth(imQMD_Soft_lw);
+  imQMD_pim_sum->SetLineColor(imQMD_Soft_mc);
+
+  imQMD_pip_sum->SetFillColorAlpha(imQMD_Soft_fc,.5);
+  imQMD_pip_sum->SetFillStyle(imQMD_Soft_fs);
+  imQMD_pip_sum->SetMarkerStyle(imQMD_Soft_ms);
+  imQMD_pip_sum->SetMarkerColor(imQMD_Soft_mc);
+  imQMD_pip_sum->SetMarkerSize(imQMD_Soft_msz);
+  imQMD_pip_sum->SetLineStyle(imQMD_Soft_ls);
+  imQMD_pip_sum->SetLineWidth(imQMD_Soft_lw);
+  imQMD_pip_sum->SetLineColor(imQMD_Soft_mc);
+
   
   tuQMD_pim_sum->SetFillColorAlpha(tuQMD_Soft_fc,.5);
   tuQMD_pim_sum->SetFillStyle(tuQMD_Soft_fs);
@@ -291,6 +318,7 @@ void publishPionsSpectra132Sn()
   leg->AddEntry(tuQMD_pim_sum,"TuQMD","FL");
   leg->AddEntry(urQMD_pim_sum,"urQMD","FL");
[…]

### 2020-08-19 chapters_altFormat/ch_experiment.tex
Second revisions to University

@@
 \begin{figure}[!htb]
-\includegraphics[width=\textwidth]{fc_overview2.png}
+\centering
+\includegraphics[width=.8\textwidth]{fc_overview2.png}
 \label{fig:fc_overview2}
@@
 \midrule
- Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & Distance to pad-plane \si{\milli\metre}& Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
+ Plane & Material & Diameter \si{\micro\metre} & Pitch \si{\milli\metre} & $D_{p.p.}$ \si{\milli\metre}& Tension \si{\newton} & Voltage \si{\volt}\\ [0.5ex] 
  Anode  & Au-plated W   &  20  &  4  &  4   &  0.5  &  1460  \\
@@
 \end{tabular}
-\caption{Wire plane properties.}
+\caption{Wire plane properties. Were $D_{p.p.}$ is the distance from the anode wire to the pad plane.}
 \label{tb:wireplane}
@@
 The gas purity was monitored with an oxygen and water monitor which are the two most concerning contaminants.
-The water never exceed  100 ppm  and the oxygen level never exceeded 10 ppm, which are the two main contaminants in the gas which can absorb primary electrons \cite{tpcAging}.
+The water never exceed  50 ppm  and the oxygen level never exceeded 50 ppm, which are the two main contaminants in the gas which can absorb primary electrons \cite{tpcAging}.
 Figure~\ref{fig:driftvel} shows the drift velocity of P10 gas at 1 atm (\SI{760}{\torr}) as a function of the reduced electric field value given in units of \si{\volt\per\centi\metre\per\torr}.

### 2020-08-19 macrosForThesis/DigiTasks/PulseShapePlot/plotPulse.C

@@ -5,6 +5,8 @@ using namespace style;
 void plotPulse()
 {
 
+  gStyle->SetLineWidth(3);
+  
   TFile *f = TFile::Open("pulses.root");
   TF1 *sat = (TF1 *)f->Get("sat");
   TF1 *not_sat = (TF1 *)f->Get("not_sat");  
@@ -65,7 +67,7 @@ void plotPulse()
   TCanvas *c2 = style::stdcvs("cvs_2",1000,500);
   TCanvas *c3 = style::stdcvs("cvs_3",1000,500);
   sat->SetLineWidth(6);
-  not_sat->SetLineWidth(6);
+  not_sat->SetLineWidth(9);
   not_sat->SetLineStyle(9);
 
   sat_b->SetLineWidth(6);
@@ -115,8 +117,9 @@ void plotPulse()
   
   sat->GetXaxis()->CenterTitle();
   sat->GetYaxis()->CenterTitle();
-
-
+  gStyle->SetLineWidth(3);
+  gPad->SetLineWidth(3);
+  gStyle->SetFrameLineWidth(3);
   //  not_sat->GetXaxis()->SetRangeUser(0,60);
   not_sat_b->GetXaxis()->SetRangeUser(0,40);
  

### 2020-08-19 macrosForThesis/PionSpectra/PlotDataTheory/publishIsoScale_pim.C

@@ -1,0 +1,414 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * Divide(TGraphErrors *a, TGraphErrors *b)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+
+     double ey_a = a->GetErrorY(i);
+     double ey_b = b->GetErrorY(i);
+      
+      x[i] = xx;
+      y[i] = y_a/y_b;
+      ex[i] = 0;
+      ey[i] = (y_a/y_b)*sqrt( pow(ey_a/y_a,2) + pow(ey_b/y_b,2)) ;
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+void publishIsoScale_pim()
+{
+  gStyle->SetLineWidth(8);
+
+  TFile *pBUU_soft_no_f_132 = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
[…]

### 2020-08-19 macrosForThesis/PionSpectra/PlotDataTheory/publishIsoScale_pip.C

@@ -1,0 +1,414 @@
+#include "/home/justin/mythesis/thesis_style.h"
+
+using namespace style;
+
+TGraphErrors * Divide(TGraphErrors *a, TGraphErrors *b)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+
+     double ey_a = a->GetErrorY(i);
+     double ey_b = b->GetErrorY(i);
+      
+      x[i] = xx;
+      y[i] = y_a/y_b;
+      ex[i] = 0;
+      ey[i] = (y_a/y_b)*sqrt( pow(ey_a/y_a,2) + pow(ey_b/y_b,2)) ;
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+
+TGraphErrors * CombineGraph(TGraphErrors *a, TGraphErrors *b, double scale = 1)
+{
+  int a_num = a->GetN();
+  int b_num = b->GetN();
+
+  double x[a_num], y[a_num], ey[a_num], ex[a_num];
+
+  if(a_num != b_num)
+    cout<<"Graph points NOT EQUAL!!!"<<endl;
+
+  for(int i = 0; i < a_num; i++)
+    {
+      double xx,y_a,y_b;
+      a->GetPoint(i,xx,y_a);
+      b->GetPoint(i,xx,y_b);
+      y_a *= scale;
+      y_b *= scale;
+
+      x[i] = xx;
+      y[i] = (y_a + y_b)/2.;
+      ex[i] = 0;
+      ey[i] = abs(y[i] - y_a);
+    }
+
+  return (new TGraphErrors(a_num,x,y,ex,ey));
+}
+
+void publishIsoScale_pip()
+{
+  gStyle->SetLineWidth(8);
+
+  TFile *pBUU_soft_no_f_132 = TFile::Open("/home/justin/Homework/pBUU/K210EM100/input_for_esym.info/pBUUSoft_r-1.root");
[…]

### 2020-08-21 MSU-thesis_estee_MaxFormat.tex
Second revisions to University
Final submit to Uni

@@
 \usepackage{rotating}
+%\usepackage{showframe}
 \usepackage{mathtools}
@@
 \sisetup{input-uncertainty-signs}
+
+\makeatletter
+\patchcmd{\@makechapterhead}{\vspace*{50\p@}}{\vspace*{60pt}}{}{}%
+\patchcmd{\@makeschapterhead}{\vspace*{50\p@}}{\vspace*{60pt}}{}{}%
+\makeatother
 
@@
 It is the density dependence of the symmetry energy that is related to the outward pressure, which supports the star, preventing it from collapsing under its own gravitation force.
+
  For years, theory and experiment  have combined to progressively constrain the Equation of State (EoS) of symmetric matter, and it was only in the last decade that the density dependence of pure neutron matter, the symmetry energy, had been constrained at normal nuclear densities and lower, $\rho < \rho_o$.
@@
 This prompted the construction of the Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) which was constructed to measure charged pions produced in neutron-rich HICs with a high efficiency.
+
 This Dissertation encompasses the first pion spectra from the most neutron-rich ($\tin{132}{124}$), and neutron-poor ($\tin{108}{112}$) systems measured at \SI{270}{\MeVA} taken at the RIBF facility at RIKEN.
@@
 To all my friends at the University of Tokyo CNS for your friendship, softball, and hospitality.
-I also wanted to thank our Polish colleagues Jerzy \L ukasik, Piotr Paw\l owski, Krzysztof Pelczar, and Pawe\l \L asko for their hard work in developing the KATANA array and for setting up the trigger in the experiment.
+I also wanted to thank our Polish colleagues Jerzy \L ukasik, Piotr Paw\l owski, Krzysztof Pelczar, and Pawe\l ~\L asko for their hard work in developing the KATANA array and for setting up the trigger in the experiment.
 
@@
 %% Uncomment ``\appendix'' for one appendix, ``\appendices'' for more than one, and neither for none.
-%%\appendix
-\appendices
+\appendix
+%\appendices
 \include{chapters_altFormat/ch_appendix}
@@
 %% just type \chapter{<appendix 1 name>}
-%% and enter the text of the appendix as you would a chapter.
+%% and enter the text of the appendix as you would a chapter.fil
 

### 2020-08-21 chapters_altFormat/ch_data_analysis.tex
Second revisions to University
Final submit to Uni

@@
 
+\caption{Example of before and after the space charge correction comparing the momentum measurement with the BDC vertex included and without it.}
  \label{fig:mom_sc}
@@
 The electron will terminate on the closest anode wire, that is the anode wire that is closest to the shifted z-position $z\textprime = z + dz$, and the final electron z position is then updated to be the same as the anode wire it terminated on.
+
  The total drift time of the electron $t$ is calculated as,
@@
 The final output charge on all the pads are the superposition of the PRFs of all drifted electrons.
+
  The MC PRF is expressed as,
@@
 Now that the pion distribution is flattened, we take bin slices along $p/q$ to determine the particle yield in a certain bin and also estimate the background contribution.
-The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0$ Dalitz decay described in Appendix~\ref{appen:dalitz}.
+The background relevant to  the charged pions comes from the $e^+$ and $e^-$ resulting from the $\pi^0 \rightarrow e^+ e^- \gamma$ Dalitz decay.
 Figure~\ref{fig:pimmom_flat} shows the Gaussian fits to the $z_i$ distribution for two different $p/q$ binned cuts around the $\pi^-$ particles, whereas Figure~\ref{fig:pipmom_flat} shows the background $\pi^+$ particles.

### 2020-08-21 chapters_altFormat/ch_data_analysis_2.tex
Submit to University
Final submit to Uni

@@
     \centering
-    \subfloat[ $\tin{108}{112}$ system. ]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system. ]{{\includegraphics[width=.9\textwidth,valign=c]{angularCut-pim-Sn108.png} \label{fig:pim108angle}}}%
     \qquad
-    \subfloat[$\tin{132}{124}$ system. ]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
+    \subfloat[$\tin{132}{124}$ system. ]{{\includegraphics[width=.9\textwidth,valign=c]{angularCut-pim-Sn132.png} \label{fig:pim132angle}}}%
 
@@
     \centering
-    \subfloat[ $\tin{108}{112}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
+    \subfloat[ $\tin{108}{112}$ system.]{{\includegraphics[width=.9\textwidth,valign=c]{angularCut-pip-Sn108.png} \label{fig:pip108angle}}}%
     \qquad
-    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.8\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
+    \subfloat[$\tin{132}{124}$ system.]{{\includegraphics[width=.9\textwidth,valign=c]{angularCut-pip-Sn132.png} \label{fig:pip132angle}}}%
      \caption{Measured $\pi^+$ angular distribution for the two spherical angles $\theta$ and $\phi$.}
@@
 
-\subsection{Correcting Pion Spectra (Efficiency and Acceptance)}
+\section{Correcting Pion Spectra (Efficiency and Acceptance)}
 \label{sec:corrSpectra}
@@
 
-\[ \left( \begin{array}{cccc}
-1 & 0 & 0 & 0\\
-0 & 1 & 0 & 0\\
-0 & 0 & \gamma & -\beta \gamma \\
+\begin{equation}
+A = \begin{pmatrix}
+1 & 0 & 0 & 0\\[-25pt]
+0 & 1 & 0 & 0\\[-25pt]
+0 & 0 & \gamma & -\beta \gamma\\[-25pt]
 0 & 0 & -\beta \gamma & \gamma
-\end{array} \right) \]
+\end{pmatrix},
+\end{equation}
 
@@
 
-\subsection{Pion Statistical Error}
+\section{Pion Statistical Error}
 
